### 2025-03-25 paper/paper.tex
added paper

% !TEX root = /Users/jonasfrugte/Desktop/Research_Project/paper/paper.tex
\documentclass[11pt]{article} % Choose a document class, like 'article', 'report', or 'revtex4-2' for scientific papers

% Packages to enhance functionality
\usepackage[utf8]{inputenc}   % Allows UTF-8 character encoding
\usepackage{amsmath}          % For advanced math typesetting
\usepackage{amssymb}          % Additional math symbols
\usepackage{graphicx}         % For including images
\usepackage{hyperref}         % For hyperlinks in the document
\usepackage{geometry}         % For setting up page geometry
\geometry{margin=1in}         % 1-inch margins all around
\usepackage{authblk}          % For author affiliations
\usepackage{siunitx}          % For consistent typesetting of units
\usepackage{subcaption}       % For subfigures within a figure
\usepackage{float}            % To control float positioning (e.g., figures/tables)
\usepackage{enumitem}         % Better control over list formatting
\usepackage{tocloft}
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usetikzlibrary{shapes.geometric, arrows, backgrounds}
\usepackage{empheq}
\usepackage{booktabs}
\usepackage{authblk}

\usepackage{listings}
\usepackage{xcolor} % For color customization (optional)

\usepackage[backend=biber, style=numeric]{biblatex}
\addbibresource{sources.bib}

\lstset{
    language=Python,
    basicstyle=\ttfamily\small,
    keywordstyle=\color{blue}\bfseries,
    stringstyle=\color{red},
    commentstyle=\color{gray}\itshape,
    showstringspaces=false,
    frame=single, % Adds a frame around the code
    breaklines=true % Wrap long lines
}

% some custom commands that should've already been in latex by default
\DeclareRobustCommand{\d}{\ifmmode\text{d}\else d\fi}
\DeclareRobustCommand{\Cov}{\ifmmode\text{Cov}\else d\fi}
\newcommand{\br}[1]{\ensuremath{\left( #1 \right)}}
\newcommand{\sbr}[1]{\ensuremath{\left[ #1 \right]}}

\setlength{\cftbeforesecskip}{5pt}

% Additional customizations (optional)
\setlength{\parindent}{0pt}   % No paragraph indentation
\setlength{\parskip}{1em}     % Add space between paragraphs

\title{\huge \textbf{Parameter Forecasts From CMB Lensing and Galaxy Lensing Power- and Bispectra}}

\author[1]{\textbf{Jonas Frugte}}
\author[2]{\textbf{P. Daniel Meerburg}}

\affil[1]{Independent Researcher $\infty$, The Netherlands}
\affil[2]{Van Swinderen Institute for Particle Physics and Gravity, University of Groningen, Nijenborgh 4, 9747 AG Groningen, The Netherlands}
\affil[ ]{}
\affil[ ]{\textit{E-mail:} \texttt{jonasfrugte@gmail.com}}

\date{20 February 2025} % Removes the default date

\begin{document}
\maketitle

% \textbf{
%     \huge \noindent 
%     Parameter Forecasts From CMB $\times$ Galaxy Lensing Powerspectra $\times$ Bispectra
%     \vspace{0.5cm}
%     \\
%     \normalsize
%     Jonas Frugte, Daan Meerburg\\
%     20 February 2025
%     }\\
% \vspace{0.5cm}
\hrule

\begin{abstract}
    (Placeholder text written with chatgpt) We present a Fisher forecast analysis of the constraints on cosmological parameters from measurements of CMB lensing and galaxy weak lensing power- and bispectra. Using simulated data consistent with upcoming surveys, we explore the sensitivity of lensing observables to parameters including the Hubble constant \(H_0\), baryon density \(\Omega_b h^2\), dark matter density \(\Omega_c h^2\), scalar spectral index \(n_s\), amplitude of scalar fluctuations \(A_s\), and the sum of neutrino masses \(\sum m_\nu\).

    Our results indicate that combining power- and bispectra improves parameter constraints compared to power spectra alone, reducing uncertainties by a factor of \(\sim X\%\). In particular, we forecast constraints of \(\sigma(H_0) = X.X\) km/s/Mpc, \(\sigma(\Omega_b h^2) = X.XXX\), \(\sigma(\Omega_c h^2) = X.XXX\), \(\sigma(n_s) = X.XXXX\), \(\sigma(A_s) = X.XX \times 10^{-9}\), and \(\sigma(\sum m_\nu) = X.XX\) eV. These results demonstrate the potential of bispectrum measurements in future surveys such as the Simons Observatory, CMB-S4, LSST, and Euclid.

    Additionally, we investigate the impact of different noise models and survey redshift distributions, finding that improvements in galaxy redshift accuracy significantly enhance constraints on \(\sum m_\nu\). We also discuss the degeneracies between parameters and assess their impact using eigenvalue analysis of the Fisher matrix.

    This study highlights the advantages of combining lensing power- and bispectra for cosmological parameter estimation and provides a roadmap for optimizing future surveys to maximize their scientific return.
\end{abstract}
\pagebreak
%\tableofcontents

% \section{Conventions}
% Use $\mathbf k$ for wave numbers (Fourier space), $l$ for multipole moments, $m$ for magnetic quantum number. $P^{XY}(k, \eta, \eta')$ and $B^{XYZ}(k_1,k_2,k_3, \eta_1, \eta_2, \eta_3)$ correspond to bi and power spectra of fields $X$, $Y$, and $Z$ in Fourier space. $P_l^{XY}(\eta,\eta')$ and $B_{l_1l_2l_3}^{XYZ}(\eta_1, \eta_2, \eta_3)$ correspond to the multipole moments of the bi and power spectra of fields $X$, $Y$, and $Z$. If all times are the same we may instead write only one time argument and if the time is current time then we omit the time argument entirely. In other words,
% $$
% P_l^{XY}(\eta):=P_l^{XY}(\eta,\eta), \quad P_l^{XY}:=P_l^{XY}(\eta_{\text{current}}), \quad \text{similarly for all other spectra}.
% $$
% If both fields are the same, we may only denote one field, e.g. 
% $$
% P_l^{X}(\eta) := P_l^{XX}(\eta).
% $$

% Our Fourier convention is
% \begin{align}
%     &\text{Forward Transform:} & \quad \tilde{f}(k) &= \int_{-\infty}^{\infty} f(x) \, e^{-i k x} \, dx\\
%     &\text{Inverse Transform:} & \quad f(x) &= \frac{1}{2\pi} \int_{-\infty}^{\infty} \tilde{f}(k) \, e^{i k x} \, dk
% \end{align}
% we choose this convention because it is in line with the convention used in the CAMB source code. We work in natural units where $c=1$.

\section{Introduction}

Weak lensing has been used as a probe to constrain cosmological parameters since the 2000's \cite{kilbinger2015cosmology,Bacon2000,Kaiser2000,Waerbeke2000,Bartelmann2001}. A range of upcoming surveys, such as the survey by the Simons Observatory (SO) \cite{Ade2019}, the Legacy Survey of Space and Time (LSST) \cite{Ivezic2019}, and Euclid \cite{Laureijs2011} aim to measure weak lensing of the CMB and of galaxies. They will do so much more accurately than previous surveys, enabling tighter constraints on parameters and progressing our understanding of physics as a whole. Typically, the main value of interest is the power spectrum (or equivalently, two point function) of the lensing potential. This can be measured in terms of the lensing convergence $\kappa$, lensing shear $\gamma$, or lensing potential $\psi$. In the weak lensing regime they all contain the same amount of information and can be easily converted into one another, see section \ref{sec:weaklensstats}. For the CMB the lensing potential is generally measured directly \cite{Planck2018Lensing}, while with galaxies one instead measures the shear, as this can be deduced from the ellipiticity distribution of the observed galaxies \cite{HoekstraJain2008}.

Even with Gaussian initial conditions, there are significant amounts of non-Gaussianity in the weak lensing signal, especially in galaxy lensing due to it being a probe of the more recent distribution of matter \cite{Bernardeau1997,Takada2003}. In addition to looking at the power spectrum, a natural next step is thus to measure the lensing bispectrum. This has already been done for signals from galaxies \cite{vanWaerbeke2002}, but no proper detection has been done of the CMB lensing bispectrum. Upcoming surveys such as the ones mentioned earlier may detect this signal. Therefore it is relevant to know the detectability of, and parameter constraints possible from, all 4 signals: CMB lensing power- and bispectra and galaxy lensing power- and bispectra. Notably, this includes looking at cross correlation as well.

In this article we aim to answer this question in the context of experimental parameters similar to those of next generation surveys such as mentioned earlier. We are especially interested in seeing if approximate parameter degeneracies can be broken by combining all data. Calculating the covariance matrix of unbiased optimal estimators of the cosmological parameters can be done by combining cosmological simulations (in our case we use the CAMB package \cite{Lewis2000}) with a Fisher matrix analysis. See figure \ref{fig:flowchart} for an overview of the calculation done in this paper. The parameters looked at are the Hubble constant, $H_0$, physical baryon density, $\Omega_b h^2$, physical cold dark matter density, $\Omega_c h^2$, scalar spectral index, $n_s$, amplitude of primordial scalar fluctuations, $A_s$, sum of neutrino masses $\sum m_\nu$, and dark energy equation of state parameter, $w_0$. Though, we will find that some of the parameters are poorly constrained even when combining all 4 signals and thus omit them from the analysis in the results. As a secondary purpose, this paper also includes a derivation for the Fisher matrix formalism in the non-trivial context of bispectra with multiple tracers. The final formula can already be found in the literature (see e.g.\ the appendix of \cite{Kalaja_2021}), but the authors of this paper were not able to find a derivation and hence included it.

The structure of this paper is as follows: we first review the theory of weak lensing and weak lensing statistics in sections \ref{sec:weaklensing} and \ref{sec:weaklensstats}. We then provide technical details of the calculation in section \ref{sec:calcdetails}, such as choices for noise and numerical approximations made in the code. In section \ref{sec:results} we present the results, consisting of signal to noise ratios and parameter constraints for different combinations of the CMB and galaxy lensing power- and bispectra. We conclude in section \ref{sec:discussion}. Appendix \ref{sec:fisher} contains a short review of the Fisher matrix formalism and gives a technical derivation of the formula used to calculate each Fisher matrix element in the case of bispectra measurements with multiple tracers. Finally, appendix \ref{sec:shear} explains how to relate the shear to the lensing potential in spherical harmonic space.

All code written for this project is uploaded on github at LINK .

\begin{figure}[h!]
    \centering
    \input{figures/calc_flowchart.tikz}
    \caption{Flowchart of steps involved in calculating the minimum error/uncertainty in cosmological parameter estimates using convergence and shear power- and bi-spectra.}
    \label{fig:flowchart}
\end{figure}

\section{Weak Lensing}\label{sec:weaklensing}
\subsection{Approximations Made / Regime}
\subsection{Perturbed Photon Paths}
Work in \textbf{conformal newtonian gauge}. In this section we'll work in units where $c=1$. Denoting conformal time and conformal radial distance by $\eta$ and $\chi$, respectively, the perturbed line element is given by
\begin{equation}
    \d s^2 = a^2(\eta)((1 + 2\Psi_N) \d \eta^2 - (1 + 2\Phi_N) \gamma_{ij}\d x^i \d x^j) 
\end{equation}

where $\gamma_{ij}$ is the unperturbed line element
\begin{equation}
    \gamma_{ij} = \d x^i \d x^j = \d \chi^2 + f^2_K(\chi)(\d \theta^2 + \sin ^2 \theta \d \phi ^2).
\end{equation}
From here of on work in first order in the scalar potentials $\Phi_N$ and $\Psi_N$. For lensing we consider null-geodesics so $\d s^2 = 0$ and we can rewrite the perturbed line element as
\begin{equation}
    \d \hat s^2 = (1 + 4\Psi)\d\eta^2 - \gamma_{ij} \d x^i \d x^j
\end{equation}
with $\Psi$ the \textbf{Weyl Potential} given by $\Psi := (\Psi_N - \Phi_N)/2$. A null geodesic $x^\mu(\hat\lambda)$ in terms of its affine parameter ($\hat \lambda$) satisfies the geodesic equation
\begin{equation}
    \frac{\d^2 x^\mu}{\d \hat\lambda^2} + \Gamma^\mu_{\nu\rho} \frac{\d x^\nu}{\d \hat\lambda}\frac{\d x^\rho}{\d\hat\lambda} = 0,
\end{equation}
with $\hat g_{\mu\nu}(\d x^\mu/\d \hat\lambda)(\d x^\nu / \d \hat \lambda) = 0$.

In the unperturbed there are incoming radial solutions of the form $\chi = \eta_0 - \eta$. We can express $\hat\lambda$ in terms of $\eta$ using The 0-component of the geodesic equation,
\begin{equation}
    \frac{\d^2\eta}{\d\hat\lambda^2} + 2(\frac{\d\eta}{\d\hat\lambda})^2\frac{\d\Psi}{\d\eta}+2\frac{\d\eta}{\d\hat\lambda}\frac{\d x^i}{\d\hat\lambda}\frac{\partial\Psi}{\partial x^i} = 0,
\end{equation}
where the derivative $\d\Psi/\d\eta = \partial_\eta\Psi+(\d x^i/\d\eta)\partial_i\Psi$ is along the perturbed ray. This allows us to rewrite the other (spatial) geodesic equations as
\begin{equation}
    \frac{\d^2x^i}{\d\eta^2} - 2\frac{\d x^i}{\d\eta}(\frac{\d\Psi}{\d\eta} + \frac{\d x^j}{\d\eta}\frac{\partial\Psi}{\partial x^j})+2\gamma^{ij}\frac{\partial\Psi}{\partial x^j}+\bar\Gamma^i_{jk}\frac{\d x^j}{\d\eta}\frac{\d x^k}{\d \eta} = 0,
\end{equation}
where $\bar\Gamma^i_{jk}$ are the connection coefficients of the unperturbed three geometry $\gamma_{ij}$. Without loss of generality we consider an observer located at the origin of the spatial coordinates, meaning we are interested in rays that end at $x^i = 0$. In that case $\d \chi/\d \eta = -1 + O(\Psi)$, $\d \theta/\d\eta = O(\Psi)$ and $\d \phi/\d\eta = O(\Psi)$ (with $\theta$ and $\phi$ angular spatial coordinates). For the case of such rays we can evaluate the connection coefficients and rewrite the spatial geodesic equations as
\begin{gather}
    \frac{\partial^2\chi}{\d\eta^2}+2\frac{\d\Psi}{\d\eta}=0,\\
    \frac{\d^2\theta}{\d\eta^2}-2\frac{\d\ln f_K(\chi)}{\d\chi}\frac{\d\theta}{\d\eta}+\frac{2}{f^2_K(\chi)}\frac{\partial\Psi}{\partial\theta}=0,\\
    \frac{\d^2\phi}{\d\eta^2} - 2\frac{\d \ln f_K(\chi)}{\d\chi}\frac{\d\phi}{\d\eta}+\frac{2}{f^2_K(\chi)}\frac{1}{\sin^2(\theta)}\frac{\partial\Psi}{\partial\phi}=0.
\end{gather}
Integrating twice and using the null condition we can rewrite the the first equation as
\begin{equation}
    \chi = \eta_0 - \eta - 2\int_{\eta_0}^\eta\Psi\d\eta',
\end{equation}
where the integral is along the ray. Since we are working at first order in $\Psi$ we can evaluate the integral along the unperturbed path, this is known as the \textbf{Born approximation}.\textit{Next follows a short discussion of why perturbations in observation time can be ignored, include?} Doing the same for the other two equations gives
% TODO: make n a proper vector (bold)
\begin{gather}
    \theta(\eta_0 - \chi_*)=\theta_0 - \int_0^{\chi_*}\d\chi\frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}2\frac{\partial}{\partial \theta}\Psi(\chi\hat n, \eta_0 - \chi),\\
    \phi(\eta_0 - \chi_*)=\theta_0 - \int_0^{\chi_*}\d\chi\frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}\frac{2}{\sin^2\theta}\frac{\partial}{\partial \phi}\Psi(\chi\hat n, \eta_0 - \chi),
\end{gather}
where $\theta_0$ and $\phi_0$ label the line of sight $\hat n$. The displacement vector $\mathbf{\alpha}$ indicates the infinitesimal displacement of the lensed rays from their unpetrubed observed angles and is related to $\theta$ and $\phi$ through
$$
\alpha_\theta = \theta - \theta_0, \quad \alpha_\phi = \sin^2\theta(\phi - \phi_0).
$$
This allows us to summarize the above results as
\begin{equation}
    \mathbf \alpha = -2 \int_0^{\chi_*}\d \chi \frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}\nabla_{\hat{\mathbf n}}\Psi(\chi\hat{\mathbf n};\eta_0 - \chi).
\end{equation}
When working in the weak field limit (i.e. small spacetime curvature) the $00$ component of Einsteins equations give us a poisson equation \cite{dodelson2020modern},
\begin{equation}
    -\nabla^2\Psi(\textbf{x}, \eta) =  4\pi G a(\eta)^2\delta\rho(\mathbf x, \eta)
\end{equation}
where $K$ is $0$, $1$, or $-1$ depending on the curvature of the universe, $\Delta = \partial_x^2 + \partial_y^2 + \partial_z^2$ is the 3 dimensional laplace operator, $\delta\bar\rho$ is the comoving total energy perturbation (evaluated in the rest-frame of the total energy). %TODO: proper definition for anisotropic stress and explain why we can ignore it

\subsection{Convergence, shear}
Quantities we are interested in measuring depend on how the lensing vector changes with the observation angle. \textbf{Magnification matrix} is defined by
\begin{equation}
    A_{ij} := \delta_{ij} + \frac{\partial}{\partial\theta_i}\alpha_i(\hat{\mathbf n}).
\end{equation}
This matrix can then be decomposed in the following form, which immediately gives us definitions for the \textbf{convergence}, $\kappa$, \textbf{shear}, $\gamma_1$ and $\gamma_2$, and \textbf{rotation}, $\omega$:
\begin{equation}
    A_ij(\hat{\mathbf n}) = \begin{pmatrix}
        1 - \kappa - \gamma_1 & -\gamma_2 + \omega \\
        -\gamma_2 - \omega & 1 - \kappa + \gamma_1
    \end{pmatrix}.
\end{equation}
At 1st order $A$ is a symetric matrix by definition and so $\omega$ has to be 0, we will ignore it from here on out. Intuitively, A tells you how a small patch in the sky transforms due to weak lensing, i.e. if you have surface brightness $I$ then
% TODO: make delta thick
\begin{equation}
    I_{\text{emitted}}(\hat{\mathbf{n}} + \mathbf{\delta}) = I_{\text{observed}}((\hat{\mathbf{n}} + \mathbf{\delta}) + \mathbf\alpha(\hat{\mathbf{n}} + \mathbf{\delta})) \approx I_{\text{observed}}((\hat{\mathbf{n}} + \mathbf{\alpha} + A_{ij}\delta_{j}).
\end{equation}
given that $|A_{ij}| = (1 - \kappa)^2 + \omega^2 - |\gamma|^2 = 1 - 2\kappa + O(\kappa^2, \gamma^2, \omega^2)$, we can interpret $\kappa$ as telling us about the overal magnification of the source. The $\gamma_i$ represent the area preserving distortion, i.e. stretching and squeezing in a specific direction.

\subsection{lensing potential}
Taking the derivative out of the integral in equation REF leads us to defining the \textbf{lensing potential}, $\psi$,
\begin{equation}
    \mathbf \alpha = \nabla_{\hat{\mathbf n}}\psi, \quad \psi(\hat{\mathbf n}) = -2 \int_0^{\chi_*} \d \chi \frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}\Psi(\chi\hat{\mathbf n}; \eta_0 - \chi).
\end{equation}

We can relate $\kappa$ and $\gamma$ directly to the lensing potential as
% TODO: should these subscripts be theta and phi?
\begin{equation}
    \kappa = \frac{1}{2}\nabla^2\psi, \quad \gamma_1 = \frac{1}{2}(\partial_\theta^2 - \partial_\phi^2)\psi, \quad \gamma_2 = \partial_\theta\partial_\phi \psi.
\end{equation}

\subsection{Regarding window functions}
The window function for light emitted from a source at fixed radius $\chi_*$ is $(\chi-\chi^*)/(\chi_*\chi)$, if the source is instead distributed over the radius as $p(\chi)$, with $p(\chi)$ normalized to integrate to 1, we generalize the window function as
\begin{equation*}
    W(\chi) = \int_\chi^{\chi_*}\d\chi' p(\chi')\frac{\chi'-\chi}{\chi'\chi}.
\end{equation*}
The integration limit is up to surface of last scattering because further than that no source can contribute to the observed image. In the case of an image of the CMB we can take $p(\chi')=\delta(\chi'-\chi_*)$, in which case the window function reduces back to $(\chi_*-\chi)/(\chi_*\chi)$. In the case of different sources, we will differentiate between their respective window functions with a subscript. In this case the factor that changes is the distribution function. The sources considered in this paper are only the CMB and selections of galaxies from galaxy surveys.

\section{Weak Lensing Statistics}\label{sec:weaklensstats}
\subsection{Lensing Potential Powerspectrum}
Consider the lensing potential, $\psi$, then decomposing into spherical harmonics gives
\begin{equation}
    \psi(\hat{\mathbf n}) = \sum_{lm} \psi_{lm}Y_{lm}(\hat{\mathbf n}).
\end{equation}
On the other hand, consider the decomposition of $\Psi$ in fourier modes with the Fourier convention $\Psi(\mathbf x, \eta)=\int\frac{\d^3 \mathbf k}{(2\pi)^{3}}\Psi(\mathbf k, \eta)e^{i\mathbf k \cdot\mathbf x}$,
\begin{equation}
    \psi(\hat{\mathbf n}) = -2 \int_0^{\chi_*}\d \chi W(\chi) \int\frac{\d^3 \mathbf k}{(2\pi)^{3}}\Psi(\mathbf k, \eta_0 - \chi)e^{i\mathbf k \cdot\hat{\mathbf n}\chi}.
\end{equation}
We can than relate the multipole modes of $\psi$ to the fourier modes of $\Psi$ through % TODO: check sign convention here
\begin{gather}
    \psi_{lm} = \langle Y_l^m | \psi \rangle = \int\d^2\hat{\mathbf n} Y_l^m(\hat{\mathbf n})^* \psi(\hat{\mathbf n})\\ 
    = -2\int\d^2\hat{\mathbf n} Y_l^m(\hat{\mathbf n})^* \int_0^{\chi_*}\d \chi W(\chi) \int\frac{\d^3 \mathbf k}{(2\pi)^{3}}\Psi(\mathbf k, \eta_0 - \chi)e^{i\mathbf k \cdot\hat{\mathbf n}\chi}
\end{gather}
Now define the power spectrum as
\begin{equation}
    \langle \Psi(\mathbf k, \eta)\Psi^*(\mathbf k',\eta')\rangle = \frac{2\pi^2}{k^3}P_\Psi(k, \eta, \eta')\delta(\mathbf k - \mathbf k'),
\end{equation}
with $\eta$ denoting the conformal time. This gives 
%TODO: define the window function
\begin{equation}
    \langle \psi(\hat{\mathbf n})\psi(\hat{\mathbf n}') \rangle = 4\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\int \frac{\d^3\mathbf k}{(2\pi)^6}\frac{2\pi^2}{k^3}P_\psi(k,\eta_0 - \chi, \eta_0-\chi')e^{i\mathbf k\cdot\hat {\mathbf n}\chi}e^{i\mathbf k\cdot\hat {\mathbf n}'\chi'},
\end{equation}
where we used that $\eta = \eta_0 - \chi$ along the unperturbed photon path (Born approximation), with $\eta_0$ the time at which the light ray hits earth. 
We can use the result
\begin{equation}
    e^{i\mathbf k \cdot \hat {\mathbf n}\chi}=4\pi\sum_{lm}i^lj_l(k\chi )Y_l^m(\hat{\mathbf n})^*Y_l^m(\hat{\mathbf k}) = 4\pi\sum_{lm}i^lj_l(k\chi )Y_l^m(\hat{\mathbf n})Y_l^m(\hat{\mathbf k})^*,
\end{equation}
where $j_l$ is the spherical Bessel function (ELABORATE ON THIS?), to rewrite the above equation. Using both versions of the identity above, we immedatialy get a factor $Y_l^m(\hat{\mathbf k})Y_{l'}^{m'}(\hat{\mathbf k})^*$ in our integral. We can factor the differential element of $\d^3\mathbf k$ into a radial and angular part as $k^2\d k\d^2\Omega_k$, with $\Omega_k$ the solid angle, to apply the orthonormality condition of the spherical harmonics. Note that we take the same sequence of steps a number of times in other parts of the derivations of the lensing spectra. We thus obtain
\begin{gather}
    \langle \psi(\hat{\mathbf n})\psi(\hat{\mathbf n}') \rangle = 4(4\pi)^2\sum_{ll'mm'}i^{l+l'}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\\\times\int \frac{k^2\d k}{(2\pi)^6}\frac{2\pi^2}{k^3}j_l(k\chi)j_{l'}(k\chi')P_\psi(k,\eta_0 - \chi, \eta_0-\chi')Y_{lm}(\hat{\mathbf n})Y_{l'm'}(\hat{\mathbf n}')^*\delta_{ll'}\delta_{mm'}. \label{eq:lenspotcorrexpanded}
\end{gather}
The angular power spectrum is defined similarly to the power spectrum, i.e.
\begin{equation}
    \langle \psi_{lm}\psi_{l'm'}^* \rangle = \delta_{ll'}\delta_{mm'}C_l^\psi.
\end{equation}
Note that the correlation is independent of $m$ and $m'$. %TODO: explain how this relates to statistical isotropy
We can thus read from equation \ref{eq:lenspotcorrexpanded} that
\begin{gather}
    C_l^\psi = 4(4\pi)^2(-1)^{l}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\int \frac{k^2\d k}{(2\pi)^6}\frac{2\pi^2}{k^3}j_l(k\chi)j_{l}(k\chi')P_\psi(k,\eta_0 - \chi, \eta_0-\chi'),
\end{gather}
which can be simplified to
\begin{gather}
    C_l^\psi = (-1)^{l}\frac{2}{\pi^2}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\int k^2\d kj_l(k\chi)j_{l}(k\chi')\frac{P_\psi(k,\eta_0 - \chi, \eta_0-\chi')}{k^3}.
\end{gather}

To further evaluate the integral we will make the standard \textbf{Limber approximation}. The Bessel functions peak sharply at $l=x$\footnote{Some sources use $x\approx l+1/2$ instead, which is slightly more accurate for larger scales (low $l$) and slightly less accurate for smaller scales.}, with the peak being increasingly sharp for higher $l$. Similarly to $\delta(x-x_0)f(x)=\delta(x-x_0)f(x_0)$, we thus take $j_l(k\chi)f(k)\approx j_l(k\chi)f(l/\chi)$. The Bessel functions satisfy an orthogonality condtion,
\begin{equation}
    \int k^2\d k j_l(k\chi)j_l(k\chi') = \frac{\pi}{2\chi^2}\delta(\chi-\chi').
\end{equation}
In combination with the Limber approximation we thus find
\begin{equation}
    \int k^2\d k j_l(k\chi)j_l(k\chi')f(k) \approx \frac{\pi}{2\chi^2}\delta(\chi-\chi')f(l/\chi).
\end{equation}
%TODO: how high? quantify limber approx, perhaps using convergence x LSS paper
It allows us to write the Limber-approximate angular spectrum as
\begin{gather}
    C^\psi(l) = (-1)^{l}\frac{2}{\pi^2}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi') \frac{\pi}{2\chi^2}\delta(\chi-\chi')\frac{\chi^3}{l^3}P_\psi(l/\chi,\eta_0 - \chi, \eta_0-\chi')\\
    = (-1)^{l}\frac{1}{l^3\pi}\int_0^{\chi_*}\chi\d \chi W(\chi)^2 P_\psi(l/\chi,\eta_0 - \chi, \eta_0-\chi).
\end{gather}
% The power spectrum of the potential can be related to the power spectrum of the matter density using the Poisson equation %TODO: create ref
% For a flat universe in atter or dark energy domination we thus find the relation
% \begin{equation}
%     P_\Psi(k;\eta, \eta)=\frac{9\omega_m^2(\eta)H^4(\eta)}{8\pi^2} \frac{P_m(k, \eta)}{k},
% \end{equation}
% where we use the shorthand $P_m(k, \eta):=P_m(k,\eta,\eta)$. The lensing power spectrum is thus finally given in terms of the matter power spectrum as
% \begin{equation}
%     C_l^\psi=\underbrace{9\omega_m^2(\eta)H^4(\eta)}_{:=C(\eta)}\frac{1}{l^4}\int_0^{\chi_*}\d\chi\chi^2 W(\chi)^2P_m(l/\chi, \eta_0-\chi).
% \end{equation}
% We will henceforth denote the $\eta$ dependent factor outside of the integral as $C(\eta)$.
\subsection{Lensing potential bispectrum}
The derivation of the bispectrum proceeds similarly to that of the power spectrum. We aim to compute the bispectrum of the lensing potential fields of 3 (possibly distinct sources), $\psi_1$, $\psi_2$, $\psi_3$. 
\begin{gather*}
    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle = 
    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{l_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}e^{i\mathbf k_i \cdot\hat{\mathbf n_i}\chi_i}} \\
    \times \langle \prod_i\Psi(\mathbf k_i, \eta_0 - \chi_i) \rangle.
\end{gather*}
Defining the bispectrum of the graviational potential as
\begin{equation*}
    \langle \prod_{i=1,2,3}\Psi(\mathbf k_i, \eta_0 - \chi_i) \rangle = (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}),
\end{equation*}
we rewrite the lensing potential bispecrum as
\begin{gather*}
    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle = 
    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{l_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}e^{i\mathbf k_i \cdot\hat{\mathbf n_i}\chi_i}}\\ \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
\end{gather*}
Now using the identity REF for the complex exponential:
\begin{gather*}
    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle \\ = 
    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{l_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi\sum_{lm}i^{l}j_{l}(k_i\chi_i)Y_{l}^{m}(\hat{\mathbf n_i})Y_{l}^{m}(\hat{\mathbf k_i})^*}\\
    \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}) \\
    = \sbr{\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)Y_{l_i}^{m_i}(\hat{\mathbf k_i})^*}} (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
\end{gather*}
We can write the 3D dirac delta function in terms of spherical harmonics as
\begin{equation*}
    \delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3) = 8\int\d^3\mathbf x\prod_{i=1,2,3}\br{ \sum_{l_jm_j} i^{l_j}j_{l_j}(k_ix)Y_{l_j}^{m_j}(\hat{\mathbf k_i}) Y_{l_j}^{m_j}(\hat{\mathbf x})^* }.
\end{equation*}
This results in
\begin{gather*}
    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle= \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)Y_{l_i}^{m_i}(\hat{\mathbf k_i})^*} \\
    \times (2\pi)^38\int\d^3\mathbf x\prod_{i}\br{ \sum_{lm} i^{l}j_{l}(k_ix)Y_{l}^{m}(\hat{\mathbf k_i}) Y_{l}^{m}(\hat{\mathbf x})^* }B^\Psi(\{k_i\}, \{\eta_0-\chi_i\})\\
    = (2\pi)^38\int\d^3\mathbf x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3}}4\pi (-1)^{l_i}j_{l_i}(k_i\chi_i)j_{l_i}(k_ix) Y_{l_i}^{m_i}(\hat{\mathbf x})^*} B^\Psi(\{k_i\}, \{\eta_0-\chi_i\})
\end{gather*}
The angular part of the $\mathbf x$ integral can be evaluated using the identity
\begin{gather*}
    \int \mathrm{d}\Omega_{\hat{n}} \, Y_{l_1 m_1}^*(\hat{x}) Y_{l_2 m_2}^*(\hat{n}) Y_{l_3 m_3}^*(\hat{n}) = (-1)^{m_1 + m_2 + m_3} 
    \int \mathrm{d}\Omega_{\hat{n}} \, Y_{l_1 -m_1}(\hat{n}) Y_{l_2 -m_2}(\hat{n})Y_{l_3 -m_3}(\hat{n}) \\
    = (-1)^{m_1 + m_2 + m_3} 
    \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix}
    \begin{pmatrix} l_1 & l_2 & l_3 \\ -m_1 & -m_2 & -m_3 \end{pmatrix} \equiv A_{\mathbf{l}}^{\mathbf{m}},
\end{gather*}
giving
\begin{gather*}
    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle 
    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3}}4\pi (-1)^{l_i}j_{l_i}(k_i\chi_i)j_{l_i}(k_ix)} \\ \times B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
\end{gather*}
Now applying the Limber approximation again:
\begin{gather*}
    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle 
    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi_i^2}\delta(x-\chi_i)4\pi (-1)^{l_i}} \\ \times B^\Psi(\{l_i/\chi_i\}, \{\eta_0-\chi_i\}) \\
    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int \chi^2\d \chi\prod_i\br{-2 W_i(\chi)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi^2}4\pi (-1)^{l_i}} B^\Psi(\{l_i/\chi\}, \eta_0-\chi).
\end{gather*}
Finally, we aim to rewrite the above in terms of the angular bispectrum of the lensing potential.

The definition for the bispectrum of any set of randomly distributed spherical harmonic components $X^k_{lm}$ is CITE Hu 2000 
\begin{equation*}
    \langle (X_1)_{l_1m_1}(X_2)_{l_2m_2}(X_3)_{x_3m_3} \rangle = \begin{pmatrix}
        l_1 & l_2 & l_3 \\ m_1 & m_2 & m_3
    \end{pmatrix}
    B_{l_1l_2l_3}^{X_1X_2X_3}.
\end{equation*}
Note the independence on $m_i$, this necesarilly follows from statistical isotropy. % TODO: add a short explanation here
If $m_1+m_2+m_3\neq 0$, the associated Wigner-3j symbol vanishes and the bispectrum is set to zero.
Also note that in this definition we have immediately generalized to include cross correlation between different fields $X_1$, $X_2$, $X_3$. This will be relevant later when we look at cross correlations between shear and convergence fields.

Using the above definition and the symmetry property
\begin{equation*}
    \begin{pmatrix}
        l_1 & l_2 & l_3\\
        -m_1 & -m_2 & -m_3
      \end{pmatrix}
      =
      (-1)^{l_1+l_2+l_3}
      \begin{pmatrix}
        l_1 & l_2 & l_3\\
        m_1 & m_2 & m_3
      \end{pmatrix}
\end{equation*}
we find
\begin{gather*}
    B^{\psi_1\psi_2\psi_3}_{l_1l_2l_3}
    = (-1)^{l_1+l_2+l_3}
    \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix}
    (2\pi)^3 8 \\
    \times \int \chi^2\d \chi\prod_i\br{-2 W_i(\chi, \chi_*)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi^2}4\pi (-1)^{l_i}} B^\Psi(\{l_i/\chi\}, \{\eta_0-\chi\}),
\end{gather*}
where we were able to drop the $(-1)^{m_1+m_2+m_3}$ factor due to the bispecrum vanishing if that sum doesn't equal $0$, as mentioned earlier. When all $m_i$ equal zero, the Wigner 3j-symbol gains a number of useful properties % TODO: discuss all of them
In particular, it vanishes if $l_1+l_2+l_3$ is odd, meaning we can drop the $(-1)^{l_1+l_2+l_3}$ factor. Additionally cancelling common factors then gives
\begin{align*}
    B^{\psi_1\psi_2\psi_3}_{l_1l_2l_3}
    =
    -\sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} 8 \int \frac{\d \chi }{\chi^4} W_1(\chi)W_2(\chi)W_3(\chi) B^\Psi(\{l_i/\chi\}, \eta_0-\chi).
\end{align*}

% \begin{gather*}
%     = \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{l_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3/2}}4\pi\sum_{l_i'm_i'}i^{l_i'}j_{l_i'}(k_i\chi_i)Y_{l_i'}^{m_i'}(\hat{\mathbf n_i})Y_{l_i'}^{m_i'}(\hat{\mathbf k_i})^*}\\ 
%     \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3) B_\Psi(k_1, k_2, k_3) \\
%     = \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3/2}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)Y_{l_i}^{m_i}(\hat{\mathbf k_i})^*}\\ 
%     \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3) B_\Psi(k_1, k_2, k_3) \\
%     = \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3/2}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)Y_{l_i}^{m_i}(\hat{\mathbf k_i})^*}\\ 
%     \times (2\pi)^3\sbr{ \int \d^3 \mathbf{x} \prod_i2\sum_{l_i'm_i'}i^{l_i'}j_{l_i'}(k_ix) Y_{l_i'm_i'}(\hat{\mathbf k}_i) Y_{l_i'm_i'}(\hat{\mathbf x})^* } B_\Psi(k_1, k_2, k_3) \\
%     = \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3/2}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)}\\ 
%     \times (2\pi)^3\sbr{ \int \d^3 \mathbf{x} \prod_i2i^{l_i}j_{l_i}(k_ix) Y_{l_im_i}(\hat{\mathbf x})^* } B_\Psi(k_1, k_2, k_3) \\
%     = A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3/2}}4\pi i^{l_i}(2i^{l_i})j_{l_i}(k_i\chi_i)j_{l_i}(k_ix)}
%     (2\pi)^3  B_\Psi(k_1, k_2, k_3) \\
%     = A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\frac{1}{(2\pi)^{3/2}}4\pi i^{l_i}(2i^{l_i})\frac{\pi}{2\chi_i^2}\delta(x-\chi_i)}
%     (2\pi)^3  B_\Psi(l_1/\chi, l_2/\chi, l_3/\chi) \\
%     = A_{\mathbf l}^{\mathbf m}\int_0^{\chi_*} x^2\d x\prod_i\br{-2 W(x, \chi_*)\frac{1}{(2\pi)^{3/2}}4\pi i^{l_i}(2i^{l_i})\frac{\pi}{2x^2}}
%     (2\pi)^3  B_\Psi(l_1/\chi, l_2/\chi, l_3/\chi) \\
%     = A_{\mathbf{l}}^{\mathbf{m}} (-1)^{1 + l_1 + l_2 + l_3} 8 (2\pi)^{9/2} \int_0^{\chi_*} \frac{[W(\chi, \chi_*)]^3}{\chi^4} \, B_\Psi\left( \frac{l_1}{\chi}, \frac{l_2}{\chi}, \frac{l_3}{\chi} \right ) \, \mathrm{d}\chi.
% \end{gather*}

% TODO: mention that the definition of the bispectrum is different form power spectrum of Psi in terms of normalization constant
% TODO: explain 3d dirac delta in terms of spherical harmonics (eq 78 thesis Luna)
% TODO: explain that we used that int d^2 Omega Y Y Y = A_vec l^vec m (eq 82 thesis Luna)
% using that d^3 k = k^2 d k d^2 Omega_k
% TODO: fix this
\subsection{Gravitational potential spectra in terms of matter spectra}
% TODO: fix references and citations 
To convert equations ... and ... to forms that we can directly evaluate we will need to replace the spectra of the graviational potential with matter spectra. Specifically, we define the matter power spectrum in agreement with the CAMB package documentation CITE and the matter bispectrum in agreement with CITE FITTING FUNC PAPER such that we can approximate it in terms of the matter power spectrum using hyperextended perturbation theory. The \textbf{density contrast} is defined as
\begin{equation}
    \delta(\mathbf x) := \frac{\rho(\mathbf x)-\bar \rho}{\bar\rho},
\end{equation}
and the matter spectra are defined in terms of the fourier transformed density contrast $\delta(\mathbf k)$ as 
\begin{gather*}
    \langle \delta(\mathbf k, \eta) \delta(\mathbf k', \eta)^* \rangle = (2\pi)^3\delta(\mathbf k - \mathbf k')P^\delta(\mathbf k, \eta),\\
    \langle \delta(\mathbf k_1, \eta) \delta(\mathbf k_2, \eta) \delta(\mathbf k_3, \eta) \rangle = (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\delta(k_1, k_2, k_3, \eta).
\end{gather*}
The mean matter density of the universe, \( \bar{\rho} \), can be expressed in terms of the matter density parameter \( \Omega_m \) and the critical density \( \rho_c \) as 
$$
\bar{\rho}(\eta) = \frac{3 \Omega_m H_0^2}{8 \pi G}\frac{1}{a(\eta)^3},
$$
where $a(\eta)$ is the only time dependent factor on the right hand side.
In the case of a flat universe ($K=0$) we can thus rewrite the Poisson equation REF (6.41 in modern cosmology) as
\begin{equation}
    -\nabla^2\Psi = 4\pi G a^2\br{\frac{3\Omega_m H_0^2}{8\pi G}\frac{1}{a^3}}\delta = \frac{3\Omega_m H_0^2}{2} \frac{1}{a}\delta \implies \Psi(k, \eta) = \frac{3\Omega_m H_0^2}{2}\frac{1}{a}\frac{\delta(k, \eta)}{k^2},
\end{equation}
where $\Psi(k,\eta)$ and $\delta(k,\eta)$ are functions in Fourier space. For the power- and bispectra we find
\begin{gather*}
    \langle \Psi(\mathbf k, \eta)\Psi^*(\mathbf k',\eta)\rangle = \frac{2\pi^2}{k^3}P^\Psi(k, \eta)\delta(\mathbf k - \mathbf k') \implies P^\Psi(k,\eta) = \frac{1}{k}(9\Omega_m^2H_0^4\pi) \frac{1}{a^2}P^\delta(k,\eta),\\
    \langle \Psi(\mathbf k_1, \eta)\Psi(\mathbf k_2,\eta)\Psi(\mathbf k_3,\eta)\rangle = (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(k_1,k_2,k_3, \eta)\\ 
    \implies B^\Psi(k_1,k_2,k_3, \eta) = \frac{1}{k_1^2k_2^2k_3^2}\br{\frac{3\Omega_m H_0^2}{2}}^3 \frac{1}{a^3}B^\delta(\{k_i\}, \eta).
\end{gather*}
We conclude this section by using the above equations to rewrite the lensing potential spectra (equations REF and REF) in terms of matter spectra 
\begin{empheq}[box=\fbox]{align*}
    P^{\psi_X\psi_Y}_l
    & = (-1)^{l}\frac{9}{l^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)P^\delta(l/\chi,\eta_0-\chi),\\
    B^{\psi_X\psi_Y\psi_Z}_{l_1l_2l_3} &= -\sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{l_1^2l_2^2l_3^2}\Omega_m^3H_0^6\\
    & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{l_i/\chi\}, \eta_0-\chi).
\end{empheq}

% \begin{gather*}
%     \langle \psi_{l_1m_1} \psi_{l_2m_2} \psi_{l_3m_3} \rangle = A_{\mathbf{l}}^{\mathbf{m}} (-1)^{l_1 + l_2 + l_3} 8 (2\pi)^{9/2} \int_0^{\chi_*} W(\chi, \chi_*)^3 \frac{\chi^2}{l_1^2l_2^2l_3^2} \br{\frac{3\Omega_m H^2}{2}}^3|_{\eta=\eta_0-\chi}\\
%     \quad \times B^\delta\br{ \frac{l_1}{\chi}, \frac{l_2}{\chi}, \frac{l_3}{\chi}; \eta_0-\chi } \mathrm{d}\chi, \\
%     C_l^\psi = \frac{72\pi^3}{l^4}\int_0^{\chi_*}\d \chi \chi^2(\Omega_m^2H^4)|_{\eta=\eta_0-\chi} C^\delta_{l/\chi}(\eta_0-\chi) W(\chi)^2.
% \end{gather*}

\subsection{Convergence and shear multipole moments}
\paragraph{Convergence}
% TODO: make notation of nabla nicer/more consistent

Taking the harmonic decomposition of the convergence,
\begin{equation}
    \kappa(\hat{\mathbf n}) = \kappa_{lm}Y_l^m(\hat{\mathbf n}),
\end{equation}
you can relate it to the multipole moments of the lensing potential through CHECK FACTOR OF 2
\begin{equation}
    \kappa_{lm} = \frac{1}{2}(\nabla^2\psi)_{lm} = \frac{l(l+1)}{2}\psi_{lm}
\end{equation}
because the spherical harmonics are eigenfunctions of the Laplacian operator.

\paragraph{Shear}
On the other hand, the shear is given by
\begin{equation}
    \gamma = \gamma_1 + i\gamma_2 = \frac{1}{2}\eth_1(\eth_0\psi)
\end{equation}
where the spin raising operator, $\eth_s$ acts on a spin $s$ function defined on $S^2$ to create a spin $s+1$ function. The above equality is proven explicitely in appendix ... %TODO: write and cite appendix proving equivalence
$\eth_s$ can be written in $(\theta, \phi)$ coordinates as
\begin{equation}
    \eth_s = -\sin^s\theta(\partial_\theta + \frac{i}{\sin\theta})\frac{1}{\sin^s\theta}.
\end{equation}
It is worth explaining that in this context a spin $s$ function referes to a function $_sf(\theta, \phi)$ that transforms under any rotation of coordinates by picking up a phase factor $e^{is\alpha}$, with $\alpha$ the angle of the rotation, i.e.
\begin{equation}
    f'(\theta', \phi') = e^{is\alpha}f(\theta, \phi).
\end{equation}
Shear is actually a spin 2 function, which can immediately be seen by observing that rotating a galaxy image stretched and squeezed through weak lensing by $180$ degrees gives the same stretching and squeezing, i.e. the same shear. It is thus expected that the shear is proportional to $\eth^2\psi$.

The spin weighted spherical harmonics, $_sY_l^m$ are eigenfunctions of the raising/lowering operator, in particular
\begin{equation}
    \eth_sY_l^m = \sqrt{(l-s)(l+s+1)} _{s+1}Y_l^m.
\end{equation}
Decomposing the shear as
\begin{equation}
    \gamma(\theta, \phi) = \gamma_{lm} {}_sY_l^m(\theta,\phi),
\end{equation}
we thus find
\begin{equation}
    \gamma_{lm} = \frac{1}{2}\sqrt{(l-1)(l)(l+1)(l+2)} \psi_{lm}
\end{equation}

\section{Calculation Details}\label{sec:calcdetails}
\subsection{Numerical Considerations}
% I used CAMB, how I took derivative, interpolation, fisher matrix sampling, matter bispectrum in terms of power spectrum

Running cosmological simulations to obtain the matter power spectra and other functions was done with the CAMB package (REF).

The derivatives of the lensing spectra were done numerically using a four point central difference formula, given by
\begin{equation*}
    f'(x) = \frac{-f(x + 2h) + 8f(x+h) - 8 f(x - h) + f(x - 2h)}{12h} + \mathcal O (h^4).
\end{equation*}
$h$ was chosen to be $2.5\%$ of the $x$ value. Testing showed this to generally be the best choise to avoid both the inacuracies from numerical errors (relevant for small $h$) and inacuracies due to the $\mathcal O(h^4)$ error inherent in the four point finite difference formula. The derivative w.r.t. $\sigma_8$ was an exception as it could not be taken by varying it directly and so was calculated implicitly using the chain rule.

To speed up the calculation of the lensing powerspectra and especially the bispectra, an interpolation approach was used. Due to being 3 dimensional, it was not feasible to interpolate the lensing bispectrum directly. We instead chose to interpolate all 1 and 2 dimensional functions that the bispectra are based on and then calculate the bispectra with the interpolated functions instead. This significantly sped up the computation at the cost of small errors in the values in the bispectrum ($\sim 1\%$) when compared with the uninterpolated values.

The time complexity of the Fisher matrix scales with the maximum multipole considered, $l_{max}$, as $\mathcal O (l_{max}^3)$, leading to time issues for realistic values of $l_{max}$. We thus calculated the $l_1 < l_2 < l_3$ part of each Fisher matrix element (see appendix (REF)) by sampling every step of 10 (CHECK IF THIS IS STILL CORRECT BEFORE PUBLISHING) in $l_1$ and $l_2$. Due to the discontinuos nature of the bispectrum (e.g. it vanishes immediately if $l_1 + l_2 + l_3$ is and odd number) we didn't sample $l_3$ values. This was done similarly in (REF).

The matter bispectrum was calculated from the matter powerspectrum using a fitting formula based on perturbation theory \cite{bispfit}. The bispectrum is given by
\begin{equation}
    B_m(k_1, k_2, k_3, \chi) = 2 F_2(k_1, k_2, z) P_m(k_1, z) P_m(k_2, z) + \text{2 perm.}
\end{equation}
where \( P_m \) is the matter power spectrum, and the kernel \( F_2 \) is given by:
\begin{equation}
    F_2(k_1, k_2, z) = \frac{5}{7} a(k_1, z) a(k_2, z) + \frac{k_1^2 + k_2^2}{2 k_1 k_2} b(k_1, z) b(k_2, z) \cos \theta + \frac{2}{7} c(k_1, z) c(k_2, z) \cos^2 \theta.
\end{equation}
The functions \( a(k, z) \), \( b(k, z) \), and \( c(k, z) \) are defined as:
\begin{equation}
    a(k, z) = \frac{1 + \sigma_8^{a_6}(z) \sqrt{0.7} Q(n_{\text{eff}}) (q^{a_1})^{n_{\text{eff}} + a_2}}{1 + (q^{a_1})^{n_{\text{eff}} + a_2}},
\end{equation}
\begin{equation}
    b(k, z) = \frac{1 + 0.2 a_3 (n_{\text{eff}} + 3) (q^{a_7})^{n_{\text{eff}} + 3 + a_8}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_8}},
\end{equation}
\begin{equation}
    c(k, z) = \frac{1 + \left[ \frac{4.5 a_4}{1.5 + (n_{\text{eff}} + 3)^4} \right] (q^{a_5})^{n_{\text{eff}} + 3 + a_9}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_9}}.
\end{equation}
Here, \( Q(n_{\text{eff}}) \) is given by:
\begin{equation}
    Q(x) = \frac{4 - 2x}{1 + 2^{x+1}}.
\end{equation}
The effective spectral index of the linear power spectrum is defined as:
\begin{equation}
    n_{\text{eff}} \equiv \frac{d \ln P_m^{\text{lin}}(k)}{d \ln k}.
\end{equation}
Additionally, \( q \) is given by:
\begin{equation}
    q = \frac{k}{k_{\text{NL}}},
\end{equation}
where \( k_{\text{NL}} \) is the scale at which nonlinearities become significant, satisfying:
\begin{equation}
    4 \pi k_{\text{NL}}^3 P_m^{\text{lin}}(k_{\text{NL}}) = 1.
\end{equation}
The coefficients \( a_i \) are:
\begin{align*}
    a_1 &= 0.484, \quad a_2 = 3.740, \quad a_3 = -0.849, \quad a_4 = 0.392, \\
    a_5 &= 1.013, \quad a_6 = -0.575, \quad a_7 = 0.128, \quad a_8 = -0.722, \quad a_9 = -0.926.
\end{align*}

\subsection{Choosing Experimental Parameters}
% lmin, lmax, galaxy redshift distribution, fiducial values (!), noise levels, 
% for SO, CMB - S4, Euclid, LSST
We chose our fiducial cosmology in line with the results from the Planck mission \cite{planckresults}.
\begin{align*}
    H_0 &= 67.4 \, \text{km/s/Mpc}, && \text{(Hubble constant)} \\
    \Omega_b h^2 &= 0.0224, && \text{(Physical baryon density parameter)} \\
    \Omega_c h^2 &= 0.120, && \text{(Physical cold dark matter density parameter)} \\
    n_s &= 0.965, && \text{(Scalar spectral index)} \\
    A_s &= 2.1 \times 10^{-9}, && \text{(Amplitude of primordial scalar fluctuations)} \\
    \sum m_\nu &= 0.06 \, \text{eV}, && \text{(Sum of neutrino masses)} \\
    w_0 &= -1. && \text{(Dark energy equation of state parameter)}
\end{align*}

For the parameter constraints we took the multipole range to be from 50 ($l_{min}$) to 1000 ($l_{max}$) (INCREASE TO 2000?) (CHECK BEFORE PUBLISHING) in line with measurement ranges from current gen surveys (PROOF BY LOOKING AT ACTUALL EXPERIMENTS).

We will consider multiple possibilities for the other calculation parameters (noise levels and redshift distribution of observed galaxies). This is done to allow us to more accurately model constraints from measurements by the Simons Observatory (SO), CMB-S4, Euclid, and Large Synoptic Survey Telescope (LSST) surveys. The first two measure CMB lensing while the latter two measure lensing shear of galaxies.

The noise levels in the CMB lensing are caused by a variety of factors, see e.g. \cite{cmblensingestimator}. The noise curves for the SO are shown alongside a simulated pure convergence power spectrum curve in figure \ref{fig:lpsplusnoise}. This is the noise source we use for CMB lensing.

\begin{figure}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/spectraplusnoise.pdf}
    \caption{CMB and galaxy lensing potential power spectra with noise. The left panel shows the galaxy lensing potential power spectrum with LSST and Euclid noise, while the right panel shows the CMB lensing potential power spectrum with OS noise.}
    \label{fig:lpsplusnoise}
\end{figure}

For shear measurements, the noise is simpler to model as the primary source is shot noise due to the discrete number of galaxies in each patch of the survey. The mean ellipiticity of the galaxies in a patch is calculated and a deviation from the mean is interpreted as lensing shear. The greater the number of galaxies the more accurately can the mean ellipticity be determined. Choosing the size of the patch is a balancing between signal to noise and resolution. Smaller patches give more uncertain shear measurements while providing better overal resolution for the shear map and thus allowing measurements of shear bi- and powerspectra to higher multipole numbers. The opposite is true for larger patch sizes. The power spectrum of the shear noise is given by
\begin{equation*}
    N^{\gamma\gamma}_{l} = \frac{\sigma^2_{rms}}{\bar n_g}
\end{equation*}
where $\bar n_g$ is the mean number of galaxies observed per sterradian accross the whole survey \cite{weaklensingnotes2020}.
and $\sigma^2_\gamma$ is the intrinsic ellipticity dispersion. 
Euclid and the LSST both have $\sigma_{rms} \approx 0.26$ \cite{euclid1overview} \cite{lsstsciencebook}.
Euclid is expected to measure a galaxy density of approximately 30 galaxies per square arcminute \cite{euclidprep1}
while LSST aims to measure 26 \cite{galaxydenslsst}
The noise levels are graphed alongside a simulated shear power spectrum in figure \ref{fig:lpsplusnoise}. We perform our calculations with Euclid noise levels throughout this paper.
The redshift distribution of the observed galaxies is commonly parameterized as \cite{bartelmann2001weak}
\begin{equation*}
    n(z) \propto z^a \exp\sbr{-\br{\frac{z}{z_0}}^b}.
\end{equation*}
We choose the parameter combination $a = 2$, $b = 3/2$ and $z_0 = 0.64$ which is similar to the expected distributions of Euclid, (which will probe primarily in the 0.2 - 2.6 redshift range \cite{euclidprep10})
and the LSST mission (which has $a \approx 2$, $b \approx 1$, and $z_0 \approx 0.3$ from predictions for the obtained data \cite{lsstsciencebookchapter3}).

\section{Results}\label{sec:results}
\subsection{SNR} 
The signal to noise ratios of the CMB and galaxy lensing power- and bispectra v.s. the maximum multipole moment measured can be found in figure \ref{fig:snrplots}. It can be seen that the CMB bispectrum signal to noise never exceeds 2 even for $l_{\max} = 2000$, meaning that this signal is not likely to be able to contribute to parameter constraints in future surveys. All other quantities should be well detectable for $l_{\max}\geq 500$.
\begin{figure}
    \includegraphics[width=\textwidth]{figures/snrplots.pdf}
    \caption{Signal to noise ratios for galaxy lensing (dashed) and CMB lensing (non-dashes) power- (in red) and bispectra (in blue). The $x$-axis labels the maximum multipole moment measured up to, starting from $l_{\min}=50$.}
    \label{fig:snrplots}
\end{figure}

\subsection{Parameter Constraints}
For the parameter constraints, even when combining all signals, most of the parameters are poorly constrained (with minimum uncertainties much bigger than their fiducial values). The combination that is well constrained is $H_0$ and $\sigma_8$. Alternatively one can also look at $H_0$ and $S_8$, where $S_8$ is defined as 
\begin{equation*}
    S_8 = \sigma_8 \sqrt{\frac{\Omega_c + \Omega_b}{0.3}}
\end{equation*}
and is commonly found to be a particularly well constrained parameter in weak lensing surveys. The results are essentially the same so we will stick with $\sigma_8$. The constraints and covariances are shown in figures \ref{fig:side_by_side} (a) and (b). In particular, you can see in  the off diagonal plot that combining the CMB and galaxy lensing removes approximate degeneracies in $H_0$, $\sigma_8$ space to create a much smaller $1\sigma$ confidence ellipse. When comparing powerspectra to bispectra constraints in figure \ref{fig:S8const} we see a more mild improvement in constraints when combining the different sources of information.
\begin{figure}[htbp]
    \centering
    \begin{subfigure}[t]{0.48\textwidth}
        \centering
        \includegraphics[width=\textwidth]{paramconstraints_sigma8.pdf}
        \caption{Constraints for $H_0$ and $\sigma_8$ for CMB lensing, galaxy lensing, and CMB + galaxy lensing combined. We use the information from both power and bispectra together with $l_{\max}=2000$ in all three cases.
        The confidence ellipses are for $1\sigma$ ($\sim70\%$) certainty.}
        \label{fig:sigma8const}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{0.48\textwidth}
        \centering
        \includegraphics[width=\textwidth]{paramconstraints_sigma8_diffspectra.pdf}
        \caption{Same as Fig.~\ref{fig:sigma8const}, except comparing constraints from only using power spectra, only bispectra, or both. $S_8$ instead of $\sigma_8$. In all three cases we use CMB $and$ galaxy lensing.}
        \label{fig:S8const}
    \end{subfigure}
    \caption{Side-by-side comparison of parameter constraints.}
    \label{fig:side_by_side}
\end{figure}

\section{Discussion and conclusion}\label{sec:discussion}
% \bibliographystyle{alpha} % Choose your citation style
\printbibliography

\appendix

\section{Fisher Matrix Analysis} \label{sec:fisher}
\subsection{Determing uncertainty in experimental parameters}

Fisher matrix formalism is used to find a lower bound on the constraints we can place on experimental parameters. It combines the Cramer-Rao Inequality \cite{casella2002statistical} with the assumption that we have unbiased estimators  following a gaussian distribution \cite{dodelson2020modern}. In particular, denoting the Fisher matrix by $F$ and the parameters as $\theta_i$, it can be shown that
\begin{equation}
    Var_{\theta_i}(\hat\theta_i) \geq \br{F^{-1}}_{\theta_i\theta_j}.
\end{equation}
In the case of $n$ potentially different random variables $x_i$, each with associated mean $\mu_{x_i}$ and error $\sigma_{x_i}$, the Fisher matrix is given as
\begin{equation}
    F_{\theta_i\theta_j} := \sum_{p, q=1}^n \frac{\partial\mu_{x_p}}{\partial\theta_i}(\tilde\theta_k)(\text{Cov}^{-1})_{x_px_q}\frac{\partial\mu_{x_q}}{\partial\theta_j}(\tilde\theta_k),
\end{equation}
where Cov is the covariance matrix associated with the random vector $(x_1, ..., x_n)$, $\text{Cov}_{x_px_q}:=\text{Cov}(x_p,x_q)$.

%%%%%%%%%%%%%%%%%%%%%

\subsection{Fisher matrices for power- bispectra with multiple tracers}
When considering power spectra, the definition of the Fisher matrix immediately gives 
\begin{equation*}
    F_{\alpha\beta} = \sum_{l_{min}\leq l, l' \leq l_{max}}\sum_{[XY][X'Y']} \partial_\alpha C_{l}^{XY} \br{\Cov^{-1}}^{XY, X'Y'}_{l, l'} \partial_\beta C^{X'Y'}_{l'}.
\end{equation*}
Here the covariance matrix is given as
\begin{equation*}
    \Cov^{XY, X'Y'}_{l, l'} = \langle \hat C^{XY}_{l} \hat C^{X'Y'}_{l'} \rangle.
\end{equation*}
The estimator of a power or bispectrum is just the product of the estimators of the appropriate tracers, $\hat C^{XY}_{l} = \hat X(l) \hat Y(l)$. The sum over $[XY]$ and $[X'Y']$ denotes a sum over possible tracer combinations \textit{without} counting permutations of tracer configuration. This is because permutations are not distinct signals, i.e. $\hat X(l) \hat Y(l) = \hat Y(l) \hat X(l)$. In fact, if we were to count these permutations as distinct signals we will get identical columns and/or rows in our covariance matrix making inversion impossible:
\begin{equation*}
    \langle \hat X(l) \hat Y (l) \hat X'(l') \hat Y' (l') \rangle = \langle \hat Y (l) \hat X (l) \hat X'(l') \hat Y' (l') \rangle, \quad \forall X', Y', l'. 
\end{equation*}

To evaluate the covariance matrix we again assume that the estimators are gaussian distributed so that we can do apply a wick contraction, as is commonly done \cite{tegmark1997cosmic}. In this case we get
\begin{equation*}
    \Cov^{XY, X'Y'}_{l, l'} = \delta_{ll'}\br{C^{XX'}_lC^{YY'}_l + C^{XY'}_lC^{YX'}_l}.
\end{equation*}
Under the Gaussian approximation the gaussian matrix thus vanishes except for $3\times 3$ block matrices (in the case of 2 tracers) on the diagonal. The fisher matrix is then
\begin{equation*}
    F_{\alpha\beta} = \sum_{l}\sum_{[XY][X'Y']} \partial_\alpha C_{l}^{XY} \br{\Cov^{-1}_l}^{XY, X'Y'} \partial_\beta C^{X'Y'}_{l}.
\end{equation*} 
$\Cov^{-1}_l$ here denotes the inverse of the block matrix at $l$.

Next, we consider the Fisher matrix for bispectra measurements.
\begin{equation*}
    F_{\alpha\beta} = \sum_{\text{distinct signals}}\sum_{\text{distinct signals prime}} B^{XYZ}_{l_1l_2l_3}\br{\Cov^{-1}}^{XYZ,X'Z'Y'}_{l_1l_2l_3,l_1'l_2'l_3'}B^{X'Y'Z'}_{l_1'l_2'l_3'}.
\end{equation*} 
Counting only distinct signals requires more care compared to the power spectra. The rule is that $B^{XYZ}_{l_1l_2l_3}$ and $B^{X'Y'Z'}_{l_1'l_2'l_3'}$ are not distinct signals if there exists a permutation $\sigma$ that simultaniously maps $X'Y'Z'$ to $XYZ$ and $l_1'l_2'l_3'$ to $l_1l_2l_3$. It turns out that we can write a sum over distinct signals out as
\begin{gather*}
    \sum_{\text{distinct signals}} = \underbrace{\sum_{l_1=l_2=l_3}\sum_{[XYZ]}}_{\text{sum }1} + \underbrace{\sum_{l_1=l_2\neq l_3}\sum_{[XY]Z}}_{\text{sum }2} + \underbrace{\sum_{l_1<l_2<l_3}\sum_{XYZ}}_{\text{sum }3}.
\end{gather*}
With the same definition again for the $[\cdot]$ notation. For example:
\begin{gather*}
    % \{\{XYZ\}|X,Y,Z\in\{\psi_\kappa, \psi_\gamma\}\} = \{\psi_\kappa\psi_\kappa\psi_\kappa, \psi_\kappa\psi_\kappa\psi_\gamma, \psi_\kappa\psi_\gamma\psi_\gamma, \psi_\gamma\psi_\gamma\psi_\gamma\}, \\
    \{[XY]Z|X,Y,Z\in\{\psi_\kappa, \psi_\gamma\}\} = \{\psi_\kappa\psi_\kappa\psi_\kappa, \psi_\kappa\psi_\gamma\psi_\kappa, \psi_\gamma\psi_\gamma\psi_\kappa,\psi_\kappa\psi_\kappa\psi_\gamma, \psi_\kappa\psi_\gamma\psi_\gamma, \psi_\gamma\psi_\gamma\psi_\gamma\}.
\end{gather*}

It follows to show that the sets that these sums sum over form a partition of the set of all distinct signals. Clearly all 3 sets are pairwise disjoint (no common elements) because of the criteria for the $l_i$'s. To show that their union covers the set of distinct signals, start by considering an arbitrary signal. Its $l$ configuration will trivially correspond to exactly one of the three sums. If it is sum 1, then we are free to permute the $XYZ$'s by virtue of the $l$'s being identical so we will be able to match the $XYZ$ configuration to one of the elements of $\{\{XYZ\}\}$. Similarly, if the $l$ configuration corresponds to sum 2, then we are free to permute the $XY$ configuration to match with one of the elements of $\{\{XY\}Z\}$. The $Z$ value does not matter because any $Z$ value is accounted for. For sum 3 we can argue that we can switch around the order of the $l$'s to satisfy $l_1 < l2 < l3$ and the corresponding $XYZ$ configuration will be accounted for in sum 3 because all $XYZ$ combinations are accounted. Finally, it is simple to verify that no distinct signal is counted more than once within each sum.

\paragraph{Covariance matrix}
Every element of the Fisher matrix can be seen as an inner product weighted by the inverse covariance matrix. We can choose how we order the vectors\footnote{The entries are the derivatives of the bispectra}. We organize the vectors according to the sum 1, 2, and 3 parts first. Then by $l$ configuration. Within each $l$ configuration we can choose any ordering for the $XYZ$ configurations. The covariance matrix now becomes a block matrix with each block corresponding to a $l_i$ and $l_i'$ configuration. When wick contracting using the gaussian approximation, every block matrix where $l_1l_2l_3$ is not a permutation of $l_1'l_2'l_3'$ vanishes. The entries of each block matrix are given as
\begin{gather*}
    \br{\Cov_{l_1l_2l_3}}^{XYZ,X'Y'Z'} = C^{XX'}_{l_1}C^{YY'}_{l_2}C^{ZZ'}_{l_3} + \delta_{l_1l_2}C^{XY'}_{l_1}C^{YX'}_{l_2}C^{ZZ'}_{l_3} +\delta_{l_2l_3}C^{XX'}_{l_1}C^{YZ'}_{l_2}C^{ZX'}_{l_3} \\ + \delta_{l_3l_1}C^{XZ'}_{l_1}C^{YY'}_{l_2}C^{ZX'}_{l_3} +\delta_{l_1l_2}\delta_{l_2l_3}\br{C^{XY'}_{l_1}C^{YZ'}_{l_2}C^{ZX'}_{l_3} +C^{XZ'}_{l_1}C^{YX'}_{l_2}C^{ZY'}_{l_3}}.
\end{gather*}
With our ordering this means that the covariance matrix is again a diagonal block matrix, now with blocks of size $4 \times 4$ (sum 1), $6 \times 6$ (sum 2), and $8 \times 8$ (sum 3).

\subsection{Explicit form for inverse covariance matrix}
The Fisher matrix above can be significantly simplified and written as
\begin{equation*}
F_{\alpha\beta} = \sum_{l_1 \leq l_2 \leq l_3} \frac{\mathcal P _{l_1l_2l_3}}{6}
\sum_{XYZ}\sum_{X'Y'Z'} 
\partial_\alpha B^{X Y Z}_{l_1 l_2 l_3} 
(C^{-1})^{X X'}_{l_1}
(C^{-1})^{Y Y'}_{l_2}
(C^{-1})^{Z Z'}_{l_3}
\partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3}
\end{equation*}
where
\begin{equation*}
    C_{l} := \begin{pmatrix}
        C^{\psi_\kappa\psi_\kappa}_l & C^{\psi_\kappa\psi_\gamma}_l \\
        C^{\psi_\kappa\psi_\gamma}_l & C^{\psi_\gamma\psi_\gamma}_l
    \end{pmatrix}
\end{equation*}
and $\mathcal P_{l_1l_2l_3}$ is defined as the number of distinct permutations that can be made with $l_1l_2l_3$. This form was, for example, used in \cite{Kalaja_2021}\footnote{Note that in \cite{Kalaja_2021} this is based on a previous equation summing over \textit{all} $l_i$ (so including permutations of each configuration) which is missing a factor of $1/6$.}.

To show that the above is the same as the formula for the Fisher matrix given earlier, first parition the sum in the same way and collect all terms that fit in the different categories.
\begin{align*}
    F_{\alpha\beta} &= \sum_{l_1=l_2=l_3} \sum_{[XYZ]}\sum_{[X'Y'Z']} \partial_{\alpha} B^{XYZ}_{l_1 l_2 l_3}\br{\frac{\mathcal P _{l_1l_2l_3}}{6}\sum_{d. p. XYZ} \sum_{d. p. X'Y'Z'}(C^{-1})^{X X'}_{l_1}
    (C^{-1})^{Y Y'}_{l_2}
    (C^{-1})^{Z Z'}_{l_3}}
    \partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3} \\
    &+ \sum_{l_1=l_2\neq l_3} \sum_{[XY]Z}\sum_{[X'Y']Z'} \partial_{\alpha} B^{XYZ}_{l_1 l_2 l_3}\br{\frac{\mathcal P _{l_1l_2l_3}}{6}\sum_{d. p. XY} \sum_{d. p. X'Y'}(C^{-1})^{X X'}_{l_1}
    (C^{-1})^{Y Y'}_{l_2}
    (C^{-1})^{Z Z'}_{l_3}}
    \partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3} \\
    &+ \sum_{l_1<l_2<l_3} \sum_{XYZ}\sum_{X'Y'Z'} \partial_{\alpha} B^{XYZ}_{l_1 l_2 l_3}(C^{-1})^{X X'}_{l_1}
    (C^{-1})^{Y Y'}_{l_2}
    (C^{-1})^{Z Z'}_{l_3}
    \partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3},
\end{align*}
where ``$d. p.$'' stands for ``distinct permutations''.
The entries above are then the entries of the inverses of the block matrices mentioned earlier. This can be checked. For example, for the $l_1 = l_2 = l_3$ sum the multiplication of block matrices corresponding to the same $l_i$ configuration can be written as:
\begin{gather*}
    \sum_{[X'Y'Z']} \br{\frac{\mathcal P _{l_1l_2l_3}}{6} \sum_{d. p. XYZ} \sum_{d. p. X'Y'Z'}(C^{-1})^{X X'}_{l}
    (C^{-1})^{Y Y'}_{l}
    (C^{-1})^{Z Z'}_{l}} \br{C^{X'X''}_l C^{Y'Y''}_l C^{Z'Z''}_l + \text{perms } X''Y''Z''} \\
    = \frac{\mathcal P _{l_1l_2l_3}}{6}\sbr{\br{\sum_{d. p. XYZ} \sum_{X'Y'Z'}(C^{-1})^{X X'}_{l}
    (C^{-1})^{Y Y'}_{l}
    (C^{-1})^{Z Z'}_{l}} C^{X'X''}_l C^{Y'Y''}_l C^{Z'Z''}_l} + \text{perms } X''Y''Z'' \\
    = \frac{\mathcal P _{l_1l_2l_3}}{6} \sum_{d. p. XYZ} \delta_{XX''}\delta_{YY''}\delta_{ZZ''} + \text{perms } X''Y''Z''
    = \frac{\mathcal P _{l_1l_2l_3}}{6} \delta_{[XYZ], [X''Y''Z'']} + \text{perms } X''Y''Z''\\
    = \delta_{[XYZ], [X''Y''Z'']} .
\end{gather*}
The sum over the different wick contractions will similarly cancel with the $\mathcal P_{l_1l_2l_3} / 6$ factor for the $l_1 = l_2 \neq l_3$ sum. For the $l_1 < l_2 < l_3$ sum the proof is similar as well except no cancellation is required.

The same type of simplification can be made in the Fisher matrix for the power spectrum, though it doesn't offer any significant benefits compared to our current $3\times 3$ block matrix form.

\subsection{Signal to Noise Ratio (SNR)}
To measure the overal detectability of the lensing spectra, we introduce an overal amplitude of our signal, $A$, with fiducial value 1 as experimental parameter and compute $F_{AA}$. Obviously, $\partial_A\br{A B^{XYZ}_{l_1l_2l_3}}|_{A=1} = B^{XYZ}_{l_1l_2l_3}$, so we find
\begin{equation*}
    \br{\frac{S}{N}}^2 := F_{AA} =
\sum_{XYZ, X'Y'Z'} \sum_{l_1\leq l_2 \leq l_3} \frac{\mathcal P _{l_1l_2l_3}}{6}
B^{X Y Z}_{l_1 l_2 l_3} 
(C^{-1})^{X X'}_{l_1}
(C^{-1})^{Y Y'}_{l_2}
(C^{-1})^{Z Z'}_{l_3}
B^{X' Y' Z'}_{l_1 l_2 l_3}.
\end{equation*}
The equation for the SNR of the power spectra is identical in form.

\subsection{Combined Fisher Matrix}
To compute the Fisher matrix of an experiment measuring both lensing power- and bispectra we are required to compute and invert the full covariance matrix. Under the Gaussian matrix this simplifies trivially. The correlation between the a power- and bispectrum estimator contains an odd (5) amount of fields and thus always vanishes. We are thus allowed to simply add the Fisher matrices of the power- and bispectra to compute the combined Fisher matrix.

\subsection{Fisher Matrices and Eigenvalues}
As explained earlier, inverting the Fisher matrix gives us a covariance matrix of the estimators, $\hat \theta_i$ of our experimental parameters. For any vector $\mathbf v$,
\begin{equation*}
    \mathbf{v}^T_j \Cov_{ij} \mathbf v_i = \langle (v_i \hat \theta_i)^2 \rangle.
\end{equation*}
In particular, if the covariance matrix thus has some large eigenvalue $\lambda_L$ (large compared to the overal accuracy of our experiment), with eigenvector $v_i$, then the corresponding estimator, $v_i\hat \theta_i$, has a large variance and thus represents an approximate degeneracy in our parameters. Clearly, when looking at the Fisher matrix we thus instead look for the smallest eigenvalue.

If we have such an abnormally small eigenvalue we are likely to run into problems with calculating and inverting the Fisher matrix due to numerical errors. These problems are that inversion becomes unstable and that the matrix as a whole might not be PSD because the small eigenvalue was (wrongly) calculated to be negative. These problems are likely because (1) the smallest eigenvalue is sensitive to numerical errors due to being small and (2) the errors on the eigenvalues is proportial to the condition number of the matrix \cite{horn2012matrix}, where the condition number is the ratio between the largest and smallest eigenvalue of the matrix \cite{trefethen1997numerical}. Specifically, for a matrix $A$ with some perturbation $\Delta A$, the corresponding  perturbation of eigenvalue $\lambda_i$ is bounded above as
\begin{equation*}
    |\Delta \lambda_i | \leq ||\Delta A || \kappa(A)|,
\end{equation*}
with $\kappa(A)$ the condition number.

Before inverting a Fisher matrix, it is thus recommended to remove rows and columns corresponding to parameters that are poorly constrained by the experiment, which can be determined by checking for small eigenvalues and their eigenvectors.

% Computing the inverse is then reduced to computing the inverse of the individual block matrices. The block matrices for sum 1 and 2 don't have a particularly nice form (cite paper about multiple tracer fisher matrices for bispectra) but for sum 3 the inverse of the block matrix corresponding to $l_1l_2l_3$ is
% \begin{equation*}
%     (\Cov_{l_1l_2l_3}^{-1})^{XYZ, X'Y'Z'} = \hat C_{l_1}^{X,X'}\hat C_{l_2}^{Y,Y'}\hat C_{l_3}^{Z,Z'}.
% \end{equation*}
% \begin{equation*}
%     (\Cov)^{XYZ, X'Y'Z'} = C_{l_1}^{X,X'} C_{l_2}^{Y,Y'} C_{l_3}^{Z,Z'}.
% \end{equation*}
% As can be seen from
% \begin{equation*}
%     \sum_{X'Y'Z'} = \hat C_{l_1}^{X,X'}\hat C_{l_2}^{Y,Y'}\hat C_{l_3}^{Z,Z'} C_{l_1}^{X'X''}C_{l_2}^{Y'Y''}C_{l_3}^{Z'Z''} = \delta_{XX''}\delta_{YY''}\delta_{ZZ''} = \delta_{XYZ, X''Y''Z''}
% \end{equation*}

% It's worth mentioning that due to all $l$'s being distinct in each configuration of sum 3 the Wick contraction corresponding to each element of the covariance matrix can only be done in one way and we thus only have one term. In the sum 2 we get two terms due to $l_1$ and $l_2$ being identical. In sum 3 we similarly get 6 terms corresponding to all possible Wick contractions being nonzero.

% \subsection{Shear noise}
% In the absence of lensing, galaxies will generally still look elliptical rather than perfectly spherical. This \textbf{intrinsic ellipticity} is a major cause of noise when measuring lensing shear. We can define the ellipticity of a galaxy image (or any type of object, for that matter) in terms of \textbf{quadrupole moments} of the surface brightness as
% \begin{equation*}
% 	e_1 = \frac{I_{xx} - I_{yy}}{I_{xx} + I_{yy}}, \quad e_2 = \frac{2I_{xy}}{I_{xx} + I_{yy}},
% \end{equation*}
% where
% \begin{equation*}
% 	I_{ij} = \int \d x\d y I(x,y)ij, \quad i,j\in\{x,y\},
% \end{equation*}
% and the galaxy is centered on $(x,y)=0$.

% The connection with lensing shear is then made by taylor expanding and using the lensing matrix. If we have unlensed ellipiticities $e_i^\text{true}$ and lensed ellipticities $e_i^{\text{obs}}$, then \cite{dodelson_modern_cosmology}
% \begin{gather*}
% 	e_1^{\text{obs}} = \br{1-2\sbr{e_1^{\text{true}}\gamma_1+e_2^{\text{true}}\gamma_2}}e_1^{\text{true}} + 2\gamma_1,\\
% 	e_2^{\text{obs}} = \br{1-2\sbr{e_1^{\text{true}}\gamma_1+e_2^{\text{true}}\gamma_2}}e_2^{\text{true}} + 2\gamma_2.
% \end{gather*}
% In the case of weak lensing, $\gamma_i, \kappa \ll 1$, thus the above reduces to $\gamma_i \approx (e_i^{\text{obs}} - e^{\text{true}}_i)/2$. The analysis thusfar was quietly done in real space, but taking the harmonic transform of $\gamma_i$ is trivial under the approximation we just made. We thus find the error due to intrinsic ellipticity in harmonic space to equal $ \langle \gamma_{i,l}^2\rangle \approx \langle \br{e_{i,l}^\text{true}}^2\rangle/4$. In a real survey we don't measure the shear for individual galaxies, but instead average it over small patches of the sky. The size of each patch is chosen to balance between signal to noise ratio of the shear measurement and resolution of the shear map. If there are $N$ galaxies observed per patch, then the noise is given by
% $$
% \langle \br{\hat \gamma_{i,l}  - \gamma_{i,l}^{\text{true}}}^2\rangle \approx \frac{\langle \br{e_{i,l}^\text{true}}^2\rangle}{4N}.
% $$
% We don't know the exact value for the right hand side either, so we use the standard estimator for that. There is no correlation between two patches, i.e. the noise is white noise and thus scale independent. Finally, we want to look at the noise in $\gamma_l\gamma_{l'}^*$. We thus get
% \begin{equation*}
% 	C^{\gamma\gamma,\;\text{obs}}_l = C^{\gamma\gamma, \;\text{true}}_l + N_l^{\gamma\gamma},
% \end{equation*} 
% with
% \begin{equation*}
% 	 N_l^{\gamma\gamma} = \frac{1}{4}\frac{e_{\text{RMS}}}{N}, \quad e_{\text{RMS}}:=\sqrt{\frac{1}{N}\sum^N_{i=1}\sbr{(e_{1, i} - \bar e_1)^2 + (e_{2,i}-\bar e_2)^2}}.
% \end{equation*}

% Typically $e_{RMS} / 4 \sim 0.3$ \cite{Kilbinger_2015}.

% The EUCLID team aims for a noise of about 20 percent and thus require at least $\sim 4 \times 10^8$ galaxies measured per patch, giving an absolute error of \cite{2024}
% $$
% N_l^{\gamma\gamma} \approx 8 \times 10^{-10} \quad \text{(EUCLID)}.
% $$

% \subsection{Convergence Noise}
% \subsubsection{Optimal Estimator}
% In practice, the lensing potential cannot be directly measured but is instead estimated based on the temperature, $\Theta$, $E$-mode polarization, $E$, and $B$-mode polarization, $B$, fields.

% Define some axes $x$ and $y$ on a small patch of the sky and let $\mathbf E$ be the electric field of the CMB radiation, then the \textbf{linear polarization Stokes parameters} are defined as
% \begin{equation*}
% 	Q := E_xE^*_x - E_yE_y^*, \quad U := E_xE_y^* + E_yE_x^*.
% \end{equation*} 
% By looking at a sufficiently small patch of the sky we can apply the \textbf{flat sky limit}, in which case we take a 2D Fourier transform over the approximately flat patch instead of working with spherical harmonics. In this case we find $Q(\mathbf k)$ and $U(\mathbf k)$. In Fourier space we can instead work with the more convenient $E$ and $B$ modes, defined as
% \begin{gather*}
% 	E(\mathbf k)=\cos 2\phi_{\mathbf k} Q(\mathbf k) + \sin 2\phi_{\mathbf k} U(\mathbf k),\\
% 	B(\mathbf k)=-\sin 2\phi_{\mathbf k} Q(\mathbf k) + \cos 2\phi_{\mathbf k} U(\mathbf k),
% \end{gather*}
% where $\phi_\textbf{k}$ is the angle of $\mathbf k$ with the $(1,0)$  vector (i.e. $k_{\hat{\mathbf{x}}}$).
% $E$ corresponds to the curl free component of the polarization field and behaves like a scalar under transformations. $B$ corresponds to the rest of the polarization field and behaves like a semi-scalar under transformations. The Fourier components are thus related to the real space fields as
% \begin{gather*}
% \Theta(\mathbf{\hat{n}}) = \int \frac{d^2 k}{(2\pi)^2} \Theta(\mathbf{k}) e^{i \mathbf{k} \cdot \mathbf{\hat{n}}}, \\
% [Q \pm iU](\mathbf{\hat{n}}) = -\int \frac{d^2 k}{(2\pi)^2} \left[ E(\mathbf{k}) \pm iB(\mathbf{k}) \right] e^{\pm 2i \varphi_{\mathbf{k}}} e^{i \mathbf{k} \cdot \mathbf{\hat{n}}}, \\
% \phi(\mathbf{\hat{n}}) = \int \frac{d^2 K}{(2\pi)^2} \phi(\mathbf{K}) e^{i \mathbf{K} \cdot \mathbf{\hat{n}}}.
% \end{gather*}

% As discussed earlier, lensing acts in real space as 
% \begin{equation*}
%     X(\hat n) = \tilde X(\hat {\mathbf n} + \nabla \phi(\hat{\mathbf n})) \approx \tilde X(\hat{\mathbf n}) + \nabla \phi \cdot \nabla \tilde X |_{\hat{\mathbf{n}}},
% \end{equation*}
% with $X$ any of $\Theta$, $Q \pm iU$ and $\tilde X$ the corresponding \textit{unlensed} field. In Fourier space this naturally takes the form of a convolution. The changes in Fourier moments due to lensing are thus given by \cite{Hu_2002}
% \begin{gather*}
%     \delta \Theta(\mathbf{k}) = \int \frac{d^2 k'}{(2\pi)^2} \tilde{\Theta}(\mathbf{k'}) W(\mathbf{k'}, \mathbf{K}), \\[8pt]
%     \delta E(\mathbf{k}) = \int \frac{d^2 k'}{(2\pi)^2} 
%     \left[ \tilde{E}(\mathbf{k'}) \cos 2\varphi_{k'k} - \tilde{B}(\mathbf{k'}) \sin 2\varphi_{k'k} \right] W(\mathbf{k'}, \mathbf{K}), \\[8pt]
%     \delta B(\mathbf{k}) = \int \frac{d^2 k'}{(2\pi)^2} 
%     \left[ \tilde{B}(\mathbf{k'}) \cos 2\varphi_{k'k} + \tilde{E}(\mathbf{k'}) \sin 2\varphi_{k'k} \right] W(\mathbf{k'}, \mathbf{K}),
% \end{gather*}

% \noindent where \( \varphi_{k'k} \equiv \varphi_{k'} - \varphi_k \), \( \mathbf{K} = \mathbf{k} - \mathbf{k'} \), and
% \begin{gather*}
%     W(\mathbf{k}, \mathbf{K}) = -[\mathbf{k} \cdot \mathbf{K}] \phi(\mathbf{K}).
% \end{gather*}

% For power spectra in 2D Fourier space we use the definition
% \begin{equation*}
%     \langle x(\mathbf k)x'(\mathbf k') \rangle = (2\pi)^2\delta(\mathbf k + \mathbf k')C_k^{xx'}
% \end{equation*}
% unless stated otherwise. We will denote unlensed powerspectra with a tilde on top. $C^{xx'}(k)$ denotes the observed power spectrum is related to the true (but still lensed) spectrum as
% \begin{equation*}
%     C^{xx', \text{ obs}}(k) = B(k)^2C^{xx', \text{ true}}(k) + \delta_{xx'}N^x(k).
% \end{equation*}
% We are then able to express the power spectra of the lensed fields in terms of the unlensed fields as
% \begin{equation*}
%     \langle x(\mathbf k)x'(\mathbf k') \rangle = f_{xx'}(\mathbf k, \mathbf k') \phi(\mathbf k + \mathbf k'),
% \end{equation*}
% where $f_{xx'}$ is a function of 2 wavenumbers and unlensed power spectra. Specific values of $f$ for different combinations of $x$ and $x'$ can be found in ... %TODO: fix this
% We can then set up an estimator for $\phi$ using a weighted integral of the Fourier modes measured for $x$ and $x'$ by allowing for a filter function $F_{xx'}(\mathbf k, \mathbf k')$. This gives the estimator
% \begin{equation*}
%     \hat\phi_{xx'}(\mathbf k) = \frac{\int \frac{\d^2k'}{(2\pi)^2}x(\mathbf k')x'(\mathbf k - \mathbf k') F_{xx'}(\mathbf k', \mathbf k - \mathbf k')}{\int \frac{\d^2k'}{(2\pi)^2}f_{xx'}(\mathbf k', \mathbf k - \mathbf k') F_{xx'}(\mathbf k', \mathbf k - \mathbf k')}.
% \end{equation*}
% It's easily checked that $\langle \hat\phi_{xx'} \rangle = \phi$ regardless of the filter function used. Minimizing for the variance of the estimator gives
% \begin{equation}
%     F_{xx'}(\mathbf{k}_1, \mathbf{k}_2) = 
%     \frac{C_{k_1}^{x'x'} C_{k_2}^{xx} f_{xx'}(\mathbf{k}_1, \mathbf{k}_2) 
%     - C_{k_1}^{xx'} C_{k_2}^{xx'} f_{xx'}(\mathbf{k}_2, \mathbf{k}_1)}
%     {C_{k_1}^{xx} C_{k_2}^{x'x'} C_{k_1}^{x'x'} C_{k_2}^{xx} 
%     - \left( C_{k_1}^{xx'} C_{k_2}^{xx'} \right)^2}.
% \end{equation}
% With this estimator we assume that the unlensed powerspectra have been determined through other means (e.g. a simulation).
% With optimal weighting the noise becomes
% \begin{equation*}
%     \langle \hat\phi_{xy}(\mathbf K) \hat\phi_{x'y'}(\mathbf K') \rangle = (2\pi)^2\delta(\mathbf K - \mathbf K') [C^{\phi \phi}_K + K^{-2}N_{xyx'y'}(K)],
% \end{equation*}
% where
% \begin{equation*}
%     N_{xyx'y'}(K):=K^{-2}A_{xy}(K)A_{x'y'}(K)\int\frac{\d^2 k'}{(2\pi)^2}F_{xy}(\mathbf k', \mathbf k - \mathbf k')\br{
%         F_{x'y'}(\mathbf k', \mathbf k - \mathbf k') C_{k'}^{xx'}C^{yy'}_{k - k'} + F_{x'y'}(\mathbf k - \mathbf k', \mathbf k') C_{k'}^{xy'}C^{yx'}_{k - k'}
%     }.
% \end{equation*}

% \subsubsection{Noise}
% We partition the sky into a finite number of pixels of size (in solid angle) $\Omega_{\text{pix}}$. We take into account detector beam width and homogenous noise, but not boundary effects or inhomogeneous noise. %TODO: justify this or take it into account
% In real space the observed field relates to the true (but still lensed) field as
% \begin{equation*}
%     X^{\text{obs}}(\hat{\mathbf n}) = \int \d\Omega' X^{\text{true}}(\hat{\mathbf n}') B(\hat{\mathbf n}, \hat{\mathbf n}') + \eta(\hat{\mathbf n}),
% \end{equation*}
% where $B$ is the beam pattern and $\eta$ is the noise map stemming from any number of experimental sources of error.

% In the case of an isotropic uniform beam, the beam pattern can be written as a function of $\hat{\mathbf n} - \hat{\mathbf n}'$ and the Fourier transform gives 
% \begin{equation*}
%     X^{\text{obs}}(\mathbf k) = X^{\text{true}}(\mathbf k)B(\mathbf k) + \eta(\mathbf k),
% \end{equation*}
% with
% \begin{equation*}
%     B(\mathbf k) = e^{-\frac{k^2\theta_{\text{FWHM}}^2}{4\ln 2}},
% \end{equation*}
% and $\theta_{\text{FWHM}}$ the angle for full width at half minimum.
% By using a likelihood approach we obtain the estimator of the lensed power spectrum given by
% \begin{equation*}
%     \hat C^{XX}(k) = B(\mathbf k)^{-2}\br{X^{\text{obs}}(\mathbf k)X^{\text{obs}}(\mathbf k)^* - N(k)}
% \end{equation*}
% with variance
% \begin{equation*}
%     \Cov(\hat C^{XX}(k), \hat C^{XX}(k')) = 2\sbr{C(k) + N(k)B(k)^{-2}}\delta(\mathbf k - \mathbf k').
% \end{equation*}
% The noise as calculated in eq ... % TODO: ref 
% is thus given by using
% \begin{equation*}
%     C^{XY, \text{ obs}}(k) = \langle \hat C^{XY}(k) \rangle = B(\mathbf k)^{-2}(C)
% \end{equation*}
% \subsection{Power and bispectrum estimators}
% \subsection{Combining experimental and theoretical (shot noise?) uncertainty}
% %Let's say you are looking at a field $X_l$ and have some experimental error $E_l$, uncorrelated with $X_l$ and with mean $0$ (so no systematic error). Then
% %\begin{equation*}
% %	\langle (X+E)(X+E)^*\rangle = \langle X X^*\rangle + \langle EE^* \rangle.
% %\end{equation*}

% In the case of an error expressed as
% \begin{gather*}
% 	\text{``We measure the amplitude of the CMB lensing} \\\text{
% 	power spectrum at state-of-the-art 2.3\% precision, [...]"}
% \end{gather*}
% we can quantify this as
% \begin{equation*}
% 	\hat P^X_l = P_l^X(1+E),
% \end{equation*}
% where $E\sim \mathcal N(0, \sigma^2)$ and in the example above $\sigma = 0.023$. It thus follows that

%TODO: mention SNR

\section{Shear equals twice spin raised lensing potential}\label{sec:shear}

Consider a point on $S^2$, $(r_0, \theta_0, \phi_0)$, at which we observe some cosmological object. We can then define a set of cartesian coordinates $(\tilde r, \tilde y, \tilde x)$ as shown in figure \ref{fig:tildecoords}.

\begin{figure}[h!]
    \centering
    \input{figures/lensing_coords.tikz}
    \caption{
        $(\tilde r, \tilde y, \tilde x)$ coordinates defined for a point on the unit sphere. These act as ordinary cartesian coordinates but rotated such that, at the associated point on $S^2$, $\hat{\tilde r}$ points straight out of the unit sphere, $\hat{\tilde y}$ is parallel to the great arc with constant $\phi$ and $\hat{\tilde x}$ is parallel to the great arc with constant $\theta$. These coordinates are used to define the shear and convergence in terms of the lensing potential.
        }
    \label{fig:tildecoords}
\end{figure}

Note that there it isn't obvious whether to define these coordinates at the point where the lensed light hits $S^2$ or the unlensed light hits $S^2$. We will assume that lensing effects are sufficiently weak that either definition works. We can then express the tilde coordinates in terms of spherical coordinates either by applying a rotation matrix or by calculating the $r, \theta, \phi$ derivatives of $(x,y,z)$ coordinates at $(r_0, \theta_0, \phi_0)$ to find $\hat{\tilde{\mathbf r}}$, $\hat{\tilde{\boldsymbol{\theta}}}$, and $\hat{\tilde{\boldsymbol{\phi}}}$ and then take inner products. Regardless, we find
\begin{align}
    \tilde r &= r\sin\theta\cos\phi\sin\theta_0\cos\phi_0+r\sin\theta\sin\phi\sin\theta_0\sin\phi_0+r\cos\theta\cos\theta_0,\\
    \tilde y &= r\sin\theta\cos\phi\cos\theta_0\cos\phi_0+r\sin\theta\sin\phi\cos\theta_0\sin\phi_0-r\cos\theta\sin\theta_0,\\
    \tilde x &= -r\sin\theta\cos\phi\sin\phi_0+r\sin\theta\sin\phi\cos\phi_0.
\end{align}
This gives derivatives
\begin{align*}
    \frac{\partial}{\partial \theta} = & \left( r \cos \theta \cos \phi \sin \theta_0 \cos \phi_0 + r \cos \theta \sin \phi \sin \theta_0 \sin \phi_0 - r \sin \theta \cos \theta_0 \right) \frac{\partial}{\partial \tilde{r}} \\
    & + \left( r \cos \theta \cos \phi \cos \theta_0 \cos \phi_0 + r \cos \theta \sin \phi \cos \theta_0 \sin \phi_0 + r \sin \theta \sin \theta_0 \right) \frac{\partial}{\partial \tilde{y}} \\
    & + \left( -r \cos \theta \cos \phi \sin \phi_0 + r \cos \theta \sin \phi \cos \phi_0 \right) \frac{\partial}{\partial \tilde{x}}.\\
    \frac{\partial}{\partial \phi} = & \left( -r \sin \theta \sin \phi \sin \theta_0 \cos \phi_0 + r \sin \theta \cos \phi \sin \theta_0 \sin \phi_0 \right) \frac{\partial}{\partial \tilde{r}} \\
& + \left( -r \sin \theta \sin \phi \cos \theta_0 \cos \phi_0 + r \sin \theta \cos \phi \cos \theta_0 \sin \phi_0 \right) \frac{\partial}{\partial \tilde{y}} \\
& + \left( r \sin \theta \sin \phi \sin \phi_0 + r \sin \theta \cos \phi \cos \phi_0 \right) \frac{\partial}{\partial \tilde{x}}.
\end{align*}
Evaluated at our point of interest we obtain $\partial_\theta=\partial_{\tilde y}$ and $\partial_\phi=\sin\theta_0\partial_{\tilde x}$. The second order derivatives can then be obtained using the first order derivatives. We can immediately evaluate them at the point to get
\begin{align*}
    \partial_\phi^2|_{(r_0,\theta_0,\phi_0)}&=-\sin^2\theta_0\partial_{\tilde r}-\sin\theta_0\cos\theta_0\partial_{\tilde y}+\sin^2\theta_0\partial^2_{\tilde x},\\
    \partial_\theta\partial_\phi|_{(r_0,\theta_0,\phi_0)}&=\cos\theta_0\partial_{\tilde x}+\sin\theta_0\partial_{\tilde x}\partial_{\tilde y},\\
    \partial_\theta^2|_{(r_0,\theta_0,\phi_0)}&=-\partial_{\tilde r}+\partial^2_{\tilde y}.
\end{align*}
Thus, at $(r_0, \theta_0, \phi_0)$,
\begin{gather*}
    \frac{1}{2}\eth_1(\eth_0\psi) = \frac{1}{2}\sin\theta(\partial_\theta+\frac{i}{\sin\theta}\partial_\phi)(\frac{1}{\sin\theta}(\partial_\theta+\frac{1}{\sin\theta}\partial_\phi))\\
    =\frac{\partial^2 \psi}{\partial \theta^2} - \frac{\cos\theta}{\sin\theta} \frac{\partial \psi}{\partial \theta} + \frac{2 i}{\sin\theta} \frac{\partial^2 \psi}{\partial \theta \partial \phi} - \frac{1}{\sin^2\theta} \frac{\partial^2 \psi}{\partial \phi^2} - 2 i \frac{\cos\theta}{\sin^2\theta} \frac{\partial \psi}{\partial \phi}\\
    =\frac{1}{2}(\partial_{\tilde y}^2 - \partial_{\tilde x}^2 + 2i\partial_{\tilde x}\partial_{\tilde y})\psi
    =\gamma_1 + i\gamma_2 = \gamma,
\end{gather*}
as required.

% \section{Fisher matrices raw data}

% Bispectra Fisher matrices:\\
% \begin{gather*}
% F^B_{\psi_c} =
% \begin{pmatrix}
% 4\times 10^{-3} & 6\times 10^{-4} & -6\times 10^{-3} & -9\times 10^{-3} & 1\times 10^{-4} & -4\times 10^{-3} & -1\times 10^{-3}\\
% 6\times 10^{-4} & 1\times 10^{-4} & -1\times 10^{-3} & -2\times 10^{-3} & 2\times 10^{-5} & -6\times 10^{-4} & -2\times 10^{-4}\\
% -6\times 10^{-3} & -1\times 10^{-3} & 1\times 10^{-2} & 1\times 10^{-2} & -2\times 10^{-4} & 6\times 10^{-3} & 2\times 10^{-3}\\
% -9\times 10^{-3} & -2\times 10^{-3} & 1\times 10^{-2} & 2\times 10^{-2} & -2\times 10^{-4} & 8\times 10^{-3} & 3\times 10^{-3}\\
% 1\times 10^{-4} & 2\times 10^{-5} & -2\times 10^{-4} & -2\times 10^{-4} & 3\times 10^{-6} & -1\times 10^{-4} & -3\times 10^{-5}\\
% -4\times 10^{-3} & -6\times 10^{-4} & 6\times 10^{-3} & 8\times 10^{-3} & -1\times 10^{-4} & 4\times 10^{-3} & 1\times 10^{-3}\\
% -1\times 10^{-3} & -2\times 10^{-4} & 2\times 10^{-3} & 3\times 10^{-3} & -3\times 10^{-5} & 1\times 10^{-3} & 4\times 10^{-4}
% \end{pmatrix},
% \\
% F^B_{\psi_s} =
% \begin{pmatrix}
% 3\times 10^{-1} & 4\times 10^{-2} & -3\times 10^{-1} & -1\times 10^{-1} & 5\times 10^{-3} & -1\times 10^{-1} & -1\times 10^{-1}\\
% 4\times 10^{-2} & 4\times 10^{-3} & -3\times 10^{-2} & -2\times 10^{-2} & 5\times 10^{-4} & -1\times 10^{-2} & -1\times 10^{-2}\\
% -3\times 10^{-1} & -3\times 10^{-2} & 2\times 10^{-1} & 1\times 10^{-1} & -3\times 10^{-3} & 1\times 10^{-1} & 1\times 10^{-1}\\
% -1\times 10^{-1} & -2\times 10^{-2} & 1\times 10^{-1} & 7\times 10^{-2} & -2\times 10^{-3} & 6\times 10^{-2} & 5\times 10^{-2}\\
% 5\times 10^{-3} & 5\times 10^{-4} & -3\times 10^{-3} & -2\times 10^{-3} & 6\times 10^{-5} & -2\times 10^{-3} & -2\times 10^{-3}\\
% -1\times 10^{-1} & -1\times 10^{-2} & 1\times 10^{-1} & 6\times 10^{-2} & -2\times 10^{-3} & 6\times 10^{-2} & 5\times 10^{-2}\\
% -1\times 10^{-1} & -1\times 10^{-2} & 1\times 10^{-1} & 5\times 10^{-2} & -2\times 10^{-3} & 5\times 10^{-2} & 5\times 10^{-2}
% \end{pmatrix},
% \\
% F^B_{\psi_{s + c}} =
% \begin{pmatrix}
% 4\times 10^{-1} & 4\times 10^{-2} & -3\times 10^{-1} & -2\times 10^{-1} & 5\times 10^{-3} & -2\times 10^{-1} & -1\times 10^{-1}\\
% 4\times 10^{-2} & 4\times 10^{-3} & -3\times 10^{-2} & -2\times 10^{-2} & 6\times 10^{-4} & -2\times 10^{-2} & -1\times 10^{-2}\\
% -3\times 10^{-1} & -3\times 10^{-2} & 2\times 10^{-1} & 2\times 10^{-1} & -4\times 10^{-3} & 1\times 10^{-1} & 1\times 10^{-1}\\
% -2\times 10^{-1} & -2\times 10^{-2} & 2\times 10^{-1} & 1\times 10^{-1} & -3\times 10^{-3} & 8\times 10^{-2} & 7\times 10^{-2}\\
% 5\times 10^{-3} & 6\times 10^{-4} & -4\times 10^{-3} & -3\times 10^{-3} & 7\times 10^{-5} & -2\times 10^{-3} & -2\times 10^{-3}\\
% -2\times 10^{-1} & -2\times 10^{-2} & 1\times 10^{-1} & 8\times 10^{-2} & -2\times 10^{-3} & 7\times 10^{-2} & 6\times 10^{-2}\\
% -1\times 10^{-1} & -1\times 10^{-2} & 1\times 10^{-1} & 7\times 10^{-2} & -2\times 10^{-3} & 6\times 10^{-2} & 5\times 10^{-2}
% \end{pmatrix}.
% \end{gather*}

% Powerspectra Fisher matrices:
% \begin{gather*}
% F^P_{\psi_c} =
% \begin{pmatrix}
% 2\times 10^{-2} & 4\times 10^{-3} & -4\times 10^{-2} & 8\times 10^{-3} & 9\times 10^{-4} & -3\times 10^{-2} & -1\times 10^{-2}\\
% 4\times 10^{-3} & 1\times 10^{-3} & -9\times 10^{-3} & -6\times 10^{-4} & 2\times 10^{-4} & -5\times 10^{-3} & -2\times 10^{-3}\\
% -4\times 10^{-2} & -9\times 10^{-3} & 8\times 10^{-2} & -9\times 10^{-3} & -2\times 10^{-3} & 6\times 10^{-2} & 2\times 10^{-2}\\
% 8\times 10^{-3} & -6\times 10^{-4} & -9\times 10^{-3} & 2\times 10^{-2} & 3\times 10^{-4} & -1\times 10^{-2} & -4\times 10^{-3}\\
% 9\times 10^{-4} & 2\times 10^{-4} & -2\times 10^{-3} & 3\times 10^{-4} & 3\times 10^{-5} & -1\times 10^{-3} & -4\times 10^{-4}\\
% -3\times 10^{-2} & -5\times 10^{-3} & 6\times 10^{-2} & -1\times 10^{-2} & -1\times 10^{-3} & 4\times 10^{-2} & 2\times 10^{-2}\\
% -1\times 10^{-2} & -2\times 10^{-3} & 2\times 10^{-2} & -4\times 10^{-3} & -4\times 10^{-4} & 2\times 10^{-2} & 6\times 10^{-3}
% \end{pmatrix},
% \\
% F^P_{\psi_s} =
% \begin{pmatrix}
% 2\times 10^{-1} & 1\times 10^{-2} & -1\times 10^{-1} & -3\times 10^{-2} & 2\times 10^{-3} & -6\times 10^{-2} & -7\times 10^{-2}\\
% 1\times 10^{-2} & 1\times 10^{-3} & -9\times 10^{-3} & -3\times 10^{-3} & 2\times 10^{-4} & -5\times 10^{-3} & -6\times 10^{-3}\\
% -1\times 10^{-1} & -9\times 10^{-3} & 7\times 10^{-2} & 2\times 10^{-2} & -1\times 10^{-3} & 4\times 10^{-2} & 4\times 10^{-2}\\
% -3\times 10^{-2} & -3\times 10^{-3} & 2\times 10^{-2} & 9\times 10^{-3} & -4\times 10^{-4} & 1\times 10^{-2} & 1\times 10^{-2}\\
% 2\times 10^{-3} & 2\times 10^{-4} & -1\times 10^{-3} & -4\times 10^{-4} & 3\times 10^{-5} & -7\times 10^{-4} & -8\times 10^{-4}\\
% -6\times 10^{-2} & -5\times 10^{-3} & 4\times 10^{-2} & 1\times 10^{-2} & -7\times 10^{-4} & 2\times 10^{-2} & 2\times 10^{-2}\\
% -7\times 10^{-2} & -6\times 10^{-3} & 4\times 10^{-2} & 1\times 10^{-2} & -8\times 10^{-4} & 2\times 10^{-2} & 3\times 10^{-2}
% \end{pmatrix},
% \\
% F^P_{\psi_{s+c}} =
% \begin{pmatrix}
% 9\times 10^{-2} & 9\times 10^{-3} & -7\times 10^{-2} & -1\times 10^{-2} & 1\times 10^{-3} & -4\times 10^{-2} & -4\times 10^{-2}\\
% 9\times 10^{-3} & 1\times 10^{-3} & -9\times 10^{-3} & -2\times 10^{-3} & 2\times 10^{-4} & -5\times 10^{-3} & -4\times 10^{-3}\\
% -7\times 10^{-2} & -9\times 10^{-3} & 7\times 10^{-2} & 8\times 10^{-3} & -1\times 10^{-3} & 5\times 10^{-2} & 3\times 10^{-2}\\
% -1\times 10^{-2} & -2\times 10^{-3} & 8\times 10^{-3} & 2\times 10^{-2} & -7\times 10^{-5} & -3\times 10^{-4} & 5\times 10^{-3}\\
% 1\times 10^{-3} & 2\times 10^{-4} & -1\times 10^{-3} & -7\times 10^{-5} & 3\times 10^{-5} & -9\times 10^{-4} & -6\times 10^{-4}\\
% -4\times 10^{-2} & -5\times 10^{-3} & 5\times 10^{-2} & -3\times 10^{-4} & -9\times 10^{-4} & 3\times 10^{-2} & 2\times 10^{-2}\\
% -4\times 10^{-2} & -4\times 10^{-3} & 3\times 10^{-2} & 5\times 10^{-3} & -6\times 10^{-4} & 2\times 10^{-2} & 2\times 10^{-2}
% \end{pmatrix}.
% \end{gather*}

\end{document}

### 2025-03-25 paper/figures/Fisher_ellipses_plots.py

@@ -1,0 +1,299 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.patches import Ellipse
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.patches import Ellipse
+from scipy.stats import norm
+from numpy.linalg import inv
+
+# # Define fiducial parameter values
+# fiducial_values = {
+#     r"$H_0$": 67.4,
+#     r"$\Omega_b h^2$": 0.0224,
+#     r"$\Omega_c h^2$": 0.120,
+#     r"$A_s$": 2.1e-9,
+#     r"$n_s$": 0.965,
+#     r"$\Sigma m_\nu$": 0.06,
+# }
+
+fiducial_values = {
+    r"$H_0$": 67.4,
+    r"$\Omega_b h^2$": 0.0224,
+    r"$\Omega_c h^2$": 0.120,
+    r"$n_s$": 0.965
+}
+
+fiducial_array = np.array([
+    fiducial_values[param] for param in fiducial_values.keys()
+])
+
+# Function to normalize covariance values
+def normalize_covariance(covariance_matrix, fiducial_array):
+    normalization_matrix = np.outer(fiducial_array, fiducial_array)
+    return (covariance_matrix / normalization_matrix) * 100
+
+# Function to plot normalized confidence ellipses
+def plot_multiple_fisher_ellipses_with_percentages_no_text(fisher_matrices, parameter_names, labels, confidence_level=0.68):
+    num_params = len(parameter_names)
+    chi2_val = {0.68: 2.3, 0.95: 5.99}[confidence_level]
+    colors = ['blue', 'red', 'green', 'purple', 'orange']
+
+    fig, axes = plt.subplots(num_params, num_params, figsize=(12, 12))
+    for i in range(num_params):
+        for j in range(num_params):
+            ax = axes[i, j]
+            if i == j:
+                # Plot 1D Gaussians for each Fisher matrix on the diagonal
+                for k, fisher_matrix in enumerate(fisher_matrices):
+                    covariance_matrix = np.linalg.inv(fisher_matrix)
+                    print([covariance_matrix[i, i] for i in range(num_params)])
+                    covariance_normalized = normalize_covariance(covariance_matrix, fiducial_array)
+                    sigma = np.sqrt(covariance_normalized[i, i]) #!!!
+                    x = np.linspace(-3 * sigma, 3 * sigma, 1000)
[…]

### 2025-03-25 paper/figures/snr_plots.py

@@ -1,0 +1,52 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+# Define data
+ell_max = [250, 500, 750, 1000, 1250, 1500, 1750, 2000]
+
+snrp_c = np.array([15989.588551558505, 31618.168645038517, 39928.86302292515, 
+                     44034.90338600421, 46105.011942758254, 47230.495325004005, 
+                     47860.80819355356, 48209.52024209044])
+
+snrp_s = np.array([22253.427444706474, 66208.20974556397, 112348.39746634613, 
+                     155146.7664808113, 192855.2357456099, 225221.0307281884, 
+                     252637.9760405973, 275744.9966060093])
+
+snrb_c = np.array([0.4662713058660203, 1.7371421603331405, 2.871500243034231, 
+                     3.65500342260037, 4.15133536761963, 4.4659414552800625, 
+                     4.664800647728199, 4.786812407063142])
+
+snrb_s = np.array([34.868772005476124, 263.4320514442226, 685.031116048985, 
+                     1215.6310184788954, 1779.353150346276, 2329.5943353466655, 
+                     2840.639407847829, 3300.6119934544467])
+
+# Set figure size
+plt.figure(figsize=(7,5))
+
+# Plot each array
+plt.plot(ell_max, np.sqrt(snrp_c), marker='x', label='CMB, powerspectrum', linestyle='--', color='blue')
+plt.plot(ell_max, np.sqrt(snrp_s), marker='x', label='galaxy, powerspectrum', color='blue')
+plt.plot(ell_max, np.sqrt(snrb_c), marker='x', label='CMB, bispectrum', linestyle='--', color='red')
+plt.plot(ell_max, np.sqrt(snrb_s), marker='x', label='galaxy, bispectrum', color='red')
+
+plt.minorticks_on()
+plt.xlabel(r'$l_{\max}$', fontsize=12)
+plt.ylabel(r'$S/N$', fontsize=12)
+
+# Increase the x-axis limit to add extra space on the right
+x_max = ell_max[-1]
+plt.xlim(ell_max[0] - 50, x_max + 250)
+
+# Annotate each line with the y-value at the right end
+offset = 50  # horizontal offset for the text
[…]

### 2025-03-25 paper/fisherstuff.py

@@ -1,0 +1,446 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.patches import Ellipse
+from scipy.stats import norm
+from numpy.linalg import inv
+
+np.set_printoptions(
+    precision=1,
+    suppress=False, 
+    formatter={'float_kind': lambda x: f"{x:.2e}"}
+)
+
+######################
+# Bispectra Matrices #
+######################
+visb_c = np.array([
+    [ 3.66134328e-03,  5.64294103e-04, -5.59024163e-03, -8.28197198e-03,  9.21839119e-05, -3.43845273e-03, -1.21009701e-03],
+    [ 5.64294103e-04,  9.39182080e-05, -8.78913206e-04, -1.41245463e-03,  1.42075865e-05, -5.24072512e-04, -1.83365409e-04],
+    [-5.59024163e-03, -8.78913206e-04,  8.58128614e-03,  1.29940930e-02, -1.40723722e-04,  5.23439712e-03,  1.84009658e-03],
+    [-8.28197198e-03, -1.41245463e-03,  1.29940930e-02,  2.15235269e-02, -2.08249724e-04,  7.65644963e-03,  2.67469517e-03],
+    [ 9.21839119e-05,  1.42075865e-05, -1.40723722e-04, -2.08249724e-04,  2.32194050e-06, -8.65918443e-05, -3.04632746e-05],
+    [-3.43845273e-03, -5.24072512e-04,  5.23439712e-03,  7.65644963e-03, -8.65918443e-05,  3.23545639e-03,  1.13828815e-03],
+    [-1.21009701e-03, -1.83365409e-04,  1.84009658e-03,  2.67469517e-03, -3.04632746e-05,  1.13828815e-03,  4.01520985e-04]
+])
+
+visb_s = np.array([
+    [ 1.43008841e+01,  1.58650740e+00, -1.14635424e+01, -8.04509013e+00,  1.95515860e-01, -5.98457532e+00, -5.21205911e+00],
+    [ 1.58650740e+00,  1.78726181e-01, -1.28311279e+00, -9.14981146e-01,  2.18306721e-02, -6.68745848e-01, -5.78208226e-01],
+    [-1.14635424e+01, -1.28311279e+00,  9.23957956e+00,  6.54726869e+00, -1.57349525e-01,  4.81985548e+00,  4.17900899e+00],
[…]

### 2025-03-25 paper/test.py

@@ -1,0 +1,272 @@
+# follows quadratic estimator of https://arxiv.org/pdf/astro-ph/0111606
+
+print('boba')
+
+import numpy as np
+from scipy.integrate import quad, dblquad
+from cosm_setup import lensing_spectra
+from scipy.interpolate import interp1d
+
+# Instantiate your spectra object (which has a camb.results instance in spectra.results)
+spectra = lensing_spectra(fiducial_k_nls=True)
+
+# Get C_l for TT, EE, BB, TE in DIMENSIONLESS UNITS!! up to some lmax
+lmax = 5000
+cls = spectra.results.get_lensed_scalar_cls(lmax=lmax, raw_cl=True)
+
+# cls[:,0] = TT, cls[:,1] = EE, cls[:,2] = BB, cls[:,3] = TE
+ells = np.arange(lmax+1)
+fTT = interp1d(ells, cls[:,0], kind='cubic', bounds_error=False, fill_value=0.0)
+fEE = interp1d(ells, cls[:,1], kind='cubic', bounds_error=False, fill_value=0.0)
+fBB = interp1d(ells, cls[:,2], kind='cubic', bounds_error=False, fill_value=0.0)
+fTE = interp1d(ells, cls[:,3], kind='cubic', bounds_error=False, fill_value=0.0)
+
+def cmbps(l, type1, type2):
+    """
+    Returns C_l in μK^2 for the given combination (type1, type2).
+    type1, type2 can be 'T', 'E', or 'B'.
+    """
+    if (type1, type2) == ('T', 'T'):
+        return fTT(l)
+    elif (type1, type2) == ('E', 'E'):
+        return fEE(l)
+    elif (type1, type2) == ('B', 'B'):
+        return fBB(l)
+    elif (type1, type2) in [('T', 'E'), ('E', 'T')]:
+        return fTE(l)
+    else:
+        # For non-standard combos like TB or EB (normally zero in standard ΛCDM),
+        # just return zero unless you have them computed.
+        return 0.0
+
+arcmintorad = np.pi / 10800 # converts arcminutes to radians
+
+def cmbps_noise(l, type1, type2, sigma, Delta_T, Delta_P):
+    noise = 0
+
+    # units to input:
+    # sigma: arcmin
[…]

### 2025-03-25 paper/test2.py

@@ -1,0 +1,150 @@
+#!/usr/bin/env python3
+import numpy as np
+import camb
+import matplotlib.pyplot as plt
+
+def get_camb_spectra(lmax=3000):
+    """
+    Returns:
+      ell: array of multipoles [0..lmax]
+      unlensed: shape (lmax+1,4) for [TT,EE,BB,TE]
+      lensed:   shape (lmax+1,4) same
+    in muK^2 units.
+    """
+    pars = camb.CAMBparams()
+    pars.set_cosmology(H0=67.5, ombh2=0.022, omch2=0.122, tau=0.06)
+    pars.InitPower.set_params(ns=0.965)
+    pars.set_for_lmax(lmax, lens_potential_accuracy=0)
+    pars.WantCls = True
+    
+    # Unlensed
+    pars.DoLensing = False
+    r1 = camb.get_results(pars)
+    unl = r1.get_cmb_power_spectra(pars, CMB_unit='muK', raw_cl=True)['unlensed_scalar']
+    
+    # Lensed
+    pars.DoLensing = True
+    r2 = camb.get_results(pars)
+    lens = r2.get_cmb_power_spectra(pars, CMB_unit='muK', raw_cl=True)['total']
+    
+    ell = np.arange(lmax+1)
+    return ell, unl[:lmax+1,:], lens[:lmax+1,:]
+
+def add_noise(lensed_cls, beam_arcmin=4.0, Delta_T=1.0, Delta_P=np.sqrt(2.0)):
+    """
+    Add white noise to the lensed spectra in muK^2.
+    beam_arcmin: FWHM in arcmin
+    Delta_T,P: noise in muK·arcmin
+    """
+    obs = np.copy(lensed_cls)
+    lmax = obs.shape[0] - 1
+    from math import log, exp
+    sigma = beam_arcmin*(np.pi/10800.)
+    dT = Delta_T*(np.pi/10800.)
+    dP = Delta_P*(np.pi/10800.)
+    
+    for L in range(2, lmax+1):
+        factor = (L*(L+1)*sigma*sigma)/(8.*log(2.))
+        nTT = dT*dT*np.exp(factor)
+        nPP = dP*dP*np.exp(factor)
+        obs[L,0] += nTT  # TT
+        obs[L,1] += nPP  # EE
+        obs[L,2] += nPP  # BB
+        # TE cross-noise = 0
+    return obs
+
[…]

### 2025-03-26 code/SNR.py
fisher matrices updated

@@ -66,6 +66,6 @@ def all_bs_SNR2():
 
     pass
 
-# ps_SNR2()
+ps_SNR2()
 single_bs_SNR2()
 # all_bs_SNR2()

### 2025-03-26 code/data_importer.pyx

@@ -54,7 +54,6 @@ cdef double lmin_cmbn = ls_cmbn[0]
 cdef int lnum_cmbn = len(ls_cmbn)
 cdef double lmax_cmbn = ls_cmbn[lnum_cmbn - 1]
 
-cdef double[:] cmbn_41sqrt2 = np.abs(np.loadtxt('cmb_noise_files/Ns_sigma4_DeltaT1_DeltaP1.41.txt')) / (ls_cmbn_np_array * (ls_cmbn_np_array + 1))
 cdef double[:] cmbn_301 = np.abs(np.loadtxt('cmb_noise_files/Ns_sigma3_DeltaT0_DeltaP1.txt')) / (ls_cmbn_np_array * (ls_cmbn_np_array + 1))
 cdef double[:] cmbn_106 = np.abs(np.loadtxt('cmb_noise_files/Ns_sigma1_DeltaT0_DeltaP6.txt')) / (ls_cmbn_np_array * (ls_cmbn_np_array + 1))
 

### 2025-04-04 code/bisp_compare_toshiya.py
misc

@@ -11,7 +11,7 @@ spectra = cs.lensing_spectra(
 import numpy as np
 import matplotlib.pyplot as plt
 
-ls = np.arange(10, 2001, 50)
+ls = np.arange(10, 2001, 20)
 
 lbs_eq_ex_tree = np.array([
     spectra.lbs_flat(l, l, l, ('c', 'c', 'c'), model = 'tree') * float(l)**10 / (2 * np.pi)**2 / 8 for l in ls

### 2025-04-04 code/cosm_setup.py

@@ -6,6 +6,7 @@ import scipy
 import scipy.integrate
 import scipy.optimize
 import scipy.interpolate
+from scipy.interpolate import CubicSpline
 
 # np.seterr(all='warn')
 
@@ -86,6 +87,12 @@ class lensing_spectra:
       print('now calculating k_nls')
       self.zs_for_k_nls = np.linspace(0, 8, 101)
       self.k_nls, self.k_nl_errors, self.zs_for_k_nls_used = self.k_nl_exact(self.zs_for_k_nls)
+      
+    self.zs_for_k_nls_used_max = self.zs_for_k_nls_used[-1]
+    print(f'Max z for extended F_2: {self.zs_for_k_nls_used_max}')
+  
+    self._k_nl = CubicSpline(self.zs_for_k_nls_used, self.k_nls, bc_type='not-a-knot')
+
 
   def obtain_results(self):
     '''
@@ -134,6 +141,7 @@ class lensing_spectra:
     self.results = camb.get_results(pars)
 
     print('Cosmological results created.')
+
     
     pass
 
@@ -161,9 +169,9 @@ class lensing_spectra:
 
     return ell, C_phi_phi
   
-  def _k_nl(self, z):
-    # used quadratic interpolation for finer k_nl values but this sometimes gives negative k_nl values resulting in errors
-    return scipy.interpolate.interp1d(self.zs_for_k_nls_used, self.k_nls, kind = 'linear', bounds_error=False, fill_value = 'extrapolate')(z)
+  # def _k_nl(self, z):
+  #   # used quadratic interpolation for finer k_nl values but this sometimes gives negative k_nl values resulting in errors
+  #   return scipy.interpolate.interp1d(self.zs_for_k_nls_used, self.k_nls, kind = 'linear', bounds_error=False, fill_value = 'extrapolate')(z)
 
   def _k_nl_toshiya_like(self, z, k_min=1e-4, k_max=1000.0, num_k_samples=5000):
     """
@@ -301,13 +309,14 @@ class lensing_spectra:
       num_samples = int(kmax / 0.001) # denominator is stepsize in ls, used to be 0.001 which didn't give any issues
 
       ls = np.linspace(0, kmax, num_samples)
[…]

### 2025-04-04 paper/old/paper_old_version/paper_old_vers.tex
paper changes

--- 2025-03-25 paper/paper.tex
@@
-% !TEX root = /Users/jonasfrugte/Desktop/Research_Project/paper/paper.tex
+% !TEX root = /Users/jonasfrugte/Desktop/ResearchProject/paper.tex
 \documentclass[11pt]{article} % Choose a document class, like 'article', 'report', or 'revtex4-2' for scientific papers
@@
 \usepackage{hyperref}         % For hyperlinks in the document
+\usepackage{cite}             % For handling references better
 \usepackage{geometry}         % For setting up page geometry
@@
 \usepackage{booktabs}
-\usepackage{authblk}
 
@@
 \usepackage{xcolor} % For color customization (optional)
-
-\usepackage[backend=biber, style=numeric]{biblatex}
-\addbibresource{sources.bib}
 
@@
 
-\title{\huge \textbf{Parameter Forecasts From CMB Lensing and Galaxy Lensing Power- and Bispectra}}
-
-\author[1]{\textbf{Jonas Frugte}}
-\author[2]{\textbf{P. Daniel Meerburg}}
-
-\affil[1]{Independent Researcher $\infty$, The Netherlands}
-\affil[2]{Van Swinderen Institute for Particle Physics and Gravity, University of Groningen, Nijenborgh 4, 9747 AG Groningen, The Netherlands}
-\affil[ ]{}
-\affil[ ]{\textit{E-mail:} \texttt{jonasfrugte@gmail.com}}
-
-\date{20 February 2025} % Removes the default date
-
 \begin{document}
-\maketitle
-
-% \textbf{
-%     \huge \noindent 
-%     Parameter Forecasts From CMB $\times$ Galaxy Lensing Powerspectra $\times$ Bispectra
-%     \vspace{0.5cm}
-%     \\
-%     \normalsize
-%     Jonas Frugte, Daan Meerburg\\
-%     20 February 2025
-%     }\\
-% \vspace{0.5cm}
+
+\textbf{
+    \Huge \noindent 
+    Parameter Forecasts from Convergence $\times$ Shear Bispectra 
+    \vspace{0.5cm}
+    \\
+    \normalsize
+    Jonas Frugte, Daan Meerburg\\
+    24 October 2024
+    }\\
+\vspace{0.5cm}
 \hrule
@@
 \begin{abstract}
-    (Placeholder text written with chatgpt) We present a Fisher forecast analysis of the constraints on cosmological parameters from measurements of CMB lensing and galaxy weak lensing power- and bispectra.
-Using simulated data consistent with upcoming surveys, we explore the sensitivity of lensing observables to parameters including the Hubble constant \(H_0\), baryon density \(\Omega_b h^2\), dark matter density \(\Omega_c h^2\), scalar spectral index \(n_s\), amplitude of scalar fluctuations \(A_s\), and the sum of neutrino masses \(\sum m_\nu\).
-
-    Our results indicate that combining power- and bispectra improves parameter constraints compared to power spectra alone, reducing uncertainties by a factor of \(\sim X\%\).
-In particular, we forecast constraints of \(\sigma(H_0) = X.X\) km/s/Mpc, \(\sigma(\Omega_b h^2) = X.XXX\), \(\sigma(\Omega_c h^2) = X.XXX\), \(\sigma(n_s) = X.XXXX\), \(\sigma(A_s) = X.XX \times 10^{-9}\), and \(\sigma(\sum m_\nu) = X.XX\) eV.
-These results demonstrate the potential of bispectrum measurements in future surveys such as the Simons Observatory, CMB-S4, LSST, and Euclid.
-
-    Additionally, we investigate the impact of different noise models and survey redshift distributions, finding that improvements in galaxy redshift accuracy significantly enhance constraints on \(\sum m_\nu\).
-We also discuss the degeneracies between parameters and assess their impact using eigenvalue analysis of the Fisher matrix.
-
-    This study highlights the advantages of combining lensing power- and bispectra for cosmological parameter estimation and provides a roadmap for optimizing future surveys to maximize their scientific return.
+    abstract text
 \end{abstract}
+\vspace{0.5cm}
+\hrule
+
+\tableofcontents
 \pagebreak
-%\tableofcontents
-
-% \section{Conventions}
-% Use $\mathbf k$ for wave numbers (Fourier space), $l$ for multipole moments, $m$ for magnetic quantum number.
+\section{Conventions}
+Use $\mathbf k$ for wave numbers (Fourier space), $l$ for multipole moments, $m$ for magnetic quantum number.
 $P^{XY}(k, \eta, \eta')$ and $B^{XYZ}(k_1,k_2,k_3, \eta_1, \eta_2, \eta_3)$ correspond to bi and power spectra of fields $X$, $Y$, and $Z$ in Fourier space.
@@
 In other words,
-% $$
-% P_l^{XY}(\eta):=P_l^{XY}(\eta,\eta), \quad P_l^{XY}:=P_l^{XY}(\eta_{\text{current}}), \quad \text{similarly for all other spectra}.
-% $$
-% If both fields are the same, we may only denote one field, e.g. 
-% $$
-% P_l^{X}(\eta) := P_l^{XX}(\eta).
-% $$
-
-% Our Fourier convention is
-% \begin{align}
-%     &\text{Forward Transform:} & \quad \tilde{f}(k) &= \int_{-\infty}^{\infty} f(x) \, e^{-i k x} \, dx\\
-%     &\text{Inverse Transform:} & \quad f(x) &= \frac{1}{2\pi} \int_{-\infty}^{\infty} \tilde{f}(k) \, e^{i k x} \, dk
-% \end{align}
-% we choose this convention because it is in line with the convention used in the CAMB source code. We work in natural units where $c=1$.
+$$
+P_l^{XY}(\eta):=P_l^{XY}(\eta,\eta), \quad P_l^{XY}:=P_l^{XY}(\eta_{\text{current}}), \quad \text{similarly for all other spectra}.
+$$
+If both fields are the same, we may only denote one field, e.g. 
+$$
+P_l^{X}(\eta) := P_l^{XX}(\eta).
+$$
+
+Our Fourier convention is
+\begin{align}
+    &\text{Forward Transform:} & \quad \tilde{f}(k) &= \int_{-\infty}^{\infty} f(x) \, e^{-i k x} \, dx\\
+    &\text{Inverse Transform:} & \quad f(x) &= \frac{1}{2\pi} \int_{-\infty}^{\infty} \tilde{f}(k) \, e^{i k x} \, dk
+\end{align}
+we choose this convention because it is in line with the convention used in the CAMB source code. We work in natural units where $c=1$.
 
 \section{Introduction}
-
-Weak lensing has been used as a probe to constrain cosmological parameters since the 2000's \cite{kilbinger2015cosmology,Bacon2000,Kaiser2000,Waerbeke2000,Bartelmann2001}.
-A range of upcoming surveys, such as the survey by the Simons Observatory (SO) \cite{Ade2019}, the Legacy Survey of Space and Time (LSST) \cite{Ivezic2019}, and Euclid \cite{Laureijs2011} aim to measure weak lensing of the CMB and of galaxies.
-They will do so much more accurately than previous surveys, enabling tighter constraints on parameters and progressing our understanding of physics as a whole.
-Typically, the main value of interest is the power spectrum (or equivalently, two point function) of the lensing potential.
-This can be measured in terms of the lensing convergence $\kappa$, lensing shear $\gamma$, or lensing potential $\psi$.
-In the weak lensing regime they all contain the same amount of information and can be easily converted into one another, see section \ref{sec:weaklensstats}.
-For the CMB the lensing potential is generally measured directly \cite{Planck2018Lensing}, while with galaxies one instead measures the shear, as this can be deduced from the ellipiticity distribution of the observed galaxies \cite{HoekstraJain2008}.
-
-Even with Gaussian initial conditions, there are significant amounts of non-Gaussianity in the weak lensing signal, especially in galaxy lensing due to it being a probe of the more recent distribution of matter \cite{Bernardeau1997,Takada2003}.
-In addition to looking at the power spectrum, a natural next step is thus to measure the lensing bispectrum.
-This has already been done for signals from galaxies \cite{vanWaerbeke2002}, but no proper detection has been done of the CMB lensing bispectrum.
-Upcoming surveys such as the ones mentioned earlier may detect this signal.
-Therefore it is relevant to know the detectability of, and parameter constraints possible from, all 4 signals: CMB lensing power- and bispectra and galaxy lensing power- and bispectra.
-Notably, this includes looking at cross correlation as well.
-
-In this article we aim to answer this question in the context of experimental parameters similar to those of next generation surveys such as mentioned earlier.
-We are especially interested in seeing if approximate parameter degeneracies can be broken by combining all data.
-Calculating the covariance matrix of unbiased optimal estimators of the cosmological parameters can be done by combining cosmological simulations (in our case we use the CAMB package \cite{Lewis2000}) with a Fisher matrix analysis.
-See figure \ref{fig:flowchart} for an overview of the calculation done in this paper.
-The parameters looked at are the Hubble constant, $H_0$, physical baryon density, $\Omega_b h^2$, physical cold dark matter density, $\Omega_c h^2$, scalar spectral index, $n_s$, amplitude of primordial scalar fluctuations, $A_s$, sum of neutrino masses $\sum m_\nu$, and dark energy equation of state parameter, $w_0$.
-Though, we will find that some of the parameters are poorly constrained even when combining all 4 signals and thus omit them from the analysis in the results.
-As a secondary purpose, this paper also includes a derivation for the Fisher matrix formalism in the non-trivial context of bispectra with multiple tracers.
-The final formula can already be found in the literature (see e.g.\ the appendix of \cite{Kalaja_2021}), but the authors of this paper were not able to find a derivation and hence included it.
-
-The structure of this paper is as follows: we first review the theory of weak lensing and weak lensing statistics in sections \ref{sec:weaklensing} and \ref{sec:weaklensstats}.
-We then provide technical details of the calculation in section \ref{sec:calcdetails}, such as choices for noise and numerical approximations made in the code.
-In section \ref{sec:results} we present the results, consisting of signal to noise ratios and parameter constraints for different combinations of the CMB and galaxy lensing power- and bispectra.
-We conclude in section \ref{sec:discussion}.
-Appendix \ref{sec:fisher} contains a short review of the Fisher matrix formalism and gives a technical derivation of the formula used to calculate each Fisher matrix element in the case of bispectra measurements with multiple tracers.
-Finally, appendix \ref{sec:shear} explains how to relate the shear to the lensing potential in spherical harmonic space.
-
-All code written for this project is uploaded on github at LINK .
 
@@
     \caption{Flowchart of steps involved in calculating the minimum error/uncertainty in cosmological parameter estimates using convergence and shear power- and bi-spectra.}
-    \label{fig:flowchart}
 \end{figure}
 
-\section{Weak Lensing}\label{sec:weaklensing}
+\section{Weak Lensing}
 \subsection{Approximations Made / Regime}
@@
 \end{equation}
-When working in the weak field limit (i.e. small spacetime curvature) the $00$ component of Einsteins equations give us a poisson equation \cite{dodelson2020modern},
+When working in the weak field limit (i.e. small spacetime curvature) the $00$ component of Einsteins equations give us a poisson equation \cite{dodelson_modern_cosmology},
 \begin{equation}
@@
 \begin{equation*}
-    W(\chi) = \int_\chi^{\chi_*}\d\chi' p(\chi')\frac{\chi'-\chi}{\chi'\chi}.
+    W(\chi) = \int_0^{\chi_*}\d\chi' p(\chi')\frac{\chi'-\chi}{\chi'\chi}.
 \end{equation*}
@@
 
-\section{Weak Lensing Statistics}\label{sec:weaklensstats}
+\section{Spherical Harmonics}
+\subsection{definition}
+The \textbf{spherical harmonics} $Y_{l}^{m}(\theta, \phi)$ with $l \geq 0$ and $-l\leq m\leq l$ are defined on a sphere in 3D space ($0\leq\theta\leq \pi$ and $0\leq\phi<2\pi$, also denoted $S^2$).
+They arise naturally when solutions of the laplace, helmholtz, and schrodinger equations in 3D space are decomposed into radial and angular parts, $f(r, \theta, \phi) = f_{\text{spher}}(r)f_{\text{angular}}(\theta, \phi)$.
+They can be written explicitely as
+\begin{equation}
+    Y^m_l(\theta, \phi) := \sqrt{\frac{2l+1}{4\pi}\frac{(l-m)!}{(l+m)!}}P_l^m(\cos\theta)e^{im\phi},
+\end{equation}
+where $P_l^m$ are \textbf{associated legendre functions}. They are defined in terms of the \textbf{legendre functions} as
+\begin{equation}
+    P_l^m(x) = \begin{cases}
+        (1-x^2)^{m/2}\frac{\d^m}{\d x^m}P_l(x), & m \geq 0,\\
+        (-1)^m\frac{(l+m)!}{(l-m)!}P_l^{-m}(x), & m < 0.
+    \end{cases}
+\end{equation}
+Finally, the legendre functions themselves are implicitly defined through the generating function
+% TODO: find source
+\begin{equation}
+    G(x, t) = \frac{1}{\sqrt{1 - 2xt + t^2}} = \sum_{l=0}^\infty P_l(x)t^l \implies P_l(x) = \frac{1}{l!}\frac{\d ^l}{\d t^l}G(x, 0).
+\end{equation}
+Here we require $x$ to be a real number between $-1$ and $1$ and $t$ a complex number with magnitude less than 1.
+
+\subsection{orthonormality}
+Defining the inner product
+\begin{equation}
+    \langle f(\theta, \phi)|g(\theta, \phi)\rangle := \int_0^{2\pi}\d\phi\int_0^\pi\sin\theta\d\theta f(\theta, \phi)^*g(\theta, \phi)
+\end{equation}
+on the space of functions from $S^2$ to $\mathbb C$, we can show that the spherical harmonics are orthonormal, i.e.
+\begin{equation}
+    \langle Y^{m'}_{l'} | Y^{m}_{l} \rangle = \delta_{mm'}\delta_{ll'}.
+\end{equation}
+
+We first show orthogonality of the associated legendre functions. They satisfy the differential equation
+\begin{equation}
+    \frac{\d}{\d x}((1-x^2)\frac{\d P^m_l(x)}{\d x}) + (l(l + 1) - \frac{m^2}{1 - x^2})P^m_l(x) = 0,
+\end{equation}
+on the interval $[-1, 1]$, which is of \textbf{Sturm-Liouville form},
+\begin{equation}
+    \frac{\d}{\d x}(p(x)\frac{\d y}{\d x}) + (-\lambda - q(x))y = 0.
+\end{equation}
+$\lambda$ is refered to as the eigenvalue of $y$ in this case. Some texts generalize the $\lambda$ term to $\lambda w(x)$, but we will not need these extra complications here. %TODO: find a ref
+It is a general result for $y_n$ and $y_m$ with distinct eigenvalues $\lambda_n$ and $\lambda_m$ that they are orthogonal.
+Consider the space of real functions on (wlog) $[-1, 1]$.
+We can define an inner product
+\begin{equation}
+    \langle f | g \rangle := \int_{-1}^1 f(x)g(x)\d x.
+\end{equation}
+In this case the sturm Liouville problem can be restated as finding the eigenvalues of a differential operator $\mathcal L$, i.e.
+\begin{equation}
+\mathcal L [y] = \lambda, \quad \mathcal L[y] := \frac{\d}{\d x}(p(x)\frac{\d y}{\d x}) - q(x)y.
+\end{equation}
+It can then be easily verified that $\mathcal L$ is a hermitian operator, i.e.
+$\langle \mathcal L f | g \rangle = \langle f | \mathcal L g \rangle $ if $p(x)$ vanishes on the boundaries.
+% TODO: verify the boundary part
+It is then a standard result that eigenvalues of the hermitian operator with distinct eigenvalues are orthogonal.
+Given that the $P^m_l$ are eigenfunctions of a sturm liouvilly problem with associated eigenvalues $l(l+1)$, they must be orthogonal.
+In the case that they are equal, we the integral can be analytically calculated, giving the result
+\begin{equation}
+    \int_{-1}^1 P^m_l(x)P^m_{l'}(x) \d x = \frac{2(l+m)!}{(2l+1)(l-m)!}\delta_{ll'}.
+\end{equation}
+Orthonormality of the spherical harmonics is than checked by using the definition in terms of the associated legendre functions. Define the normalisation constant of $Y_m^l$ as $N^m_l$, then
+\begin{gather}
+    \int \d\theta\d \phi \sin\theta (Y^m_l)^* Y^{m'}_{l'} = N_l^mN_{l'}^{m'} \int_0^{2\pi} e^{-im\phi}e^{im'\phi}\int_0^\pi P^m_l(\cos\theta)P^{m'}_{l'}(\cos\theta)\sin\theta\d\theta \\
+    = (N_l^m)^2 \delta_{mm'}\frac{2(l+m)!}{(2l+1)(l-m)!}\delta_{ll'} = \delta_{mm'}\delta_{ll'}.
+\end{gather}
+
+\subsection{completenes}
+Take an arbitrary function $f(\theta, \phi)$.
+Define harmonic coefficients $c_{lm} = \langle Y_l^m | f \rangle$, then we aim to prove that, for continuous $f$, $f(\theta, \phi) = \sum_{lm} c_{lm}Y^m_l(\theta\phi)$.
+We first require 2 intermediate results.
+The legendre addition formula is a well known resulting stating that
+\begin{equation}
+    P_l(\cos\gamma) = \frac{4\pi}{2l + 1} \sum_{-l\leq m \leq l} Y^m_l(\theta', \phi')^* Y^m_l(\theta, \phi),
+\end{equation}
+where $\gamma$ is the angle between the vectors pointing in the directions $(\theta', \phi')$ and $(\theta, \phi)$. Additionally, in E W Thompson % TODO: get proper ref
+it is shown that, for points at which $f$ is continuous,
+\begin{equation}
+   f(\theta, \phi) = \sum_{l\geq 0} \frac{2l + 1}{4\pi} \int^\pi_0\int^{2\pi}_{0}f(\theta',\phi')P_l(\cos\gamma)\sin\theta'\d\theta'\d\phi'.
+\end{equation}
+Putting these two results together we write 
+\begin{gather}
+    \sum_{lm} c_{lm}Y^m_l(\cos\theta\phi) = \sum_{lm} \int \d\phi'\d\theta'\sin\theta'Y^m_l(\theta',\phi')^*f(\theta',\phi')Y^m_l(\theta,\phi) \\
+    = \sum_l \int \d\theta'\d\phi'\sin\theta' P_l(\cos\gamma)\frac{2l+1}{4\pi}f(\theta',\phi') = f(\theta, \phi).
+\end{gather}
+Thus the spherical harmonics form a complete orthonormal basis of the real functions on $S^2$
+
+\section{Weak Lensing Statistics}
 \subsection{Lensing Potential Powerspectrum}
@@
 
-\section{Calculation Details}\label{sec:calcdetails}
-\subsection{Numerical Considerations}
-% I used CAMB, how I took derivative, interpolation, fisher matrix sampling, matter bispectrum in terms of power spectrum
-
-Running cosmological simulations to obtain the matter power spectra and other functions was done with the CAMB package (REF).
-
-The derivatives of the lensing spectra were done numerically using a four point central difference formula, given by
-\begin{equation*}
-    f'(x) = \frac{-f(x + 2h) + 8f(x+h) - 8 f(x - h) + f(x - 2h)}{12h} + \mathcal O (h^4).
-\end{equation*}
-$h$ was chosen to be $2.5\%$ of the $x$ value.
-Testing showed this to generally be the best choise to avoid both the inacuracies from numerical errors (relevant for small $h$) and inacuracies due to the $\mathcal O(h^4)$ error inherent in the four point finite difference formula.
-The derivative w.r.t.
-$\sigma_8$ was an exception as it could not be taken by varying it directly and so was calculated implicitly using the chain rule.
-
-To speed up the calculation of the lensing powerspectra and especially the bispectra, an interpolation approach was used.
-Due to being 3 dimensional, it was not feasible to interpolate the lensing bispectrum directly.
-We instead chose to interpolate all 1 and 2 dimensional functions that the bispectra are based on and then calculate the bispectra with the interpolated functions instead.
-This significantly sped up the computation at the cost of small errors in the values in the bispectrum ($\sim 1\%$) when compared with the uninterpolated values.
-
-The time complexity of the Fisher matrix scales with the maximum multipole considered, $l_{max}$, as $\mathcal O (l_{max}^3)$, leading to time issues for realistic values of $l_{max}$.
-We thus calculated the $l_1 < l_2 < l_3$ part of each Fisher matrix element (see appendix (REF)) by sampling every step of 10 (CHECK IF THIS IS STILL CORRECT BEFORE PUBLISHING) in $l_1$ and $l_2$.
-Due to the discontinuos nature of the bispectrum (e.g.
-it vanishes immediately if $l_1 + l_2 + l_3$ is and odd number) we didn't sample $l_3$ values.
-This was done similarly in (REF).
-
-The matter bispectrum was calculated from the matter powerspectrum using a fitting formula based on perturbation theory \cite{bispfit}. The bispectrum is given by
-\begin{equation}
-    B_m(k_1, k_2, k_3, \chi) = 2 F_2(k_1, k_2, z) P_m(k_1, z) P_m(k_2, z) + \text{2 perm.}
-\end{equation}
-where \( P_m \) is the matter power spectrum, and the kernel \( F_2 \) is given by:
-\begin{equation}
-    F_2(k_1, k_2, z) = \frac{5}{7} a(k_1, z) a(k_2, z) + \frac{k_1^2 + k_2^2}{2 k_1 k_2} b(k_1, z) b(k_2, z) \cos \theta + \frac{2}{7} c(k_1, z) c(k_2, z) \cos^2 \theta.
-\end{equation}
-The functions \( a(k, z) \), \( b(k, z) \), and \( c(k, z) \) are defined as:
-\begin{equation}
-    a(k, z) = \frac{1 + \sigma_8^{a_6}(z) \sqrt{0.7} Q(n_{\text{eff}}) (q^{a_1})^{n_{\text{eff}} + a_2}}{1 + (q^{a_1})^{n_{\text{eff}} + a_2}},
-\end{equation}
-\begin{equation}
-    b(k, z) = \frac{1 + 0.2 a_3 (n_{\text{eff}} + 3) (q^{a_7})^{n_{\text{eff}} + 3 + a_8}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_8}},
-\end{equation}
-\begin{equation}
-    c(k, z) = \frac{1 + \left[ \frac{4.5 a_4}{1.5 + (n_{\text{eff}} + 3)^4} \right] (q^{a_5})^{n_{\text{eff}} + 3 + a_9}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_9}}.
-\end{equation}
-Here, \( Q(n_{\text{eff}}) \) is given by:
-\begin{equation}
-    Q(x) = \frac{4 - 2x}{1 + 2^{x+1}}.
-\end{equation}
-The effective spectral index of the linear power spectrum is defined as:
-\begin{equation}
-    n_{\text{eff}} \equiv \frac{d \ln P_m^{\text{lin}}(k)}{d \ln k}.
-\end{equation}
-Additionally, \( q \) is given by:
-\begin{equation}
-    q = \frac{k}{k_{\text{NL}}},
-\end{equation}
-where \( k_{\text{NL}} \) is the scale at which nonlinearities become significant, satisfying:
-\begin{equation}
-    4 \pi k_{\text{NL}}^3 P_m^{\text{lin}}(k_{\text{NL}}) = 1.
-\end{equation}
-The coefficients \( a_i \) are:
-\begin{align*}
-    a_1 &= 0.484, \quad a_2 = 3.740, \quad a_3 = -0.849, \quad a_4 = 0.392, \\
-    a_5 &= 1.013, \quad a_6 = -0.575, \quad a_7 = 0.128, \quad a_8 = -0.722, \quad a_9 = -0.926.
-\end{align*}
-
-\subsection{Choosing Experimental Parameters}
-% lmin, lmax, galaxy redshift distribution, fiducial values (!), noise levels, 
-% for SO, CMB - S4, Euclid, LSST
-We chose our fiducial cosmology in line with the results from the Planck mission \cite{planckresults}.
-\begin{align*}
-    H_0 &= 67.4 \, \text{km/s/Mpc}, && \text{(Hubble constant)} \\
-    \Omega_b h^2 &= 0.0224, && \text{(Physical baryon density parameter)} \\
-    \Omega_c h^2 &= 0.120, && \text{(Physical cold dark matter density parameter)} \\
-    n_s &= 0.965, && \text{(Scalar spectral index)} \\
-    A_s &= 2.1 \times 10^{-9}, && \text{(Amplitude of primordial scalar fluctuations)} \\
-    \sum m_\nu &= 0.06 \, \text{eV}, && \text{(Sum of neutrino masses)} \\
-    w_0 &= -1. && \text{(Dark energy equation of state parameter)}
-\end{align*}
-
-For the parameter constraints we took the multipole range to be from 50 ($l_{min}$) to 1000 ($l_{max}$) (INCREASE TO 2000?) (CHECK BEFORE PUBLISHING) in line with measurement ranges from current gen surveys (PROOF BY LOOKING AT ACTUALL EXPERIMENTS).
-
-We will consider multiple possibilities for the other calculation parameters (noise levels and redshift distribution of observed galaxies).
-This is done to allow us to more accurately model constraints from measurements by the Simons Observatory (SO), CMB-S4, Euclid, and Large Synoptic Survey Telescope (LSST) surveys.
-The first two measure CMB lensing while the latter two measure lensing shear of galaxies.
-
-The noise levels in the CMB lensing are caused by a variety of factors, see e.g.
-\cite{cmblensingestimator}.
-The noise curves for the SO are shown alongside a simulated pure convergence power spectrum curve in figure \ref{fig:lpsplusnoise}.
-This is the noise source we use for CMB lensing.
-
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=\textwidth]{figures/spectraplusnoise.pdf}
-    \caption{CMB and galaxy lensing potential power spectra with noise.
-The left panel shows the galaxy lensing potential power spectrum with LSST and Euclid noise, while the right panel shows the CMB lensing potential power spectrum with OS noise.}
-    \label{fig:lpsplusnoise}
-\end{figure}
-
-For shear measurements, the noise is simpler to model as the primary source is shot noise due to the discrete number of galaxies in each patch of the survey.
-The mean ellipiticity of the galaxies in a patch is calculated and a deviation from the mean is interpreted as lensing shear.
-The greater the number of galaxies the more accurately can the mean ellipticity be determined.
-Choosing the size of the patch is a balancing between signal to noise and resolution.
-Smaller patches give more uncertain shear measurements while providing better overal resolution for the shear map and thus allowing measurements of shear bi- and powerspectra to higher multipole numbers.
-The opposite is true for larger patch sizes.
-The power spectrum of the shear noise is given by
-\begin{equation*}
-    N^{\gamma\gamma}_{l} = \frac{\sigma^2_{rms}}{\bar n_g}
-\end{equation*}
-where $\bar n_g$ is the mean number of galaxies observed per sterradian accross the whole survey \cite{weaklensingnotes2020}.
-and $\sigma^2_\gamma$ is the intrinsic ellipticity dispersion. 
-Euclid and the LSST both have $\sigma_{rms} \approx 0.26$ \cite{euclid1overview} \cite{lsstsciencebook}.
-Euclid is expected to measure a galaxy density of approximately 30 galaxies per square arcminute \cite{euclidprep1}
-while LSST aims to measure 26 \cite{galaxydenslsst}
-The noise levels are graphed alongside a simulated shear power spectrum in figure \ref{fig:lpsplusnoise}. We perform our calculations with Euclid noise levels throughout this paper.
-The redshift distribution of the observed galaxies is commonly parameterized as \cite{bartelmann2001weak}
-\begin{equation*}
-    n(z) \propto z^a \exp\sbr{-\br{\frac{z}{z_0}}^b}.
-\end{equation*}
-We choose the parameter combination $a = 2$, $b = 3/2$ and $z_0 = 0.64$ which is similar to the expected distributions of Euclid, (which will probe primarily in the 0.2 - 2.6 redshift range \cite{euclidprep10})
-and the LSST mission (which has $a \approx 2$, $b \approx 1$, and $z_0 \approx 0.3$ from predictions for the obtained data \cite{lsstsciencebookchapter3}).
-
-\section{Results}\label{sec:results}
-\subsection{SNR} 
-The signal to noise ratios of the CMB and galaxy lensing power- and bispectra v.s.
-the maximum multipole moment measured can be found in figure \ref{fig:snrplots}.
-It can be seen that the CMB bispectrum signal to noise never exceeds 2 even for $l_{\max} = 2000$, meaning that this signal is not likely to be able to contribute to parameter constraints in future surveys.
-All other quantities should be well detectable for $l_{\max}\geq 500$.
-\begin{figure}
-    \includegraphics[width=\textwidth]{figures/snrplots.pdf}
-    \caption{Signal to noise ratios for galaxy lensing (dashed) and CMB lensing (non-dashes) power- (in red) and bispectra (in blue).
-The $x$-axis labels the maximum multipole moment measured up to, starting from $l_{\min}=50$.}
-    \label{fig:snrplots}
-\end{figure}
-
-\subsection{Parameter Constraints}
-For the parameter constraints, even when combining all signals, most of the parameters are poorly constrained (with minimum uncertainties much bigger than their fiducial values).
-The combination that is well constrained is $H_0$ and $\sigma_8$.
-Alternatively one can also look at $H_0$ and $S_8$, where $S_8$ is defined as
-\begin{equation*}
-    S_8 = \sigma_8 \sqrt{\frac{\Omega_c + \Omega_b}{0.3}}
-\end{equation*}
-and is commonly found to be a particularly well constrained parameter in weak lensing surveys.
-The results are essentially the same so we will stick with $\sigma_8$.
-The constraints and covariances are shown in figures \ref{fig:side_by_side} (a) and (b).
-In particular, you can see in  the off diagonal plot that combining the CMB and galaxy lensing removes approximate degeneracies in $H_0$, $\sigma_8$ space to create a much smaller $1\sigma$ confidence ellipse.
-When comparing powerspectra to bispectra constraints in figure \ref{fig:S8const} we see a more mild improvement in constraints when combining the different sources of information.
-\begin{figure}[htbp]
-    \centering
-    \begin{subfigure}[t]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{paramconstraints_sigma8.pdf}
-        \caption{Constraints for $H_0$ and $\sigma_8$ for CMB lensing, galaxy lensing, and CMB + galaxy lensing combined.
-We use the information from both power and bispectra together with $l_{\max}=2000$ in all three cases.
-        The confidence ellipses are for $1\sigma$ ($\sim70\%$) certainty.}
-        \label{fig:sigma8const}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{paramconstraints_sigma8_diffspectra.pdf}
-        \caption{Same as Fig.~\ref{fig:sigma8const}, except comparing constraints from only using power spectra, only bispectra, or both.
-$S_8$ instead of $\sigma_8$.
-In all three cases we use CMB $and$ galaxy lensing.}
-        \label{fig:S8const}
-    \end{subfigure}
-    \caption{Side-by-side comparison of parameter constraints.}
-    \label{fig:side_by_side}
-\end{figure}
-
-\section{Discussion and conclusion}\label{sec:discussion}
-% \bibliographystyle{alpha} % Choose your citation style
-\printbibliography
-
-\appendix
-
-\section{Fisher Matrix Analysis} \label{sec:fisher}
-\subsection{Determing uncertainty in experimental parameters}
-
-Fisher matrix formalism is used to find a lower bound on the constraints we can place on experimental parameters.
-It combines the Cramer-Rao Inequality \cite{casella2002statistical} with the assumption that we have unbiased estimators  following a gaussian distribution \cite{dodelson2020modern}.
-In particular, denoting the Fisher matrix by $F$ and the parameters as $\theta_i$, it can be shown that
-\begin{equation}
+\section{Cosmological Perturbation Theory}
+
+\section{Fisher Matrix Analysis}
+\subsection{Determining experimental parameters}
+Let's say we are conduction some experiment with parameters $p_i$ and we make a set of observations $d_j$. Then the \textbf{likelihood function}, $\mathcal L$, is defined as
+\begin{equation}
+    \mathcal(d_j|p_i) := P(d_j|p_i).
+\end{equation}
+We of course want to determine the reverse, i.e. the probability distribution of the parameters given some set of observations. Bayes' theorem tells us that
+\begin{equation}
+    P(w_i|d_j)=\frac{P(d_j|w_i)P(w_i)}{P(d_j)}.
+\end{equation}
+We can ignore the denominator (which we don't know) and get that
+\begin{equation}
+    P(w_i|d_j)\propto\mathcal L (d_j|w_i)P_{\text{prior}}(w_i).
+\end{equation}
+with the proportianality constant of course being determined by requiring the pdf to integrate to 1.
+The probability distribution of the parameters without any measurements is called the \textbf{prior probability} and is usually taken to be a uniform distribution, i.e.
+we can ignore it.
+If we have prior information about the constants, e.g.
+some minimum and maximum values of the parameters, then we can always incorporate this into our results by changing the prior function.
+The pdf on the left hand side of the above equation is also called the \textbf{posterior for }$w_i$.
+The parameters are then by finding the maximum of the probability distribution, i.e.
+by solving
+\begin{equation}
+    \frac{\partial \mathcal L}{\partial w_i}|_{w_{i,\text{ best fit}}}=0.
+\end{equation}
+If, for example, we only have one parameters $w$ and it represents the average of the data, we obtain the familiar value $w=\frac{1}{n}\sum_{i=1}^n d_i$.
+\subsection{Determing uncertainty in experimental parameters (Fisher matrix analysis)}
+% proof of cramer rao bound from "statistical inference" from george casella, Roger l berger
+The \textbf{Cramer-Rao Inequality} is based on the \textbf{Cauchy-Schwarz inequality} ($|Cov(X,Y)|^2\leq Var(X)Var(Y)$) and gives us a lower bound on the variance of an estimator.
+Specifically, let $X_1, ..., X_n$ be a sample with pdf $f(\mathbf x | \theta)$ and let $W(\mathbf X)=W(X_1, ..., X_n)$ be a sufficiently well behaved estimator.
+Then
+\begin{equation}
+    Var_\theta(W(\mathbf X)) \geq \frac{\br{\frac{\d}{\d\theta}E_\theta \sbr{ W(\mathbf X) }}^2}{E_\theta\sbr{\br{\frac{\partial}{\partial\theta}\log f(\mathbf X | \theta)}^2}}.
+\label{eq:cramerraobound}
+\end{equation}
+
+Now let's assume the following: we have only one parameter $\theta$, $W=\hat\theta$ is an \textbf{unbiased estimator} of $\theta$, we conduct $n$ measurements, and measurements are going to equal $\mu(\theta)$ for $\mu$ as some function of $\theta$ plus some parameter independent gaussian error $\sigma$, i.e.
+$$f(x|\theta)\propto e^{-\frac{-(x-\mu(\theta))^2}{2\sigma^2}}.$$
+In this case we can rewrite equation \eqref{eq:cramerraobound} as
+\begin{gather}
+    Var_\theta(\hat\theta(\mathbf X)) 
+    \geq \br{\frac{\d}{\d\theta}\theta}^2\br{E_\theta\sbr{\br{\sum_{i=1}^n \frac{(x_i-\mu(\theta))}{\sigma^2}\frac{\partial\mu(\theta)}{\partial\theta}}^2}}^{-1}
+    = \br{n\frac{\br{\frac{\partial\mu(\theta)}{\partial\theta}}^2}{\sigma^2}}^{-1}.
+\end{gather}
+A few comments are in order.
+First notice that the variance of the estimator of $\theta$ will in general depend on the true value of $\theta$.
+Given that we don't know the true value with certainty, the error is estimated by setting $\theta=\tilde\theta$, which is the most accurate estimate that we have, called the \textbf{fiducial value} of $\theta$.
+Second, the not so elegant form that we left the result in is on purpose.
+In the case of $m$ parameters $\theta_i$, $\mu$ will become a $\mathbb R^m \rightarrow \mathbb R$ function and we can generalize the above procedure to define the $m\times m$ \textbf{Fisher matrix} as
+\begin{equation}
+    F_{\theta_i\theta_j} := n\frac{\frac{\partial\mu}{\partial\theta_i}(\tilde\theta_k)\frac{\partial\mu}{\partial\theta_j}(\tilde\theta_k)}{\sigma^2},
+\end{equation}
+such that \begin{equation}
     Var_{\theta_i}(\hat\theta_i) \geq \br{F^{-1}}_{\theta_i\theta_j}.
 \end{equation}
-In the case of $n$ potentially different random variables $x_i$, each with associated mean $\mu_{x_i}$ and error $\sigma_{x_i}$, the Fisher matrix is given as
+To further generalize to the case of $n$ potentially different random variables $x_i$, each with associated $\mu_{x_i}$ error $\sigma_{x_i}$, we need to take into account that these RV's may be correlated and thus define the Fisher matrix as
 \begin{equation}
@@
 where Cov is the covariance matrix associated with the random vector $(x_1, ..., x_n)$, $\text{Cov}_{x_px_q}:=\text{Cov}(x_p,x_q)$.
-
-%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Fisher matrices for power- bispectra with multiple tracers}
-When considering power spectra, the definition of the Fisher matrix immediately gives 
-\begin{equation*}
-    F_{\alpha\beta} = \sum_{l_{min}\leq l, l' \leq l_{max}}\sum_{[XY][X'Y']} \partial_\alpha C_{l}^{XY} \br{\Cov^{-1}}^{XY, X'Y'}_{l, l'} \partial_\beta C^{X'Y'}_{l'}.
-\end{equation*}
-Here the covariance matrix is given as
-\begin{equation*}
-    \Cov^{XY, X'Y'}_{l, l'} = \langle \hat C^{XY}_{l} \hat C^{X'Y'}_{l'} \rangle.
-\end{equation*}
-The estimator of a power or bispectrum is just the product of the estimators of the appropriate tracers, $\hat C^{XY}_{l} = \hat X(l) \hat Y(l)$.
-The sum over $[XY]$ and $[X'Y']$ denotes a sum over possible tracer combinations \textit{without} counting permutations of tracer configuration.
-This is because permutations are not distinct signals, i.e.
-$\hat X(l) \hat Y(l) = \hat Y(l) \hat X(l)$.
-In fact, if we were to count these permutations as distinct signals we will get identical columns and/or rows in our covariance matrix making inversion impossible:
-\begin{equation*}
-    \langle \hat X(l) \hat Y (l) \hat X'(l') \hat Y' (l') \rangle = \langle \hat Y (l) \hat X (l) \hat X'(l') \hat Y' (l') \rangle, \quad \forall X', Y', l'. 
-\end{equation*}
-
-To evaluate the covariance matrix we again assume that the estimators are gaussian distributed so that we can do apply a wick contraction, as is commonly done \cite{tegmark1997cosmic}.
-In this case we get
-\begin{equation*}
-    \Cov^{XY, X'Y'}_{l, l'} = \delta_{ll'}\br{C^{XX'}_lC^{YY'}_l + C^{XY'}_lC^{YX'}_l}.
-\end{equation*}
-Under the Gaussian approximation the gaussian matrix thus vanishes except for $3\times 3$ block matrices (in the case of 2 tracers) on the diagonal. The fisher matrix is then
-\begin{equation*}
-    F_{\alpha\beta} = \sum_{l}\sum_{[XY][X'Y']} \partial_\alpha C_{l}^{XY} \br{\Cov^{-1}_l}^{XY, X'Y'} \partial_\beta C^{X'Y'}_{l}.
-\end{equation*} 
-$\Cov^{-1}_l$ here denotes the inverse of the block matrix at $l$.
-
-Next, we consider the Fisher matrix for bispectra measurements.
-\begin{equation*}
-    F_{\alpha\beta} = \sum_{\text{distinct signals}}\sum_{\text{distinct signals prime}} B^{XYZ}_{l_1l_2l_3}\br{\Cov^{-1}}^{XYZ,X'Z'Y'}_{l_1l_2l_3,l_1'l_2'l_3'}B^{X'Y'Z'}_{l_1'l_2'l_3'}.
-\end{equation*} 
-Counting only distinct signals requires more care compared to the power spectra.
-The rule is that $B^{XYZ}_{l_1l_2l_3}$ and $B^{X'Y'Z'}_{l_1'l_2'l_3'}$ are not distinct signals if there exists a permutation $\sigma$ that simultaniously maps $X'Y'Z'$ to $XYZ$ and $l_1'l_2'l_3'$ to $l_1l_2l_3$.
-It turns out that we can write a sum over distinct signals out as
-\begin{gather*}
-    \sum_{\text{distinct signals}} = \underbrace{\sum_{l_1=l_2=l_3}\sum_{[XYZ]}}_{\text{sum }1} + \underbrace{\sum_{l_1=l_2\neq l_3}\sum_{[XY]Z}}_{\text{sum }2} + \underbrace{\sum_{l_1<l_2<l_3}\sum_{XYZ}}_{\text{sum }3}.
-\end{gather*}
-With the same definition again for the $[\cdot]$ notation. For example:
-\begin{gather*}
-    % \{\{XYZ\}|X,Y,Z\in\{\psi_\kappa, \psi_\gamma\}\} = \{\psi_\kappa\psi_\kappa\psi_\kappa, \psi_\kappa\psi_\kappa\psi_\gamma, \psi_\kappa\psi_\gamma\psi_\gamma, \psi_\gamma\psi_\gamma\psi_\gamma\}, \\
-    \{[XY]Z|X,Y,Z\in\{\psi_\kappa, \psi_\gamma\}\} = \{\psi_\kappa\psi_\kappa\psi_\kappa, \psi_\kappa\psi_\gamma\psi_\kappa, \psi_\gamma\psi_\gamma\psi_\kappa,\psi_\kappa\psi_\kappa\psi_\gamma, \psi_\kappa\psi_\gamma\psi_\gamma, \psi_\gamma\psi_\gamma\psi_\gamma\}.
+This result is sufficiently general for our purposes and is what we will be using.
+\subsection{Fisher Matrices and Eigenvalues}
+As explained earlier, inverting the Fisher matrix gives us a (best case scenario) covariance matrix of the estimators of our experimental parameters. For any vector $\mathbf v$,
+\begin{equation*}
+    \mathbf{v}^T_j \Cov_{ij} \mathbf v_i = \langle (v_i \hat p_i)^2 \rangle.
+\end{equation*}
+In particular, if the covariance matrix thus has some very large eigenvalue $\lambda_L$ (large compared to the overal accuracy of our experiment) then our the corresponding parameter $v_i\hat p_i$ has a large variance and thus represents an approximate degeneracy in our parameters.
+Of course when looking at the Fisher matrix we instead look for the smallest eigenvalue.
+
+Another consideration related to the above is that if we have such an abnormally small eigenvalue we are likely to run into problems with inverting the Fisher matrix due in a context with numerical errors.
+For one, we are likely to have an eigenvalue go negative, which leads to a covariance matrix that is not positive semi definite (PSD) and will thus give unusable results.
+This is because (1) the smallest eigenvalue is sensitive to numerical errors due to being small (2) the errors on the eigenvalues is proportial to the condition number of the matrix \cite{Horn2012MatrixAnalysis}, specifically for a matrix $A$ with some perturbation $\Delta A$, the perturbation of eigenvalue $\lambda_i$ is bounded above as
+\begin{equation*}
+    |\Delta \lambda_i | \leq ||\Delta A || \kappa(A).
+\end{equation*}
+Here $\kappa(A)$ is the \textbf{condition number} defined as the largest eigenvalue divided by the smallest eigenvalue. If a ``small'' eigenvalue exists the condition number will thus be ``large''.
+
+\subsection{Adapting fisher matrix to our use case}
+Literally applying definition REF gives our fisher matrix element $F_{\alpha\beta}$, with $\alpha$ and $\beta$ parameters of our cosmology, as
+\begin{equation*}
+    F_{\alpha\beta} = \sum_{\text{distinct signals}}\sum_{\text{distinct signals prime}} B^{XYZ}_{l_1l_2l_3}\br{\Cov^{-1}}^{XYZ,X'Z'Y'}_{l_1l_2l_3,l_1'l_2'l_3'}B^{X'Y'Z'}_{l_1'l_2'l_3'}
+\end{equation*}
+here the sum over $XYZ$ and $l_1l_2l_3$ as well as over $X'Y'Z'$ and $l_1'l_2'l_3'$ is over distinct combinations, i.e.
+applying the same permutation on, e.g., $XYZ$ and $l_1l_2l_3$ yields a new term that isn't counted because it's not actually a distinct signal.
+
+The sum can be broken up as
+\begin{gather*}
+    \sum_{\text{distinct signals}} = \underbrace{\sum_{l_1=l_2=l_3}\sum_{\{XYZ\}}}_{\text{sum }1} + \underbrace{\sum_{l_1=l_2\neq l_3}\sum_{\{XY\}Z}}_{\text{sum }2} + \underbrace{\sum_{l_1<l_2<l_3}\sum_{XYZ}}_{\text{sum }3}.
+\end{gather*}
+Here we use set notation $\{\cdot\}$ to denote that we don't count a permutation of these two indices as a distinct signal. In other words,
+\begin{gather*}
+    \{\{XYZ\}|X,Y,Z\in\{\psi_\kappa, \psi_\gamma\}\} = \{\psi_\kappa\psi_\kappa\psi_\kappa, \psi_\kappa\psi_\kappa\psi_\gamma, \psi_\kappa\psi_\gamma\psi_\gamma, \psi_\gamma\psi_\gamma\psi_\gamma\}, \\
+    \{\{XY\}Z|X,Y,Z\in\{\psi_\kappa, \psi_\gamma\}\} = \{\psi_\kappa\psi_\kappa\psi_\kappa, \psi_\kappa\psi_\gamma\psi_\kappa, \psi_\gamma\psi_\gamma\psi_\kappa,\psi_\kappa\psi_\kappa\psi_\gamma, \psi_\kappa\psi_\gamma\psi_\gamma, \psi_\gamma\psi_\gamma\psi_\gamma\}.
 \end{gather*}
@@
 Clearly all 3 sets are pairwise disjoint (no common elements) because of the criteria for the $l_i$'s.
-To show that their union covers the set of distinct signals, start by considering an arbitrary signal.
-Its $l$ configuration will trivially correspond to exactly one of the three sums.
+To show that their union covers the set of distinct signals, consider any signal.
+It's $l$ configuration will trivially correspond to exactly one of the three sums.
 If it is sum 1, then we are free to permute the $XYZ$'s by virtue of the $l$'s being identical so we will be able to match the $XYZ$ configuration to one of the elements of $\{\{XYZ\}\}$.
 Similarly, if the $l$ configuration corresponds to sum 2, then we are free to permute the $XY$ configuration to match with one of the elements of $\{\{XY\}Z\}$.
-The $Z$ value does not matter because any $Z$ value is accounted for.
-For sum 3 we can argue that we can switch around the order of the $l$'s to satisfy $l_1 < l2 < l3$ and the corresponding $XYZ$ configuration will be accounted for in sum 3 because all $XYZ$ combinations are accounted.
-Finally, it is simple to verify that no distinct signal is counted more than once within each sum.
+The $Z$ value doesn't matter because any $Z$ value is accounted for.
+For sum 3 we can argue that we can switch around the order of the $l$'s to satisfy $l_1 < l2 < l3$ and the corresponding $XYZ$ configuration will be in summed over set because all $XYZ$ combinations are counted here.
+Finally it is easy to see that distinct signal is counted more than once.
 
 \paragraph{Covariance matrix}
-Every element of the Fisher matrix can be seen as an inner product weighted by the inverse covariance matrix.
-We can choose how we order the vectors\footnote{The entries are the derivatives of the bispectra}.
-We organize the vectors according to the sum 1, 2, and 3 parts first.
-Then by $l$ configuration.
-Within each $l$ configuration we can choose any ordering for the $XYZ$ configurations.
-The covariance matrix now becomes a block matrix with each block corresponding to a $l_i$ and $l_i'$ configuration.
-When wick contracting using the gaussian approximation, every block matrix where $l_1l_2l_3$ is not a permutation of $l_1'l_2'l_3'$ vanishes.
-The entries of each block matrix are given as
-\begin{gather*}
-    \br{\Cov_{l_1l_2l_3}}^{XYZ,X'Y'Z'} = C^{XX'}_{l_1}C^{YY'}_{l_2}C^{ZZ'}_{l_3} + \delta_{l_1l_2}C^{XY'}_{l_1}C^{YX'}_{l_2}C^{ZZ'}_{l_3} +\delta_{l_2l_3}C^{XX'}_{l_1}C^{YZ'}_{l_2}C^{ZX'}_{l_3} \\ + \delta_{l_3l_1}C^{XZ'}_{l_1}C^{YY'}_{l_2}C^{ZX'}_{l_3} +\delta_{l_1l_2}\delta_{l_2l_3}\br{C^{XY'}_{l_1}C^{YZ'}_{l_2}C^{ZX'}_{l_3} +C^{XZ'}_{l_1}C^{YX'}_{l_2}C^{ZY'}_{l_3}}.
-\end{gather*}
-With our ordering this means that the covariance matrix is again a diagonal block matrix, now with blocks of size $4 \times 4$ (sum 1), $6 \times 6$ (sum 2), and $8 \times 8$ (sum 3).
-
-\subsection{Explicit form for inverse covariance matrix}
-The Fisher matrix above can be significantly simplified and written as
-\begin{equation*}
-F_{\alpha\beta} = \sum_{l_1 \leq l_2 \leq l_3} \frac{\mathcal P _{l_1l_2l_3}}{6}
-\sum_{XYZ}\sum_{X'Y'Z'} 
-\partial_\alpha B^{X Y Z}_{l_1 l_2 l_3} 
-(C^{-1})^{X X'}_{l_1}
-(C^{-1})^{Y Y'}_{l_2}
-(C^{-1})^{Z Z'}_{l_3}
-\partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3}
-\end{equation*}
-where
-\begin{equation*}
-    C_{l} := \begin{pmatrix}
-        C^{\psi_\kappa\psi_\kappa}_l & C^{\psi_\kappa\psi_\gamma}_l \\
-        C^{\psi_\kappa\psi_\gamma}_l & C^{\psi_\gamma\psi_\gamma}_l
-    \end{pmatrix}
-\end{equation*}
-and $\mathcal P_{l_1l_2l_3}$ is defined as the number of distinct permutations that can be made with $l_1l_2l_3$.
-This form was, for example, used in \cite{Kalaja_2021}\footnote{Note that in \cite{Kalaja_2021} this is based on a previous equation summing over \textit{all} $l_i$ (so including permutations of each configuration) which is missing a factor of $1/6$.}.
-
-To show that the above is the same as the formula for the Fisher matrix given earlier, first parition the sum in the same way and collect all terms that fit in the different categories.
-\begin{align*}
-    F_{\alpha\beta} &= \sum_{l_1=l_2=l_3} \sum_{[XYZ]}\sum_{[X'Y'Z']} \partial_{\alpha} B^{XYZ}_{l_1 l_2 l_3}\br{\frac{\mathcal P _{l_1l_2l_3}}{6}\sum_{d.
-p.
-XYZ} \sum_{d.
-p.
-X'Y'Z'}(C^{-1})^{X X'}_{l_1}
-    (C^{-1})^{Y Y'}_{l_2}
-    (C^{-1})^{Z Z'}_{l_3}}
-    \partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3} \\
-    &+ \sum_{l_1=l_2\neq l_3} \sum_{[XY]Z}\sum_{[X'Y']Z'} \partial_{\alpha} B^{XYZ}_{l_1 l_2 l_3}\br{\frac{\mathcal P _{l_1l_2l_3}}{6}\sum_{d. p. XY} \sum_{d. p. X'Y'}(C^{-1})^{X X'}_{l_1}
-    (C^{-1})^{Y Y'}_{l_2}
-    (C^{-1})^{Z Z'}_{l_3}}
-    \partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3} \\
-    &+ \sum_{l_1<l_2<l_3} \sum_{XYZ}\sum_{X'Y'Z'} \partial_{\alpha} B^{XYZ}_{l_1 l_2 l_3}(C^{-1})^{X X'}_{l_1}
-    (C^{-1})^{Y Y'}_{l_2}
-    (C^{-1})^{Z Z'}_{l_3}
-    \partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3},
-\end{align*}
-where ``$d. p.$'' stands for ``distinct permutations''.
-The entries above are then the entries of the inverses of the block matrices mentioned earlier.
-This can be checked.
-For example, for the $l_1 = l_2 = l_3$ sum the multiplication of block matrices corresponding to the same $l_i$ configuration can be written as:
-\begin{gather*}
-    \sum_{[X'Y'Z']} \br{\frac{\mathcal P _{l_1l_2l_3}}{6} \sum_{d. p. XYZ} \sum_{d. p. X'Y'Z'}(C^{-1})^{X X'}_{l}
-    (C^{-1})^{Y Y'}_{l}
-    (C^{-1})^{Z Z'}_{l}} \br{C^{X'X''}_l C^{Y'Y''}_l C^{Z'Z''}_l + \text{perms } X''Y''Z''} \\
-    = \frac{\mathcal P _{l_1l_2l_3}}{6}\sbr{\br{\sum_{d. p. XYZ} \sum_{X'Y'Z'}(C^{-1})^{X X'}_{l}
-    (C^{-1})^{Y Y'}_{l}
-    (C^{-1})^{Z Z'}_{l}} C^{X'X''}_l C^{Y'Y''}_l C^{Z'Z''}_l} + \text{perms } X''Y''Z'' \\
-    = \frac{\mathcal P _{l_1l_2l_3}}{6} \sum_{d. p. XYZ} \delta_{XX''}\delta_{YY''}\delta_{ZZ''} + \text{perms } X''Y''Z''
-    = \frac{\mathcal P _{l_1l_2l_3}}{6} \delta_{[XYZ], [X''Y''Z'']} + \text{perms } X''Y''Z''\\
-    = \delta_{[XYZ], [X''Y''Z'']} .
-\end{gather*}
-The sum over the different wick contractions will similarly cancel with the $\mathcal P_{l_1l_2l_3} / 6$ factor for the $l_1 = l_2 \neq l_3$ sum.
-For the $l_1 < l_2 < l_3$ sum the proof is similar as well except no cancellation is required.
-
-The same type of simplification can be made in the Fisher matrix for the power spectrum, though it doesn't offer any significant benefits compared to our current $3\times 3$ block matrix form.
-
-\subsection{Signal to Noise Ratio (SNR)}
-To measure the overal detectability of the lensing spectra, we introduce an overal amplitude of our signal, $A$, with fiducial value 1 as experimental parameter and compute $F_{AA}$.
-Obviously, $\partial_A\br{A B^{XYZ}_{l_1l_2l_3}}|_{A=1} = B^{XYZ}_{l_1l_2l_3}$, so we find
-\begin{equation*}
-    \br{\frac{S}{N}}^2 := F_{AA} =
-\sum_{XYZ, X'Y'Z'} \sum_{l_1\leq l_2 \leq l_3} \frac{\mathcal P _{l_1l_2l_3}}{6}
-B^{X Y Z}_{l_1 l_2 l_3} 
-(C^{-1})^{X X'}_{l_1}
-(C^{-1})^{Y Y'}_{l_2}
-(C^{-1})^{Z Z'}_{l_3}
-B^{X' Y' Z'}_{l_1 l_2 l_3}.
-\end{equation*}
-The equation for the SNR of the power spectra is identical in form.
-
-\subsection{Combined Fisher Matrix}
-To compute the Fisher matrix of an experiment measuring both lensing power- and bispectra we are required to compute and invert the full covariance matrix.
-Under the Gaussian matrix this simplifies trivially.
-The correlation between the a power- and bispectrum estimator contains an odd (5) amount of fields and thus always vanishes.
-We are thus allowed to simply add the Fisher matrices of the power- and bispectra to compute the combined Fisher matrix.
-
-\subsection{Fisher Matrices and Eigenvalues}
-As explained earlier, inverting the Fisher matrix gives us a covariance matrix of the estimators, $\hat \theta_i$ of our experimental parameters. For any vector $\mathbf v$,
-\begin{equation*}
-    \mathbf{v}^T_j \Cov_{ij} \mathbf v_i = \langle (v_i \hat \theta_i)^2 \rangle.
-\end{equation*}
-In particular, if the covariance matrix thus has some large eigenvalue $\lambda_L$ (large compared to the overal accuracy of our experiment), with eigenvector $v_i$, then the corresponding estimator, $v_i\hat \theta_i$, has a large variance and thus represents an approximate degeneracy in our parameters.
-Clearly, when looking at the Fisher matrix we thus instead look for the smallest eigenvalue.
-
-If we have such an abnormally small eigenvalue we are likely to run into problems with calculating and inverting the Fisher matrix due to numerical errors.
-These problems are that inversion becomes unstable and that the matrix as a whole might not be PSD because the small eigenvalue was (wrongly) calculated to be negative.
-These problems are likely because (1) the smallest eigenvalue is sensitive to numerical errors due to being small and (2) the errors on the eigenvalues is proportial to the condition number of the matrix \cite{horn2012matrix}, where the condition number is the ratio between the largest and smallest eigenvalue of the matrix \cite{trefethen1997numerical}.
-Specifically, for a matrix $A$ with some perturbation $\Delta A$, the corresponding  perturbation of eigenvalue $\lambda_i$ is bounded above as
-\begin{equation*}
-    |\Delta \lambda_i | \leq ||\Delta A || \kappa(A)|,
-\end{equation*}
-with $\kappa(A)$ the condition number.
-
-Before inverting a Fisher matrix, it is thus recommended to remove rows and columns corresponding to parameters that are poorly constrained by the experiment, which can be determined by checking for small eigenvalues and their eigenvectors.
-
-% Computing the inverse is then reduced to computing the inverse of the individual block matrices.
+Every element of the Fisher matrix is essentially an inner product weighted by the inverse covariance matrix.
+We can choose how we order vectors of bispectra derivatives.
+Organize the vectors according to the sum 1, 2, and 3 parts first.
+We then order by $l$ configuration first.
+Within each $l$ configuration we can choose any ordering for the $XYZ$ orderings which doesn't change our calculation as long as we stay consistent.
+In this ordering, the covariance matrix can be seen as a block matrix, with every block corresponding to a specific combination of 2 $l$ configurations.
+With the way we decomposed the sum, it's easy to see that no $l$ configuration is the permutation of another.
+If we compute the covariance according to the purely gaussian approximation, the covariance between the bispectrum estimators of two $l$ configurations that are not a permutation of eachother vanishes.
+This means that the covariance matrix becomes a diagonal block matrix blocks of size $4 \times 4$ (sum 1), $6 \times 6$ (sum 2), and $8 \times 8$ (sum 3).
+Computing the inverse is then reduced to computing the inverse of the individual block matrices.
 The block matrices for sum 1 and 2 don't have a particularly nice form (cite paper about multiple tracer fisher matrices for bispectra) but for sum 3 the inverse of the block matrix corresponding to $l_1l_2l_3$ is
-% \begin{equation*}
-%     (\Cov_{l_1l_2l_3}^{-1})^{XYZ, X'Y'Z'} = \hat C_{l_1}^{X,X'}\hat C_{l_2}^{Y,Y'}\hat C_{l_3}^{Z,Z'}.
-% \end{equation*}
-% \begin{equation*}
-%     (\Cov)^{XYZ, X'Y'Z'} = C_{l_1}^{X,X'} C_{l_2}^{Y,Y'} C_{l_3}^{Z,Z'}.
-% \end{equation*}
-% As can be seen from
-% \begin{equation*}
-%     \sum_{X'Y'Z'} = \hat C_{l_1}^{X,X'}\hat C_{l_2}^{Y,Y'}\hat C_{l_3}^{Z,Z'} C_{l_1}^{X'X''}C_{l_2}^{Y'Y''}C_{l_3}^{Z'Z''} = \delta_{XX''}\delta_{YY''}\delta_{ZZ''} = \delta_{XYZ, X''Y''Z''}
-% \end{equation*}
-
-% It's worth mentioning that due to all $l$'s being distinct in each configuration of sum 3 the Wick contraction corresponding to each element of the covariance matrix can only be done in one way and we thus only have one term.
+\begin{equation*}
+    (\Cov_{l_1l_2l_3}^{-1})^{XYZ, X'Y'Z'} = \hat C_{l_1}^{X,X'}\hat C_{l_2}^{Y,Y'}\hat C_{l_3}^{Z,Z'}.
+\end{equation*}
+\begin{equation*}
+    (\Cov)^{XYZ, X'Y'Z'} = C_{l_1}^{X,X'} C_{l_2}^{Y,Y'} C_{l_3}^{Z,Z'}.
+\end{equation*}
+As can be seen from
+\begin{equation*}
+    \sum_{X'Y'Z'} = \hat C_{l_1}^{X,X'}\hat C_{l_2}^{Y,Y'}\hat C_{l_3}^{Z,Z'} C_{l_1}^{X'X''}C_{l_2}^{Y'Y''}C_{l_3}^{Z'Z''} = \delta_{XX''}\delta_{YY''}\delta_{ZZ''} = \delta_{XYZ, X''Y''Z''}
+\end{equation*}
+
+It's worth mentioning that due to all $l$'s being distinct in each configuration of sum 3 the Wick contraction corresponding to each element of the covariance matrix can only be done in one way and we thus only have one term.
 In the sum 2 we get two terms due to $l_1$ and $l_2$ being identical.
@@
 
-% \subsection{Shear noise}
-% In the absence of lensing, galaxies will generally still look elliptical rather than perfectly spherical.
+% With some carefull consideration, it can be established that the sum can be understood as a sum over all possible $XYZ$ configurations (counting different permutations as different terms) and a sum over all possible $l_1l_2l_3$ configurations satisfying $l1\leq l2 \leq l3$ (i.e.
+not counting permutations) with a factor $O_{l1l2l3}^{XYZ}$ to take into account additional overcounting, i.e.
+% \begin{equation*}
+%     \sum_{\text{distinct signals}} B_{l_1l_2l_3}^{XYZ} = \sum_{X,Y,Z \in \{\psi_\kappa, \psi_\gamma\}}\sum_{l_1\leq l_2\leq l_3} (OB^{XYZ}_{l_1l_2l_3})^{-1} B_{l_1l_2l_3}^{XYZ}.
+% \end{equation*}
+% $OB$ counts the distinct permutations of $XYZ$ that leave $(l_1, l_2, l_3)$ the same, (``O'' for overcounting and ``B'' for bispecrum).
+It's explicit form is not very pretty and is given as e.g.
+a Python function as
+% \begin{lstlisting}
+% def OB(X : str, Y : str, Z : str, l1 : int, l2 : int, l3 : int) -> int:
+%     '''
+%     counts the amount of ways to order XYZ that leave l1l2l3 the same
+%     (!) assumes that XYZ only have two different possible values that they can take (psi_kappa and psi_gamma)
+%     '''
+%     if l1 == l2 and l2 == l3:
+%         if not (X == Y and Y == Z):
+%             return 3
+%         else:
+%             return 1
+%     elif l1 == l2:
+%         if X == Y:
+%             return 1
+%         else:
+%             return 2
+%     elif l2 == l3:
+%         if Y == Z:
+%             return 1
+%         else:
+%             return 2
+%     elif l1 == l3:
+%         if X == Z:
+%             return 1
+%         else:
+%             return 2
+%     else:
+%         return 1
+% \end{lstlisting}
+% Because bispectra with different unordered $l_1l_2l_3$ configurations will have zero covariance with eachother (at least in the gaussian approximation) the covariance matrix will take the form of block matrices on the diagonal and zero elsewhere.
+Each block matrix corresponds to a different $l_1l_2l_3$ configurations and each the entries correspond to the different choices for $XYZ$.
+We thus now have
+% \begin{equation*}
+%     F_{\alpha\beta} = \sum_{l_1 \leq l_2 \leq l_3} \sum_{X,Y,Z,X',Y',Z'\in{\psi_{\gamma}, \psi_{\kappa}}} (OB^{XYZ}_{l_1l_2l_3})^{-1}B^{XYZ}_{l_1l_2l_3}
+%     \br{\Cov^{-1}_{l_1l_2l_3}}^{XYZ,X'Z'Y'}
+%     (OB^{X'Y'Z'}_{l_1'l_2'l_3'})^{-1}B^{X'Y'Z'}_{l_1'l_2'l_3'}.
+% \end{equation*}
+% Notice that the covariance matrix notation now changed as well.
+Here we are referring to the inverse of the ($8\times 8$) block matrix corresponding to $l_1l_2l_3$ configuration.
+We next aim to calculate the inverse of this matrix.
+Under the gaussian approximation % TODO : discuss this
+% we can wick contract to get
+% \begin{gather*}
+%     \br{\Cov_{l_1l_2l_3}}^{XYZ,X'Z'Y'} = C^{XX'}_{l1} C^{YY'}_{l2} C^{ZZ'}_{l3} + \delta_{l_1l_2} C^{XY'}_{l1} C^{YX'}_{l2} C^{ZZ'}_{l3} + \delta_{l_1l_3} C^{XZ'}_{l1} C^{YY'}_{l2} C^{ZX'}_{l3} \\
+%     + \delta_{l_2l_3} C^{XX'}_{l1} C^{YZ'}_{l2} C^{ZY'}_{l3} + \delta_{l_1l_2}\delta_{l_2l_3} (C^{XZ'}_{l1} C^{YX'}_{l2} C^{ZY'}_{l3} + C^{XY'}_{l1} C^{YZ'}_{l2} C^{ZX'}_{l3})
+% \end{gather*}
+% CHECK IF CORRECT AND IF I CAN STILL SIMPLIFY THIS
+
+% The inverse of this matrix is
+% \begin{equation*}
+%     M^{\alpha\beta} = \frac{1}{\Delta^{l_1l_2l_3}} \sum_{\text{dist. perms ijk}} \hat C^{ip} \hat C^{jq} \hat C^{kr}, \quad \alpha = ijk, \; \beta = pqr,
+% \end{equation*}
+% with
+% \begin{equation*}
+%     \hat C := \begin{pmatrix}
+%         C^{\psi_\kappa\psi_\kappa} & C^{\psi_\kappa\psi_\gamma}\\
+%         C^{\psi_\kappa\psi_\gamma} & C^{\psi_\gamma\psi_\gamma}
+%     \end{pmatrix}^{-1}.
+% \end{equation*}
+% At this point the Fisher matrix can be written as
+% \begin{gather*}
+%     F_{\theta_1\theta_2}=\sum_{\alpha\beta}\sum_{\text{distinct $l$ configs}}\sum_{\text{distinct $l'$ configs}}\partial_{\theta_1}B^\alpha_{l_1l_2l_3}M^{\alpha\beta}_{l_1l_2l_3}\partial_{\theta_2}B^\beta_{l_1'l_2'l_3'}\delta_{\{l_1l_2l_3\}\{l_1'l_2'l_3'\}}
+% \end{gather*}
+
+% For the powerspectra we get a similar but simpler equation.
+% \begin{gather*}
+%     F_{\alpha\beta} = \sum_{\text{distinc signals}} P^{XY}_{l}(\Cov^{-1})_{l, l'}^{XY, X'Y'}P^{X'Y'}_{l'} \\
+%     = \sum_{l} \sum_{X,Y,X',Y'\in\{\psi_\kappa,\psi_\gamma\}}(OP^{XY})^{-1}P^{XY}_{l}(\Cov_l^{-1})^{XY,X'Y'}(OP^{X'Y'})^{-1}P^{X'Y'}_{l}
+% \end{gather*}
+% with $OP$ given below.
+% \begin{lstlisting}
+%     def OP(X : str, Y : str) -> int:
+%         '''
+%         counts the amount of ways to order XY that leave ll the same
+%         in contrast to OB, behaviour doesn't change for different l values so they aren't taken as an argument
+%         (!) assumes that XY only have two different possible values that they can take (psi_kappa and psi_gamma)
+%         '''
+%         if X == Y:
+%             return 1
+%         else:
+%             return 2
+%     \end{lstlisting}
+% Each (uninverted) block matrix is of the form
+% \begin{equation*}
+%     (\Cov_l)^{XY} = P^{XX'}_l P^{YY'}_l + P^{XY'}_l P^{YX'}_l.
+% \end{equation*}
+
+% \subsection{More on the fisher}
+% Start with the following sum:
+% \begin{equation}
+%     \sum_{ijkpqr}\sum_{l_1l_2l_3}\frac{1}{6}B^{pqr}_{l_1l_2l_3}(C^{-1})^{ip}_{l_1}(C^{-1})^{jq}_{l_2}(C^{-1})^{kr}_{l_3} B^{ijk}_{l_3}.
+% \end{equation}
+% The sum over $l_1l_2l_3$ makes the summand of the $ijkpqr$ sum symmetric in its $ijk$ and $pqr$ indices.
+We can thus express the above as a sum over unordered $ijk$ and $pqr$ denoted by $\alpha$ and $\beta$ as
+% \begin{equation}
+%     \cdots = \sum_{\alpha\beta}\sum_{l_1l_2l_3} B^\alpha_{l_1l_2l_3} (M_1)^{\alpha\beta}_{l_1l_2l_3} B^\beta_{l_1l_2l_3},
+% \end{equation}
+% where
+% \begin{gather*}
+%     M_1^{\alpha\beta} = \\ \frac{1}{6}
+% \begin{pmatrix}
+% \hat{C}^{TT}\hat{C}^{TT}\hat{C}^{TT} & 3\hat{C}^{TT}\hat{C}^{TT}\hat{C}^{TE} & 3\hat{C}^{TT}\hat{C}^{TE}\hat{C}^{TE} & \hat{C}^{TE}\hat{C}^{TE}\hat{C}^{TE} \\
+% 3\hat{C}^{TT}\hat{C}^{TT}\hat{C}^{TE} & 3\hat{C}^{EE}\hat{C}^{TT}\hat{C}^{TT} + 6\hat{C}^{TE}\hat{C}^{TE}\hat{C}^{TT} & 3\hat{C}^{TE}\hat{C}^{TE}\hat{C}^{TE} + 6\hat{C}^{TT}\hat{C}^{EE}\hat{C}^{TE} & 3\hat{C}^{TE}\hat{C}^{TE}\hat{C}^{EE} \\
+% 3\hat{C}^{TT}\hat{C}^{TE}\hat{C}^{TE} & 3\hat{C}^{TE}\hat{C}^{TE}\hat{C}^{TE} + 6\hat{C}^{TT}\hat{C}^{EE}\hat{C}^{TE} & 6\hat{C}^{EE}\hat{C}^{TE}\hat{C}^{TE} + 3\hat{C}^{TT}\hat{C}^{EE}\hat{C}^{EE} & 3\hat{C}^{TE}\hat{C}^{EE}\hat{C}^{EE} \\
+% \hat{C}^{TE}\hat{C}^{TE}\hat{C}^{TE} & 3\hat{C}^{TE}\hat{C}^{TE}\hat{C}^{EE} & 3\hat{C}^{TE}\hat{C}^{EE}\hat{C}^{EE} & \hat{C}^{EE}\hat{C}^{EE}\hat{C}^{EE}
+% \end{pmatrix}.
+% \end{gather*}
+% This matrix is essentially of the form
+% \begin{gather*}
+%     M_1^{\alpha\beta} = \frac{1}{6} \sum_{\text{dist. perms ijk}} \sum_{\text{dist. perms pqr}} \hat C^{ip} \hat C^{jq} \hat C^{kr}
+% \end{gather*}
+% The matrix given by
+% \begin{gather*}
+%     M_2^{\alpha\beta} = \sum_{\text{all perms ijk}} C^{ip} C^{jq} C^{kr}
+% \end{gather*}
+% is the inverse of $M_1$. This can be checked. By definition:
+% \begin{gather*}
+%     M_1^{\alpha\beta}M_2^{\beta\gamma} = \frac{1}{6}\sum_{ijk}\br{\sum_{\text{dist.
+perms ijk}} \sum_{\text{dist.
+perms pqr}} \hat C^{ip} \hat C^{jq} \hat C^{kr}}\br{\sum_{\text{all perms ijk}} C^{pn} C^{qm} C^{rw}}.
+% \end{gather*}
+% Now notice that on the 2nd sum of left factor the amount of terms will be
+% \begin{equation*}
+%     \frac{(\text{total \# of indices (3)})!}{(\text{\# indicies of type }\psi_\gamma)!(\text{\# indicies of type }\psi_\kappa)!}.
+% \end{equation*}
+% On the other hand, when summing over the $pqr$'s, the each term will ``line up'' with the ordering with some of the terms of the second factor. The amount of factors it will line up with equals
+% \begin{equation*}
+%     (\text{\# indicies of type }\psi_\gamma)!(\text{\# indicies of type }\psi_\kappa)!.
+% \end{equation*} 
+% So multiplying the two (i.e. counting the total number of indices) we get $3!=6$ identical terms, which cancels with the $1/6$ factor. We thus find
+% \begin{equation*}
+%     M_1^{\alpha\beta}M_2^{\beta\gamma} = \sum_{\text{dist. perms ijk}} \delta_{in}\delta_{jm}\delta_{kw} = \delta_{\alpha\beta}.
+% \end{equation*}
+% In the last step we used that if we only look at $ijk$ configurations as distinct if they are not permutations of each other, then the sum above is just the normal delta factor in an identity matrix.
+
+\subsection{Shear noise}
+In the absence of lensing, galaxies will generally still look elliptical rather than perfectly spherical.
 This \textbf{intrinsic ellipticity} is a major cause of noise when measuring lensing shear.
 We can define the ellipticity of a galaxy image (or any type of object, for that matter) in terms of \textbf{quadrupole moments} of the surface brightness as
-% \begin{equation*}
-% 	e_1 = \frac{I_{xx} - I_{yy}}{I_{xx} + I_{yy}}, \quad e_2 = \frac{2I_{xy}}{I_{xx} + I_{yy}},
-% \end{equation*}
-% where
-% \begin{equation*}
-% 	I_{ij} = \int \d x\d y I(x,y)ij, \quad i,j\in\{x,y\},
-% \end{equation*}
-% and the galaxy is centered on $(x,y)=0$.
-
-% The connection with lensing shear is then made by taylor expanding and using the lensing matrix.
+\begin{equation*}
+	e_1 = \frac{I_{xx} - I_{yy}}{I_{xx} + I_{yy}}, \quad e_2 = \frac{2I_{xy}}{I_{xx} + I_{yy}},
+\end{equation*}
+where
+\begin{equation*}
+	I_{ij} = \int \d x\d y I(x,y)ij, \quad i,j\in\{x,y\},
+\end{equation*}
+and the galaxy is centered on $(x,y)=0$.
+
+The connection with lensing shear is then made by taylor expanding and using the lensing matrix.
 If we have unlensed ellipiticities $e_i^\text{true}$ and lensed ellipticities $e_i^{\text{obs}}$, then \cite{dodelson_modern_cosmology}
-% \begin{gather*}
-% 	e_1^{\text{obs}} = \br{1-2\sbr{e_1^{\text{true}}\gamma_1+e_2^{\text{true}}\gamma_2}}e_1^{\text{true}} + 2\gamma_1,\\
-% 	e_2^{\text{obs}} = \br{1-2\sbr{e_1^{\text{true}}\gamma_1+e_2^{\text{true}}\gamma_2}}e_2^{\text{true}} + 2\gamma_2.
-% \end{gather*}
-% In the case of weak lensing, $\gamma_i, \kappa \ll 1$, thus the above reduces to $\gamma_i \approx (e_i^{\text{obs}} - e^{\text{true}}_i)/2$.
+\begin{gather*}
+	e_1^{\text{obs}} = \br{1-2\sbr{e_1^{\text{true}}\gamma_1+e_2^{\text{true}}\gamma_2}}e_1^{\text{true}} + 2\gamma_1,\\
+	e_2^{\text{obs}} = \br{1-2\sbr{e_1^{\text{true}}\gamma_1+e_2^{\text{true}}\gamma_2}}e_2^{\text{true}} + 2\gamma_2.
+\end{gather*}
+In the case of weak lensing, $\gamma_i, \kappa \ll 1$, thus the above reduces to $\gamma_i \approx (e_i^{\text{obs}} - e^{\text{true}}_i)/2$.
 The analysis thusfar was quietly done in real space, but taking the harmonic transform of $\gamma_i$ is trivial under the approximation we just made.
@@
 If there are $N$ galaxies observed per patch, then the noise is given by
-% $$
-% \langle \br{\hat \gamma_{i,l}  - \gamma_{i,l}^{\text{true}}}^2\rangle \approx \frac{\langle \br{e_{i,l}^\text{true}}^2\rangle}{4N}.
-% $$
-% We don't know the exact value for the right hand side either, so we use the standard estimator for that.
+$$
+\langle \br{\hat \gamma_{i,l}  - \gamma_{i,l}^{\text{true}}}^2\rangle \approx \frac{\langle \br{e_{i,l}^\text{true}}^2\rangle}{4N}.
+$$
+We don't know the exact value for the right hand side either, so we use the standard estimator for that.
 There is no correlation between two patches, i.e.
@@
 We thus get
-% \begin{equation*}
-% 	C^{\gamma\gamma,\;\text{obs}}_l = C^{\gamma\gamma, \;\text{true}}_l + N_l^{\gamma\gamma},
-% \end{equation*} 
-% with
-% \begin{equation*}
-% 	 N_l^{\gamma\gamma} = \frac{1}{4}\frac{e_{\text{RMS}}}{N}, \quad e_{\text{RMS}}:=\sqrt{\frac{1}{N}\sum^N_{i=1}\sbr{(e_{1, i} - \bar e_1)^2 + (e_{2,i}-\bar e_2)^2}}.
-% \end{equation*}
-
-% Typically $e_{RMS} / 4 \sim 0.3$ \cite{Kilbinger_2015}.
-
-% The EUCLID team aims for a noise of about 20 percent and thus require at least $\sim 4 \times 10^8$ galaxies measured per patch, giving an absolute error of \cite{2024}
-% $$
-% N_l^{\gamma\gamma} \approx 8 \times 10^{-10} \quad \text{(EUCLID)}.
-% $$
-
-% \subsection{Convergence Noise}
-% \subsubsection{Optimal Estimator}
-% In practice, the lensing potential cannot be directly measured but is instead estimated based on the temperature, $\Theta$, $E$-mode polarization, $E$, and $B$-mode polarization, $B$, fields.
-
-% Define some axes $x$ and $y$ on a small patch of the sky and let $\mathbf E$ be the electric field of the CMB radiation, then the \textbf{linear polarization Stokes parameters} are defined as
-% \begin{equation*}
-% 	Q := E_xE^*_x - E_yE_y^*, \quad U := E_xE_y^* + E_yE_x^*.
-% \end{equation*} 
-% By looking at a sufficiently small patch of the sky we can apply the \textbf{flat sky limit}, in which case we take a 2D Fourier transform over the approximately flat patch instead of working with spherical harmonics.
+\begin{equation*}
+	C^{\gamma\gamma,\;\text{obs}}_l = C^{\gamma\gamma, \;\text{true}}_l + N_l^{\gamma\gamma},
+\end{equation*} 
+with
+\begin{equation*}
+	 N_l^{\gamma\gamma} = \frac{1}{4}\frac{e_{\text{RMS}}}{N}, \quad e_{\text{RMS}}:=\sqrt{\frac{1}{N}\sum^N_{i=1}\sbr{(e_{1, i} - \bar e_1)^2 + (e_{2,i}-\bar e_2)^2}}.
+\end{equation*}
+
+Typically $e_{RMS} / 4 \sim 0.3$ \cite{Kilbinger_2015}.
+
+The EUCLID team aims for a noise of about 20 percent and thus require at least $\sim 4 \times 10^8$ galaxies measured per patch, giving an absolute error of \cite{2024}
+$$
+N_l^{\gamma\gamma} \approx 8 \times 10^{-10} \quad \text{(EUCLID)}.
+$$
+
+\subsection{Convergence Noise}
+\subsubsection{Optimal Estimator}
+In practice, the lensing potential cannot be directly measured but is instead estimated based on the temperature, $\Theta$, $E$-mode polarization, $E$, and $B$-mode polarization, $B$, fields.
+
+Define some axes $x$ and $y$ on a small patch of the sky and let $\mathbf E$ be the electric field of the CMB radiation, then the \textbf{linear polarization Stokes parameters} are defined as
+\begin{equation*}
+	Q := E_xE^*_x - E_yE_y^*, \quad U := E_xE_y^* + E_yE_x^*.
+\end{equation*} 
+By looking at a sufficiently small patch of the sky we can apply the \textbf{flat sky limit}, in which case we take a 2D Fourier transform over the approximately flat patch instead of working with spherical harmonics.
 In this case we find $Q(\mathbf k)$ and $U(\mathbf k)$.
 In Fourier space we can instead work with the more convenient $E$ and $B$ modes, defined as
-% \begin{gather*}
-% 	E(\mathbf k)=\cos 2\phi_{\mathbf k} Q(\mathbf k) + \sin 2\phi_{\mathbf k} U(\mathbf k),\\
-% 	B(\mathbf k)=-\sin 2\phi_{\mathbf k} Q(\mathbf k) + \cos 2\phi_{\mathbf k} U(\mathbf k),
-% \end{gather*}
-% where $\phi_\textbf{k}$ is the angle of $\mathbf k$ with the $(1,0)$  vector (i.e. $k_{\hat{\mathbf{x}}}$).
-% $E$ corresponds to the curl free component of the polarization field and behaves like a scalar under transformations.
+\begin{gather*}
+	E(\mathbf k)=\cos 2\phi_{\mathbf k} Q(\mathbf k) + \sin 2\phi_{\mathbf k} U(\mathbf k),\\
+	B(\mathbf k)=-\sin 2\phi_{\mathbf k} Q(\mathbf k) + \cos 2\phi_{\mathbf k} U(\mathbf k),
+\end{gather*}
+where $\phi_\textbf{k}$ is the angle of $\mathbf k$ with the $(1,0)$  vector (i.e. $k_{\hat{\mathbf{x}}}$).
+$E$ corresponds to the curl free component of the polarization field and behaves like a scalar under transformations.
 $B$ corresponds to the rest of the polarization field and behaves like a semi-scalar under transformations.
 The Fourier components are thus related to the real space fields as
-% \begin{gather*}
-% \Theta(\mathbf{\hat{n}}) = \int \frac{d^2 k}{(2\pi)^2} \Theta(\mathbf{k}) e^{i \mathbf{k} \cdot \mathbf{\hat{n}}}, \\
-% [Q \pm iU](\mathbf{\hat{n}}) = -\int \frac{d^2 k}{(2\pi)^2} \left[ E(\mathbf{k}) \pm iB(\mathbf{k}) \right] e^{\pm 2i \varphi_{\mathbf{k}}} e^{i \mathbf{k} \cdot \mathbf{\hat{n}}}, \\
-% \phi(\mathbf{\hat{n}}) = \int \frac{d^2 K}{(2\pi)^2} \phi(\mathbf{K}) e^{i \mathbf{K} \cdot \mathbf{\hat{n}}}.
-% \end{gather*}
-
-% As discussed earlier, lensing acts in real space as 
-% \begin{equation*}
-%     X(\hat n) = \tilde X(\hat {\mathbf n} + \nabla \phi(\hat{\mathbf n})) \approx \tilde X(\hat{\mathbf n}) + \nabla \phi \cdot \nabla \tilde X |_{\hat{\mathbf{n}}},
-% \end{equation*}
-% with $X$ any of $\Theta$, $Q \pm iU$ and $\tilde X$ the corresponding \textit{unlensed} field.
+\begin{gather*}
+\Theta(\mathbf{\hat{n}}) = \int \frac{d^2 k}{(2\pi)^2} \Theta(\mathbf{k}) e^{i \mathbf{k} \cdot \mathbf{\hat{n}}}, \\
+[Q \pm iU](\mathbf{\hat{n}}) = -\int \frac{d^2 k}{(2\pi)^2} \left[ E(\mathbf{k}) \pm iB(\mathbf{k}) \right] e^{\pm 2i \varphi_{\mathbf{k}}} e^{i \mathbf{k} \cdot \mathbf{\hat{n}}}, \\
+\phi(\mathbf{\hat{n}}) = \int \frac{d^2 K}{(2\pi)^2} \phi(\mathbf{K}) e^{i \mathbf{K} \cdot \mathbf{\hat{n}}}.
+\end{gather*}
+
+As discussed earlier, lensing acts in real space as 
+\begin{equation*}
+    X(\hat n) = \tilde X(\hat {\mathbf n} + \nabla \phi(\hat{\mathbf n})) \approx \tilde X(\hat{\mathbf n}) + \nabla \phi \cdot \nabla \tilde X |_{\hat{\mathbf{n}}},
+\end{equation*}
+with $X$ any of $\Theta$, $Q \pm iU$ and $\tilde X$ the corresponding \textit{unlensed} field.
 In Fourier space this naturally takes the form of a convolution.
 The changes in Fourier moments due to lensing are thus given by \cite{Hu_2002}
-% \begin{gather*}
-%     \delta \Theta(\mathbf{k}) = \int \frac{d^2 k'}{(2\pi)^2} \tilde{\Theta}(\mathbf{k'}) W(\mathbf{k'}, \mathbf{K}), \\[8pt]
-%     \delta E(\mathbf{k}) = \int \frac{d^2 k'}{(2\pi)^2} 
-%     \left[ \tilde{E}(\mathbf{k'}) \cos 2\varphi_{k'k} - \tilde{B}(\mathbf{k'}) \sin 2\varphi_{k'k} \right] W(\mathbf{k'}, \mathbf{K}), \\[8pt]
-%     \delta B(\mathbf{k}) = \int \frac{d^2 k'}{(2\pi)^2} 
-%     \left[ \tilde{B}(\mathbf{k'}) \cos 2\varphi_{k'k} + \tilde{E}(\mathbf{k'}) \sin 2\varphi_{k'k} \right] W(\mathbf{k'}, \mathbf{K}),
-% \end{gather*}
-
-% \noindent where \( \varphi_{k'k} \equiv \varphi_{k'} - \varphi_k \), \( \mathbf{K} = \mathbf{k} - \mathbf{k'} \), and
-% \begin{gather*}
-%     W(\mathbf{k}, \mathbf{K}) = -[\mathbf{k} \cdot \mathbf{K}] \phi(\mathbf{K}).
-% \end{gather*}
-
-% For power spectra in 2D Fourier space we use the definition
-% \begin{equation*}
-%     \langle x(\mathbf k)x'(\mathbf k') \rangle = (2\pi)^2\delta(\mathbf k + \mathbf k')C_k^{xx'}
-% \end{equation*}
-% unless stated otherwise. We will denote unlensed powerspectra with a tilde on top. $C^{xx'}(k)$ denotes the observed power spectrum is related to the true (but still lensed) spectrum as
-% \begin{equation*}
-%     C^{xx', \text{ obs}}(k) = B(k)^2C^{xx', \text{ true}}(k) + \delta_{xx'}N^x(k).
-% \end{equation*}
-% We are then able to express the power spectra of the lensed fields in terms of the unlensed fields as
-% \begin{equation*}
-%     \langle x(\mathbf k)x'(\mathbf k') \rangle = f_{xx'}(\mathbf k, \mathbf k') \phi(\mathbf k + \mathbf k'),
-% \end{equation*}
-% where $f_{xx'}$ is a function of 2 wavenumbers and unlensed power spectra. Specific values of $f$ for different combinations of $x$ and $x'$ can be found in ... %TODO: fix this
-% We can then set up an estimator for $\phi$ using a weighted integral of the Fourier modes measured for $x$ and $x'$ by allowing for a filter function $F_{xx'}(\mathbf k, \mathbf k')$.
+\begin{gather*}
+    \delta \Theta(\mathbf{k}) = \int \frac{d^2 k'}{(2\pi)^2} \tilde{\Theta}(\mathbf{k'}) W(\mathbf{k'}, \mathbf{K}), \\[8pt]
+    \delta E(\mathbf{k}) = \int \frac{d^2 k'}{(2\pi)^2} 
+    \left[ \tilde{E}(\mathbf{k'}) \cos 2\varphi_{k'k} - \tilde{B}(\mathbf{k'}) \sin 2\varphi_{k'k} \right] W(\mathbf{k'}, \mathbf{K}), \\[8pt]
+    \delta B(\mathbf{k}) = \int \frac{d^2 k'}{(2\pi)^2} 
+    \left[ \tilde{B}(\mathbf{k'}) \cos 2\varphi_{k'k} + \tilde{E}(\mathbf{k'}) \sin 2\varphi_{k'k} \right] W(\mathbf{k'}, \mathbf{K}),
+\end{gather*}
+
+\noindent where \( \varphi_{k'k} \equiv \varphi_{k'} - \varphi_k \), \( \mathbf{K} = \mathbf{k} - \mathbf{k'} \), and
+\begin{gather*}
+    W(\mathbf{k}, \mathbf{K}) = -[\mathbf{k} \cdot \mathbf{K}] \phi(\mathbf{K}).
+\end{gather*}
+
+For power spectra in 2D Fourier space we use the definition
+\begin{equation*}
+    \langle x(\mathbf k)x'(\mathbf k') \rangle = (2\pi)^2\delta(\mathbf k + \mathbf k')C_k^{xx'}
+\end{equation*}
+unless stated otherwise. We will denote unlensed powerspectra with a tilde on top. $C^{xx'}(k)$ denotes the observed power spectrum is related to the true (but still lensed) spectrum as
+\begin{equation*}
+    C^{xx', \text{ obs}}(k) = B(k)^2C^{xx', \text{ true}}(k) + \delta_{xx'}N^x(k).
+\end{equation*}
+We are then able to express the power spectra of the lensed fields in terms of the unlensed fields as
+\begin{equation*}
+    \langle x(\mathbf k)x'(\mathbf k') \rangle = f_{xx'}(\mathbf k, \mathbf k') \phi(\mathbf k + \mathbf k'),
+\end{equation*}
+where $f_{xx'}$ is a function of 2 wavenumbers and unlensed power spectra. Specific values of $f$ for different combinations of $x$ and $x'$ can be found in ... %TODO: fix this
+We can then set up an estimator for $\phi$ using a weighted integral of the Fourier modes measured for $x$ and $x'$ by allowing for a filter function $F_{xx'}(\mathbf k, \mathbf k')$.
 This gives the estimator
-% \begin{equation*}
-%     \hat\phi_{xx'}(\mathbf k) = \frac{\int \frac{\d^2k'}{(2\pi)^2}x(\mathbf k')x'(\mathbf k - \mathbf k') F_{xx'}(\mathbf k', \mathbf k - \mathbf k')}{\int \frac{\d^2k'}{(2\pi)^2}f_{xx'}(\mathbf k', \mathbf k - \mathbf k') F_{xx'}(\mathbf k', \mathbf k - \mathbf k')}.
-% \end{equation*}
-% It's easily checked that $\langle \hat\phi_{xx'} \rangle = \phi$ regardless of the filter function used. Minimizing for the variance of the estimator gives
-% \begin{equation}
-%     F_{xx'}(\mathbf{k}_1, \mathbf{k}_2) = 
-%     \frac{C_{k_1}^{x'x'} C_{k_2}^{xx} f_{xx'}(\mathbf{k}_1, \mathbf{k}_2) 
-%     - C_{k_1}^{xx'} C_{k_2}^{xx'} f_{xx'}(\mathbf{k}_2, \mathbf{k}_1)}
-%     {C_{k_1}^{xx} C_{k_2}^{x'x'} C_{k_1}^{x'x'} C_{k_2}^{xx} 
-%     - \left( C_{k_1}^{xx'} C_{k_2}^{xx'} \right)^2}.
-% \end{equation}
-% With this estimator we assume that the unlensed powerspectra have been determined through other means (e.g. a simulation).
-% With optimal weighting the noise becomes
-% \begin{equation*}
-%     \langle \hat\phi_{xy}(\mathbf K) \hat\phi_{x'y'}(\mathbf K') \rangle = (2\pi)^2\delta(\mathbf K - \mathbf K') [C^{\phi \phi}_K + K^{-2}N_{xyx'y'}(K)],
-% \end{equation*}
-% where
-% \begin{equation*}
-%     N_{xyx'y'}(K):=K^{-2}A_{xy}(K)A_{x'y'}(K)\int\frac{\d^2 k'}{(2\pi)^2}F_{xy}(\mathbf k', \mathbf k - \mathbf k')\br{
-%         F_{x'y'}(\mathbf k', \mathbf k - \mathbf k') C_{k'}^{xx'}C^{yy'}_{k - k'} + F_{x'y'}(\mathbf k - \mathbf k', \mathbf k') C_{k'}^{xy'}C^{yx'}_{k - k'}
-%     }.
-% \end{equation*}
-
-% \subsubsection{Noise}
-% We partition the sky into a finite number of pixels of size (in solid angle) $\Omega_{\text{pix}}$.
+\begin{equation*}
+    \hat\phi_{xx'}(\mathbf k) = \frac{\int \frac{\d^2k'}{(2\pi)^2}x(\mathbf k')x'(\mathbf k - \mathbf k') F_{xx'}(\mathbf k', \mathbf k - \mathbf k')}{\int \frac{\d^2k'}{(2\pi)^2}f_{xx'}(\mathbf k', \mathbf k - \mathbf k') F_{xx'}(\mathbf k', \mathbf k - \mathbf k')}.
+\end{equation*}
+It's easily checked that $\langle \hat\phi_{xx'} \rangle = \phi$ regardless of the filter function used. Minimizing for the variance of the estimator gives
+\begin{equation}
+    F_{xx'}(\mathbf{k}_1, \mathbf{k}_2) = 
+    \frac{C_{k_1}^{x'x'} C_{k_2}^{xx} f_{xx'}(\mathbf{k}_1, \mathbf{k}_2) 
+    - C_{k_1}^{xx'} C_{k_2}^{xx'} f_{xx'}(\mathbf{k}_2, \mathbf{k}_1)}
+    {C_{k_1}^{xx} C_{k_2}^{x'x'} C_{k_1}^{x'x'} C_{k_2}^{xx} 
+    - \left( C_{k_1}^{xx'} C_{k_2}^{xx'} \right)^2}.
+\end{equation}
+With this estimator we assume that the unlensed powerspectra have been determined through other means (e.g. a simulation).
+With optimal weighting the noise becomes
+\begin{equation*}
+    \langle \hat\phi_{xy}(\mathbf K) \hat\phi_{x'y'}(\mathbf K') \rangle = (2\pi)^2\delta(\mathbf K - \mathbf K') [C^{\phi \phi}_K + K^{-2}N_{xyx'y'}(K)],
+\end{equation*}
+where
+\begin{equation*}
+    N_{xyx'y'}(K):=K^{-2}A_{xy}(K)A_{x'y'}(K)\int\frac{\d^2 k'}{(2\pi)^2}F_{xy}(\mathbf k', \mathbf k - \mathbf k')\br{
+        F_{x'y'}(\mathbf k', \mathbf k - \mathbf k') C_{k'}^{xx'}C^{yy'}_{k - k'} + F_{x'y'}(\mathbf k - \mathbf k', \mathbf k') C_{k'}^{xy'}C^{yx'}_{k - k'}
+    }.
+\end{equation*}
+
+\subsubsection{Noise}
+We partition the sky into a finite number of pixels of size (in solid angle) $\Omega_{\text{pix}}$.
 We take into account detector beam width and homogenous noise, but not boundary effects or inhomogeneous noise.
 %TODO: justify this or take it into account
-% In real space the observed field relates to the true (but still lensed) field as
-% \begin{equation*}
-%     X^{\text{obs}}(\hat{\mathbf n}) = \int \d\Omega' X^{\text{true}}(\hat{\mathbf n}') B(\hat{\mathbf n}, \hat{\mathbf n}') + \eta(\hat{\mathbf n}),
-% \end{equation*}
-% where $B$ is the beam pattern and $\eta$ is the noise map stemming from any number of experimental sources of error.
-
-% In the case of an isotropic uniform beam, the beam pattern can be written as a function of $\hat{\mathbf n} - \hat{\mathbf n}'$ and the Fourier transform gives 
-% \begin{equation*}
-%     X^{\text{obs}}(\mathbf k) = X^{\text{true}}(\mathbf k)B(\mathbf k) + \eta(\mathbf k),
-% \end{equation*}
-% with
-% \begin{equation*}
-%     B(\mathbf k) = e^{-\frac{k^2\theta_{\text{FWHM}}^2}{4\ln 2}},
-% \end{equation*}
-% and $\theta_{\text{FWHM}}$ the angle for full width at half minimum.
-% By using a likelihood approach we obtain the estimator of the lensed power spectrum given by
-% \begin{equation*}
-%     \hat C^{XX}(k) = B(\mathbf k)^{-2}\br{X^{\text{obs}}(\mathbf k)X^{\text{obs}}(\mathbf k)^* - N(k)}
-% \end{equation*}
-% with variance
-% \begin{equation*}
-%     \Cov(\hat C^{XX}(k), \hat C^{XX}(k')) = 2\sbr{C(k) + N(k)B(k)^{-2}}\delta(\mathbf k - \mathbf k').
-% \end{equation*}
-% The noise as calculated in eq ... % TODO: ref 
-% is thus given by using
-% \begin{equation*}
-%     C^{XY, \text{ obs}}(k) = \langle \hat C^{XY}(k) \rangle = B(\mathbf k)^{-2}(C)
-% \end{equation*}
-% \subsection{Power and bispectrum estimators}
-% \subsection{Combining experimental and theoretical (shot noise?) uncertainty}
-% %Let's say you are looking at a field $X_l$ and have some experimental error $E_l$, uncorrelated with $X_l$ and with mean $0$ (so no systematic error). Then
-% %\begin{equation*}
-% %	\langle (X+E)(X+E)^*\rangle = \langle X X^*\rangle + \langle EE^* \rangle.
-% %\end{equation*}
-
-% In the case of an error expressed as
-% \begin{gather*}
-% 	\text{``We measure the amplitude of the CMB lensing} \\\text{
-% 	power spectrum at state-of-the-art 2.3\% precision, [...]"}
-% \end{gather*}
-% we can quantify this as
-% \begin{equation*}
-% 	\hat P^X_l = P_l^X(1+E),
-% \end{equation*}
-% where $E\sim \mathcal N(0, \sigma^2)$ and in the example above $\sigma = 0.023$. It thus follows that
+In real space the observed field relates to the true (but still lensed) field as
+\begin{equation*}
+    X^{\text{obs}}(\hat{\mathbf n}) = \int \d\Omega' X^{\text{true}}(\hat{\mathbf n}') B(\hat{\mathbf n}, \hat{\mathbf n}') + \eta(\hat{\mathbf n}),
+\end{equation*}
+where $B$ is the beam pattern and $\eta$ is the noise map stemming from any number of experimental sources of error.
+
+In the case of an isotropic uniform beam, the beam pattern can be written as a function of $\hat{\mathbf n} - \hat{\mathbf n}'$ and the Fourier transform gives 
+\begin{equation*}
+    X^{\text{obs}}(\mathbf k) = X^{\text{true}}(\mathbf k)B(\mathbf k) + \eta(\mathbf k),
+\end{equation*}
+with
+\begin{equation*}
+    B(\mathbf k) = e^{-\frac{k^2\theta_{\text{FWHM}}^2}{4\ln 2}},
+\end{equation*}
+and $\theta_{\text{FWHM}}$ the angle for full width at half minimum.
+By using a likelihood approach we obtain the estimator of the lensed power spectrum given by
+\begin{equation*}
+    \hat C^{XX}(k) = B(\mathbf k)^{-2}\br{X^{\text{obs}}(\mathbf k)X^{\text{obs}}(\mathbf k)^* - N(k)}
+\end{equation*}
+with variance
+\begin{equation*}
+    \Cov(\hat C^{XX}(k), \hat C^{XX}(k')) = 2\sbr{C(k) + N(k)B(k)^{-2}}\delta(\mathbf k - \mathbf k').
+\end{equation*}
+The noise as calculated in eq ... % TODO: ref 
+is thus given by using
+\begin{equation*}
+    C^{XY, \text{ obs}}(k) = \langle \hat C^{XY}(k) \rangle = B(\mathbf k)^{-2}(C)
+\end{equation*}
+\subsection{Power and bispectrum estimators}
+\subsection{Combining experimental and theoretical (shot noise?) uncertainty}
+%Let's say you are looking at a field $X_l$ and have some experimental error $E_l$, uncorrelated with $X_l$ and with mean $0$ (so no systematic error). Then
+%\begin{equation*}
+%	\langle (X+E)(X+E)^*\rangle = \langle X X^*\rangle + \langle EE^* \rangle.
+%\end{equation*}
+
+In the case of an error expressed as
+\begin{gather*}
+	\text{``We measure the amplitude of the CMB lensing} \\\text{
+	power spectrum at state-of-the-art 2.3\% precision, [...]"}
+\end{gather*}
+we can quantify this as
+\begin{equation*}
+	\hat P^X_l = P_l^X(1+E),
+\end{equation*}
+where $E\sim \mathcal N(0, \sigma^2)$ and in the example above $\sigma = 0.023$. It thus follows that
 
@@
 
-\section{Shear equals twice spin raised lensing potential}\label{sec:shear}
+\section{Current/Future Experiments Considered}
+\subsection{Atacama Cosmology Telescope (ACT)}
+Using the results of the sixth data release of the ACT, DR6, the lensing power spectrum was determined with state of the art accuracy in the range $40<l<763$, \cite{ACT_lps_measurement} .
+They worked with averages over non-overlapping multipole bins seen in \ref{fig:act_lps_errors}.
+Running a number of simulatutions of their measurement, they found the full covariance matrix of the lensing power spectrum measurement, also seen in figure \ref{fig:act_lps_errors}.
+\begin{figure}
+\centering
+\begin{subfigure}{0.45\textwidth}
+    %\label{tab:lensing_bandpowers}
+    \begin{tabular}{ccc}
+        \toprule
+        \([L_{\text{min}} ~ L_{\text{max}}]\) & \( L_b \) & \( 10^7 [L(L+1)]^2 C_L^{\phi\phi} / 4 \) \\
+        \midrule
+        \([40 ~ 66]\)   & 53.0  & \( 2.354 \pm 0.157 \) \\
+        \([67 ~ 101]\)  & 83.5  & \( 1.822 \pm 0.096 \) \\
+        \([102 ~ 145]\) & 123.0 & \( 1.368 \pm 0.068 \) \\
+        \([146 ~ 199]\) & 172.0 & \( 1.000 \pm 0.054 \) \\
+        \([200 ~ 264]\) & 231.5 & \( 0.758 \pm 0.047 \) \\
+        \([265 ~ 339]\) & 301.5 & \( 0.617 \pm 0.048 \) \\
+        \([340 ~ 426]\) & 382.5 & \( 0.439 \pm 0.039 \) \\
+        \([427 ~ 526]\) & 476.0 & \( 0.409 \pm 0.032 \) \\
+        \([527 ~ 638]\) & 582.0 & \( 0.249 \pm 0.027 \) \\
+        \([639 ~ 763]\) & 700.5 & \( 0.156 \pm 0.026 \) \\
+        \bottomrule
+    \end{tabular}
+%    \caption{Lensing bandpowers in the baseline range, for our minimum-variance (MV) estimator.
+Values are given for \( 10^7 [L(L + 1)]^2 C_L^{\phi\phi} / 4 = 10^7 C_L^{\kappa\kappa} \) averaged over non-interlapping bins with bin edges given in the first column and band centers \( L_b \) in the second.
+Source is \cite{ACT_lps_measurement}}
+\end{subfigure}
+% \begin{subfigure}{0.45\textwidth}
+% 	\includegraphics[width=\textwidth]{figures/cov_mat_act.png}
+% \end{subfigure}
+\label{fig:act_lps_errors}
+\caption{Errors and covariances from ACT DR6 lensing power spectrum measurements \cite{ACT_lps_measurement}}
+\end{figure}
+
+In this case we can thus model the estimator of the lensing power spectrum as a random variable according to
+
+\begin{equation*}
+	\hat P^X_l = P^X_l(1+E_l), \quad E \sim \mathcal N(0, \sigma^2_l).
+\end{equation*}
+
+The denominators in the terms of the Fisher matrix thus now equal
+\begin{equation*}
+	\langle \hat P^X_{l_1} \hat P^X_{l_2} \hat P^X_{l_3} \rangle =P^X_{l_1}P^X_{l_1}P^X_{l_1} (1 + \langle E_{l_1}E_{l_2}\rangle + \langle E_{l_2}E_{l_3}\rangle + \langle E_{l_3}E_{l_1}\rangle),
+\end{equation*}
+the $\langle E_i E_j \rangle$ terms then equal the covariances seen in figure \ref{fig:act_lps_errors}.
+Our calculation of the Fisher bounds assume a measurement of each individual multipole.
+We make the assumption that the variances / covariances stay roughly the same for multipoles in the same bin.
+In this case
+\begin{equation*}
+	\langle E_{l_{A,i}}E_{l_{B,j}} \rangle = \frac{1}{P^\psi_{l_{A,i}}P^\psi_{l_{B,j}}}\Cov (\hat P^\psi_{l_{A,i}},\hat P^\psi_{l_{B,j}}) = N_AN_B \underbrace{ \frac{1}{P^\psi_AP^\psi_B}\Cov\br{ P^\psi_A , P_B^\psi }}_{\text{from figure \ref{fig:act_lps_errors}}},
+\end{equation*}
+where $l_{A,i}$ is a multipole from bin $A$, $N_A$ is the amount of multipoles in bin $A$, and
+$$
+P^\psi_A:=\sum_{l_{A,i}\in \text{bin} A}P^\psi_{l_{A,i}}
+$$
+Similarly for $B$.
+
+\subsection{Dark Energy Survey (DES)}
+\subsection{Simons Observatory}
+\subsection{Vera C. Rubin Observatory's Legacy Survey of Space and Time (LSST)}
+\subsection{Euclid Mission}
+\subsection{Nancy Grace Roman Space Telescope}
+\section{Results and Discussion}
+\subsection{Results without experimental errors}
+% \begin{figure}[h!]
+% 	\centering
+% 	\includegraphics[width=\textwidth]{figures/fisher_results_theoretical.png}
+% 	\label{fig:fisher_results_theoretical}
+% 	\caption{Lower bound on the relative error in determining cosmological parameters as a function of the maximum multipole moment, $l_{\text{max}}$, measured by an experiment, with a minimum multipole moment of $l_{\text{min}}=50$.
+The analysis considers the impact of measuring only convergence (blue), only shear (orange), or both convergence and shear combined (green).
+These results represent an idealized scenario, assuming perfect measurements without experimental errors.}
+% \end{figure}
+% \begin{figure}[h!]
+% 	\centering
+% 	\includegraphics[width=0.6\textwidth]{figures/SNR_results_theoretical.png}
+% 	\label{fig:SNR_results_theoretical}
+% 	\caption{Relative uncertainty in the measurement of the lensing bispectra amplitude (SNR) as a function of $l_{\text{max}}$.
+The setup and assumptions are identical to those described in the previous figure \ref{fig:fisher_results_theoretical}.}
+% \end{figure}
+
+\bibliographystyle{alpha} % Choose your citation style
+\bibliography{sources}
+
+\appendix
+\section{Shear equals twice spin raised lensing potential}
 
@@
 
-% \section{Fisher matrices raw data}
-
-% Bispectra Fisher matrices:\\
-% \begin{gather*}
-% F^B_{\psi_c} =
-% \begin{pmatrix}
-% 4\times 10^{-3} & 6\times 10^{-4} & -6\times 10^{-3} & -9\times 10^{-3} & 1\times 10^{-4} & -4\times 10^{-3} & -1\times 10^{-3}\\
-% 6\times 10^{-4} & 1\times 10^{-4} & -1\times 10^{-3} & -2\times 10^{-3} & 2\times 10^{-5} & -6\times 10^{-4} & -2\times 10^{-4}\\
-% -6\times 10^{-3} & -1\times 10^{-3} & 1\times 10^{-2} & 1\times 10^{-2} & -2\times 10^{-4} & 6\times 10^{-3} & 2\times 10^{-3}\\
-% -9\times 10^{-3} & -2\times 10^{-3} & 1\times 10^{-2} & 2\times 10^{-2} & -2\times 10^{-4} & 8\times 10^{-3} & 3\times 10^{-3}\\
-% 1\times 10^{-4} & 2\times 10^{-5} & -2\times 10^{-4} & -2\times 10^{-4} & 3\times 10^{-6} & -1\times 10^{-4} & -3\times 10^{-5}\\
-% -4\times 10^{-3} & -6\times 10^{-4} & 6\times 10^{-3} & 8\times 10^{-3} & -1\times 10^{-4} & 4\times 10^{-3} & 1\times 10^{-3}\\
-% -1\times 10^{-3} & -2\times 10^{-4} & 2\times 10^{-3} & 3\times 10^{-3} & -3\times 10^{-5} & 1\times 10^{-3} & 4\times 10^{-4}
-% \end{pmatrix},
-% \\
-% F^B_{\psi_s} =
-% \begin{pmatrix}
-% 3\times 10^{-1} & 4\times 10^{-2} & -3\times 10^{-1} & -1\times 10^{-1} & 5\times 10^{-3} & -1\times 10^{-1} & -1\times 10^{-1}\\
-% 4\times 10^{-2} & 4\times 10^{-3} & -3\times 10^{-2} & -2\times 10^{-2} & 5\times 10^{-4} & -1\times 10^{-2} & -1\times 10^{-2}\\
-% -3\times 10^{-1} & -3\times 10^{-2} & 2\times 10^{-1} & 1\times 10^{-1} & -3\times 10^{-3} & 1\times 10^{-1} & 1\times 10^{-1}\\
-% -1\times 10^{-1} & -2\times 10^{-2} & 1\times 10^{-1} & 7\times 10^{-2} & -2\times 10^{-3} & 6\times 10^{-2} & 5\times 10^{-2}\\
-% 5\times 10^{-3} & 5\times 10^{-4} & -3\times 10^{-3} & -2\times 10^{-3} & 6\times 10^{-5} & -2\times 10^{-3} & -2\times 10^{-3}\\
-% -1\times 10^{-1} & -1\times 10^{-2} & 1\times 10^{-1} & 6\times 10^{-2} & -2\times 10^{-3} & 6\times 10^{-2} & 5\times 10^{-2}\\
-% -1\times 10^{-1} & -1\times 10^{-2} & 1\times 10^{-1} & 5\times 10^{-2} & -2\times 10^{-3} & 5\times 10^{-2} & 5\times 10^{-2}
-% \end{pmatrix},
-% \\
-% F^B_{\psi_{s + c}} =
-% \begin{pmatrix}
-% 4\times 10^{-1} & 4\times 10^{-2} & -3\times 10^{-1} & -2\times 10^{-1} & 5\times 10^{-3} & -2\times 10^{-1} & -1\times 10^{-1}\\
-% 4\times 10^{-2} & 4\times 10^{-3} & -3\times 10^{-2} & -2\times 10^{-2} & 6\times 10^{-4} & -2\times 10^{-2} & -1\times 10^{-2}\\
-% -3\times 10^{-1} & -3\times 10^{-2} & 2\times 10^{-1} & 2\times 10^{-1} & -4\times 10^{-3} & 1\times 10^{-1} & 1\times 10^{-1}\\
-% -2\times 10^{-1} & -2\times 10^{-2} & 2\times 10^{-1} & 1\times 10^{-1} & -3\times 10^{-3} & 8\times 10^{-2} & 7\times 10^{-2}\\
-% 5\times 10^{-3} & 6\times 10^{-4} & -4\times 10^{-3} & -3\times 10^{-3} & 7\times 10^{-5} & -2\times 10^{-3} & -2\times 10^{-3}\\
-% -2\times 10^{-1} & -2\times 10^{-2} & 1\times 10^{-1} & 8\times 10^{-2} & -2\times 10^{-3} & 7\times 10^{-2} & 6\times 10^{-2}\\
-% -1\times 10^{-1} & -1\times 10^{-2} & 1\times 10^{-1} & 7\times 10^{-2} & -2\times 10^{-3} & 6\times 10^{-2} & 5\times 10^{-2}
-% \end{pmatrix}.
-% \end{gather*}
-
-% Powerspectra Fisher matrices:
-% \begin{gather*}
-% F^P_{\psi_c} =
-% \begin{pmatrix}
-% 2\times 10^{-2} & 4\times 10^{-3} & -4\times 10^{-2} & 8\times 10^{-3} & 9\times 10^{-4} & -3\times 10^{-2} & -1\times 10^{-2}\\
-% 4\times 10^{-3} & 1\times 10^{-3} & -9\times 10^{-3} & -6\times 10^{-4} & 2\times 10^{-4} & -5\times 10^{-3} & -2\times 10^{-3}\\
-% -4\times 10^{-2} & -9\times 10^{-3} & 8\times 10^{-2} & -9\times 10^{-3} & -2\times 10^{-3} & 6\times 10^{-2} & 2\times 10^{-2}\\
-% 8\times 10^{-3} & -6\times 10^{-4} & -9\times 10^{-3} & 2\times 10^{-2} & 3\times 10^{-4} & -1\times 10^{-2} & -4\times 10^{-3}\\
-% 9\times 10^{-4} & 2\times 10^{-4} & -2\times 10^{-3} & 3\times 10^{-4} & 3\times 10^{-5} & -1\times 10^{-3} & -4\times 10^{-4}\\
-% -3\times 10^{-2} & -5\times 10^{-3} & 6\times 10^{-2} & -1\times 10^{-2} & -1\times 10^{-3} & 4\times 10^{-2} & 2\times 10^{-2}\\
-% -1\times 10^{-2} & -2\times 10^{-3} & 2\times 10^{-2} & -4\times 10^{-3} & -4\times 10^{-4} & 2\times 10^{-2} & 6\times 10^{-3}
-% \end{pmatrix},
-% \\
-% F^P_{\psi_s} =
-% \begin{pmatrix}
-% 2\times 10^{-1} & 1\times 10^{-2} & -1\times 10^{-1} & -3\times 10^{-2} & 2\times 10^{-3} & -6\times 10^{-2} & -7\times 10^{-2}\\
-% 1\times 10^{-2} & 1\times 10^{-3} & -9\times 10^{-3} & -3\times 10^{-3} & 2\times 10^{-4} & -5\times 10^{-3} & -6\times 10^{-3}\\
-% -1\times 10^{-1} & -9\times 10^{-3} & 7\times 10^{-2} & 2\times 10^{-2} & -1\times 10^{-3} & 4\times 10^{-2} & 4\times 10^{-2}\\
-% -3\times 10^{-2} & -3\times 10^{-3} & 2\times 10^{-2} & 9\times 10^{-3} & -4\times 10^{-4} & 1\times 10^{-2} & 1\times 10^{-2}\\
-% 2\times 10^{-3} & 2\times 10^{-4} & -1\times 10^{-3} & -4\times 10^{-4} & 3\times 10^{-5} & -7\times 10^{-4} & -8\times 10^{-4}\\
-% -6\times 10^{-2} & -5\times 10^{-3} & 4\times 10^{-2} & 1\times 10^{-2} & -7\times 10^{-4} & 2\times 10^{-2} & 2\times 10^{-2}\\
-% -7\times 10^{-2} & -6\times 10^{-3} & 4\times 10^{-2} & 1\times 10^{-2} & -8\times 10^{-4} & 2\times 10^{-2} & 3\times 10^{-2}
-% \end{pmatrix},
-% \\
-% F^P_{\psi_{s+c}} =
-% \begin{pmatrix}
-% 9\times 10^{-2} & 9\times 10^{-3} & -7\times 10^{-2} & -1\times 10^{-2} & 1\times 10^{-3} & -4\times 10^{-2} & -4\times 10^{-2}\\
-% 9\times 10^{-3} & 1\times 10^{-3} & -9\times 10^{-3} & -2\times 10^{-3} & 2\times 10^{-4} & -5\times 10^{-3} & -4\times 10^{-3}\\
-% -7\times 10^{-2} & -9\times 10^{-3} & 7\times 10^{-2} & 8\times 10^{-3} & -1\times 10^{-3} & 5\times 10^{-2} & 3\times 10^{-2}\\
-% -1\times 10^{-2} & -2\times 10^{-3} & 8\times 10^{-3} & 2\times 10^{-2} & -7\times 10^{-5} & -3\times 10^{-4} & 5\times 10^{-3}\\
-% 1\times 10^{-3} & 2\times 10^{-4} & -1\times 10^{-3} & -7\times 10^{-5} & 3\times 10^{-5} & -9\times 10^{-4} & -6\times 10^{-4}\\
-% -4\times 10^{-2} & -5\times 10^{-3} & 5\times 10^{-2} & -3\times 10^{-4} & -9\times 10^{-4} & 3\times 10^{-2} & 2\times 10^{-2}\\
-% -4\times 10^{-2} & -4\times 10^{-3} & 3\times 10^{-2} & 5\times 10^{-3} & -6\times 10^{-4} & 2\times 10^{-2} & 2\times 10^{-2}
-% \end{pmatrix}.
-% \end{gather*}
-
 \end{document}

### 2025-04-04 paper/paper.tex
fisher matrices updated
paper changes

@@
-% !TEX root = /Users/jonasfrugte/Desktop/Research_Project/paper/paper.tex
+% !TEX root = /Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/paper/paper.tex
 \documentclass[11pt]{article} % Choose a document class, like 'article', 'report', or 'revtex4-2' for scientific papers
@@
 
-\affil[1]{Independent Researcher $\infty$, The Netherlands}
+\affil[1]{Independent Researcher, The Netherlands}
 \affil[2]{Van Swinderen Institute for Particle Physics and Gravity, University of Groningen, Nijenborgh 4, 9747 AG Groningen, The Netherlands}
@@
 \begin{document}
+\begin{center}
+    \textit{Please note:}\\ This is still a draft version of the paper I am planning to finish April 2025.
+It thus may still contain mistakes, missing citations, typo's, etc.
+that will not be present in the final version.
+The code used to write this paper can be found on my GitHub: \url{https://github.com/Jonas-Frugte/fisher_calc_weak_lensing}.
+\end{center}
+\pagebreak
 \maketitle
@@
 \begin{abstract}
-    (Placeholder text written with chatgpt) We present a Fisher forecast analysis of the constraints on cosmological parameters from measurements of CMB lensing and galaxy weak lensing power- and bispectra.
-Using simulated data consistent with upcoming surveys, we explore the sensitivity of lensing observables to parameters including the Hubble constant \(H_0\), baryon density \(\Omega_b h^2\), dark matter density \(\Omega_c h^2\), scalar spectral index \(n_s\), amplitude of scalar fluctuations \(A_s\), and the sum of neutrino masses \(\sum m_\nu\).
-
-    Our results indicate that combining power- and bispectra improves parameter constraints compared to power spectra alone, reducing uncertainties by a factor of \(\sim X\%\).
-In particular, we forecast constraints of \(\sigma(H_0) = X.X\) km/s/Mpc, \(\sigma(\Omega_b h^2) = X.XXX\), \(\sigma(\Omega_c h^2) = X.XXX\), \(\sigma(n_s) = X.XXXX\), \(\sigma(A_s) = X.XX \times 10^{-9}\), and \(\sigma(\sum m_\nu) = X.XX\) eV.
-These results demonstrate the potential of bispectrum measurements in future surveys such as the Simons Observatory, CMB-S4, LSST, and Euclid.
-
-    Additionally, we investigate the impact of different noise models and survey redshift distributions, finding that improvements in galaxy redshift accuracy significantly enhance constraints on \(\sum m_\nu\).
-We also discuss the degeneracies between parameters and assess their impact using eigenvalue analysis of the Fisher matrix.
-
-    This study highlights the advantages of combining lensing power- and bispectra for cosmological parameter estimation and provides a roadmap for optimizing future surveys to maximize their scientific return.
+    Weak gravitational lensing of the Cosmic Microwave Background (CMB) and distant galaxies provides a powerful probe of the large-scale structure and cosmic expansion history.
+Upcoming Stage-IV surveys like the Simons Observatory, CMB-S4, LSST, and Euclid promise unprecedented precision, motivating detailed forecasts of their constraining power.
+While lensing power spectra are standard cosmological tools, non-Gaussianities induced by structure formation encode additional information in the bispectrum.
+In this work, we perform a comprehensive Fisher matrix forecast to quantify the detectability of these signals and the cosmological constraints achievable by combining power spectrum and bispectrum measurements from both CMB lensing and galaxy shear, including their cross-correlations.
+We model Stage-IV-like experimental specifications and forecast constraints on the standard $\Lambda$CDM parameters, the sum of neutrino masses ($\Sigma m_\nu$), and the dark energy equation of state ($w_0$).
+We demonstrate that incorporating bispectra significantly enhances constraints, reducing marginalized parameter uncertainties by up to $\sim \texttt{X}\%$ compared to power spectra alone.
+We project competitive constraints, including $\sigma(H_0) = \texttt{X.X~km/s/Mpc}$, $\sigma(\sigma_8) = \texttt{X.XXX}$, and $\sigma(\Sigma m_\nu) = \texttt{X.XX~eV}$.
     \end{abstract}
@@
 
-All code written for this project is uploaded on github at LINK .
+All code written for this project is uploaded on github at \url{https://github.com/Jonas-Frugte/fisher_calc_weak_lensing}.
 
@@
 \end{equation}
-you can relate it to the multipole moments of the lensing potential through CHECK FACTOR OF 2
+you can relate it to the multipole moments of the lensing potential through
 \begin{equation}
@@
 The time complexity of the Fisher matrix scales with the maximum multipole considered, $l_{max}$, as $\mathcal O (l_{max}^3)$, leading to time issues for realistic values of $l_{max}$.
-We thus calculated the $l_1 < l_2 < l_3$ part of each Fisher matrix element (see appendix (REF)) by sampling every step of 10 (CHECK IF THIS IS STILL CORRECT BEFORE PUBLISHING) in $l_1$ and $l_2$.
+We thus calculated the $l_1 < l_2 < l_3$ part of each Fisher matrix element (see the appendix) by sampling every step of 10 %(CHECK IF THIS IS STILL CORRECT BEFORE PUBLISHING)
+in $l_1$ and $l_2$.
 Due to the discontinuos nature of the bispectrum (e.g.
@@
 
-For the parameter constraints we took the multipole range to be from 50 ($l_{min}$) to 1000 ($l_{max}$) (INCREASE TO 2000?) (CHECK BEFORE PUBLISHING) in line with measurement ranges from current gen surveys (PROOF BY LOOKING AT ACTUALL EXPERIMENTS).
-
-We will consider multiple possibilities for the other calculation parameters (noise levels and redshift distribution of observed galaxies).
+For the parameter constraints we took the multipole range to be from 2 ($l_{min}$) to 2000 ($l_{max}$).
+
+% We will consider multiple possibilities for the other calculation parameters (noise levels and redshift distribution of observed galaxies).
 This is done to allow us to more accurately model constraints from measurements by the Simons Observatory (SO), CMB-S4, Euclid, and Large Synoptic Survey Telescope (LSST) surveys.
@@
 
-The noise levels in the CMB lensing are caused by a variety of factors, see e.g.
-\cite{cmblensingestimator}.
-The noise curves for the SO are shown alongside a simulated pure convergence power spectrum curve in figure \ref{fig:lpsplusnoise}.
-This is the noise source we use for CMB lensing.
+For noise in CMB lensing we consider lensing estimated by an experiment measuring polarization with white noise $\Delta_P$ and beam spread $\sigma$.
+We then use the quadratic estimator from \cite{Hu_2002} to estimate the resulting noise in the lensing power spectrum.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/spectraplusnoise.pdf}
-    \caption{CMB and galaxy lensing potential power spectra with noise.
-The left panel shows the galaxy lensing potential power spectrum with LSST and Euclid noise, while the right panel shows the CMB lensing potential power spectrum with OS noise.}
+    \includegraphics[width=\textwidth]{/Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/code/plots/spectraplusnoise.pdf}
+    \caption{CMB (right) and galaxy (left) lensing potential power spectra compared to associated experimental noise.
+Current (stage 3) noise values are displayed as well as near future (stage 4) noi se values.
+The CMB lensing experiment does not measure temperature anisotropies, only polarization.
+CMB noise values are chosen in accordance with \cite{Namikawa_2016}.
+Shear noise values are chosen to be similar to e.g.
+Euclid measurements for stage 4 and e.g.
+KiDS for stage 3.}
     \label{fig:lpsplusnoise}
@@
 and $\sigma^2_\gamma$ is the intrinsic ellipticity dispersion. 
-Euclid and the LSST both have $\sigma_{rms} \approx 0.26$ \cite{euclid1overview} \cite{lsstsciencebook}.
-Euclid is expected to measure a galaxy density of approximately 30 galaxies per square arcminute \cite{euclidprep1}
-while LSST aims to measure 26 \cite{galaxydenslsst}
-The noise levels are graphed alongside a simulated shear power spectrum in figure \ref{fig:lpsplusnoise}. We perform our calculations with Euclid noise levels throughout this paper.
+Experiments such as Euclid and the LSST both have $\sigma_{rms} \approx 0.3$ \cite{euclid1overview} \cite{lsstsciencebook} and 
+$ n_g\approx 30 \text{ arcmin}^{-2}$.
+
+The noise values are calculated for stage 3 (current) surveys and stage 4 (near future) surveys.
+We will use the stage 4 noise values for parameter constraints.
+A comparison of stage 3 noise, stage 4 noise, and lensing signal can be seen in figure \ref{fig:lpsplusnoise}.
+
 The redshift distribution of the observed galaxies is commonly parameterized as \cite{bartelmann2001weak}
@@
 \subsection{SNR}
-The signal to noise ratios of the CMB and galaxy lensing power- and bispectra v.s.
+Lensing power spectra are well detectable in nearly all cases. 
+The signal to noise ratios of the CMB and galaxy lensing bispectra v.s.
 the maximum multipole moment measured can be found in figure \ref{fig:snrplots}.
-It can be seen that the CMB bispectrum signal to noise never exceeds 2 even for $l_{\max} = 2000$, meaning that this signal is not likely to be able to contribute to parameter constraints in future surveys.
+Shear bispectra can be measured well by both stage 3 and stage 4 experiments, while CMB lensing is only semi detectable by a stage 3 experiment but well detectable by stage 4 experiments.
+%It can be seen that the CMB bispectrum signal to noise never exceeds 2 even for $l_{\max} = 2000$, meaning that this signal is not likely to be able to contribute to parameter constraints in future surveys.
 All other quantities should be well detectable for $l_{\max}\geq 500$.
@@
     \caption{Signal to noise ratios for galaxy lensing (dashed) and CMB lensing (non-dashes) power- (in red) and bispectra (in blue).
-The $x$-axis labels the maximum multipole moment measured up to, starting from $l_{\min}=50$.}
+The $x$-axis labels the maximum multipole moment measured up to, starting from $l_{\min}=2$.}
     \label{fig:snrplots}

### 2025-04-04 paper/figures/fisher_vs_lmaxes_plot copy.py

@@ -1,0 +1,112 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+SNR2_c = np.array([21752.37109375, 42467.37109375, 61431.392578125, 80951.72265625, 103566.880859375, 127987.7421875])
+
+SNR2_s = np.array([58.471466064453125, 209.129638671875, 472.9197998046875, 857.3757019042969, 1365.5498962402344, 1985.4261779785156])
+
+SNR2_full = np.array([15973925.273764089, 21700691.52666128, 25006818.371009313, 27387274.95562951, 29266159.18653201, 31273214.900456935])
+
+# Given Data
+l_values = np.array([500, 1000, 1500, 2000, 2500, 3000])
+
+Fisher_c = {
+    "H0": np.array([14.077168464660645, 31.897847175598145, 50.746644020080566, 72.83032321929932, 99.00564861297607, 127.41051578521729]),
+    "ombh2": np.array([4112962.0, 9336375.0, 20095801.0, 26203080.5, 38208011.5, 53317144.5]),
+    "omch2": np.array([13600145.0, 32862975.0, 52410357.0, 77972975.0, 110291517.0, 147623545.0]),
+    "As": np.array([1.9673580907200617e+22, 4.143315935600502e+22, 6.3830039936278815e+22, 8.84181108864527e+22, 1.1733016440005867e+23, 1.5001670052396254e+23]),
+    "ns": np.array([12025.2470703125, 28862.3662109375, 55617.6474609375, 118703.4716796875, 219686.6748046875, 355107.7529296875]),
+    "mnu": np.array([20274.103515625, 44399.357421875, 67587.205078125, 92397.80859375, 121286.11328125, 154003.509765625]),
+}
+
+Fisher_s = {
+    "H0": np.array([0.27174660563468933, 0.980530709028244, 2.1956381499767303, 3.912103980779648, 6.173850864171982, 8.896407455205917]),
+    "ombh2": np.array([43218.79296875, 174591.40234375, 407098.21484375, 743700.71484375, 1195916.71484375, 1761281.83984375]),
+    "omch2": np.array([62721.1875, 276152.34375, 634106.46875, 1145792.4375, 1821605.5625, 2635287.3125]),
[…]

### 2025-04-04 paper/figures/snr_plots.py

@@ -1,0 +1,54 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+# Define data
+ell_max = [250, 500, 750, 1000, 1250, 1500, 1750, 2000]
+
+# snrp_c = [28177.879112088674, 96206.55433780684, 175853.35058235622, 243509.61866955983, 289830.6864329934, 317440.70691750693, 332633.4570946259, 340837.0890864015]
+
+# snrp_s = [23441.8759542758, 67396.65825513331, 113536.84597591546, 156335.21499038063, 194043.68425517925, 226409.47923775774, 253826.42455016664, 276933.44511557865]
+
+# new
+# Convergence bispectrum SNR^2:
+# [0.4606180302137393, 2.428277643136825, 4.556412710716516, 5.924272288700067, 6.620955446735307, 6.958473094521213, 7.12003774130501, 7.201717334076159]
+# Shear bispectrum SNR^2:
+# [14.937583099693954, 44.263581102561155, 70.4841150784228, 91.60588005582235, 108.3741775299598, 121.44133496518877, 131.79346525781503, 139.98497349255402]
+
+snrb_c_s3 = [0.4606180302137393, 2.428277643136825, 4.556412710716516, 5.924272288700067, 6.620955446735307, 6.958473094521213, 7.12003774130501, 7.201717334076159]
+
+snrb_s_s3 = [14.937583099693954, 44.263581102561155, 70.4841150784228, 91.60588005582235, 108.3741775299598, 121.44133496518877, 131.79346525781503, 139.98497349255402]
+
+snrb_c_s4 = [0.7032144111436794, 6.542889778749563, 20.500697932075894, 39.87427410664932, 59.254915437097765, 74.95646527206883, 86.07133128729137, 93.3095234367019]
+
+snrb_s_s4 = [50.95719570421603, 307.30117549383294, 748.571492858229, 1289.9230963915802, 1863.479395132727, 2419.0805380400534, 2928.1361348954356, 3387.6648308426447]
+
+# Set figure size
+plt.figure(figsize=(7,5))
+
+# Plot each array
+plt.plot(ell_max, np.sqrt(snrb_c_s3), marker='x', label='CMB, stage 3', linestyle = '--', color='red')
[…]

### 2025-04-04 paper/fisherstuff.py

@@ -1,0 +1,401 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.patches import Ellipse
+from scipy.stats import norm
+from numpy.linalg import inv
+
+np.set_printoptions(
+    precision=1,
+    suppress=False, 
+    formatter={'float_kind': lambda x: f"{x:.2e}"}
+)
+
+fisher_matrices_dir = '/Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/code/fisher_matrices'
+
+######################
+# Bispectra Matrices #
+######################
+visb_c = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_c.txt')
+
+visb_s = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_s.txt')
+
+visb_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_both.txt')
+
+########################
+# Powerspectra Matrices
+########################
+visp_c = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_c.txt')
+
+visp_s = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_s.txt')
+
+visp_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_both.txt')
+
+########### NORMALIZATION ###########
+
+param_values_normalization = np.array([
+    67.4,         # H_0
+    0.0224,       # Omega_b h^2
+    0.120,        # Omega_c h^2
+    0.965,        # n_s
+    0.06,         # m_\nu
+    2.1e-9,       # A_s
+    -1.0          # w_0
+])
+
+normalization_matrix = np.outer(param_values_normalization, param_values_normalization)
+# If needed, multiply each matrix by this normalization_matrix.
+
+########### SIGMA8 ################
+
+def append_row_column(mats, vec):
+    """
+    For each matrix in 'mats':
+      - Enlarge the matrix by 1 row + 1 column
+      - New column entries: old_row(mat) · vec
+      - New row same as new column
+      - New bottom-right corner = sum( mat * outer(vec, vec) )
+      - Returns an updated list (same references replaced).
+    """
+    matsnew = mats
+    for i in range(len(mats)):
+        mat = mats[i]
+        N = mat.shape[0]
+        
+        # Compute new col/row
+        new_col = mat @ vec  # shape (N,)
[…]

### 2025-04-04 paper/old/fisherstuff OLD.py

@@ -1,0 +1,11 @@
+import itertools
+import numpy as np
+
+mat = np.zeros((3, 3))
+
+for i, j in itertools.product(range(3), repeat = 2):
+    if i >= j:
+        x = np.random.rand()
+        mat[i, j] = x
+        # mat[j, i] = x
+print(mat)

### 2025-04-04 paper/old/fisherstuff_nonoise.py

@@ -1,0 +1,326 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.patches import Ellipse
+from scipy.stats import norm
+from numpy.linalg import inv
+
+np.set_printoptions(
+    precision=1,
+    suppress=False, 
+    formatter={'float_kind': lambda x: f"{x:.2e}"}
+)
+
+######################
+# Bispectra Matrices #
+######################
+# Convergence (bisp)
+visb_c = np.array([
+    [ 1.03291163e+00,  2.06865762e-01, -1.73748438e+00, -3.36334319e+00,  2.57507000e-02, -9.02272360e-01, -3.31060410e-01],
+    [ 2.06865762e-01,  4.21529585e-02, -3.47999633e-01, -6.88880813e-01,  5.15774102e-03, -1.79857024e-01, -6.60428685e-02],
+    [-1.73748438e+00, -3.47999633e-01,  2.92890430e+00,  5.69889944e+00, -4.32792348e-02,  1.51535768e+00,  5.56382986e-01],
+    [-3.36334319e+00, -6.88880813e-01,  5.69889944e+00,  1.12884346e+01, -8.38662127e-02,  2.92025774e+00,  1.07246662e+00],
+    [ 2.57507000e-02,  5.15774102e-03, -4.32792348e-02, -8.38662127e-02,  6.41959141e-04, -2.24938576e-02, -8.25204814e-03],
+    [-9.02272360e-01, -1.79857024e-01,  1.51535768e+00,  2.92025774e+00, -2.24938576e-02,  7.89137341e-01,  2.89453449e-01],
+    [-3.31060410e-01, -6.60428685e-02,  5.56382986e-01,  1.07246662e+00, -8.25204814e-03,  2.89453449e-01,  1.06244427e-01]
+])
+
+# Shear (bisp)
+visb_s = np.array([
+    [ 1.88500828e+02,  2.23183949e+01, -1.56671938e+02, -1.21036153e+02,  2.64358967e+00, -8.11044392e+01, -6.82077843e+01],
+    [ 2.23183949e+01,  2.67274426e+00, -1.86760429e+01, -1.45708725e+01,  3.14533131e-01, -9.65684766e+00, -8.07652633e+00],
+    [-1.56671938e+02, -1.86760429e+01,  1.30780412e+02,  1.01553923e+02, -2.20358748e+00,  6.76397718e+01,  5.66880738e+01],
[…]

### 2025-04-04 paper/sigma8_constraint.py

@@ -1,0 +1,11 @@
+import numpy as np
+import fisherstuff
+
+visbpp_f = fisherstuff.visb_f # + fisherstuff.visp_f
+
+s8_ders = np.array([2.76945054e-03, -6.50570238e+00,  4.30234556e+00,  3.06272435e-01, 1.93167857e+08, -2.15681667e-01, -2.01175433e-01])
+
+s8_const = np.sum(np.outer(s8_ders, s8_ders) * visbpp_f)
+print(s8_const)
+
+print((visbpp_f[1, 1] + 2 * visbpp_f[1, 2] + visbpp_f[2, 2])**(-1/2))

### 2025-04-05 paper/paper_draft.tex
paper changes

--- 2025-04-04 paper/paper.tex
@@
-% !TEX root = /Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/paper/paper.tex
+% !TEX root = /Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/paper/paper_draft.tex
 \documentclass[11pt]{article} % Choose a document class, like 'article', 'report', or 'revtex4-2' for scientific papers
@@
 \affil[ ]{\textit{E-mail:} \texttt{jonasfrugte@gmail.com}}
-
-\date{20 February 2025} % Removes the default date
+ % Removes the default date
 
 \begin{document}
-\begin{center}
-    \textit{Please note:}\\ This is still a draft version of the paper I am planning to finish April 2025.
-It thus may still contain mistakes, missing citations, typo's, etc.
-that will not be present in the final version.
-The code used to write this paper can be found on my GitHub: \url{https://github.com/Jonas-Frugte/fisher_calc_weak_lensing}.
-\end{center}
-\pagebreak
 \maketitle
@@
 % \vspace{0.5cm}
+
+\begin{center}
+    \textbf{NB: Currently still in draft.}
+\end{center}
 \hrule
@@
 \begin{abstract}
-    Weak gravitational lensing of the Cosmic Microwave Background (CMB) and distant galaxies provides a powerful probe of the large-scale structure and cosmic expansion history.
+    % Weak gravitational lensing of the Cosmic Microwave Background (CMB) and distant galaxies provides a powerful probe of the large-scale structure and cosmic expansion history.
 Upcoming Stage-IV surveys like the Simons Observatory, CMB-S4, LSST, and Euclid promise unprecedented precision, motivating detailed forecasts of their constraining power.

### 2025-04-08 paper/paper_draft.tex
paper changes

@@
 \usepackage[backend=biber, style=numeric]{biblatex}
-\addbibresource{sources.bib}
+\addbibresource{/Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/paper/sources.bib}
 
@@
 \DeclareRobustCommand{\Cov}{\ifmmode\text{Cov}\else d\fi}
+\DeclareRobustCommand{\CMB}{\ifmmode\text{CMB}\else d\fi}
+\DeclareRobustCommand{\gal}{\ifmmode\text{gal}\else d\fi}
+
 \newcommand{\br}[1]{\ensuremath{\left( #1 \right)}}
@@
 
-\section{Weak Lensing}\label{sec:weaklensing}
-\subsection{Approximations Made / Regime}
-\subsection{Perturbed Photon Paths}
-Work in \textbf{conformal newtonian gauge}.
-In this section we'll work in units where $c=1$.
-Denoting conformal time and conformal radial distance by $\eta$ and $\chi$, respectively, the perturbed line element is given by
-\begin{equation}
-    \d s^2 = a^2(\eta)((1 + 2\Psi_N) \d \eta^2 - (1 + 2\Phi_N) \gamma_{ij}\d x^i \d x^j) 
-\end{equation}
-
-where $\gamma_{ij}$ is the unperturbed line element
-\begin{equation}
-    \gamma_{ij} = \d x^i \d x^j = \d \chi^2 + f^2_K(\chi)(\d \theta^2 + \sin ^2 \theta \d \phi ^2).
-\end{equation}
-From here of on work in first order in the scalar potentials $\Phi_N$ and $\Psi_N$. For lensing we consider null-geodesics so $\d s^2 = 0$ and we can rewrite the perturbed line element as
-\begin{equation}
-    \d \hat s^2 = (1 + 4\Psi)\d\eta^2 - \gamma_{ij} \d x^i \d x^j
-\end{equation}
-with $\Psi$ the \textbf{Weyl Potential} given by $\Psi := (\Psi_N - \Phi_N)/2$. A null geodesic $x^\mu(\hat\lambda)$ in terms of its affine parameter ($\hat \lambda$) satisfies the geodesic equation
-\begin{equation}
-    \frac{\d^2 x^\mu}{\d \hat\lambda^2} + \Gamma^\mu_{\nu\rho} \frac{\d x^\nu}{\d \hat\lambda}\frac{\d x^\rho}{\d\hat\lambda} = 0,
-\end{equation}
-with $\hat g_{\mu\nu}(\d x^\mu/\d \hat\lambda)(\d x^\nu / \d \hat \lambda) = 0$.
-
-In the unperturbed there are incoming radial solutions of the form $\chi = \eta_0 - \eta$. We can express $\hat\lambda$ in terms of $\eta$ using The 0-component of the geodesic equation,
-\begin{equation}
-    \frac{\d^2\eta}{\d\hat\lambda^2} + 2(\frac{\d\eta}{\d\hat\lambda})^2\frac{\d\Psi}{\d\eta}+2\frac{\d\eta}{\d\hat\lambda}\frac{\d x^i}{\d\hat\lambda}\frac{\partial\Psi}{\partial x^i} = 0,
-\end{equation}
-where the derivative $\d\Psi/\d\eta = \partial_\eta\Psi+(\d x^i/\d\eta)\partial_i\Psi$ is along the perturbed ray. This allows us to rewrite the other (spatial) geodesic equations as
-\begin{equation}
-    \frac{\d^2x^i}{\d\eta^2} - 2\frac{\d x^i}{\d\eta}(\frac{\d\Psi}{\d\eta} + \frac{\d x^j}{\d\eta}\frac{\partial\Psi}{\partial x^j})+2\gamma^{ij}\frac{\partial\Psi}{\partial x^j}+\bar\Gamma^i_{jk}\frac{\d x^j}{\d\eta}\frac{\d x^k}{\d \eta} = 0,
-\end{equation}
-where $\bar\Gamma^i_{jk}$ are the connection coefficients of the unperturbed three geometry $\gamma_{ij}$.
-Without loss of generality we consider an observer located at the origin of the spatial coordinates, meaning we are interested in rays that end at $x^i = 0$.
-In that case $\d \chi/\d \eta = -1 + O(\Psi)$, $\d \theta/\d\eta = O(\Psi)$ and $\d \phi/\d\eta = O(\Psi)$ (with $\theta$ and $\phi$ angular spatial coordinates).
-For the case of such rays we can evaluate the connection coefficients and rewrite the spatial geodesic equations as
-\begin{gather}
-    \frac{\partial^2\chi}{\d\eta^2}+2\frac{\d\Psi}{\d\eta}=0,\\
-    \frac{\d^2\theta}{\d\eta^2}-2\frac{\d\ln f_K(\chi)}{\d\chi}\frac{\d\theta}{\d\eta}+\frac{2}{f^2_K(\chi)}\frac{\partial\Psi}{\partial\theta}=0,\\
-    \frac{\d^2\phi}{\d\eta^2} - 2\frac{\d \ln f_K(\chi)}{\d\chi}\frac{\d\phi}{\d\eta}+\frac{2}{f^2_K(\chi)}\frac{1}{\sin^2(\theta)}\frac{\partial\Psi}{\partial\phi}=0.
-\end{gather}
-Integrating twice and using the null condition we can rewrite the the first equation as
-\begin{equation}
-    \chi = \eta_0 - \eta - 2\int_{\eta_0}^\eta\Psi\d\eta',
-\end{equation}
-where the integral is along the ray.
-Since we are working at first order in $\Psi$ we can evaluate the integral along the unperturbed path, this is known as the \textbf{Born approximation}.\textit{Next follows a short discussion of why perturbations in observation time can be ignored, include?} Doing the same for the other two equations gives
-% TODO: make n a proper vector (bold)
-\begin{gather}
-    \theta(\eta_0 - \chi_*)=\theta_0 - \int_0^{\chi_*}\d\chi\frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}2\frac{\partial}{\partial \theta}\Psi(\chi\hat n, \eta_0 - \chi),\\
-    \phi(\eta_0 - \chi_*)=\theta_0 - \int_0^{\chi_*}\d\chi\frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}\frac{2}{\sin^2\theta}\frac{\partial}{\partial \phi}\Psi(\chi\hat n, \eta_0 - \chi),
-\end{gather}
-where $\theta_0$ and $\phi_0$ label the line of sight $\hat n$.
-The displacement vector $\mathbf{\alpha}$ indicates the infinitesimal displacement of the lensed rays from their unpetrubed observed angles and is related to $\theta$ and $\phi$ through
-$$
-\alpha_\theta = \theta - \theta_0, \quad \alpha_\phi = \sin^2\theta(\phi - \phi_0).
-$$
-This allows us to summarize the above results as
-\begin{equation}
-    \mathbf \alpha = -2 \int_0^{\chi_*}\d \chi \frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}\nabla_{\hat{\mathbf n}}\Psi(\chi\hat{\mathbf n};\eta_0 - \chi).
-\end{equation}
-When working in the weak field limit (i.e. small spacetime curvature) the $00$ component of Einsteins equations give us a poisson equation \cite{dodelson2020modern},
-\begin{equation}
-    -\nabla^2\Psi(\textbf{x}, \eta) =  4\pi G a(\eta)^2\delta\rho(\mathbf x, \eta)
-\end{equation}
-where $K$ is $0$, $1$, or $-1$ depending on the curvature of the universe, $\Delta = \partial_x^2 + \partial_y^2 + \partial_z^2$ is the 3 dimensional laplace operator, $\delta\bar\rho$ is the comoving total energy perturbation (evaluated in the rest-frame of the total energy).
-%TODO: proper definition for anisotropic stress and explain why we can ignore it
-
-\subsection{Convergence, shear}
-Quantities we are interested in measuring depend on how the lensing vector changes with the observation angle. \textbf{Magnification matrix} is defined by
-\begin{equation}
-    A_{ij} := \delta_{ij} + \frac{\partial}{\partial\theta_i}\alpha_i(\hat{\mathbf n}).
-\end{equation}
-This matrix can then be decomposed in the following form, which immediately gives us definitions for the \textbf{convergence}, $\kappa$, \textbf{shear}, $\gamma_1$ and $\gamma_2$, and \textbf{rotation}, $\omega$:
-\begin{equation}
-    A_ij(\hat{\mathbf n}) = \begin{pmatrix}
-        1 - \kappa - \gamma_1 & -\gamma_2 + \omega \\
-        -\gamma_2 - \omega & 1 - \kappa + \gamma_1
-    \end{pmatrix}.
-\end{equation}
-At 1st order $A$ is a symetric matrix by definition and so $\omega$ has to be 0, we will ignore it from here on out.
-Intuitively, A tells you how a small patch in the sky transforms due to weak lensing, i.e.
-if you have surface brightness $I$ then
-% TODO: make delta thick
-\begin{equation}
-    I_{\text{emitted}}(\hat{\mathbf{n}} + \mathbf{\delta}) = I_{\text{observed}}((\hat{\mathbf{n}} + \mathbf{\delta}) + \mathbf\alpha(\hat{\mathbf{n}} + \mathbf{\delta})) \approx I_{\text{observed}}((\hat{\mathbf{n}} + \mathbf{\alpha} + A_{ij}\delta_{j}).
-\end{equation}
-given that $|A_{ij}| = (1 - \kappa)^2 + \omega^2 - |\gamma|^2 = 1 - 2\kappa + O(\kappa^2, \gamma^2, \omega^2)$, we can interpret $\kappa$ as telling us about the overal magnification of the source.
-The $\gamma_i$ represent the area preserving distortion, i.e.
-stretching and squeezing in a specific direction.
-
-\subsection{lensing potential}
-Taking the derivative out of the integral in equation REF leads us to defining the \textbf{lensing potential}, $\psi$,
-\begin{equation}
-    \mathbf \alpha = \nabla_{\hat{\mathbf n}}\psi, \quad \psi(\hat{\mathbf n}) = -2 \int_0^{\chi_*} \d \chi \frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}\Psi(\chi\hat{\mathbf n}; \eta_0 - \chi).
-\end{equation}
-
-We can relate $\kappa$ and $\gamma$ directly to the lensing potential as
-% TODO: should these subscripts be theta and phi?
-\begin{equation}
-    \kappa = \frac{1}{2}\nabla^2\psi, \quad \gamma_1 = \frac{1}{2}(\partial_\theta^2 - \partial_\phi^2)\psi, \quad \gamma_2 = \partial_\theta\partial_\phi \psi.
-\end{equation}
-
-\subsection{Regarding window functions}
-The window function for light emitted from a source at fixed radius $\chi_*$ is $(\chi-\chi^*)/(\chi_*\chi)$, if the source is instead distributed over the radius as $p(\chi)$, with $p(\chi)$ normalized to integrate to 1, we generalize the window function as
-\begin{equation*}
-    W(\chi) = \int_\chi^{\chi_*}\d\chi' p(\chi')\frac{\chi'-\chi}{\chi'\chi}.
-\end{equation*}
-The integration limit is up to surface of last scattering because further than that no source can contribute to the observed image.
-In the case of an image of the CMB we can take $p(\chi')=\delta(\chi'-\chi_*)$, in which case the window function reduces back to $(\chi_*-\chi)/(\chi_*\chi)$.
-In the case of different sources, we will differentiate between their respective window functions with a subscript.
-In this case the factor that changes is the distribution function.
-The sources considered in this paper are only the CMB and selections of galaxies from galaxy surveys.
-
-\section{Weak Lensing Statistics}\label{sec:weaklensstats}
-\subsection{Lensing Potential Powerspectrum}
-Consider the lensing potential, $\psi$, then decomposing into spherical harmonics gives
-\begin{equation}
-    \psi(\hat{\mathbf n}) = \sum_{lm} \psi_{lm}Y_{lm}(\hat{\mathbf n}).
-\end{equation}
-On the other hand, consider the decomposition of $\Psi$ in fourier modes with the Fourier convention $\Psi(\mathbf x, \eta)=\int\frac{\d^3 \mathbf k}{(2\pi)^{3}}\Psi(\mathbf k, \eta)e^{i\mathbf k \cdot\mathbf x}$,
-\begin{equation}
-    \psi(\hat{\mathbf n}) = -2 \int_0^{\chi_*}\d \chi W(\chi) \int\frac{\d^3 \mathbf k}{(2\pi)^{3}}\Psi(\mathbf k, \eta_0 - \chi)e^{i\mathbf k \cdot\hat{\mathbf n}\chi}.
-\end{equation}
-We can than relate the multipole modes of $\psi$ to the fourier modes of $\Psi$ through % TODO: check sign convention here
-\begin{gather}
-    \psi_{lm} = \langle Y_l^m | \psi \rangle = \int\d^2\hat{\mathbf n} Y_l^m(\hat{\mathbf n})^* \psi(\hat{\mathbf n})\\ 
-    = -2\int\d^2\hat{\mathbf n} Y_l^m(\hat{\mathbf n})^* \int_0^{\chi_*}\d \chi W(\chi) \int\frac{\d^3 \mathbf k}{(2\pi)^{3}}\Psi(\mathbf k, \eta_0 - \chi)e^{i\mathbf k \cdot\hat{\mathbf n}\chi}
-\end{gather}
-Now define the power spectrum as
-\begin{equation}
-    \langle \Psi(\mathbf k, \eta)\Psi^*(\mathbf k',\eta')\rangle = \frac{2\pi^2}{k^3}P_\Psi(k, \eta, \eta')\delta(\mathbf k - \mathbf k'),
-\end{equation}
-with $\eta$ denoting the conformal time. This gives 
-%TODO: define the window function
-\begin{equation}
-    \langle \psi(\hat{\mathbf n})\psi(\hat{\mathbf n}') \rangle = 4\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\int \frac{\d^3\mathbf k}{(2\pi)^6}\frac{2\pi^2}{k^3}P_\psi(k,\eta_0 - \chi, \eta_0-\chi')e^{i\mathbf k\cdot\hat {\mathbf n}\chi}e^{i\mathbf k\cdot\hat {\mathbf n}'\chi'},
-\end{equation}
-where we used that $\eta = \eta_0 - \chi$ along the unperturbed photon path (Born approximation), with $\eta_0$ the time at which the light ray hits earth. 
-We can use the result
-\begin{equation}
-    e^{i\mathbf k \cdot \hat {\mathbf n}\chi}=4\pi\sum_{lm}i^lj_l(k\chi )Y_l^m(\hat{\mathbf n})^*Y_l^m(\hat{\mathbf k}) = 4\pi\sum_{lm}i^lj_l(k\chi )Y_l^m(\hat{\mathbf n})Y_l^m(\hat{\mathbf k})^*,
-\end{equation}
-where $j_l$ is the spherical Bessel function (ELABORATE ON THIS?), to rewrite the above equation.
-Using both versions of the identity above, we immedatialy get a factor $Y_l^m(\hat{\mathbf k})Y_{l'}^{m'}(\hat{\mathbf k})^*$ in our integral.
-We can factor the differential element of $\d^3\mathbf k$ into a radial and angular part as $k^2\d k\d^2\Omega_k$, with $\Omega_k$ the solid angle, to apply the orthonormality condition of the spherical harmonics.
-Note that we take the same sequence of steps a number of times in other parts of the derivations of the lensing spectra.
-We thus obtain
-\begin{gather}
-    \langle \psi(\hat{\mathbf n})\psi(\hat{\mathbf n}') \rangle = 4(4\pi)^2\sum_{ll'mm'}i^{l+l'}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\\\times\int \frac{k^2\d k}{(2\pi)^6}\frac{2\pi^2}{k^3}j_l(k\chi)j_{l'}(k\chi')P_\psi(k,\eta_0 - \chi, \eta_0-\chi')Y_{lm}(\hat{\mathbf n})Y_{l'm'}(\hat{\mathbf n}')^*\delta_{ll'}\delta_{mm'}.
-\label{eq:lenspotcorrexpanded}
-\end{gather}
-The angular power spectrum is defined similarly to the power spectrum, i.e.
-\begin{equation}
-    \langle \psi_{lm}\psi_{l'm'}^* \rangle = \delta_{ll'}\delta_{mm'}C_l^\psi.
-\end{equation}
-Note that the correlation is independent of $m$ and $m'$. %TODO: explain how this relates to statistical isotropy
-We can thus read from equation \ref{eq:lenspotcorrexpanded} that
-\begin{gather}
-    C_l^\psi = 4(4\pi)^2(-1)^{l}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\int \frac{k^2\d k}{(2\pi)^6}\frac{2\pi^2}{k^3}j_l(k\chi)j_{l}(k\chi')P_\psi(k,\eta_0 - \chi, \eta_0-\chi'),
-\end{gather}
-which can be simplified to
-\begin{gather}
-    C_l^\psi = (-1)^{l}\frac{2}{\pi^2}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\int k^2\d kj_l(k\chi)j_{l}(k\chi')\frac{P_\psi(k,\eta_0 - \chi, \eta_0-\chi')}{k^3}.
-\end{gather}
-
-To further evaluate the integral we will make the standard \textbf{Limber approximation}.
-The Bessel functions peak sharply at $l=x$\footnote{Some sources use $x\approx l+1/2$ instead, which is slightly more accurate for larger scales (low $l$) and slightly less accurate for smaller scales.}, with the peak being increasingly sharp for higher $l$.
-Similarly to $\delta(x-x_0)f(x)=\delta(x-x_0)f(x_0)$, we thus take $j_l(k\chi)f(k)\approx j_l(k\chi)f(l/\chi)$.
-The Bessel functions satisfy an orthogonality condtion,
-\begin{equation}
-    \int k^2\d k j_l(k\chi)j_l(k\chi') = \frac{\pi}{2\chi^2}\delta(\chi-\chi').
-\end{equation}
-In combination with the Limber approximation we thus find
-\begin{equation}
-    \int k^2\d k j_l(k\chi)j_l(k\chi')f(k) \approx \frac{\pi}{2\chi^2}\delta(\chi-\chi')f(l/\chi).
-\end{equation}
-%TODO: how high? quantify limber approx, perhaps using convergence x LSS paper
-It allows us to write the Limber-approximate angular spectrum as
-\begin{gather}
-    C^\psi(l) = (-1)^{l}\frac{2}{\pi^2}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi') \frac{\pi}{2\chi^2}\delta(\chi-\chi')\frac{\chi^3}{l^3}P_\psi(l/\chi,\eta_0 - \chi, \eta_0-\chi')\\
-    = (-1)^{l}\frac{1}{l^3\pi}\int_0^{\chi_*}\chi\d \chi W(\chi)^2 P_\psi(l/\chi,\eta_0 - \chi, \eta_0-\chi).
-\end{gather}
-% The power spectrum of the potential can be related to the power spectrum of the matter density using the Poisson equation %TODO: create ref
-% For a flat universe in atter or dark energy domination we thus find the relation
-% \begin{equation}
-%     P_\Psi(k;\eta, \eta)=\frac{9\omega_m^2(\eta)H^4(\eta)}{8\pi^2} \frac{P_m(k, \eta)}{k},
-% \end{equation}
-% where we use the shorthand $P_m(k, \eta):=P_m(k,\eta,\eta)$. The lensing power spectrum is thus finally given in terms of the matter power spectrum as
-% \begin{equation}
-%     C_l^\psi=\underbrace{9\omega_m^2(\eta)H^4(\eta)}_{:=C(\eta)}\frac{1}{l^4}\int_0^{\chi_*}\d\chi\chi^2 W(\chi)^2P_m(l/\chi, \eta_0-\chi).
-% \end{equation}
-% We will henceforth denote the $\eta$ dependent factor outside of the integral as $C(\eta)$.
-\subsection{Lensing potential bispectrum}
-The derivation of the bispectrum proceeds similarly to that of the power spectrum.
-We aim to compute the bispectrum of the lensing potential fields of 3 (possibly distinct sources), $\psi_1$, $\psi_2$, $\psi_3$.
-\begin{gather*}
-    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle = 
-    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{l_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}e^{i\mathbf k_i \cdot\hat{\mathbf n_i}\chi_i}} \\
-    \times \langle \prod_i\Psi(\mathbf k_i, \eta_0 - \chi_i) \rangle.
-\end{gather*}
-Defining the bispectrum of the graviational potential as
-\begin{equation*}
-    \langle \prod_{i=1,2,3}\Psi(\mathbf k_i, \eta_0 - \chi_i) \rangle = (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}),
-\end{equation*}
-we rewrite the lensing potential bispecrum as
-\begin{gather*}
-    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle = 
-    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{l_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}e^{i\mathbf k_i \cdot\hat{\mathbf n_i}\chi_i}}\\ \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
-\end{gather*}
-Now using the identity REF for the complex exponential:
-\begin{gather*}
-    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle \\ = 
-    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{l_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi\sum_{lm}i^{l}j_{l}(k_i\chi_i)Y_{l}^{m}(\hat{\mathbf n_i})Y_{l}^{m}(\hat{\mathbf k_i})^*}\\
-    \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}) \\
-    = \sbr{\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)Y_{l_i}^{m_i}(\hat{\mathbf k_i})^*}} (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
-\end{gather*}
-We can write the 3D dirac delta function in terms of spherical harmonics as
-\begin{equation*}
-    \delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3) = 8\int\d^3\mathbf x\prod_{i=1,2,3}\br{ \sum_{l_jm_j} i^{l_j}j_{l_j}(k_ix)Y_{l_j}^{m_j}(\hat{\mathbf k_i}) Y_{l_j}^{m_j}(\hat{\mathbf x})^* }.
-\end{equation*}
-This results in
-\begin{gather*}
-    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle= \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)Y_{l_i}^{m_i}(\hat{\mathbf k_i})^*} \\
-    \times (2\pi)^38\int\d^3\mathbf x\prod_{i}\br{ \sum_{lm} i^{l}j_{l}(k_ix)Y_{l}^{m}(\hat{\mathbf k_i}) Y_{l}^{m}(\hat{\mathbf x})^* }B^\Psi(\{k_i\}, \{\eta_0-\chi_i\})\\
-    = (2\pi)^38\int\d^3\mathbf x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3}}4\pi (-1)^{l_i}j_{l_i}(k_i\chi_i)j_{l_i}(k_ix) Y_{l_i}^{m_i}(\hat{\mathbf x})^*} B^\Psi(\{k_i\}, \{\eta_0-\chi_i\})
-\end{gather*}
-The angular part of the $\mathbf x$ integral can be evaluated using the identity
-\begin{gather*}
-    \int \mathrm{d}\Omega_{\hat{n}} \, Y_{l_1 m_1}^*(\hat{x}) Y_{l_2 m_2}^*(\hat{n}) Y_{l_3 m_3}^*(\hat{n}) = (-1)^{m_1 + m_2 + m_3} 
-    \int \mathrm{d}\Omega_{\hat{n}} \, Y_{l_1 -m_1}(\hat{n}) Y_{l_2 -m_2}(\hat{n})Y_{l_3 -m_3}(\hat{n}) \\
-    = (-1)^{m_1 + m_2 + m_3} 
-    \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix}
-    \begin{pmatrix} l_1 & l_2 & l_3 \\ -m_1 & -m_2 & -m_3 \end{pmatrix} \equiv A_{\mathbf{l}}^{\mathbf{m}},
-\end{gather*}
-giving
-\begin{gather*}
-    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle 
-    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3}}4\pi (-1)^{l_i}j_{l_i}(k_i\chi_i)j_{l_i}(k_ix)} \\ \times B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
-\end{gather*}
-Now applying the Limber approximation again:
-\begin{gather*}
-    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle 
-    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi_i^2}\delta(x-\chi_i)4\pi (-1)^{l_i}} \\ \times B^\Psi(\{l_i/\chi_i\}, \{\eta_0-\chi_i\}) \\
-    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int \chi^2\d \chi\prod_i\br{-2 W_i(\chi)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi^2}4\pi (-1)^{l_i}} B^\Psi(\{l_i/\chi\}, \eta_0-\chi).
-\end{gather*}
-Finally, we aim to rewrite the above in terms of the angular bispectrum of the lensing potential.
-
-The definition for the bispectrum of any set of randomly distributed spherical harmonic components $X^k_{lm}$ is CITE Hu 2000 
-\begin{equation*}
-    \langle (X_1)_{l_1m_1}(X_2)_{l_2m_2}(X_3)_{x_3m_3} \rangle = \begin{pmatrix}
-        l_1 & l_2 & l_3 \\ m_1 & m_2 & m_3
-    \end{pmatrix}
-    B_{l_1l_2l_3}^{X_1X_2X_3}.
-\end{equation*}
-Note the independence on $m_i$, this necesarilly follows from statistical isotropy. % TODO: add a short explanation here
-If $m_1+m_2+m_3\neq 0$, the associated Wigner-3j symbol vanishes and the bispectrum is set to zero.
-Also note that in this definition we have immediately generalized to include cross correlation between different fields $X_1$, $X_2$, $X_3$.
-This will be relevant later when we look at cross correlations between shear and convergence fields.
-
-Using the above definition and the symmetry property
-\begin{equation*}
-    \begin{pmatrix}
-        l_1 & l_2 & l_3\\
-        -m_1 & -m_2 & -m_3
-      \end{pmatrix}
-      =
-      (-1)^{l_1+l_2+l_3}
-      \begin{pmatrix}
-        l_1 & l_2 & l_3\\
-        m_1 & m_2 & m_3
-      \end{pmatrix}
-\end{equation*}
-we find
-\begin{gather*}
-    B^{\psi_1\psi_2\psi_3}_{l_1l_2l_3}
-    = (-1)^{l_1+l_2+l_3}
-    \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix}
-    (2\pi)^3 8 \\
-    \times \int \chi^2\d \chi\prod_i\br{-2 W_i(\chi, \chi_*)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi^2}4\pi (-1)^{l_i}} B^\Psi(\{l_i/\chi\}, \{\eta_0-\chi\}),
-\end{gather*}
-where we were able to drop the $(-1)^{m_1+m_2+m_3}$ factor due to the bispecrum vanishing if that sum doesn't equal $0$, as mentioned earlier.
-When all $m_i$ equal zero, the Wigner 3j-symbol gains a number of useful properties % TODO: discuss all of them
-In particular, it vanishes if $l_1+l_2+l_3$ is odd, meaning we can drop the $(-1)^{l_1+l_2+l_3}$ factor. Additionally cancelling common factors then gives
-\begin{align*}
-    B^{\psi_1\psi_2\psi_3}_{l_1l_2l_3}
-    =
-    -\sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} 8 \int \frac{\d \chi }{\chi^4} W_1(\chi)W_2(\chi)W_3(\chi) B^\Psi(\{l_i/\chi\}, \eta_0-\chi).
-\end{align*}
-
-% \begin{gather*}
-%     = \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{l_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3/2}}4\pi\sum_{l_i'm_i'}i^{l_i'}j_{l_i'}(k_i\chi_i)Y_{l_i'}^{m_i'}(\hat{\mathbf n_i})Y_{l_i'}^{m_i'}(\hat{\mathbf k_i})^*}\\
-%     \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3) B_\Psi(k_1, k_2, k_3) \\
-%     = \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3/2}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)Y_{l_i}^{m_i}(\hat{\mathbf k_i})^*}\\ 
-%     \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3) B_\Psi(k_1, k_2, k_3) \\
-%     = \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3/2}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)Y_{l_i}^{m_i}(\hat{\mathbf k_i})^*}\\ 
-%     \times (2\pi)^3\sbr{ \int \d^3 \mathbf{x} \prod_i2\sum_{l_i'm_i'}i^{l_i'}j_{l_i'}(k_ix) Y_{l_i'm_i'}(\hat{\mathbf k}_i) Y_{l_i'm_i'}(\hat{\mathbf x})^* } B_\Psi(k_1, k_2, k_3) \\
-%     = \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3/2}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)}\\ 
-%     \times (2\pi)^3\sbr{ \int \d^3 \mathbf{x} \prod_i2i^{l_i}j_{l_i}(k_ix) Y_{l_im_i}(\hat{\mathbf x})^* } B_\Psi(k_1, k_2, k_3) \\
-%     = A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3/2}}4\pi i^{l_i}(2i^{l_i})j_{l_i}(k_i\chi_i)j_{l_i}(k_ix)}
-%     (2\pi)^3  B_\Psi(k_1, k_2, k_3) \\
-%     = A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\frac{1}{(2\pi)^{3/2}}4\pi i^{l_i}(2i^{l_i})\frac{\pi}{2\chi_i^2}\delta(x-\chi_i)}
-%     (2\pi)^3  B_\Psi(l_1/\chi, l_2/\chi, l_3/\chi) \\
-%     = A_{\mathbf l}^{\mathbf m}\int_0^{\chi_*} x^2\d x\prod_i\br{-2 W(x, \chi_*)\frac{1}{(2\pi)^{3/2}}4\pi i^{l_i}(2i^{l_i})\frac{\pi}{2x^2}}
-%     (2\pi)^3  B_\Psi(l_1/\chi, l_2/\chi, l_3/\chi) \\
-%     = A_{\mathbf{l}}^{\mathbf{m}} (-1)^{1 + l_1 + l_2 + l_3} 8 (2\pi)^{9/2} \int_0^{\chi_*} \frac{[W(\chi, \chi_*)]^3}{\chi^4} \, B_\Psi\left( \frac{l_1}{\chi}, \frac{l_2}{\chi}, \frac{l_3}{\chi} \right ) \, \mathrm{d}\chi.
-% \end{gather*}
-
-% TODO: mention that the definition of the bispectrum is different form power spectrum of Psi in terms of normalization constant
-% TODO: explain 3d dirac delta in terms of spherical harmonics (eq 78 thesis Luna)
-% TODO: explain that we used that int d^2 Omega Y Y Y = A_vec l^vec m (eq 82 thesis Luna)
-% using that d^3 k = k^2 d k d^2 Omega_k
-% TODO: fix this
-\subsection{Gravitational potential spectra in terms of matter spectra}
-% TODO: fix references and citations 
-To convert equations ...
-and ...
-to forms that we can directly evaluate we will need to replace the spectra of the graviational potential with matter spectra.
-Specifically, we define the matter power spectrum in agreement with the CAMB package documentation CITE and the matter bispectrum in agreement with CITE FITTING FUNC PAPER such that we can approximate it in terms of the matter power spectrum using hyperextended perturbation theory.
-The \textbf{density contrast} is defined as
-\begin{equation}
-    \delta(\mathbf x) := \frac{\rho(\mathbf x)-\bar \rho}{\bar\rho},
-\end{equation}
-and the matter spectra are defined in terms of the fourier transformed density contrast $\delta(\mathbf k)$ as 
-\begin{gather*}
-    \langle \delta(\mathbf k, \eta) \delta(\mathbf k', \eta)^* \rangle = (2\pi)^3\delta(\mathbf k - \mathbf k')P^\delta(\mathbf k, \eta),\\
-    \langle \delta(\mathbf k_1, \eta) \delta(\mathbf k_2, \eta) \delta(\mathbf k_3, \eta) \rangle = (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\delta(k_1, k_2, k_3, \eta).
-\end{gather*}
-The mean matter density of the universe, \( \bar{\rho} \), can be expressed in terms of the matter density parameter \( \Omega_m \) and the critical density \( \rho_c \) as 
-$$
-\bar{\rho}(\eta) = \frac{3 \Omega_m H_0^2}{8 \pi G}\frac{1}{a(\eta)^3},
-$$
-where $a(\eta)$ is the only time dependent factor on the right hand side.
-In the case of a flat universe ($K=0$) we can thus rewrite the Poisson equation REF (6.41 in modern cosmology) as
-\begin{equation}
-    -\nabla^2\Psi = 4\pi G a^2\br{\frac{3\Omega_m H_0^2}{8\pi G}\frac{1}{a^3}}\delta = \frac{3\Omega_m H_0^2}{2} \frac{1}{a}\delta \implies \Psi(k, \eta) = \frac{3\Omega_m H_0^2}{2}\frac{1}{a}\frac{\delta(k, \eta)}{k^2},
-\end{equation}
-where $\Psi(k,\eta)$ and $\delta(k,\eta)$ are functions in Fourier space. For the power- and bispectra we find
-\begin{gather*}
-    \langle \Psi(\mathbf k, \eta)\Psi^*(\mathbf k',\eta)\rangle = \frac{2\pi^2}{k^3}P^\Psi(k, \eta)\delta(\mathbf k - \mathbf k') \implies P^\Psi(k,\eta) = \frac{1}{k}(9\Omega_m^2H_0^4\pi) \frac{1}{a^2}P^\delta(k,\eta),\\
-    \langle \Psi(\mathbf k_1, \eta)\Psi(\mathbf k_2,\eta)\Psi(\mathbf k_3,\eta)\rangle = (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(k_1,k_2,k_3, \eta)\\ 
-    \implies B^\Psi(k_1,k_2,k_3, \eta) = \frac{1}{k_1^2k_2^2k_3^2}\br{\frac{3\Omega_m H_0^2}{2}}^3 \frac{1}{a^3}B^\delta(\{k_i\}, \eta).
-\end{gather*}
-We conclude this section by using the above equations to rewrite the lensing potential spectra (equations REF and REF) in terms of matter spectra 
+\section{Theory}
+Radiation from cosmological objects is distorted due to gravitational lensing.
+In the vast majority of cases this lensing is a relatively weak effect and is thus referred to as weak lensing.
+Weak lensing is quantified through the deflection field $\mathbf d(\hat {\mathbf n})$ which equals the difference between the observed angle of a point on the sky and the true (unlensed) angle.
+This field is the gradient of the lensing potential $\psi(\hat {\mathbf n})$ given as a weighted integral over the mass distribution along the line of sight between earth and the observed point.
+
+In the case of CMB surveys, lensing alters the statistical properties of the temperature and polarization fields and can thus be calculated by comparing to the expected unlensed signal, see e.g.
+\cite{cmblensingestimator}.
+In galaxy surveys, lensing alters the ellipiticities of observed galaxies.
+If a large enough number of galaxies are observed, this effect can be seperated from the intrinsic ellipiticities of the galaxies and thus the lensing potential can be estimated.
+
+To constrain cosmological parameters we can then look at the lensing potential of the CMB, $\psi_\CMB$, and of galaxy surveys, $\psi_\gal$.
+These are directly related to the matter power and bispectra as
 \begin{empheq}[box=\fbox]{align*}
     P^{\psi_X\psi_Y}_l
-    & = (-1)^{l}\frac{9}{l^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)P^\delta(l/\chi,\eta_0-\chi),\\
-    B^{\psi_X\psi_Y\psi_Z}_{l_1l_2l_3} &= -\sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{l_1^2l_2^2l_3^2}\Omega_m^3H_0^6\\
-    & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{l_i/\chi\}, \eta_0-\chi).
+    & = \frac{9}{l^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)P^\delta(l/\chi,\eta_0-\chi),\\
+    B^{\psi_X\psi_Y\psi_Z}_{l_1l_2l_3} &= \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{l_1^2l_2^2l_3^2}\Omega_m^3H_0^6\\
+    & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{l_i/\chi\}, \eta_0-\chi),
 \end{empheq}
-
-% \begin{gather*}
-%     \langle \psi_{l_1m_1} \psi_{l_2m_2} \psi_{l_3m_3} \rangle = A_{\mathbf{l}}^{\mathbf{m}} (-1)^{l_1 + l_2 + l_3} 8 (2\pi)^{9/2} \int_0^{\chi_*} W(\chi, \chi_*)^3 \frac{\chi^2}{l_1^2l_2^2l_3^2} \br{\frac{3\Omega_m H^2}{2}}^3|_{\eta=\eta_0-\chi}\\
-%     \quad \times B^\delta\br{ \frac{l_1}{\chi}, \frac{l_2}{\chi}, \frac{l_3}{\chi}; \eta_0-\chi } \mathrm{d}\chi, \\
-%     C_l^\psi = \frac{72\pi^3}{l^4}\int_0^{\chi_*}\d \chi \chi^2(\Omega_m^2H^4)|_{\eta=\eta_0-\chi} C^\delta_{l/\chi}(\eta_0-\chi) W(\chi)^2.
-% \end{gather*}
-
-\subsection{Convergence and shear multipole moments}
-\paragraph{Convergence}
-% TODO: make notation of nabla nicer/more consistent
-
-Taking the harmonic decomposition of the convergence,
-\begin{equation}
-    \kappa(\hat{\mathbf n}) = \kappa_{lm}Y_l^m(\hat{\mathbf n}),
-\end{equation}
-you can relate it to the multipole moments of the lensing potential through
-\begin{equation}
-    \kappa_{lm} = \frac{1}{2}(\nabla^2\psi)_{lm} = \frac{l(l+1)}{2}\psi_{lm}
-\end{equation}
-because the spherical harmonics are eigenfunctions of the Laplacian operator.
-
-\paragraph{Shear}
-On the other hand, the shear is given by
-\begin{equation}
-    \gamma = \gamma_1 + i\gamma_2 = \frac{1}{2}\eth_1(\eth_0\psi)
-\end{equation}
-where the spin raising operator, $\eth_s$ acts on a spin $s$ function defined on $S^2$ to create a spin $s+1$ function.
-The above equality is proven explicitely in appendix ...
-%TODO: write and cite appendix proving equivalence
-$\eth_s$ can be written in $(\theta, \phi)$ coordinates as
-\begin{equation}
-    \eth_s = -\sin^s\theta(\partial_\theta + \frac{i}{\sin\theta})\frac{1}{\sin^s\theta}.
-\end{equation}
-It is worth explaining that in this context a spin $s$ function referes to a function $_sf(\theta, \phi)$ that transforms under any rotation of coordinates by picking up a phase factor $e^{is\alpha}$, with $\alpha$ the angle of the rotation, i.e.
-\begin{equation}
-    f'(\theta', \phi') = e^{is\alpha}f(\theta, \phi).
-\end{equation}
-Shear is actually a spin 2 function, which can immediately be seen by observing that rotating a galaxy image stretched and squeezed through weak lensing by $180$ degrees gives the same stretching and squeezing, i.e.
-the same shear.
-It is thus expected that the shear is proportional to $\eth^2\psi$.
-
-The spin weighted spherical harmonics, $_sY_l^m$ are eigenfunctions of the raising/lowering operator, in particular
-\begin{equation}
-    \eth_sY_l^m = \sqrt{(l-s)(l+s+1)} _{s+1}Y_l^m.
-\end{equation}
-Decomposing the shear as
-\begin{equation}
-    \gamma(\theta, \phi) = \gamma_{lm} {}_sY_l^m(\theta,\phi),
-\end{equation}
-we thus find
-\begin{equation}
-    \gamma_{lm} = \frac{1}{2}\sqrt{(l-1)(l)(l+1)(l+2)} \psi_{lm}
-\end{equation}
+with $X, Y, Z \in \{\CMB, \gal\}$.
+$\Omega_m$ is the present day matter density parameter, $H_0$ is the present day Hubble constant, $a(\eta)$ is the scale factor, $\eta_0$ is the conformal time today, $\chi$ is the comoving radial distance, $\chi_*$ the distance to surface of last scattering, $P^\delta(k, \eta)$ the matter power spectrum, $B^\delta(k_1, k_2, k_3, \eta)$ the matter bispectrum, $W_X(\chi)$ the window function, and
+$$
+\begin{pmatrix}
+    l_1&l_2&l_3 \\ m_1 & m_2 & m_3
+\end{pmatrix}
+$$ 
+the Wigner 3-j symbol. The only factor that differs between CMB and galaxy lensing is the window function, defined as
+\begin{equation*}
+    W_X(\chi) = \int_\chi^{\infty}\d\chi' p_X(\chi')\frac{\chi'-\chi}{\chi'\chi},
+\end{equation*}
+with $p_X$ the radius distribution of the radiation source. For the CMB we take $p(\chi) = \delta(\chi - \chi_*)$. Notice that we can also look at cross correlation between CMB and galaxy lensing.
+
+For a derivation of the lensing potential and its power- and bispectra, see appendices \ref{sec:weaklensing} and \ref{sec:weaklensstats}.
 
@@
 \appendix
+
+\section{Weak Lensing}\label{sec:weaklensing}
+\subsection{Approximations Made / Regime}
+\subsection{Perturbed Photon Paths}
+Work in \textbf{conformal newtonian gauge}.
+In this section we'll work in units where $c=1$.
+Denoting conformal time and conformal radial distance by $\eta$ and $\chi$, respectively, the perturbed line element is given by
+\begin{equation}
+    \d s^2 = a^2(\eta)((1 + 2\Psi_N) \d \eta^2 - (1 + 2\Phi_N) \gamma_{ij}\d x^i \d x^j) 
+\end{equation}
+
+where $\gamma_{ij}$ is the unperturbed line element
+\begin{equation}
+    \gamma_{ij} = \d x^i \d x^j = \d \chi^2 + f^2_K(\chi)(\d \theta^2 + \sin ^2 \theta \d \phi ^2).
+\end{equation}
+From here of on work in first order in the scalar potentials $\Phi_N$ and $\Psi_N$. For lensing we consider null-geodesics so $\d s^2 = 0$ and we can rewrite the perturbed line element as
+\begin{equation}
+    \d \hat s^2 = (1 + 4\Psi)\d\eta^2 - \gamma_{ij} \d x^i \d x^j
+\end{equation}
+with $\Psi$ the \textbf{Weyl Potential} given by $\Psi := (\Psi_N - \Phi_N)/2$. A null geodesic $x^\mu(\hat\lambda)$ in terms of its affine parameter ($\hat \lambda$) satisfies the geodesic equation
+\begin{equation}
+    \frac{\d^2 x^\mu}{\d \hat\lambda^2} + \Gamma^\mu_{\nu\rho} \frac{\d x^\nu}{\d \hat\lambda}\frac{\d x^\rho}{\d\hat\lambda} = 0,
+\end{equation}
+with $\hat g_{\mu\nu}(\d x^\mu/\d \hat\lambda)(\d x^\nu / \d \hat \lambda) = 0$.
+
+In the unperturbed there are incoming radial solutions of the form $\chi = \eta_0 - \eta$. We can express $\hat\lambda$ in terms of $\eta$ using The 0-component of the geodesic equation,
+\begin{equation}
+    \frac{\d^2\eta}{\d\hat\lambda^2} + 2(\frac{\d\eta}{\d\hat\lambda})^2\frac{\d\Psi}{\d\eta}+2\frac{\d\eta}{\d\hat\lambda}\frac{\d x^i}{\d\hat\lambda}\frac{\partial\Psi}{\partial x^i} = 0,
+\end{equation}
+where the derivative $\d\Psi/\d\eta = \partial_\eta\Psi+(\d x^i/\d\eta)\partial_i\Psi$ is along the perturbed ray. This allows us to rewrite the other (spatial) geodesic equations as
+\begin{equation}
+    \frac{\d^2x^i}{\d\eta^2} - 2\frac{\d x^i}{\d\eta}(\frac{\d\Psi}{\d\eta} + \frac{\d x^j}{\d\eta}\frac{\partial\Psi}{\partial x^j})+2\gamma^{ij}\frac{\partial\Psi}{\partial x^j}+\bar\Gamma^i_{jk}\frac{\d x^j}{\d\eta}\frac{\d x^k}{\d \eta} = 0,
+\end{equation}
+where $\bar\Gamma^i_{jk}$ are the connection coefficients of the unperturbed three geometry $\gamma_{ij}$.
+Without loss of generality we consider an observer located at the origin of the spatial coordinates, meaning we are interested in rays that end at $x^i = 0$.
+In that case $\d \chi/\d \eta = -1 + O(\Psi)$, $\d \theta/\d\eta = O(\Psi)$ and $\d \phi/\d\eta = O(\Psi)$ (with $\theta$ and $\phi$ angular spatial coordinates).
+For the case of such rays we can evaluate the connection coefficients and rewrite the spatial geodesic equations as
+\begin{gather}
+    \frac{\partial^2\chi}{\d\eta^2}+2\frac{\d\Psi}{\d\eta}=0,\\
+    \frac{\d^2\theta}{\d\eta^2}-2\frac{\d\ln f_K(\chi)}{\d\chi}\frac{\d\theta}{\d\eta}+\frac{2}{f^2_K(\chi)}\frac{\partial\Psi}{\partial\theta}=0,\\
+    \frac{\d^2\phi}{\d\eta^2} - 2\frac{\d \ln f_K(\chi)}{\d\chi}\frac{\d\phi}{\d\eta}+\frac{2}{f^2_K(\chi)}\frac{1}{\sin^2(\theta)}\frac{\partial\Psi}{\partial\phi}=0.
+\end{gather}
+Integrating twice and using the null condition we can rewrite the the first equation as
+\begin{equation}
+    \chi = \eta_0 - \eta - 2\int_{\eta_0}^\eta\Psi\d\eta',
+\end{equation}
+where the integral is along the ray.
+Since we are working at first order in $\Psi$ we can evaluate the integral along the unperturbed path, this is known as the \textbf{Born approximation}.\textit{Next follows a short discussion of why perturbations in observation time can be ignored, include?} Doing the same for the other two equations gives
+% TODO: make n a proper vector (bold)
+\begin{gather}
+    \theta(\eta_0 - \chi_*)=\theta_0 - \int_0^{\chi_*}\d\chi\frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}2\frac{\partial}{\partial \theta}\Psi(\chi\hat n, \eta_0 - \chi),\\
+    \phi(\eta_0 - \chi_*)=\theta_0 - \int_0^{\chi_*}\d\chi\frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}\frac{2}{\sin^2\theta}\frac{\partial}{\partial \phi}\Psi(\chi\hat n, \eta_0 - \chi),
+\end{gather}
+where $\theta_0$ and $\phi_0$ label the line of sight $\hat n$.
+The displacement vector $\mathbf{\alpha}$ indicates the infinitesimal displacement of the lensed rays from their unpetrubed observed angles and is related to $\theta$ and $\phi$ through
+$$
+\alpha_\theta = \theta - \theta_0, \quad \alpha_\phi = \sin^2\theta(\phi - \phi_0).
+$$
+This allows us to summarize the above results as
+\begin{equation}
+    \mathbf \alpha = -2 \int_0^{\chi_*}\d \chi \frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}\nabla_{\hat{\mathbf n}}\Psi(\chi\hat{\mathbf n};\eta_0 - \chi).
+\end{equation}
+When working in the weak field limit (i.e. small spacetime curvature) the $00$ component of Einsteins equations give us a poisson equation \cite{dodelson2020modern},
+\begin{equation}
+    -\nabla^2\Psi(\textbf{x}, \eta) =  4\pi G a(\eta)^2\delta\rho(\mathbf x, \eta)
+\end{equation}
+where $K$ is $0$, $1$, or $-1$ depending on the curvature of the universe, $\Delta = \partial_x^2 + \partial_y^2 + \partial_z^2$ is the 3 dimensional laplace operator, $\delta\bar\rho$ is the comoving total energy perturbation (evaluated in the rest-frame of the total energy).
+%TODO: proper definition for anisotropic stress and explain why we can ignore it
+
+\subsection{Convergence, shear}
+Quantities we are interested in measuring depend on how the lensing vector changes with the observation angle. \textbf{Magnification matrix} is defined by
+\begin{equation}
+    A_{ij} := \delta_{ij} + \frac{\partial}{\partial\theta_i}\alpha_i(\hat{\mathbf n}).
+\end{equation}
+This matrix can then be decomposed in the following form, which immediately gives us definitions for the \textbf{convergence}, $\kappa$, \textbf{shear}, $\gamma_1$ and $\gamma_2$, and \textbf{rotation}, $\omega$:
+\begin{equation}
+    A_ij(\hat{\mathbf n}) = \begin{pmatrix}
+        1 - \kappa - \gamma_1 & -\gamma_2 + \omega \\
+        -\gamma_2 - \omega & 1 - \kappa + \gamma_1
+    \end{pmatrix}.
+\end{equation}
+At 1st order $A$ is a symetric matrix by definition and so $\omega$ has to be 0, we will ignore it from here on out.
+Intuitively, A tells you how a small patch in the sky transforms due to weak lensing, i.e.
+if you have surface brightness $I$ then
+% TODO: make delta thick
+\begin{equation}
+    I_{\text{emitted}}(\hat{\mathbf{n}} + \mathbf{\delta}) = I_{\text{observed}}((\hat{\mathbf{n}} + \mathbf{\delta}) + \mathbf\alpha(\hat{\mathbf{n}} + \mathbf{\delta})) \approx I_{\text{observed}}((\hat{\mathbf{n}} + \mathbf{\alpha} + A_{ij}\delta_{j}).
+\end{equation}
+given that $|A_{ij}| = (1 - \kappa)^2 + \omega^2 - |\gamma|^2 = 1 - 2\kappa + O(\kappa^2, \gamma^2, \omega^2)$, we can interpret $\kappa$ as telling us about the overal magnification of the source.
+The $\gamma_i$ represent the area preserving distortion, i.e.
+stretching and squeezing in a specific direction.
+
+\subsection{lensing potential}
+Taking the derivative out of the integral in equation REF leads us to defining the \textbf{lensing potential}, $\psi$,
+\begin{equation}
+    \mathbf \alpha = \nabla_{\hat{\mathbf n}}\psi, \quad \psi(\hat{\mathbf n}) = -2 \int_0^{\chi_*} \d \chi \frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}\Psi(\chi\hat{\mathbf n}; \eta_0 - \chi).
+\end{equation}
+
+We can relate $\kappa$ and $\gamma$ directly to the lensing potential as
+% TODO: should these subscripts be theta and phi?
+\begin{equation}
+    \kappa = \frac{1}{2}\nabla^2\psi, \quad \gamma_1 = \frac{1}{2}(\partial_\theta^2 - \partial_\phi^2)\psi, \quad \gamma_2 = \partial_\theta\partial_\phi \psi.
+\end{equation}
+
+\subsection{Regarding window functions}
+The window function for light emitted from a source at fixed radius $\chi_*$ is $(\chi-\chi^*)/(\chi_*\chi)$, if the source is instead distributed over the radius as $p(\chi)$, with $p(\chi)$ normalized to integrate to 1, we generalize the window function as
+\begin{equation*}
+    W(\chi) = \int_\chi^{\infty}\d\chi' p(\chi')\frac{\chi'-\chi}{\chi'\chi}.
+\end{equation*}
+The integration limit is up to surface of last scattering because further than that no source can contribute to the observed image.
+In the case of an image of the CMB we can take $p(\chi')=\delta(\chi'-\chi_*)$, in which case the window function reduces back to $(\chi_*-\chi)/(\chi_*\chi)$.
+In the case of different sources, we will differentiate between their respective window functions with a subscript.
+In this case the factor that changes is the distribution function.
+The sources considered in this paper are only the CMB and selections of galaxies from galaxy surveys.
+
+\section{Weak Lensing Statistics}\label{sec:weaklensstats}
+\subsection{Lensing Potential Powerspectrum}
+Consider the lensing potential, $\psi$, then decomposing into spherical harmonics gives
+\begin{equation}
+    \psi(\hat{\mathbf n}) = \sum_{lm} \psi_{lm}Y_{lm}(\hat{\mathbf n}).
+\end{equation}
+On the other hand, consider the decomposition of $\Psi$ in fourier modes with the Fourier convention $\Psi(\mathbf x, \eta)=\int\frac{\d^3 \mathbf k}{(2\pi)^{3}}\Psi(\mathbf k, \eta)e^{i\mathbf k \cdot\mathbf x}$,
+\begin{equation}
+    \psi(\hat{\mathbf n}) = -2 \int_0^{\chi_*}\d \chi W(\chi) \int\frac{\d^3 \mathbf k}{(2\pi)^{3}}\Psi(\mathbf k, \eta_0 - \chi)e^{i\mathbf k \cdot\hat{\mathbf n}\chi}.
+\end{equation}
+We can than relate the multipole modes of $\psi$ to the fourier modes of $\Psi$ through % TODO: check sign convention here
+\begin{gather}
+    \psi_{lm} = \langle Y_l^m | \psi \rangle = \int\d^2\hat{\mathbf n} Y_l^m(\hat{\mathbf n})^* \psi(\hat{\mathbf n})\\ 
+    = -2\int\d^2\hat{\mathbf n} Y_l^m(\hat{\mathbf n})^* \int_0^{\chi_*}\d \chi W(\chi) \int\frac{\d^3 \mathbf k}{(2\pi)^{3}}\Psi(\mathbf k, \eta_0 - \chi)e^{i\mathbf k \cdot\hat{\mathbf n}\chi}
+\end{gather}
+Now define the power spectrum as
+\begin{equation}
+    \langle \Psi(\mathbf k, \eta)\Psi(\mathbf k',\eta')\rangle = \frac{2\pi^2}{k^3}P_\Psi(k, \eta, \eta')\delta(\mathbf k + \mathbf k'),
+\end{equation}
+with $\eta$ denoting the conformal time. This gives 
+%TODO: define the window function
+\begin{equation}
+    \langle \psi(\hat{\mathbf n})\psi(\hat{\mathbf n}') \rangle = 4\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\int \frac{\d^3\mathbf k}{(2\pi)^6}\frac{2\pi^2}{k^3}P_\psi(k,\eta_0 - \chi, \eta_0-\chi')e^{i\mathbf k\cdot\hat {\mathbf n}\chi}e^{-i\mathbf k\cdot\hat {\mathbf n}'\chi'},
+\end{equation}
+where we used that $\eta = \eta_0 - \chi$ along the unperturbed photon path (Born approximation), with $\eta_0$ the time at which the light ray hits earth. 
+We can use the result
+\begin{equation}
+    e^{i\mathbf k \cdot \hat {\mathbf n}\chi}=4\pi\sum_{lm}i^lj_l(k\chi )Y_l^m(\hat{\mathbf n})^*Y_l^m(\hat{\mathbf k}) = 4\pi\sum_{lm}i^lj_l(k\chi )Y_l^m(\hat{\mathbf n})Y_l^m(\hat{\mathbf k})^*,
+\end{equation}
+where $j_l$ is the spherical Bessel function, to rewrite the above equation.
+Using both versions of the identity above, we immediately get a factor $Y_l^m(\hat{\mathbf k})Y_{l'}^{m'}(\hat{\mathbf k})^*$ in our integral.
+We can factor the differential element of $\d^3\mathbf k$ into a radial and angular part as $k^2\d k\d^2\Omega_k$, with $\Omega_k$ the solid angle, to apply the orthonormality condition of the spherical harmonics.
+Note that we take the same sequence of steps a number of times in other parts of the derivations of the lensing spectra.
+We thus obtain
+\begin{gather}
+    \langle \psi(\hat{\mathbf n})\psi(\hat{\mathbf n}') \rangle = 4(4\pi)^2\sum_{ll'mm'}i^{l-l'}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\\\times\int \frac{k^2\d k}{(2\pi)^6}\frac{2\pi^2}{k^3}j_l(k\chi)j_{l'}(k\chi')P_\psi(k,\eta_0 - \chi, \eta_0-\chi')Y_{lm}(\hat{\mathbf n})Y_{l'm'}(\hat{\mathbf n}')^*\delta_{ll'}\delta_{mm'}.
+\label{eq:lenspotcorrexpanded}
+\end{gather}
+The angular power spectrum is defined similarly to the power spectrum, i.e.
+\begin{equation}
+    \langle \psi_{lm}\psi_{l'm'}^* \rangle = \delta_{ll'}\delta_{mm'}C_l^\psi.
+\end{equation}
+Note that the correlation is independent of $m$ and $m'$. %TODO: explain how this relates to statistical isotropy
+We can thus read from equation \ref{eq:lenspotcorrexpanded} that
+\begin{gather}
+    C_l^\psi = 4(4\pi)^2\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\int \frac{k^2\d k}{(2\pi)^6}\frac{2\pi^2}{k^3}j_l(k\chi)j_{l}(k\chi')P_\psi(k,\eta_0 - \chi, \eta_0-\chi'),
+\end{gather}
+which can be simplified to
+\begin{gather}
+    C_l^\psi = \frac{2}{\pi^2}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\int k^2\d kj_l(k\chi)j_{l}(k\chi')\frac{P_\psi(k,\eta_0 - \chi, \eta_0-\chi')}{k^3}.
+\end{gather}
+
+To further evaluate the integral we will make the standard \textbf{Limber approximation}.
+The Bessel functions peak sharply at $l=x$\footnote{Some sources use $x\approx l+1/2$ instead, which is slightly more accurate for larger scales (low $l$) and slightly less accurate for smaller scales.}, with the peak being increasingly sharp for higher $l$.
+Similarly to $\delta(x-x_0)f(x)=\delta(x-x_0)f(x_0)$, we thus take $j_l(k\chi)f(k)\approx j_l(k\chi)f(l/\chi)$.
+The Bessel functions satisfy an orthogonality condtion,
+\begin{equation}
+    \int k^2\d k j_l(k\chi)j_l(k\chi') = \frac{\pi}{2\chi^2}\delta(\chi-\chi').
+\end{equation}
+In combination with the Limber approximation we thus find
+\begin{equation}
+    \int k^2\d k j_l(k\chi)j_l(k\chi')f(k) \approx \frac{\pi}{2\chi^2}\delta(\chi-\chi')f(l/\chi).
+\end{equation}
+%TODO: how high? quantify limber approx, perhaps using convergence x LSS paper
+It allows us to write the Limber-approximate angular spectrum as
+\begin{gather}
+    C^\psi(l) = \frac{2}{\pi^2}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi') \frac{\pi}{2\chi^2}\delta(\chi-\chi')\frac{\chi^3}{l^3}P_\psi(l/\chi,\eta_0 - \chi, \eta_0-\chi')\\
+    = \frac{1}{l^3\pi}\int_0^{\chi_*}\chi\d \chi W(\chi)^2 P_\psi(l/\chi,\eta_0 - \chi, \eta_0-\chi).
+\end{gather}
+% The power spectrum of the potential can be related to the power spectrum of the matter density using the Poisson equation %TODO: create ref
+% For a flat universe in atter or dark energy domination we thus find the relation
+% \begin{equation}
+%     P_\Psi(k;\eta, \eta)=\frac{9\omega_m^2(\eta)H^4(\eta)}{8\pi^2} \frac{P_m(k, \eta)}{k},
+% \end{equation}
+% where we use the shorthand $P_m(k, \eta):=P_m(k,\eta,\eta)$. The lensing power spectrum is thus finally given in terms of the matter power spectrum as
+% \begin{equation}
+%     C_l^\psi=\underbrace{9\omega_m^2(\eta)H^4(\eta)}_{:=C(\eta)}\frac{1}{l^4}\int_0^{\chi_*}\d\chi\chi^2 W(\chi)^2P_m(l/\chi, \eta_0-\chi).
+% \end{equation}
+% We will henceforth denote the $\eta$ dependent factor outside of the integral as $C(\eta)$.
+\subsection{Lensing potential bispectrum}
+The derivation of the bispectrum proceeds similarly to that of the power spectrum.
+We aim to compute the bispectrum of the lensing potential fields of 3 (possibly distinct sources), $\psi_1$, $\psi_2$, $\psi_3$.
+\begin{gather*}
+    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle = 
+    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{l_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}e^{i\mathbf k_i \cdot\hat{\mathbf n_i}\chi_i}} \\
+    \times \langle \prod_i\Psi(\mathbf k_i, \eta_0 - \chi_i) \rangle.
+\end{gather*}
+Defining the bispectrum of the graviational potential as
+\begin{equation*}
+    \langle \prod_{i=1,2,3}\Psi(\mathbf k_i, \eta_0 - \chi_i) \rangle = (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}),
+\end{equation*}
+we rewrite the lensing potential bispecrum as
+\begin{gather*}
+    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle = 
+    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{l_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}e^{i\mathbf k_i \cdot\hat{\mathbf n_i}\chi_i}}\\ \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
+\end{gather*}
+Now using the identity REF for the complex exponential:
+\begin{gather*}
+    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle \\ = 
+    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{l_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi\sum_{lm}i^{l}j_{l}(k_i\chi_i)Y_{l}^{m}(\hat{\mathbf n_i})Y_{l}^{m}(\hat{\mathbf k_i})^*}\\
+    \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}) \\
+    = \sbr{\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)Y_{l_i}^{m_i}(\hat{\mathbf k_i})^*}} (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
+\end{gather*}
+We can write the 3D dirac delta function in terms of spherical harmonics as
+\begin{equation*}
+    \delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3) = 8\int\d^3\mathbf x\prod_{i=1,2,3}\br{ \sum_{l_jm_j} i^{l_j}j_{l_j}(k_ix)Y_{l_j}^{m_j}(\hat{\mathbf k_i}) Y_{l_j}^{m_j}(\hat{\mathbf x})^* }.
+\end{equation*}
+This results in
+\begin{gather*}
+    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle= \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)Y_{l_i}^{m_i}(\hat{\mathbf k_i})^*} \\
+    \times (2\pi)^38\int\d^3\mathbf x\prod_{i}\br{ \sum_{lm} i^{l}j_{l}(k_ix)Y_{l}^{m}(\hat{\mathbf k_i}) Y_{l}^{m}(\hat{\mathbf x})^* }B^\Psi(\{k_i\}, \{\eta_0-\chi_i\})\\
+    = (2\pi)^38\int\d^3\mathbf x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3}}4\pi (-1)^{l_i}j_{l_i}(k_i\chi_i)j_{l_i}(k_ix) Y_{l_i}^{m_i}(\hat{\mathbf x})^*} B^\Psi(\{k_i\}, \{\eta_0-\chi_i\})
+\end{gather*}
+The angular part of the $\mathbf x$ integral can be evaluated using the identity
+\begin{gather*}
+    \int \mathrm{d}\Omega_{\hat{n}} \, Y_{l_1 m_1}^*(\hat{x}) Y_{l_2 m_2}^*(\hat{n}) Y_{l_3 m_3}^*(\hat{n}) = (-1)^{m_1 + m_2 + m_3} 
+    \int \mathrm{d}\Omega_{\hat{n}} \, Y_{l_1 -m_1}(\hat{n}) Y_{l_2 -m_2}(\hat{n})Y_{l_3 -m_3}(\hat{n}) \\
+    = (-1)^{m_1 + m_2 + m_3} 
+    \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix}
+    \begin{pmatrix} l_1 & l_2 & l_3 \\ -m_1 & -m_2 & -m_3 \end{pmatrix} \equiv A_{\mathbf{l}}^{\mathbf{m}},
+\end{gather*}
+giving
+\begin{gather*}
+    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle 
+    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3}}4\pi (-1)^{l_i}j_{l_i}(k_i\chi_i)j_{l_i}(k_ix)} \\ \times B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
+\end{gather*}
+Now applying the Limber approximation again:
+\begin{gather*}
+    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle 
+    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi_i^2}\delta(x-\chi_i)4\pi (-1)^{l_i}} \\ \times B^\Psi(\{l_i/\chi_i\}, \{\eta_0-\chi_i\}) \\
+    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int \chi^2\d \chi\prod_i\br{-2 W_i(\chi)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi^2}4\pi (-1)^{l_i}} B^\Psi(\{l_i/\chi\}, \eta_0-\chi).
+\end{gather*}
+Finally, we aim to rewrite the above in terms of the angular bispectrum of the lensing potential.
+
+The definition for the bispectrum of any set of randomly distributed spherical harmonic components $X^k_{lm}$ is CITE Hu 2000 
+\begin{equation*}
+    \langle (X_1)_{l_1m_1}(X_2)_{l_2m_2}(X_3)_{x_3m_3} \rangle = \begin{pmatrix}
+        l_1 & l_2 & l_3 \\ m_1 & m_2 & m_3
+    \end{pmatrix}
+    B_{l_1l_2l_3}^{X_1X_2X_3}.
+\end{equation*}
+Note the independence on $m_i$, this necesarilly follows from statistical isotropy. % TODO: add a short explanation here
+If $m_1+m_2+m_3\neq 0$, the associated Wigner-3j symbol vanishes and the bispectrum is set to zero.
+Also note that in this definition we have immediately generalized to include cross correlation between different fields $X_1$, $X_2$, $X_3$.
+This will be relevant later when we look at cross correlations between shear and convergence fields.
+
+Using the above definition and the symmetry property
+\begin{equation*}
+    \begin{pmatrix}
+        l_1 & l_2 & l_3\\
+        -m_1 & -m_2 & -m_3
+      \end{pmatrix}
+      =
+      (-1)^{l_1+l_2+l_3}
+      \begin{pmatrix}
+        l_1 & l_2 & l_3\\
+        m_1 & m_2 & m_3
+      \end{pmatrix}
+\end{equation*}
+we find
+\begin{gather*}
+    B^{\psi_1\psi_2\psi_3}_{l_1l_2l_3}
+    = (-1)^{l_1+l_2+l_3}
+    \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix}
+    (2\pi)^3 8 \\
+    \times \int \chi^2\d \chi\prod_i\br{-2 W_i(\chi, \chi_*)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi^2}4\pi (-1)^{l_i}} B^\Psi(\{l_i/\chi\}, \{\eta_0-\chi\}),
+\end{gather*}
+where we were able to drop the $(-1)^{m_1+m_2+m_3}$ factor due to the bispecrum vanishing if that sum doesn't equal $0$, as mentioned earlier.
+When all $m_i$ equal zero, the Wigner 3j-symbol gains a number of useful properties % TODO: discuss all of them
+In particular, it vanishes if $l_1+l_2+l_3$ is odd, meaning we can drop the $(-1)^{l_1+l_2+l_3}$ factor. Additionally cancelling common factors then gives
+\begin{align*}
+    B^{\psi_1\psi_2\psi_3}_{l_1l_2l_3}
+    =
+    -\sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} 8 \int \frac{\d \chi }{\chi^4} W_1(\chi)W_2(\chi)W_3(\chi) B^\Psi(\{l_i/\chi\}, \eta_0-\chi).
+\end{align*}
+
+% \begin{gather*}
+%     = \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{l_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3/2}}4\pi\sum_{l_i'm_i'}i^{l_i'}j_{l_i'}(k_i\chi_i)Y_{l_i'}^{m_i'}(\hat{\mathbf n_i})Y_{l_i'}^{m_i'}(\hat{\mathbf k_i})^*}\\
+%     \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3) B_\Psi(k_1, k_2, k_3) \\
+%     = \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3/2}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)Y_{l_i}^{m_i}(\hat{\mathbf k_i})^*}\\ 
+%     \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3) B_\Psi(k_1, k_2, k_3) \\
+%     = \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3/2}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)Y_{l_i}^{m_i}(\hat{\mathbf k_i})^*}\\ 
+%     \times (2\pi)^3\sbr{ \int \d^3 \mathbf{x} \prod_i2\sum_{l_i'm_i'}i^{l_i'}j_{l_i'}(k_ix) Y_{l_i'm_i'}(\hat{\mathbf k}_i) Y_{l_i'm_i'}(\hat{\mathbf x})^* } B_\Psi(k_1, k_2, k_3) \\
+%     = \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3/2}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)}\\ 
+%     \times (2\pi)^3\sbr{ \int \d^3 \mathbf{x} \prod_i2i^{l_i}j_{l_i}(k_ix) Y_{l_im_i}(\hat{\mathbf x})^* } B_\Psi(k_1, k_2, k_3) \\
+%     = A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3/2}}4\pi i^{l_i}(2i^{l_i})j_{l_i}(k_i\chi_i)j_{l_i}(k_ix)}
+%     (2\pi)^3  B_\Psi(k_1, k_2, k_3) \\
+%     = A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\frac{1}{(2\pi)^{3/2}}4\pi i^{l_i}(2i^{l_i})\frac{\pi}{2\chi_i^2}\delta(x-\chi_i)}
+%     (2\pi)^3  B_\Psi(l_1/\chi, l_2/\chi, l_3/\chi) \\
+%     = A_{\mathbf l}^{\mathbf m}\int_0^{\chi_*} x^2\d x\prod_i\br{-2 W(x, \chi_*)\frac{1}{(2\pi)^{3/2}}4\pi i^{l_i}(2i^{l_i})\frac{\pi}{2x^2}}
+%     (2\pi)^3  B_\Psi(l_1/\chi, l_2/\chi, l_3/\chi) \\
+%     = A_{\mathbf{l}}^{\mathbf{m}} (-1)^{1 + l_1 + l_2 + l_3} 8 (2\pi)^{9/2} \int_0^{\chi_*} \frac{[W(\chi, \chi_*)]^3}{\chi^4} \, B_\Psi\left( \frac{l_1}{\chi}, \frac{l_2}{\chi}, \frac{l_3}{\chi} \right ) \, \mathrm{d}\chi.
+% \end{gather*}
+
+% TODO: mention that the definition of the bispectrum is different form power spectrum of Psi in terms of normalization constant
+% TODO: explain 3d dirac delta in terms of spherical harmonics (eq 78 thesis Luna)
+% TODO: explain that we used that int d^2 Omega Y Y Y = A_vec l^vec m (eq 82 thesis Luna)
+% using that d^3 k = k^2 d k d^2 Omega_k
+% TODO: fix this
+\subsection{Gravitational potential spectra in terms of matter spectra}
+% TODO: fix references and citations 
+To convert equations ...
+and ...
+to forms that we can directly evaluate we will need to replace the spectra of the graviational potential with matter spectra.
+Specifically, we define the matter power spectrum in agreement with the CAMB package documentation CITE and the matter bispectrum in agreement with CITE FITTING FUNC PAPER such that we can approximate it in terms of the matter power spectrum using hyperextended perturbation theory.
+The \textbf{density contrast} is defined as
+\begin{equation}
+    \delta(\mathbf x) := \frac{\rho(\mathbf x)-\bar \rho}{\bar\rho},
+\end{equation}
+and the matter spectra are defined in terms of the fourier transformed density contrast $\delta(\mathbf k)$ as 
+\begin{gather*}
+    \langle \delta(\mathbf k, \eta) \delta(\mathbf k', \eta)^* \rangle = (2\pi)^3\delta(\mathbf k - \mathbf k')P^\delta(\mathbf k, \eta),\\
+    \langle \delta(\mathbf k_1, \eta) \delta(\mathbf k_2, \eta) \delta(\mathbf k_3, \eta) \rangle = (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\delta(k_1, k_2, k_3, \eta).
+\end{gather*}
+The mean matter density of the universe, \( \bar{\rho} \), can be expressed in terms of the matter density parameter \( \Omega_m \) and the critical density \( \rho_c \) as 
+$$
+\bar{\rho}(\eta) = \frac{3 \Omega_m H_0^2}{8 \pi G}\frac{1}{a(\eta)^3},
+$$
+where $a(\eta)$ is the only time dependent factor on the right hand side.
+In the case of a flat universe ($K=0$) we can thus rewrite the Poisson equation REF as
+% in modern cosm eq 6.41 they say - nabla^2 Phi = ..., however here we use Psi instead and Psi = - Phi in our regime so hence why we don't take the - sign
+\begin{equation}
+    \nabla^2\Psi = 4\pi G a^2\br{\frac{3\Omega_m H_0^2}{8\pi G}\frac{1}{a^3}}\delta = \frac{3\Omega_m H_0^2}{2} \frac{1}{a}\delta \implies \Psi(k, \eta) = -\frac{3\Omega_m H_0^2}{2}\frac{1}{a}\frac{\delta(k, \eta)}{k^2},
+\end{equation}
+where $\Psi(k,\eta)$ and $\delta(k,\eta)$ are functions in Fourier space. For the power- and bispectra we find
+\begin{gather*}
+    \langle \Psi(\mathbf k, \eta)\Psi^*(\mathbf k',\eta)\rangle = \frac{2\pi^2}{k^3}P^\Psi(k, \eta)\delta(\mathbf k - \mathbf k') \implies P^\Psi(k,\eta) = \frac{1}{k}(9\Omega_m^2H_0^4\pi) \frac{1}{a^2}P^\delta(k,\eta),\\
+    \langle \Psi(\mathbf k_1, \eta)\Psi(\mathbf k_2,\eta)\Psi(\mathbf k_3,\eta)\rangle = -(2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(k_1,k_2,k_3, \eta)\\ 
+    \implies B^\Psi(k_1,k_2,k_3, \eta) = -\frac{1}{k_1^2k_2^2k_3^2}\br{\frac{3\Omega_m H_0^2}{2}}^3 \frac{1}{a^3}B^\delta(\{k_i\}, \eta).
+\end{gather*}
+We conclude this section by using the above equations to rewrite the lensing potential spectra (equations REF and REF) in terms of matter spectra 
+\begin{empheq}[box=\fbox]{align*}
+    P^{\psi_X\psi_Y}_l
+    & = \frac{9}{l^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)P^\delta(l/\chi,\eta_0-\chi),\\
+    B^{\psi_X\psi_Y\psi_Z}_{l_1l_2l_3} &= \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{l_1^2l_2^2l_3^2}\Omega_m^3H_0^6\\
+    & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{l_i/\chi\}, \eta_0-\chi).
+\end{empheq}
+
+% \begin{gather*}
+%     \langle \psi_{l_1m_1} \psi_{l_2m_2} \psi_{l_3m_3} \rangle = A_{\mathbf{l}}^{\mathbf{m}} (-1)^{l_1 + l_2 + l_3} 8 (2\pi)^{9/2} \int_0^{\chi_*} W(\chi, \chi_*)^3 \frac{\chi^2}{l_1^2l_2^2l_3^2} \br{\frac{3\Omega_m H^2}{2}}^3|_{\eta=\eta_0-\chi}\\
+%     \quad \times B^\delta\br{ \frac{l_1}{\chi}, \frac{l_2}{\chi}, \frac{l_3}{\chi}; \eta_0-\chi } \mathrm{d}\chi, \\
+%     C_l^\psi = \frac{72\pi^3}{l^4}\int_0^{\chi_*}\d \chi \chi^2(\Omega_m^2H^4)|_{\eta=\eta_0-\chi} C^\delta_{l/\chi}(\eta_0-\chi) W(\chi)^2.
+% \end{gather*}
+
+\subsection{Convergence and shear multipole moments}
+\paragraph{Convergence}
+% TODO: make notation of nabla nicer/more consistent
+
+Taking the harmonic decomposition of the convergence,
+\begin{equation}
+    \kappa(\hat{\mathbf n}) = \kappa_{lm}Y_l^m(\hat{\mathbf n}),
+\end{equation}
+you can relate it to the multipole moments of the lensing potential through
+\begin{equation}
+    \kappa_{lm} = \frac{1}{2}(\nabla^2\psi)_{lm} = \frac{l(l+1)}{2}\psi_{lm}
+\end{equation}
+because the spherical harmonics are eigenfunctions of the Laplacian operator.
+
+\paragraph{Shear}
+On the other hand, the shear is given by
+\begin{equation}
+    \gamma = \gamma_1 + i\gamma_2 = \frac{1}{2}\eth_1(\eth_0\psi)
+\end{equation}
+where the spin raising operator, $\eth_s$ acts on a spin $s$ function defined on $S^2$ to create a spin $s+1$ function.
+The above equality is proven explicitely in appendix ...
+%TODO: write and cite appendix proving equivalence
+$\eth_s$ can be written in $(\theta, \phi)$ coordinates as
+\begin{equation}
+    \eth_s = -\sin^s\theta(\partial_\theta + \frac{i}{\sin\theta})\frac{1}{\sin^s\theta}.
+\end{equation}
+It is worth explaining that in this context a spin $s$ function referes to a function $_sf(\theta, \phi)$ that transforms under any rotation of coordinates by picking up a phase factor $e^{is\alpha}$, with $\alpha$ the angle of the rotation, i.e.
+\begin{equation}
+    f'(\theta', \phi') = e^{is\alpha}f(\theta, \phi).
+\end{equation}
+Shear is actually a spin 2 function, which can immediately be seen by observing that rotating a galaxy image stretched and squeezed through weak lensing by $180$ degrees gives the same stretching and squeezing, i.e.
+the same shear.
+It is thus expected that the shear is proportional to $\eth^2\psi$.
+
+The spin weighted spherical harmonics, $_sY_l^m$ are eigenfunctions of the raising/lowering operator, in particular
+\begin{equation}
+    \eth_sY_l^m = \sqrt{(l-s)(l+s+1)} _{s+1}Y_l^m.
+\end{equation}
+Decomposing the shear as
+\begin{equation}
+    \gamma(\theta, \phi) = \gamma_{lm} {}_sY_l^m(\theta,\phi),
+\end{equation}
+we thus find
+\begin{equation}
+    \gamma_{lm} = \frac{1}{2}\sqrt{(l-1)(l)(l+1)(l+2)} \psi_{lm}
+\end{equation}
 

### 2025-04-10 paper/paper_draft.tex
paper changes

@@
 \usepackage{authblk}
+\usepackage{multirow} % needed for merging cells vertically
+\usepackage{booktabs}
 
@@
 \section{Theory}
+\subsection{Weak lensing spectra}
 Radiation from cosmological objects is distorted due to gravitational lensing.
@@
 
+\subsection{Nonlinear matter powerspectrum}
+Determining the nonlinear matter powerspectrum can be done using numerical codes such as CAMB.
+The nonlinear matter bispectrum was calculated from the powerspectrum using a fitting formula based on perturbation theory in \cite{bispfit}.
+It is given by
+\begin{equation}
+    B_m(k_1, k_2, k_3, \chi) = 2 F_2(k_1, k_2, z) P_m(k_1, z) P_m(k_2, z) + \text{2 perm.}
+\end{equation}
+where \( P_m \) is the nonlinear matter power spectrum\footnote{Compare to the tree level bispectrum where we instead use the linear powerspectrum.}, and the kernel \( F_2 \) is modified from the tree level result with factors \( a(k, z) \), \( b(k, z) \), and \( c(k, z) \):
+\begin{equation}
+    F_2(k_1, k_2, z) = \frac{5}{7} a(k_1, z) a(k_2, z) + \frac{k_1^2 + k_2^2}{2 k_1 k_2} b(k_1, z) b(k_2, z) \cos \theta + \frac{2}{7} c(k_1, z) c(k_2, z) \cos^2 \theta.
+\end{equation}
+They are defined as:
+\begin{equation}
+    a(k, z) = \frac{1 + \sigma_8^{a_6}(z) \sqrt{0.7} Q(n_{\text{eff}}) (q^{a_1})^{n_{\text{eff}} + a_2}}{1 + (q^{a_1})^{n_{\text{eff}} + a_2}},
+\end{equation}
+\begin{equation}
+    b(k, z) = \frac{1 + 0.2 a_3 (n_{\text{eff}} + 3) (q^{a_7})^{n_{\text{eff}} + 3 + a_8}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_8}},
+\end{equation}
+\begin{equation}
+    c(k, z) = \frac{1 + \left[ \frac{4.5 a_4}{1.5 + (n_{\text{eff}} + 3)^4} \right] (q^{a_5})^{n_{\text{eff}} + 3 + a_9}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_9}}.
+\end{equation}
+Here, \( Q(n_{\text{eff}}) \) is given by:
+\begin{equation}
+    Q(x) = \frac{4 - 2x}{1 + 2^{x+1}}.
+\end{equation}
+The effective spectral index of the linear power spectrum is defined as:
+\begin{equation}
+    n_{\text{eff}} \equiv \frac{d \ln P_m^{\text{lin}}(k)}{d \ln k}.
+\end{equation}
+Additionally, \( q \) is given by:
+\begin{equation}
+    q = \frac{k}{k_{\text{NL}}},
+\end{equation}
+where \( k_{\text{NL}} \) is the scale at which nonlinearities become significant, satisfying:
+\begin{equation}
+    4 \pi k_{\text{NL}}^3 P_m^{\text{lin}}(k_{\text{NL}}, 0) = 1.
+\end{equation}
+The coefficients \( a_i \) are:
+\begin{align*}
+    a_1 &= 0.484, \quad a_2 = 3.740, \quad a_3 = -0.849, \quad a_4 = 0.392, \\
+    a_5 &= 1.013, \quad a_6 = -0.575, \quad a_7 = 0.128, \quad a_8 = -0.722, \quad a_9 = -0.926.
+\end{align*}
+
+\subsection{Fisher matrix analysis}
+SNR and parameter constraints are obtained through a Fisher matrix analysis as is standard.
+A full derivation of below statements can be found in appendix \ref{sec:fisher}.
+The auto and cross power spectrum Fisher matrix is given as
+\begin{equation*}
+    F_{\alpha\beta} = \sum_{l} \sum_{XY}\sum_{X'Y'}(2l+1)\partial_\alpha C_l^{XY} (C^{-1})_l^{XX'}(C^{-1})_l^{YY'} \partial_\beta C^{X'Y'}_l
+\end{equation*}
+where
+\begin{equation*}
+    C_{l} := \begin{pmatrix}
+        C^{\psi_\kappa\psi_\kappa}_l & C^{\psi_\kappa\psi_\gamma}_l \\
+        C^{\psi_\kappa\psi_\gamma}_l & C^{\psi_\gamma\psi_\gamma}_l
+    \end{pmatrix}.
+\end{equation*}
+For the bispectra we instead get 
+\begin{equation*}
+    F_{\alpha\beta} = \sum_{l_1 \leq l_2 \leq l_3} \frac{\mathcal P _{l_1l_2l_3}}{6}
+    \sum_{XYZ}\sum_{X'Y'Z'} 
+    \partial_\alpha B^{X Y Z}_{l_1 l_2 l_3} 
+    (C^{-1})^{X X'}_{l_1}
+    (C^{-1})^{Y Y'}_{l_2}
+    (C^{-1})^{Z Z'}_{l_3}
+    \partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3},
+\end{equation*}
+with $\mathcal P_{l_1l_2l_3}$ defined as the number of distinct permutations that can be made with $l_1l_2l_3$
+
+When only considering autospectra the formulas simplify to
+\begin{equation*}
+    F_{\alpha\beta} = \sum_{l} \sum_{XY}\sum_{X'Y'}(2l+1)\frac{\partial_{\alpha}C_l^{XX} \partial_\beta C^{XX}_l}{(C_l^{XX})^2},
+\end{equation*}
+and, for bispectra,
+\begin{equation*}
+    F_{\alpha\beta} = \sum_{l_1 \leq l_2 \leq l_3} \frac{\mathcal P _{l_1l_2l_3}}{6}
+    \frac{\partial_\alpha B^{X X X}_{l_1 l_2 l_3} 
+    \partial_\beta B^{XXX}_{l_1 l_2 l_3}}{C^{XX}_{l_1}C^{XX}_{l_2}C^{XX}_{l_3}}.
+\end{equation*}
+
 \section{Calculation Details}\label{sec:calcdetails}
@@
 Running cosmological simulations to obtain the matter power spectra and other functions was done with the CAMB package (REF).
-
-The derivatives of the lensing spectra were done numerically using a four point central difference formula, given by
-\begin{equation*}
-    f'(x) = \frac{-f(x + 2h) + 8f(x+h) - 8 f(x - h) + f(x - 2h)}{12h} + \mathcal O (h^4).
-\end{equation*}
-$h$ was chosen to be $2.5\%$ of the $x$ value.
-Testing showed this to generally be the best choise to avoid both the inacuracies from numerical errors (relevant for small $h$) and inacuracies due to the $\mathcal O(h^4)$ error inherent in the four point finite difference formula.
-The derivative w.r.t.
-$\sigma_8$ was an exception as it could not be taken by varying it directly and so was calculated implicitly using the chain rule.
 
@@
 
-The matter bispectrum was calculated from the matter powerspectrum using a fitting formula based on perturbation theory \cite{bispfit}. The bispectrum is given by
-\begin{equation}
-    B_m(k_1, k_2, k_3, \chi) = 2 F_2(k_1, k_2, z) P_m(k_1, z) P_m(k_2, z) + \text{2 perm.}
-\end{equation}
-where \( P_m \) is the matter power spectrum, and the kernel \( F_2 \) is given by:
-\begin{equation}
-    F_2(k_1, k_2, z) = \frac{5}{7} a(k_1, z) a(k_2, z) + \frac{k_1^2 + k_2^2}{2 k_1 k_2} b(k_1, z) b(k_2, z) \cos \theta + \frac{2}{7} c(k_1, z) c(k_2, z) \cos^2 \theta.
-\end{equation}
-The functions \( a(k, z) \), \( b(k, z) \), and \( c(k, z) \) are defined as:
-\begin{equation}
-    a(k, z) = \frac{1 + \sigma_8^{a_6}(z) \sqrt{0.7} Q(n_{\text{eff}}) (q^{a_1})^{n_{\text{eff}} + a_2}}{1 + (q^{a_1})^{n_{\text{eff}} + a_2}},
-\end{equation}
-\begin{equation}
-    b(k, z) = \frac{1 + 0.2 a_3 (n_{\text{eff}} + 3) (q^{a_7})^{n_{\text{eff}} + 3 + a_8}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_8}},
-\end{equation}
-\begin{equation}
-    c(k, z) = \frac{1 + \left[ \frac{4.5 a_4}{1.5 + (n_{\text{eff}} + 3)^4} \right] (q^{a_5})^{n_{\text{eff}} + 3 + a_9}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_9}}.
-\end{equation}
-Here, \( Q(n_{\text{eff}}) \) is given by:
-\begin{equation}
-    Q(x) = \frac{4 - 2x}{1 + 2^{x+1}}.
-\end{equation}
-The effective spectral index of the linear power spectrum is defined as:
-\begin{equation}
-    n_{\text{eff}} \equiv \frac{d \ln P_m^{\text{lin}}(k)}{d \ln k}.
-\end{equation}
-Additionally, \( q \) is given by:
-\begin{equation}
-    q = \frac{k}{k_{\text{NL}}},
-\end{equation}
-where \( k_{\text{NL}} \) is the scale at which nonlinearities become significant, satisfying:
-\begin{equation}
-    4 \pi k_{\text{NL}}^3 P_m^{\text{lin}}(k_{\text{NL}}) = 1.
-\end{equation}
-The coefficients \( a_i \) are:
-\begin{align*}
-    a_1 &= 0.484, \quad a_2 = 3.740, \quad a_3 = -0.849, \quad a_4 = 0.392, \\
-    a_5 &= 1.013, \quad a_6 = -0.575, \quad a_7 = 0.128, \quad a_8 = -0.722, \quad a_9 = -0.926.
-\end{align*}
-
 \subsection{Choosing Experimental Parameters}
+
 % lmin, lmax, galaxy redshift distribution, fiducial values (!), noise levels, 
@@
 
-For the parameter constraints we took the multipole range to be from 2 ($l_{min}$) to 2000 ($l_{max}$).
-
-% We will consider multiple possibilities for the other calculation parameters (noise levels and redshift distribution of observed galaxies).
-This is done to allow us to more accurately model constraints from measurements by the Simons Observatory (SO), CMB-S4, Euclid, and Large Synoptic Survey Telescope (LSST) surveys.
-The first two measure CMB lensing while the latter two measure lensing shear of galaxies.
-
-For noise in CMB lensing we consider lensing estimated by an experiment measuring polarization with white noise $\Delta_P$ and beam spread $\sigma$.
-We then use the quadratic estimator from \cite{Hu_2002} to estimate the resulting noise in the lensing power spectrum.
+% For the parameter constraints we took the multipole range to be from 2 ($l_{min}$) to 2000 ($l_{max}$).
+
+This paper considers noise levels for stage 3 and stage 4 weak lensing surveys.
+Values can be found in table \ref{tab:noiselevels}.
+A comparison of noise power spectrum to lensing power spectrum can be found in figure \ref{fig:lpsplusnoise}.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{|c|c|c|c|}
+    \hline
+    \textbf{source} & \textbf{survey stage} & \textbf{noise vals} & \textbf{comparable experiments} \\
+    \hline 
+    \multirow{2}{*}{CMB} 
+        & stage 3 & $\sigma = 1'$, $\Delta_P = 6' \; \mu \text{K}$ & Advanced ACTPol, SPT-3G \\
+    \cline{2-4}
+        & stage 4 & $\sigma = 3'$, $\Delta_P = 1' \; \mu \text{K}$ & CMB-S4, LiteBIRD \\
+    \hline
+    \multirow{2}{*}{galaxies} 
+        & stage 3 & $\sigma_{\text{rms}} = 0.3$, $n_g = 5 \text{arcmin}^{-2}$ & DES, KiDS \\
+    \cline{2-4}
+        & stage 4 & $\sigma_{\text{rms}} = 0.3$, $n_g = 30 \text{arcmin}^{-2}$ & LSST, Euclid \\
+    \hline
+    \end{tabular}
+    \caption{Noise levels considered for weak lensing of galaxies and the CMB.
+$\sigma$ (beam width) and $\Delta_P$ (polarization white noise) describe CMB survey specifications, while $\sigma_{\text{rms}}$ (intrinsic galaxy ellipticity) and $n_g$ (observed galaxy density) refer to galaxy shear surveys.}
+    \label{tab:noiselevels}
+\end{table}
+
+CMB noise is estimated through the quadratic estimator developed in \cite{cmblensingestimator}.
+We do not take temperature measurements into account, as these are subleading for stage 4 experiments (REF).
+This is similarly done in \cite{Namikawa_2016}.
+The parameters characterizing the noise levels are thus beam width, $\sigma$ and polarization white noise, $\Delta_P$.
+Galaxy lensing is determined by measuring lensing shear (see appendices \ref{sec:weaklensing} and \ref{sec:shear}).
+The noise in this measurement is dominated by scale independent shot noise and has associated noise power spectrum $N_l^{\text{shear}} = \sigma_{\text{rms}}^2 / n_g$.
+$n_g$ is the amount of galaxies observed per unit angle.
+The noise power spectrum for the lensing potential then equals
+\begin{equation*}
+    N_l^{\text{lens. potential}} = \frac{(l-1)l(l+1)(l+2)}{4}N_l^{\text{shear}}.
+\end{equation*}
+For parameter constraints we use stage 4 noise levels.
 
@@
 
-For shear measurements, the noise is simpler to model as the primary source is shot noise due to the discrete number of galaxies in each patch of the survey.
+% For noise in CMB lensing we consider lensing estimated by an experiment measuring polarization with white noise $\Delta_P$ and beam spread $\sigma$.
+We then use the quadratic estimator from \cite{Hu_2002} to estimate the resulting noise in the lensing power spectrum.
+
+% For shear measurements, the noise is simpler to model as the primary source is shot noise due to the discrete number of galaxies in each patch of the survey.
 The mean ellipiticity of the galaxies in a patch is calculated and a deviation from the mean is interpreted as lensing shear.
@@
 The power spectrum of the shear noise is given by
-\begin{equation*}
-    N^{\gamma\gamma}_{l} = \frac{\sigma^2_{rms}}{\bar n_g}
-\end{equation*}
-where $\bar n_g$ is the mean number of galaxies observed per sterradian accross the whole survey \cite{weaklensingnotes2020}.
-and $\sigma^2_\gamma$ is the intrinsic ellipticity dispersion. 
-Experiments such as Euclid and the LSST both have $\sigma_{rms} \approx 0.3$ \cite{euclid1overview} \cite{lsstsciencebook} and 
-$ n_g\approx 30 \text{ arcmin}^{-2}$.
-
-The noise values are calculated for stage 3 (current) surveys and stage 4 (near future) surveys.
+% \begin{equation*}
+%     N^{\gamma\gamma}_{l} = \frac{\sigma^2_{rms}}{\bar n_g}
+% \end{equation*}
+% where $\bar n_g$ is the mean number of galaxies observed per sterradian accross the whole survey \cite{weaklensingnotes2020}.
+% and $\sigma^2_\gamma$ is the intrinsic ellipticity dispersion. 
+% Experiments such as Euclid and the LSST both have $\sigma_{rms} \approx 0.3$ \cite{euclid1overview} \cite{lsstsciencebook} and 
+% $ n_g\approx 30 \text{ arcmin}^{-2}$.
+
+% The noise values are calculated for stage 3 (current) surveys and stage 4 (near future) surveys.
 We will use the stage 4 noise values for parameter constraints.

### 2025-04-10 code/Fisher_calc.c
misc

@@ -19452,7 +19452,7 @@ static int __pyx_f_11Fisher_calc_Delta(int __pyx_v_l1, int __pyx_v_l2, int __pyx
  *     else:
  *         return 1             # <<<<<<<<<<<<<<
  * 
- * # cdef double Fisher_mat_single_term(int l1, int l2, int l3, char* type, char* par1, char* par2, int num_samples) noexcept nogil:
+ * cdef double Fisher_mat_single_term(int l1, int l2, int l3, char* type, char* par1, char* par2, int num_samples) noexcept nogil:
  */
   /*else*/ {
     __pyx_r = 1;
@@ -19472,33 +19472,33 @@ static int __pyx_f_11Fisher_calc_Delta(int __pyx_v_l1, int __pyx_v_l2, int __pyx
   return __pyx_r;
 }
 
-/* "Fisher_calc.pyx":64
- * #         * lps_f_obs(l2, type, type)**(-1) * lps_f_obs(l3, type, type)**(-1) * lbs_der(l1, l2, l3, type, type, type, num_samples, par2) / Delta(l1, l2, l3)
+/* "Fisher_calc.pyx":60
+ *         return 1
  * 
  * cdef double Fisher_mat_single_term(int l1, int l2, int l3, char* type, char* par1, char* par2, int num_samples) noexcept nogil:             # <<<<<<<<<<<<<<
- *         return 1.
- * 
+ *         return lbs_der(l1, l2, l3, type, type, type, num_samples, par1) * lps_f_obs(l1, type, type)**(-1) \
+ *         * lps_f_obs(l2, type, type)**(-1) * lps_f_obs(l3, type, type)**(-1) * lbs_der(l1, l2, l3, type, type, type, num_samples, par2) / Delta(l1, l2, l3)
  */
 
[…]

### 2025-04-10 code/bisp_compare_toshiya.py

@@ -1,12 +1,12 @@
 import cosm_setup as cs
 
-spectra = cs.lensing_spectra(
-    ombh2=0.0223,
-    omch2=0.119,
-    As=2.13e-9,
-    ns=0.965,
-    fiducial_k_nls=False
-    )
+spectra = cs.lensing_spectra()
+# spectra = cs.lensing_spectra(
+#     # ombh2=0.0223,
+#     # omch2=0.119,
+#     # As=2.13e-9,
+#     # ns=0.965,
+#     )
 
 import numpy as np
 import matplotlib.pyplot as plt

### 2025-04-10 code/cosm_setup.py

@@ -1,14 +1,8 @@
-print(f'now running setup.py as {__name__}')
+print(f'Now running setup.py as: {__name__}.')
 
 import camb
 import numpy as np
 import scipy
-import scipy.integrate
-import scipy.optimize
-import scipy.interpolate
-from scipy.interpolate import CubicSpline
-
-# np.seterr(all='warn')
 
 #======================================
 
@@ -18,7 +12,7 @@ class lensing_spectra:
   '''
   Calculates relevant features of a cosmology given input parameters and matter bispectrum fitting constants
   '''
-  def __init__(self, H0=67.4, ombh2=0.0223, omch2=0.119, ns=0.965, As=2.13e-9, mnu=0.06, w0=-1, wa=0, redshifts=np.linspace(1100, 0, 200), matter_bispec_fit_params = matter_bispec_fit_params, fiducial_k_nls = False):
+  def __init__(self, H0=67.4, ombh2=0.0223, omch2=0.119, ns=0.965, As=2.13e-9, mnu=0.06, w0=-1, redshifts=np.linspace(1100, 0, 200), matter_bispec_fit_params = matter_bispec_fit_params):
     '''
     Initialize instance of class, set results and some methods/attributes, default inputs are for fiducial cosmology
 
@@ -77,21 +71,8 @@ class lensing_spectra:
     self.z_galaxy_source = 5
     self.chi_galaxy_source = self.results.comoving_radial_distance(self.z_galaxy_source)
 
-    if fiducial_k_nls:
-      print('took fiducial k_nls')
-      k_nl_data = np.loadtxt('/home3/p319950/ResearchProject/fisher_calc_weak_lensing/code/fiducial_k_nls.txt') # needs to be updated
-      self.zs_for_k_nls_used = k_nl_data[:, 0]
-      self.k_nls = k_nl_data[:, 1]
-    
-    else:
-      print('now calculating k_nls')
-      self.zs_for_k_nls = np.linspace(0, 8, 101)
-      self.k_nls, self.k_nl_errors, self.zs_for_k_nls_used = self.k_nl_exact(self.zs_for_k_nls)
-      
-    self.zs_for_k_nls_used_max = self.zs_for_k_nls_used[-1]
-    print(f'Max z for extended F_2: {self.zs_for_k_nls_used_max}')
-  
-    self._k_nl = CubicSpline(self.zs_for_k_nls_used, self.k_nls, bc_type='not-a-knot')
[…]

### 2025-04-10 code/interpolation.pyx

@@ -12,11 +12,7 @@ cpdef double linear_interp(double x, double xmin, double xmax, int n_points, dou
     cdef double dx = x - (xmin + i * (xmax - xmin) / (n_points - 1))
     cdef double dydx = (ys[i + 1] - ys[i]) / ((xmax - xmin) / (n_points - 1))
 
-    # if i >= ys.shape[0]:
-    #     printf("oh noooo :( (%d and %d)", i, ys.shape[0])
 
-    # printf("niggershape: %d\n", ys.shape[0])
-    # printf("i: %d\n", i)
     if isnan(ys[i] + dx * dydx) or isinf(ys[i] + dx * dydx):
         printf('nonono')
     return ys[i] + dx * dydx

### 2025-04-11 paper/paper_draft.tex
paper changes

@@
 \begin{equation*}
-    F_{\alpha\beta} = \sum_{l} \sum_{XY}\sum_{X'Y'}(2l+1)\frac{\partial_{\alpha}C_l^{XX} \partial_\beta C^{XX}_l}{(C_l^{XX})^2},
+    F_{\alpha\beta} = \sum_{l} \sum_{XY}\sum_{X'Y'}\frac{2l+1}{2}\frac{\partial_{\alpha}C_l^{XX} \partial_\beta C^{XX}_l}{(C_l^{XX})^2},
 \end{equation*}
@@
 \section{Calculation Details}\label{sec:calcdetails}
-\subsection{Numerical Considerations}
-% I used CAMB, how I took derivative, interpolation, fisher matrix sampling, matter bispectrum in terms of power spectrum
-
-Running cosmological simulations to obtain the matter power spectra and other functions was done with the CAMB package (REF).
-
-To speed up the calculation of the lensing powerspectra and especially the bispectra, an interpolation approach was used.
+% \subsection{Numerical Considerations}
+% % I used CAMB, how I took derivative, interpolation, fisher matrix sampling, matter bispectrum in terms of power spectrum
+
+% Running cosmological simulations to obtain the matter power spectra and other functions was done with the CAMB package (REF).
+
+% To speed up the calculation of the lensing powerspectra and especially the bispectra, an interpolation approach was used.
 Due to being 3 dimensional, it was not feasible to interpolate the lensing bispectrum directly.
@@
 
-The time complexity of the Fisher matrix scales with the maximum multipole considered, $l_{max}$, as $\mathcal O (l_{max}^3)$, leading to time issues for realistic values of $l_{max}$.
+% The time complexity of the Fisher matrix scales with the maximum multipole considered, $l_{max}$, as $\mathcal O (l_{max}^3)$, leading to time issues for realistic values of $l_{max}$.
 We thus calculated the $l_1 < l_2 < l_3$ part of each Fisher matrix element (see the appendix) by sampling every step of 10 %(CHECK IF THIS IS STILL CORRECT BEFORE PUBLISHING)
-in $l_1$ and $l_2$.
+% in $l_1$ and $l_2$.
 Due to the discontinuos nature of the bispectrum (e.g.
@@
 
-\subsection{Choosing Experimental Parameters}
+% \subsection{Choosing Experimental Parameters}
 
@@
     H_0 &= 67.4 \, \text{km/s/Mpc}, && \text{(Hubble constant)} \\
-    \Omega_b h^2 &= 0.0224, && \text{(Physical baryon density parameter)} \\
-    \Omega_c h^2 &= 0.120, && \text{(Physical cold dark matter density parameter)} \\
+    \Omega_b h^2 &= 0.0223, && \text{(Physical baryon density parameter)} \\
+    \Omega_c h^2 &= 0.119, && \text{(Physical cold dark matter density parameter)} \\
     n_s &= 0.965, && \text{(Scalar spectral index)} \\
-    A_s &= 2.1 \times 10^{-9}, && \text{(Amplitude of primordial scalar fluctuations)} \\
+    A_s &= 2.13 \times 10^{-9}, && \text{(Amplitude of primordial scalar fluctuations)} \\
     \sum m_\nu &= 0.06 \, \text{eV}, && \text{(Sum of neutrino masses)} \\
@@
 \begin{equation}
-    e^{i\mathbf k \cdot \hat {\mathbf n}\chi}=4\pi\sum_{lm}i^lj_l(k\chi )Y_l^m(\hat{\mathbf n})^*Y_l^m(\hat{\mathbf k}) = 4\pi\sum_{lm}i^lj_l(k\chi )Y_l^m(\hat{\mathbf n})Y_l^m(\hat{\mathbf k})^*,
+    e^{i\mathbf k \cdot \hat {\mathbf n}\chi}=4\pi\sum_{lm}i^lj_l(k\chi )Y_l^m(\hat{\mathbf n})^*Y_l^m(\hat{\mathbf k}) = 4\pi\sum_{lm}i^lj_l(k\chi )Y_l^m(\hat{\mathbf n})Y_l^m(\hat{\mathbf k})^*, \label{eq:complexexp}
 \end{equation}
@@
 Similarly to $\delta(x-x_0)f(x)=\delta(x-x_0)f(x_0)$, we thus take $j_l(k\chi)f(k)\approx j_l(k\chi)f(l/\chi)$.
-The Bessel functions satisfy an orthogonality condtion,
+The Bessel functions satisfy an orthogonality condition,
 \begin{equation}
@@
     C^\psi(l) = \frac{2}{\pi^2}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi') \frac{\pi}{2\chi^2}\delta(\chi-\chi')\frac{\chi^3}{l^3}P_\psi(l/\chi,\eta_0 - \chi, \eta_0-\chi')\\
-    = \frac{1}{l^3\pi}\int_0^{\chi_*}\chi\d \chi W(\chi)^2 P_\psi(l/\chi,\eta_0 - \chi, \eta_0-\chi).
+    = \frac{1}{l^3\pi}\int_0^{\chi_*}\chi\d \chi W(\chi)^2 P_\psi(l/\chi,\eta_0 - \chi, \eta_0-\chi). \label{eq:lpsintermsofpsi}
 \end{gather}
@@
 \end{gather*}
-Now using the identity REF for the complex exponential:
+Now using equation \ref{eq:complexexp} to rewrite the complex exponential:
 \begin{gather*}
@@
 
-The definition for the bispectrum of any set of randomly distributed spherical harmonic components $X^k_{lm}$ is CITE Hu 2000 
+The definition for the bispectrum of any set of randomly distributed spherical harmonic components $X^k_{lm}$ is \cite{Hu2000}
 \begin{equation*}
@@
 In particular, it vanishes if $l_1+l_2+l_3$ is odd, meaning we can drop the $(-1)^{l_1+l_2+l_3}$ factor. Additionally cancelling common factors then gives
-\begin{align*}
+\begin{align}
     B^{\psi_1\psi_2\psi_3}_{l_1l_2l_3}
@@
     -\sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} 8 \int \frac{\d \chi }{\chi^4} W_1(\chi)W_2(\chi)W_3(\chi) B^\Psi(\{l_i/\chi\}, \eta_0-\chi).
-\end{align*}
+\label{eq:lbsintermsofpsi}
+\end{align}
 
@@
 % TODO: fix references and citations 
-To convert equations ...
-and ...
-to forms that we can directly evaluate we will need to replace the spectra of the graviational potential with matter spectra.
-Specifically, we define the matter power spectrum in agreement with the CAMB package documentation CITE and the matter bispectrum in agreement with CITE FITTING FUNC PAPER such that we can approximate it in terms of the matter power spectrum using hyperextended perturbation theory.
+We can rewrite equations \ref{eq:lpsintermsofpsi} and \ref{eq:lbsintermsofpsi} in terms of the matter spectra instead of the graviational potential spectra using the poisson equation.
+This then allows us to numerically evaluate these objects using CAMB.
 The \textbf{density contrast} is defined as
@@
 where $a(\eta)$ is the only time dependent factor on the right hand side.
-In the case of a flat universe ($K=0$) we can thus rewrite the Poisson equation REF as
+In the case of a flat universe ($K=0$) we can thus rewrite the Poisson equation \cite{dodelson2020modern} as
 % in modern cosm eq 6.41 they say - nabla^2 Phi = ..., however here we use Psi instead and Psi = - Phi in our regime so hence why we don't take the - sign
@@
 \end{gather*}
-We conclude this section by using the above equations to rewrite the lensing potential spectra (equations REF and REF) in terms of matter spectra 
+Finally, we obtain:
 \begin{empheq}[box=\fbox]{align*}

### 2025-04-13 paper/paper_draft.tex
paper changes

@@
 
-\subsection{Nonlinear matter powerspectrum}
+\subsection{Nonlinear matter bispectrum}
 Determining the nonlinear matter powerspectrum can be done using numerical codes such as CAMB.
@@
 \end{equation*}
-with $\mathcal P_{l_1l_2l_3}$ defined as the number of distinct permutations that can be made with $l_1l_2l_3$
-
-When only considering autospectra the formulas simplify to
+with $\mathcal P_{l_1l_2l_3}$ defined as the number of distinct permutations that can be made with $l_1l_2l_3$. When only considering autospectra the formulas simplify to
 \begin{equation*}
@@
 \end{equation*}
-and, for bispectra,
+and (for bispectra)
 \begin{equation*}
@@
 \section{Weak Lensing}\label{sec:weaklensing}
-\subsection{Approximations Made / Regime}
 \subsection{Perturbed Photon Paths}
-Work in \textbf{conformal newtonian gauge}.
-In this section we'll work in units where $c=1$.
-Denoting conformal time and conformal radial distance by $\eta$ and $\chi$, respectively, the perturbed line element is given by
-\begin{equation}
-    \d s^2 = a^2(\eta)((1 + 2\Psi_N) \d \eta^2 - (1 + 2\Phi_N) \gamma_{ij}\d x^i \d x^j) 
-\end{equation}
-
+We work in the conformal newtonian gauge and with natural units.
+Denoting conformal time and conformal radial distance by $\eta$ and $\chi$, respectively, the perturbed line element in FLRW spacetime is given by
+\begin{equation}
+    \d s^2 = a^2(\eta)((1 + 2\Psi_N) \d \eta^2 - (1 + 2\Phi_N) \gamma_{ij}\d x^i \d x^j),
+\end{equation}
 where $\gamma_{ij}$ is the unperturbed line element
 \begin{equation}
-    \gamma_{ij} = \d x^i \d x^j = \d \chi^2 + f^2_K(\chi)(\d \theta^2 + \sin ^2 \theta \d \phi ^2).
-\end{equation}
-From here of on work in first order in the scalar potentials $\Phi_N$ and $\Psi_N$. For lensing we consider null-geodesics so $\d s^2 = 0$ and we can rewrite the perturbed line element as
-\begin{equation}
-    \d \hat s^2 = (1 + 4\Psi)\d\eta^2 - \gamma_{ij} \d x^i \d x^j
-\end{equation}
-with $\Psi$ the \textbf{Weyl Potential} given by $\Psi := (\Psi_N - \Phi_N)/2$. A null geodesic $x^\mu(\hat\lambda)$ in terms of its affine parameter ($\hat \lambda$) satisfies the geodesic equation
-\begin{equation}
-    \frac{\d^2 x^\mu}{\d \hat\lambda^2} + \Gamma^\mu_{\nu\rho} \frac{\d x^\nu}{\d \hat\lambda}\frac{\d x^\rho}{\d\hat\lambda} = 0,
-\end{equation}
-with $\hat g_{\mu\nu}(\d x^\mu/\d \hat\lambda)(\d x^\nu / \d \hat \lambda) = 0$.
-
-In the unperturbed there are incoming radial solutions of the form $\chi = \eta_0 - \eta$. We can express $\hat\lambda$ in terms of $\eta$ using The 0-component of the geodesic equation,
-\begin{equation}
-    \frac{\d^2\eta}{\d\hat\lambda^2} + 2(\frac{\d\eta}{\d\hat\lambda})^2\frac{\d\Psi}{\d\eta}+2\frac{\d\eta}{\d\hat\lambda}\frac{\d x^i}{\d\hat\lambda}\frac{\partial\Psi}{\partial x^i} = 0,
-\end{equation}
-where the derivative $\d\Psi/\d\eta = \partial_\eta\Psi+(\d x^i/\d\eta)\partial_i\Psi$ is along the perturbed ray. This allows us to rewrite the other (spatial) geodesic equations as
-\begin{equation}
-    \frac{\d^2x^i}{\d\eta^2} - 2\frac{\d x^i}{\d\eta}(\frac{\d\Psi}{\d\eta} + \frac{\d x^j}{\d\eta}\frac{\partial\Psi}{\partial x^j})+2\gamma^{ij}\frac{\partial\Psi}{\partial x^j}+\bar\Gamma^i_{jk}\frac{\d x^j}{\d\eta}\frac{\d x^k}{\d \eta} = 0,
-\end{equation}
-where $\bar\Gamma^i_{jk}$ are the connection coefficients of the unperturbed three geometry $\gamma_{ij}$.
+    \gamma_{ij} = \d x^i \d x^j = \d \chi^2 + f^2_K(\chi)(\d \theta^2 + \sin ^2 \theta \d \phi ^2),
+\end{equation}
+and $f_K(\chi)$ is the comoving angular diameter distance.
+We will hereafter only consider a flat universe, so that $f_K(\chi) = \chi$.
+Weak lensing of a point source can be quantified by looking at the deflection field $\mathbf d (\hat{\mathbf n}) = \mathbf \theta_{\text{obs}} - \mathbf \theta_{\text{true}}$, i.e.
+the (small) difference between the angle at which we see the object and the angle at which we would see the object had no lensing occured.
+To first order in $\Psi_N$ and $\Phi_N$, the deflection is given as \cite{dodelson2020modern}
+\begin{equation}
+    \mathbf d(\hat{\mathbf n}) = -2 \int_0^{\chi_*}\d \chi \frac{\chi_* - \chi}{\chi_*\chi}\nabla_{\hat{\mathbf n}}\Psi(\chi\hat{\mathbf n};\eta_0 - \chi),
+\end{equation}
+where $\Psi$ is the Weyl Potential, $\Psi := (\Psi_N - \Phi_N)/2$, and $\chi_*$ is the conformal distance to the source.
+$\nabla_{\hat{\mathbf n}}$ is the derivative along the axes orthogonal to the line of sight.
+The above equation can be written in terms of the lensing potential, $\psi$, as $\mathbf d (\hat{\mathbf n}) = \nabla_{\hat{\mathbf n}}$, with
+\begin{equation}
+    \psi(\hat {\mathbf n}) := -2 \int_0^{\chi_*}\d \chi \frac{\chi_* - \chi}{\chi_*\chi}\Psi(\chi\hat{\mathbf n};\eta_0 - \chi).
+\end{equation}
+
+If the source is instead distributed radial distances according to some distribution function $p(\chi)$, with $p(\chi)$ normalized to integrate to 1, the $(\chi_* - \chi)/(\chi_*\chi)$ factor is changed as
+\begin{equation*}
+    \frac{\chi_*-\chi}{\chi_*\chi} \rightarrow W(\chi) := \int_\chi^{\infty}\d\chi' p(\chi')\frac{\chi'-\chi}{\chi'\chi}.
+\end{equation*}
+$W(\chi)$ is then called the window function. In the most general case, the lensing potential is thus given by
+\begin{equation}
+    \psi(\hat {\mathbf n}) := -2 \int_0^{\infty}\d \chi W(\chi)\Psi(\chi\hat{\mathbf n};\eta_0 - \chi).
+\end{equation}
+The integration limit is sometimes also taken to be the surface of last scattering, as any window function vanishes after that distance.
+In the case of CMB lensing we can take $p(\chi')=\delta(\chi'-\chi_*)$, in which case the window function reduces to $H(\chi_* - \chi)(\chi_*-\chi)/(\chi_*\chi)$, with $H(\chi)$ the Heaviside step function.
+
+% From here of on work in first order in the scalar potentials $\Phi_N$ and $\Psi_N$. For lensing we consider null-geodesics so $\d s^2 = 0$ and we can rewrite the perturbed line element as
+% \begin{equation}
+%     \d \hat s^2 = (1 + 4\Psi)\d\eta^2 - \gamma_{ij} \d x^i \d x^j
+% \end{equation}
+% with $\Psi$ the \textbf{Weyl Potential} given by $\Psi := (\Psi_N - \Phi_N)/2$. A null geodesic $x^\mu(\hat\lambda)$ in terms of its affine parameter ($\hat \lambda$) satisfies the geodesic equation
+% \begin{equation}
+%     \frac{\d^2 x^\mu}{\d \hat\lambda^2} + \Gamma^\mu_{\nu\rho} \frac{\d x^\nu}{\d \hat\lambda}\frac{\d x^\rho}{\d\hat\lambda} = 0,
+% \end{equation}
+% with $\hat g_{\mu\nu}(\d x^\mu/\d \hat\lambda)(\d x^\nu / \d \hat \lambda) = 0$.
+
+% In the unperturbed there are incoming radial solutions of the form $\chi = \eta_0 - \eta$. We can express $\hat\lambda$ in terms of $\eta$ using The 0-component of the geodesic equation,
+% \begin{equation}
+%     \frac{\d^2\eta}{\d\hat\lambda^2} + 2(\frac{\d\eta}{\d\hat\lambda})^2\frac{\d\Psi}{\d\eta}+2\frac{\d\eta}{\d\hat\lambda}\frac{\d x^i}{\d\hat\lambda}\frac{\partial\Psi}{\partial x^i} = 0,
+% \end{equation}
+% where the derivative $\d\Psi/\d\eta = \partial_\eta\Psi+(\d x^i/\d\eta)\partial_i\Psi$ is along the perturbed ray. This allows us to rewrite the other (spatial) geodesic equations as
+% \begin{equation}
+%     \frac{\d^2x^i}{\d\eta^2} - 2\frac{\d x^i}{\d\eta}(\frac{\d\Psi}{\d\eta} + \frac{\d x^j}{\d\eta}\frac{\partial\Psi}{\partial x^j})+2\gamma^{ij}\frac{\partial\Psi}{\partial x^j}+\bar\Gamma^i_{jk}\frac{\d x^j}{\d\eta}\frac{\d x^k}{\d \eta} = 0,
+% \end{equation}
+% where $\bar\Gamma^i_{jk}$ are the connection coefficients of the unperturbed three geometry $\gamma_{ij}$.
 Without loss of generality we consider an observer located at the origin of the spatial coordinates, meaning we are interested in rays that end at $x^i = 0$.
@@
 For the case of such rays we can evaluate the connection coefficients and rewrite the spatial geodesic equations as
-\begin{gather}
-    \frac{\partial^2\chi}{\d\eta^2}+2\frac{\d\Psi}{\d\eta}=0,\\
-    \frac{\d^2\theta}{\d\eta^2}-2\frac{\d\ln f_K(\chi)}{\d\chi}\frac{\d\theta}{\d\eta}+\frac{2}{f^2_K(\chi)}\frac{\partial\Psi}{\partial\theta}=0,\\
-    \frac{\d^2\phi}{\d\eta^2} - 2\frac{\d \ln f_K(\chi)}{\d\chi}\frac{\d\phi}{\d\eta}+\frac{2}{f^2_K(\chi)}\frac{1}{\sin^2(\theta)}\frac{\partial\Psi}{\partial\phi}=0.
-\end{gather}
-Integrating twice and using the null condition we can rewrite the the first equation as
-\begin{equation}
-    \chi = \eta_0 - \eta - 2\int_{\eta_0}^\eta\Psi\d\eta',
-\end{equation}
-where the integral is along the ray.
+% \begin{gather}
+%     \frac{\partial^2\chi}{\d\eta^2}+2\frac{\d\Psi}{\d\eta}=0,\\
+%     \frac{\d^2\theta}{\d\eta^2}-2\frac{\d\ln f_K(\chi)}{\d\chi}\frac{\d\theta}{\d\eta}+\frac{2}{f^2_K(\chi)}\frac{\partial\Psi}{\partial\theta}=0,\\
+%     \frac{\d^2\phi}{\d\eta^2} - 2\frac{\d \ln f_K(\chi)}{\d\chi}\frac{\d\phi}{\d\eta}+\frac{2}{f^2_K(\chi)}\frac{1}{\sin^2(\theta)}\frac{\partial\Psi}{\partial\phi}=0.
+% \end{gather}
+% Integrating twice and using the null condition we can rewrite the the first equation as
+% \begin{equation}
+%     \chi = \eta_0 - \eta - 2\int_{\eta_0}^\eta\Psi\d\eta',
+% \end{equation}
+% where the integral is along the ray.
 Since we are working at first order in $\Psi$ we can evaluate the integral along the unperturbed path, this is known as the \textbf{Born approximation}.\textit{Next follows a short discussion of why perturbations in observation time can be ignored, include?} Doing the same for the other two equations gives
-% TODO: make n a proper vector (bold)
-\begin{gather}
-    \theta(\eta_0 - \chi_*)=\theta_0 - \int_0^{\chi_*}\d\chi\frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}2\frac{\partial}{\partial \theta}\Psi(\chi\hat n, \eta_0 - \chi),\\
-    \phi(\eta_0 - \chi_*)=\theta_0 - \int_0^{\chi_*}\d\chi\frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}\frac{2}{\sin^2\theta}\frac{\partial}{\partial \phi}\Psi(\chi\hat n, \eta_0 - \chi),
-\end{gather}
-where $\theta_0$ and $\phi_0$ label the line of sight $\hat n$.
+% % TODO: make n a proper vector (bold)
+% \begin{gather}
+%     \theta(\eta_0 - \chi_*)=\theta_0 - \int_0^{\chi_*}\d\chi\frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}2\frac{\partial}{\partial \theta}\Psi(\chi\hat n, \eta_0 - \chi),\\
+%     \phi(\eta_0 - \chi_*)=\theta_0 - \int_0^{\chi_*}\d\chi\frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}\frac{2}{\sin^2\theta}\frac{\partial}{\partial \phi}\Psi(\chi\hat n, \eta_0 - \chi),
+% \end{gather}
+% where $\theta_0$ and $\phi_0$ label the line of sight $\hat n$.
 The displacement vector $\mathbf{\alpha}$ indicates the infinitesimal displacement of the lensed rays from their unpetrubed observed angles and is related to $\theta$ and $\phi$ through
-$$
-\alpha_\theta = \theta - \theta_0, \quad \alpha_\phi = \sin^2\theta(\phi - \phi_0).
-$$
-This allows us to summarize the above results as
-\begin{equation}
-    \mathbf \alpha = -2 \int_0^{\chi_*}\d \chi \frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}\nabla_{\hat{\mathbf n}}\Psi(\chi\hat{\mathbf n};\eta_0 - \chi).
-\end{equation}
-When working in the weak field limit (i.e. small spacetime curvature) the $00$ component of Einsteins equations give us a poisson equation \cite{dodelson2020modern},
-\begin{equation}
-    -\nabla^2\Psi(\textbf{x}, \eta) =  4\pi G a(\eta)^2\delta\rho(\mathbf x, \eta)
-\end{equation}
-where $K$ is $0$, $1$, or $-1$ depending on the curvature of the universe, $\Delta = \partial_x^2 + \partial_y^2 + \partial_z^2$ is the 3 dimensional laplace operator, $\delta\bar\rho$ is the comoving total energy perturbation (evaluated in the rest-frame of the total energy).
+% $$
+% \alpha_\theta = \theta - \theta_0, \quad \alpha_\phi = \sin^2\theta(\phi - \phi_0).
+% $$
+% An object 
+% \begin{equation}
+%     \mathbf \alpha = -2 \int_0^{\chi_*}\d \chi \frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}\nabla_{\hat{\mathbf n}}\Psi(\chi\hat{\mathbf n};\eta_0 - \chi).
+% \end{equation}
+% When working in the weak field limit (i.e. small spacetime curvature) the $00$ component of Einsteins equations give us a poisson equation \cite{dodelson2020modern},
+% \begin{equation}
+%     -\nabla^2\Psi(\textbf{x}, \eta) =  4\pi G a(\eta)^2\delta\rho(\mathbf x, \eta)
+% \end{equation}
+% where $K$ is $0$, $1$, or $-1$ depending on the curvature of the universe, $\Delta = \partial_x^2 + \partial_y^2 + \partial_z^2$ is the 3 dimensional laplace operator, $\delta\bar\rho$ is the comoving total energy perturbation (evaluated in the rest-frame of the total energy).
 %TODO: proper definition for anisotropic stress and explain why we can ignore it
 
-\subsection{Convergence, shear}
-Quantities we are interested in measuring depend on how the lensing vector changes with the observation angle. \textbf{Magnification matrix} is defined by
+\subsection{Convergence and Shear}
+Quantities we are interested in measuring depend on how the lensing vector changes with the observation angle. The magnification matrix is defined as
 \begin{equation}
@@
 
-\subsection{lensing potential}
-Taking the derivative out of the integral in equation REF leads us to defining the \textbf{lensing potential}, $\psi$,
-\begin{equation}
-    \mathbf \alpha = \nabla_{\hat{\mathbf n}}\psi, \quad \psi(\hat{\mathbf n}) = -2 \int_0^{\chi_*} \d \chi \frac{f_K(\chi_* - \chi)}{f_K(\chi_*)f_K(\chi)}\Psi(\chi\hat{\mathbf n}; \eta_0 - \chi).
-\end{equation}
-
 We can relate $\kappa$ and $\gamma$ directly to the lensing potential as
@@
 \end{equation}
-
-\subsection{Regarding window functions}
-The window function for light emitted from a source at fixed radius $\chi_*$ is $(\chi-\chi^*)/(\chi_*\chi)$, if the source is instead distributed over the radius as $p(\chi)$, with $p(\chi)$ normalized to integrate to 1, we generalize the window function as
-\begin{equation*}
-    W(\chi) = \int_\chi^{\infty}\d\chi' p(\chi')\frac{\chi'-\chi}{\chi'\chi}.
-\end{equation*}
-The integration limit is up to surface of last scattering because further than that no source can contribute to the observed image.
-In the case of an image of the CMB we can take $p(\chi')=\delta(\chi'-\chi_*)$, in which case the window function reduces back to $(\chi_*-\chi)/(\chi_*\chi)$.
-In the case of different sources, we will differentiate between their respective window functions with a subscript.
-In this case the factor that changes is the distribution function.
-The sources considered in this paper are only the CMB and selections of galaxies from galaxy surveys.
 

### 2025-04-17 paper/paper_draft.tex
paper changes

@@
 When comparing powerspectra to bispectra constraints in figure \ref{fig:S8const} we see a more mild improvement in constraints when combining the different sources of information.
-\begin{figure}[htbp]
+% \begin{figure}[htbp]
+%     \centering
+%     \begin{subfigure}[t]{0.48\textwidth}
+%         \centering
+%         \includegraphics[width=\textwidth]{paramconstraints_sigma8.pdf}
+%         
+%         \label{fig:sigma8const}
+%     \end{subfigure}
+%     \hfill
+%     \begin{subfigure}[t]{0.48\textwidth}
+%         \centering
+%         \includegraphics[width=\textwidth]{paramconstraints_sigma8_diffspectra.pdf}
+%         \caption{Same as Fig.~\ref{fig:sigma8const}, except comparing constraints from only using power spectra, only bispectra, or both.
+$S_8$ instead of $\sigma_8$.
+In all three cases we use CMB $and$ galaxy lensing.}
+%         \label{fig:S8const}
+%     \end{subfigure}
+%     \caption{Side-by-side comparison of parameter constraints.}
+%     \label{fig:side_by_side}
+% \end{figure}
+
+\begin{figure}
     \centering
-    \begin{subfigure}[t]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{paramconstraints_sigma8.pdf}
+    \includegraphics[width=\textwidth]{figures/paramconstraints_difftracers.pdf}
     \caption{Constraints for $H_0$ and $\sigma_8$ for CMB lensing, galaxy lensing, and CMB + galaxy lensing combined.
 We use the information from both power and bispectra together with $l_{\max}=2000$ in all three cases.
-        The confidence ellipses are for $1\sigma$ ($\sim70\%$) certainty.}
-        \label{fig:sigma8const}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[t]{0.48\textwidth}
+The confidence ellipses are for $1\sigma$ certainty.}
+    \label{fig:difftracer}
+\end{figure}
+\begin{figure}
     \centering
-        \includegraphics[width=\textwidth]{paramconstraints_sigma8_diffspectra.pdf}
-        \caption{Same as Fig.~\ref{fig:sigma8const}, except comparing constraints from only using power spectra, only bispectra, or both.
-$S_8$ instead of $\sigma_8$.
-In all three cases we use CMB $and$ galaxy lensing.}
-        \label{fig:S8const}
-    \end{subfigure}
-    \caption{Side-by-side comparison of parameter constraints.}
-    \label{fig:side_by_side}
+    \includegraphics[width=\textwidth]{figures/paramconstraints_diffspectra.pdf}
+    \caption{Same as Fig. \ref{fig:difftracer}, except comparing constraints from only using power spectra, only bispectra, or both. In all three cases we use CMB $and$ galaxy lensing.}
+    \label{fig:diffspectra}
 \end{figure}

### 2025-04-17 paper/fisherstuff.py

@@ -20,7 +20,7 @@ visb_c = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_c.txt')
 
 visb_s = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_s.txt')
 
-visb_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_both.txt')
+visb_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_both.txt')
 
 ########################
 # Powerspectra Matrices
@@ -141,8 +141,8 @@ def print_sorted_eigens(matrix_name, mat):
         print(vecs[:, i])
 
 # Indices to remove: 6,1,4,5 (0-based). This means we keep [0,2,3].
-keep_indices = [0, 7, 9]  # We'll keep H_0 and sigma_8
-keep_indices = [3, 6]  # We'll keep H_0 and sigma_8
+keep_indices = [0, 7, 9]
+# keep_indices = [3, 6]
 
 # Reduced copies for bispectra
 visb_c_reduced = visb_c[np.ix_(keep_indices, keep_indices)]
@@ -310,10 +310,10 @@ def plot_corner(
             # Default: ±5σ around param_values
             sigma_i = np.sqrt(cov_matrices[0][i, i])
             sigma_j = np.sqrt(cov_matrices[0][j, j])
-            x_min = param_values[i] - 1*sigma_i
-            x_max = param_values[i] + 1*sigma_i
-            y_min = param_values[j] - 1*sigma_j
-            y_max = param_values[j] + 1*sigma_j
+            x_min = param_values[i] - 1.5*sigma_i
+            x_max = param_values[i] + 1.5*sigma_i
+            y_min = param_values[j] - 1.5*sigma_j
+            y_max = param_values[j] + 1.5*sigma_j
 
             ax_low.set_xlim(x_min, x_max)
             ax_low.set_ylim(y_min, y_max)
@@ -377,13 +377,13 @@ if __name__ == "__main__":
 
     labels = ["CMB", "Galaxy", "CMB + Galaxy"]
 
-    cov_matrices = [
-        inv(visp_f_reduced),
-        inv(visb_f_reduced),
-        inv(visb_f_reduced + visp_f_reduced)
-    ]
+    # cov_matrices = [
+    #     inv(visp_f_reduced),
+    #     inv(visb_f_reduced),
+    #     inv(visb_f_reduced + visp_f_reduced)
+    # ]
[…]

### 2025-04-17 paper/tempCodeRunnerFile.py

@@ -1,0 +1,7 @@
+    # cov_matrices = [
+    #     inv(visp_f_reduced),
+    #     inv(visb_f_reduced),
+    #     inv(visb_f_reduced + visp_f_reduced)
+    # ]
+
+    # labels = ['powerspectra', 'bispectra', 'power- $+$ bispectra']

### 2025-04-22 paper/paper_draft.tex
paper changes

@@
 \begin{empheq}[box=\fbox]{align*}
-    P^{\psi_X\psi_Y}_l
-    & = \frac{9}{l^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)P^\delta(l/\chi,\eta_0-\chi),\\
+    C^{\psi_X\psi_Y}_l
+    & = \frac{9}{l^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)C^\delta(l/\chi,\eta_0-\chi),\\
     B^{\psi_X\psi_Y\psi_Z}_{l_1l_2l_3} &= \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{l_1^2l_2^2l_3^2}\Omega_m^3H_0^6\\
@@
 \begin{equation}
-    B_m(k_1, k_2, k_3, \chi) = 2 F_2(k_1, k_2, z) P_m(k_1, z) P_m(k_2, z) + \text{2 perm.}
-\end{equation}
-where \( P_m \) is the nonlinear matter power spectrum\footnote{Compare to the tree level bispectrum where we instead use the linear powerspectrum.}, and the kernel \( F_2 \) is modified from the tree level result with factors \( a(k, z) \), \( b(k, z) \), and \( c(k, z) \):
+    B^\delta(k_1, k_2, k_3, \chi) = 2 F_2(k_1, k_2, z) C^\delta(k_1, z) C^\delta(k_2, z) + \text{2 perm.}
+\end{equation}
+where \( C^\delta \) is the nonlinear matter power spectrum\footnote{Compare to the tree level bispectrum where we instead use the linear powerspectrum.}, and the kernel \( F_2 \) is modified from the tree level result with factors \( a(k, z) \), \( b(k, z) \), and \( c(k, z) \):
 \begin{equation}
@@
 \begin{equation}
-    n_{\text{eff}} \equiv \frac{d \ln P_m^{\text{lin}}(k)}{d \ln k}.
+    n_{\text{eff}} \equiv \frac{d \ln C^\delta_{\text{lin}}(k)}{d \ln k}.
 \end{equation}
@@
 \begin{equation}
-    4 \pi k_{\text{NL}}^3 P_m^{\text{lin}}(k_{\text{NL}}, 0) = 1.
+    4 \pi k_{\text{NL}}^3 C^\delta_{\text{lin}}(k_{\text{NL}}, 0) = 1.
 \end{equation}
@@
 $\nabla_{\hat{\mathbf n}}$ is the derivative along the axes orthogonal to the line of sight.
-The above equation can be written in terms of the lensing potential, $\psi$, as $\mathbf d (\hat{\mathbf n}) = \nabla_{\hat{\mathbf n}}$, with
+The above equation can be written in terms of the lensing potential, $\psi$, as $\mathbf d (\hat{\mathbf n}) = \nabla_{\hat{\mathbf n}} \psi(\hat{\mathbf n})$, with
 \begin{equation}
@@
 
-If the source is instead distributed radial distances according to some distribution function $p(\chi)$, with $p(\chi)$ normalized to integrate to 1, the $(\chi_* - \chi)/(\chi_*\chi)$ factor is changed as
+If the source is instead distributed over radial distance according to some distribution function $p(\chi)$, with $p(\chi)$ normalized to integrate to 1, the $(\chi_* - \chi)/(\chi_*\chi)$ factor is changed as
 \begin{equation*}
@@
 \subsection{Convergence and Shear}
-Quantities we are interested in measuring depend on how the lensing vector changes with the observation angle. The magnification matrix is defined as
-\begin{equation}
-    A_{ij} := \delta_{ij} + \frac{\partial}{\partial\theta_i}\alpha_i(\hat{\mathbf n}).
-\end{equation}
-This matrix can then be decomposed in the following form, which immediately gives us definitions for the \textbf{convergence}, $\kappa$, \textbf{shear}, $\gamma_1$ and $\gamma_2$, and \textbf{rotation}, $\omega$:
-\begin{equation}
-    A_ij(\hat{\mathbf n}) = \begin{pmatrix}
+The magnification matrix is defined as
+\begin{equation}
+    A_{ij} := \delta_{ij} + \frac{\partial}{\partial n_j}d_i(\hat{\mathbf n}).
+\end{equation}
+This matrix can be decomposed in the following form, which immediately gives us definitions for the \textbf{convergence}, $\kappa$, \textbf{shear}, $\gamma_1$ and $\gamma_2$, and \textbf{rotation}, $\omega$:
+\begin{equation}
+    A_{ij}(\hat{\mathbf n}) = \begin{pmatrix}
         1 - \kappa - \gamma_1 & -\gamma_2 + \omega \\
@@
 \end{equation}
-At 1st order $A$ is a symetric matrix by definition and so $\omega$ has to be 0, we will ignore it from here on out.
-Intuitively, A tells you how a small patch in the sky transforms due to weak lensing, i.e.
-if you have surface brightness $I$ then
-% TODO: make delta thick
-\begin{equation}
-    I_{\text{emitted}}(\hat{\mathbf{n}} + \mathbf{\delta}) = I_{\text{observed}}((\hat{\mathbf{n}} + \mathbf{\delta}) + \mathbf\alpha(\hat{\mathbf{n}} + \mathbf{\delta})) \approx I_{\text{observed}}((\hat{\mathbf{n}} + \mathbf{\alpha} + A_{ij}\delta_{j}).
-\end{equation}
-given that $|A_{ij}| = (1 - \kappa)^2 + \omega^2 - |\gamma|^2 = 1 - 2\kappa + O(\kappa^2, \gamma^2, \omega^2)$, we can interpret $\kappa$ as telling us about the overal magnification of the source.
+At 1st order $A$ is a symetric matrix by definition and $\omega$ vanishes, we will ignore it from here on out.
+Intuitively, $A$ tells you how a small patch in the sky transforms due to lensing.
+If we change the unlensed direction of a light source by $\delta\hat{\mathbf n}$, then the corresponding change in direction in the lensed image can be calculated as
+\begin{equation}
+\hat{\mathbf{n}} + \delta \hat{\mathbf{n}} \rightarrow \hat{\mathbf{n}} + \delta \hat{\mathbf{n}} + \mathbf d (\hat{\mathbf{n}} + \delta \hat{\mathbf{n}}) = \hat{\mathbf{n}} + \mathbf d (\hat{\mathbf n}) + \delta \hat{\mathbf n} + A_{ij}(\delta \hat{\mathbf{n}})_j.
+\end{equation}
+For an image on the sky, $A_{ij}$ thus introduces distortion.
+Note that $|A_{ij}| = (1 - \kappa)^2 + \omega^2 - |\gamma|^2 = 1 - 2\kappa + O(\kappa^2, \gamma^2, \omega^2)$.
+We can thus interpret $\kappa$ as telling us about the overal magnification of the source.
 The $\gamma_i$ represent the area preserving distortion, i.e.
@@
 \begin{equation}
-    \kappa = \frac{1}{2}\nabla^2\psi, \quad \gamma_1 = \frac{1}{2}(\partial_\theta^2 - \partial_\phi^2)\psi, \quad \gamma_2 = \partial_\theta\partial_\phi \psi.
+    \kappa = \frac{1}{2}\nabla^2\psi, \quad \gamma_1 = \frac{1}{2}(\partial_{n_1}^2 - \partial_{n_2}^2)\psi, \quad \gamma_2 = \partial_{n_1}\partial_{n_2} \psi.
 \end{equation}
@@
 \subsection{Lensing Potential Powerspectrum}
-Consider the lensing potential, $\psi$, then decomposing into spherical harmonics gives
+The lensing potential can be decomposed into spherical harmonics as
 \begin{equation}
@@
 \end{equation}
-On the other hand, consider the decomposition of $\Psi$ in fourier modes with the Fourier convention $\Psi(\mathbf x, \eta)=\int\frac{\d^3 \mathbf k}{(2\pi)^{3}}\Psi(\mathbf k, \eta)e^{i\mathbf k \cdot\mathbf x}$,
+On the other hand, consider the decomposition of $\Psi$ in Fourier modes with the Fourier convention $\Psi(\mathbf x, \eta)=\int\frac{\d^3 \mathbf k}{(2\pi)^{3}}\Psi(\mathbf k, \eta)e^{i\mathbf k \cdot\mathbf x}$,
 \begin{equation}
@@
 \end{equation}
-where we used that $\eta = \eta_0 - \chi$ along the unperturbed photon path (Born approximation), with $\eta_0$ the time at which the light ray hits earth. 
+where we used that $\eta = \eta_0 - \chi$ along the unperturbed photon path (this is known as the Born approximation), with $\eta_0$ the time at which the light ray hits earth. 
 We can use the result
@@
 
-To further evaluate the integral we will make the standard \textbf{Limber approximation}.
+To further evaluate the integral we will make the Limber approximation.
 The Bessel functions peak sharply at $l=x$\footnote{Some sources use $x\approx l+1/2$ instead, which is slightly more accurate for larger scales (low $l$) and slightly less accurate for smaller scales.}, with the peak being increasingly sharp for higher $l$.
@@
 Also note that in this definition we have immediately generalized to include cross correlation between different fields $X_1$, $X_2$, $X_3$.
-This will be relevant later when we look at cross correlations between shear and convergence fields.
-
+This is relevant when we look at cross correlations between CMB and galaxy lensing.
+6758
 Using the above definition and the symmetry property
@@
         m_1 & m_2 & m_3
-      \end{pmatrix}
+      \end{pmatrix},
 \end{equation*}
@@
 % TODO: fix references and citations 
-We can rewrite equations \ref{eq:lpsintermsofpsi} and \ref{eq:lbsintermsofpsi} in terms of the matter spectra instead of the graviational potential spectra using the poisson equation.
-This then allows us to numerically evaluate these objects using CAMB.
-The \textbf{density contrast} is defined as
+We can rewrite equations \ref{eq:lpsintermsofpsi} and \ref{eq:lbsintermsofpsi} in terms of the matter spectra instead of the $\psi$ spectra using the poisson equation.
+This allows us to numerically evaluate these lensing spectra using CAMB.
+The density contrast is defined as
 \begin{equation}
@@
 \end{gather*}
-The mean matter density of the universe, \( \bar{\rho} \), can be expressed in terms of the matter density parameter \( \Omega_m \) and the critical density \( \rho_c \) as 
+The mean matter density of the universe, \( \bar{\rho} \) is given as 
 $$
@@
 where $a(\eta)$ is the only time dependent factor on the right hand side.
-In the case of a flat universe ($K=0$) we can thus rewrite the Poisson equation \cite{dodelson2020modern} as
+The poisson equation relates $\Psi$ to the density contrast as \cite{dodelson2020modern}
 % in modern cosm eq 6.41 they say - nabla^2 Phi = ..., however here we use Psi instead and Psi = - Phi in our regime so hence why we don't take the - sign
 \begin{equation}
-    \nabla^2\Psi = 4\pi G a^2\br{\frac{3\Omega_m H_0^2}{8\pi G}\frac{1}{a^3}}\delta = \frac{3\Omega_m H_0^2}{2} \frac{1}{a}\delta \implies \Psi(k, \eta) = -\frac{3\Omega_m H_0^2}{2}\frac{1}{a}\frac{\delta(k, \eta)}{k^2},
+    \nabla^2\Psi(\mathbf x) = 4\pi G a^2\br{\frac{3\Omega_m H_0^2}{8\pi G}\frac{1}{a^3}}\delta(\mathbf x) = \frac{3\Omega_m H_0^2}{2} \frac{1}{a}\delta(\mathbf x) \implies \Psi(k, \eta) = -\frac{3\Omega_m H_0^2}{2}\frac{1}{a}\frac{\delta(k, \eta)}{k^2},
 \end{equation}
@@
 \begin{gather*}
-    \langle \Psi(\mathbf k, \eta)\Psi^*(\mathbf k',\eta)\rangle = \frac{2\pi^2}{k^3}P^\Psi(k, \eta)\delta(\mathbf k - \mathbf k') \implies P^\Psi(k,\eta) = \frac{1}{k}(9\Omega_m^2H_0^4\pi) \frac{1}{a^2}P^\delta(k,\eta),\\
+    \langle \Psi(\mathbf k, \eta)\Psi^*(\mathbf k',\eta)\rangle = \frac{2\pi^2}{k^3}C^\Psi(k, \eta)\delta(\mathbf k - \mathbf k') \implies C^\Psi(k,\eta) = \frac{1}{k}(9\Omega_m^2H_0^4\pi) \frac{1}{a^2}C^\delta(k,\eta),\\
     \langle \Psi(\mathbf k_1, \eta)\Psi(\mathbf k_2,\eta)\Psi(\mathbf k_3,\eta)\rangle = -(2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(k_1,k_2,k_3, \eta)\\ 

### 2025-04-22 code/Fisher_matrix_approx.py
fish mat changes

@@ -26,19 +26,18 @@ def fisher_calc_wrapper(args, tracers):
 num_pars = len(pars)
 
 for tracer in ['c', 's', 'both']:
-    if tracer == 'both': # !!!
-        mat = np.zeros((num_pars, num_pars))
-        counter = 1
-        for i in range(num_pars):
-            for j in range(i, num_pars):
-                result = fisher_calc_wrapper((i, j), tracer)
-                mat[i, j] = result
-                mat[j, i] = result  # Symmetric assignment
-                print(f'{tracer}:', counter, '/', num_pars * (num_pars + 1) / 2)
-                counter += 1
-        
-        np.savetxt(f'fisher_matrices/fish_mat_bisp_approx_{tracer}.txt', mat)
-        print(mat)
+    mat = np.zeros((num_pars, num_pars))
+    counter = 1
+    for i in range(num_pars):
+        for j in range(i, num_pars):
+            result = fisher_calc_wrapper((i, j), tracer)
+            mat[i, j] = result
+            mat[j, i] = result  # Symmetric assignment
+            print(f'{tracer}:', counter, '/', num_pars * (num_pars + 1) / 2)
+            counter += 1
+    
+    np.savetxt(f'fisher_matrices/fish_mat_bisp_approx_{tracer}_5p.txt', mat)
+    print(mat)
 
 
 # for python implementation

### 2025-04-22 code/bisp_compare_neutrinos.py

@@ -1,0 +1,25 @@
+import cosm_setup as cs
+
+
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+ls = np.arange(10, 2001, 20)
+
+for mnu in (0, 0.75*0.06, 0.9*0.06, 0.95 * 0.06, 0.06, 1.05*0.06, 1.1*0.06, 1.25*0.06, 0.2):
+    spectra = cs.lensing_spectra(mnu = mnu)
+
+    lbs_eq_ex_nonl = np.array([
+        spectra.lbs_flat(l, l, l, ('c', 'c', 'c')) * float(l)**10 / (2 * np.pi)**2 / 8 for l in ls
+    ])
+
+    lbs_fold_ex_nonl = np.array([
+        spectra.lbs_flat(l, l/2, l/2, ('c', 'c', 'c')) * float(l)**10 / 2**4 / (2 * np.pi)**2 / 8 for l in ls
+    ])
+
+    plt.loglog(ls, lbs_eq_ex_nonl, label=f'eq, mnu={mnu}')
+    plt.loglog(ls, lbs_fold_ex_nonl, linestyle = 'dotted', label=f'fold, mnu={mnu}')
+
+plt.legend()
+plt.savefig('/home3/p319950/ResearchProject/fisher_calc_weak_lensing/code/toshiya_bisp_curves/plot.png', dpi = 300)

### 2025-04-22 code/cosm_setup.py

@@ -92,6 +92,8 @@ class lensing_spectra:
     pars.set_dark_energy(w=self.w0)
     self.pars = pars
 
+    kmax_interp = 2000
+
     # power spectrum in units of Mpc^3, k in units of Mpc^-1, with input (z, k) on RHS 
     print('Creating linear mps')
     linear_mps_swapped = camb.get_matter_power_interpolator(
@@ -101,7 +103,7 @@ class lensing_spectra:
       k_hunit=False,
       zmin = 0,
       zmax = 1100,
-      kmax = 10000,
+      kmax = kmax_interp,
       extrap_kmax=True).P
     
     print('Creating mps')
@@ -112,7 +114,7 @@ class lensing_spectra:
       k_hunit=False,
       zmin = 0,
       zmax = 1100,
-      kmax = 10000,
+      kmax = kmax_interp,
       extrap_kmax=True).P
     
     self.mps = lambda k, z : mps_swapped(z, k)

### 2025-04-22 code/data_exporter.py

@@ -40,7 +40,7 @@ def data_export(folder_name, cosm_par, lps = True, a_create = True, b_create = T
     # 'rough' version has an interpolation error of around 4 percent and is mainly for testing stuff
     if rough:
         exp_par = exp_par_rough
-        filepath = '/scratch/p319950/data_rough/' + folder_name
+        filepath = '/scratch/p319950/data_rough_5p/' + folder_name
 
     # 'fine' version has an error of around 1 percent and is for final calculations
     if not rough:
@@ -178,7 +178,7 @@ par_names = ('H', 'ombh2', 'omch2', 'ns', 'As', 'mnu', 'w0')
 fiducial_cosm_par = (67.4, 0.0223, 0.119, 0.965, 2.13e-9, 0.06, -1)
 num_pars = len(par_names)
 
-dx = 0.025 # if you change this here it also needs to be changed at end of data_importer.pyx !!!
+dx = 0.05 # if you change this here it also needs to be changed at end of data_importer.pyx !!!
 
 dx_coeffs = [2, 1, -1, -2]
 
@@ -197,9 +197,9 @@ def dx_coeffs_to_str(coeff):
 # data_export(folder_name, cosm_par, lps = True, a_create = True, b_create = True, c_create = True, mps = True, rest = True, exp_par = exp_par)
 
 # for 1 + 8 x 4 x 3 = 97 core config
-# which_to_create = [[True, False, False, False, True, True], [False, True, False, False, False, False], [False, False, True, True, False, False]]
+which_to_create = [[True, False, False, False, True, True], [False, True, False, False, False, False], [False, False, True, True, False, False]]
 # which_to_create = [[False, True, False, False, False, False], [False, False, True, False, False, False], [False, False, False, True, False, False]]
-which_to_create = [[True, False, False, False, True, True]]
+#which_to_create = [[True, False, False, False, True, True]]
 # which_to_create = [[False, False, False, True, False, False]]
 
 # for 1 + 8 x 4 x 1 = 33 core config

### 2025-04-22 code/data_importer.c

@@ -3873,9 +3873,9 @@ static const char __pyx_k_lensing_power_spectrum_cc[] = "/lensing_power_spectrum
 static const char __pyx_k_lensing_power_spectrum_cs[] = "/lensing_power_spectrum_cs";
 static const char __pyx_k_lensing_power_spectrum_ss[] = "/lensing_power_spectrum_ss";
 static const char __pyx_k_Index_out_of_bounds_axis_d[] = "Index out of bounds (axis %d)";
-static const char __pyx_k_scratch_p319950_data_rough[] = "/scratch/p319950/data_rough/";
 static const char __pyx_k_Step_may_not_be_zero_axis_d[] = "Step may not be zero (axis %d)";
 static const char __pyx_k_itemsize_0_for_cython_array[] = "itemsize <= 0 for cython.array";
+static const char __pyx_k_scratch_p319950_data_rough_5p[] = "/scratch/p319950/data_rough_5p/";
 static const char __pyx_k_unable_to_allocate_array_data[] = "unable to allocate array data.";
 static const char __pyx_k_strided_and_direct_or_indirect[] = "<strided and direct or indirect>";
 static const char __pyx_k_All_dimensions_preceding_dimensi[] = "All dimensions preceding dimension %d must be indexed and not sliced";
@@ -4333,7 +4333,7 @@ typedef struct {
   PyObject *__pyx_kp_u_scale_factor;
   PyObject *__pyx_n_s_scale_factor_data;
   PyObject *__pyx_n_s_scale_factor_test;
-  PyObject *__pyx_kp_u_scratch_p319950_data_rough;
+  PyObject *__pyx_kp_u_scratch_p319950_data_rough_5p;
   PyObject *__pyx_n_s_setstate;
   PyObject *__pyx_n_s_setstate_cython;
   PyObject *__pyx_n_s_shape;
@@ -4849,7 +4849,7 @@ static int __pyx_m_clear(PyObject *m) {
   Py_CLEAR(clear_module_state->__pyx_kp_u_scale_factor);
   Py_CLEAR(clear_module_state->__pyx_n_s_scale_factor_data);
   Py_CLEAR(clear_module_state->__pyx_n_s_scale_factor_test);
[…]

### 2025-04-22 code/data_importer.pyx

@@ -37,12 +37,12 @@ cdef int conv_noise_type = 2
 # type values correspond to:
 # 0: S0 noise curves (old, remove)
 # 1: sigma = 1, Delta_P = 6 (stage 3 toshiya)
-# 2: sigma = 3, Delta_P = 1 (stage 4 toshiya)
+# 2: sigma = 3, Delta_P = 1 (stage 4 toshiya)§
 
 # (units are in microKelvin arcmin and arcmin)
 
 # SO noise
-cdef str folder_file_path = '/scratch/p319950/data_rough/'
+cdef str folder_file_path = '/scratch/p319950/data_rough_5p/'
 cdef str filepath_convergence_noise_file_path = folder_file_path + 'conv_noise.dat'
 conv_noise_data_array = np.loadtxt(filepath_convergence_noise_file_path)
 cdef double[:, :] conv_noise_data = conv_noise_data_array
@@ -1217,7 +1217,7 @@ cpdef double lps_w02m(int l, char* type1, char* type2) noexcept nogil:
 cdef double der_2o(double f2p, double f1p, double f1m, double f2m, double dx) noexcept nogil:
     return (-1 * f2p + 8 * f1p - 8 * f1m + f2m) / (12 * dx)
 
-cdef double dx_frac = 0.025
+cdef double dx_frac = 0.05
 cdef double[:] fiducial_cosm_par = np.array([67.4, 0.0224, 0.120, 0.965, 2.1e-9, 0.06])
 
 

### 2025-04-24 paper/paper_draft.tex
paper changes

@@
 We can relate $\kappa$ and $\gamma$ directly to the lensing potential as
-% TODO: should these subscripts be theta and phi?
 \begin{equation}
@@
 \end{equation}
+
+It is shown in appendix \ref{sec:shear} that
+\begin{equation}
+    \gamma := \gamma_1 + i\gamma_2 = \frac{1}{2}\eth_1(\eth_0\psi)
+\end{equation}
+where the spin raising operator, $\eth_s$ acts on a spin $s$ function defined on $S^2$ to create a spin $s+1$ function.
+$\eth_s$ can be written in spherical coordinates $(\theta, \phi)$ as\footnote{We use the physics convention for the definition of $\theta$ and $\phi$ here.}
+\begin{equation}
+    \eth_s = -\sin^s\theta(\partial_\theta + \frac{i}{\sin\theta})\frac{1}{\sin^s\theta}.
+\end{equation}
+In this context a spin $s$ function referes to a function $_sf(\theta, \phi)$ that transforms under any rotation of coordinates by picking up a phase factor $e^{is\alpha}$, with $\alpha$ the angle of the rotation, i.e.
+\begin{equation}
+    f'(\theta', \phi') = e^{is\alpha}f(\theta, \phi).
+\end{equation}
+Shear is thus a spin 2 function, which can be checked by noting that rotating a galaxy image stretched and squeezed through weak lensing by $180$ degrees gives the same stretching and squeezing, i.e.
+the same shear.
+
+% The spin weighted spherical harmonics, $_sY_l^m$ are eigenfunctions of the raising/lowering operator, in particular
+% \begin{equation}
+%     \eth_sY_l^m = \sqrt{(l-s)(l+s+1)} _{s+1}Y_l^m.
+% \end{equation}
+The spherical harmonics are eigenfunctions of $\nabla^2$ and the spin raising and lowering operators. Using this property, the corresponding relations in spherical harmonic space can be shown to be 
+\begin{equation*}
+    \kappa_{lm} = \frac{l(l+1)}{2}\psi_{lm}, \quad \gamma_{lm} = \frac{\sqrt{(l-1)l(l+1)(l+2)}}{2} \psi_{lm}.
+\end{equation*}
 
@@
 \end{empheq}
-
-% \begin{gather*}
-%     \langle \psi_{l_1m_1} \psi_{l_2m_2} \psi_{l_3m_3} \rangle = A_{\mathbf{l}}^{\mathbf{m}} (-1)^{l_1 + l_2 + l_3} 8 (2\pi)^{9/2} \int_0^{\chi_*} W(\chi, \chi_*)^3 \frac{\chi^2}{l_1^2l_2^2l_3^2} \br{\frac{3\Omega_m H^2}{2}}^3|_{\eta=\eta_0-\chi}\\
-%     \quad \times B^\delta\br{ \frac{l_1}{\chi}, \frac{l_2}{\chi}, \frac{l_3}{\chi}; \eta_0-\chi } \mathrm{d}\chi, \\
-%     C_l^\psi = \frac{72\pi^3}{l^4}\int_0^{\chi_*}\d \chi \chi^2(\Omega_m^2H^4)|_{\eta=\eta_0-\chi} C^\delta_{l/\chi}(\eta_0-\chi) W(\chi)^2.
-% \end{gather*}
-
-\subsection{Convergence and shear multipole moments}
-\paragraph{Convergence}
-% TODO: make notation of nabla nicer/more consistent
-
-Taking the harmonic decomposition of the convergence,
-\begin{equation}
-    \kappa(\hat{\mathbf n}) = \kappa_{lm}Y_l^m(\hat{\mathbf n}),
-\end{equation}
-you can relate it to the multipole moments of the lensing potential through
-\begin{equation}
-    \kappa_{lm} = \frac{1}{2}(\nabla^2\psi)_{lm} = \frac{l(l+1)}{2}\psi_{lm}
-\end{equation}
-because the spherical harmonics are eigenfunctions of the Laplacian operator.
-
-\paragraph{Shear}
-On the other hand, the shear is given by
-\begin{equation}
-    \gamma = \gamma_1 + i\gamma_2 = \frac{1}{2}\eth_1(\eth_0\psi)
-\end{equation}
-where the spin raising operator, $\eth_s$ acts on a spin $s$ function defined on $S^2$ to create a spin $s+1$ function.
-The above equality is proven explicitely in appendix ...
-%TODO: write and cite appendix proving equivalence
-$\eth_s$ can be written in $(\theta, \phi)$ coordinates as
-\begin{equation}
-    \eth_s = -\sin^s\theta(\partial_\theta + \frac{i}{\sin\theta})\frac{1}{\sin^s\theta}.
-\end{equation}
-It is worth explaining that in this context a spin $s$ function referes to a function $_sf(\theta, \phi)$ that transforms under any rotation of coordinates by picking up a phase factor $e^{is\alpha}$, with $\alpha$ the angle of the rotation, i.e.
-\begin{equation}
-    f'(\theta', \phi') = e^{is\alpha}f(\theta, \phi).
-\end{equation}
-Shear is actually a spin 2 function, which can immediately be seen by observing that rotating a galaxy image stretched and squeezed through weak lensing by $180$ degrees gives the same stretching and squeezing, i.e.
-the same shear.
-It is thus expected that the shear is proportional to $\eth^2\psi$.
-
-The spin weighted spherical harmonics, $_sY_l^m$ are eigenfunctions of the raising/lowering operator, in particular
-\begin{equation}
-    \eth_sY_l^m = \sqrt{(l-s)(l+s+1)} _{s+1}Y_l^m.
-\end{equation}
-Decomposing the shear as
-\begin{equation}
-    \gamma(\theta, \phi) = \gamma_{lm} {}_sY_l^m(\theta,\phi),
-\end{equation}
-we thus find
-\begin{equation}
-    \gamma_{lm} = \frac{1}{2}\sqrt{(l-1)(l)(l+1)(l+2)} \psi_{lm}
-\end{equation}
 

### 2025-04-25 paper/paper_draft.tex
paper changes

@@
 
-Fisher matrix formalism is used to find a lower bound on the constraints we can place on experimental parameters.
+The Fisher matrix formalism is used to find a lower bound on the constraints we can place on experimental parameters.
 It combines the Cramer-Rao Inequality \cite{casella2002statistical} with the assumption that we have unbiased estimators  following a gaussian distribution \cite{dodelson2020modern}.
-In particular, denoting the Fisher matrix by $F$ and the parameters as $\theta_i$, it can be shown that
-\begin{equation}
-    Var_{\theta_i}(\hat\theta_i) \geq \br{F^{-1}}_{\theta_i\theta_j}.
-\end{equation}
-In the case of $n$ potentially different random variables $x_i$, each with associated mean $\mu_{x_i}$ and error $\sigma_{x_i}$, the Fisher matrix is given as
-\begin{equation}
-    F_{\theta_i\theta_j} := \sum_{p, q=1}^n \frac{\partial\mu_{x_p}}{\partial\theta_i}(\tilde\theta_k)(\text{Cov}^{-1})_{x_px_q}\frac{\partial\mu_{x_q}}{\partial\theta_j}(\tilde\theta_k),
+In particular, denoting the Fisher matrix by $F_{\theta_i\theta_j}$, the parameters as $\theta_i$, and their estimators as $\hat\theta_i$, it can be shown that
+\begin{equation}
+    \text{Var}(\hat\theta_i) \geq \br{F^{-1}}_{\theta_i\theta_i}.
+\end{equation}
+In the case of $n$ measurements whose outcomes are realizations of random variables $x_i$, each with associated mean $\mu_{x_i}$, the Fisher matrix is given as
+\begin{equation}
+    F_{\theta_i\theta_j} := \sum_{p, q=1}^n \frac{\partial\mu_{x_p}}{\partial\theta_i}(\tilde\theta_k)(\text{Cov}^{-1})_{x_px_q}(\tilde \theta_k)\frac{\partial\mu_{x_q}}{\partial\theta_j}(\tilde\theta_k),
 \end{equation}
 where Cov is the covariance matrix associated with the random vector $(x_1, ..., x_n)$, $\text{Cov}_{x_px_q}:=\text{Cov}(x_p,x_q)$.
+The derivative of the mean measurement outcomes and measurement covariances will in general depend on the true, unknown, values of the experimental parameters.
+We therefore evaluate these quantities with for our best guess of the experimental parameters given some outside information.
+These are known as the ``fiducial'' values.
 
@@
 
-\subsection{Fisher matrices for power- bispectra with multiple tracers}
-When considering power spectra, the definition of the Fisher matrix immediately gives 
+\subsection{Fisher matrices for power- and bispectra with multiple tracers}
+For power spectra, the definition of the Fisher matrix gives 
 \begin{equation*}
@@
 \end{equation*}
-The estimator of a power or bispectrum is just the product of the estimators of the appropriate tracers, $\hat C^{XY}_{l} = \hat X(l) \hat Y(l)$.
+The estimator of a power or bispectrum is the product of the estimators of the appropriate tracers, e.g.
+$\hat C^{XY}_{l} = \hat X(l) \hat Y(l)$.
 The sum over $[XY]$ and $[X'Y']$ denotes a sum over possible tracer combinations \textit{without} counting permutations of tracer configuration.
@@
 
-To evaluate the covariance matrix we again assume that the estimators are gaussian distributed so that we can do apply a wick contraction, as is commonly done \cite{tegmark1997cosmic}.
+To evaluate the covariance matrix we again assume that the estimators are gaussian distributed so that we can apply a wick contraction, as is commonly done \cite{tegmark1997cosmic}.
 In this case we get
 \begin{equation*}
-    \Cov^{XY, X'Y'}_{l, l'} = \delta_{ll'}\br{C^{XX'}_lC^{YY'}_l + C^{XY'}_lC^{YX'}_l}.
-\end{equation*}
-Under the Gaussian approximation the gaussian matrix thus vanishes except for $3\times 3$ block matrices (in the case of 2 tracers) on the diagonal. The fisher matrix is then
+    \Cov^{XY, X'Y'}_{l, l'} = \frac{1}{2l+1}\delta_{ll'}\br{\tilde C^{XX'}_l \tilde C^{YY'}_l + \tilde C^{XY'}_l \tilde C^{YX'}_l}.
+\end{equation*}
+Two remarks are in order.
+\begin{enumerate}
+    \item By definition
+    \begin{equation*}
+        \langle X_{lm}X'_{l'm'} \rangle = (2\pi)^3 \delta_{ll'}\delta_{mm'}C_l^{XX'},
+    \end{equation*}
+    so you can average over the measurements done for different values of $m$, i.e.
+$X_{l(-l)}$, $X_{l(-l+1)}$, $\cdots$, $X_{l(l-1)}$, $X_{ll}$.
+This results in the $(2l+1)^{-1}$ factor in the power spectrum covariance.
+    \item The tilde is used to denote the powerspectrum as calculated earlier plus the noise power spectrum, $N^{XX'}_l$, which is the power spectrum of the noise associated with the estimator of the field.
+This is where experimental noise is incorporated into the calculation.
+\end{enumerate}
+Under the Gaussian approximation the covariance matrix vanishes except for $3\times 3$ block matrices (in the case of 2 tracers) on the diagonal. The fisher matrix is then
 \begin{equation*}
@@
 The rule is that $B^{XYZ}_{l_1l_2l_3}$ and $B^{X'Y'Z'}_{l_1'l_2'l_3'}$ are not distinct signals if there exists a permutation $\sigma$ that simultaniously maps $X'Y'Z'$ to $XYZ$ and $l_1'l_2'l_3'$ to $l_1l_2l_3$.
-It turns out that we can write a sum over distinct signals out as
+It turns out that we can write a sum over distinct signals explicitly as
 \begin{gather*}
@@
 The $Z$ value does not matter because any $Z$ value is accounted for.
-For sum 3 we can argue that we can switch around the order of the $l$'s to satisfy $l_1 < l2 < l3$ and the corresponding $XYZ$ configuration will be accounted for in sum 3 because all $XYZ$ combinations are accounted.
+For sum 3 we can argue that we can switch around the order of the $l$'s to satisfy $l_1 < l2 < l3$ and the corresponding $XYZ$ configuration will be accounted for in sum 3 because all $XYZ$ combinations are counted.
 Finally, it is simple to verify that no distinct signal is counted more than once within each sum.
 
-\paragraph{Covariance matrix}
+To calculate the elements of the covariance matrix consider the following.
 Every element of the Fisher matrix can be seen as an inner product weighted by the inverse covariance matrix.
@@
 Within each $l$ configuration we can choose any ordering for the $XYZ$ configurations.
-The covariance matrix now becomes a block matrix with each block corresponding to a $l_i$ and $l_i'$ configuration.
-When wick contracting using the gaussian approximation, every block matrix where $l_1l_2l_3$ is not a permutation of $l_1'l_2'l_3'$ vanishes.
-The entries of each block matrix are given as
-\begin{gather*}
-    \br{\Cov_{l_1l_2l_3}}^{XYZ,X'Y'Z'} = C^{XX'}_{l_1}C^{YY'}_{l_2}C^{ZZ'}_{l_3} + \delta_{l_1l_2}C^{XY'}_{l_1}C^{YX'}_{l_2}C^{ZZ'}_{l_3} +\delta_{l_2l_3}C^{XX'}_{l_1}C^{YZ'}_{l_2}C^{ZX'}_{l_3} \\ + \delta_{l_3l_1}C^{XZ'}_{l_1}C^{YY'}_{l_2}C^{ZX'}_{l_3} +\delta_{l_1l_2}\delta_{l_2l_3}\br{C^{XY'}_{l_1}C^{YZ'}_{l_2}C^{ZX'}_{l_3} +C^{XZ'}_{l_1}C^{YX'}_{l_2}C^{ZY'}_{l_3}}.
+The covariance matrix now becomes a block matrix with each block corresponding to an $l_i$ and $l_i'$ configuration.
+When wick contracting using the gaussian approximation, every block matrix where $(l_1,l_2,l_3)\neq (l_1',l_2',l_3')$ vanishes.
+It can then be shown that the entries of each block matrix are given as
+\begin{gather*}
+    \br{\Cov_{l_1l_2l_3}}^{XYZ,X'Y'Z'} = \tilde C^{XX'}_{l_1}\tilde C^{YY'}_{l_2}\tilde C^{ZZ'}_{l_3} + \delta_{l_1l_2}\tilde C^{XY'}_{l_1}\tilde C^{YX'}_{l_2}\tilde C^{ZZ'}_{l_3} +\delta_{l_2l_3}\tilde C^{XX'}_{l_1}\tilde C^{YZ'}_{l_2}\tilde C^{ZX'}_{l_3} \\ + \delta_{l_3l_1}\tilde C^{XZ'}_{l_1}\tilde C^{YY'}_{l_2}\tilde C^{ZX'}_{l_3} +\delta_{l_1l_2}\delta_{l_2l_3}\br{\tilde C^{XY'}_{l_1}\tilde C^{YZ'}_{l_2}\tilde C^{ZX'}_{l_3} +\tilde C^{XZ'}_{l_1}\tilde C^{YX'}_{l_2}\tilde C^{ZY'}_{l_3}}.
 \end{gather*}
@@
 \partial_\alpha B^{X Y Z}_{l_1 l_2 l_3} 
-(C^{-1})^{X X'}_{l_1}
-(C^{-1})^{Y Y'}_{l_2}
-(C^{-1})^{Z Z'}_{l_3}
+(\tilde C^{-1})^{X X'}_{l_1}
+(\tilde C^{-1})^{Y Y'}_{l_2}
+(\tilde C^{-1})^{Z Z'}_{l_3}
 \partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3}
 \end{equation*}
-where
+where, in the case of two tracers, 
 \begin{equation*}
     C_{l} := \begin{pmatrix}
-        C^{\psi_\kappa\psi_\kappa}_l & C^{\psi_\kappa\psi_\gamma}_l \\
-        C^{\psi_\kappa\psi_\gamma}_l & C^{\psi_\gamma\psi_\gamma}_l
+        \tilde C^{\psi_1\psi_1}_l & \tilde C^{\psi_1\psi_2}_l \\
+        \tilde C^{\psi_1\psi_2}_l & \tilde C^{\psi_2\psi_2}_l
     \end{pmatrix}
@@
 p.
-X'Y'Z'}(C^{-1})^{X X'}_{l_1}
-    (C^{-1})^{Y Y'}_{l_2}
-    (C^{-1})^{Z Z'}_{l_3}}
+X'Y'Z'}(\tilde C^{-1})^{X X'}_{l_1}
+    (\tilde C^{-1})^{Y Y'}_{l_2}
+    (\tilde C^{-1})^{Z Z'}_{l_3}}
     \partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3} \\
-    &+ \sum_{l_1=l_2\neq l_3} \sum_{[XY]Z}\sum_{[X'Y']Z'} \partial_{\alpha} B^{XYZ}_{l_1 l_2 l_3}\br{\frac{\mathcal P _{l_1l_2l_3}}{6}\sum_{d. p. XY} \sum_{d. p. X'Y'}(C^{-1})^{X X'}_{l_1}
-    (C^{-1})^{Y Y'}_{l_2}
-    (C^{-1})^{Z Z'}_{l_3}}
+    &+ \sum_{l_1=l_2\neq l_3} \sum_{[XY]Z}\sum_{[X'Y']Z'} \partial_{\alpha} B^{XYZ}_{l_1 l_2 l_3}\br{\frac{\mathcal P _{l_1l_2l_3}}{6}\sum_{d. p. XY} \sum_{d. p. X'Y'}(\tilde C^{-1})^{X X'}_{l_1}
+    (\tilde C^{-1})^{Y Y'}_{l_2}
+    (\tilde C^{-1})^{Z Z'}_{l_3}}
     \partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3} \\
-    &+ \sum_{l_1<l_2<l_3} \sum_{XYZ}\sum_{X'Y'Z'} \partial_{\alpha} B^{XYZ}_{l_1 l_2 l_3}(C^{-1})^{X X'}_{l_1}
-    (C^{-1})^{Y Y'}_{l_2}
-    (C^{-1})^{Z Z'}_{l_3}
+    &+ \sum_{l_1<l_2<l_3} \sum_{XYZ}\sum_{X'Y'Z'} \partial_{\alpha} B^{XYZ}_{l_1 l_2 l_3}(\tilde C^{-1})^{X X'}_{l_1}
+    (\tilde C^{-1})^{Y Y'}_{l_2}
+    (\tilde C^{-1})^{Z Z'}_{l_3}
     \partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3},
@@
 \begin{gather*}
-    \sum_{[X'Y'Z']} \br{\frac{\mathcal P _{l_1l_2l_3}}{6} \sum_{d. p. XYZ} \sum_{d. p. X'Y'Z'}(C^{-1})^{X X'}_{l}
-    (C^{-1})^{Y Y'}_{l}
-    (C^{-1})^{Z Z'}_{l}} \br{C^{X'X''}_l C^{Y'Y''}_l C^{Z'Z''}_l + \text{perms } X''Y''Z''} \\
-    = \frac{\mathcal P _{l_1l_2l_3}}{6}\sbr{\br{\sum_{d. p. XYZ} \sum_{X'Y'Z'}(C^{-1})^{X X'}_{l}
-    (C^{-1})^{Y Y'}_{l}
-    (C^{-1})^{Z Z'}_{l}} C^{X'X''}_l C^{Y'Y''}_l C^{Z'Z''}_l} + \text{perms } X''Y''Z'' \\
+    \sum_{[X'Y'Z']} \br{\frac{\mathcal P _{l_1l_2l_3}}{6} \sum_{d. p. XYZ} \sum_{d. p. X'Y'Z'}(\tilde C^{-1})^{X X'}_{l}
+    (\tilde C^{-1})^{Y Y'}_{l}
+    (\tilde C^{-1})^{Z Z'}_{l}} \br{\tilde C^{X'X''}_l \tilde C^{Y'Y''}_l \tilde C^{Z'Z''}_l + \text{perms } X''Y''Z''} \\
+    = \frac{\mathcal P _{l_1l_2l_3}}{6}\sbr{\br{\sum_{d. p. XYZ} \sum_{X'Y'Z'}(\tilde C^{-1})^{X X'}_{l}
+    (\tilde C^{-1})^{Y Y'}_{l}
+    (\tilde C^{-1})^{Z Z'}_{l}} \tilde C^{X'X''}_l \tilde C^{Y'Y''}_l \tilde C^{Z'Z''}_l} + \text{perms } X''Y''Z'' \\
     = \frac{\mathcal P _{l_1l_2l_3}}{6} \sum_{d. p. XYZ} \delta_{XX''}\delta_{YY''}\delta_{ZZ''} + \text{perms } X''Y''Z''
@@
 \subsection{Signal to Noise Ratio (SNR)}
-To measure the overal detectability of the lensing spectra, we introduce an overal amplitude of our signal, $A$, with fiducial value 1 as experimental parameter and compute $F_{AA}$.
+To quantify the detectability of the lensing spectra, we introduce an overal amplitude of our signal, $A$, with fiducial value 1 as experimental parameter and compute $F_{AA}$.
 Obviously, $\partial_A\br{A B^{XYZ}_{l_1l_2l_3}}|_{A=1} = B^{XYZ}_{l_1l_2l_3}$, so we find
@@
 B^{X Y Z}_{l_1 l_2 l_3} 
-(C^{-1})^{X X'}_{l_1}
-(C^{-1})^{Y Y'}_{l_2}
-(C^{-1})^{Z Z'}_{l_3}
+(\tilde C^{-1})^{X X'}_{l_1}
+(\tilde C^{-1})^{Y Y'}_{l_2}
+(\tilde C^{-1})^{Z Z'}_{l_3}
 B^{X' Y' Z'}_{l_1 l_2 l_3}.
@@
 
-\subsection{Combined Fisher Matrix}
+\subsection{Fisher matrix of power- + bispectra}
 To compute the Fisher matrix of an experiment measuring both lensing power- and bispectra we are required to compute and invert the full covariance matrix.
-Under the Gaussian matrix this simplifies trivially.
-The correlation between the a power- and bispectrum estimator contains an odd (5) amount of fields and thus always vanishes.
+If we keep assuming that the measurements are close enough to Gaussian to be able to use wick contractions to a good approximation, the full covariance matrix simplifies trivially.
+The correlation between a power- and bispectrum estimator contains an odd (5) amount of fields and thus always vanishes.
 We are thus allowed to simply add the Fisher matrices of the power- and bispectra to compute the combined Fisher matrix.
 
-\subsection{Fisher Matrices and Eigenvalues}
-As explained earlier, inverting the Fisher matrix gives us a covariance matrix of the estimators, $\hat \theta_i$ of our experimental parameters. For any vector $\mathbf v$,
-\begin{equation*}
-    \mathbf{v}^T_j \Cov_{ij} \mathbf v_i = \langle (v_i \hat \theta_i)^2 \rangle.
-\end{equation*}
-In particular, if the covariance matrix thus has some large eigenvalue $\lambda_L$ (large compared to the overal accuracy of our experiment), with eigenvector $v_i$, then the corresponding estimator, $v_i\hat \theta_i$, has a large variance and thus represents an approximate degeneracy in our parameters.
+% \subsection{Fisher Matrices and Eigenvalues}
+% As explained earlier, inverting the Fisher matrix gives us a covariance matrix of the estimators, $\hat \theta_i$ of our experimental parameters. For any vector $\mathbf v$,
+% \begin{equation*}
+%     \mathbf{v}^T_j \Cov_{ij} \mathbf v_i = \langle (v_i \hat \theta_i)^2 \rangle.
+% \end{equation*}
+% In particular, if the covariance matrix thus has some large eigenvalue $\lambda_L$ (large compared to the overal accuracy of our experiment), with eigenvector $v_i$, then the corresponding estimator, $v_i\hat \theta_i$, has a large variance and thus represents an approximate degeneracy in our parameters.
 Clearly, when looking at the Fisher matrix we thus instead look for the smallest eigenvalue.
 
-If we have such an abnormally small eigenvalue we are likely to run into problems with calculating and inverting the Fisher matrix due to numerical errors.
+% If we have such an abnormally small eigenvalue we are likely to run into problems with calculating and inverting the Fisher matrix due to numerical errors.
 These problems are that inversion becomes unstable and that the matrix as a whole might not be PSD because the small eigenvalue was (wrongly) calculated to be negative.
@@
 Specifically, for a matrix $A$ with some perturbation $\Delta A$, the corresponding  perturbation of eigenvalue $\lambda_i$ is bounded above as
-\begin{equation*}
-    |\Delta \lambda_i | \leq ||\Delta A || \kappa(A)|,
-\end{equation*}
-with $\kappa(A)$ the condition number.
-
-Before inverting a Fisher matrix, it is thus recommended to remove rows and columns corresponding to parameters that are poorly constrained by the experiment, which can be determined by checking for small eigenvalues and their eigenvectors.
+% \begin{equation*}
+%     |\Delta \lambda_i | \leq ||\Delta A || \kappa(A)|,
+% \end{equation*}
+% with $\kappa(A)$ the condition number.
+
+% Before inverting a Fisher matrix, it is thus recommended to remove rows and columns corresponding to parameters that are poorly constrained by the experiment, which can be determined by checking for small eigenvalues and their eigenvectors.
 
@@
     =\frac{1}{2}(\partial_{\tilde y}^2 - \partial_{\tilde x}^2 + 2i\partial_{\tilde x}\partial_{\tilde y})\psi
-    =\gamma_1 + i\gamma_2 = \gamma,
-\end{gather*}
-as required.
+    =\gamma_1 + i\gamma_2 = \gamma.
+\end{gather*}
 

### 2025-04-29 paper/paper_draft.tex
paper changes

@@
 
-% \section{Fisher matrices raw data}
-
-% Bispectra Fisher matrices:\\
-% \begin{gather*}
-% F^B_{\psi_c} =
-% \begin{pmatrix}
-% 4\times 10^{-3} & 6\times 10^{-4} & -6\times 10^{-3} & -9\times 10^{-3} & 1\times 10^{-4} & -4\times 10^{-3} & -1\times 10^{-3}\\
-% 6\times 10^{-4} & 1\times 10^{-4} & -1\times 10^{-3} & -2\times 10^{-3} & 2\times 10^{-5} & -6\times 10^{-4} & -2\times 10^{-4}\\
-% -6\times 10^{-3} & -1\times 10^{-3} & 1\times 10^{-2} & 1\times 10^{-2} & -2\times 10^{-4} & 6\times 10^{-3} & 2\times 10^{-3}\\
-% -9\times 10^{-3} & -2\times 10^{-3} & 1\times 10^{-2} & 2\times 10^{-2} & -2\times 10^{-4} & 8\times 10^{-3} & 3\times 10^{-3}\\
-% 1\times 10^{-4} & 2\times 10^{-5} & -2\times 10^{-4} & -2\times 10^{-4} & 3\times 10^{-6} & -1\times 10^{-4} & -3\times 10^{-5}\\
-% -4\times 10^{-3} & -6\times 10^{-4} & 6\times 10^{-3} & 8\times 10^{-3} & -1\times 10^{-4} & 4\times 10^{-3} & 1\times 10^{-3}\\
-% -1\times 10^{-3} & -2\times 10^{-4} & 2\times 10^{-3} & 3\times 10^{-3} & -3\times 10^{-5} & 1\times 10^{-3} & 4\times 10^{-4}
-% \end{pmatrix},
-% \\
-% F^B_{\psi_s} =
-% \begin{pmatrix}
-% 3\times 10^{-1} & 4\times 10^{-2} & -3\times 10^{-1} & -1\times 10^{-1} & 5\times 10^{-3} & -1\times 10^{-1} & -1\times 10^{-1}\\
-% 4\times 10^{-2} & 4\times 10^{-3} & -3\times 10^{-2} & -2\times 10^{-2} & 5\times 10^{-4} & -1\times 10^{-2} & -1\times 10^{-2}\\
-% -3\times 10^{-1} & -3\times 10^{-2} & 2\times 10^{-1} & 1\times 10^{-1} & -3\times 10^{-3} & 1\times 10^{-1} & 1\times 10^{-1}\\
-% -1\times 10^{-1} & -2\times 10^{-2} & 1\times 10^{-1} & 7\times 10^{-2} & -2\times 10^{-3} & 6\times 10^{-2} & 5\times 10^{-2}\\
-% 5\times 10^{-3} & 5\times 10^{-4} & -3\times 10^{-3} & -2\times 10^{-3} & 6\times 10^{-5} & -2\times 10^{-3} & -2\times 10^{-3}\\
-% -1\times 10^{-1} & -1\times 10^{-2} & 1\times 10^{-1} & 6\times 10^{-2} & -2\times 10^{-3} & 6\times 10^{-2} & 5\times 10^{-2}\\
-% -1\times 10^{-1} & -1\times 10^{-2} & 1\times 10^{-1} & 5\times 10^{-2} & -2\times 10^{-3} & 5\times 10^{-2} & 5\times 10^{-2}
-% \end{pmatrix},
-% \\
-% F^B_{\psi_{s + c}} =
-% \begin{pmatrix}
-% 4\times 10^{-1} & 4\times 10^{-2} & -3\times 10^{-1} & -2\times 10^{-1} & 5\times 10^{-3} & -2\times 10^{-1} & -1\times 10^{-1}\\
-% 4\times 10^{-2} & 4\times 10^{-3} & -3\times 10^{-2} & -2\times 10^{-2} & 6\times 10^{-4} & -2\times 10^{-2} & -1\times 10^{-2}\\
-% -3\times 10^{-1} & -3\times 10^{-2} & 2\times 10^{-1} & 2\times 10^{-1} & -4\times 10^{-3} & 1\times 10^{-1} & 1\times 10^{-1}\\
-% -2\times 10^{-1} & -2\times 10^{-2} & 2\times 10^{-1} & 1\times 10^{-1} & -3\times 10^{-3} & 8\times 10^{-2} & 7\times 10^{-2}\\
-% 5\times 10^{-3} & 6\times 10^{-4} & -4\times 10^{-3} & -3\times 10^{-3} & 7\times 10^{-5} & -2\times 10^{-3} & -2\times 10^{-3}\\
-% -2\times 10^{-1} & -2\times 10^{-2} & 1\times 10^{-1} & 8\times 10^{-2} & -2\times 10^{-3} & 7\times 10^{-2} & 6\times 10^{-2}\\
-% -1\times 10^{-1} & -1\times 10^{-2} & 1\times 10^{-1} & 7\times 10^{-2} & -2\times 10^{-3} & 6\times 10^{-2} & 5\times 10^{-2}
-% \end{pmatrix}.
-% \end{gather*}
-
-% Powerspectra Fisher matrices:
-% \begin{gather*}
-% F^P_{\psi_c} =
-% \begin{pmatrix}
-% 2\times 10^{-2} & 4\times 10^{-3} & -4\times 10^{-2} & 8\times 10^{-3} & 9\times 10^{-4} & -3\times 10^{-2} & -1\times 10^{-2}\\
-% 4\times 10^{-3} & 1\times 10^{-3} & -9\times 10^{-3} & -6\times 10^{-4} & 2\times 10^{-4} & -5\times 10^{-3} & -2\times 10^{-3}\\
-% -4\times 10^{-2} & -9\times 10^{-3} & 8\times 10^{-2} & -9\times 10^{-3} & -2\times 10^{-3} & 6\times 10^{-2} & 2\times 10^{-2}\\
-% 8\times 10^{-3} & -6\times 10^{-4} & -9\times 10^{-3} & 2\times 10^{-2} & 3\times 10^{-4} & -1\times 10^{-2} & -4\times 10^{-3}\\
-% 9\times 10^{-4} & 2\times 10^{-4} & -2\times 10^{-3} & 3\times 10^{-4} & 3\times 10^{-5} & -1\times 10^{-3} & -4\times 10^{-4}\\
-% -3\times 10^{-2} & -5\times 10^{-3} & 6\times 10^{-2} & -1\times 10^{-2} & -1\times 10^{-3} & 4\times 10^{-2} & 2\times 10^{-2}\\
-% -1\times 10^{-2} & -2\times 10^{-3} & 2\times 10^{-2} & -4\times 10^{-3} & -4\times 10^{-4} & 2\times 10^{-2} & 6\times 10^{-3}
-% \end{pmatrix},
-% \\
-% F^P_{\psi_s} =
-% \begin{pmatrix}
-% 2\times 10^{-1} & 1\times 10^{-2} & -1\times 10^{-1} & -3\times 10^{-2} & 2\times 10^{-3} & -6\times 10^{-2} & -7\times 10^{-2}\\
-% 1\times 10^{-2} & 1\times 10^{-3} & -9\times 10^{-3} & -3\times 10^{-3} & 2\times 10^{-4} & -5\times 10^{-3} & -6\times 10^{-3}\\
-% -1\times 10^{-1} & -9\times 10^{-3} & 7\times 10^{-2} & 2\times 10^{-2} & -1\times 10^{-3} & 4\times 10^{-2} & 4\times 10^{-2}\\
-% -3\times 10^{-2} & -3\times 10^{-3} & 2\times 10^{-2} & 9\times 10^{-3} & -4\times 10^{-4} & 1\times 10^{-2} & 1\times 10^{-2}\\
-% 2\times 10^{-3} & 2\times 10^{-4} & -1\times 10^{-3} & -4\times 10^{-4} & 3\times 10^{-5} & -7\times 10^{-4} & -8\times 10^{-4}\\
-% -6\times 10^{-2} & -5\times 10^{-3} & 4\times 10^{-2} & 1\times 10^{-2} & -7\times 10^{-4} & 2\times 10^{-2} & 2\times 10^{-2}\\
-% -7\times 10^{-2} & -6\times 10^{-3} & 4\times 10^{-2} & 1\times 10^{-2} & -8\times 10^{-4} & 2\times 10^{-2} & 3\times 10^{-2}
-% \end{pmatrix},
-% \\
-% F^P_{\psi_{s+c}} =
-% \begin{pmatrix}
-% 9\times 10^{-2} & 9\times 10^{-3} & -7\times 10^{-2} & -1\times 10^{-2} & 1\times 10^{-3} & -4\times 10^{-2} & -4\times 10^{-2}\\
-% 9\times 10^{-3} & 1\times 10^{-3} & -9\times 10^{-3} & -2\times 10^{-3} & 2\times 10^{-4} & -5\times 10^{-3} & -4\times 10^{-3}\\
-% -7\times 10^{-2} & -9\times 10^{-3} & 7\times 10^{-2} & 8\times 10^{-3} & -1\times 10^{-3} & 5\times 10^{-2} & 3\times 10^{-2}\\
-% -1\times 10^{-2} & -2\times 10^{-3} & 8\times 10^{-3} & 2\times 10^{-2} & -7\times 10^{-5} & -3\times 10^{-4} & 5\times 10^{-3}\\
-% 1\times 10^{-3} & 2\times 10^{-4} & -1\times 10^{-3} & -7\times 10^{-5} & 3\times 10^{-5} & -9\times 10^{-4} & -6\times 10^{-4}\\
-% -4\times 10^{-2} & -5\times 10^{-3} & 5\times 10^{-2} & -3\times 10^{-4} & -9\times 10^{-4} & 3\times 10^{-2} & 2\times 10^{-2}\\
-% -4\times 10^{-2} & -4\times 10^{-3} & 3\times 10^{-2} & 5\times 10^{-3} & -6\times 10^{-4} & 2\times 10^{-2} & 2\times 10^{-2}
-% \end{pmatrix}.
-% \end{gather*}
+\section{Numerical derivative}
+The derivatives with respect to cosmological parameters were taken with a central difference formula, i.e.
+\begin{equation*}
+    f'(x) = \frac{f(x+h) - f(x-h)}{2h} + \mathcal O (h^2).
+\end{equation*}
+Each change of the cosmological parameters requires a recalculation of the entire cosmology, making it computationally expensive.
+For the approximation to be accurate a balance needs to be found between numerical errors for small $h$ and a larger $\mathcal O(h^2)$ error for larger $h$.
+The $h$ values chosen for each parameter are shown in table \ref{tab:cosmo-params-diff} and are similar to the values used in \cite{Namikawa_2016}\footnote{As confirmed during a conversation with the author.}.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{lcc}
+    \toprule
+    Parameter & Fiducial value & Finite difference ($h$) \\
+    \midrule
+    $H$            & $67.4$               & fiducial\,$\times0.1$   \\
+    $\Omega_b h^2$ & $0.0223$             & fiducial\,$\times0.1$   \\
+    $\Omega_c h^2$ & $0.119$              & fiducial\,$\times0.005$ \\
+    $n_s$          & $0.965$              & fiducial\,$\times0.005$ \\
+    $A_s$          & $2.13\times10^{-9}$  & fiducial\,$\times0.1$   \\
+    $m_\nu$        & $0.06$               & fiducial\,$\times0.1$   \\
+    $w_0$          & $-1$                 & $0.03$                  \\
+    \bottomrule
+    \end{tabular}
+    \caption{Fiducial cosmological parameters and their finite-difference steps}
+    \label{tab:cosmo-params-diff}
+\end{table}
+
+To test the accuracy, we vary $h$ by $\pm 5\%$ and $\pm 10\%$ and plot the relative change in the derivative of the equilateral bispectrum in figure \ref{fig:deraccuracy}.
+As long as numerical noise doesn't dominate, it can be shown that the relative error in our approximation is approximately 5 times the relative difference that taking $h \rightarrow h(1 \pm 0.1)$ leads to.
+We thus conclude that, based on this test, the derivatives are almost always computed with up to one percent of error.
+
+\begin{figure}[t]
+    \centering
+    \includegraphics[height=\textheight]{/Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/code/plots/derivative_accuracy.png}
+    \caption{Central difference approximation of derivative of equilateral CMB weak lensing bispectrum for differences as specified in table \ref{tab:cosmo-params-diff} as well as for differences $5\%$ and $10\%$ larger and smaller.}
+    \label{fig:deraccuracy}
+\end{figure}
 

### 2025-04-30 code/data_exporter.py
misc

@@ -141,7 +141,7 @@ def data_export(folder_name, cosm_par, lps = True, a_create = True, b_create = T
             print(f'creating a for {folder_name}')
             # a (2d: k, z)
             with open(filepath + '/a', 'w') as file:
-                data = [[cosm_data.a(k, z) for z in zs_fine] for k in ks_log_fine]
+                data = [[cosm_data.a(k, z) for k in ks_log_fine] for z in zs_fine]
                 for row in data:
                     file.write(' '.join(map(str, row)) + '\n')
             print(f'a created to {folder_name}')
@@ -150,7 +150,7 @@ def data_export(folder_name, cosm_par, lps = True, a_create = True, b_create = T
             print(f'creating b for {folder_name}')
             # b (2d: k, z)
             with open(filepath + '/b', 'w') as file:
-                data = [[cosm_data.b(k, z) for z in zs_fine] for k in ks_log_fine]
+                data = [[cosm_data.b(k, z) for k in ks_log_fine] for z in zs_fine]
                 for row in data:
                     file.write(' '.join(map(str, row)) + '\n')
             print(f'b created to {folder_name}')
@@ -159,7 +159,7 @@ def data_export(folder_name, cosm_par, lps = True, a_create = True, b_create = T
             print(f'creating c for {folder_name}')
             # c (2d: k, z)
             with open(filepath + '/c', 'w') as file:
-                data = [[cosm_data.c(k, z) for z in zs_fine] for k in ks_log_fine]
+                data = [[cosm_data.c(k, z) for k in ks_log_fine] for z in zs_fine]
                 for row in data:
                     file.write(' '.join(map(str, row)) + '\n')
             print(f'c created to {folder_name}')

### 2025-04-30 code/data_exporter_new.py

@@ -113,17 +113,17 @@ par_names = ('H', 'ombh2', 'omch2', 'ns', 'As', 'mnu', 'w0')
 fiducial_cosm_par = np.array([67.4, 0.0223, 0.119, 0.965, 2.13e-9, 0.06, -1])
 # based on values that toshiya told me about
 cosm_par_delta = np.array([fiducial_cosm_par[0] * 0.1,
-                           fiducial_cosm_par[1] * 0.1,
-                           fiducial_cosm_par[2] * 0.005,
-                           fiducial_cosm_par[3] * 0.005,
+                           fiducial_cosm_par[1] * 0.15,
+                           fiducial_cosm_par[2] * 0.05,
+                           fiducial_cosm_par[3] * 0.01,
                            fiducial_cosm_par[4] * 0.1,
-                           fiducial_cosm_par[5] * 0.1, # mnu h should probably be smaller
-                           0.03])
+                           fiducial_cosm_par[5] * 0.05, # mnu h should probably be smaller
+                           0.06])
 
 num_pars = len(par_names)
 
 # perturb the delta used to calculate derivative in order to test stability
-delta_delta_coeffs = [2, 1, -1, -2]
+delta_delta_coeffs = [2, 1, 0, -1, -2]
 
 def delta_delta_coeffs_to_str(coeff):
     if coeff == 2:

### 2025-04-30 paper/paper_draft.tex
paper changes

@@
 \subsection{Parameter Constraints}
-For the parameter constraints, even when combining all signals, most of the parameters are poorly constrained (with minimum uncertainties much bigger than their fiducial values).
-The combination that is well constrained is $H_0$ and $\sigma_8$.
-Alternatively one can also look at $H_0$ and $S_8$, where $S_8$ is defined as
-\begin{equation*}
-    S_8 = \sigma_8 \sqrt{\frac{\Omega_c + \Omega_b}{0.3}}
-\end{equation*}
-and is commonly found to be a particularly well constrained parameter in weak lensing surveys.
-The results are essentially the same so we will stick with $\sigma_8$.
-The constraints and covariances are shown in figures \ref{fig:side_by_side} (a) and (b).
-In particular, you can see in  the off diagonal plot that combining the CMB and galaxy lensing removes approximate degeneracies in $H_0$, $\sigma_8$ space to create a much smaller $1\sigma$ confidence ellipse.
-When comparing powerspectra to bispectra constraints in figure \ref{fig:S8const} we see a more mild improvement in constraints when combining the different sources of information.
-% \begin{figure}[htbp]
-%     \centering
-%     \begin{subfigure}[t]{0.48\textwidth}
-%         \centering
-%         \includegraphics[width=\textwidth]{paramconstraints_sigma8.pdf}
-%         
-%         \label{fig:sigma8const}
-%     \end{subfigure}
-%     \hfill
-%     \begin{subfigure}[t]{0.48\textwidth}
-%         \centering
-%         \includegraphics[width=\textwidth]{paramconstraints_sigma8_diffspectra.pdf}
-%         \caption{Same as Fig.~\ref{fig:sigma8const}, except comparing constraints from only using power spectra, only bispectra, or both.
-$S_8$ instead of $\sigma_8$.
-In all three cases we use CMB $and$ galaxy lensing.}
-%         \label{fig:S8const}
-%     \end{subfigure}
-%     \caption{Side-by-side comparison of parameter constraints.}
-%     \label{fig:side_by_side}
-% \end{figure}
+Parameter constraints for the full set of $\Lambda$ CDM parameters are shown in figure \ref{fig:paramconstraintsall}.
+The combination $H_0$, $\sigma_8$, and $\Omega_m$ is particularly well constrained, constraints for these parameters alone (i.e.
+marginalized over all other parameters) are found in figure \ref{fig:paramconstraintstight}.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/paramconstraints_difftracers.pdf}
-    \caption{Constraints for $H_0$ and $\sigma_8$ for CMB lensing, galaxy lensing, and CMB + galaxy lensing combined.
-We use the information from both power and bispectra together with $l_{\max}=2000$ in all three cases.
-The confidence ellipses are for $1\sigma$ certainty.}
-    \label{fig:difftracer}
+    \includegraphics[width=\textwidth]{figures/param_constraints_tight.pdf}
+    \caption{Constraints for $H_0$, $\sigma_8$, and $\Omega_m$ for different combinations of weak lensing spectra.
+Valid for ``stage 4'' experiments.
+We use $l_{\max}=2000$.
+The confidence ellipses are for $1\sigma$ certainty.
+The standard deviation of each Gaussian is given explicitly in the top left corners of the diagonal plots.}
+    \label{fig:paramconstraintstight}
 \end{figure}
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/paramconstraints_diffspectra.pdf}
-    \caption{Same as Fig. \ref{fig:difftracer}, except comparing constraints from only using power spectra, only bispectra, or both. In all three cases we use CMB $and$ galaxy lensing.}
-    \label{fig:diffspectra}
+    \includegraphics[width=\textwidth]{figures/param_constraints_all.pdf}
+    \caption{Same as Fig. \ref{fig:paramconstraintstight}, except comparing constraints for all parameters of our $\Lambda$CDM cosmology.}
+    \label{fig:paramconstraintsall}
 \end{figure}
@@
 \section{Discussion and conclusion}\label{sec:discussion}
-% \bibliographystyle{alpha} % Choose your citation style
+For stage 4 surveys weak lensing matter and power spectra can all be measured with high significance.
+Additionally, parameter constraints are significantly improved by including bispectrum information in nearly all cases.
+This falls in line with work by Toshiya Namikawa \cite{Namikawa_2016}, where a 33 percent and 35 percent improvement was found for $m_\nu$ and $w_0$ by measuring the CMB ,lensing bispectrum in addition to the powerspectrum.
+In particular, from the confidence ellipses it can be seen that these improvements do not just come from additional information, but also because the bispectrum has different (approximate) parameter degeneracies compared to the power spectrum and thus they synergize when combined together.
+
+There are a number of limitations in this work.
+The most obvious is that the fitting formula used for the nonlinear matter bispectrum is not very accurate, only up to about 10 percent.
+Additionally, certain parameters that affect nonlinear large scale structure, such as neutrino masses, likely change the fitting parameters of the bispectrum when varied.
+This is currently not taking into account, making the derivatives less accurate.
+Another limitation is taking into account full sky effects.
+We assumed throughout this paper that surveys would measure multipoles as low as $l=2$, however our analytical formulas to calculate lensing spectra use the Limber approximation, which is only really accurate after $l=50$.
+Not using this approximation would make the calculation more complicated due to requireing knowledge of the uneeven time matter power and bispectra and computationally much longer due to having to evaluate a double integral.
+With the current methods it would thus not be feasible to calculate Fisher matrices without employing the Limber approximation, even if we would be able to calculate the unequel time matter spectra.
+Perhaps methods used in \cite{Chen:2021vba} could be used to solve this problem.
+Additionally, to calculate the Fisher matrices within a managable timeframe, the derivative of the lensing spectra with respect to cosmological parameters was done using a central difference formula, introducing errors of at most a few percent (CHECK EXACT VALUE LATER) (see appendix \ref{sec:derivatives} for further discussion) and functions such as the matter power spectrum, fitting parameters of the matter bispectrum, etc.
+were approximated using pregenerated data and interpolation, also introducing at most a few percent of error.
+Currently these errors are not significant because this paper only aims to give a rough / qualitative understanding of how usefull CMB and galaxy weak lensing spectra are, however if a similar calculation is done aiming to give precise quantitative results, these factors will need to be looked at again and improved.
+Furthermore, the current models for calculating lensing noise are on the basic side.
+Analysing real life data requires a better understanding of underlying systematics.
+For instance, determining the redshift distribution of galaxies is well known to be a difficult process (CITE SOMETHING HERE) so these errors would have to be taken into account as well.
+A better estimator can also be used.
+For example, the iterative estimator developed in CITE leads to a significantly better signal to noise for the CMB lensing bispectrum, as shown in \cite{Namikawa_2016}.
+Finally, for the covariance of the spectra estimators the Gaussian approximation was used, meaning we ignored any contributions not coming from standard Wick contractions, see appendix \ref{sec:fisher}.
+A next step could thus be to take these additional contributions into account, which could perhaps be done by running N body simulations and ray tracing simulations.
+Notably, this could potentially (depending on how accurate the simulations are) also allow us to calculate the lensing spectra much more accurately, i.e.
+without having to use a fitting formula for the matter bispectrum or Limber approximation.
+
+\pagebreak
 \printbibliography
@@
 \section{Numerical derivative}
+\label{sec:derivatives}
 The derivatives with respect to cosmological parameters were taken with a central difference formula, i.e.

### 2025-04-30 paper/fisherstuff.py

@@ -16,11 +16,11 @@ fisher_matrices_dir = '/Users/jonasfrugte/Desktop/Research_Project/fisher_calc_w
 ######################
 # Bispectra Matrices #
 ######################
-visb_c = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_c.txt')
+visb_c = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_c.txt')
 
-visb_s = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_s.txt')
+visb_s = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_s.txt')
 
-visb_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_both.txt')
+visb_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_both.txt')
 
 ########################
 # Powerspectra Matrices
@@ -35,7 +35,7 @@ visp_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_both.txt')
 # Prior Matrices #
 ######################
 
-prior = np.diag((0.6, 0.00015, 0.0014, 0.0044, 0.01, 0.034*1e-9, 0.1))
+#prior = np.diag((0.6, 0.00015, 0.0014, 0.0044, 0.01, 0.034*1e-9, 0.1))
 
 ########### NORMALIZATION ###########
 
@@ -147,8 +147,8 @@ def print_sorted_eigens(matrix_name, mat):
         print(vecs[:, i])
 
 # Indices to remove: 6,1,4,5 (0-based). This means we keep [0,2,3].
-#keep_indices = [0, 7, 9]
-keep_indices = [0, 1, 2, 3, 4, 5, 6]
+keep_indices = [0, 7, 9]
+#keep_indices = [0, 1, 2, 3, 4, 5, 6]
 #keep_indices = [4, 6]
 
 # Reduced copies for bispectra
@@ -161,7 +161,7 @@ visp_c_reduced = visp_c[np.ix_(keep_indices, keep_indices)]
 visp_s_reduced = visp_s[np.ix_(keep_indices, keep_indices)]
 visp_f_reduced = visp_f[np.ix_(keep_indices, keep_indices)]
 
-prior_reduced = prior[np.ix_(keep_indices, keep_indices)]
+#prior_reduced = prior[np.ix_(keep_indices, keep_indices)]
 
[…]

### 2025-04-30 paper/tempCodeRunnerFile.py

@@ -1,9 +1,8 @@
 
-    # cov_matrices = [
-    #     planck_reduced,
-    #     inv(inv(planck_reduced) + visp_f_reduced),
-    #     #planck_reduced + inv(visb_f_reduced),
-    #     inv(inv(planck_reduced) + visb_f_reduced + visp_f_reduced)
-    # ]
+    cov_matrices = [
+        inv(visp_f_reduced),
+        inv(visb_f_reduced),
+        inv(visb_f_reduced + visp_f_reduced)
+    ]
 
-    # labels = ['planck only', 'powerspectra', 'power- $+$ bispectra']
+    labels = ['powerspectr

### 2025-05-02 paper/paper_draft.tex
added tables

@@
     \centering
+    \tiny
+    \begin{tabular}{lrrrrr}
+        \hline
+         Par            &   CMB + Gal Powersp &   CMB + Gal Bisp &   CMB Power- + Bisp &   Gal Power- + Bisp &   CMB + Gal Power- + Bisp \\
+        \hline
+         $H_0$          &                41.0 &              1.9 &                 7.4 &                 4.8 &                       0.7 \\
+         $\Omega_b h^2$ &               109.8 &             18.4 &                10.1 &                17.3 &                       7.1 \\
+         $\Omega_c h^2$ &                58.3 &              2.4 &                 2.2 &                 2.7 &                       1.1 \\
+         $n_s$          &                11.4 &              1.9 &                 1.8 &                 3.2 &                       1.0 \\
+         $m_\nu$        &               603.3 &            373.9 &               216.3 &               260.1 &                     102.3 \\
+         $A_s$          &                39.1 &              7.0 &                 3.4 &                 7.8 &                       2.4 \\
+         $w_0$          &                46.1 &              3.1 &                18.3 &                12.7 &                       1.6 \\
+        \hline
+    \end{tabular}
+\caption{Parameter constraints in percentage of the fiducial values for different combinations of weak lensing information.}
+\label{tab:paramconstraintsall}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/param_constraints_all.pdf}
+    \caption{Same as Fig. \ref{fig:paramconstraintstight}, except comparing constraints for all parameters of our $\Lambda$CDM cosmology.}
+    \label{fig:paramconstraintsall}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \tiny
+    \begin{tabular}{lrrrrr}
+        \hline
+         Par        &   CMB + Gal Powersp &   CMB + Gal Bisp &   CMB Power- + Bisp &   Gal Power- + Bisp &   CMB + Gal Power- + Bisp \\
+        \hline
+         $H_0$      &        3.3022621043 &     0.9658765234 &        0.1360204496 &        0.6433515918 &              0.1173950908 \\
+         $\sigma_8$ &        0.0000000178 &     0.0000000148 &        0.0000000023 &        0.0000000122 &              0.0000000020 \\
+         $\Omega_m$ &        0.4126866976 &     0.2388283763 &        0.1579156300 &        0.1996632391 &              0.1122269336 \\
+        \hline
+    \end{tabular}
+\caption{Same as table \ref{tab:paramconstraintsall}, except for a smaller selection of parameters well constrained by lensing.}
+\label{tab:paramconstraintstight}
+\end{figure}
+
+\begin{figure}
+    \centering
     \includegraphics[width=\textwidth]{figures/param_constraints_tight.pdf}
@@
     \label{fig:paramconstraintstight}
-\end{figure}
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/param_constraints_all.pdf}
-    \caption{Same as Fig. \ref{fig:paramconstraintstight}, except comparing constraints for all parameters of our $\Lambda$CDM cosmology.}
-    \label{fig:paramconstraintsall}
 \end{figure}
@@
 We thus conclude that, based on this test, the derivatives are almost always computed with up to one percent of error.
+The exception is the derivative with respect to neutrino masses, which introduces a larger error.
 
@@
     \includegraphics[height=\textheight]{/Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/code/plots/derivative_accuracy.png}
-    \caption{Central difference approximation of derivative of equilateral CMB weak lensing bispectrum for differences as specified in table \ref{tab:cosmo-params-diff} as well as for differences $5\%$ and $10\%$ larger and smaller.}
+    \caption{Central difference approximation of derivative of equilateral CMB weak lensing bispectrum (left) and powerspectrum (right) for differences as specified in table \ref{tab:cosmo-params-diff} as well as for differences $5\%$ and $10\%$ larger and smaller.}
     \label{fig:deraccuracy}

### 2025-05-02 paper/fisherstuff.py

@@ -148,7 +148,7 @@ def print_sorted_eigens(matrix_name, mat):
 
 # Indices to remove: 6,1,4,5 (0-based). This means we keep [0,2,3].
 keep_indices = [0, 7, 9]
-#keep_indices = [0, 1, 2, 3, 4, 5, 6]
+# keep_indices = [0, 1, 2, 3, 4, 5, 6]
 #keep_indices = [4, 6]
 
 # Reduced copies for bispectra
@@ -350,6 +350,21 @@ def plot_corner(
     plt.tight_layout()
     return fig, axes
 
+from tabulate import tabulate
+
+def save_table(param_names, param_vals, which_pars, constraints, labels):
+    table = [[0 for i in range(len(constraints) + 1)] for j in range(len(which_pars))]
+    for i in range(len(which_pars)):
+        # param names and fiducial values
+        table[i][0] = param_names[i]
+        #table[i][1] = param_vals[i]
+        #constraints
+        for j in range(len(constraints)):
+            table[i][j+1] = np.abs(np.sqrt(constraints[j][i][i]) / param_vals[i]) * 100
+    labels = ['Par'] + labels
+    print(tabulate(table, headers=labels, tablefmt='latex_raw', floatfmt='.10f'))
+        
+    
 
 if __name__ == "__main__":
     # Parameter info
@@ -409,7 +424,7 @@ if __name__ == "__main__":
         inv(visb_f_reduced + visp_f_reduced)
     ]
 
-    labels = ['CMB + Galaxy Powerspec', 'CMB + Galaxy Bispectra', 'CMB Power- + Bispectra', 'Galaxy Power- + Bispectra', 'CMB + Galaxy Power- + Bispectra']
+    labels = ['CMB + Gal Powersp', 'CMB + Gal Bisp', 'CMB Power- + Bisp', 'Gal Power- + Bisp', 'CMB + Gal Power- + Bisp']
 
     fig, axes = plot_corner(
         cov_matrices=cov_matrices,
@@ -424,4 +439,6 @@ if __name__ == "__main__":
 
     #print(np.linalg.inv(visp_f_reduced + visb_f_reduced))
     # plt.show()
-    plt.savefig('/Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/paper/figures/param_constraints_tight.pdf', dpi = 300)
[…]

### 2025-05-02 code/Fisher_matrix_approx.py
new fishes

@@ -5,7 +5,7 @@ from itertools import *
 
 lmin = 2
 lmax = 2000
-stepsizes = [1 * 8, 5 * 8, 10 * 8]
+stepsizes = [1 * 8 * 2, 5 * 8 * 2, 10 * 8 * 2]
 num_bispec_samples = 100
 num_cores = 32
 
@@ -36,27 +36,5 @@ for tracer in ['c', 's', 'both']:
             print(f'{tracer}:', counter, '/', num_pars * (num_pars + 1) / 2)
             counter += 1
     
-    np.savetxt(f'fisher_matrices/fish_mat_bisp_approx_{tracer}_5p.txt', mat)
+    np.savetxt(f'fisher_matrices/fish_mat_bisp_approx_{tracer}.txt', mat)
     print(mat)
-
-
-# for python implementation
-# if __name__ == "__main__":
-#     num_cores = multiprocessing.cpu_count()
-    
-#     mat = np.zeros((len(pars), len(pars)))
-    
-#     # Generate index pairs for the lower triangular part of the matrix (excluding diagonal)
-#     index_pairs = [(i, j) for i in range(len(pars)) for j in range(i + 1)]
-    
-#     with multiprocessing.Pool(processes=num_cores) as pool:
-#         results = pool.map(fisher_calc_wrapper, index_pairs)
-    
-#     # Fill the matrix with computed values
-#     for i, j, result in results:
-#         mat[i, j] = result
-#         mat[j, i] = result  # Symmetric assignment
-
-#     print(mat)
-    
-#     print("Fisher matrix computed successfully!")

### 2025-05-06 code/Fisher_matrix_approx.py
added planck-like noise

@@ -2,12 +2,13 @@ import numpy as np
 import multiprocessing
 import Fisher_calc as vis
 from itertools import *
-
+print('snoopie')
 lmin = 2
 lmax = 2000
 stepsizes = [1 * 8 * 2, 5 * 8 * 2, 10 * 8 * 2]
+
 num_bispec_samples = 100
-num_cores = 32
+num_cores = 4
 
 pars = [b'H', b'ombh2', b'omch2', b'ns', b'mnu', b'As', b'w0']
 
@@ -22,19 +23,28 @@ def fisher_calc_wrapper(args, tracers):
         return vis.Fisher_mat_single(lmin, 0, 200, stepsizes[0], num_bispec_samples, pars[i], pars[j], num_cores, b's') + vis.Fisher_mat_single(lmin, 200, 1000, stepsizes[1], num_bispec_samples, pars[i], pars[j], num_cores, b's') + vis.Fisher_mat_single(lmin, 1000, lmax, stepsizes[2], num_bispec_samples, pars[i], pars[j], num_cores, b's')
     if tracers == 'both':
         return vis.Fisher_mat_full(lmin, 0, 200, stepsizes[0], num_bispec_samples, pars[i], pars[j], num_cores) + vis.Fisher_mat_full(lmin, 200, 1000, stepsizes[1], num_bispec_samples, pars[i], pars[j], num_cores) + vis.Fisher_mat_full(lmin, 1000, lmax, stepsizes[2], num_bispec_samples, pars[i], pars[j], num_cores)
+    
+def fisher_calc_wrapper_planck(args, tracers):
+    i, j = args
+    # Fisher_mat_full(int lmin, int lminbin, int lmax, int triangle_step_size, int num_bispec_samples, char* par1, char* par2, int num_cores)
+    # Fisher_mat_single(int lmin, int lminbin, int lmax, int triangle_step_size, int num_bispec_samples, char* par1, char* par2, int num_cores, char* type)
+    if tracers == 'c':
+        return vis.Fisher_mat_single(8, 0, 400, 8, num_bispec_samples, pars[i], pars[j], num_cores, b'c')
 
 num_pars = len(pars)
+print('snoopie2')
 
-for tracer in ['c', 's', 'both']:
+for tracer in ['c']: # ['c', 's', 'both']:
+    print('snoopie: ', tracer)
     mat = np.zeros((num_pars, num_pars))
     counter = 1
     for i in range(num_pars):
         for j in range(i, num_pars):
[…]

### 2025-05-06 code/cmb_noise_fast_runner.py

@@ -7,9 +7,10 @@ import numpy as np
 import os
 
 #############################################################################################
-# configs = ((1, 1), (1, 2), (1, 3), (2, 2), (2, 3))#, (3, 3))
+configs = ((1, 1), (1, 2), (1, 3), (2, 2), (2, 3))#, (3, 3))
 # configs = ((2, 2), (2, 3))
-configs = [[2, 3]]
+# configs = [[2, 3]]
+# configs = [[1, 1]]
 #############################################################################################
 
 num_configs = len(configs)
@@ -66,7 +67,7 @@ if __name__ == '__main__':
         import matplotlib.pyplot as plt
         plt.loglog(Ls, np.array(noise_vals) * Ls * (Ls + 1) / 2 / np.pi)
         plt.xlim(1, 2000)
-        plt.savefig('test3.png')
+        plt.savefig('test.png')
     
     # Noise configurations: (sigma, Delta_T, Delta_P)
     # noise_configs = [
@@ -77,7 +78,7 @@ if __name__ == '__main__':
     # ]
 
     noise_configs = [
-        (4, 1, 1.41),
+        (5, 30, 52),
     ]
 
     # noise_configs = [

### 2025-05-06 code/data_importer_new.pyx

@@ -27,11 +27,12 @@ def get_k_max():
     return k_max
 
 # specifies the noise that will be used for cmb lensing
-cdef int conv_noise_type = 2
+cdef int conv_noise_type = 3
 # type values correspond to:
 # 0: S0 noise curves (old, remove)
 # 1: sigma = 1, Delta_P = 6 (stage 3 toshiya)
-# 2: sigma = 3, Delta_P = 1 (stage 4 toshiya)§
+# 2: sigma = 3, Delta_P = 1 (stage 4 toshiya)
+# 3: sigma = 5, Delta_T = 30, Delta_P = 52 (planck (double check))
 
 # (units are in microKelvin arcmin and arcmin)
 
@@ -50,6 +51,7 @@ cdef double lmax_cmbn = ls_cmbn[lnum_cmbn - 1]
 
 cdef double[:] cmbn_301 = np.abs(np.loadtxt('cmb_noise_files/Ns_sigma3_DeltaT0_DeltaP1.txt')) / (ls_cmbn_np_array * (ls_cmbn_np_array + 1))
 cdef double[:] cmbn_106 = np.abs(np.loadtxt('cmb_noise_files/Ns_sigma1_DeltaT0_DeltaP6.txt')) / (ls_cmbn_np_array * (ls_cmbn_np_array + 1))
+cdef double[:] planck_noise = np.abs(np.loadtxt('cmb_noise_files/Ns_sigma_planck.txt')) / (ls_cmbn_np_array * (ls_cmbn_np_array + 1))
 
 # perhaps we can export data as c arrays instead of as memory views in the future so that we can specify return types like below and can avoid having to 
 # declare types of all data before every data import
@@ -315,12 +317,16 @@ cdef double lps_noise(int l, char* type1, char* type2) noexcept nogil:
         elif conv_noise_type == 2:
             # stage 4 toshiya
             noise = interp.logspace_linear_interp(l, lmin_cmbn, lmax_cmbn, lnum_cmbn, cmbn_301)
+        elif conv_noise_type == 3:
+            noise = interp.logspace_linear_interp(l, lmin_cmbn, lmax_cmbn, lnum_cmbn, planck_noise)
 
     if type1[0] == b's' and type2[0] == b's':
         if conv_noise_type == 1:
             noise = 4. * ((l - 1.) * l * (l + 1.) * (l + 2.))**(-1) * 0.3**2 / 5
         if conv_noise_type == 2:
[…]

### 2025-05-06 code/interp_settings.py

@@ -1,11 +1,11 @@
 exp_par = {
     'k_min' : 1e-4,
     'k_max' : 2000,
-    'k_num' : 300,
+    'k_num' : 300 * 2,
     'chi_min' : 1e-8,
     'chi_max' : 14000, # slightly more than chi_last_scatter
     'chi_num' : 100 * 50,
     'z_min' : 1e-12,
     'z_max' : 1100, # slightly more than z_last_scatter
-    'z_num' : 100
+    'z_num' : 100 * 2
 }

### 2025-05-06 paper/paper_draft.tex
paper changes

@@
         \hline
-         $H_0$          &                41.0 &              1.9 &                 7.4 &                 4.8 &                       0.7 \\
-         $\Omega_b h^2$ &               109.8 &             18.4 &                10.1 &                17.3 &                       7.1 \\
-         $\Omega_c h^2$ &                58.3 &              2.4 &                 2.2 &                 2.7 &                       1.1 \\
-         $n_s$          &                11.4 &              1.9 &                 1.8 &                 3.2 &                       1.0 \\
-         $m_\nu$        &               603.3 &            373.9 &               216.3 &               260.1 &                     102.3 \\
-         $A_s$          &                39.1 &              7.0 &                 3.4 &                 7.8 &                       2.4 \\
-         $w_0$          &                46.1 &              3.1 &                18.3 &                12.7 &                       1.6 \\
+         $H_0$          &        9.8607392131 &     1.9228760331 &        5.0596593228 &        4.9998059321 &              0.6566865031 \\
+         $\Omega_b h^2$ &       21.1047938757 &    22.3422408475 &        9.9278715395 &       19.0626128005 &              6.6783756779 \\
+         $\Omega_c h^2$ &        0.0000125702 &     1.9971113528 &        0.0000125243 &        2.4722397353 &              0.0000125239 \\
+         $n_s$          &        2.5555063255 &     2.0884251995 &        1.0353938531 &        3.4233966612 &              0.6771684184 \\
+         $m_\nu$        &      540.3486045452 &   397.7301866598 &      197.9074769610 &      274.0779514021 &            110.8336892534 \\
+         $A_s$          &       17.8092869822 &     7.5399323519 &        4.1275268227 &        7.9422103452 &              2.0233577134 \\
+         $w_0$          &       43.5247638551 &     3.8394789067 &       12.3789001621 &       13.7550208605 &              1.7287298348 \\
         \hline
@@
     \includegraphics[width=\textwidth]{figures/param_constraints_all.pdf}
-    \caption{Same as Fig. \ref{fig:paramconstraintstight}, except comparing constraints for all parameters of our $\Lambda$CDM cosmology.}
+    \caption{Parameter constraints and confidence ellipses for $\Lambda$CDM model as well as neutrino mass and dark energy equation of state parameter.
+Using ``stage 4'' noise with multipole range from $2$ to $2000$.
+Different weak lensing information combinations considered.
+No priors used.
+The confidence ellipses are for $1\sigma$ certainty.
+They show approximate degeneracies in the information gained from a survey if they are ``stretched''.
+When the information of ellipses with degeneracies in different directions is combined, the degeneracies are removed and the constraints typically becomes much better on the relevant parameters.}
     \label{fig:paramconstraintsall}
@@
         \hline
-         $H_0$      &        3.3022621043 &     0.9658765234 &        0.1360204496 &        0.6433515918 &              0.1173950908 \\
-         $\sigma_8$ &        0.0000000178 &     0.0000000148 &        0.0000000023 &        0.0000000122 &              0.0000000020 \\
-         $\Omega_m$ &        0.4126866976 &     0.2388283763 &        0.1579156300 &        0.1996632391 &              0.1122269336 \\
+         $H_0$          &        5.3683958191 &     0.8819917551 &        0.1393060101 &        0.6113966629 &              0.1212982914 \\
+         $\sigma_8$     &        0.0000000157 &     0.0000000157 &        0.0000000057 &        0.0000000099 &              0.0000000046 \\
+         $\Omega_b h^2$ &        5.5823437092 &     3.3491075426 &        2.8087163845 &        3.9650003059 &              2.2514916319 \\
         \hline
@@
     \includegraphics[width=\textwidth]{figures/param_constraints_tight.pdf}
-    \caption{Constraints for $H_0$, $\sigma_8$, and $\Omega_m$ for different combinations of weak lensing spectra.
-Valid for ``stage 4'' experiments.
-We use $l_{\max}=2000$.
-The confidence ellipses are for $1\sigma$ certainty.
-The standard deviation of each Gaussian is given explicitly in the top left corners of the diagonal plots.}
+    \caption{Same as \ref{fig:paramconstraintsall}, except showing constraints for a set of parameters that can be determined particularly accurately with weak lensing surveys.
+We chose Hubble's constant, $H_0$, baryonic + dark matter density, $\Omega_m$, and the root mean square density variation when the matter distribution is smoothed on a scale of 8 megaparsecs, commonly known as $\sigma_8$.}
     \label{fig:paramconstraintstight}
@@
 
-\pagebreak
 \printbibliography
@@
     $H$            & $67.4$               & fiducial\,$\times0.1$   \\
-    $\Omega_b h^2$ & $0.0223$             & fiducial\,$\times0.1$   \\
+    $\Omega_b h^2$ & $0.0223$             & fiducial\,$\times0.2$   \\
     $\Omega_c h^2$ & $0.119$              & fiducial\,$\times0.005$ \\
-    $n_s$          & $0.965$              & fiducial\,$\times0.005$ \\
-    $A_s$          & $2.13\times10^{-9}$  & fiducial\,$\times0.1$   \\
-    $m_\nu$        & $0.06$               & fiducial\,$\times0.1$   \\
-    $w_0$          & $-1$                 & $0.03$                  \\
+    $n_s$          & $0.965$              & fiducial\,$\times0.02$ \\
+    $A_s$          & $2.13\times10^{-9}$  & fiducial\,$\times0.15$   \\
+    $m_\nu$        & $0.06$               & fiducial\,$\times0.15$   \\
+    $w_0$          & $-1$                 & $0.06$                  \\
     \bottomrule

### 2025-05-06 paper/fisherstuff.py

@@ -4,6 +4,7 @@ import matplotlib.pyplot as plt
 from matplotlib.patches import Ellipse
 from scipy.stats import norm
 from numpy.linalg import inv
+import matplotlib.lines as mlines
 
 np.set_printoptions(
     precision=1,
@@ -147,8 +148,8 @@ def print_sorted_eigens(matrix_name, mat):
         print(vecs[:, i])
 
 # Indices to remove: 6,1,4,5 (0-based). This means we keep [0,2,3].
-keep_indices = [0, 7, 9]
-# keep_indices = [0, 1, 2, 3, 4, 5, 6]
+# keep_indices = [0, 7, 9]
+keep_indices = [0, 1, 2, 3, 4, 5, 6]
 #keep_indices = [4, 6]
 
 # Reduced copies for bispectra
@@ -282,18 +283,17 @@ def plot_corner(
             sigma = np.sqrt(cov[i, i])
             x_min = param_values[i] - 100*sigma
             x_max = param_values[i] + 100*sigma
-            x_plot = np.linspace(x_min, x_max, 1000)
+            x_plot = np.linspace(x_min, x_max, 5000)
             pdf = norm.pdf(x_plot, loc=param_values[i], scale=sigma)
 
             ax.fill_between(x_plot, pdf, color=color, lw=1, alpha = 0.2)
             ax.plot(x_plot, pdf, color=color, lw=1, alpha = 1)
             
             # Print sigma^2 in matching color, offset each line a bit
-            sigma_sq = sigma**2
             ax.text(
                 0.05,
                 0.90 - 0.08*k,
-                rf"${sigma_sq:.1e}$",
+                rf"${sigma:.1e}$",
                 color=color,
                 transform=ax.transAxes
             )
@@ -345,7 +345,8 @@ def plot_corner(
                 ax_low.set_ylabel(param_names[j])
 
     # Make a legend in the top-right corner
-    fig.legend(labels, loc='upper right', bbox_to_anchor=(0.98, 0.98))
+    lines = [mlines.Line2D([], [], color = colors[i], label = labels[i]) for i in range(len(colors))]
+    fig.legend(handles=lines, loc='upper right', bbox_to_anchor=(0.98, 0.98))
 
     plt.tight_layout()
     return fig, axes
@@ -398,23 +399,21 @@ if __name__ == "__main__":
[…]

### 2025-05-06 paper/tempCodeRunnerFile.py

@@ -1,8 +1,0 @@
-
-    cov_matrices = [
-        inv(visp_f_reduced),
-        inv(visb_f_reduced),
-        inv(visb_f_reduced + visp_f_reduced)
-    ]
-
-    labels = ['powerspectr

### 2025-05-07 paper/paper_draft.tex
fixed fisher manip

@@
         \hline
-         $H_0$          &        9.8607392131 &     1.9228760331 &        5.0596593228 &        4.9998059321 &              0.6566865031 \\
-         $\Omega_b h^2$ &       21.1047938757 &    22.3422408475 &        9.9278715395 &       19.0626128005 &              6.6783756779 \\
-         $\Omega_c h^2$ &        0.0000125702 &     1.9971113528 &        0.0000125243 &        2.4722397353 &              0.0000125239 \\
-         $n_s$          &        2.5555063255 &     2.0884251995 &        1.0353938531 &        3.4233966612 &              0.6771684184 \\
-         $m_\nu$        &      540.3486045452 &   397.7301866598 &      197.9074769610 &      274.0779514021 &            110.8336892534 \\
-         $A_s$          &       17.8092869822 &     7.5399323519 &        4.1275268227 &        7.9422103452 &              2.0233577134 \\
-         $w_0$          &       43.5247638551 &     3.8394789067 &       12.3789001621 &       13.7550208605 &              1.7287298348 \\
+         $H_0$                     &              102.87 &             1.94 &                9.63 &                5.22 &                      0.94 \\
+         $\Omega_b h^2$            &              234.33 &            21.39 &               25.53 &               20.30 &                     13.50 \\
+         $\Omega_c h^2$            &              127.06 &             2.43 &                3.96 &                3.13 &                      1.42 \\
+         $n_s$                     &               25.44 &             1.94 &                3.25 &                3.82 &                      1.51 \\
+         $m_\nu$                   &             3121.97 &           307.15 &              384.96 &              277.55 &                    178.07 \\
+         $A_s$                     &               69.24 &             6.39 &                7.96 &                8.44 &                      4.45 \\
+         $w_0$                     &              103.74 &             2.90 &               23.14 &               13.87 &                      2.33 \\
+         \hline
+         \hline
+         $\sigma_8$                &               31.82 &             3.85 &                9.16 &                4.56 &                      1.84 \\
+         $S_8$                     &               37.67 &             2.08 &                2.52 &                2.27 &                      1.14 \\
+         $\Omega_m$                &              172.99 &             5.65 &               27.29 &               14.45 &                      3.57 \\
+         $\sigma_8\Omega_m^{0.25}$ &               32.15 &             2.65 &                2.82 &                1.99 &                      1.29 \\
         \hline
         \end{tabular}
-\caption{Parameter constraints in percentage of the fiducial values for different combinations of weak lensing information.}
+\caption{Parameter constraints in percentage of the fiducial values for different combinations of weak lensing information. Below the double line are constraints for derived parameters.}
 \label{tab:paramconstraintsall}
@@
 
-\begin{figure}
-    \centering
-    \tiny
-    \begin{tabular}{lrrrrr}
-        \hline
-         Par            &   CMB + Gal Powersp &   CMB + Gal Bisp &   CMB Power- + Bisp &   Gal Power- + Bisp &   CMB + Gal Power- + Bisp \\
-        \hline
-         $H_0$          &        5.3683958191 &     0.8819917551 &        0.1393060101 &        0.6113966629 &              0.1212982914 \\
-         $\sigma_8$     &        0.0000000157 &     0.0000000157 &        0.0000000057 &        0.0000000099 &              0.0000000046 \\
-         $\Omega_b h^2$ &        5.5823437092 &     3.3491075426 &        2.8087163845 &        3.9650003059 &              2.2514916319 \\
-        \hline
-        \end{tabular}
-\caption{Same as table \ref{tab:paramconstraintsall}, except for a smaller selection of parameters well constrained by lensing.}
-\label{tab:paramconstraintstight}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \tiny
+%     \begin{tabular}{lrrrrr}
+%         \hline
+%          Par        &   CMB + Gal Powersp &   CMB + Gal Bisp &   CMB Power- + Bisp &   Gal Power- + Bisp &   CMB + Gal Power- + Bisp \\
+%         \hline
+%          $H_0$      &      102.8657993427 &     1.9383910910 &        9.6317154638 &        5.2219296015 &              0.9395018958 \\
+%          $\sigma_8$ &       31.8199604416 &     3.8499102909 &        9.1631304725 &        4.5639390699 &              1.8357116464 \\
+%          $\Omega_m$ &      172.9940493951 &     5.6526767874 &       27.2943604041 &       14.4462387878 &              3.5663315554 \\
+%         \hline
+%     \end{tabular}
+% \caption{Same as table \ref{tab:paramconstraintsall}, except for a smaller selection of parameters well constrained by lensing.}
+% \label{tab:paramconstraintstight}
+% \end{figure}
 

### 2025-05-07 paper/playing_with_fish.py

@@ -38,7 +38,9 @@ visp_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_both.txt')
 
 planck = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_c_planck.txt')
 
-planck_prior = np.diag([0.0, 1/(0.0005)**2, 0, 1/(0.02)**2, 0, 0, 0])
+planck_prior = np.diag([12 / 0.6**2, 1/(0.0005)**2, 12 / 1**2, 1/(0.02)**2, 1e6, 1e19, 1e6])
+
+planck_p_prior = planck + planck_prior
 
 # pars = [b'H', b'ombh2', b'omch2', b'ns', b'mnu', b'As', b'w0']
 
@@ -54,9 +56,6 @@ param_values_normalization = np.array([
     2.1e-9,       # A_s
     -1.0          # w_0
 ])
-
-normalization_matrix = np.outer(param_values_normalization, param_values_normalization)
-# If needed, multiply each matrix by this normalization_matrix.
 
 ########### SIGMA8 ################
 
@@ -100,40 +99,28 @@ def append_row_column(mats, vec):
     return matsnew
 
 s8_ders = np.array([
-    1 / 2.76945054e-03, # H_0
-    1 / -6.50570238e+00, # Omega_b h^2 
-    1 / 4.30234556e+00,  # Omega_c h^2
-    1 / 3.06272435e-01,# n_s
-    1 / -2.15681667e-01, # m_\nu
-    1 / 1.93167857e+08, # A_s
-    1 / -2.01175433e-01  # w_0
+    2.76945054e-03, # H_0
+    -6.50570238e+00, # Omega_b h^2 
+    4.30234556e+00,  # Omega_c h^2
+    3.06272435e-01,# n_s
+    -2.15681667e-01, # m_\nu
+    1.93167857e+08, # A_s
+    -2.01175433e-01  # w_0
     ])
 
-# pars = [b'H', b'ombh2', b'omch2', b'ns', b'mnu', b'As', b'w0']
-
-visb_c, visb_s, visb_f, visp_c, visp_s, visp_f, planck, planck_prior = append_row_column(
-    [visb_c, visb_s, visb_f, visp_c, visp_s, visp_f, planck, planck_prior],
-    s8_ders
-)
-
[…]

### 2025-05-07 paper/tempCodeRunnerFile.py

@@ -1,0 +1,4 @@
+visb_c, visb_s, visb_f, visp_c, visp_s, visp_f, planck, planck_p_prior = append_row_column(
+#     [visb_c, visb_s, visb_f, visp_c, visp_s, visp_f, planck, planck_p_prior],
+#     s8_ders
+# )

### 2025-05-08 code/Fisher_matrix_approx.py
new planck fish matrix

@@ -5,7 +5,7 @@ from itertools import *
 print('snoopie')
 lmin = 2
 lmax = 2000
-stepsizes = [1 * 8 * 2, 5 * 8 * 2, 10 * 8 * 2]
+stepsizes = [1 * 8 * 4, 5 * 8 * 4, 10 * 8 * 4]
 
 num_bispec_samples = 100
 num_cores = 4
@@ -31,20 +31,27 @@ def fisher_calc_wrapper_planck(args, tracers):
     if tracers == 'c':
         return vis.Fisher_mat_single(8, 0, 400, 8, num_bispec_samples, pars[i], pars[j], num_cores, b'c')
 
+def fisher_calc_wrapper_takada_jain(args, tracers):
+    i, j = args
+    # Fisher_mat_full(int lmin, int lminbin, int lmax, int triangle_step_size, int num_bispec_samples, char* par1, char* par2, int num_cores)
+    # Fisher_mat_single(int lmin, int lminbin, int lmax, int triangle_step_size, int num_bispec_samples, char* par1, char* par2, int num_cores, char* type)
+    if tracers == 's':
+        # lmax should actually be 3000, but interpolation currently doesn't go that high
+        return vis.Fisher_mat_single(50, 0, 200, stepsizes[0], num_bispec_samples, pars[i], pars[j], num_cores, b's') + vis.Fisher_mat_single(50, 200, 1000, stepsizes[1], num_bispec_samples, pars[i], pars[j], num_cores, b's') + vis.Fisher_mat_single(50, 1000, 2000, stepsizes[2], num_bispec_samples, pars[i], pars[j], num_cores, b's')
 num_pars = len(pars)
 print('snoopie2')
 
-for tracer in ['c']: # ['c', 's', 'both']:
+for tracer in ['s']: # ['c', 's', 'both']:
     print('snoopie: ', tracer)
     mat = np.zeros((num_pars, num_pars))
     counter = 1
     for i in range(num_pars):
         for j in range(i, num_pars):
-            result = fisher_calc_wrapper_planck((i, j), tracer)
+            result = fisher_calc_wrapper_takada_jain((i, j), tracer)
             mat[i, j] = result
             mat[j, i] = result  # Symmetric assignment
             print(f'{tracer}:', counter, '/', num_pars * (num_pars + 1) / 2)
             counter += 1
     
[…]

### 2025-05-08 code/Fisher_matrix_powerspectra.py

@@ -18,12 +18,20 @@ def fisher_calc_wrapper(args, tracers):
     if tracers == 'both':
         return vispy.Fisher_powersp(lmin, lmax, pars[i], pars[j])
 
-for tracer in ['c', 's', 'both']:
+def fisher_calc_wrapper_planck(args, tracers):
+    print('Planck simulation')
+    i, j = args
+    # Fisher_mat_full(int lmin, int lminbin, int lmax, int triangle_step_size, int num_bispec_samples, char* par1, char* par2, int num_cores)
+    # Fisher_mat_single(int lmin, int lminbin, int lmax, int triangle_step_size, int num_bispec_samples, char* par1, char* par2, int num_cores, char* type)
+    if tracers == 'c':
+        return vispy.Fisher_powersp_single(8, 400, b'c', par1 = pars[i], par2 = pars[j])
+
+for tracer in ['c']: #['c', 's', 'both']:
     mat = np.zeros((len(pars), len(pars)))
     for i, j in product(range(len(pars)), repeat = 2):
-        result = fisher_calc_wrapper((i, j), tracer)
+        result = fisher_calc_wrapper_planck((i, j), tracer)
         mat[i, j] = result
         mat[j, i] = result  # Symmetric assignment
 
-    np.savetxt(f'fisher_matrices/fish_mat_powersp_{tracer}.txt', mat)
+    np.savetxt(f'fisher_matrices/fish_mat_powersp_{tracer}_planck.txt', mat)
     print(mat)

### 2025-05-08 code/cosm_setup.py

@@ -44,7 +44,7 @@ class lensing_spectra:
     self.mnu = mnu        # Sum of neutrino masses
     self.w0 = w0
     self.matter_bispec_fit_params = matter_bispec_fit_params # params of matter bispectrum fitting function as (a_1, ..., a_9)
-    self.k_nl_max = 2000
+    self.k_nl_max = 1000
 
     self.obtain_results()
 
@@ -83,7 +83,7 @@ class lensing_spectra:
     print('Obtaining cosmological results.')
 
     pars = camb.CAMBparams()
-    lmax = 2000
+    lmax = 1000
     pars.set_cosmology(H0=self.H0, ombh2=self.ombh2, omch2=self.omch2, mnu=self.mnu, omk=0, tau=0.063, neutrino_hierarchy='normal')
     pars.InitPower.set_params(As=self.As, ns=self.ns, r=0)
     pars.InitPower.pivot_scalar = 0.05
@@ -92,7 +92,7 @@ class lensing_spectra:
     pars.set_dark_energy(w=self.w0)
     self.pars = pars
 
-    kmax_interp = 2000
+    kmax_interp = 1000
 
     # power spectrum in units of Mpc^3, k in units of Mpc^-1, with input (z, k) on RHS 
     print('Creating linear mps')

### 2025-05-08 code/data_importer_new.pyx

@@ -27,7 +27,7 @@ def get_k_max():
     return k_max
 
 # specifies the noise that will be used for cmb lensing
-cdef int conv_noise_type = 3
+cdef int noise_type = 3
 # type values correspond to:
 # 0: S0 noise curves (old, remove)
 # 1: sigma = 1, Delta_P = 6 (stage 3 toshiya)
@@ -309,24 +309,24 @@ cdef double lps_f(int l, char* type1, char* type2) noexcept nogil:
 cdef double lps_noise(int l, char* type1, char* type2) noexcept nogil:
     cdef float noise = 0.
     if type1[0] == b'c' and type2[0] == b'c':
-        if conv_noise_type == 0:
+        if noise_type == 0:
             noise = 4. * (l * 1.0)**(-2) * (l + 1.0)**(-2) * conv_noise_data[l-2, 7]
-        elif conv_noise_type == 1:
+        elif noise_type == 1:
             # stage 3 wide toshiya
             noise = interp.logspace_linear_interp(l, lmin_cmbn, lmax_cmbn, lnum_cmbn, cmbn_106)
-        elif conv_noise_type == 2:
+        elif noise_type == 2:
             # stage 4 toshiya
             noise = interp.logspace_linear_interp(l, lmin_cmbn, lmax_cmbn, lnum_cmbn, cmbn_301)
-        elif conv_noise_type == 3:
+        elif noise_type == 3:
             noise = interp.logspace_linear_interp(l, lmin_cmbn, lmax_cmbn, lnum_cmbn, planck_noise)
 
     if type1[0] == b's' and type2[0] == b's':
-        if conv_noise_type == 1:
+        if noise_type == 1:
             noise = 4. * ((l - 1.) * l * (l + 1.) * (l + 2.))**(-1) * 0.3**2 / 5
-        if conv_noise_type == 2:
+        if noise_type == 2:
             noise = 4. * ((l - 1.) * l * (l + 1.) * (l + 2.))**(-1) * 0.3**2 / 30 # 8 * 10.0 ** (-10) # this should actually be in sterradain, but that gives wrong results so for now we have it like this
-        if conv_noise_type == 3:
-            noise = 4. * ((l - 1.) * l * (l + 1.) * (l + 2.))**(-1) * 0.3**2 / 30
+        if noise_type == 3:
[…]

### 2025-05-08 code/sigma8_ders.py

@@ -5,7 +5,7 @@ par_names = ('H', 'ombh2', 'omch2', 'ns', 'As', 'mnu', 'w0')
 fiducial_cosm_par = (67.4, 0.0224, 0.120, 0.965, 2.1e-9, 0.06, -1)
 num_pars = len(par_names)
 
-dx = 0.025 # if you change this here it also needs to be changed at end of data_importer.pyx !!!
+dx = 0.05 # if you change this here it also needs to be changed at end of data_importer.pyx !!!
 
 dx_coeffs = [2, 1, -1, -2]
 
@@ -26,7 +26,7 @@ sigma8s = np.zeros((num_pars, len(dx_coeffs)))
 for j in range(len(dx_coeffs)):
     for i in range(num_pars):
         perturbed_params = [fiducial_cosm_par[k] * (1 + dx_coeffs[j] * dx * int(k == i)) for k in range(num_pars)]
-        spectra = cs.lensing_spectra(*perturbed_params, fiducial_k_nls=True)
+        spectra = cs.lensing_spectra(*perturbed_params)
         sigma8s[i, j] = spectra.results.get_sigma8_0()
 
 print(sigma8s)

### 2025-05-08 paper/playing_with_fish.py
paper changes

@@ -42,6 +42,8 @@ planck_prior = np.diag([12 / 0.6**2, 1/(0.0005)**2, 12 / 1**2, 1/(0.02)**2, 1e6,
 
 planck_p_prior = planck + planck_prior
 
+toshiya_prior = np.diag([1 / 0.54**2, 1/(0.00015)**2, 1/(0.0014)**2, 1/(0.0042)**2, 0.05 / 0.06**2, 1e21, 1 / 0.5**2])
+
 # pars = [b'H', b'ombh2', b'omch2', b'ns', b'mnu', b'As', b'w0']
 
 
@@ -143,10 +145,10 @@ sigmaomm_ders = [0,
            (0.25 * sigma8 * omm**(-0.75))]
 
 
-keep_indices = [0, 7, 9]
+# keep_indices = [0, 7, 9]
 # keep_indices=[10]
 # keep_indices = [0, 1, 2, 3, 4, 5, 6]
-#keep_indices = [4, 6]
+keep_indices = [4, 6]
 
 def process_fishes(fishes, keep_indices, derived_param_derivss = [s8_ders, S8_ders, omm_ders, sigmaomm_ders]):
     fishes_inv = [inv(fish) for fish in fishes]
@@ -393,24 +395,24 @@ if __name__ == "__main__":
 
     # labels = ["CMB", "Galaxy", "CMB + Galaxy"]
 
-    # cov_matrices = [
-    #     inv(visp_c_reduced),
-    #     inv(visb_c_reduced),
-    #     inv(visb_c_reduced + visp_c_reduced)
-    # ]
-
-    # labels = ['cmb powerp', 'cmb bisp', 'cmb power + bisp']
-
-
     fish_matrices = [
-        visp_c + visb_c,
-        visp_s + visb_s,
-        visp_f,
-        visb_f,
-        visb_f + visp_f
+        toshiya_prior + 1.5 * visp_c,
+        toshiya_prior + 1.5 * visb_c,
+        toshiya_prior + 1.5 * visb_c + 1.5 * visp_c
     ]
 
-    labels = ['CMB + Gal Powersp', 'CMB + Gal Bisp', 'CMB Power- + Bisp', 'Gal Power- + Bisp', 'CMB + Gal Power- + Bisp']
+    labels = ['cmb powerp', 'cmb bisp', 'cmb power + bisp']
+
+
+    # fish_matrices = [
+    #     visp_c + visb_c,
+    #     visp_s + visb_s,
+    #     visp_f,
+    #     visb_f,
[…]

### 2025-05-08 paper/playing_with_fish.py
paper changes

@@ -35,8 +35,7 @@ visp_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_both.txt')
 ######################
 # Prior Matrices #
 ######################
-
-planck = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_c_planck.txt')
+planck = np.loadtxt(fisher_matrices_dir + '/fish_mat_powerspec_c_planck_temp.txt')
 
 planck_prior = np.diag([12 / 0.6**2, 1/(0.0005)**2, 12 / 1**2, 1/(0.02)**2, 1e6, 1e19, 1e6])
 
@@ -146,9 +145,9 @@ sigmaomm_ders = [0,
 
 
 # keep_indices = [0, 7, 9]
-# keep_indices=[10]
+keep_indices=[10]
 # keep_indices = [0, 1, 2, 3, 4, 5, 6]
-keep_indices = [4, 6]
+#keep_indices = [4, 6]
 
 def process_fishes(fishes, keep_indices, derived_param_derivss = [s8_ders, S8_ders, omm_ders, sigmaomm_ders]):
     fishes_inv = [inv(fish) for fish in fishes]
@@ -163,11 +162,11 @@ def process_fishes(fishes, keep_indices, derived_param_derivss = [s8_ders, S8_de
     
     return fishes_inv
 
-# print(process_fishes(
-#     [planck_prior, planck, planck_p_prior],
-#     [10],   
-#     derived_param_derivss = [s8_ders, S8_ders, omm_ders, sigmaomm_ders]
-#     ))
+print(np.sqrt(process_fishes(
+    [planck_prior, planck, planck_p_prior],
+    [10],   
+    derived_param_derivss = [s8_ders, S8_ders, omm_ders, sigmaomm_ders]
+    )))
 
 ######### PLOTS #########
 
@@ -414,13 +413,13 @@ if __name__ == "__main__":
 
     # labels = ['CMB + Gal Powersp', 'CMB + Gal Bisp', 'CMB Power- + Bisp', 'Gal Power- + Bisp', 'CMB + Gal Power- + Bisp']
 
-    # cov_matrices = [
-    #     inv(planck_reduced),
-    #     inv(planck_prior_reduced),
-    #     inv(planck_prior_reduced + planck_reduced)
-    # ]
+    fish_matrices = [
+        planck,
+        planck_prior,
+        planck_prior + planck
+    ]
 
[…]

### 2025-05-08 paper/tempCodeRunnerFile.py

@@ -1,4 +1,5 @@
-visb_c, visb_s, visb_f, visp_c, visp_s, visp_f, planck, planck_p_prior = append_row_column(
-#     [visb_c, visb_s, visb_f, visp_c, visp_s, visp_f, planck, planck_p_prior],
-#     s8_ders
-# )
+print(process_fishes(
+    [planck_prior, planck, planck_p_prior],
+    [10],   
+    derived_param_derivss = [s8_ders, S8_ders, omm_ders, sigmaomm_ders]
+    ))

### 2025-05-08 paper/playing_with_fish.py
misc

@@ -35,7 +35,7 @@ visp_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_both.txt')
 ######################
 # Prior Matrices #
 ######################
-planck = np.loadtxt(fisher_matrices_dir + '/fish_mat_powerspec_c_planck_temp.txt')
+planck = np.loadtxt(fisher_matrices_dir + '/fish_mat_powerspec_c_planck.txt')
 
 planck_prior = np.diag([12 / 0.6**2, 1/(0.0005)**2, 12 / 1**2, 1/(0.02)**2, 1e6, 1e19, 1e6])
 

### 2025-05-08 code/Fisher_matrix_powerspectra.py
takada jain fisher spectra

@@ -25,13 +25,21 @@ def fisher_calc_wrapper_planck(args, tracers):
     # Fisher_mat_single(int lmin, int lminbin, int lmax, int triangle_step_size, int num_bispec_samples, char* par1, char* par2, int num_cores, char* type)
     if tracers == 'c':
         return vispy.Fisher_powersp_single(8, 400, b'c', par1 = pars[i], par2 = pars[j])
+    
+def fisher_calc_wrapper_takada_jain(args, tracers):
+    i, j = args
+    # Fisher_mat_full(int lmin, int lminbin, int lmax, int triangle_step_size, int num_bispec_samples, char* par1, char* par2, int num_cores)
+    # Fisher_mat_single(int lmin, int lminbin, int lmax, int triangle_step_size, int num_bispec_samples, char* par1, char* par2, int num_cores, char* type)
+    if tracers == 's':
+        # lmax should actually be 3000, but interpolation currently doesn't go that high
+        return vispy.Fisher_powersp_single(50, 2000, b's', par1 = pars[i], par2 = pars[j])
 
-for tracer in ['c']: #['c', 's', 'both']:
+for tracer in ['s']: #['c', 's', 'both']:
     mat = np.zeros((len(pars), len(pars)))
     for i, j in product(range(len(pars)), repeat = 2):
-        result = fisher_calc_wrapper_planck((i, j), tracer)
+        result = fisher_calc_wrapper_takada_jain((i, j), tracer)
         mat[i, j] = result
         mat[j, i] = result  # Symmetric assignment
 
-    np.savetxt(f'fisher_matrices/fish_mat_powersp_{tracer}_planck.txt', mat)
+    np.savetxt(f'fisher_matrices/fish_mat_powersp_{tracer}_takada_jain.txt', mat)
     print(mat)

### 2025-05-11 code/Fisher_calc_python_imp.c
cmb fish created

@@ -2210,18 +2210,20 @@ static PyObject *__pyx_builtin_sum;
 /* #### Code section: string_decls ### */
 static const char __pyx_k_T[] = "T";
 static const char __pyx_k_c[] = "c";
+static const char __pyx_k_e[] = "e";
 static const char __pyx_k_i[] = "i";
 static const char __pyx_k_j[] = "j";
 static const char __pyx_k_l[] = "l";
 static const char __pyx_k_s[] = "s";
+static const char __pyx_k_t[] = "t";
 static const char __pyx_k_x[] = "x";
 static const char __pyx_k__9[] = ", ";
 static const char __pyx_k_l1[] = "l1";
 static const char __pyx_k_l2[] = "l2";
 static const char __pyx_k_l3[] = "l3";
 static const char __pyx_k_np[] = "np";
-static const char __pyx_k__31[] = "*";
-static const char __pyx_k__51[] = "?";
+static const char __pyx_k__35[] = "*";
+static const char __pyx_k__56[] = "?";
 static const char __pyx_k_inv[] = "inv";
 static const char __pyx_k_mat[] = "mat";
 static const char __pyx_k_min[] = "min";
@@ -2288,6 +2290,7 @@ static const char __pyx_k_cov_mat_inv_th[] = "cov_mat_inv_th";
 static const char __pyx_k_condition_check[] = "condition_check";
 static const char __pyx_k_Condition_numbers[] = "Condition numbers:";
 static const char __pyx_k_Fisher_bisp_single[] = "Fisher_bisp_single";
+static const char __pyx_k_Fisher_powersp_cmb[] = "Fisher_powersp_cmb";
 static const char __pyx_k_asyncio_coroutines[] = "asyncio.coroutines";
 static const char __pyx_k_cline_in_traceback[] = "cline_in_traceback";
 static const char __pyx_k_num_bispec_samples[] = "num_bispec_samples";
@@ -2312,8 +2315,9 @@ static PyObject *__pyx_pf_22Fisher_calc_python_imp_4condition_check(CYTHON_UNUSE
[…]

### 2025-05-11 code/Fisher_calc_python_imp.pyx

@@ -273,6 +273,41 @@ def Fisher_powersp(lmin, lmax, par1 = b'snr', par2 = b'snr'):
     
     return result1
 
+def Fisher_powersp_cmb(lmin, lmax, par1 = b'snr', par2 = b'snr'):
+
+    result1 = 0
+
+    # l1 = l2 sum
+    for l in range(lmin, lmax + 1):
+        xyz_configs = ((b't', b't'), (b't', b'e'), (b'e', b'e'))
+
+        vec_l = np.zeros(3)
+        vec_r = np.zeros(3)
+
+        # SNR_case
+        if par1 == b'snr' and par2 == b'snr':
+            for config_num in range(len(xyz_configs)):
+                vec_r[config_num] = lps_f(l, xyz_configs[config_num][0], xyz_configs[config_num][1])
+                vec_l = vec_r
+        # Fisher matrix case
+        else:
+            for config_num in range(len(xyz_configs)):
+                vec_r[config_num] = lps_der(l, xyz_configs[config_num][0], xyz_configs[config_num][1], par2, dd)
+                vec_l[config_num] = lps_der(l, xyz_configs[config_num][0], xyz_configs[config_num][1], par1, dd)
+
+        cov_mat = np.zeros((3, 3))
+        for i, j in product(range(len(xyz_configs)), repeat=2):
+            xyz_config_i = xyz_configs[i]
+            xyz_config_j = xyz_configs[j]
+            for x in permutations((0, 1), 2):
+                cov_mat[i, j] += lps_f_obs(l, xyz_config_i[0], xyz_config_j[x[0]]) * lps_f_obs(l, xyz_config_i[1], xyz_config_j[x[1]])
+
+        condition_check(cov_mat)
+
+        result1 += (2 * l + 1) * vec_l.T @ np.linalg.solve(cov_mat, vec_r) # 2 l + 1 factor is new
+    
+    return result1
+
 def Fisher_powersp_single(lmin, lmax, tracer, par1 = b'snr', par2 = b'snr'):
 
     result1 = 0

### 2025-05-11 code/Fisher_matrix_powerspectra.py

@@ -33,13 +33,25 @@ def fisher_calc_wrapper_takada_jain(args, tracers):
     if tracers == 's':
         # lmax should actually be 3000, but interpolation currently doesn't go that high
         return vispy.Fisher_powersp_single(50, 2000, b's', par1 = pars[i], par2 = pars[j])
+    
+def fisher_calc_wrapper_cmb(args, tracers):
+    i, j = args
+    # Fisher_mat_full(int lmin, int lminbin, int lmax, int triangle_step_size, int num_bispec_samples, char* par1, char* par2, int num_cores)
+    # Fisher_mat_single(int lmin, int lminbin, int lmax, int triangle_step_size, int num_bispec_samples, char* par1, char* par2, int num_cores, char* type)
+    if tracers == 't':
+        return vispy.Fisher_powersp_single(lmin, lmax, b't', par1 = pars[i], par2 = pars[j])
+    if tracers == 'e':
+        return vispy.Fisher_powersp_single(lmin, lmax, b'e', par1 = pars[i], par2 = pars[j])
+    if tracers == 'both':
+        # lmax should actually be 3000, but interpolation currently doesn't go that high
+        return vispy.Fisher_powersp_cmb(2, 2000, par1 = pars[i], par2 = pars[j])
 
-for tracer in ['s']: #['c', 's', 'both']:
+for tracer in ['both']: #['c', 's', 'both']:
     mat = np.zeros((len(pars), len(pars)))
     for i, j in product(range(len(pars)), repeat = 2):
-        result = fisher_calc_wrapper_takada_jain((i, j), tracer)
+        result = fisher_calc_wrapper_cmb((i, j), tracer)
         mat[i, j] = result
         mat[j, i] = result  # Symmetric assignment
 
-    np.savetxt(f'fisher_matrices/fish_mat_powersp_{tracer}_takada_jain.txt', mat)
+    np.savetxt(f'fisher_matrices/fish_mat_powersp_{tracer}_cmb.txt', mat)
     print(mat)

### 2025-05-11 code/code_writer.py

@@ -18,14 +18,15 @@ cdef double[:] window_c_data_{cosm_par}_{pm}_{delta_delta_coeff} = np.zeros(chi_
 cdef double[:] window_s_data_{cosm_par}_{pm}_{delta_delta_coeff} = np.zeros(chi_num, dtype=np.float64)
 cdef double[:, :] mps_data_{cosm_par}_{pm}_{delta_delta_coeff} = np.zeros((k_num_fine, z_num_fine), dtype=np.float64)
 cdef double[:] z_at_chi_data_{cosm_par}_{pm}_{delta_delta_coeff} = np.zeros(chi_num, dtype=np.float64)
+cdef double[:, :] cmbps_{cosm_par}_{pm}_{delta_delta_coeff} = np.zeros((lmax_cmbps+1, 5), dtype=np.float64)
 
-cosm_par_{cosm_par}_{pm}_{delta_delta_coeff}, C_{cosm_par}_{pm}_{delta_delta_coeff}, a_data_{cosm_par}_{pm}_{delta_delta_coeff}, b_data_{cosm_par}_{pm}_{delta_delta_coeff}, c_data_{cosm_par}_{pm}_{delta_delta_coeff}, lps_cc_data_{cosm_par}_{pm}_{delta_delta_coeff}, lps_cs_data_{cosm_par}_{pm}_{delta_delta_coeff}, lps_ss_data_{cosm_par}_{pm}_{delta_delta_coeff}, scale_factor_data_{cosm_par}_{pm}_{delta_delta_coeff}, window_c_data_{cosm_par}_{pm}_{delta_delta_coeff}, window_s_data_{cosm_par}_{pm}_{delta_delta_coeff}, mps_data_{cosm_par}_{pm}_{delta_delta_coeff}, z_at_chi_data_{cosm_par}_{pm}_{delta_delta_coeff} = data_import_func('data_{cosm_par}_{pm}_{delta_delta_coeff}')
[…]

### 2025-05-11 code/cosm_setup.py

@@ -44,7 +44,7 @@ class lensing_spectra:
     self.mnu = mnu        # Sum of neutrino masses
     self.w0 = w0
     self.matter_bispec_fit_params = matter_bispec_fit_params # params of matter bispectrum fitting function as (a_1, ..., a_9)
-    self.k_nl_max = 1000
+    self.k_nl_max = 3000
 
     self.obtain_results()
 
@@ -83,7 +83,7 @@ class lensing_spectra:
     print('Obtaining cosmological results.')
 
     pars = camb.CAMBparams()
-    lmax = 1000
+    lmax = 3000
     pars.set_cosmology(H0=self.H0, ombh2=self.ombh2, omch2=self.omch2, mnu=self.mnu, omk=0, tau=0.063, neutrino_hierarchy='normal')
     pars.InitPower.set_params(As=self.As, ns=self.ns, r=0)
     pars.InitPower.pivot_scalar = 0.05
@@ -92,7 +92,7 @@ class lensing_spectra:
     pars.set_dark_energy(w=self.w0)
     self.pars = pars
 
-    kmax_interp = 1000
+    kmax_interp = 3000
 
     # power spectrum in units of Mpc^3, k in units of Mpc^-1, with input (z, k) on RHS 
     print('Creating linear mps')

### 2025-05-11 code/data_exporter_new.py

@@ -67,6 +67,16 @@ def data_export(folder_name, cosm_par, lps = True, a_create = True, b_create = T
         np.save(filepath + '/z_at_chi', [cosm_data.results.redshift_at_comoving_radial_distance(chi) for chi in chis])
         print(f'z at chi created to {folder_name}')
 
+        lmax = 3000
+        cmb_ps = cosm_data.results.get_lensed_scalar_cls(lmax = lmax, raw_cl=True)
+
+        cmb_ps_with_ls = np.zeros((np.shape(cmb_ps)[0], np.shape(cmb_ps)[1] + 1))
+        cmb_ps_with_ls[:, 0] = np.arange(lmax + 1)[:]
+        cmb_ps_with_ls[:, 1:] = cmb_ps[:, :]
+
+        np.save(filepath + '/cmb_ps_with_ls', cmb_ps_with_ls)
+        print(f'cmb_ps_with_ls created to {folder_name}')
+
     if lps:
         # lensing power spectrum (convergence, convergence) (1d: k)
         np.save(filepath + '/lensing_power_spectrum_cc', [cosm_data.lps(k, ('convergence', 'convergence')) for k in ks_log_fine])
@@ -150,7 +160,7 @@ delta_delta = 0.05
 # data_export(folder_name, cosm_par, lps = True, a_create = True, b_create = True, c_create = True, mps = True, rest = True, exp_par = exp_par)
 
 # allows control over which data to create
-which_to_create = [[True, True, True, True, True, True]]
+which_to_create = [[False, False, False, False, False, True]]
 
 exports = [['data_fiducial', fiducial_cosm_par, *create_settings] for create_settings in which_to_create]
 

### 2025-05-11 code/data_importer_new.pyx

@@ -27,14 +27,17 @@ def get_k_max():
     return k_max
 
 # specifies the noise that will be used for cmb lensing
-cdef int noise_type = 3
-# type values correspond to:
+cdef int cmb_noise_type = 2
+cdef int galaxy_noise_type = 2
+# CMB type values correspond to:
 # 0: S0 noise curves (old, remove)
 # 1: sigma = 1, Delta_P = 6 (stage 3 toshiya)
 # 2: sigma = 3, Delta_P = 1 (stage 4 toshiya)
 # 3: sigma = 5, Delta_T = 30, Delta_P = 52 (planck (double check))
 
 # (units are in microKelvin arcmin and arcmin)
+
+cdef int lmax_cmbps = 3000
 
 # SO noise
 cdef str folder_file_path = '/scratch/p319950/data/'
@@ -73,13 +76,14 @@ def data_import_func(folder_name):
     cdef double[:] window_s_data = np.load(filepath + '/window_func_s.npy')
     cdef double[:, :] mps_data = np.load(filepath + '/matter_power_spectrum.npy')
     cdef double[:] z_at_chi_data = np.load(filepath + '/z_at_chi.npy')
+    cdef double[:, :] cmbps = np.load(filepath + '/cmb_ps_with_ls.npy')
 
     # all 2d funcs are made such that order of arguments is (k, z)
     C = 0.
     with open(filepath + '/rho_bar', 'r') as file:
         C = float(file.read())
 
-    return cosm_par, C, a_data, b_data, c_data, lps_cc_data, lps_cs_data, lps_ss_data, scale_factor_data, window_c_data, window_s_data, mps_data, z_at_chi_data
+    return cosm_par, C, a_data, b_data, c_data, lps_cc_data, lps_cs_data, lps_ss_data, scale_factor_data, window_c_data, window_s_data, mps_data, z_at_chi_data, cmbps
 
 
 
@@ -94,16 +98,25 @@ cpdef double c(double k, double z, double[:, :] c_data) noexcept nogil:
 
 # without gil cython doesn't like strings bc they are python objects, you can use char* C objects (C like strings)
 
[…]

### 2025-05-11 code/import_code.py

@@ -13,14 +13,15 @@ cdef double[:] window_c_data_H_p_2m = np.zeros(chi_num, dtype=np.float64)
 cdef double[:] window_s_data_H_p_2m = np.zeros(chi_num, dtype=np.float64)
 cdef double[:, :] mps_data_H_p_2m = np.zeros((k_num_fine, z_num_fine), dtype=np.float64)
 cdef double[:] z_at_chi_data_H_p_2m = np.zeros(chi_num, dtype=np.float64)
+cdef double[:, :] cmbps_H_p_2m = np.zeros((lmax_cmbps+1, 5), dtype=np.float64)
 
-cosm_par_H_p_2m, C_H_p_2m, a_data_H_p_2m, b_data_H_p_2m, c_data_H_p_2m, lps_cc_data_H_p_2m, lps_cs_data_H_p_2m, lps_ss_data_H_p_2m, scale_factor_data_H_p_2m, window_c_data_H_p_2m, window_s_data_H_p_2m, mps_data_H_p_2m, z_at_chi_data_H_p_2m = data_import_func('data_H_p_2m')
+cosm_par_H_p_2m, C_H_p_2m, a_data_H_p_2m, b_data_H_p_2m, c_data_H_p_2m, lps_cc_data_H_p_2m, lps_cs_data_H_p_2m, lps_ss_data_H_p_2m, scale_factor_data_H_p_2m, window_c_data_H_p_2m, window_s_data_H_p_2m, mps_data_H_p_2m, z_at_chi_data_H_p_2m, cmbps_H_p_2m = data_import_func('data_H_p_2m')
 
 cdef double lbs_H_p_2m(int k1, int k2, int k3, char* type1, char* type2, char* type3, int num_samples) noexcept nogil:
     return lbs(k1, k2, k3, type1, type2, type3, num_samples, C_H_p_2m, a_data_H_p_2m, b_data_H_p_2m, c_data_H_p_2m, scale_factor_data_H_p_2m, window_c_data_H_p_2m, window_s_data_H_p_2m, mps_data_H_p_2m, z_at_chi_data_H_p_2m)
 
 cdef double lps_H_p_2m(int l, char* type1, char* type2) noexcept nogil:
-    return lps(l, type1, type2, lps_cc_data_H_p_2m, lps_cs_data_H_p_2m, lps_ss_data_H_p_2m)
[…]

### 2025-05-11 paper/playing_with_fish.py
misc

@@ -33,15 +33,25 @@ visp_s = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_s.txt')
 visp_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_both.txt')
 
 ######################
-# Prior Matrices #
+# Planck #
 ######################
-planck = np.loadtxt(fisher_matrices_dir + '/fish_mat_powerspec_c_planck.txt')
+planck = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_c_planck.txt')
 
-planck_prior = np.diag([12 / 0.6**2, 1/(0.0005)**2, 12 / 1**2, 1/(0.02)**2, 1e6, 1e19, 1e6])
+planck_prior = np.diag([12 / 60**2, 1/(0.0005)**2, 12 / 1**2, 1/(0.02)**2, 1e6, 1e19, 1e6])
 
 planck_p_prior = planck + planck_prior
 
-toshiya_prior = np.diag([1 / 0.54**2, 1/(0.00015)**2, 1/(0.0014)**2, 1/(0.0042)**2, 0.05 / 0.06**2, 1e21, 1 / 0.5**2])
+viscmb_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_both_cmb.txt')
+
+######################
+# Takada & Jain #
+######################
+
+visp_s_takada_jain = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_s_takada_jain.txt')
+
+visb_s_takada_jain = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_s_takada_jain.txt')
+
+prior_takada_jain = np.diag([1 / 13**2, 1/(0.0002)**2, 0, 1/(0.008)**2, 1e8, 0, 0])
 
 # pars = [b'H', b'ombh2', b'omch2', b'ns', b'mnu', b'As', b'w0']
 
@@ -145,9 +155,9 @@ sigmaomm_ders = [0,
 
 
 # keep_indices = [0, 7, 9]
-keep_indices=[10]
+#keep_indices=[10]
 # keep_indices = [0, 1, 2, 3, 4, 5, 6]
-#keep_indices = [4, 6]
+keep_indices = [4, 6]
 
 def process_fishes(fishes, keep_indices, derived_param_derivss = [s8_ders, S8_ders, omm_ders, sigmaomm_ders]):
     fishes_inv = [inv(fish) for fish in fishes]
[…]

### 2025-05-11 paper/tempCodeRunnerFile.py

@@ -1,5 +1,5 @@
-print(process_fishes(
-    [planck_prior, planck, planck_p_prior],
-    [10],   
-    derived_param_derivss = [s8_ders, S8_ders, omm_ders, sigmaomm_ders]
-    ))
+print(np.sqrt(process_fishes(
+#     [planck_prior, planck, planck_p_prior],
+#     [10],   
+#     derived_param_derivss = [s8_ders, S8_ders, omm_ders, sigmaomm_ders]
+#     )))

### 2025-05-15 paper/paper_draft.tex
paper changes
misc

@@
 
-\author[1]{\textbf{Jonas Frugte}}
-\author[2]{\textbf{P. Daniel Meerburg}}
-
-\affil[1]{Independent Researcher, The Netherlands}
-\affil[2]{Van Swinderen Institute for Particle Physics and Gravity, University of Groningen, Nijenborgh 4, 9747 AG Groningen, The Netherlands}
+\author[ ]{\textbf{Jonas Frugte}}
+\author[1]{\textbf{P. Daniel Meerburg}}
+
+% \affil[1]{Independent Researcher, The Netherlands}
+\affil[1]{Van Swinderen Institute for Particle Physics and Gravity, University of Groningen, Nijenborgh 4, 9747 AG Groningen, The Netherlands}
 \affil[ ]{}
@@
 They will do so much more accurately than previous surveys, enabling tighter constraints on parameters and progressing our understanding of physics as a whole.
-Typically, the main value of interest is the power spectrum (or equivalently, two point function) of the lensing potential.
+Typically, the statistic of interest is the power spectrum (or equivalently, two point function) of the lensing potential.
 This can be measured in terms of the lensing convergence $\kappa$, lensing shear $\gamma$, or lensing potential $\psi$.
-In the weak lensing regime they all contain the same amount of information and can be easily converted into one another, see section \ref{sec:weaklensstats}.
-For the CMB the lensing potential is generally measured directly \cite{Planck2018Lensing}, while with galaxies one instead measures the shear, as this can be deduced from the ellipiticity distribution of the observed galaxies \cite{HoekstraJain2008}.
+In the weak lensing regime they all contain the same amount of information and can be trivially converted into one another, see appendix \ref{sec:weaklensstats}.
+In CMB surveys the temperature and polarization anisotropies are used to estimate the lensing potential \cite{Planck2018Lensing}, while with galaxies one instead calculates lensing effects by measuring the lensing shear, which can be deduced from the ellipiticity distribution of the observed galaxies \cite{HoekstraJain2008}.
 
@@
 In addition to looking at the power spectrum, a natural next step is thus to measure the lensing bispectrum.
-This has already been done for signals from galaxies \cite{vanWaerbeke2002}, but no proper detection has been done of the CMB lensing bispectrum.
+This has already been done for signals from galaxies \cite{vanWaerbeke2002}, but we do not yet have any proper detection of the CMB lensing bispectrum.
 Upcoming surveys such as the ones mentioned earlier may detect this signal.
-Therefore it is relevant to know the detectability of, and parameter constraints possible from, all 4 signals: CMB lensing power- and bispectra and galaxy lensing power- and bispectra.
-Notably, this includes looking at cross correlation as well.
-
-In this article we aim to answer this question in the context of experimental parameters similar to those of next generation surveys such as mentioned earlier.
+Therefore it is timely to consider the detectability, as well as ability to constrain cosmological parameter of CMB lensing power- and bispectra, galaxy lensing power- and bispectra, and cross correlations between the two weak lensing observables.
+
+In this paper we aim to answer this question in the context of experimental parameters similar to those of next generation surveys.
 We are especially interested in seeing if approximate parameter degeneracies can be broken by combining all data.
-Calculating the covariance matrix of unbiased optimal estimators of the cosmological parameters can be done by combining cosmological simulations (in our case we use the CAMB package \cite{Lewis2000}) with a Fisher matrix analysis.
+Calculating the covariance matrix of unbiased optimal estimators of the cosmological parameters can be done by combining a Boltzmann code for cosmological perturbations with a Fisher matrix analysis.
 See figure \ref{fig:flowchart} for an overview of the calculation done in this paper.
-The parameters looked at are the Hubble constant, $H_0$, physical baryon density, $\Omega_b h^2$, physical cold dark matter density, $\Omega_c h^2$, scalar spectral index, $n_s$, amplitude of primordial scalar fluctuations, $A_s$, sum of neutrino masses $\sum m_\nu$, and dark energy equation of state parameter, $w_0$.
-Though, we will find that some of the parameters are poorly constrained even when combining all 4 signals and thus omit them from the analysis in the results.
-As a secondary purpose, this paper also includes a derivation for the Fisher matrix formalism in the non-trivial context of bispectra with multiple tracers.
+We look at $\lambda$CDM parameters as well as the sum of neutrino masses $\sum m_\nu$ and dark energy equation of state parameter, $w_0$.
+
+Usefull derivations relevant to this paper can be found in ose, this paper also includes a derivation for the Fisher matrix formalism in the non-trivial context of bispectra with multiple tracers.
 The final formula can already be found in the literature (see e.g.\ the appendix of \cite{Kalaja_2021}), but the authors of this paper were not able to find a derivation and hence included it.
 
-The structure of this paper is as follows: we first review the theory of weak lensing and weak lensing statistics in sections \ref{sec:weaklensing} and \ref{sec:weaklensstats}.
-We then provide technical details of the calculation in section \ref{sec:calcdetails}, such as choices for noise and numerical approximations made in the code.
-In section \ref{sec:results} we present the results, consisting of signal to noise ratios and parameter constraints for different combinations of the CMB and galaxy lensing power- and bispectra.
-We conclude in section \ref{sec:discussion}.
-Appendix \ref{sec:fisher} contains a short review of the Fisher matrix formalism and gives a technical derivation of the formula used to calculate each Fisher matrix element in the case of bispectra measurements with multiple tracers.
-Finally, appendix \ref{sec:shear} explains how to relate the shear to the lensing potential in spherical harmonic space.
+The structure of this paper is as follows: we introduce relevant formulas for weak lensing statistics, nonlinear matter bispectrum, and fisher matrix formalism in section \ref{sec:theory}.
+We specify details such as the fiducial values of the parameters and noise models used in section \ref{sec:calcdetails}.
+Parameter constraints and signal to noise ratios are presented in section \ref{sec:results} and discussed in section \ref{sec:discussion}.
+The appendices review the derivations of the formulas for weak lensing power- and bispectra.
+We also derive the Fisher matrix formalism, in particular for the more involved case of measuring bispectra with multiple tracers.
 
@@
 \section{Theory}
+\label{sec:theory}
 \subsection{Weak lensing spectra}
@@
     C^{\psi_X\psi_Y}_l
-    & = \frac{9}{l^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)C^\delta(l/\chi,\eta_0-\chi),\\
-    B^{\psi_X\psi_Y\psi_Z}_{l_1l_2l_3} &= \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{l_1^2l_2^2l_3^2}\Omega_m^3H_0^6\\
+    & = \frac{9}{\ell^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)C^\delta(\ell/\chi,\eta_0-\chi),\\
+    B^{\psi_X\psi_Y\psi_Z}_{\ell_1\ell_2\ell_3} &= \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{l_1^2l_2^2l_3^2}\Omega_m^3H_0^6\\
     & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{l_i/\chi\}, \eta_0-\chi),
@@
 with $X, Y, Z \in \{\CMB, \gal\}$.
-$\Omega_m$ is the present day matter density parameter, $H_0$ is the present day Hubble constant, $a(\eta)$ is the scale factor, $\eta_0$ is the conformal time today, $\chi$ is the comoving radial distance, $\chi_*$ the distance to surface of last scattering, $P^\delta(k, \eta)$ the matter power spectrum, $B^\delta(k_1, k_2, k_3, \eta)$ the matter bispectrum, $W_X(\chi)$ the window function, and
+A derivation can be found in appendices \ref{sec:weaklensing} and \ref{sec:weaklensstats}.
+$\Omega_m$ is the present day matter density parameter, $H_0$ is the present day Hubble constant, $a(\eta)$ is the scale factor, $\eta_0$ is the conformal time today, $\chi$ is the comoving radial distance, $\chi_*$ the distance to surface of last scattering, $C^\delta(k, \eta)$ the matter power spectrum, $B^\delta(k_1, k_2, k_3, \eta)$ the matter bispectrum, $W_X(\chi)$ the window function, and
 $$
@@
     l_1&l_2&l_3 \\ m_1 & m_2 & m_3
-\end{pmatrix}
+\end{pmatrix},
 $$ 
@@
 \end{equation*}
-with $p_X$ the radius distribution of the radiation source. For the CMB we take $p(\chi) = \delta(\chi - \chi_*)$. Notice that we can also look at cross correlation between CMB and galaxy lensing.
-
-For a derivation of the lensing potential and its power- and bispectra, see appendices \ref{sec:weaklensing} and \ref{sec:weaklensstats}.
+with $p_X$ the radius distribution of the radiation source. For the CMB we take $p(\chi) = \delta(\chi - \chi_*)$.
 
@@
 \begin{equation}
-    B^\delta(k_1, k_2, k_3, \chi) = 2 F_2(k_1, k_2, z) C^\delta(k_1, z) C^\delta(k_2, z) + \text{2 perm.}
+    B^\delta(k_1, k_2, k_3, \chi) = 2 F_2(k_1, k_2, z) C^\delta(k_1, z) C^\delta(k_2, z) + \text{2 perm},
 \end{equation}
@@
 \begin{equation*}
-    F_{\alpha\beta} = \sum_{l} \sum_{XY}\sum_{X'Y'}(2l+1)\partial_\alpha C_l^{XY} (C^{-1})_l^{XX'}(C^{-1})_l^{YY'} \partial_\beta C^{X'Y'}_l
+    F_{\alpha\beta} = \sum_{l} \sum_{XY}\sum_{X'Y'}(2l+1)\partial_\alpha C_l^{XY} (C^{-1})_l^{XX'}(C^{-1})_l^{YY'} \partial_\beta C^{X'Y'}_l,
 \end{equation*}
@@
 % for SO, CMB - S4, Euclid, LSST
-We chose our fiducial cosmology in line with the results from the Planck mission \cite{planckresults}.
-\begin{align*}
-    H_0 &= 67.4 \, \text{km/s/Mpc}, && \text{(Hubble constant)} \\
-    \Omega_b h^2 &= 0.0223, && \text{(Physical baryon density parameter)} \\
-    \Omega_c h^2 &= 0.119, && \text{(Physical cold dark matter density parameter)} \\
-    n_s &= 0.965, && \text{(Scalar spectral index)} \\
-    A_s &= 2.13 \times 10^{-9}, && \text{(Amplitude of primordial scalar fluctuations)} \\
-    \sum m_\nu &= 0.06 \, \text{eV}, && \text{(Sum of neutrino masses)} \\
-    w_0 &= -1. && \text{(Dark energy equation of state parameter)}
-\end{align*}
+We chose our fiducial cosmology in line with the results from the Planck mission \cite{planckresults}, see table \ref{tab:fiducialpars}.
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{|c|c|c|}
+        \hline
+        \textbf{parameter} & \textbf{notation} & \textbf{fiducial value} \\
+        \hline
+        Hubble constant & $H_0$ & $67.4 \, \text{km/s/Mpc}$ \\
+        \hline
+        Physical baryon density parameter & $\Omega_b h^2$ & $0.0223$ \\
+        \hline
+        Physical cold dark matter density parameter & $\Omega_c h^2$ & $0.119$\\
+        \hline
+        Scalar spectral index & $n_s$ & 0.965 \\
+        \hline
+        Amplitude of primordial scalar fluctuations & $A_s$ & $2.13 \times 10^{-9}$ \\
+        \hline
+        Sum of neutrino masses & $\sum m_\nu$ & $0.06$\\
+        \hline
+        Dark energy equation of state parameter & $w_0$ & -1 \\ \hline
+    \end{tabular}
+    \caption{Cosmological parameters considered ($\Lambda \text{CDM} + w_0 + m_\nu$) and their fiducial values used.}
+    \label{tab:fiducialpars}
+\end{table}
 
@@
 
-This paper considers noise levels for stage 3 and stage 4 weak lensing surveys.
-Values can be found in table \ref{tab:noiselevels}.
+This paper considers noise levels for ``stage 3'' and ``stage 4'' weak lensing surveys.
+Values can be found in table \ref{tab:noiselevels}, including the experiments they are representative of.
 A comparison of noise power spectrum to lensing power spectrum can be found in figure \ref{fig:lpsplusnoise}.
@@
 
-CMB noise is estimated through the quadratic estimator developed in \cite{cmblensingestimator}.
-We do not take temperature measurements into account, as these are subleading for stage 4 experiments (REF).
+CMB lensing noise is estimated through the quadratic estimator developed in \cite{cmblensingestimator}.
+We consider experiments that only measure polarization noise.
 This is similarly done in \cite{Namikawa_2016}.
@@
     \caption{CMB (right) and galaxy (left) lensing potential power spectra compared to associated experimental noise.
-Current (stage 3) noise values are displayed as well as near future (stage 4) noi se values.
-The CMB lensing experiment does not measure temperature anisotropies, only polarization.
+Current (stage 3) noise values are displayed as well as near future (stage 4) noise values.
+The CMB lensing experiment uses only polarization.
 CMB noise values are chosen in accordance with \cite{Namikawa_2016}.
@@
 \section{Results}\label{sec:results}
-\subsection{SNR}
 Lensing power spectra are well detectable in nearly all cases. 
@@
 All other quantities should be well detectable for $l_{\max}\geq 500$.
-\begin{figure}
+\begin{figure}[h!]
     \includegraphics[width=\textwidth]{figures/snrplots.pdf}
@@
 
-\subsection{Parameter Constraints}
 Parameter constraints for the full set of $\Lambda$ CDM parameters are shown in figure \ref{fig:paramconstraintsall}.
-The combination $H_0$, $\sigma_8$, and $\Omega_m$ is particularly well constrained, constraints for these parameters alone (i.e.
-marginalized over all other parameters) are found in figure \ref{fig:paramconstraintstight}.
+The combination $H_0$, $\sigma_8$, and $\Omega_m$ is particularly well constrained.
+Constraints for these parameters alone are found in figure \ref{fig:paramconstraintstight}.
+Constraints are significantly improved by including bispectrum information in nearly all cases.
+This falls in line with a study in ref \cite{Namikawa_2016}, where a 33 percent and 35 percent improvement was found for $m_\nu$ and $w_0$ by measuring the CMB lensing bispectrum in addition to the powerspectrum.
+Constraints also improve significantly, though not quite as much, when combining CMB and galaxy lensing, compared to only measuring one of these signals.
+Some key observations are:
+\begin{itemize}
+    \item Lensing powerspectra alone are not sufficient to competitively constrain $\Lambda\mathrm{CDM} + w_0 + m_\nu$ even in the case of stage 4 surveys.
+    \item For low-enough noise values (in particular for stage 4 noise) the bispectra by themselves can competitively constrain the parameters and contain more information than the power spectra.
+If we denote the multipole moment at which the noise surpasses the powerspectrum signal as $\ell_{N>C}$, than the fisher matrices for the bispectra roughly scale as $F^{B}_{\alpha\beta}\propto \ell_{N>C}^3$ while for the powerspectra they scale as $F^{C}_{\alpha\beta}\propto \ell_{N>C}^2$ (with an extra multipole factor due to the $2\ell + 1$ factor in the powerspectrum estimator covariance).
+The powerspectrum is sufficiently bigger than the bispectrum that it contains more information given current day noise levels, however as surveys get better the bispectra may eventually contain more information because of the more favorable $\ell_{N>C}$ scaling.
+
+    \item The information from bispectra is sensitive to the cosmology in a different way than that of the powerspectra.
+The main parameter combinations where approximate degeneracies are broken are $(n_s,\Omega_b h^2),\;(\Omega_b h^2,H_0),\;(\Omega_c h^2,H_0),\;(\Omega_c h^2,\Omega_b h^2)\;(n_s,\Omega_c h^2),\;(m_\nu,H_0),\;(m_\nu,n_s),$ $(A_s,\Omega_b h^2)$.
+These means that even if stage 4 surveys surveys aren't able to reach the noise levels assumed in this paper, the bispectra will likely still lead to significantly tighter constraints.
+
+    \item Stage 4 CMB and galaxy weak lensing surveys have similar constraining power if both measure powerspectra and bispectra.
+Galaxy surveys are usually somewhat better, but not significantly so.
+Some degeneracies are broken but not quite as much as when looking at combining power and bispectra.
+The main parameter combinations where this happens are  $(n_s, H_0),\;(n_s,\Omega_b h^2),\;(m_\nu,H_0),\;(m_\nu,\Omega_s),\;(A_s,H_0)\;(w_0,A_s),\;(w_0, m_\nu),\;(w_0,\Omega_s)$.
+  \end{itemize}
 
@@
         \hline
-         Par                       &   CMB + Gal Powersp &   CMB + Gal Bisp &   CMB Power- + Bisp &   Gal Power- + Bisp &   CMB + Gal Power- + Bisp \\
+         Par                       &   CMB-L-$C$ + G-L-$C$ &   CMB-L-$C$ + G-L-$B$ &   CMB-L-$C$ + CMB-L-$B$ &   G-L-$C$ + G-L-$B$ &   all \\
         \hline
@@
         \end{tabular}
-\caption{Parameter constraints in percentage of the fiducial values for different combinations of weak lensing information. Below the double line are constraints for derived parameters.}
+\caption{Parameter constraints in percentage of the fiducial values for different combinations of weak lensing information.
+Below the double line are constraints for derived parameters.
+L stands for ``lensing'', G stands for ``galaxy'', $C$ stands for powerspectra, and $B$ for bispectra.}
 \label{tab:paramconstraintsall}
@@
 
+\pagebreak
+
 \section{Discussion and conclusion}\label{sec:discussion}
-For stage 4 surveys weak lensing matter and power spectra can all be measured with high significance.
-Additionally, parameter constraints are significantly improved by including bispectrum information in nearly all cases.
-This falls in line with work by Toshiya Namikawa \cite{Namikawa_2016}, where a 33 percent and 35 percent improvement was found for $m_\nu$ and $w_0$ by measuring the CMB ,lensing bispectrum in addition to the powerspectrum.
-In particular, from the confidence ellipses it can be seen that these improvements do not just come from additional information, but also because the bispectrum has different (approximate) parameter degeneracies compared to the power spectrum and thus they synergize when combined together.
+In this work, we have presented parameter forecasts from CMB lensing and galaxy lensing power- and bispectra, considering experimental setups similar to Stage 3 and Stage 4 surveys.
+Our analysis shows that while lensing power spectra are generally well detectable, the inclusion of bispectra offers significant improvements.
+Notably, shear bispectra are found to be measurable with both Stage 3 and Stage 4 experimental noise levels, whereas CMB lensing bispectra are semi-detectable with Stage 3 and well-detectable with Stage 4 noise levels.
+
+The primary impact of incorporating bispectrum information is a significant tightening of cosmological parameter constraints, particularly for parameters like $H_0$, $\sigma_8$, and $\Omega_m$, and in breaking key parameter degeneracies.
+We found that for sufficiently low noise levels, characteristic of Stage 4 surveys, the bispectra themselves can offer constraining power comparable to or even exceeding that of the power spectra, due to a more favorable scaling of information content with the maximum usable multipole\footnote{The multipole at which the noise powerspectrum starts exceeding the cosmological powerspectrum.}.
+Furthermore, the combination of CMB and galaxy lensing probes, especially when both power spectra and bispectra are utilized, further enhances constraints and aids in mitigating degeneracies.
+However, lensing power spectra alone, even for Stage 4 surveys, appear insufficient to competitively constrain an 8-parameter $\Lambda\text{CDM} + w_0 + \sum m_\nu$ model without priors.
 
 There are a number of limitations in this work.
-The most obvious is that the fitting formula used for the nonlinear matter bispectrum is not very accurate, only up to about 10 percent.
-Additionally, certain parameters that affect nonlinear large scale structure, such as neutrino masses, likely change the fitting parameters of the bispectrum when varied.
-This is currently not taking into account, making the derivatives less accurate.
-Another limitation is taking into account full sky effects.
-We assumed throughout this paper that surveys would measure multipoles as low as $l=2$, however our analytical formulas to calculate lensing spectra use the Limber approximation, which is only really accurate after $l=50$.
-Not using this approximation would make the calculation more complicated due to requireing knowledge of the uneeven time matter power and bispectra and computationally much longer due to having to evaluate a double integral.
-With the current methods it would thus not be feasible to calculate Fisher matrices without employing the Limber approximation, even if we would be able to calculate the unequel time matter spectra.
-Perhaps methods used in \cite{Chen:2021vba} could be used to solve this problem.
-Additionally, to calculate the Fisher matrices within a managable timeframe, the derivative of the lensing spectra with respect to cosmological parameters was done using a central difference formula, introducing errors of at most a few percent (CHECK EXACT VALUE LATER) (see appendix \ref{sec:derivatives} for further discussion) and functions such as the matter power spectrum, fitting parameters of the matter bispectrum, etc.
-were approximated using pregenerated data and interpolation, also introducing at most a few percent of error.
-Currently these errors are not significant because this paper only aims to give a rough / qualitative understanding of how usefull CMB and galaxy weak lensing spectra are, however if a similar calculation is done aiming to give precise quantitative results, these factors will need to be looked at again and improved.
-Furthermore, the current models for calculating lensing noise are on the basic side.
-Analysing real life data requires a better understanding of underlying systematics.
-For instance, determining the redshift distribution of galaxies is well known to be a difficult process (CITE SOMETHING HERE) so these errors would have to be taken into account as well.
-A better estimator can also be used.
-For example, the iterative estimator developed in CITE leads to a significantly better signal to noise for the CMB lensing bispectrum, as shown in \cite{Namikawa_2016}.
-Finally, for the covariance of the spectra estimators the Gaussian approximation was used, meaning we ignored any contributions not coming from standard Wick contractions, see appendix \ref{sec:fisher}.
-A next step could thus be to take these additional contributions into account, which could perhaps be done by running N body simulations and ray tracing simulations.
-Notably, this could potentially (depending on how accurate the simulations are) also allow us to calculate the lensing spectra much more accurately, i.e.
-without having to use a fitting formula for the matter bispectrum or Limber approximation.
+The limitations inherent to the fisher matrix formalism and to non physical noise models such as those considered in this paper are already well known.
+The main limitations specific to this work are:
+\begin{itemize}
+     \item \textbf{Nonlinear Matter Bispectrum Fitting Formula:} The fitting formula used for the nonlinear matter bispectrum (from \cite{bispfit}) has an estimated accuracy of only up to about 10 percent.
+This intrinsic inaccuracy in the model for $B^{\delta}$ will propagate to the lensing bispectra.
+     \item \textbf{Dependence of Fitting Parameters on Cosmology:} Certain cosmological parameters that affect nonlinear large-scale structure, such as neutrino masses, likely alter the fitting parameters of the matter bispectrum formula when varied.
+In the literature this is currently not considered, which can make the derivatives of the bispectrum with respect to these cosmological parameters less accurate.
+     \item \textbf{Limber Approximation and Sky Coverage:} We assumed throughout this paper that surveys would measure multipoles as low as $\ell = 2$.
+However, our analytical formulas to calculate lensing spectra utilize the Limber approximation, which is only robust for $\ell \gtrsim 50$.
+Avoiding the Limber approximation would significantly complicate calculations, requiring knowledge of unequal-time matter power and bispectra, and would be computationally much more demanding due to the need to evaluate double integrals instead of single integrals.
+With the methods used in this paper, recalculating Fisher matrices without the Limber approximation would not be feasible, even if unequal-time matter spectra were readily available.
+Methods such as those explored in reference \cite{Chen:2021vba} might offer a path to address this.
+
+     \item \textbf{Gaussian Estimator Assumption for Covariance:} As noted in appendix \ref{sec:fisher}, the Gaussian approximation was used for the covariance of the spectra estimators, i.e.
+we ignored any contributions not arising from standard Wick contractions.
+A next step could be to account for these additional contributions, perhaps through N-body and ray-tracing simulations.
+Notably, such simulations could potentially also allow for a more accurate calculation of the lensing spectra themselves, without relying on fitting formulas for the matter bispectrum or the Limber approximation, depending on simulation accuracy.
+
+    \item \textbf{CMB Lensing Noise Estimator} The CMB lensing noise was estimated via a quadratic estimator.
+While standard, such estimators may not be strictly optimal, especially with complex foregrounds or non-Gaussianities.
+Better estimators exist; for example, the iterative estimator developed in \cite{Smith_2012} can lead to a significantly better signal-to-noise ratio for the CMB lensing bispectrum, as shown in \cite{Namikawa_2016}.
+    \item \textbf{Numerical Derivatives and Interpolation:} To calculate the Fisher matrices within a manageable timeframe, derivatives of the lensing spectra with respect to cosmological parameters were computed using a central difference formula.
+This introduces errors (estimated at most a few percent, though larger for neutrino mass derivatives, see appendix \ref{sec:derivatives}).
+Functions such as the matter power spectrum and fitting parameters for the matter bispectrum were approximated using pre-generated data and interpolation, also introducing errors of at most a few percent.
+Currently, these numerical errors are not considered significant because this paper primarily aims to provide a qualitative understanding of the utility of CMB and galaxy weak lensing spectra.
+However, for precise quantitative results, these factors would need to be revisited and improved.
+\end{itemize}
+
+Despite these limitations, this work provides valuable insight into the potential of future weak lensing surveys.
+Addressing these limitations will be crucial for obtaining robust cosmological constraints from upcoming observational data.
 
@@
 
+\section{Papers compared to}
+Takada Jain 2005:
+\begin{itemize}
+    \item $ 50 < l < 3000$
+    \item powerspectrum + bispectrum
+    \item galaxy lensing
+    \item $\epsilon = 0.4$, $n_g = 100$, $f_{sky}=0.1$
+    \item derivative step: 5 percent.
+    \item priors:
+    \begin{gather*}
+        \sigma(\ln \Omega_B h^2) = 0.01\\
+        \sigma(n) = 0.008\\
+        \sigma(H_0) = 13
+        \sigma(m_\nu) = \infty
+    \end{gather*}
+    \item constraints obtained by paper:
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{lccc}
+          \hline
+          Quantity                & PS   & Bisp & PS+Bisp \\
+          \hline
+          $\sigma(\Omega_{\rm de})$ & 0.14 & 0.13 & 0.023   \\
+          $\sigma(\sigma_8)$        & 0.12 & 0.12 & 0.023   \\
+          $\sigma(w_0)$             & 4.0  & 3.6  & 1.1     \\
+          $\sigma(w_a)$             & 20   & 15   & 5.1     \\
+          \hline
+        \end{tabular}
+      \end{table}
+      \item constraints we got:
+      \begin{table}[h!]
+        \centering
+        \begin{tabular}{lccc}
+          \hline
+          Quantity                & PS   & Bisp & PS+Bisp \\
+          \hline
+          $\sigma(\sigma_8)$        & 0.09 & 0.09 & 0.05   \\
+          $\sigma(w_0)$             & 0.37  & 0.37  & 0.21     \\
+          \hline
+        \end{tabular}
+      \end{table}
+\end{itemize}
+
+Planck VII:
+\begin{itemize}
+    \item $ 8 < l < 400$
+    \item powerspectrum only
+    \item cmb lensing
+    \item noise: 5 beamwidth, $\Delta_T = 30$, $\Delta_P = 52$.
+    \item derivative step: ??
+    \item priors:
+    \begin{gather*}
+        \sigma(\Omega_b h^2) = 0.0005\\
+        \sigma(\Omega_c h^2) = [0.01, 0.99]\\
+        \sigma(n_s) = 0.02\\
+        \sigma(10^10A_s) = [1.91, 3.91]\\
+        \sigma(H_0) = [0.4, 1.0]\\
+        \sigma(m_\nu) = \infty\\
+        \sigma(w_0) = \infty
+    \end{gather*}
+    \item constraints obtained by paper:
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{lc}
+          \hline
+          Quantity & PS  \\
+          \hline
+          $\sigma(\sigma_8\Omega_m^{0.25})$ & 0.020 \\
+          \hline
+        \end{tabular}
+      \end{table}
+      \item constrain that we obtained: 0.0275
+      \item notes: actually different beamwidth for polarization and temperature
+\end{itemize}
+
 \end{document}

### 2025-05-15 paper/playing_with_fish.py

@@ -154,10 +154,10 @@ sigmaomm_ders = [0,
            (0.25 * sigma8 * omm**(-0.75))]
 
 
-# keep_indices = [0, 7, 9]
+keep_indices = [0, 7, 9]
 #keep_indices=[10]
-# keep_indices = [0, 1, 2, 3, 4, 5, 6]
-keep_indices = [4, 6]
+#keep_indices = [0, 1, 2, 3, 4, 5, 6]
+#keep_indices = [4, 6]
 
 def process_fishes(fishes, keep_indices, derived_param_derivss = [s8_ders, S8_ders, omm_ders, sigmaomm_ders]):
     fishes_inv = [inv(fish) for fish in fishes]
@@ -189,7 +189,7 @@ print(0.67**(-0.5) * np.sqrt(process_fishes(
 plt_scale = 0.5
 
 # new version that returns flipped version which I think is correct
-def confidence_ellipse(ax, mean, cov, color='blue', nstd=1.0, **kwargs):
+def confidence_ellipse(ax, mean, cov, color='blue', nstd=1.0, isbig=False, **kwargs):
     """
     Plot the 2D confidence ellipse of a Gaussian distribution given by mean & covariance
     onto the Matplotlib axes 'ax'.
@@ -197,12 +197,19 @@ def confidence_ellipse(ax, mean, cov, color='blue', nstd=1.0, **kwargs):
     - cov: (2x2) covariance matrix
     - color: color for the ellipse
     - nstd: number of standard deviations (1 => ~68% region, 2 => ~95%, etc.)
-    """
+    """        
     vals, vecs = np.linalg.eigh(cov)
     # Angle to rotate the ellipse
     angle = 45 - (np.degrees(np.arctan2(*vecs[:, 1][::-1])) - 45)
     # Major/minor axes
     width, height = 2 * nstd * np.sqrt(vals)
+    if isbig == True:
+        width_small, height_small = (0.5 * width, 0.5 * height)
+        ellip = Ellipse(
+        xy=mean, width=width_small, height=height_small,
+        angle=angle, edgecolor=color, facecolor='none', lw=1, linestyle = '--', **kwargs
+        )
+        ax.add_patch(ellip)
 
     ellip = Ellipse(
         xy=mean, width=width, height=height,
@@ -321,9 +328,13 @@ def plot_corner(
             for (cov, color) in zip(cov_matrices, colors):
[…]

### 2025-05-15 paper/vistest.py

@@ -1,0 +1,418 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.patches import Ellipse
+from scipy.stats import norm
+from numpy.linalg import inv
+import matplotlib.lines as mlines
+
+np.set_printoptions(
+    precision=1,
+    suppress=False, 
+    formatter={'float_kind': lambda x: f"{x:.2e}"}
+)
+
+fisher_matrices_dir = '/Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/code/fisher_matrices'
+
+######################
+# Bispectra Matrices #
+######################
+visb_c = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_c.txt')
+
+visb_s = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_s.txt')
+
+visb_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_both.txt')
+
+########################
+# Powerspectra Matrices
+########################
+visp_c = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_c.txt')
+
+visp_s = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_s.txt')
+
+visp_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_both.txt')
+
+######################
+# Planck #
+######################
+planck = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_c_planck.txt')
+
+planck_prior = np.diag([12 / 60**2, 1/(0.0005)**2, 12 / 1**2, 1/(0.02)**2, 1e6, 1e19, 1e6])
+
+planck_p_prior = planck + planck_prior
+
+viscmb_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_both_cmb.txt')
+
+######################
+# Takada & Jain #
+######################
+
+visp_s_takada_jain = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_s_takada_jain.txt')
+
+visb_s_takada_jain = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_s_takada_jain.txt')
+
+prior_takada_jain = np.diag([1 / 13**2, 1/(0.0002)**2, 0, 1/(0.008)**2, 1e8, 0, 0])
+
[…]

### 2025-05-18 paper/paper_draft.tex
paper changes

@@
-% !TEX root = /Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/paper/paper_draft.tex
+% !TEX root = ../paper/paper_draft.tex
 \documentclass[11pt]{article} % Choose a document class, like 'article', 'report', or 'revtex4-2' for scientific papers
 
-% Packages to enhance functionality
+\usepackage{jcappub}
+
 \usepackage[utf8]{inputenc}   % Allows UTF-8 character encoding
@@
 \geometry{margin=1in}         % 1-inch margins all around
-\usepackage{authblk}          % For author affiliations
 \usepackage{siunitx}          % For consistent typesetting of units
-\usepackage{subcaption}       % For subfigures within a figure
 \usepackage{float}            % To control float positioning (e.g., figures/tables)
@@
 \usepackage{booktabs}
-\usepackage{authblk}
 \usepackage{multirow} % needed for merging cells vertically
@@
 
-\usepackage[backend=biber, style=numeric]{biblatex}
-\addbibresource{/Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/paper/sources.bib}
+\usepackage[backend=bibtex]{biblatex}
+\addbibresource{sources.bib}
 
@@
 \setlength{\parskip}{1em}     % Add space between paragraphs
-
-\title{\huge \textbf{Parameter Forecasts From CMB Lensing and Galaxy Lensing Power- and Bispectra}}
-
-\author[ ]{\textbf{Jonas Frugte}}
-\author[1]{\textbf{P. Daniel Meerburg}}
-
-% \affil[1]{Independent Researcher, The Netherlands}
-\affil[1]{Van Swinderen Institute for Particle Physics and Gravity, University of Groningen, Nijenborgh 4, 9747 AG Groningen, The Netherlands}
-\affil[ ]{}
-\affil[ ]{\textit{E-mail:} \texttt{jonasfrugte@gmail.com}}
- % Removes the default date
+\title{Parameter Forecasts From CMB Lensing and Galaxy Lensing Power- and Bispectra}
+
+\author[1]{Jonas Frugte}
+\author[1]{P. Daniel Meerburg}
+
+\affiliation[1]{Van Swinderen Institute for Particle Physics and Gravity, University of Groningen, Nijenborgh 4, 9747 AG Groningen, The Netherlands}
+% \affil[ ]{}
+% \affil[ ]{\textit{E-mail:} \texttt{jonasfrugte@gmail.com}}
+
+\emailAdd{jonasfrugte@gmail.com}
+\emailAdd{p.d.meerburg@rug.nl}
+\abstract{}
 
 \begin{document}
+
 \maketitle
@@
 
-\begin{center}
-    \textbf{NB: Currently still in draft.}
-\end{center}
-\hrule
-
-\begin{abstract}
-    % Weak gravitational lensing of the Cosmic Microwave Background (CMB) and distant galaxies provides a powerful probe of the large-scale structure and cosmic expansion history.
-Upcoming Stage-IV surveys like the Simons Observatory, CMB-S4, LSST, and Euclid promise unprecedented precision, motivating detailed forecasts of their constraining power.
-While lensing power spectra are standard cosmological tools, non-Gaussianities induced by structure formation encode additional information in the bispectrum.
-In this work, we perform a comprehensive Fisher matrix forecast to quantify the detectability of these signals and the cosmological constraints achievable by combining power spectrum and bispectrum measurements from both CMB lensing and galaxy shear, including their cross-correlations.
-We model Stage-IV-like experimental specifications and forecast constraints on the standard $\Lambda$CDM parameters, the sum of neutrino masses ($\Sigma m_\nu$), and the dark energy equation of state ($w_0$).
-We demonstrate that incorporating bispectra significantly enhances constraints, reducing marginalized parameter uncertainties by up to $\sim \texttt{X}\%$ compared to power spectra alone.
-We project competitive constraints, including $\sigma(H_0) = \texttt{X.X~km/s/Mpc}$, $\sigma(\sigma_8) = \texttt{X.XXX}$, and $\sigma(\Sigma m_\nu) = \texttt{X.XX~eV}$.
-    \end{abstract}
-\pagebreak
 %\tableofcontents
@@
 A range of upcoming surveys, such as the survey by the Simons Observatory (SO) \cite{Ade2019}, the Legacy Survey of Space and Time (LSST) \cite{Ivezic2019}, and Euclid \cite{Laureijs2011} aim to measure weak lensing of the CMB and of galaxies.
-They will do so much more accurately than previous surveys, enabling tighter constraints on parameters and progressing our understanding of physics as a whole.
+These upcoming surveys are expected to achieve significantly higher accuracy than their predecessors, leading to tighter constraints on cosmological parameters and advancing our overall understanding of physics.
 Typically, the statistic of interest is the power spectrum (or equivalently, two point function) of the lensing potential.
 This can be measured in terms of the lensing convergence $\kappa$, lensing shear $\gamma$, or lensing potential $\psi$.
-In the weak lensing regime they all contain the same amount of information and can be trivially converted into one another, see appendix \ref{sec:weaklensstats}.
+In the weak lensing regime they all contain the same amount of information and can be directly converted into one another, see appendix \ref{sec:weaklensstats}.
 In CMB surveys the temperature and polarization anisotropies are used to estimate the lensing potential \cite{Planck2018Lensing}, while with galaxies one instead calculates lensing effects by measuring the lensing shear, which can be deduced from the ellipiticity distribution of the observed galaxies \cite{HoekstraJain2008}.
@@
 In addition to looking at the power spectrum, a natural next step is thus to measure the lensing bispectrum.
-This has already been done for signals from galaxies \cite{vanWaerbeke2002}, but we do not yet have any proper detection of the CMB lensing bispectrum.
+While the bispectrum has been measured for galaxy lensing signals \cite{vanWaerbeke2002}, a definitive detection of the CMB lensing bispectrum is still pending.
 Upcoming surveys such as the ones mentioned earlier may detect this signal.
@@
 
-Usefull derivations relevant to this paper can be found in ose, this paper also includes a derivation for the Fisher matrix formalism in the non-trivial context of bispectra with multiple tracers.
-The final formula can already be found in the literature (see e.g.\ the appendix of \cite{Kalaja_2021}), but the authors of this paper were not able to find a derivation and hence included it.
+Useful derivations relevant to this paper can be found in the appendices.
+In particular, they include a derivation for the Fisher matrix formalism for the non-trivial context of bispectra with multiple tracers.
+While the final formula is available in the literature (e.g., the appendix of \cite{Kalaja_2021}), this paper includes a derivation as the authors were unable to locate one elsewhere.
 
@@
 Radiation from cosmological objects is distorted due to gravitational lensing.
-In the vast majority of cases this lensing is a relatively weak effect and is thus referred to as weak lensing.
+Typically, this lensing is a relatively weak effect and is thus referred to as weak lensing.
 Weak lensing is quantified through the deflection field $\mathbf d(\hat {\mathbf n})$ which equals the difference between the observed angle of a point on the sky and the true (unlensed) angle.
-This field is the gradient of the lensing potential $\psi(\hat {\mathbf n})$ given as a weighted integral over the mass distribution along the line of sight between earth and the observed point.
+This field is the gradient of the lensing potential, $\psi(\hat {\mathbf n})$, which is expressed as a weighted integral of the mass distribution along the line of sight from the observer to the source.
 
@@
 In galaxy surveys, lensing alters the ellipiticities of observed galaxies.
-If a large enough number of galaxies are observed, this effect can be seperated from the intrinsic ellipiticities of the galaxies and thus the lensing potential can be estimated.
+If a large enough number of galaxies are observed, this effect can be separated from the intrinsic ellipiticities of the galaxies and thus the lensing potential can be estimated.
 
@@
 with $X, Y, Z \in \{\CMB, \gal\}$.
-A derivation can be found in appendices \ref{sec:weaklensing} and \ref{sec:weaklensstats}.
+A derivation can be found in appendices \ref{sec:weaklensing} and \ref{sec:weaklensstats}, or in the literature, see e.g.
+\cite{bartelmann2001weak}.
 $\Omega_m$ is the present day matter density parameter, $H_0$ is the present day Hubble constant, $a(\eta)$ is the scale factor, $\eta_0$ is the conformal time today, $\chi$ is the comoving radial distance, $\chi_*$ the distance to surface of last scattering, $C^\delta(k, \eta)$ the matter power spectrum, $B^\delta(k_1, k_2, k_3, \eta)$ the matter bispectrum, $W_X(\chi)$ the window function, and
@@
 \subsection{Nonlinear matter bispectrum}
-Determining the nonlinear matter powerspectrum can be done using numerical codes such as CAMB.
+Determining the nonlinear matter powerspectrum can be done using numerical codes such as CAMB \cite{Lewis2000}.
 The nonlinear matter bispectrum was calculated from the powerspectrum using a fitting formula based on perturbation theory in \cite{bispfit}.
@@
 \subsection{Fisher matrix analysis}
-SNR and parameter constraints are obtained through a Fisher matrix analysis as is standard.
+SNR and parameter constraints are obtained through a Fisher matrix analysis as is standard \cite{dodelson2020modern}.
 A full derivation of below statements can be found in appendix \ref{sec:fisher}.
@@
 \section{Calculation Details}\label{sec:calcdetails}
-% \subsection{Numerical Considerations}
-% % I used CAMB, how I took derivative, interpolation, fisher matrix sampling, matter bispectrum in terms of power spectrum
-
-% Running cosmological simulations to obtain the matter power spectra and other functions was done with the CAMB package (REF).
-
-% To speed up the calculation of the lensing powerspectra and especially the bispectra, an interpolation approach was used.
-Due to being 3 dimensional, it was not feasible to interpolate the lensing bispectrum directly.
-We instead chose to interpolate all 1 and 2 dimensional functions that the bispectra are based on and then calculate the bispectra with the interpolated functions instead.
-This significantly sped up the computation at the cost of small errors in the values in the bispectrum ($\sim 1\%$) when compared with the uninterpolated values.
-
-% The time complexity of the Fisher matrix scales with the maximum multipole considered, $l_{max}$, as $\mathcal O (l_{max}^3)$, leading to time issues for realistic values of $l_{max}$.
-We thus calculated the $l_1 < l_2 < l_3$ part of each Fisher matrix element (see the appendix) by sampling every step of 10 %(CHECK IF THIS IS STILL CORRECT BEFORE PUBLISHING)
-% in $l_1$ and $l_2$.
-Due to the discontinuos nature of the bispectrum (e.g.
-it vanishes immediately if $l_1 + l_2 + l_3$ is and odd number) we didn't sample $l_3$ values.
-This was done similarly in (REF).
-
-% \subsection{Choosing Experimental Parameters}
-
-% lmin, lmax, galaxy redshift distribution, fiducial values (!), noise levels, 
-% for SO, CMB - S4, Euclid, LSST
+% TODO: find proper references for stage 3 and stage 4 names as well as experimental parameters and real life surveys
 We chose our fiducial cosmology in line with the results from the Planck mission \cite{planckresults}, see table \ref{tab:fiducialpars}.
+For all parameter constraints we use a prior similar to the one used in \cite{Planck2018Lensing}, see table \ref{tab:fiducialpars}.
+This prior only restricts $\Omega_b h^2$ and $n_s$ significantly.
+The power spectra, in particular, exhibit complete insensitivity to certain parameters.
+Consequently, relying solely on them to constrain the entire cosmological model results in very poor constraints.
+Future lensing surveys will of course use information from surveys measuring other statistics such as galaxy distributions or primary CMB anisotropies, so not using any priors is not representative of what stage 4 experiments could achieve.
+Even a weak prior improves the constraining power of our power spectra considerably.
 \begin{table}[h!]
     \centering
-    \begin{tabular}{|c|c|c|}
-        \hline
-        \textbf{parameter} & \textbf{notation} & \textbf{fiducial value} \\
-        \hline
-        Hubble constant & $H_0$ & $67.4 \, \text{km/s/Mpc}$ \\
-        \hline
-        Physical baryon density parameter & $\Omega_b h^2$ & $0.0223$ \\
-        \hline
-        Physical cold dark matter density parameter & $\Omega_c h^2$ & $0.119$\\
-        \hline
-        Scalar spectral index & $n_s$ & 0.965 \\
-        \hline
-        Amplitude of primordial scalar fluctuations & $A_s$ & $2.13 \times 10^{-9}$ \\
-        \hline
-        Sum of neutrino masses & $\sum m_\nu$ & $0.06$\\
-        \hline
-        Dark energy equation of state parameter & $w_0$ & -1 \\ \hline
+    \begin{tabular}{|c|c|c|c|}
+        \hline
+        \textbf{parameter} & \textbf{notation} & \textbf{fiducial value} & \textbf{$\sigma$ of prior}\\
+        \hline
+        Hubble constant & $H_0$ & $67.4 \, \text{km/s/Mpc}$ & 17.3\\
+        \hline
+        Physical baryon density parameter & $\Omega_b h^2$ & $0.0223$ & 0.0005\\
+        \hline
+        Physical cold dark matter density parameter & $\Omega_c h^2$ & $0.119$ & $0.288$\\
+        \hline
+        Scalar spectral index & $n_s$ & 0.965 & 0.02 \\
+        \hline
+        Amplitude of primordial scalar fluctuations & $A_s$ & $2.13 \times 10^{-9}$ & $1\times 10^{-9}$ \\
+        \hline
+        Sum of neutrino masses & $\sum m_\nu$ & $0.06$ & $\infty$\\
+        \hline
+        Dark energy equation of state parameter & $w_0$ & -1 & $\infty$\\ \hline
     \end{tabular}
-    \caption{Cosmological parameters considered ($\Lambda \text{CDM} + w_0 + m_\nu$) and their fiducial values used.}
+    \caption{Cosmological parameters varied ($\Lambda \text{CDM} + w_0 + m_\nu$).
+We use a prior similar to the one used by the Planck weak lensing results \cite{Planck2018Lensing} which essentially only restricts $\Omega_bh^2$ and $n_s$ significantly.}
     \label{tab:fiducialpars}
 \end{table}
-
-% For the parameter constraints we took the multipole range to be from 2 ($l_{min}$) to 2000 ($l_{max}$).
 
@@
 
-\begin{table}[h]
+\begin{table}[h!]
     \centering
@@
 Galaxy lensing is determined by measuring lensing shear (see appendices \ref{sec:weaklensing} and \ref{sec:shear}).
-The noise in this measurement is dominated by scale independent shot noise and has associated noise power spectrum $N_l^{\text{shear}} = \sigma_{\text{rms}}^2 / n_g$.
+The noise in this measurement is dominated by scale independent shot noise and has associated noise power spectrum $N_l^{\text{shear}} = \sigma_{\text{rms}}^2 / n_g$, \cite{bartelmann2001weak}.
 $n_g$ is the amount of galaxies observed per unit angle.
@@
     \centering
-    \includegraphics[width=\textwidth]{/Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/code/plots/spectraplusnoise.pdf}
+    \includegraphics[width=\textwidth]{../code/plots/spectraplusnoise.pdf}
     \caption{CMB (right) and galaxy (left) lensing potential power spectra compared to associated experimental noise.
@@
 
+Derivatives are calculated with a central difference formula, more information can be found in appendix \ref{sec:derivatives}.
+Obtaining accurate derivatives was challenging.
+In order to check the accuracy of our results we calculated constraints for similar experimental parameters as in references \cite{Planck2018Lensing, Takada2003, Namikawa_2016} and found good agreement in all cases.
+
 \section{Results}\label{sec:results}
 Lensing power spectra are well detectable in nearly all cases. 
-The signal to noise ratios of the CMB and galaxy lensing bispectra v.s.
-the maximum multipole moment measured can be found in figure \ref{fig:snrplots}.
-Shear bispectra can be measured well by both stage 3 and stage 4 experiments, while CMB lensing is only semi detectable by a stage 3 experiment but well detectable by stage 4 experiments.
+The signal to noise ratios of the CMB and galaxy lensing bispectra versus the maximum multipole moment measured can be found in figure \ref{fig:snrplots}.
+Shear bispectra can be measured well by both stage 3 and stage 4 experiments, while CMB lensing is only marginally detectable by a stage 3 experiment but well detectable by stage 4 experiments.
+CMB $S/N$ results match those of reference \cite{Namikawa_2016}.
 %It can be seen that the CMB bispectrum signal to noise never exceeds 2 even for $l_{\max} = 2000$, meaning that this signal is not likely to be able to contribute to parameter constraints in future surveys.
@@
 
-Parameter constraints for the full set of $\Lambda$ CDM parameters are shown in figure \ref{fig:paramconstraintsall}.
-The combination $H_0$, $\sigma_8$, and $\Omega_m$ is particularly well constrained.
-Constraints for these parameters alone are found in figure \ref{fig:paramconstraintstight}.
-Constraints are significantly improved by including bispectrum information in nearly all cases.
-This falls in line with a study in ref \cite{Namikawa_2016}, where a 33 percent and 35 percent improvement was found for $m_\nu$ and $w_0$ by measuring the CMB lensing bispectrum in addition to the powerspectrum.
+Below we show parameter constraints for a selection of variables that have the potential to be particularly well constrained by weak lensing, i.e.
+$m_\nu$, $w_0$, $\sigma_8$, and $\Omega_m$.
+Constraints for $\Lambda$CDM parameters can be found in appendix \ref{sec:lambdacdmconstraints}.
+Figures \ref{fig:paramconstraintstightcmb} and \ref{fig:paramconstraintstightgal} show confidence ellipses from stage 4 CMB and galaxy lensing surveys, respectively.
+Table \ref{tab:paramconstraintstight} contains the forecasted constraints as uncertainties in percentage of the fiducial values.
+Despite picking priors that significantly increase powerspectrum constraining power, constraints are significantly improved by including bispectrum information in nearly all cases.
+Again in agreement with reference \cite{Namikawa_2016}, where a 33 percent and 35 percent improvement was found for $m_\nu$ and $w_0$, respectively, by measuring the CMB lensing bispectrum in addition to the powerspectrum.
 Constraints also improve significantly, though not quite as much, when combining CMB and galaxy lensing, compared to only measuring one of these signals.
-Some key observations are:
+Other key observations are:
 \begin{itemize}
-    \item Lensing powerspectra alone are not sufficient to competitively constrain $\Lambda\mathrm{CDM} + w_0 + m_\nu$ even in the case of stage 4 surveys.
+    \item Lensing powerspectra alone are not sufficient to competitively constrain $\Lambda\mathrm{CDM} + w_0 + m_\nu$ without priors even in the case of stage 4 surveys.
     \item For low-enough noise values (in particular for stage 4 noise) the bispectra by themselves can competitively constrain the parameters and contain more information than the power spectra.
@@
 The main parameter combinations where approximate degeneracies are broken are $(n_s,\Omega_b h^2),\;(\Omega_b h^2,H_0),\;(\Omega_c h^2,H_0),\;(\Omega_c h^2,\Omega_b h^2)\;(n_s,\Omega_c h^2),\;(m_\nu,H_0),\;(m_\nu,n_s),$ $(A_s,\Omega_b h^2)$.
-These means that even if stage 4 surveys surveys aren't able to reach the noise levels assumed in this paper, the bispectra will likely still lead to significantly tighter constraints.
+This means that even if stage 4 surveys aren't able to reach the noise levels assumed in this paper, the bispectra will likely still lead to significantly tighter constraints.
 
@@
 
+\begin{table}
+    \centering
+    \scriptsize
+    \begin{tabular}{|l|rrr|rrr|rrr|}
+        \hline
+        & \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
+        \hline
+         Par        &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
+        \hline
+         $m_\nu$    &    1262.54 &                    1630.18 &                              779.06 &     435.57 &                    1015.84 &                              268.14 &     304.87 &                     250.06 &                              169.29 \\
+         $w_0$      &      77.72 &                     103.75 &                               57.12 &      19.05 &                      14.94 &                                2.46 &      17.58 &                       9.37 &                                2.26 \\
+         $\sigma_8$ &      15.62 &                      13.64 &                               12.35 &       7.92 &                      14.67 &                                3.26 &       6.73 &                       3.47 &                                1.47 \\
+         $\Omega_m$ &      64.20 &                      57.93 &                               52.32 &      21.14 &                      18.26 &                                4.44 &      20.46 &                      10.21 &                                2.74 \\
+         \hline
+        \end{tabular}
+\caption{Parameter constraints in percentage of the fiducial values for different combinations of weak lensing information.
+$C$ stands for powerspectra, and $B$ for bispectra.
+All surveys are assumed to be stage 4 surveys with multipole range $2 \leq \ell \leq 2000$.
+We use the prior described in table \ref{tab:fiducialpars}.}
+\label{tab:paramconstraintstight}
+\end{table}
+
 \begin{figure}
     \centering
-    \tiny
-    \begin{tabular}{lrrrrr}
-        \hline
-         Par                       &   CMB-L-$C$ + G-L-$C$ &   CMB-L-$C$ + G-L-$B$ &   CMB-L-$C$ + CMB-L-$B$ &   G-L-$C$ + G-L-$B$ &   all \\
-        \hline
-         $H_0$                     &              102.87 &             1.94 &                9.63 &                5.22 &                      0.94 \\
-         $\Omega_b h^2$            &              234.33 &            21.39 &               25.53 &               20.30 &                     13.50 \\
-         $\Omega_c h^2$            &              127.06 &             2.43 &                3.96 &                3.13 &                      1.42 \\
-         $n_s$                     &               25.44 &             1.94 &                3.25 &                3.82 &                      1.51 \\
-         $m_\nu$                   &             3121.97 &           307.15 &              384.96 &              277.55 &                    178.07 \\
-         $A_s$                     &               69.24 &             6.39 &                7.96 &                8.44 &                      4.45 \\
-         $w_0$                     &              103.74 &             2.90 &               23.14 &               13.87 &                      2.33 \\
-         \hline
-         \hline
-         $\sigma_8$                &               31.82 &             3.85 &                9.16 &                4.56 &                      1.84 \\
-         $S_8$                     &               37.67 &             2.08 &                2.52 &                2.27 &                      1.14 \\
-         $\Omega_m$                &              172.99 &             5.65 &               27.29 &               14.45 &                      3.57 \\
-         $\sigma_8\Omega_m^{0.25}$ &               32.15 &             2.65 &                2.82 &                1.99 &                      1.29 \\
-        \hline
-        \end{tabular}
-\caption{Parameter constraints in percentage of the fiducial values for different combinations of weak lensing information.
-Below the double line are constraints for derived parameters.
-L stands for ``lensing'', G stands for ``galaxy'', $C$ stands for powerspectra, and $B$ for bispectra.}
-\label{tab:paramconstraintsall}
+    \includegraphics[width=0.7\textwidth]{figures/param_constraints_tight_cmb.pdf}
+    \caption{Parameter constraints and confidence ellipses for $\mu_\nu$, $w_0$, $\sigma_8$, and $\Omega_m$.
+Using ``stage 4'' noise with multipole range from $2$ to $2000$.
+The colored plots show CMB lensing power- and/or bispectrum constraints.
+Priors used can be found in table \ref{tab:fiducialpars}.
+The confidence ellipses are for $1\sigma$ certainty.
+They show approximate degeneracies in the information gained from a survey if they are ``stretched''.
+When the information of ellipses with degeneracies in different directions is combined, the degeneracies are removed and the constraints typically becomes much better on the relevant parameters.
+The black plots show the constraints using all lensing information, i.e.
+CMB lensing spectra, galaxy lensing spectram, and all cross spectra.}
+    \label{fig:paramconstraintstightcmb}
 \end{figure}
-
 \begin{figure}
     \centering
-    \includegraphics[width=\textwidth]{figures/param_constraints_all.pdf}
-    \caption{Parameter constraints and confidence ellipses for $\Lambda$CDM model as well as neutrino mass and dark energy equation of state parameter.
-Using ``stage 4'' noise with multipole range from $2$ to $2000$.
-Different weak lensing information combinations considered.
-No priors used.
-The confidence ellipses are for $1\sigma$ certainty.
-They show approximate degeneracies in the information gained from a survey if they are ``stretched''.
-When the information of ellipses with degeneracies in different directions is combined, the degeneracies are removed and the constraints typically becomes much better on the relevant parameters.}
-    \label{fig:paramconstraintsall}
+    \includegraphics[width=0.7\textwidth]{figures/param_constraints_tight_gal.pdf}
+    \caption{Same as figure \ref{fig:paramconstraintstightcmb}, except here the colored plots show galaxy lensing constraints.}
+    \label{fig:paramconstraintstightgal}
 \end{figure}
@@
 %     \end{tabular}
-% \caption{Same as table \ref{tab:paramconstraintsall}, except for a smaller selection of parameters well constrained by lensing.}
+% \caption{Same as table \ref{tab:paramconstraintstight}, except for a smaller selection of parameters well constrained by lensing.}
 % \label{tab:paramconstraintstight}
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figures/param_constraints_tight.pdf}
-    \caption{Same as \ref{fig:paramconstraintsall}, except showing constraints for a set of parameters that can be determined particularly accurately with weak lensing surveys.
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=\textwidth]{figures/param_constraints_tight.pdf}
+%     \caption{Same as \ref{fig:paramconstraintsall}, except showing constraints for a set of parameters that can be determined particularly accurately with weak lensing surveys.
 We chose Hubble's constant, $H_0$, baryonic + dark matter density, $\Omega_m$, and the root mean square density variation when the matter distribution is smoothed on a scale of 8 megaparsecs, commonly known as $\sigma_8$.}
-    \label{fig:paramconstraintstight}
-\end{figure}
+%     \label{fig:paramconstraintstight}
+% \end{figure}
 
@@
 \begin{itemize}
-     \item \textbf{Nonlinear Matter Bispectrum Fitting Formula:} The fitting formula used for the nonlinear matter bispectrum (from \cite{bispfit}) has an estimated accuracy of only up to about 10 percent.
+     \item The fitting formula used for the nonlinear matter bispectrum (from \cite{bispfit}) has an estimated accuracy of only up to about 10 percent.
 This intrinsic inaccuracy in the model for $B^{\delta}$ will propagate to the lensing bispectra.
-     \item \textbf{Dependence of Fitting Parameters on Cosmology:} Certain cosmological parameters that affect nonlinear large-scale structure, such as neutrino masses, likely alter the fitting parameters of the matter bispectrum formula when varied.
+     \item Certain cosmological parameters that affect nonlinear large-scale structure, such as neutrino masses, likely alter the fitting parameters of the matter bispectrum formula when varied.
 In the literature this is currently not considered, which can make the derivatives of the bispectrum with respect to these cosmological parameters less accurate.
-     \item \textbf{Limber Approximation and Sky Coverage:} We assumed throughout this paper that surveys would measure multipoles as low as $\ell = 2$.
+     \item We assumed throughout this paper that surveys would measure multipoles as low as $\ell = 2$.
 However, our analytical formulas to calculate lensing spectra utilize the Limber approximation, which is only robust for $\ell \gtrsim 50$.
@@
 
-     \item \textbf{Gaussian Estimator Assumption for Covariance:} As noted in appendix \ref{sec:fisher}, the Gaussian approximation was used for the covariance of the spectra estimators, i.e.
+     \item As noted in appendix \ref{sec:fisher}, the Gaussian approximation was used for the covariance of the spectra estimators, i.e.
 we ignored any contributions not arising from standard Wick contractions.
@@
 
-    \item \textbf{CMB Lensing Noise Estimator} The CMB lensing noise was estimated via a quadratic estimator.
+    \item The CMB lensing noise was estimated via a quadratic estimator.
 While standard, such estimators may not be strictly optimal, especially with complex foregrounds or non-Gaussianities.
 Better estimators exist; for example, the iterative estimator developed in \cite{Smith_2012} can lead to a significantly better signal-to-noise ratio for the CMB lensing bispectrum, as shown in \cite{Namikawa_2016}.
-    \item \textbf{Numerical Derivatives and Interpolation:} To calculate the Fisher matrices within a manageable timeframe, derivatives of the lensing spectra with respect to cosmological parameters were computed using a central difference formula.
+    \item To calculate the Fisher matrices within a manageable timeframe, derivatives of the lensing spectra with respect to cosmological parameters were computed using a central difference formula.
 This introduces errors (estimated at most a few percent, though larger for neutrino mass derivatives, see appendix \ref{sec:derivatives}).
@@
 Weak lensing of a point source can be quantified by looking at the deflection field $\mathbf d (\hat{\mathbf n}) = \mathbf \theta_{\text{obs}} - \mathbf \theta_{\text{true}}$, i.e.
-the (small) difference between the angle at which we see the object and the angle at which we would see the object had no lensing occured.
+the (small) difference between the angle at which we see the object and the angle at which we would see the object had no lensing occued.
 To first order in $\Psi_N$ and $\Phi_N$, the deflection is given as \cite{dodelson2020modern}
@@
     \centering
-    \includegraphics[height=\textheight]{/Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/code/plots/derivative_accuracy.png}
+    \includegraphics[height=\textheight]{../code/plots/derivative_accuracy.png}
     \caption{Central difference approximation of derivative of equilateral CMB weak lensing bispectrum (left) and powerspectrum (right) for differences as specified in table \ref{tab:cosmo-params-diff} as well as for differences $5\%$ and $10\%$ larger and smaller.}
@@
 
-\section{Papers compared to}
-Takada Jain 2005:
-\begin{itemize}
-    \item $ 50 < l < 3000$
-    \item powerspectrum + bispectrum
-    \item galaxy lensing
-    \item $\epsilon = 0.4$, $n_g = 100$, $f_{sky}=0.1$
-    \item derivative step: 5 percent.
-    \item priors:
-    \begin{gather*}
-        \sigma(\ln \Omega_B h^2) = 0.01\\
-        \sigma(n) = 0.008\\
-        \sigma(H_0) = 13
-        \sigma(m_\nu) = \infty
-    \end{gather*}
-    \item constraints obtained by paper:
+\section{$\Lambda$CDM constraints}
+\label{sec:lambdacdmconstraints}
+
+We do not expect to use weak lensing surveys to competitively constrain the main $\Lambda$CDM parameters.
+Despite this, we include these constraints for completeness in figures \ref{fig:paramconstraintsallcmb} and \ref{fig:paramconstraintsallgal} for CMB and galaxy lensing, respectively, and in table \ref{tab:paramconstraintsall}.
+CMB and galaxy lensing respectively.
+All survey parameters are the same as in section \ref{sec:results}.
 \begin{table}[h!]
     \centering
-        \begin{tabular}{lccc}
-          \hline
-          Quantity                & PS   & Bisp & PS+Bisp \\
-          \hline
-          $\sigma(\Omega_{\rm de})$ & 0.14 & 0.13 & 0.023   \\
-          $\sigma(\sigma_8)$        & 0.12 & 0.12 & 0.023   \\
-          $\sigma(w_0)$             & 4.0  & 3.6  & 1.1     \\
-          $\sigma(w_a)$             & 20   & 15   & 5.1     \\
+    \scriptsize
+    \begin{tabular}{|l|r|rrr|rrr|rrr|}
+        \hline
+        && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
+        \hline
+         Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
+        \hline
+         $H_0$          &   25.70 &      25.56 &                      24.62 &                               22.49 &       8.23 &                       7.02 &                                1.90 &       7.52 &                       3.56 &                                0.92 \\
+         $\Omega_b h^2$ &    2.23 &       2.23 &                       2.23 &                                2.23 &       2.23 &                       2.22 &                                2.21 &       2.22 &                       2.20 &                                2.19 \\
+         $\Omega_c h^2$ &  240.56 &       8.91 &                      11.16 &                                7.83 &       4.51 &                       3.22 &                                2.15 &       3.10 &                       2.01 &                                1.33 \\
+         $n_s$          &    2.07 &       2.06 &                       2.07 &                                2.05 &       2.03 &                       1.90 &                                0.88 &       1.44 &                       1.44 &                                0.67 \\
+         $A_s$          &   47.62 &      45.54 &                      44.79 &                               29.26 &      10.63 &                       4.57 &                                3.50 &       6.93 &                       4.01 &                                2.84 \\
         \hline
         \end{tabular}
+        \caption{Same as table \ref{tab:paramconstraintstight}, except for $\Lambda$CDM parameters.}
+        \label{tab:paramconstraintsall}
 \end{table}
-      \item constraints we got:
-      \begin{table}[h!]
+
+\begin{figure}[h!]
     \centering
-        \begin{tabular}{lccc}
-          \hline
-          Quantity                & PS   & Bisp & PS+Bisp \\
-          \hline
-          $\sigma(\sigma_8)$        & 0.09 & 0.09 & 0.05   \\
-          $\sigma(w_0)$             & 0.37  & 0.37  & 0.21     \\
-          \hline
-        \end{tabular}
-      \end{table}
-\end{itemize}
-
-Planck VII:
-\begin{itemize}
-    \item $ 8 < l < 400$
-    \item powerspectrum only
-    \item cmb lensing
-    \item noise: 5 beamwidth, $\Delta_T = 30$, $\Delta_P = 52$.
-    \item derivative step: ??
-    \item priors:
-    \begin{gather*}
-        \sigma(\Omega_b h^2) = 0.0005\\
-        \sigma(\Omega_c h^2) = [0.01, 0.99]\\
-        \sigma(n_s) = 0.02\\
-        \sigma(10^10A_s) = [1.91, 3.91]\\
-        \sigma(H_0) = [0.4, 1.0]\\
-        \sigma(m_\nu) = \infty\\
-        \sigma(w_0) = \infty
-    \end{gather*}
-    \item constraints obtained by paper:
-    \begin{table}[h!]
+    \includegraphics[width=0.5\textwidth]{figures/param_constraints_all_cmb.pdf}
+    \caption{Same as figure \ref{fig:paramconstraintstightcmb}, except for the standard $\Lambda$CDM parameters.}
+    \label{fig:paramconstraintsallcmb}
+\end{figure}
+\begin{figure}[h!]
     \centering
-        \begin{tabular}{lc}
-          \hline
-          Quantity & PS  \\
-          \hline
-          $\sigma(\sigma_8\Omega_m^{0.25})$ & 0.020 \\
-          \hline
-        \end{tabular}
-      \end{table}
-      \item constrain that we obtained: 0.0275
-      \item notes: actually different beamwidth for polarization and temperature
-\end{itemize}
+    \includegraphics[width=0.5\textwidth]{figures/param_constraints_all_gal.pdf}
+    \caption{Same as figure \ref{fig:paramconstraintstightgal}, except for the standard $\Lambda$CDM parameters.}
+    \label{fig:paramconstraintsallgal}
+\end{figure}
+
+% NOT SUPPOSED TO STAY IN THE PAPER, JUST FOR REFERENCE
+% \section{Papers compared to}
+% Takada Jain 2005:
+% \begin{itemize}
+%     \item $ 50 < l < 3000$
+%     \item powerspectrum + bispectrum
+%     \item galaxy lensing
+%     \item $\epsilon = 0.4$, $n_g = 100$, $f_{sky}=0.1$
+%     \item derivative step: 5 percent.
+%     \item priors:
+%     \begin{gather*}
+%         \sigma(\ln \Omega_B h^2) = 0.01\\
+%         \sigma(n) = 0.008\\
+%         \sigma(H_0) = 13
+%         \sigma(m_\nu) = \infty
+%     \end{gather*}
+%     \item constraints obtained by paper:
+%     \begin{table}[h!]
+%         \centering
+%         \begin{tabular}{lccc}
+%           \hline
+%           Quantity                & PS   & Bisp & PS+Bisp \\
+%           \hline
+%           $\sigma(\Omega_{\rm de})$ & 0.14 & 0.13 & 0.023   \\
+%           $\sigma(\sigma_8)$        & 0.12 & 0.12 & 0.023   \\
+%           $\sigma(w_0)$             & 4.0  & 3.6  & 1.1     \\
+%           $\sigma(w_a)$             & 20   & 15   & 5.1     \\
+%           \hline
+%         \end{tabular}
+%       \end{table}
+%       \item constraints we got:
+%       \begin{table}[h!]
+%         \centering
+%         \begin{tabular}{lccc}
+%           \hline
+%           Quantity                & PS   & Bisp & PS+Bisp \\
+%           \hline
+%           $\sigma(\sigma_8)$        & 0.09 & 0.09 & 0.05   \\
+%           $\sigma(w_0)$             & 0.37  & 0.37  & 0.21     \\
+%           \hline
+%         \end{tabular}
+%       \end{table}
+% \end{itemize}
+
+% Planck VII:
+% \begin{itemize}
+%     \item $ 8 < l < 400$
+%     \item powerspectrum only
+%     \item cmb lensing
+%     \item noise: 5 beamwidth, $\Delta_T = 30$, $\Delta_P = 52$.
+%     \item derivative step: ??
+%     \item priors:
+%     \begin{gather*}
+%         \sigma(\Omega_b h^2) = 0.0005\\
+%         \sigma(\Omega_c h^2) = [0.01, 0.99]\\
+%         \sigma(n_s) = 0.02\\
+%         \sigma(10^10A_s) = [1.91, 3.91]\\
+%         \sigma(H_0) = [0.4, 1.0]\\
+%         \sigma(m_\nu) = \infty\\
+%         \sigma(w_0) = \infty
+%     \end{gather*}
+%     \item constraints obtained by paper:
+%     \begin{table}[h!]
+%         \centering
+%         \begin{tabular}{lc}
+%           \hline
+%           Quantity & PS  \\
+%           \hline
+%           $\sigma(\sigma_8\Omega_m^{0.25})$ & 0.020 \\
+%           \hline
+%         \end{tabular}
+%       \end{table}
+%       \item constrain that we obtained: 0.0275
+%       \item notes: actually different beamwidth for polarization and temperature
+% \end{itemize}
 

### 2025-05-18 paper/fisherstuff.py

@@ -1,494 +1,0 @@
-import numpy as np
-import matplotlib
-import matplotlib.pyplot as plt
-from matplotlib.patches import Ellipse
-from scipy.stats import norm
-from numpy.linalg import inv
-import matplotlib.lines as mlines
-
-np.set_printoptions(
-    precision=1,
-    suppress=False, 
-    formatter={'float_kind': lambda x: f"{x:.2e}"}
-)
-
-fisher_matrices_dir = '/Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/code/fisher_matrices'
-
-######################
-# Bispectra Matrices #
-######################
-visb_c = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_c.txt')
-
-visb_s = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_s.txt')
-
-visb_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_both.txt')
-
-########################
-# Powerspectra Matrices
-########################
-visp_c = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_c.txt')
-
-visp_s = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_s.txt')
-
-visp_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_both.txt')
-
-######################
-# Prior Matrices #
-######################
-
-planck = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_c_planck.txt')
-
-planck_prior = np.diag([0.0, 1/(0.0005)**2, 0, 1/(0.02)**2, 0, 0, 0])
-
-# pars = [b'H', b'ombh2', b'omch2', b'ns', b'mnu', b'As', b'w0']
-
-
-########### NORMALIZATION ###########
-
-param_values_normalization = np.array([
-    67.4,         # H_0
-    0.0224,       # Omega_b h^2
-    0.120,        # Omega_c h^2
-    0.965,        # n_s
-    0.06,         # m_\nu
-    2.1e-9,       # A_s
-    -1.0          # w_0
-])
-
-normalization_matrix = np.outer(param_values_normalization, param_values_normalization)
-# If needed, multiply each matrix by this normalization_matrix.
-
-########### SIGMA8 ################
-
-def append_row_column(mats, vec):
-    """
-    For each matrix in 'mats':
-      - Enlarge the matrix by 1 row + 1 column
[…]

### 2025-05-18 paper/playing_with_fish.py

@@ -37,7 +37,7 @@ visp_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_both.txt')
 ######################
 planck = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_c_planck.txt')
 
-planck_prior = np.diag([12 / 60**2, 1/(0.0005)**2, 12 / 1**2, 1/(0.02)**2, 1e6, 1e19, 1e6])
+planck_prior = np.diag([12 / 60**2, 1/(0.0005)**2, 12 / 1**2, 1/(0.02)**2, 0.01, 1e18, 0.01])
 
 planck_p_prior = planck + planck_prior
 
@@ -154,9 +154,9 @@ sigmaomm_ders = [0,
            (0.25 * sigma8 * omm**(-0.75))]
 
 
-keep_indices = [0, 7, 9]
+#keep_indices = [4, 6, 7, 9]
 #keep_indices=[10]
-#keep_indices = [0, 1, 2, 3, 4, 5, 6]
+keep_indices = [0, 1, 2, 3, 5]
 #keep_indices = [4, 6]
 
 def process_fishes(fishes, keep_indices, derived_param_derivss = [s8_ders, S8_ders, omm_ders, sigmaomm_ders]):
@@ -172,11 +172,11 @@ def process_fishes(fishes, keep_indices, derived_param_derivss = [s8_ders, S8_de
     
     return fishes_inv
 
-print(0.67**(-0.5) * np.sqrt(process_fishes(
-    [planck_prior, planck, planck_p_prior],
-    [10],   
-    derived_param_derivss = [s8_ders, S8_ders, omm_ders, sigmaomm_ders]
-    )))
+# print(0.67**(-0.5) * np.sqrt(process_fishes(
+#     [planck_prior, planck, planck_p_prior],
+#     [10],   
+#     derived_param_derivss = [s8_ders, S8_ders, omm_ders, sigmaomm_ders]
+#     )))
 
 # print(0.1**(-0.5) * np.sqrt(process_fishes(
 #     [prior_takada_jain + visp_s_takada_jain, prior_takada_jain + visp_s_takada_jain, prior_takada_jain + visb_s_takada_jain + visb_s_takada_jain],
@@ -186,7 +186,7 @@ print(0.67**(-0.5) * np.sqrt(process_fishes(
 
 ######### PLOTS #########
 
-plt_scale = 0.5
+plt_scale = 1
 
[…]

### 2025-05-18 paper/tempCodeRunnerFile.py

@@ -1,5 +1,1 @@
-print(np.sqrt(process_fishes(
-#     [planck_prior, planck, planck_p_prior],
-#     [10],   
-#     derived_param_derivss = [s8_ders, S8_ders, omm_ders, sigmaomm_ders]
-#     )))
+keep_in

### 2025-06-04 paper/paper_draft.tex

@@
 \setlength{\parskip}{1em}     % Add space between paragraphs
-\title{Parameter Forecasts From CMB Lensing and Galaxy Lensing Power- and Bispectra}
+\title{Future Parameter constraints from weak lensing CMB and Galaxy Lensing Power- and Bi-spectra}
 
@@
 $P^{XY}(k, \eta, \eta')$ and $B^{XYZ}(k_1,k_2,k_3, \eta_1, \eta_2, \eta_3)$ correspond to bi and power spectra of fields $X$, $Y$, and $Z$ in Fourier space.
-$P_l^{XY}(\eta,\eta')$ and $B_{l_1l_2l_3}^{XYZ}(\eta_1, \eta_2, \eta_3)$ correspond to the multipole moments of the bi and power spectra of fields $X$, $Y$, and $Z$.
+$P_l^{XY}(\eta,\eta')$ and $B_{\ell_1\ell_2\ell_3}^{XYZ}(\eta_1, \eta_2, \eta_3)$ correspond to the multipole moments of the bi and power spectra of fields $X$, $Y$, and $Z$.
 If all times are the same we may instead write only one time argument and if the time is current time then we omit the time argument entirely.
@@
 
-Weak lensing has been used as a probe to constrain cosmological parameters since the 2000's \cite{kilbinger2015cosmology,Bacon2000,Kaiser2000,Waerbeke2000,Bartelmann2001}.
-A range of upcoming surveys, such as the survey by the Simons Observatory (SO) \cite{Ade2019}, the Legacy Survey of Space and Time (LSST) \cite{Ivezic2019}, and Euclid \cite{Laureijs2011} aim to measure weak lensing of the CMB and of galaxies.
-These upcoming surveys are expected to achieve significantly higher accuracy than their predecessors, leading to tighter constraints on cosmological parameters and advancing our overall understanding of physics.
-Typically, the statistic of interest is the power spectrum (or equivalently, two point function) of the lensing potential.
+Weak lensing has been used as a probe to constrain cosmological parameters for over a decade \cite{kilbinger2015cosmology,Bacon2000,Kaiser2000,Waerbeke2000,Bartelmann2001}.
+A range of upcoming surveys, such as the survey by the Simons Observatory (SO) \cite{Ade2019}, the Legacy Survey of Space and Time (LSST) \cite{Ivezic2019}, and Euclid \cite{Laureijs2011} aims to measure weak lensing of the cosmic microwave background (CMB) and galaxies respectively.
+These upcoming surveys are expected to achieve significantly higher accuracy than their predecessors, leading to tighter constraints on cosmological parameters and advancing our overall understanding of cosmology.
+Typically, the standard summary statistic of interest is the power spectrum (or equivalently, two-point function) of the lensing potential (as well as cross-correlations with galaxy clustering).
 This can be measured in terms of the lensing convergence $\kappa$, lensing shear $\gamma$, or lensing potential $\psi$.
-In the weak lensing regime they all contain the same amount of information and can be directly converted into one another, see appendix \ref{sec:weaklensstats}.
-In CMB surveys the temperature and polarization anisotropies are used to estimate the lensing potential \cite{Planck2018Lensing}, while with galaxies one instead calculates lensing effects by measuring the lensing shear, which can be deduced from the ellipiticity distribution of the observed galaxies \cite{HoekstraJain2008}.
+In the weak lensing regime they are all equivalent, in the sense that they can be directly converted into one another (appendix \ref{sec:weaklensstats}).
+In CMB surveys, the temperature and polarization anisotropies are used to reconstruct the lensing potential \cite{Hu_2002}, while with galaxies one instead calculates lensing effects by measuring the lensing shear, which can be deduced from the ellipticity distribution of the observed galaxies \cite{HoekstraJain2008}.
 
 Even with Gaussian initial conditions, there are significant amounts of non-Gaussianity in the weak lensing signal, especially in galaxy lensing due to it being a probe of the more recent distribution of matter \cite{Bernardeau1997,Takada2003}.
-In addition to looking at the power spectrum, a natural next step is thus to measure the lensing bispectrum.
+In addition to looking at the power spectrum, a natural next step is thus to measure weak lensing bispectra.
 While the bispectrum has been measured for galaxy lensing signals \cite{vanWaerbeke2002}, a definitive detection of the CMB lensing bispectrum is still pending.
-Upcoming surveys such as the ones mentioned earlier may detect this signal.
-Therefore it is timely to consider the detectability, as well as ability to constrain cosmological parameter of CMB lensing power- and bispectra, galaxy lensing power- and bispectra, and cross correlations between the two weak lensing observables.
-
-In this paper we aim to answer this question in the context of experimental parameters similar to those of next generation surveys.
-We are especially interested in seeing if approximate parameter degeneracies can be broken by combining all data.
-Calculating the covariance matrix of unbiased optimal estimators of the cosmological parameters can be done by combining a Boltzmann code for cosmological perturbations with a Fisher matrix analysis.
-See figure \ref{fig:flowchart} for an overview of the calculation done in this paper.
-We look at $\lambda$CDM parameters as well as the sum of neutrino masses $\sum m_\nu$ and dark energy equation of state parameter, $w_0$.
-
+Upcoming surveys, such as the ones mentioned earlier may be able to detect this signal.
+Therefore it is timely to consider the potential of higher order spectra to constrain cosmological parameters.
+
+In this paper, we aim to answer this question by adopting experimental parameters similar to those of next-generation surveys.
+We are especially interested in seeing if approximate parameter degeneracies can be broken by combining CMB and galaxy weak lensing power- and bi-spectra.
+%Calculating the covariance matrix of unbiased optimal estimators of the cosmological parameters can be done by combining a Boltzmann code for cosmological perturbations with a Fisher matrix analysis.
+%See figure \ref{fig:flowchart} for an overview of the calculation done in this paper. 
+We will consider $\Lambda$CDM parameters and extensions thereof. Of particular interest are the sum of neutrino masses $\sum m_\nu$ and dark energy equation of state parameter, $w_0$.
+
+% While the final formula is available in the literature (e.g., the appendix of \cite{Kalaja_2021}), this paper includes a derivation as the authors were unable to locate one elsewhere.
+
+The structure of this paper is as follows: we introduce relevant formulas for weak lensing statistics, nonlinear matter bispectrum, and Fisher matrix formalism in section \ref{sec:theory}.
+We specify details such as the fiducial values of the parameters and noise models used in section \ref{sec:calcdetails}.
+Parameter constraints and signal-to-noise ratios are presented in section \ref{sec:results} and discussed in section \ref{sec:discussion}.
 Useful derivations relevant to this paper can be found in the appendices.
 In particular, they include a derivation for the Fisher matrix formalism for the non-trivial context of bispectra with multiple tracers.
-While the final formula is available in the literature (e.g., the appendix of \cite{Kalaja_2021}), this paper includes a derivation as the authors were unable to locate one elsewhere.
-
-The structure of this paper is as follows: we introduce relevant formulas for weak lensing statistics, nonlinear matter bispectrum, and fisher matrix formalism in section \ref{sec:theory}.
-We specify details such as the fiducial values of the parameters and noise models used in section \ref{sec:calcdetails}.
-Parameter constraints and signal to noise ratios are presented in section \ref{sec:results} and discussed in section \ref{sec:discussion}.
-The appendices review the derivations of the formulas for weak lensing power- and bispectra.
-We also derive the Fisher matrix formalism, in particular for the more involved case of measuring bispectra with multiple tracers.
-
-All code written for this project is uploaded on github at \url{https://github.com/Jonas-Frugte/fisher_calc_weak_lensing}.
-
-\begin{figure}[h!]
-    \centering
-    \input{figures/calc_flowchart.tikz}
-    \caption{Flowchart of steps involved in calculating the minimum error/uncertainty in cosmological parameter estimates using convergence and shear power- and bi-spectra.}
-    \label{fig:flowchart}
-\end{figure}
-
-\section{Theory}
+
+All code used to derive the resukts in this paper is publically available on GitHub at \url{https://github.com/Jonas-Frugte/fisher_calc_weak_lensing}.
+
+% \begin{figure}[h!]
+%     \centering
+%     \input{figures/calc_flowchart.tikz}
+%     \caption{Flowchart of steps involved in calculating the minimum error/uncertainty in cosmological parameter estimates using convergence and shear power- and bi-spectra.}
+%     \label{fig:flowchart}
+% \end{figure}
+
+\section{Background}
 \label{sec:theory}
@@
 Radiation from cosmological objects is distorted due to gravitational lensing.
-Typically, this lensing is a relatively weak effect and is thus referred to as weak lensing.
-Weak lensing is quantified through the deflection field $\mathbf d(\hat {\mathbf n})$ which equals the difference between the observed angle of a point on the sky and the true (unlensed) angle.
-This field is the gradient of the lensing potential, $\psi(\hat {\mathbf n})$, which is expressed as a weighted integral of the mass distribution along the line of sight from the observer to the source.
-
-In the case of CMB surveys, lensing alters the statistical properties of the temperature and polarization fields and can thus be calculated by comparing to the expected unlensed signal, see e.g.
-\cite{cmblensingestimator}.
-In galaxy surveys, lensing alters the ellipiticities of observed galaxies.
-If a large enough number of galaxies are observed, this effect can be separated from the intrinsic ellipiticities of the galaxies and thus the lensing potential can be estimated.
+Due to the low density of the cosmic web, the average deflection of a photon as it propagates through the universe is relatively weak and is thus referred to as weak lensing.
+Weak lensing is quantified through the deflection field $\mathbf d(\hat {\mathbf n})$ which equals the difference between the observed angle of a point in the sky and the true (unlensed) angle.
+This field is the gradient of the lensing potential, $\psi(\hat {\mathbf n})$, which is expressed as a weighted integral of the mass distribution along the line-of-sight from the observer to the source.
+
+In the case of CMB surveys, lensing alters the statistical properties of the temperature and polarization fields and can thus be calculated by comparing the observed signal to the expected unlensed signal(see e.g.
+\cite{cmblensingestimator}).
+In galaxy surveys, lensing alters the ellipticities of observed galaxies.
+If a large enough number of galaxies are observed, this effect can be separated from the intrinsic ellipticities of the galaxies which allows us to estimate the lensing potential.
 
@@
 \begin{empheq}[box=\fbox]{align*}
-    C^{\psi_X\psi_Y}_l
-    & = \frac{9}{\ell^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)C^\delta(\ell/\chi,\eta_0-\chi),\\
-    B^{\psi_X\psi_Y\psi_Z}_{\ell_1\ell_2\ell_3} &= \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{l_1^2l_2^2l_3^2}\Omega_m^3H_0^6\\
-    & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{l_i/\chi\}, \eta_0-\chi),
+    C^{\psi_X\psi_Y}_\ell
+    & = \frac{9}{\ell^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)P^\delta(\ell/\chi,\eta_0-\chi),\\
+    B^{\psi_X\psi_Y\psi_Z}_{\el\ell_1\el\ell_2\el\ell_3} &= \sqrt{\frac{(2\ell_1 + 1)(2\ell_2 + 1)(2\ell_3 + 1)}{4\pi}} \begin{pmatrix} \ell_1 & \ell_2 & \ell_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{\ell_1^2\ell_2^2\ell_3^2}\Omega_m^3H_0^6\\
+    & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{\ell_i/\chi\}, \eta_0-\chi),
 \end{empheq}
 with $X, Y, Z \in \{\CMB, \gal\}$.
-A derivation can be found in appendices \ref{sec:weaklensing} and \ref{sec:weaklensstats}, or in the literature, see e.g.
+A derivation can be found in appendices \ref{sec:weaklensing} and \ref{sec:weaklensstats}, or in the literature; see, e.g.
 \cite{bartelmann2001weak}.
-$\Omega_m$ is the present day matter density parameter, $H_0$ is the present day Hubble constant, $a(\eta)$ is the scale factor, $\eta_0$ is the conformal time today, $\chi$ is the comoving radial distance, $\chi_*$ the distance to surface of last scattering, $C^\delta(k, \eta)$ the matter power spectrum, $B^\delta(k_1, k_2, k_3, \eta)$ the matter bispectrum, $W_X(\chi)$ the window function, and
+$\Omega_m$ is the present day matter density parameter.
+$H_0$ is the present-day Hubble constant.
+$a(\eta)$ is the scale factor.
+$\eta_0$ is the conformal time today.
+$\chi$ is the comoving radial distance.
+$\chi_*$ is the distance to surface of last scattering.
+$P^\delta(k, \eta)$ is the matter power spectrum.
+$B^\delta(k_1, k_2, k_3, \eta)$ is the matter bispectrum.
+$W_X(\chi)$ is the window function.
+Finally,
 $$
 \begin{pmatrix}
-    l_1&l_2&l_3 \\ m_1 & m_2 & m_3
+    \ell_1&\ell_2&\ell_3 \\ m_1 & m_2 & m_3
 \end{pmatrix},
 $$ 
-the Wigner 3-j symbol. The only factor that differs between CMB and galaxy lensing is the window function, defined as
+is the Wigner 3-j symbol. The difference between CMB and galaxy lensing is the window function, defined as
 \begin{equation*}
@@
 \end{equation*}
-with $p_X$ the radius distribution of the radiation source. For the CMB we take $p(\chi) = \delta(\chi - \chi_*)$.
+with $p_X$ the radius distribution of the radiation source.
+For the CMB we take $p(\chi) = \delta(\chi - \chi_*)$.
+The window function presents the distribution of redshifts at which a CMB photon or galaxy photon is deflected.
+As a result, the CMB lensing window function is broader and peaks at lower redshift compared to the galaxy window function.
+
+Our convention for defining the linear and non-linear matter power spectrum is
+\begin{equation*}
+    \langle \delta(\mathbf k, \eta) \delta(\mathbf{k}',\eta) \rangle = (2\pi)^3\delta(\mathbf k + \mathbf k ')P^\delta(\mathbf k, \eta),
+\end{equation*}
+and for the bispectrum it is 
+\begin{equation*}
+    \langle \delta(\mathbf k_1, \eta) \delta(\mathbf k_2, \eta)\delta(\mathbf k_3, \eta) \rangle = (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\delta(k_1,k_2,k_3, \eta),
+\end{equation*}
+with $\delta$ the fractional matter density perturbation field.
 
 \subsection{Nonlinear matter bispectrum}
-Determining the nonlinear matter powerspectrum can be done using numerical codes such as CAMB \cite{Lewis2000}.
-The nonlinear matter bispectrum was calculated from the powerspectrum using a fitting formula based on perturbation theory in \cite{bispfit}.
+
+Determining the nonlinear matter power spectrum can be done using numerical codes such as CAMB \cite{Lewis2000}.
+The nonlinear matter bispectrum was calculated from the power spectrum using a fitting formula based on perturbation theory in \cite{bispfit}.
 It is given by
 \begin{equation}
-    B^\delta(k_1, k_2, k_3, \chi) = 2 F_2(k_1, k_2, z) C^\delta(k_1, z) C^\delta(k_2, z) + \text{2 perm},
-\end{equation}
-where \( C^\delta \) is the nonlinear matter power spectrum\footnote{Compare to the tree level bispectrum where we instead use the linear powerspectrum.}, and the kernel \( F_2 \) is modified from the tree level result with factors \( a(k, z) \), \( b(k, z) \), and \( c(k, z) \):
+    B^\delta(k_1, k_2, k_3, \chi) = 2 F_2(k_1, k_2, z) P^\delta(k_1, z) P^\delta(k_2, z) + \text{2 perm},
+\end{equation}
+where \( C^\delta \) is the nonlinear matter power spectrum\footnote{Compare to the tree level bispectrum where we instead use the linear power spectrum.}, and the kernel \( F_2 \) is modified from the tree level result with factors \( a(k, z) \), \( b(k, z) \), and \( c(k, z) \):
 \begin{equation}
@@
 \begin{equation}
-    n_{\text{eff}} \equiv \frac{d \ln C^\delta_{\text{lin}}(k)}{d \ln k}.
+    n_{\text{eff}} \equiv \frac{d \ln P^\delta_{\text{lin}}(k)}{d \ln k}.
 \end{equation}
@@
 \begin{equation}
-    4 \pi k_{\text{NL}}^3 C^\delta_{\text{lin}}(k_{\text{NL}}, 0) = 1.
+    4 \pi k_{\text{NL}}^3 P^\delta_{\text{lin}}(k_{\text{NL}}, 0) = 1.
 \end{equation}
@@
 \subsection{Fisher matrix analysis}
-SNR and parameter constraints are obtained through a Fisher matrix analysis as is standard \cite{dodelson2020modern}.
-A full derivation of below statements can be found in appendix \ref{sec:fisher}.
-The auto and cross power spectrum Fisher matrix is given as
-\begin{equation*}
-    F_{\alpha\beta} = \sum_{l} \sum_{XY}\sum_{X'Y'}(2l+1)\partial_\alpha C_l^{XY} (C^{-1})_l^{XX'}(C^{-1})_l^{YY'} \partial_\beta C^{X'Y'}_l,
+The signal-to-noise (SNR) and parameter constraints are obtained through a Fisher matrix analysis as is standard in cosmology\cite{dodelson2020modern}.
+A full derivation of the equations used below can be found in appendix \ref{sec:fisher}.
+The auto- and cross- power spectrum Fisher matrix is given as
+\begin{equation*}
+    F_{\alpha\beta} = \sum_{\ell} \sum_{XY}\sum_{X'Y'}(2\ell+1)\partial_\alpha C_\ell^{XY} (C^{-1})_\ell^{XX'}(C^{-1})_\ell^{YY'} \partial_\beta C^{X'Y'}_\ell,
 \end{equation*}
@@
 \begin{equation*}
-    C_{l} := \begin{pmatrix}
-        C^{\psi_\kappa\psi_\kappa}_l & C^{\psi_\kappa\psi_\gamma}_l \\
-        C^{\psi_\kappa\psi_\gamma}_l & C^{\psi_\gamma\psi_\gamma}_l
+    C_{\ell} := \begin{pmatrix}
+        C^{\psi_\kappa\psi_\kappa}_\ell & C^{\psi_\kappa\psi_\gamma}_\ell \\
+        C^{\psi_\kappa\psi_\gamma}_\ell & C^{\psi_\gamma\psi_\gamma}_\ell
     \end{pmatrix}.
 \end{equation*}
-For the bispectra we instead get 
-\begin{equation*}
-    F_{\alpha\beta} = \sum_{l_1 \leq l_2 \leq l_3} \frac{\mathcal P _{l_1l_2l_3}}{6}
+For the bispectra we instead have 
+\begin{equation*}
+    F_{\alpha\beta} = \sum_{\ell_1 \leq \ell_2 \leq \ell_3} \frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6}
     \sum_{XYZ}\sum_{X'Y'Z'} 
-    \partial_\alpha B^{X Y Z}_{l_1 l_2 l_3} 
-    (C^{-1})^{X X'}_{l_1}
-    (C^{-1})^{Y Y'}_{l_2}
-    (C^{-1})^{Z Z'}_{l_3}
-    \partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3},
-\end{equation*}
-with $\mathcal P_{l_1l_2l_3}$ defined as the number of distinct permutations that can be made with $l_1l_2l_3$. When only considering autospectra the formulas simplify to
-\begin{equation*}
-    F_{\alpha\beta} = \sum_{l} \sum_{XY}\sum_{X'Y'}\frac{2l+1}{2}\frac{\partial_{\alpha}C_l^{XX} \partial_\beta C^{XX}_l}{(C_l^{XX})^2},
-\end{equation*}
-and (for bispectra)
-\begin{equation*}
-    F_{\alpha\beta} = \sum_{l_1 \leq l_2 \leq l_3} \frac{\mathcal P _{l_1l_2l_3}}{6}
-    \frac{\partial_\alpha B^{X X X}_{l_1 l_2 l_3} 
-    \partial_\beta B^{XXX}_{l_1 l_2 l_3}}{C^{XX}_{l_1}C^{XX}_{l_2}C^{XX}_{l_3}}.
+    \partial_\alpha B^{X Y Z}_{\ell_1 \ell_2 \ell_3} 
+    (C^{-1})^{X X'}_{\ell_1}
+    (C^{-1})^{Y Y'}_{\ell_2}
+    (C^{-1})^{Z Z'}_{\ell_3}
+    \partial_\beta B^{X' Y' Z'}_{\ell_1 \ell_2 \ell_3},
+\end{equation*}
+with $\mathcal P_{\ell_1\ell_2\ell_3}$ defined as the number of distinct permutations that can be made with $\ell_1\ell_2\ell_3$. When only considering auto spectra the formulas simplify to:
+\begin{equation*}
+    F_{\alpha\beta} = \sum_{\ell} \sum_{XY}\sum_{X'Y'}\frac{2\ell+1}{2}\frac{\partial_{\alpha}C_\ell^{XX} \partial_\beta C^{XX}_\ell}{(C_\ell^{XX})^2},
+\end{equation*}
+and (for bispectra, assuming Gaussian, diagonal, covariance)
+\begin{equation*}
+    F_{\alpha\beta} = \sum_{\ell_1 \leq \ell_2 \leq \ell_3} \frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6}
+    \frac{\partial_\alpha B^{X X X}_{\ell_1 \ell_2 \ell_3} 
+    \partial_\beta B^{XXX}_{\ell_1 \ell_2 \ell_3}}{C^{XX}_{\ell_1}C^{XX}_{\ell_2}C^{XX}_{\ell_3}}.
 \end{equation*}
@@
 % TODO: find proper references for stage 3 and stage 4 names as well as experimental parameters and real life surveys
-We chose our fiducial cosmology in line with the results from the Planck mission \cite{planckresults}, see table \ref{tab:fiducialpars}.
-For all parameter constraints we use a prior similar to the one used in \cite{Planck2018Lensing}, see table \ref{tab:fiducialpars}.
+Our fiducial cosmology is derived from the Planck results \cite{planckresults} (see table \ref{tab:fiducialpars}).
+For all parameter constraints we use a prior similar to the one used in \cite{Planck2018Lensing} (see table \ref{tab:fiducialpars}).
 This prior only restricts $\Omega_b h^2$ and $n_s$ significantly.
@@
 Consequently, relying solely on them to constrain the entire cosmological model results in very poor constraints.
-Future lensing surveys will of course use information from surveys measuring other statistics such as galaxy distributions or primary CMB anisotropies, so not using any priors is not representative of what stage 4 experiments could achieve.
+Future lensing surveys will of course use information from surveys measuring other statistics such as galaxy distributions or primary CMB anisotropies, so not using any priors is not representative of what stage 3 and 4 experiments could achieve.
 Even a weak prior improves the constraining power of our power spectra considerably.
@@
 Values can be found in table \ref{tab:noiselevels}, including the experiments they are representative of.
-A comparison of noise power spectrum to lensing power spectrum can be found in figure \ref{fig:lpsplusnoise}.
+A comparison of the noise power spectra to the lensing power spectra are shown     in figure \ref{fig:lpsplusnoise}.
+Effectively, for galaxy lensing, the power spectrum is signal-dominated up until $\ell \sim 200$ (stage 2) and $\ell \sim 700$ (stage 4).
+For CMB lensing, the power spectrum is signal-dominated for $\ell \sim 300$ (stage 3) and $\ell \sim 1000$ (stage 4).
 
@@
 
-CMB lensing noise is estimated through the quadratic estimator developed in \cite{cmblensingestimator}.
+CMB lensing noise is estimated using quadratic estimator \cite{cmblensingestimator}.
 We consider experiments that only measure polarization noise.
 This is similarly done in \cite{Namikawa_2016}.
-The parameters characterizing the noise levels are thus beam width, $\sigma$ and polarization white noise, $\Delta_P$.
+The parameters characterizing the noise levels are thus beam width, $\sigma$, and polarization white noise, $\Delta_P$.
 Galaxy lensing is determined by measuring lensing shear (see appendices \ref{sec:weaklensing} and \ref{sec:shear}).
-The noise in this measurement is dominated by scale independent shot noise and has associated noise power spectrum $N_l^{\text{shear}} = \sigma_{\text{rms}}^2 / n_g$, \cite{bartelmann2001weak}.
+The noise in this measurement is dominated by scale-independent shot noise and has associated noise power spectrum $N_l^{\text{shear}} = \sigma_{\text{rms}}^2 / n_g$, \cite{bartelmann2001weak}.
 $n_g$ is the amount of galaxies observed per unit angle.
@@
 \begin{equation*}
-    N_l^{\text{lens. potential}} = \frac{(l-1)l(l+1)(l+2)}{4}N_l^{\text{shear}}.
-\end{equation*}
-For parameter constraints we use stage 4 noise levels.
+    N_\ell^{\text{lens. potential}} = \frac{(\ell-1)\ell(\ell+1)(\ell+2)}{4}N_\ell^{\text{shear}}.
+\end{equation*}
+For parameter constraints, we will consider only stage 4 noise levels.
 
@@
 CMB noise values are chosen in accordance with \cite{Namikawa_2016}.
+For comparison we also how the noise reconstruction for the Simons Observatory.
 Shear noise values are chosen to be similar to e.g.
@@
 and the LSST mission (which has $a \approx 2$, $b \approx 1$, and $z_0 \approx 0.3$ from predictions for the obtained data \cite{lsstsciencebookchapter3}).
-
-Derivatives are calculated with a central difference formula, more information can be found in appendix \ref{sec:derivatives}.
-Obtaining accurate derivatives was challenging.
-In order to check the accuracy of our results we calculated constraints for similar experimental parameters as in references \cite{Planck2018Lensing, Takada2003, Namikawa_2016} and found good agreement in all cases.
+For simplicity, we do not implement a tomographic binning of the source galaxies, and instead model the population with a single effective redshift distribution.
+
+Derivatives are calculated with a central difference formula (see appendix \ref{sec:derivatives} for convergence).
+To check the accuracy of our results we calculated constraints for similar experimental parameters as in references \cite{Planck2018Lensing, Takada2003, Namikawa_2016} and found good agreement in all cases.
 
 \section{Results}\label{sec:results}
-Lensing power spectra are well detectable in nearly all cases. 
-The signal to noise ratios of the CMB and galaxy lensing bispectra versus the maximum multipole moment measured can be found in figure \ref{fig:snrplots}.
-Shear bispectra can be measured well by both stage 3 and stage 4 experiments, while CMB lensing is only marginally detectable by a stage 3 experiment but well detectable by stage 4 experiments.
+Lensing power spectra have high signal-to-noise in all cases. 
+The signal-to-noise of the CMB and galaxy lensing bispectra versus the maximum multipole moment measured is shown in figure \ref{fig:snrplots}.
+In the absence of systematics, shear bispectra can be detected at high signal-to-noise with both stage 3 and stage 4 experiments.
+CMB lensing bispectra could be  detectable by a stage 3 experiment (see e.g.
+ref.~\cite{} for the effect of noise biases) and with high signal-to-noise with a stage 4 experiment.
 CMB $S/N$ results match those of reference \cite{Namikawa_2016}.
@@
     \includegraphics[width=\textwidth]{figures/snrplots.pdf}
-    \caption{Signal to noise ratios for galaxy lensing (dashed) and CMB lensing (non-dashes) power- (in red) and bispectra (in blue).
-The $x$-axis labels the maximum multipole moment measured up to, starting from $l_{\min}=2$.}
+    \caption{Signal-to-noise ratios for galaxy lensing (dashed) and CMB lensing (non-dashes) power- (in red) and bispectra (in blue).
+The $x$-axis labels the maximum multipole moment measured up to, starting from $\ell_{\min}=2$.
+SAY SOMETHING ABOUT LMIN FOR RECONSTRUCTION The maximum multipole used for CMB lensing reconstruction is $\ell = 10^4$.
+Using lower maximum multipole such as 3000 doesn't significantly affect results.}
     \label{fig:snrplots}
@@
 
-Below we show parameter constraints for a selection of variables that have the potential to be particularly well constrained by weak lensing, i.e.
-$m_\nu$, $w_0$, $\sigma_8$, and $\Omega_m$.
-Constraints for $\Lambda$CDM parameters can be found in appendix \ref{sec:lambdacdmconstraints}.
+Next, we consider cosmological parameter constraints.
+We include both the power spectrum and the bispectrum of the weak lensing.
+We emphezize on several two extension of $\Lambda CDM$, $m_\nu$ and $w_0$ as well as $\sigma_8$, and $\Omega_m$.
+Constraints on other parameters can be found found in appendix \ref{sec:lambdacdmconstraints}.
+Results shown include marganlization over all parameters.
 Figures \ref{fig:paramconstraintstightcmb} and \ref{fig:paramconstraintstightgal} show confidence ellipses from stage 4 CMB and galaxy lensing surveys, respectively.
-Table \ref{tab:paramconstraintstight} contains the forecasted constraints as uncertainties in percentage of the fiducial values.
-Despite picking priors that significantly increase powerspectrum constraining power, constraints are significantly improved by including bispectrum information in nearly all cases.
-Again in agreement with reference \cite{Namikawa_2016}, where a 33 percent and 35 percent improvement was found for $m_\nu$ and $w_0$, respectively, by measuring the CMB lensing bispectrum in addition to the powerspectrum.
+Table \ref{tab:paramconstraintstight} contains the forecasted constraints as uncertainties in the percentage of the fiducial values.
+
+!!
+By itsef lensing power spectra are not very constraining, hence we have to assume weak priors. inflated constraints from lps, however still good bispectrum info. 
+!!
+
+Despite using priors that significantly increase power spectrum constraining power, constraints are significantly improved by including bispectrum information in nearly all cases.
+Again in agreement with reference \cite{Namikawa_2016}, in which case a 33 percent and 35 percent improvement was found for $m_\nu$ and $w_0$, respectively, by measuring the CMB lensing bispectrum in addition to the power spectrum.
 Constraints also improve significantly, though not quite as much, when combining CMB and galaxy lensing, compared to only measuring one of these signals.
-Other key observations are:
+
+Besides these obvious findings, we make several other conclusions:
 \begin{itemize}
-    \item Lensing powerspectra alone are not sufficient to competitively constrain $\Lambda\mathrm{CDM} + w_0 + m_\nu$ without priors even in the case of stage 4 surveys.
-    \item For low-enough noise values (in particular for stage 4 noise) the bispectra by themselves can competitively constrain the parameters and contain more information than the power spectra.
-If we denote the multipole moment at which the noise surpasses the powerspectrum signal as $\ell_{N>C}$, than the fisher matrices for the bispectra roughly scale as $F^{B}_{\alpha\beta}\propto \ell_{N>C}^3$ while for the powerspectra they scale as $F^{C}_{\alpha\beta}\propto \ell_{N>C}^2$ (with an extra multipole factor due to the $2\ell + 1$ factor in the powerspectrum estimator covariance).
-The powerspectrum is sufficiently bigger than the bispectrum that it contains more information given current day noise levels, however as surveys get better the bispectra may eventually contain more information because of the more favorable $\ell_{N>C}$ scaling.
-
-    \item The information from bispectra is sensitive to the cosmology in a different way than that of the powerspectra.
+    \item Lensing powers pectra alone are not sufficient to competitively constrain $\Lambda\mathrm{CDM} + w_0 + m_\nu$ without priors, even in the case of stage 4 surveys.
+    \item For low-enough noise values (in particular for stage 4 noise) the bispectra can contain more information than the power spectra .
+If we denote the multipole moment at which the noise surpasses the power spectrum signal as $\ell_{N>C}$, than the fisher matrices for the bispectra roughly scale as $F^{B}_{\alpha\beta}\propto \ell_{N>C}^3$ while for the powerspectra they scale as $F^{C}_{\alpha\beta}\propto \ell_{N>C}^2$ (with an extra multipole factor due to the $2\ell + 1$ factor in the power spectrum estimator covariance).
+The power spectrum is sufficiently bigger than the bispectrum that it contains more information given current-day noise levels, however as surveys get better the bispectra may eventually contain more information because of the more favorable $\ell_{N>C}$ scaling.
+REMOVE L SCALING ARGUMENT YAP ABOUT NON GAUSSIANITY INSTEAD AND EXPLAIN ABOUT DIFFERENCE BETWEEN KERNELS
+
+    \item The information from bispectra is sensitive to cosmology in a different way than that of the power spectra.
 The main parameter combinations where approximate degeneracies are broken are $(n_s,\Omega_b h^2),\;(\Omega_b h^2,H_0),\;(\Omega_c h^2,H_0),\;(\Omega_c h^2,\Omega_b h^2)\;(n_s,\Omega_c h^2),\;(m_\nu,H_0),\;(m_\nu,n_s),$ $(A_s,\Omega_b h^2)$.
@@
 
-    \item Stage 4 CMB and galaxy weak lensing surveys have similar constraining power if both measure powerspectra and bispectra.
-Galaxy surveys are usually somewhat better, but not significantly so.
-Some degeneracies are broken but not quite as much as when looking at combining power and bispectra.
-The main parameter combinations where this happens are  $(n_s, H_0),\;(n_s,\Omega_b h^2),\;(m_\nu,H_0),\;(m_\nu,\Omega_s),\;(A_s,H_0)\;(w_0,A_s),\;(w_0, m_\nu),\;(w_0,\Omega_s)$.
+    \item Stage 4 galaxy weak lensing surveys contain significantly more information than CMB surveys.
+Comparing the confidence ellipses, it is clear that the two types of surveys depend on the underlying parameters in different ways.
+This leads to substantially better constraints for $\sigma_8$, $\Omega_m$, and $m_\nu$, (but not for $w_0$) when combining CMB and galaxy lensing.
 \end{itemize}
@@
     \scriptsize
-    \begin{tabular}{|l|rrr|rrr|rrr|}
-        \hline
-        & \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
-        \hline
-         Par        &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
-        \hline
-         $m_\nu$    &    1262.54 &                    1630.18 &                              779.06 &     435.57 &                    1015.84 &                              268.14 &     304.87 &                     250.06 &                              169.29 \\
-         $w_0$      &      77.72 &                     103.75 &                               57.12 &      19.05 &                      14.94 &                                2.46 &      17.58 &                       9.37 &                                2.26 \\
-         $\sigma_8$ &      15.62 &                      13.64 &                               12.35 &       7.92 &                      14.67 &                                3.26 &       6.73 &                       3.47 &                                1.47 \\
-         $\Omega_m$ &      64.20 &                      57.93 &                               52.32 &      21.14 &                      18.26 &                                4.44 &      20.46 &                      10.21 &                                2.74 \\
+    \begin{tabular}{|l|r|rrr|rrr|rrr|}
+        \hline
+        \multicolumn{11}{|c|}{weak priors} \\
+        \hline
+        && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
+        \hline
+        Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
+        \hline
+        $m_\nu$    &     1.00e+01 &  2.82e-01 &  1.13e+00 &        2.26e-01 & 2.60e-01 & 4.07e-01 &      1.85e-01 &      9.54e-02 &      1.81e-01 &             6.46e-02 \\
+        $w_0$      &     1.00e+01 &  5.42e-01 &  1.44e+00 &        3.87e-01 & 2.14e-01 & 1.28e-01 &      3.22e-02 &      1.51e-01 &      1.13e-01 &             2.45e-02 \\
+        $\sigma_8$ &     2.46e+00 &  1.18e-01 &  1.95e-01 &        5.69e-02 & 5.36e-02 & 5.43e-02 &      1.73e-02 &      3.48e-02 &      3.32e-02 &             9.20e-03 \\
+        $\Omega_m$ &     6.70e-01 &  1.95e-01 &  1.84e-01 &        7.73e-02 & 7.57e-02 & 5.88e-02 &      1.81e-02 &      5.15e-02 &      3.76e-02 &             8.57e-03 \\
+       \hline
+        \multicolumn{11}{|c|}{CMB $T + E$ priors} \\
+        \hline
+        && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
+        \hline
+         Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
+        \hline
+        $m_\nu$    &    1.54e-01 &  5.23e-02 &  8.76e-02 &        5.18e-02 & 1.17e-01 & 1.13e-01 &      8.79e-02 &      1.87e-02 &      7.47e-02 &             1.62e-02 \\
+        $w_0$      &    4.64e-02 &  4.18e-02 &  4.54e-02 &        4.18e-02 & 4.06e-02 & 4.14e-02 &      2.25e-02 &      3.34e-02 &      3.90e-02 &             1.99e-02 \\
+        $\sigma_8$ &    1.38e-02 &  1.06e-02 &  1.16e-02 &        1.06e-02 & 9.86e-03 & 1.00e-02 &      6.99e-03 &      8.98e-03 &      9.72e-03 &             5.28e-03 \\
+        $\Omega_m$ &    1.39e-02 &  1.20e-02 &  1.26e-02 &        1.19e-02 & 1.26e-02 & 1.15e-02 &      9.03e-03 &      1.13e-02 &      1.07e-02 &             6.16e-03 \\
        \hline
@@
 \section{Discussion and conclusion}\label{sec:discussion}
-In this work, we have presented parameter forecasts from CMB lensing and galaxy lensing power- and bispectra, considering experimental setups similar to Stage 3 and Stage 4 surveys.
-Our analysis shows that while lensing power spectra are generally well detectable, the inclusion of bispectra offers significant improvements.
-Notably, shear bispectra are found to be measurable with both Stage 3 and Stage 4 experimental noise levels, whereas CMB lensing bispectra are semi-detectable with Stage 3 and well-detectable with Stage 4 noise levels.
-
-The primary impact of incorporating bispectrum information is a significant tightening of cosmological parameter constraints, particularly for parameters like $H_0$, $\sigma_8$, and $\Omega_m$, and in breaking key parameter degeneracies.
-We found that for sufficiently low noise levels, characteristic of Stage 4 surveys, the bispectra themselves can offer constraining power comparable to or even exceeding that of the power spectra, due to a more favorable scaling of information content with the maximum usable multipole\footnote{The multipole at which the noise powerspectrum starts exceeding the cosmological powerspectrum.}.
+In this work, we have presented parameter forecasts from CMB lensing and galaxy lensing power- and bispectra, considering experimental parameters representative of Stage 3 and Stage 4 CMB and galaxy surveys.
+Our analysis shows that while lensing power spectra are generally well detectable, the inclusion of bispectra offers significant improvements when considering paramater constraints.
+Notably, both CMB and galaxy bispectra are found to be measurable, even with Stage 3 experimental noise levels.
+
+The primary impact of incorporating bispectrum information is a tightening of cosmological parameter constraints, particularly for parameters like $H_0$, $\sigma_8$, and $\Omega_m$, and in breaking key parameter degeneracies.
+We found that for sufficiently low noise levels, characteristic of Stage 4 surveys, the bispectra themselves can offer constraining power comparable to or even exceeding that of the power spectra, due to a more favorable scaling of information content with the maximum usable multipole\footnote{The multipole at which the noise power spectrum starts exceeding the cosmological power spectrum.}.
 Furthermore, the combination of CMB and galaxy lensing probes, especially when both power spectra and bispectra are utilized, further enhances constraints and aids in mitigating degeneracies.
@@
 
-There are a number of limitations in this work.
-The limitations inherent to the fisher matrix formalism and to non physical noise models such as those considered in this paper are already well known.
+There are a number of limitations to the results presented in this paper.
+The limitations inherent to the Fisher matrix formalism and to non-physical noise models such as those considered in this paper are already well known.
 The main limitations specific to this work are:
@@
      \item Certain cosmological parameters that affect nonlinear large-scale structure, such as neutrino masses, likely alter the fitting parameters of the matter bispectrum formula when varied.
-In the literature this is currently not considered, which can make the derivatives of the bispectrum with respect to these cosmological parameters less accurate.
+In the literature, this is currently not considered, which can make the derivatives of the bispectrum with respect to these cosmological parameters less accurate.
      \item We assumed throughout this paper that surveys would measure multipoles as low as $\ell = 2$.
 However, our analytical formulas to calculate lensing spectra utilize the Limber approximation, which is only robust for $\ell \gtrsim 50$.
-Avoiding the Limber approximation would significantly complicate calculations, requiring knowledge of unequal-time matter power and bispectra, and would be computationally much more demanding due to the need to evaluate double integrals instead of single integrals.
+Avoiding the Limber approximation would complicate calculations, requiring knowledge of unequal-time matter power and bispectra, and would be computationally much more demanding due to the need to evaluate double integrals instead of single integrals.
 With the methods used in this paper, recalculating Fisher matrices without the Limber approximation would not be feasible, even if unequal-time matter spectra were readily available.
@@
 Notably, such simulations could potentially also allow for a more accurate calculation of the lensing spectra themselves, without relying on fitting formulas for the matter bispectrum or the Limber approximation, depending on simulation accuracy.
+This might also explain the fact that in some instances the bispectra appear more constraining compared to the power spectra.
 
@@
 Better estimators exist; for example, the iterative estimator developed in \cite{Smith_2012} can lead to a significantly better signal-to-noise ratio for the CMB lensing bispectrum, as shown in \cite{Namikawa_2016}.
-    \item To calculate the Fisher matrices within a manageable timeframe, derivatives of the lensing spectra with respect to cosmological parameters were computed using a central difference formula.
+    \item To calculate the Fisher matrices, derivatives of the lensing spectra with respect to cosmological parameters were computed using a central difference formula.
 This introduces errors (estimated at most a few percent, though larger for neutrino mass derivatives, see appendix \ref{sec:derivatives}).
@@
 \subsection{Perturbed Photon Paths}
-We work in the conformal newtonian gauge and with natural units.
+We work in the conformal Newtonian gauge and with natural units.
 Denoting conformal time and conformal radial distance by $\eta$ and $\chi$, respectively, the perturbed line element in FLRW spacetime is given by
@@
 and $f_K(\chi)$ is the comoving angular diameter distance.
-We will hereafter only consider a flat universe, so that $f_K(\chi) = \chi$.
+We will hereafter only consider a flat universe so that $f_K(\chi) = \chi$.
 Weak lensing of a point source can be quantified by looking at the deflection field $\mathbf d (\hat{\mathbf n}) = \mathbf \theta_{\text{obs}} - \mathbf \theta_{\text{true}}$, i.e.
-the (small) difference between the angle at which we see the object and the angle at which we would see the object had no lensing occued.
+the (small) difference between the angle at which we see the object and the angle at which we would see the object had no lensing occurred.
 To first order in $\Psi_N$ and $\Phi_N$, the deflection is given as \cite{dodelson2020modern}
@@
 \end{equation}
-The integration limit is sometimes also taken to be the surface of last scattering, as any window function vanishes after that distance.
+The integration limit is sometimes also taken to be the surface of the last scattering, as any window function vanishes after that distance.
 In the case of CMB lensing we can take $p(\chi')=\delta(\chi'-\chi_*)$, in which case the window function reduces to $H(\chi_* - \chi)(\chi_*-\chi)/(\chi_*\chi)$, with $H(\chi)$ the Heaviside step function.
@@
 \end{equation}
-At 1st order $A$ is a symetric matrix by definition and $\omega$ vanishes, we will ignore it from here on out.
+At 1st order, $A$ is a symmetric matrix by definition and $\omega$ vanishes, we will ignore it from here on out.
 Intuitively, $A$ tells you how a small patch in the sky transforms due to lensing.
@@
 \end{equation}
-For an image on the sky, $A_{ij}$ thus introduces distortion.
+For an image of the sky, $A_{ij}$ thus introduces distortion.
 Note that $|A_{ij}| = (1 - \kappa)^2 + \omega^2 - |\gamma|^2 = 1 - 2\kappa + O(\kappa^2, \gamma^2, \omega^2)$.
-We can thus interpret $\kappa$ as telling us about the overal magnification of the source.
-The $\gamma_i$ represent the area preserving distortion, i.e.
+We can thus interpret $\kappa$ as telling us about the overall magnification of the source.
+The $\gamma_i$ represents the area-preserving distortion, i.e.
 stretching and squeezing in a specific direction.
@@
 \end{equation}
-In this context a spin $s$ function referes to a function $_sf(\theta, \phi)$ that transforms under any rotation of coordinates by picking up a phase factor $e^{is\alpha}$, with $\alpha$ the angle of the rotation, i.e.
+In this context, a spin $s$ function refers to a function $_sf(\theta, \phi)$ that transforms under any rotation of coordinates by picking up a phase factor $e^{is\alpha}$, with $\alpha$ the angle of the rotation, i.e.
 \begin{equation}
@@
 \section{Weak Lensing Statistics}\label{sec:weaklensstats}
-\subsection{Lensing Potential Powerspectrum}
+\subsection{Lensing Potential Power spectrum}
 The lensing potential can be decomposed into spherical harmonics as
 \begin{equation}
-    \psi(\hat{\mathbf n}) = \sum_{lm} \psi_{lm}Y_{lm}(\hat{\mathbf n}).
+    \psi(\hat{\mathbf n}) = \sum_{\ell m} \psi_{\ell m}Y_{\ell m}(\hat{\mathbf n}).
 \end{equation}
@@
 \end{equation}
-We can than relate the multipole modes of $\psi$ to the fourier modes of $\Psi$ through % TODO: check sign convention here
+We can then relate the multipole modes of $\psi$ to the Fourier modes of $\Psi$ through % TODO: check sign convention here
 \begin{gather}
-    \psi_{lm} = \langle Y_l^m | \psi \rangle = \int\d^2\hat{\mathbf n} Y_l^m(\hat{\mathbf n})^* \psi(\hat{\mathbf n})\\ 
+    \psi_{lm} = \langle Y_\ell^m | \psi \rangle = \int\d^2\hat{\mathbf n} Y_\ell^m(\hat{\mathbf n})^* \psi(\hat{\mathbf n})\\ 
     = -2\int\d^2\hat{\mathbf n} Y_l^m(\hat{\mathbf n})^* \int_0^{\chi_*}\d \chi W(\chi) \int\frac{\d^3 \mathbf k}{(2\pi)^{3}}\Psi(\mathbf k, \eta_0 - \chi)e^{i\mathbf k \cdot\hat{\mathbf n}\chi}
@@
 \end{equation}
-where we used that $\eta = \eta_0 - \chi$ along the unperturbed photon path (this is known as the Born approximation), with $\eta_0$ the time at which the light ray hits earth. 
+where we used that $\eta = \eta_0 - \chi$ along the unperturbed photon path (this is known as the Born approximation), with $\eta_0$ the time at which the light ray hits Earth. 
 We can use the result
 \begin{equation}
-    e^{i\mathbf k \cdot \hat {\mathbf n}\chi}=4\pi\sum_{lm}i^lj_l(k\chi )Y_l^m(\hat{\mathbf n})^*Y_l^m(\hat{\mathbf k}) = 4\pi\sum_{lm}i^lj_l(k\chi )Y_l^m(\hat{\mathbf n})Y_l^m(\hat{\mathbf k})^*, \label{eq:complexexp}
+    e^{i\mathbf k \cdot \hat {\mathbf n}\chi}=4\pi\sum_{\ell m}i^\ell j_\ell(k\chi )Y_\ell^m(\hat{\mathbf n})^*Y_\ell^m(\hat{\mathbf k}) = 4\pi\sum_{\ell m}i^\ell j_\ell(k\chi )Y_\ell^m(\hat{\mathbf n})Y_\ell^m(\hat{\mathbf k})^*, \label{eq:complexexp}
 \end{equation}
 where $j_l$ is the spherical Bessel function, to rewrite the above equation.
-Using both versions of the identity above, we immediately get a factor $Y_l^m(\hat{\mathbf k})Y_{l'}^{m'}(\hat{\mathbf k})^*$ in our integral.
+Using both versions of the identity above, we immediately get a factor $Y_\ell^m(\hat{\mathbf k})Y_{\ell'}^{m'}(\hat{\mathbf k})^*$ in our integral.
 We can factor the differential element of $\d^3\mathbf k$ into a radial and angular part as $k^2\d k\d^2\Omega_k$, with $\Omega_k$ the solid angle, to apply the orthonormality condition of the spherical harmonics.
@@
 \begin{gather}
-    \langle \psi(\hat{\mathbf n})\psi(\hat{\mathbf n}') \rangle = 4(4\pi)^2\sum_{ll'mm'}i^{l-l'}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\\\times\int \frac{k^2\d k}{(2\pi)^6}\frac{2\pi^2}{k^3}j_l(k\chi)j_{l'}(k\chi')P_\psi(k,\eta_0 - \chi, \eta_0-\chi')Y_{lm}(\hat{\mathbf n})Y_{l'm'}(\hat{\mathbf n}')^*\delta_{ll'}\delta_{mm'}.
+    \langle \psi(\hat{\mathbf n})\psi(\hat{\mathbf n}') \rangle = 4(4\pi)^2\sum_{ll'mm'}i^{l-l'}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\\\times\int \frac{k^2\d k}{(2\pi)^6}\frac{2\pi^2}{k^3}j_\ell(k\chi)j_{\ell'}(k\chi')P_\psi(k,\eta_0 - \chi, \eta_0-\chi')Y_{\ell m}(\hat{\mathbf n})Y_{\ell'm'}(\hat{\mathbf n}')^*\delta_{\ell \ell'}\delta_{mm'}.
 \label{eq:lenspotcorrexpanded}
@@
 \begin{equation}
-    \langle \psi_{lm}\psi_{l'm'}^* \rangle = \delta_{ll'}\delta_{mm'}C_l^\psi.
+    \langle \psi_{\ell m}\psi_{\ell'm'}^* \rangle = \delta_{\ell \ell'}\delta_{mm'}C_\ell^\psi.
 \end{equation}
@@
 \begin{gather}
-    C_l^\psi = 4(4\pi)^2\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\int \frac{k^2\d k}{(2\pi)^6}\frac{2\pi^2}{k^3}j_l(k\chi)j_{l}(k\chi')P_\psi(k,\eta_0 - \chi, \eta_0-\chi'),
+    C_l^\psi = 4(4\pi)^2\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\int \frac{k^2\d k}{(2\pi)^6}\frac{2\pi^2}{k^3}j_\ell(k\chi)j_{\ell}(k\chi')P_\psi(k,\eta_0 - \chi, \eta_0-\chi'),
 \end{gather}
@@
 \begin{gather}
-    C_l^\psi = \frac{2}{\pi^2}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\int k^2\d kj_l(k\chi)j_{l}(k\chi')\frac{P_\psi(k,\eta_0 - \chi, \eta_0-\chi')}{k^3}.
+    C_\ell^\psi = \frac{2}{\pi^2}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi')\int k^2\d kj_\ell(k\chi)j_{\ell}(k\chi')\frac{P_\psi(k,\eta_0 - \chi, \eta_0-\chi')}{k^3}.
 \end{gather}
@@
 \begin{equation}
-    \int k^2\d k j_l(k\chi)j_l(k\chi') = \frac{\pi}{2\chi^2}\delta(\chi-\chi').
+    \int k^2\d k j_\ell(k\chi)j_\ell(k\chi') = \frac{\pi}{2\chi^2}\delta(\chi-\chi').
 \end{equation}
@@
 \begin{equation}
-    \int k^2\d k j_l(k\chi)j_l(k\chi')f(k) \approx \frac{\pi}{2\chi^2}\delta(\chi-\chi')f(l/\chi).
+    \int k^2\d k j_\ell(k\chi)j_\ell(k\chi')f(k) \approx \frac{\pi}{2\chi^2}\delta(\chi-\chi')f(\ell/\chi).
 \end{equation}
@@
 \begin{gather}
-    C^\psi(l) = \frac{2}{\pi^2}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi') \frac{\pi}{2\chi^2}\delta(\chi-\chi')\frac{\chi^3}{l^3}P_\psi(l/\chi,\eta_0 - \chi, \eta_0-\chi')\\
-    = \frac{1}{l^3\pi}\int_0^{\chi_*}\chi\d \chi W(\chi)^2 P_\psi(l/\chi,\eta_0 - \chi, \eta_0-\chi). \label{eq:lpsintermsofpsi}
+    C^\psi_{\ell} = \frac{2}{\pi^2}\int_0^{\chi_*}\d \chi\int_0^{\chi_*}\d \chi' W(\chi)W(\chi') \frac{\pi}{2\chi^2}\delta(\chi-\chi')\frac{\chi^3}{\ell^3}P_\psi(\ell/\chi,\eta_0 - \chi, \eta_0-\chi')\\
+    = \frac{1}{\ell^3\pi}\int_0^{\chi_*}\chi\d \chi W(\chi)^2 P_\psi(\ell/\chi,\eta_0 - \chi, \eta_0-\chi). \label{eq:lpsintermsofpsi}
 \end{gather}
@@
 \begin{gather*}
-    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle = 
-    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{l_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}e^{i\mathbf k_i \cdot\hat{\mathbf n_i}\chi_i}} \\
+    \langle (\psi_1)_{\ell_1m_1}(\psi_2)_{\ell_2m_2}(\psi_3)_{\ell_3m_3} \rangle = 
+    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{\ell_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}e^{i\mathbf k_i \cdot\hat{\mathbf n_i}\chi_i}} \\
     \times \langle \prod_i\Psi(\mathbf k_i, \eta_0 - \chi_i) \rangle.
 \end{gather*}
-Defining the bispectrum of the graviational potential as
+Defining the bispectrum of the gravitational potential as
 \begin{equation*}
@@
 \end{equation*}
-we rewrite the lensing potential bispecrum as
+we rewrite the lensing potential bispectrum as
 \begin{gather*}
-    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle = 
-    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{l_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}e^{i\mathbf k_i \cdot\hat{\mathbf n_i}\chi_i}}\\ \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
+    \langle (\psi_1)_{\ell_1m_1}(\psi_2)_{\ell_2m_2}(\psi_3)_{\ell_3m_3} \rangle = 
+    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{\ell_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}e^{i\mathbf k_i \cdot\hat{\mathbf n_i}\chi_i}}\\ \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
 \end{gather*}
@@
 \begin{gather*}
-    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle \\ = 
-    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{l_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi\sum_{lm}i^{l}j_{l}(k_i\chi_i)Y_{l}^{m}(\hat{\mathbf n_i})Y_{l}^{m}(\hat{\mathbf k_i})^*}\\
+    \langle (\psi_1)_{\ell_1m_1}(\psi_2)_{\ell_2m_2}(\psi_3)_{\ell_3m_3} \rangle \\ = 
+    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{\ell_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi\sum_{\ellm}i^{\ell}j_{l}(k_i\chi_i)Y_{\ell}^{m}(\hat{\mathbf n_i})Y_{\ell}^{m}(\hat{\mathbf k_i})^*}\\
     \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}) \\
-    = \sbr{\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)Y_{l_i}^{m_i}(\hat{\mathbf k_i})^*}} (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
+    = \sbr{\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi i^{\ell_i}j_{\ell_i}(k_i\chi_i)Y_{\ell_i}^{m_i}(\hat{\mathbf k_i})^*}} (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
 \end{gather*}
-We can write the 3D dirac delta function in terms of spherical harmonics as
-\begin{equation*}
-    \delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3) = 8\int\d^3\mathbf x\prod_{i=1,2,3}\br{ \sum_{l_jm_j} i^{l_j}j_{l_j}(k_ix)Y_{l_j}^{m_j}(\hat{\mathbf k_i}) Y_{l_j}^{m_j}(\hat{\mathbf x})^* }.
+We can write the 3D Dirac delta function in terms of spherical harmonics as
+\begin{equation*}
+    \delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3) = 8\int\d^3\mathbf x\prod_{i=1,2,3}\br{ \sum_{\ell_jm_j} i^{\ell_j}j_{\ell_j}(k_ix)Y_{\ell_j}^{m_j}(\hat{\mathbf k_i}) Y_{\ell_j}^{m_j}(\hat{\mathbf x})^* }.
 \end{equation*}
@@
 \begin{gather*}
-    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle= \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi i^{l_i}j_{l_i}(k_i\chi_i)Y_{l_i}^{m_i}(\hat{\mathbf k_i})^*} \\
-    \times (2\pi)^38\int\d^3\mathbf x\prod_{i}\br{ \sum_{lm} i^{l}j_{l}(k_ix)Y_{l}^{m}(\hat{\mathbf k_i}) Y_{l}^{m}(\hat{\mathbf x})^* }B^\Psi(\{k_i\}, \{\eta_0-\chi_i\})\\
-    = (2\pi)^38\int\d^3\mathbf x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3}}4\pi (-1)^{l_i}j_{l_i}(k_i\chi_i)j_{l_i}(k_ix) Y_{l_i}^{m_i}(\hat{\mathbf x})^*} B^\Psi(\{k_i\}, \{\eta_0-\chi_i\})
+    \langle (\psi_1)_{\ell_1m_1}(\psi_2)_{\ell_2m_2}(\psi_3)_{\ell_3m_3} \rangle= \prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi i^{\ell_i}j_{\ell_i}(k_i\chi_i)Y_{\ell_i}^{m_i}(\hat{\mathbf k_i})^*} \\
+    \times (2\pi)^38\int\d^3\mathbf x\prod_{i}\br{ \sum_{\ell m} i^{\ell}j_{\ell}(k_ix)Y_{\ell}^{m}(\hat{\mathbf k_i}) Y_{\ell}^{m}(\hat{\mathbf x})^* }B^\Psi(\{k_i\}, \{\eta_0-\chi_i\})\\
+    = (2\pi)^38\int\d^3\mathbf x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3}}4\pi (-1)^{\ell_i}j_{\ell_i}(k_i\chi_i)j_{\ell_i}(k_ix) Y_{\ell_i}^{m_i}(\hat{\mathbf x})^*} B^\Psi(\{k_i\}, \{\eta_0-\chi_i\})
 \end{gather*}
@@
 \begin{gather*}
-    \int \mathrm{d}\Omega_{\hat{n}} \, Y_{l_1 m_1}^*(\hat{x}) Y_{l_2 m_2}^*(\hat{n}) Y_{l_3 m_3}^*(\hat{n}) = (-1)^{m_1 + m_2 + m_3} 
-    \int \mathrm{d}\Omega_{\hat{n}} \, Y_{l_1 -m_1}(\hat{n}) Y_{l_2 -m_2}(\hat{n})Y_{l_3 -m_3}(\hat{n}) \\
+    \int \mathrm{d}\Omega_{\hat{n}} \, Y_{\ell_1 m_1}^*(\hat{x}) Y_{\ell_2 m_2}^*(\hat{n}) Y_{\ell_3 m_3}^*(\hat{n}) = (-1)^{m_1 + m_2 + m_3} 
+    \int \mathrm{d}\Omega_{\hat{n}} \, Y_{\ell_1 -m_1}(\hat{n}) Y_{\ell_2 -m_2}(\hat{n})Y_{\ell_3 -m_3}(\hat{n}) \\
     = (-1)^{m_1 + m_2 + m_3} 
-    \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix}
-    \begin{pmatrix} l_1 & l_2 & l_3 \\ -m_1 & -m_2 & -m_3 \end{pmatrix} \equiv A_{\mathbf{l}}^{\mathbf{m}},
+    \sqrt{\frac{(2\ell_1 + 1)(2\ell_2 + 1)(2\ell_3 + 1)}{4\pi}} \begin{pmatrix} \ell_1 & \ell_2 & \ell_3 \\ 0 & 0 & 0 \end{pmatrix}
+    \begin{pmatrix} \ell_1 & \ell_2 & \ell_3 \\ -m_1 & -m_2 & -m_3 \end{pmatrix} \equiv A_{\mathbf{l}}^{\mathbf{m}},
 \end{gather*}
@@
 \begin{gather*}
-    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle 
-    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3}}4\pi (-1)^{l_i}j_{l_i}(k_i\chi_i)j_{l_i}(k_ix)} \\ \times B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
+    \langle (\psi_1)_{\ell_1m_1}(\psi_2)_{\ell_2m_2}(\psi_3)_{\ell_3m_3} \rangle 
+    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3}}4\pi (-1)^{\elli}j_{\ell_i}(k_i\chi_i)j_{\ell_i}(k_ix)} \\ \times B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
 \end{gather*}
@@
 \begin{gather*}
-    \langle (\psi_1)_{l_1m_1}(\psi_2)_{l_2m_2}(\psi_3)_{l_3m_3} \rangle 
-    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi_i^2}\delta(x-\chi_i)4\pi (-1)^{l_i}} \\ \times B^\Psi(\{l_i/\chi_i\}, \{\eta_0-\chi_i\}) \\
-    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int \chi^2\d \chi\prod_i\br{-2 W_i(\chi)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi^2}4\pi (-1)^{l_i}} B^\Psi(\{l_i/\chi\}, \eta_0-\chi).
+    \langle (\psi_1)_{\ell_1m_1}(\psi_2)_{\ell_2m_2}(\psi_3)_{\ell_3m_3} \rangle 
+    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi_i^2}\delta(x-\chi_i)4\pi (-1)^{\ell_i}} \\ \times B^\Psi(\{\ell_i/\chi_i\}, \{\eta_0-\chi_i\}) \\
+    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int \chi^2\d \chi\prod_i\br{-2 W_i(\chi)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi^2}4\pi (-1)^{\ell_i}} B^\Psi(\{\ell_i/\chi\}, \eta_0-\chi).
 \end{gather*}
@@
 
-The definition for the bispectrum of any set of randomly distributed spherical harmonic components $X^k_{lm}$ is \cite{Hu2000}
-\begin{equation*}
-    \langle (X_1)_{l_1m_1}(X_2)_{l_2m_2}(X_3)_{x_3m_3} \rangle = \begin{pmatrix}
-        l_1 & l_2 & l_3 \\ m_1 & m_2 & m_3
+The definition for the bispectrum of any set of randomly distributed spherical harmonic components $X^k_{\ell m}$ is \cite{Hu2000}
+\begin{equation*}
+    \langle (X_1)_{\ell_1m_1}(X_2)_{\ell_2m_2}(X_3)_{x_3m_3} \rangle = \begin{pmatrix}
+        \ell_1 & \ell_2 & \ell_3 \\ m_1 & m_2 & m_3
     \end{pmatrix}
-    B_{l_1l_2l_3}^{X_1X_2X_3}.
-\end{equation*}
-Note the independence on $m_i$, this necesarilly follows from statistical isotropy. % TODO: add a short explanation here
+    B_{\ell_1\ell_2\ell_3}^{X_1X_2X_3}.
+\end{equation*}
+Note the independence on $m_i$, this necessarily follows from statistical isotropy. % TODO: add a short explanation here
 If $m_1+m_2+m_3\neq 0$, the associated Wigner-3j symbol vanishes and the bispectrum is set to zero.
 Also note that in this definition we have immediately generalized to include cross correlation between different fields $X_1$, $X_2$, $X_3$.
-This is relevant when we look at cross correlations between CMB and galaxy lensing.
+This is relevant when we look at cross-correlations between CMB and galaxy lensing.
 6758
@@
     \begin{pmatrix}
-        l_1 & l_2 & l_3\\
+        \ell_1 & \ell_2 & \ell_3\\
         -m_1 & -m_2 & -m_3
@@
       =
-      (-1)^{l_1+l_2+l_3}
+      (-1)^{\ell_1+\ell_2+\ell_3}
       \begin{pmatrix}
-        l_1 & l_2 & l_3\\
+        \ell_1 & \ell_2 & \ell_3\\
         m_1 & m_2 & m_3
@@
 \begin{gather*}
-    B^{\psi_1\psi_2\psi_3}_{l_1l_2l_3}
-    = (-1)^{l_1+l_2+l_3}
-    \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix}
+    B^{\psi_1\psi_2\psi_3}_{\ell_1\ell_2\ell_3}
+    = (-1)^{\ell_1+\ell_2+\ell_3}
+    \sqrt{\frac{(2\ell_1 + 1)(2\ell_2 + 1)(2\ell_3 + 1)}{4\pi}} \begin{pmatrix} \ell_1 & \ell_2 & \ell_3 \\ 0 & 0 & 0 \end{pmatrix}
     (2\pi)^3 8 \\
-    \times \int \chi^2\d \chi\prod_i\br{-2 W_i(\chi, \chi_*)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi^2}4\pi (-1)^{l_i}} B^\Psi(\{l_i/\chi\}, \{\eta_0-\chi\}),
+    \times \int \chi^2\d \chi\prod_i\br{-2 W_i(\chi, \chi_*)\frac{1}{(2\pi)^{3}}\frac{\pi}{2\chi^2}4\pi (-1)^{\ell_i}} B^\Psi(\{\ell_i/\chi\}, \{\eta_0-\chi\}),
 \end{gather*}
-where we were able to drop the $(-1)^{m_1+m_2+m_3}$ factor due to the bispecrum vanishing if that sum doesn't equal $0$, as mentioned earlier.
+where we were able to drop the $(-1)^{m_1+m_2+m_3}$ factor due to the bispectrum vanishing if that sum does not equal $0$, as mentioned earlier.
 When all $m_i$ equal zero, the Wigner 3j-symbol gains a number of useful properties % TODO: discuss all of them
-In particular, it vanishes if $l_1+l_2+l_3$ is odd, meaning we can drop the $(-1)^{l_1+l_2+l_3}$ factor. Additionally cancelling common factors then gives
+In particular, it vanishes if $\ell_1+\ell_2+\ell_3$ is odd, meaning we can drop the $(-1)^{\ell_1+\ell_2+\ell_3}$ factor. Additionally cancelling common factors then gives
 \begin{align}
-    B^{\psi_1\psi_2\psi_3}_{l_1l_2l_3}
+    B^{\psi_1\psi_2\psi_3}_{\ell_1\ell_2\ell_3}
     =
-    -\sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} 8 \int \frac{\d \chi }{\chi^4} W_1(\chi)W_2(\chi)W_3(\chi) B^\Psi(\{l_i/\chi\}, \eta_0-\chi).
+    -\sqrt{\frac{(2\ell_1 + 1)(2\ell_2 + 1)(2\ell_3 + 1)}{4\pi}} \begin{pmatrix} \ell_1 & \ell_2 & \ell_3 \\ 0 & 0 & 0 \end{pmatrix} 8 \int \frac{\d \chi }{\chi^4} W_1(\chi)W_2(\chi)W_3(\chi) B^\Psi(\{l_i/\chi\}, \eta_0-\chi).
 \label{eq:lbsintermsofpsi}
@@
 %     = A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\frac{1}{(2\pi)^{3/2}}4\pi i^{l_i}(2i^{l_i})\frac{\pi}{2\chi_i^2}\delta(x-\chi_i)}
-%     (2\pi)^3  B_\Psi(l_1/\chi, l_2/\chi, l_3/\chi) \\
+%     (2\pi)^3  B_\Psi(\ell_1/\chi, \ell_2/\chi, \ell_3/\chi) \\
 %     = A_{\mathbf l}^{\mathbf m}\int_0^{\chi_*} x^2\d x\prod_i\br{-2 W(x, \chi_*)\frac{1}{(2\pi)^{3/2}}4\pi i^{l_i}(2i^{l_i})\frac{\pi}{2x^2}}
-%     (2\pi)^3  B_\Psi(l_1/\chi, l_2/\chi, l_3/\chi) \\
-%     = A_{\mathbf{l}}^{\mathbf{m}} (-1)^{1 + l_1 + l_2 + l_3} 8 (2\pi)^{9/2} \int_0^{\chi_*} \frac{[W(\chi, \chi_*)]^3}{\chi^4} \, B_\Psi\left( \frac{l_1}{\chi}, \frac{l_2}{\chi}, \frac{l_3}{\chi} \right ) \, \mathrm{d}\chi.
+%     (2\pi)^3  B_\Psi(\ell_1/\chi, \ell_2/\chi, \ell_3/\chi) \\
+%     = A_{\mathbf{l}}^{\mathbf{m}} (-1)^{1 + \ell_1 + \ell_2 + \ell_3} 8 (2\pi)^{9/2} \int_0^{\chi_*} \frac{[W(\chi, \chi_*)]^3}{\chi^4} \, B_\Psi\left( \frac{\ell_1}{\chi}, \frac{\ell_2}{\chi}, \frac{\ell_3}{\chi} \right ) \, \mathrm{d}\chi.
 % \end{gather*}
@@
 \begin{empheq}[box=\fbox]{align*}
-    C^{\psi_X\psi_Y}_l
-    & = \frac{9}{l^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)C^\delta(l/\chi,\eta_0-\chi),\\
-    B^{\psi_X\psi_Y\psi_Z}_{l_1l_2l_3} &= \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{l_1^2l_2^2l_3^2}\Omega_m^3H_0^6\\
-    & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{l_i/\chi\}, \eta_0-\chi).
+    C^{\psi_X\psi_Y}_\ell
+    & = \frac{9}{\ell^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)P^\delta(\ell/\chi,\eta_0-\chi),\\
+    B^{\psi_X\psi_Y\psi_Z}_{\ell_1\ell_2\ell_3} &= \sqrt{\frac{(2\ell_1 + 1)(2\ell_2 + 1)(2\ell_3 + 1)}{4\pi}} \begin{pmatrix} \ell_1 & \ell_2 & \ell_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{\ell_1^2\ell_2^2\ell_3^2}\Omega_m^3H_0^6\\
+    & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{\ell_i/\chi\}, \eta_0-\chi).
 \end{empheq}
@@
 \begin{equation*}
-    F_{\alpha\beta} = \sum_{l_{min}\leq l, l' \leq l_{max}}\sum_{[XY][X'Y']} \partial_\alpha C_{l}^{XY} \br{\Cov^{-1}}^{XY, X'Y'}_{l, l'} \partial_\beta C^{X'Y'}_{l'}.
+    F_{\alpha\beta} = \sum_{l_{min}\leq \ell, \ell' \leq \ell_{max}}\sum_{[XY][X'Y']} \partial_\alpha C_{\ell}^{XY} \br{\Cov^{-1}}^{XY, X'Y'}_{\ell, \ell'} \partial_\beta C^{X'Y'}_{\ell'}.
 \end{equation*}
@@
 \begin{equation*}
-    \Cov^{XY, X'Y'}_{l, l'} = \langle \hat C^{XY}_{l} \hat C^{X'Y'}_{l'} \rangle.
-\end{equation*}
-The estimator of a power or bispectrum is the product of the estimators of the appropriate tracers, e.g.
-$\hat C^{XY}_{l} = \hat X(l) \hat Y(l)$.
+    \Cov^{XY, X'Y'}_{\ell, \ell'} = \langle \hat C^{XY}_{\ell} \hat C^{X'Y'}_{\ell'} \rangle.
+\end{equation*}
+The estimator of the power- or bi-spectrum is the product of the estimators of the appropriate tracers, e.g.
+$\hat C^{XY}_{\ell} = \hat X(\ell) \hat Y(\ell)$.
 The sum over $[XY]$ and $[X'Y']$ denotes a sum over possible tracer combinations \textit{without} counting permutations of tracer configuration.
 This is because permutations are not distinct signals, i.e.
-$\hat X(l) \hat Y(l) = \hat Y(l) \hat X(l)$.
-In fact, if we were to count these permutations as distinct signals we will get identical columns and/or rows in our covariance matrix making inversion impossible:
-\begin{equation*}
-    \langle \hat X(l) \hat Y (l) \hat X'(l') \hat Y' (l') \rangle = \langle \hat Y (l) \hat X (l) \hat X'(l') \hat Y' (l') \rangle, \quad \forall X', Y', l'. 
-\end{equation*}
-
-To evaluate the covariance matrix we again assume that the estimators are gaussian distributed so that we can apply a wick contraction, as is commonly done \cite{tegmark1997cosmic}.
+$\hat X(\ell) \hat Y(\ell) = \hat Y(\ell) \hat X(\ell)$.
+In fact, if we were to count these permutations as distinct signals we would get identical columns and/or rows in our covariance matrix making inversion impossible:
+\begin{equation*}
+    \langle \hat X(\ell) \hat Y (\ell) \hat X'(\ell') \hat Y' (\ell') \rangle = \langle \hat Y (\ell) \hat X (\ell) \hat X'(\ell') \hat Y' (\ell') \rangle, \quad \forall X', Y', \ell'. 
+\end{equation*}
+
+To evaluate the covariance matrix we again assume that the estimators are Gaussian distributed so that we can apply a wick contraction, as is commonly done \cite{tegmark1997cosmic}.
 In this case we get
 \begin{equation*}
-    \Cov^{XY, X'Y'}_{l, l'} = \frac{1}{2l+1}\delta_{ll'}\br{\tilde C^{XX'}_l \tilde C^{YY'}_l + \tilde C^{XY'}_l \tilde C^{YX'}_l}.
+    \Cov^{XY, X'Y'}_{\ell, \ell'} = \frac{1}{2\ell+1}\delta_{\ell \ell'}\br{\tilde C^{XX'}_\ell \tilde C^{YY'}_\ell + \tilde C^{XY'}_\ell \tilde C^{YX'}_\ell}.
 \end{equation*}
@@
     \begin{equation*}
-        \langle X_{lm}X'_{l'm'} \rangle = (2\pi)^3 \delta_{ll'}\delta_{mm'}C_l^{XX'},
+        \langle X_{\ell m}X'_{\ell'm'} \rangle = (2\pi)^3 \delta_{\ell\ell'}\delta_{mm'}C_\ell^{XX'},
     \end{equation*}
     so you can average over the measurements done for different values of $m$, i.e.
-$X_{l(-l)}$, $X_{l(-l+1)}$, $\cdots$, $X_{l(l-1)}$, $X_{ll}$.
-This results in the $(2l+1)^{-1}$ factor in the power spectrum covariance.
-    \item The tilde is used to denote the powerspectrum as calculated earlier plus the noise power spectrum, $N^{XX'}_l$, which is the power spectrum of the noise associated with the estimator of the field.
+$X_{\ell(-\ell)}$, $X_{\ell(-\ell+1)}$, $\cdots$, $X_{\ell(\ell-1)}$, $X_{\ell \ell}$.
+This results in the $(2\ell+1)^{-1}$ factor in the power spectrum covariance.
+    \item The tilde is used to denote the power spectrum as calculated earlier plus the noise power spectrum, $N^{XX'}_\ell$, which is the power spectrum of the noise associated with the estimator of the field.
 This is where experimental noise is incorporated into the calculation.
 \end{enumerate}
-Under the Gaussian approximation the covariance matrix vanishes except for $3\times 3$ block matrices (in the case of 2 tracers) on the diagonal. The fisher matrix is then
-\begin{equation*}
-    F_{\alpha\beta} = \sum_{l}\sum_{[XY][X'Y']} \partial_\alpha C_{l}^{XY} \br{\Cov^{-1}_l}^{XY, X'Y'} \partial_\beta C^{X'Y'}_{l}.
-\end{equation*} 
-$\Cov^{-1}_l$ here denotes the inverse of the block matrix at $l$.
+Under the Gaussian approximation, the covariance matrix vanishes except for $3\times 3$ block matrices (in the case of 2 tracers) on the diagonal. The fisher matrix is then
+\begin{equation*}
+    F_{\alpha\beta} = \sum_{\ell}\sum_{[XY][X'Y']} \partial_\alpha C_{\ell}^{XY} \br{\Cov^{-1}_\ell}^{XY, X'Y'} \partial_\beta C^{X'Y'}_{\ell}.
+\end{equation*} 
+$\Cov^{-1}_\ell$ here denotes the inverse of the block matrix at $l$.
 
@@
 \begin{equation*}
-    F_{\alpha\beta} = \sum_{\text{distinct signals}}\sum_{\text{distinct signals prime}} B^{XYZ}_{l_1l_2l_3}\br{\Cov^{-1}}^{XYZ,X'Z'Y'}_{l_1l_2l_3,l_1'l_2'l_3'}B^{X'Y'Z'}_{l_1'l_2'l_3'}.
+    F_{\alpha\beta} = \sum_{\text{distinct signals}}\sum_{\text{distinct signals prime}} B^{XYZ}_{\ell_1\ell_2\ell_3}\br{\Cov^{-1}}^{XYZ,X'Z'Y'}_{\ell_1\ell_2\ell_3,\ell_1'\ell_2'\ell_3'}B^{X'Y'Z'}_{\ell_1'\ell_2'\ell_3'}.
 \end{equation*} 
 Counting only distinct signals requires more care compared to the power spectra.
-The rule is that $B^{XYZ}_{l_1l_2l_3}$ and $B^{X'Y'Z'}_{l_1'l_2'l_3'}$ are not distinct signals if there exists a permutation $\sigma$ that simultaniously maps $X'Y'Z'$ to $XYZ$ and $l_1'l_2'l_3'$ to $l_1l_2l_3$.
+The rule is that $B^{XYZ}_{\ell_1\ell_2\ell_3}$ and $B^{X'Y'Z'}_{\ell_1'\ell_2'\ell_3'}$ are not distinct signals if there exists a permutation $\sigma$ that simultaniously maps $X'Y'Z'$ to $XYZ$ and $\ell_1'\ell_2'\ell_3'$ to $\ell_1\ell_2\ell_3$.
 It turns out that we can write a sum over distinct signals explicitly as
 \begin{gather*}
-    \sum_{\text{distinct signals}} = \underbrace{\sum_{l_1=l_2=l_3}\sum_{[XYZ]}}_{\text{sum }1} + \underbrace{\sum_{l_1=l_2\neq l_3}\sum_{[XY]Z}}_{\text{sum }2} + \underbrace{\sum_{l_1<l_2<l_3}\sum_{XYZ}}_{\text{sum }3}.
+    \sum_{\text{distinct signals}} = \underbrace{\sum_{\ell_1=\ell_2=\ell_3}\sum_{[XYZ]}}_{\text{sum }1} + \underbrace{\sum_{\ell_1=\ell_2\neq \ell_3}\sum_{[XY]Z}}_{\text{sum }2} + \underbrace{\sum_{\ell_1<\ell_2<\ell_3}\sum_{XYZ}}_{\text{sum }3}.
 \end{gather*}
@@
 The $Z$ value does not matter because any $Z$ value is accounted for.
-For sum 3 we can argue that we can switch around the order of the $l$'s to satisfy $l_1 < l2 < l3$ and the corresponding $XYZ$ configuration will be accounted for in sum 3 because all $XYZ$ combinations are counted.
+For sum 3 we can argue that we can switch around the order of the $l$'s to satisfy $\ell_1 < l2 < l3$ and the corresponding $XYZ$ configuration will be accounted for in sum 3 because all $XYZ$ combinations are counted.
 Finally, it is simple to verify that no distinct signal is counted more than once within each sum.
@@
 The covariance matrix now becomes a block matrix with each block corresponding to an $l_i$ and $l_i'$ configuration.
-When wick contracting using the gaussian approximation, every block matrix where $(l_1,l_2,l_3)\neq (l_1',l_2',l_3')$ vanishes.
+When wick contracting using the gaussian approximation, every block matrix where $(\ell_1,\ell_2,\ell_3)\neq (\ell_1',\ell_2',\ell_3')$ vanishes.
 It can then be shown that the entries of each block matrix are given as
 \begin{gather*}
-    \br{\Cov_{l_1l_2l_3}}^{XYZ,X'Y'Z'} = \tilde C^{XX'}_{l_1}\tilde C^{YY'}_{l_2}\tilde C^{ZZ'}_{l_3} + \delta_{l_1l_2}\tilde C^{XY'}_{l_1}\tilde C^{YX'}_{l_2}\tilde C^{ZZ'}_{l_3} +\delta_{l_2l_3}\tilde C^{XX'}_{l_1}\tilde C^{YZ'}_{l_2}\tilde C^{ZX'}_{l_3} \\ + \delta_{l_3l_1}\tilde C^{XZ'}_{l_1}\tilde C^{YY'}_{l_2}\tilde C^{ZX'}_{l_3} +\delta_{l_1l_2}\delta_{l_2l_3}\br{\tilde C^{XY'}_{l_1}\tilde C^{YZ'}_{l_2}\tilde C^{ZX'}_{l_3} +\tilde C^{XZ'}_{l_1}\tilde C^{YX'}_{l_2}\tilde C^{ZY'}_{l_3}}.
+    \br{\Cov_{\ell_1\ell_2\ell_3}}^{XYZ,X'Y'Z'} = \tilde C^{XX'}_{\ell_1}\tilde C^{YY'}_{\ell_2}\tilde C^{ZZ'}_{\ell_3} + \delta_{\ell_1\ell_2}\tilde C^{XY'}_{\ell_1}\tilde C^{YX'}_{\ell_2}\tilde C^{ZZ'}_{\ell_3} +\delta_{\ell_2\ell_3}\tilde C^{XX'}_{\ell_1}\tilde C^{YZ'}_{\ell_2}\tilde C^{ZX'}_{\ell_3} \\ + \delta_{\ell_3\ell_1}\tilde C^{XZ'}_{\ell_1}\tilde C^{YY'}_{\ell_2}\tilde C^{ZX'}_{\ell_3} +\delta_{\ell_1\ell_2}\delta_{\ell_2\ell_3}\br{\tilde C^{XY'}_{\ell_1}\tilde C^{YZ'}_{\ell_2}\tilde C^{ZX'}_{\ell_3} +\tilde C^{XZ'}_{\ell_1}\tilde C^{YX'}_{\ell_2}\tilde C^{ZY'}_{\ell_3}}.
 \end{gather*}
@@
 \begin{equation*}
-F_{\alpha\beta} = \sum_{l_1 \leq l_2 \leq l_3} \frac{\mathcal P _{l_1l_2l_3}}{6}
+F_{\alpha\beta} = \sum_{\ell_1 \leq \ell_2 \leq \ell_3} \frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6}
 \sum_{XYZ}\sum_{X'Y'Z'} 
-\partial_\alpha B^{X Y Z}_{l_1 l_2 l_3} 
-(\tilde C^{-1})^{X X'}_{l_1}
-(\tilde C^{-1})^{Y Y'}_{l_2}
-(\tilde C^{-1})^{Z Z'}_{l_3}
-\partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3}
+\partial_\alpha B^{X Y Z}_{\ell_1 \ell_2 \ell_3} 
+(\tilde C^{-1})^{X X'}_{\ell_1}
+(\tilde C^{-1})^{Y Y'}_{\ell_2}
+(\tilde C^{-1})^{Z Z'}_{\ell_3}
+\partial_\beta B^{X' Y' Z'}_{\ell_1 \ell_2 \ell_3}
 \end{equation*}
@@
 \end{equation*}
-and $\mathcal P_{l_1l_2l_3}$ is defined as the number of distinct permutations that can be made with $l_1l_2l_3$.
+and $\mathcal P_{\ell_1\ell_2\ell_3}$ is defined as the number of distinct permutations that can be made with $\ell_1\ell_2\ell_3$.
 This form was, for example, used in \cite{Kalaja_2021}\footnote{Note that in \cite{Kalaja_2021} this is based on a previous equation summing over \textit{all} $l_i$ (so including permutations of each configuration) which is missing a factor of $1/6$.}.
@@
 \begin{align*}
-    F_{\alpha\beta} &= \sum_{l_1=l_2=l_3} \sum_{[XYZ]}\sum_{[X'Y'Z']} \partial_{\alpha} B^{XYZ}_{l_1 l_2 l_3}\br{\frac{\mathcal P _{l_1l_2l_3}}{6}\sum_{d.
+    F_{\alpha\beta} &= \sum_{\ell_1=\ell_2=\ell_3} \sum_{[XYZ]}\sum_{[X'Y'Z']} \partial_{\alpha} B^{XYZ}_{\ell_1 \ell_2 \ell_3}\br{\frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6}\sum_{d.
 p.
@@
 p.
-X'Y'Z'}(\tilde C^{-1})^{X X'}_{l_1}
-    (\tilde C^{-1})^{Y Y'}_{l_2}
-    (\tilde C^{-1})^{Z Z'}_{l_3}}
-    \partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3} \\
-    &+ \sum_{l_1=l_2\neq l_3} \sum_{[XY]Z}\sum_{[X'Y']Z'} \partial_{\alpha} B^{XYZ}_{l_1 l_2 l_3}\br{\frac{\mathcal P _{l_1l_2l_3}}{6}\sum_{d. p. XY} \sum_{d. p. X'Y'}(\tilde C^{-1})^{X X'}_{l_1}
-    (\tilde C^{-1})^{Y Y'}_{l_2}
-    (\tilde C^{-1})^{Z Z'}_{l_3}}
-    \partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3} \\
-    &+ \sum_{l_1<l_2<l_3} \sum_{XYZ}\sum_{X'Y'Z'} \partial_{\alpha} B^{XYZ}_{l_1 l_2 l_3}(\tilde C^{-1})^{X X'}_{l_1}
-    (\tilde C^{-1})^{Y Y'}_{l_2}
-    (\tilde C^{-1})^{Z Z'}_{l_3}
-    \partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3},
+X'Y'Z'}(\tilde C^{-1})^{X X'}_{\ell_1}
+    (\tilde C^{-1})^{Y Y'}_{\ell_2}
+    (\tilde C^{-1})^{Z Z'}_{\ell_3}}
+    \partial_\beta B^{X' Y' Z'}_{\ell_1 \ell_2 \ell_3} \\
+    &+ \sum_{\ell_1=\ell_2\neq \ell_3} \sum_{[XY]Z}\sum_{[X'Y']Z'} \partial_{\alpha} B^{XYZ}_{\ell_1 \ell_2 \ell_3}\br{\frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6}\sum_{d.
+p.
+XY} \sum_{d.
+p.
+X'Y'}(\tilde C^{-1})^{X X'}_{\ell_1}
+    (\tilde C^{-1})^{Y Y'}_{\ell_2}
+    (\tilde C^{-1})^{Z Z'}_{\ell_3}}
+    \partial_\beta B^{X' Y' Z'}_{\ell_1 \ell_2 \ell_3} \\
+    &+ \sum_{\ell_1<\ell_2<\ell_3} \sum_{XYZ}\sum_{X'Y'Z'} \partial_{\alpha} B^{XYZ}_{\ell_1 \ell_2 \ell_3}(\tilde C^{-1})^{X X'}_{\ell_1}
+    (\tilde C^{-1})^{Y Y'}_{\ell_2}
+    (\tilde C^{-1})^{Z Z'}_{\ell_3}
+    \partial_\beta B^{X' Y' Z'}_{\ell_1 \ell_2 \ell_3},
 \end{align*}
@@
 This can be checked.
-For example, for the $l_1 = l_2 = l_3$ sum the multiplication of block matrices corresponding to the same $l_i$ configuration can be written as:
+For example, for the $\ell_1 = \ell_2 = \ell_3$ sum the multiplication of block matrices corresponding to the same $l_i$ configuration can be written as:
 \begin{gather*}
-    \sum_{[X'Y'Z']} \br{\frac{\mathcal P _{l_1l_2l_3}}{6} \sum_{d. p. XYZ} \sum_{d. p. X'Y'Z'}(\tilde C^{-1})^{X X'}_{l}
+    \sum_{[X'Y'Z']} \br{\frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6} \sum_{d. p. XYZ} \sum_{d. p. X'Y'Z'}(\tilde C^{-1})^{X X'}_{l}
     (\tilde C^{-1})^{Y Y'}_{l}
     (\tilde C^{-1})^{Z Z'}_{l}} \br{\tilde C^{X'X''}_l \tilde C^{Y'Y''}_l \tilde C^{Z'Z''}_l + \text{perms } X''Y''Z''} \\
-    = \frac{\mathcal P _{l_1l_2l_3}}{6}\sbr{\br{\sum_{d. p. XYZ} \sum_{X'Y'Z'}(\tilde C^{-1})^{X X'}_{l}
+    = \frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6}\sbr{\br{\sum_{d. p. XYZ} \sum_{X'Y'Z'}(\tilde C^{-1})^{X X'}_{l}
     (\tilde C^{-1})^{Y Y'}_{l}
     (\tilde C^{-1})^{Z Z'}_{l}} \tilde C^{X'X''}_l \tilde C^{Y'Y''}_l \tilde C^{Z'Z''}_l} + \text{perms } X''Y''Z'' \\
-    = \frac{\mathcal P _{l_1l_2l_3}}{6} \sum_{d. p. XYZ} \delta_{XX''}\delta_{YY''}\delta_{ZZ''} + \text{perms } X''Y''Z''
-    = \frac{\mathcal P _{l_1l_2l_3}}{6} \delta_{[XYZ], [X''Y''Z'']} + \text{perms } X''Y''Z''\\
+    = \frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6} \sum_{d. p. XYZ} \delta_{XX''}\delta_{YY''}\delta_{ZZ''} + \text{perms } X''Y''Z''
+    = \frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6} \delta_{[XYZ], [X''Y''Z'']} + \text{perms } X''Y''Z''\\
     = \delta_{[XYZ], [X''Y''Z'']} .
 \end{gather*}
-The sum over the different wick contractions will similarly cancel with the $\mathcal P_{l_1l_2l_3} / 6$ factor for the $l_1 = l_2 \neq l_3$ sum.
-For the $l_1 < l_2 < l_3$ sum the proof is similar as well except no cancellation is required.
+The sum over the different wick contractions will similarly cancel with the $\mathcal P_{\ell_1\ell_2\ell_3} / 6$ factor for the $\ell_1 = \ell_2 \neq \ell_3$ sum.
+For the $\ell_1 < \ell_2 < \ell_3$ sum the proof is similar as well except no cancellation is required.
 
@@
 To quantify the detectability of the lensing spectra, we introduce an overal amplitude of our signal, $A$, with fiducial value 1 as experimental parameter and compute $F_{AA}$.
-Obviously, $\partial_A\br{A B^{XYZ}_{l_1l_2l_3}}|_{A=1} = B^{XYZ}_{l_1l_2l_3}$, so we find
+Obviously, $\partial_A\br{A B^{XYZ}_{\ell_1\ell_2\ell_3}}|_{A=1} = B^{XYZ}_{\ell_1\ell_2\ell_3}$, so we find
 \begin{equation*}
     \br{\frac{S}{N}}^2 := F_{AA} =
-\sum_{XYZ, X'Y'Z'} \sum_{l_1\leq l_2 \leq l_3} \frac{\mathcal P _{l_1l_2l_3}}{6}
-B^{X Y Z}_{l_1 l_2 l_3} 
-(\tilde C^{-1})^{X X'}_{l_1}
-(\tilde C^{-1})^{Y Y'}_{l_2}
-(\tilde C^{-1})^{Z Z'}_{l_3}
-B^{X' Y' Z'}_{l_1 l_2 l_3}.
+\sum_{XYZ, X'Y'Z'} \sum_{\ell_1\leq \ell_2 \leq \ell_3} \frac{\mathcal P _{\ell_1\ell_2\ell_3}}{6}
+B^{X Y Z}_{\ell_1 \ell_2 \ell_3} 
+(\tilde C^{-1})^{X X'}_{\ell_1}
+(\tilde C^{-1})^{Y Y'}_{\ell_2}
+(\tilde C^{-1})^{Z Z'}_{\ell_3}
+B^{X' Y' Z'}_{\ell_1 \ell_2 \ell_3}.
 \end{equation*}
@@
 % Computing the inverse is then reduced to computing the inverse of the individual block matrices.
-The block matrices for sum 1 and 2 don't have a particularly nice form (cite paper about multiple tracer fisher matrices for bispectra) but for sum 3 the inverse of the block matrix corresponding to $l_1l_2l_3$ is
-% \begin{equation*}
-%     (\Cov_{l_1l_2l_3}^{-1})^{XYZ, X'Y'Z'} = \hat C_{l_1}^{X,X'}\hat C_{l_2}^{Y,Y'}\hat C_{l_3}^{Z,Z'}.
-% \end{equation*}
-% \begin{equation*}
-%     (\Cov)^{XYZ, X'Y'Z'} = C_{l_1}^{X,X'} C_{l_2}^{Y,Y'} C_{l_3}^{Z,Z'}.
+The block matrices for sum 1 and 2 don't have a particularly nice form (cite paper about multiple tracer fisher matrices for bispectra) but for sum 3 the inverse of the block matrix corresponding to $\ell_1\ell_2\ell_3$ is
+% \begin{equation*}
+%     (\Cov_{\ell_1\ell_2\ell_3}^{-1})^{XYZ, X'Y'Z'} = \hat C_{\ell_1}^{X,X'}\hat C_{\ell_2}^{Y,Y'}\hat C_{\ell_3}^{Z,Z'}.
+% \end{equation*}
+% \begin{equation*}
+%     (\Cov)^{XYZ, X'Y'Z'} = C_{\ell_1}^{X,X'} C_{\ell_2}^{Y,Y'} C_{\ell_3}^{Z,Z'}.
 % \end{equation*}
@@
 % \begin{equation*}
-%     \sum_{X'Y'Z'} = \hat C_{l_1}^{X,X'}\hat C_{l_2}^{Y,Y'}\hat C_{l_3}^{Z,Z'} C_{l_1}^{X'X''}C_{l_2}^{Y'Y''}C_{l_3}^{Z'Z''} = \delta_{XX''}\delta_{YY''}\delta_{ZZ''} = \delta_{XYZ, X''Y''Z''}
+%     \sum_{X'Y'Z'} = \hat C_{\ell_1}^{X,X'}\hat C_{\ell_2}^{Y,Y'}\hat C_{\ell_3}^{Z,Z'} C_{\ell_1}^{X'X''}C_{\ell_2}^{Y'Y''}C_{\ell_3}^{Z'Z''} = \delta_{XX''}\delta_{YY''}\delta_{ZZ''} = \delta_{XYZ, X''Y''Z''}
 % \end{equation*}
@@
 % It's worth mentioning that due to all $l$'s being distinct in each configuration of sum 3 the Wick contraction corresponding to each element of the covariance matrix can only be done in one way and we thus only have one term.
-In the sum 2 we get two terms due to $l_1$ and $l_2$ being identical.
+In the sum 2 we get two terms due to $\ell_1$ and $\ell_2$ being identical.
 In sum 3 we similarly get 6 terms corresponding to all possible Wick contractions being nonzero.
@@
 
-Note that there it isn't obvious whether to define these coordinates at the point where the lensed light hits $S^2$ or the unlensed light hits $S^2$.
+Note that it is not obvious whether to define these coordinates at the point where the lensed light hits $S^2$ or the unlensed light hits $S^2$.
 We will assume that lensing effects are sufficiently weak that either definition works.
@@
 Evaluated at our point of interest we obtain $\partial_\theta=\partial_{\tilde y}$ and $\partial_\phi=\sin\theta_0\partial_{\tilde x}$.
-The second order derivatives can then be obtained using the first order derivatives.
+The second-order derivatives can then be obtained using the first-order derivatives.
 We can immediately evaluate them at the point to get
@@
 As long as numerical noise doesn't dominate, it can be shown that the relative error in our approximation is approximately 5 times the relative difference that taking $h \rightarrow h(1 \pm 0.1)$ leads to.
-We thus conclude that, based on this test, the derivatives are almost always computed with up to one percent of error.
+We thus conclude that, based on this test, the derivatives are almost always computed with up to one percent error.
 The exception is the derivative with respect to neutrino masses, which introduces a larger error.
@@
     \includegraphics[height=\textheight]{../code/plots/derivative_accuracy.png}
-    \caption{Central difference approximation of derivative of equilateral CMB weak lensing bispectrum (left) and powerspectrum (right) for differences as specified in table \ref{tab:cosmo-params-diff} as well as for differences $5\%$ and $10\%$ larger and smaller.}
+    \caption{Central difference approximation of derivative of equilateral CMB weak lensing bispectrum (left) and power spectrum (right) for differences as specified in table \ref{tab:cosmo-params-diff} as well as for differences $5\%$ and $10\%$ larger and smaller.}
     \label{fig:deraccuracy}
@@
     \centering
-    \scriptsize
+    \tiny
     \begin{tabular}{|l|r|rrr|rrr|rrr|}
         \hline
+        \multicolumn{11}{|c|}{weak priors} \\
+        \hline
         && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
@@
         \hline
-         $H_0$          &   25.70 &      25.56 &                      24.62 &                               22.49 &       8.23 &                       7.02 &                                1.90 &       7.52 &                       3.56 &                                0.92 \\
-         $\Omega_b h^2$ &    2.23 &       2.23 &                       2.23 &                                2.23 &       2.23 &                       2.22 &                                2.21 &       2.22 &                       2.20 &                                2.19 \\
-         $\Omega_c h^2$ &  240.56 &       8.91 &                      11.16 &                                7.83 &       4.51 &                       3.22 &                                2.15 &       3.10 &                       2.01 &                                1.33 \\
-         $n_s$          &    2.07 &       2.06 &                       2.07 &                                2.05 &       2.03 &                       1.90 &                                0.88 &       1.44 &                       1.44 &                                0.67 \\
-         $A_s$          &   47.62 &      45.54 &                      44.79 &                               29.26 &      10.63 &                       4.57 &                                3.50 &       6.93 &                       4.01 &                                2.84 \\
+        $H_0$          &     1.73e+01 &  1.65e+01 &  1.65e+01 &        6.91e+00 & 6.39e+00 & 5.04e+00 &      1.65e+00 &      4.13e+00 &      2.88e+00 &             6.65e-01 \\
+        $\Omega_b h^2$ &     5.00e-04 &  5.00e-04 &  5.00e-04 &        5.00e-04 & 5.00e-04 & 4.98e-04 &      4.95e-04 &      4.86e-04 &      4.94e-04 &             4.65e-04 \\
+        $\Omega_c h^2$ &     2.89e-01 &  6.99e-03 &  1.35e-02 &        6.42e-03 & 5.22e-03 & 4.74e-03 &      2.04e-03 &      1.07e-03 &      2.63e-03 &             8.62e-04 \\
+        $n_s$          &     2.00e-02 &  1.90e-02 &  1.99e-02 &        1.82e-02 & 8.04e-03 & 1.89e-02 &      5.54e-03 &      2.98e-03 &      1.52e-02 &             2.34e-03 \\
+        $\tau$         &     6.30e-02 &  4.69e-04 &  6.30e-02 &        4.46e-04 & 6.30e-02 & 6.30e-02 &      6.30e-02 &      5.20e-05 &      6.30e-02 &             5.07e-05 \\
+        $A_s$          &     1.00e-09 &  3.64e-10 &  9.29e-10 &        3.07e-10 & 2.56e-10 & 1.16e-10 &      7.27e-11 &      8.39e-11 &      1.10e-10 &             4.51e-11 \\
+        \hline
+        \multicolumn{11}{|c|}{CMB $T + E$ priors} \\
+        \hline
+        && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
+        \hline
+         Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
+        \hline
+        $H_0$          &    1.16e+00 &  1.00e+00 &  1.05e+00 &        1.00e+00 & 1.02e+00 & 9.47e-01 &      7.09e-01 &      9.32e-01 &      9.03e-01 &             5.17e-01 \\
+        $\Omega_b h^2$ &    5.62e-05 &  5.16e-05 &  5.58e-05 &        5.15e-05 & 5.38e-05 & 5.26e-05 &      5.10e-05 &      4.73e-05 &      5.08e-05 &             4.69e-05 \\
+        $\Omega_c h^2$ &    7.13e-04 &  6.03e-04 &  7.10e-04 &        5.94e-04 & 4.76e-04 & 5.70e-04 &      3.87e-04 &      1.63e-04 &      4.41e-04 &             1.54e-04 \\
+        $n_s$          &    2.29e-03 &  2.13e-03 &  2.25e-03 &        2.12e-03 & 2.00e-03 & 2.22e-03 &      1.96e-03 &      1.43e-03 &      2.01e-03 &             1.38e-03 \\
+        $\tau$         &    2.44e-03 &  4.30e-05 &  2.44e-03 &        4.27e-05 & 2.41e-03 & 2.42e-03 &      2.40e-03 &      2.92e-05 &      2.42e-03 &             2.80e-05 \\
+        $A_s$          &    1.05e-11 &  4.76e-12 &  1.04e-11 &        4.71e-12 & 1.02e-11 & 1.03e-11 &      9.95e-12 &      2.81e-12 &      1.01e-11 &             2.65e-12 \\
        \hline
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figures/param_constraints_all_cmb.pdf}
+    \includegraphics[width=\textwidth]{figures/param_constraints_all_cmb.pdf}
     \caption{Same as figure \ref{fig:paramconstraintstightcmb}, except for the standard $\Lambda$CDM parameters.}
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figures/param_constraints_all_gal.pdf}
+    \includegraphics[width=\textwidth]{figures/param_constraints_all_gal.pdf}
     \caption{Same as figure \ref{fig:paramconstraintstightgal}, except for the standard $\Lambda$CDM parameters.}

### 2025-06-04 paper/playing_with_fish.py
Update playing_with_fish.py

@@ -17,11 +17,11 @@ fisher_matrices_dir = 'code/fisher_matrices'
 ######################
 # Bispectra Matrices
 ######################
-visb_c = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_c.txt')
+visb_c = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_c.txt')
 
-visb_s = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_s.txt')
+visb_s = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_s.txt')
 
-visb_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_approx_both.txt')
+visb_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_both.txt')
 
 ########################
 # Powerspectra Matrices
@@ -36,11 +36,11 @@ visp_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_both.txt')
 ######################
 # Planck
 ######################
-planck = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_c_planck.txt')
+#planck = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_c_planck.txt')
 
-planck_prior = np.diag([12 / 60**2, 1/(0.0005)**2, 12 / 1**2, 1/(0.02)**2, 0.01, 1e18, 0.01])
+planck_prior = np.diag([12 / 60**2, 1/(0.0005)**2, 12 / 1**2, 1/(0.02)**2, 0.01, 0.063**(-2), 1e18, 0.01])
 
-planck_p_prior = planck + planck_prior
+#planck_p_prior = planck + planck_prior
 
 ######################
 # CMB Stage 4
@@ -120,15 +120,16 @@ def append_row_column(mats, vec):
     return matsnew
 
 # update this
+
 s8_ders = np.array([
-    2.76945054e-03, # H_0
-    -6.50570238e+00, # Omega_b h^2 
-    4.30234556e+00,  # Omega_c h^2
-    3.06272435e-01,# n_s
-    -2.15681667e-01, # m_\nu
-    0,              # tau
-    1.93167857e+08, # A_s
-    -2.01175433e-01  # w_0
+    2.81584114e-03, # H_0
+    -6.50119625e+00, # Omega_b h^2 
[…]

### 2025-06-06 paper/paper_draft.tex
paper changes

@@
 
-Weak lensing has been used as a probe to constrain cosmological parameters for over a decade \cite{kilbinger2015cosmology,Bacon2000,Kaiser2000,Waerbeke2000,Bartelmann2001}.
+Weak lensing has been used as a probe to constrain cosmological parameters for over a decade \cite{kilbinger2015cosmology,Bacon2000,Kaiser2000,Waerbeke2000,Bartelmann2001,Planck2018Lensing}.
 A range of upcoming surveys, such as the survey by the Simons Observatory (SO) \cite{Ade2019}, the Legacy Survey of Space and Time (LSST) \cite{Ivezic2019}, and Euclid \cite{Laureijs2011} aims to measure weak lensing of the cosmic microwave background (CMB) and galaxies respectively.
@@
 In addition to looking at the power spectrum, a natural next step is thus to measure weak lensing bispectra.
-While the bispectrum has been measured for galaxy lensing signals \cite{vanWaerbeke2002}, a definitive detection of the CMB lensing bispectrum is still pending.
+While the bispectrum has been measured for galaxy lensing signals \cite{vanWaerbeke2002}, a definitive detection of the CMB lensing bispectrum is still pending \cite{Kalaja_2023,Namikawa_2016}.
 Upcoming surveys, such as the ones mentioned earlier may be able to detect this signal.
@@
 
-All code used to derive the resukts in this paper is publically available on GitHub at \url{https://github.com/Jonas-Frugte/fisher_calc_weak_lensing}.
+All code used to derive the results in this paper is publically available on GitHub at \url{https://github.com/Jonas-Frugte/fisher_calc_weak_lensing}.
 
@@
 These are directly related to the matter power and bispectra as
-\begin{empheq}[box=\fbox]{align*}
+\begin{align}%[box=\fbox]{align*}
     C^{\psi_X\psi_Y}_\ell
     & = \frac{9}{\ell^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)P^\delta(\ell/\chi,\eta_0-\chi),\\
-    B^{\psi_X\psi_Y\psi_Z}_{\el\ell_1\el\ell_2\el\ell_3} &= \sqrt{\frac{(2\ell_1 + 1)(2\ell_2 + 1)(2\ell_3 + 1)}{4\pi}} \begin{pmatrix} \ell_1 & \ell_2 & \ell_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{\ell_1^2\ell_2^2\ell_3^2}\Omega_m^3H_0^6\\
+    B^{\psi_X\psi_Y\psi_Z}_{\ell\ell_1\ell\ell_2\ell\ell_3} &= \sqrt{\frac{(2\ell_1 + 1)(2\ell_2 + 1)(2\ell_3 + 1)}{4\pi}} \begin{pmatrix} \ell_1 & \ell_2 & \ell_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{\ell_1^2\ell_2^2\ell_3^2}\Omega_m^3H_0^6\\
     & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{\ell_i/\chi\}, \eta_0-\chi),
-\end{empheq}
+\end{align}
 with $X, Y, Z \in \{\CMB, \gal\}$.
@@
 \end{equation*}
+If only some fraction of the sky, $f_{\text{sky}}$, is measured, all Fisher matrices are multiplied by $f_{\text{sky}}$ \cite{Takada2003}.
 
@@
 % TODO: find proper references for stage 3 and stage 4 names as well as experimental parameters and real life surveys
-Our fiducial cosmology is derived from the Planck results \cite{planckresults} (see table \ref{tab:fiducialpars}).
-For all parameter constraints we use a prior similar to the one used in \cite{Planck2018Lensing} (see table \ref{tab:fiducialpars}).
+Our fiducial cosmology is based on the Planck results \cite{planckresults} (see table \ref{tab:fiducialpars}).
+Throughout this paper we consider two priors.
+The first we refer to as our ``weak prior'' and is similar to the one used in \cite{Planck2018Lensing} (see table \ref{tab:fiducialpars}).
 This prior only restricts $\Omega_b h^2$ and $n_s$ significantly.
+The purpose of the weak prior is as follows.
 The power spectra, in particular, exhibit complete insensitivity to certain parameters.
 Consequently, relying solely on them to constrain the entire cosmological model results in very poor constraints.
-Future lensing surveys will of course use information from surveys measuring other statistics such as galaxy distributions or primary CMB anisotropies, so not using any priors is not representative of what stage 3 and 4 experiments could achieve.
-Even a weak prior improves the constraining power of our power spectra considerably.
+Future lensing surveys will of course use information from surveys measuring other statistics such as galaxy distributions or primary CMB anisotropies, so these poor constraints are not representative of what stage 3 and 4 experiments could achieve.
+By adding a prior our calculations become more realistic.
+Additionally, by having the prior be conservative, it stays clear how the different lensing spectra complement each other.
+The other prior we will use is based on the CMB temperature and $E$-mode polarization cross- and auto-powerspectra from $\ell=30$ to $\ell = 2000$.
+With the same noise values as used for the CMB weak lensing spectra.
+These results will be more realistic and serve to test whether weak lensing spectra are able to further tighten constraints in the presence of surveys measuring more standard statistics (such as the primary CMB anisotropies) with similar noise values.
 \begin{table}[h!]
     \centering
-    \begin{tabular}{|c|c|c|c|}
-        \hline
-        \textbf{parameter} & \textbf{notation} & \textbf{fiducial value} & \textbf{$\sigma$ of prior}\\
-        \hline
-        Hubble constant & $H_0$ & $67.4 \, \text{km/s/Mpc}$ & 17.3\\
-        \hline
-        Physical baryon density parameter & $\Omega_b h^2$ & $0.0223$ & 0.0005\\
-        \hline
-        Physical cold dark matter density parameter & $\Omega_c h^2$ & $0.119$ & $0.288$\\
-        \hline
-        Scalar spectral index & $n_s$ & 0.965 & 0.02 \\
-        \hline
-        Amplitude of primordial scalar fluctuations & $A_s$ & $2.13 \times 10^{-9}$ & $1\times 10^{-9}$ \\
-        \hline
-        Sum of neutrino masses & $\sum m_\nu$ & $0.06$ & $\infty$\\
-        \hline
-        Dark energy equation of state parameter & $w_0$ & -1 & $\infty$\\ \hline
+    \scriptsize
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        \textbf{parameter} & \textbf{notation} & \textbf{fiducial value} & \textbf{$\sigma$ of weak prior} & \textbf{$\sigma$ of $T + E$ prior}\\
+        \hline
+        Hubble constant & $H_0$ & $67.4 \, \text{km/s/Mpc}$ & 17.3 & 1.21\\
+        \hline
+        Physical baryon density parameter & $\Omega_b h^2$ & $0.0223$ & 0.0005 & 0.000057\\
+        \hline
+        Physical cold dark matter density parameter & $\Omega_c h^2$ & $0.119$ & $0.288$ & 0.00083\\
+        \hline
+        Scalar spectral index & $n_s$ & 0.965 & 0.02 & 0.0025 \\
+        \hline
+        Reionization depth & $\tau$ & 0.063 & 0.063 & 0.0013 \\
+        \hline
+        Amplitude of primordial scalar fluctuations & $A_s$ & $2.13 \times 10^{-9}$ & $1\times 10^{-9}$ & $5.5\times 10^{-11}$\\
+        \hline
+        Sum of neutrino masses & $\sum m_\nu$ & $0.06$ & $\infty$ & 0.22\\
+        \hline
+        Dark energy equation of state parameter & $w_0$ & -1 & $\infty$ & 0.058 \\ 
+        \hline
     \end{tabular}
@@
 A comparison of the noise power spectra to the lensing power spectra are shown     in figure \ref{fig:lpsplusnoise}.
-Effectively, for galaxy lensing, the power spectrum is signal-dominated up until $\ell \sim 200$ (stage 2) and $\ell \sim 700$ (stage 4).
+Effectively, for galaxy lensing, the power spectrum is signal-dominated up until $\ell \sim 200$ (stage 3) and $\ell \sim 700$ (stage 4).
 For CMB lensing, the power spectrum is signal-dominated for $\ell \sim 300$ (stage 3) and $\ell \sim 1000$ (stage 4).
@@
 
-CMB lensing noise is estimated using quadratic estimator \cite{cmblensingestimator}.
-We consider experiments that only measure polarization noise.
-This is similarly done in \cite{Namikawa_2016}.
-The parameters characterizing the noise levels are thus beam width, $\sigma$, and polarization white noise, $\Delta_P$.
+CMB lensing noise is estimated using the quadratic estimator from \cite{cmblensingestimator}.
+% We consider experiments that only measure polarization noise. This is similarly done in \cite{Namikawa_2016}. 
+The parameters characterizing the noise levels are beam width, $\sigma$, polarization white noise, $\Delta_P$, and temperature white noise $\Delta_T$.
+We take $\Delta_T = \Delta_P / \sqrt 2$ throughout.
 Galaxy lensing is determined by measuring lensing shear (see appendices \ref{sec:weaklensing} and \ref{sec:shear}).
@@
 \end{equation*}
+In all cases we assume that the proportion of the sky surveyed, $f_{\text{sky}}$, equals 0.5.
 For parameter constraints, we will consider only stage 4 noise levels.
@@
 Derivatives are calculated with a central difference formula (see appendix \ref{sec:derivatives} for convergence).
-To check the accuracy of our results we calculated constraints for similar experimental parameters as in references \cite{Planck2018Lensing, Takada2003, Namikawa_2016} and found good agreement in all cases.
+To check the accuracy of our results we calculated constraints for similar experimental parameters as in references \cite{Planck2018Lensing, Takada2003, cmbs4sciencebook,neutrinoconstraints} and found good agreement in all cases.
 
@@
 CMB lensing bispectra could be  detectable by a stage 3 experiment (see e.g.
-ref.~\cite{} for the effect of noise biases) and with high signal-to-noise with a stage 4 experiment.
+ref.
+\cite{Kalaja_2023} for the effect of noise biases) and with high signal-to-noise with a stage 4 experiment.
 CMB $S/N$ results match those of reference \cite{Namikawa_2016}.
@@
     \includegraphics[width=\textwidth]{figures/snrplots.pdf}
-    \caption{Signal-to-noise ratios for galaxy lensing (dashed) and CMB lensing (non-dashes) power- (in red) and bispectra (in blue).
-The $x$-axis labels the maximum multipole moment measured up to, starting from $\ell_{\min}=2$.
-SAY SOMETHING ABOUT LMIN FOR RECONSTRUCTION The maximum multipole used for CMB lensing reconstruction is $\ell = 10^4$.
-Using lower maximum multipole such as 3000 doesn't significantly affect results.}
+    \caption{Signal-to-noise ratios for galaxy lensing (left) and CMB lensing (right) for stage 3 (dashed) and stage 4 (not dashed) surveys as a function of maximum multipole measured.
+The minimum multipole is always $l_{\text{min}}=2$.
+The multipole range used for CMB lensing reconstruction is $2\leq \ell \leq 10^4$.
+Using a more conservative range such as $30\leq \ell \leq 5000$ used to calculate Simons Observatory noise curves don't significantly affect the results.}
     \label{fig:snrplots}
@@
 We include both the power spectrum and the bispectrum of the weak lensing.
-We emphezize on several two extension of $\Lambda CDM$, $m_\nu$ and $w_0$ as well as $\sigma_8$, and $\Omega_m$.
-Constraints on other parameters can be found found in appendix \ref{sec:lambdacdmconstraints}.
-Results shown include marganlization over all parameters.
-Figures \ref{fig:paramconstraintstightcmb} and \ref{fig:paramconstraintstightgal} show confidence ellipses from stage 4 CMB and galaxy lensing surveys, respectively.
-Table \ref{tab:paramconstraintstight} contains the forecasted constraints as uncertainties in the percentage of the fiducial values.
-
-!!
-By itsef lensing power spectra are not very constraining, hence we have to assume weak priors. inflated constraints from lps, however still good bispectrum info. 
-!!
-
-Despite using priors that significantly increase power spectrum constraining power, constraints are significantly improved by including bispectrum information in nearly all cases.
-Again in agreement with reference \cite{Namikawa_2016}, in which case a 33 percent and 35 percent improvement was found for $m_\nu$ and $w_0$, respectively, by measuring the CMB lensing bispectrum in addition to the power spectrum.
-Constraints also improve significantly, though not quite as much, when combining CMB and galaxy lensing, compared to only measuring one of these signals.
+We look at two extension of $\Lambda CDM$, $m_\nu$ and $w_0$ as well as the derived parameters $\sigma_8$, and $\Omega_m$.
+Constraints on the full set of $\Lambda$CDM parameters can be found found in appendix \ref{sec:lambdacdmconstraints}.
+Results shown include marginalization over all parameters.
+Figures \ref{fig:paramconstraintstightcmb} and \ref{fig:paramconstraintstightgal} show confidence ellipses from stage 4 CMB and galaxy lensing surveys with the weak prior, respectively.
+Figures \ref{fig:paramconstraintstightcmbcmbprior} and \ref{fig:paramconstraintstightgalcmbprior} show confidence ellipses in the case of our CMB $T+E$ prior.
+Table \ref{tab:paramconstraintstight} lists all forecasted constraints .
+
+Constraints are significantly improved by including bispectrum information in nearly all cases when using weak priors.
+When including the CMB $T+E$ priors, both the CMB lensing power- and bispectra significantly improve prior constraints, however combining them doesn't lead to significant improvements compared to only using lensing power spectra.
+For galaxy lensing, the power- and bispectra yield nearly identical improvements compared to the CMB $T+E$ prior.
+In this case, combining both statistics leads to significantly tighter constraints (in contrast to CMB lensing).
+When combining CMB and galaxy lensing we see that the bispectrum is slightly less competitive than the lensing power spectrum, but can still significantly improve constraints compared to only using the lensing power spectrum.
+% Again in agreement with reference \cite{Namikawa_2016}, in which case a 33 percent and 35 percent improvement was found for $m_\nu$ and $w_0$, respectively, by measuring the CMB lensing bispectrum in addition to the power spectrum.
+
+Furthermore, it is clear that, in general, combining CMB and galaxy lensing surveys also leads to significant improvements in constraining power, even in the case of the CMB $T+E$ prior.
+The improvement is strongest for the neutrino constraint.
 
@@
     \item Lensing powers pectra alone are not sufficient to competitively constrain $\Lambda\mathrm{CDM} + w_0 + m_\nu$ without priors, even in the case of stage 4 surveys.
-    \item For low-enough noise values (in particular for stage 4 noise) the bispectra can contain more information than the power spectra .
-If we denote the multipole moment at which the noise surpasses the power spectrum signal as $\ell_{N>C}$, than the fisher matrices for the bispectra roughly scale as $F^{B}_{\alpha\beta}\propto \ell_{N>C}^3$ while for the powerspectra they scale as $F^{C}_{\alpha\beta}\propto \ell_{N>C}^2$ (with an extra multipole factor due to the $2\ell + 1$ factor in the power spectrum estimator covariance).
-The power spectrum is sufficiently bigger than the bispectrum that it contains more information given current-day noise levels, however as surveys get better the bispectra may eventually contain more information because of the more favorable $\ell_{N>C}$ scaling.
-REMOVE L SCALING ARGUMENT YAP ABOUT NON GAUSSIANITY INSTEAD AND EXPLAIN ABOUT DIFFERENCE BETWEEN KERNELS
+
+    \item For lower noise values (in particular for stage 4 noise) the bispectra becomes increasingly important compared to the powerspectra.
+The lensing bispectra are directly related to the amount of non-gaussianity in the matter distribution.
+As noise levels becomes lower, we are are able to measure the matter distribution on small enough scales where nonlinear matter evolution becomes apparent.
+Additionally, for galaxy lensing, the window function peaks at earlier redshifts than compared to CMB lensing.
+The bispectrum is largest at late times, and thus the galaxy lensing bispectrum is significantly easier to detect and is better at constraining parameters then the CMB lensing bispectrum.
 
     \item The information from bispectra is sensitive to cosmology in a different way than that of the power spectra.
-The main parameter combinations where approximate degeneracies are broken are $(n_s,\Omega_b h^2),\;(\Omega_b h^2,H_0),\;(\Omega_c h^2,H_0),\;(\Omega_c h^2,\Omega_b h^2)\;(n_s,\Omega_c h^2),\;(m_\nu,H_0),\;(m_\nu,n_s),$ $(A_s,\Omega_b h^2)$.
+The main parameter combinations where approximate degeneracies are broken are $(n_s,\Omega_b h^2),\;(\Omega_b h^2,H_0),\;(\Omega_c h^2,H_0),\;(\Omega_c h^2,\Omega_b h^2)\;(n_s,\Omega_c h^2),\;(m_\nu,H_0),\;(m_\nu,n_s),$ $(A_s,\Omega_b h^2)$ (see also appendix \ref{sec:lambdacdmconstraints}).
 This means that even if stage 4 surveys aren't able to reach the noise levels assumed in this paper, the bispectra will likely still lead to significantly tighter constraints.
 
-    \item Stage 4 galaxy weak lensing surveys contain significantly more information than CMB surveys.
-Comparing the confidence ellipses, it is clear that the two types of surveys depend on the underlying parameters in different ways.
-This leads to substantially better constraints for $\sigma_8$, $\Omega_m$, and $m_\nu$, (but not for $w_0$) when combining CMB and galaxy lensing.
+    \item Stage 4 galaxy weak lensing surveys generally contain more information than CMB surveys. 
+    % Comparing the confidence ellipses, it is clear that the two types of surveys depend on the underlying parameters in different ways.
+This leads to substantially better constraints for $\sigma_8$, $\Omega_m$, $m_\nu$, and $w_0$ when combining CMB and galaxy lensing.
 \end{itemize}
@@
         \hline
-        $m_\nu$    &    1.54e-01 &  5.23e-02 &  8.76e-02 &        5.18e-02 & 1.17e-01 & 1.13e-01 &      8.79e-02 &      1.87e-02 &      7.47e-02 &             1.62e-02 \\
-        $w_0$      &    4.64e-02 &  4.18e-02 &  4.54e-02 &        4.18e-02 & 4.06e-02 & 4.14e-02 &      2.25e-02 &      3.34e-02 &      3.90e-02 &             1.99e-02 \\
-        $\sigma_8$ &    1.38e-02 &  1.06e-02 &  1.16e-02 &        1.06e-02 & 9.86e-03 & 1.00e-02 &      6.99e-03 &      8.98e-03 &      9.72e-03 &             5.28e-03 \\
-        $\Omega_m$ &    1.39e-02 &  1.20e-02 &  1.26e-02 &        1.19e-02 & 1.26e-02 & 1.15e-02 &      9.03e-03 &      1.13e-02 &      1.07e-02 &             6.16e-03 \\
+        $m_\nu$    &    2.16e-01 &  5.28e-02 &  1.21e-01 &        5.22e-02 & 1.48e-01 & 1.49e-01 &      1.12e-01 &      1.88e-02 &      1.02e-01 &             1.62e-02 \\
+        $w_0$      &    5.78e-02 &  4.19e-02 &  5.30e-02 &        4.19e-02 & 4.66e-02 & 4.77e-02 &      2.55e-02 &      3.35e-02 &      4.33e-02 &             1.99e-02 \\
+        $\sigma_8$ &    1.55e-02 &  1.06e-02 &  1.50e-02 &        1.06e-02 & 1.18e-02 & 1.21e-02 &      8.18e-03 &      8.99e-03 &      1.19e-02 &             5.28e-03 \\
+        $\Omega_m$ &    1.50e-02 &  1.20e-02 &  1.46e-02 &        1.20e-02 & 1.34e-02 & 1.23e-02 &      9.32e-03 &      1.13e-02 &      1.19e-02 &             6.16e-03 \\
        \hline
         \end{tabular}
-\caption{Parameter constraints in percentage of the fiducial values for different combinations of weak lensing information.
+\caption{Parameter constraints for different combinations of weak lensing information.
 $C$ stands for powerspectra, and $B$ for bispectra.
 All surveys are assumed to be stage 4 surveys with multipole range $2 \leq \ell \leq 2000$.
-We use the prior described in table \ref{tab:fiducialpars}.}
+We use the weak and CMB priors described in table \ref{tab:fiducialpars}.}
 \label{tab:paramconstraintstight}
@@
 
+% CONFIDENCE ELLIPSES
+
 \begin{figure}
     \centering
-    \includegraphics[width=0.7\textwidth]{figures/param_constraints_tight_cmb.pdf}
+    \includegraphics[width=0.7\textwidth]{figures/param_constraints_tight_cmb_weak_prior.pdf}
     \caption{Parameter constraints and confidence ellipses for $\mu_\nu$, $w_0$, $\sigma_8$, and $\Omega_m$.
@@
 The colored plots show CMB lensing power- and/or bispectrum constraints.
-Priors used can be found in table \ref{tab:fiducialpars}.
+We use the weak priors listed in table \ref{tab:fiducialpars}.
 The confidence ellipses are for $1\sigma$ certainty.
@@
 \end{figure}
+
 \begin{figure}
     \centering
-    \includegraphics[width=0.7\textwidth]{figures/param_constraints_tight_gal.pdf}
+    \includegraphics[width=0.7\textwidth]{figures/param_constraints_tight_cmb_cmb_prior.pdf}
+    \caption{Same as figure \ref{fig:paramconstraintstightcmb}, except we use the CMB temperature and polarization based priors listed in table \ref{tab:fiducialpars}.}
+    \label{fig:paramconstraintstightcmbcmbprior}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.7\textwidth]{figures/param_constraints_tight_gal_weak_prior.pdf}
     \caption{Same as figure \ref{fig:paramconstraintstightcmb}, except here the colored plots show galaxy lensing constraints.}
     \label{fig:paramconstraintstightgal}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.7\textwidth]{figures/param_constraints_tight_gal_cmb_prior.pdf}
+    \caption{Same as figure \ref{fig:paramconstraintstightcmbcmbprior}, except here the colored plots show galaxy lensing constraints.}
+    \label{fig:paramconstraintstightgalcmbprior}
 \end{figure}
@@
 In this work, we have presented parameter forecasts from CMB lensing and galaxy lensing power- and bispectra, considering experimental parameters representative of Stage 3 and Stage 4 CMB and galaxy surveys.
-Our analysis shows that while lensing power spectra are generally well detectable, the inclusion of bispectra offers significant improvements when considering paramater constraints.
+Our analysis shows that while lensing power spectra are well detectable, the inclusion of bispectra generally offers significant improvements when considering parameter constraints.
 Notably, both CMB and galaxy bispectra are found to be measurable, even with Stage 3 experimental noise levels.
@@
 Better estimators exist; for example, the iterative estimator developed in \cite{Smith_2012} can lead to a significantly better signal-to-noise ratio for the CMB lensing bispectrum, as shown in \cite{Namikawa_2016}.
+Using such an iterative estimator may also result in the CMB bispectrum being able to significantly improve parameter constraints when combined with lensing power spectra and a CMB $T+E$ prior.
+
     \item To calculate the Fisher matrices, derivatives of the lensing spectra with respect to cosmological parameters were computed using a central difference formula.
@@
 However, for precise quantitative results, these factors would need to be revisited and improved.
+
+    \item A recent paper \cite{postborn} has shown that the Born approximation used to derive the formulas for the lensing power and bispectra, is no longer negligible for the lensing bispectrum.
+The post-Born corrections calculated in that paper only changed the lensing power spectra by $\leq 0.2\%$, while for the bispectra this effect was much larger, in particular for folded configurations.
+For future work it would thus be worthwile to analyze how these post-Born corrections affect the results presented in this paper.
 \end{itemize}
@@
 Now using equation \ref{eq:complexexp} to rewrite the complex exponential:
+%TODO: check \ell m here
 \begin{gather*}
     \langle (\psi_1)_{\ell_1m_1}(\psi_2)_{\ell_2m_2}(\psi_3)_{\ell_3m_3} \rangle \\ = 
-    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{\ell_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi\sum_{\ellm}i^{\ell}j_{l}(k_i\chi_i)Y_{\ell}^{m}(\hat{\mathbf n_i})Y_{\ell}^{m}(\hat{\mathbf k_i})^*}\\
+    \prod_i\br{-2\int\d^2\hat{\mathbf n_i}(Y^{m_i}_{\ell_i}(\hat{\mathbf n_i}))^*\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{\d^3\mathbf k_i}{(2\pi)^{3}}4\pi\sum_{\ell m}i^{\ell}j_{l}(k_i\chi_i)Y_{\ell}^{m}(\hat{\mathbf n_i})Y_{\ell}^{m}(\hat{\mathbf k_i})^*}\\
     \times (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}) \\
@@
     \langle (\psi_1)_{\ell_1m_1}(\psi_2)_{\ell_2m_2}(\psi_3)_{\ell_3m_3} \rangle 
-    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3}}4\pi (-1)^{\elli}j_{\ell_i}(k_i\chi_i)j_{\ell_i}(k_ix)} \\ \times B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
+    = (2\pi)^3 8 A_{\mathbf l}^{\mathbf m}\int x^2\d x\prod_i\br{-2\int_0^{\chi_*}\d\chi_i W_i(\chi_i)\int\frac{k_i^2\d k_i}{(2\pi)^{3}}4\pi (-1)^{\ell_i}j_{\ell_i}(k_i\chi_i)j_{\ell_i}(k_ix)} \\ \times B^\Psi(\{k_i\}, \{\eta_0-\chi_i\}).
 \end{gather*}
@@
     \centering
-    \begin{tabular}{lcc}
-    \toprule
+    \begin{tabular}{|l|c|c|}
+    \hline
     Parameter & Fiducial value & Finite difference ($h$) \\
-    \midrule
+    \hline
     $H$            & $67.4$               & fiducial\,$\times0.1$   \\
-    $\Omega_b h^2$ & $0.0223$             & fiducial\,$\times0.2$   \\
+    $\Omega_b h^2$ & $0.0223$             & fiducial\,$\times0.1$   \\
     $\Omega_c h^2$ & $0.119$              & fiducial\,$\times0.005$ \\
-    $n_s$          & $0.965$              & fiducial\,$\times0.02$ \\
-    $A_s$          & $2.13\times10^{-9}$  & fiducial\,$\times0.15$   \\
-    $m_\nu$        & $0.06$               & fiducial\,$\times0.15$   \\
-    $w_0$          & $-1$                 & $0.06$                  \\
-    \bottomrule
+    $n_s$          & $0.965$              & fiducial\,$\times0.005$ \\
+    $A_s$          & $2.13\times10^{-9}$  & fiducial\,$\times0.1$   \\
+    $\tau$         & $0.063$              & fiducial\,$\times 0.1$   \\
+    $m_\nu$        & $0.06$               & fiducial\,$\times0.1$   \\
+    $w_0$          & $-1$                 & $0.03$                  \\
+    \hline
     \end{tabular}
@@
 
-To test the accuracy, we vary $h$ by $\pm 5\%$ and $\pm 10\%$ and plot the relative change in the derivative of the equilateral bispectrum in figure \ref{fig:deraccuracy}.
+To test the accuracy, we varied $h$ by $\pm 5\%$ and $\pm 10\%$ and compared the relative change in the derivative of the equilateral lensing bispectra and the lensing powerspectra.
 As long as numerical noise doesn't dominate, it can be shown that the relative error in our approximation is approximately 5 times the relative difference that taking $h \rightarrow h(1 \pm 0.1)$ leads to.
-We thus conclude that, based on this test, the derivatives are almost always computed with up to one percent error.
-The exception is the derivative with respect to neutrino masses, which introduces a larger error.
-
-\begin{figure}[t]
-    \centering
-    \includegraphics[height=\textheight]{../code/plots/derivative_accuracy.png}
-    \caption{Central difference approximation of derivative of equilateral CMB weak lensing bispectrum (left) and power spectrum (right) for differences as specified in table \ref{tab:cosmo-params-diff} as well as for differences $5\%$ and $10\%$ larger and smaller.}
-    \label{fig:deraccuracy}
-\end{figure}
+We thus conclude that, based on the tests conducted, the derivatives are almost always computed with up to one percent error.
+The exception is the derivative with respect to neutrino masses, which introduces a larger error of around 10 percent.
+
+% \begin{figure}[t]
+%     \centering
+%     \includegraphics[height=\textheight]{../code/plots/derivative_accuracy.png}
+%     \caption{Central difference approximation of derivative of equilateral CMB weak lensing bispectrum (left) and power spectrum (right) for differences as specified in table \ref{tab:cosmo-params-diff} as well as for differences $5\%$ and $10\%$ larger and smaller.}
+%     \label{fig:deraccuracy}
+% \end{figure}
 
@@
         \hline
-        $H_0$          &    1.16e+00 &  1.00e+00 &  1.05e+00 &        1.00e+00 & 1.02e+00 & 9.47e-01 &      7.09e-01 &      9.32e-01 &      9.03e-01 &             5.17e-01 \\
-        $\Omega_b h^2$ &    5.62e-05 &  5.16e-05 &  5.58e-05 &        5.15e-05 & 5.38e-05 & 5.26e-05 &      5.10e-05 &      4.73e-05 &      5.08e-05 &             4.69e-05 \\
-        $\Omega_c h^2$ &    7.13e-04 &  6.03e-04 &  7.10e-04 &        5.94e-04 & 4.76e-04 & 5.70e-04 &      3.87e-04 &      1.63e-04 &      4.41e-04 &             1.54e-04 \\
-        $n_s$          &    2.29e-03 &  2.13e-03 &  2.25e-03 &        2.12e-03 & 2.00e-03 & 2.22e-03 &      1.96e-03 &      1.43e-03 &      2.01e-03 &             1.38e-03 \\
-        $\tau$         &    2.44e-03 &  4.30e-05 &  2.44e-03 &        4.27e-05 & 2.41e-03 & 2.42e-03 &      2.40e-03 &      2.92e-05 &      2.42e-03 &             2.80e-05 \\
-        $A_s$          &    1.05e-11 &  4.76e-12 &  1.04e-11 &        4.71e-12 & 1.02e-11 & 1.03e-11 &      9.95e-12 &      2.81e-12 &      1.01e-11 &             2.65e-12 \\
+        $H_0$          &    1.21e+00 &  1.00e+00 &  1.18e+00 &        1.00e+00 & 1.07e+00 & 9.93e-01 &      7.20e-01 &      9.32e-01 &      9.68e-01 &             5.17e-01 \\
+        $\Omega_b h^2$ &    5.69e-05 &  5.16e-05 &  5.67e-05 &        5.15e-05 & 5.49e-05 & 5.29e-05 &      5.18e-05 &      4.74e-05 &      5.15e-05 &             4.69e-05 \\
+        $\Omega_c h^2$ &    8.27e-04 &  6.08e-04 &  8.15e-04 &        5.99e-04 & 5.89e-04 & 6.57e-04 &      5.27e-04 &      1.63e-04 &      6.03e-04 &             1.54e-04 \\
+        $n_s$          &    2.50e-03 &  2.17e-03 &  2.50e-03 &        2.15e-03 & 2.12e-03 & 2.42e-03 &      2.11e-03 &      1.44e-03 &      2.33e-03 &             1.38e-03 \\
+        $\tau$         &    1.34e-02 &  4.37e-05 &  1.22e-02 &        4.33e-05 & 1.08e-02 & 1.13e-02 &      9.73e-03 &      2.94e-05 &      1.08e-02 &             2.82e-05 \\
+        $A_s$          &    5.45e-11 &  4.82e-12 &  4.94e-11 &        4.77e-12 & 4.39e-11 & 4.58e-11 &      3.92e-11 &      2.83e-12 &      4.32e-11 &             2.66e-12 \\
        \hline

### 2025-06-06 code/cosm_setup.py

@@ -45,7 +45,7 @@ class lensing_spectra:
     self.mnu = mnu        # Sum of neutrino masses
     self.w0 = w0
     self.matter_bispec_fit_params = matter_bispec_fit_params # params of matter bispectrum fitting function as (a_1, ..., a_9)
-    self.k_nl_max = 505
+    self.k_nl_max = 3000
 
     self.obtain_results()
 
@@ -84,7 +84,7 @@ class lensing_spectra:
     print('Obtaining cosmological results.')
 
     pars = camb.CAMBparams()
-    lmax = 505
+    lmax = 3000
     pars.set_cosmology(H0=self.H0, ombh2=self.ombh2, omch2=self.omch2, mnu=self.mnu, omk=0, tau=self.tau, neutrino_hierarchy='normal')
     pars.InitPower.set_params(As=self.As, ns=self.ns, r=0)
     pars.InitPower.pivot_scalar = 0.05
@@ -93,7 +93,7 @@ class lensing_spectra:
     pars.set_dark_energy(w=self.w0)
     self.pars = pars
 
-    kmax_interp = 505
+    kmax_interp = 3000
 
     # power spectrum in units of Mpc^3, k in units of Mpc^-1, with input (z, k) on RHS 
     print('Creating linear mps')

### 2025-06-06 code/lps_vs_noise_plot.py

@@ -6,15 +6,15 @@ fig, axs = plt.subplots(1, 2, figsize = (10, 5))
 
 # LENSING SPECTRA
 l_values = np.logspace(np.log10(2), np.log10(3000), 100)
-spectra = cs.lensing_spectra() # don't need matter bispectrum
+spectra = cs.lensing_spectra()
 gal_lps = np.array([spectra.lps(l, ('s', 's')) for l in l_values])  # Shear-shear power spectrum
 cmb_lps = np.array([spectra.lps(l, ('c', 'c')) for l in l_values])
 
 # CMB NOISE QUAD ESTIMATOR
 # original noise is of deflection signal, so divide by l(l+1) to convert to lens potential
 ls_cmbn = np.loadtxt('cmb_noise_files/ls_1_3000_64.txt')
-cmbn_301 = np.abs(np.loadtxt('cmb_noise_files/Ns_sigma3_DeltaT0_DeltaP1.txt')) / (ls_cmbn * (ls_cmbn + 1)) # S4
-cmbn_106 = np.abs(np.loadtxt('cmb_noise_files/Ns_sigma1_DeltaT0_DeltaP6.txt')) / (ls_cmbn * (ls_cmbn + 1)) # S3
+cmbn_301 = np.abs(np.loadtxt('cmb_noise_files/Ns_sigma3_DeltaT0.71_DeltaP1.txt')) / (ls_cmbn * (ls_cmbn + 1)) # S4
+cmbn_106 = np.abs(np.loadtxt('cmb_noise_files/Ns_sigma1_DeltaT4.2_DeltaP6.txt')) / (ls_cmbn * (ls_cmbn + 1)) # S3
 
 # to cross check with paper
 cmbn_41sqrt2 = np.abs(np.loadtxt('cmb_noise_files/Ns_sigma4_DeltaT1_DeltaP1.41.txt')) / (ls_cmbn * (ls_cmbn + 1)) 
@@ -29,16 +29,16 @@ cmbn_test_vals = [cmbn_TT, cmbn_TE, cmbn_TB, cmbn_EE, cmbn_EB]
 names = ['TT', 'TE', 'TB', 'EE', 'EB']
  
 # SIMONS OBSERVATORY NOISE LEVELS
-SO_noise_baseline_file_path = '/scratch/p319950/data/conv_noise.dat'
-SO_noise_goal_file_path = '/scratch/p319950/data/conv_noise_goal.dat'
[…]

### 2025-06-06 paper/figures/snr_plots.py

@@ -5,20 +5,22 @@ from scipy.interpolate import interp1d
 # Define data
 ell_max = [250, 500, 750, 1000, 1250, 1500, 1750, 2000]
 
-snrb_c_s3 = [15.07615989323126, 60.14746748173047, 103.12636014205, 129.35899320252307, 142.68006253688625, 149.21360488222513, 152.46213942042544, 154.16826858416584]
+f_sky = 0.5
 
-snrb_s_s3 = [531.1622436479552, 1725.2752469768607, 2949.2997415134773, 4030.1790464195456, 4935.08531114439, 5676.463829544848, 6279.344472866087, 6769.365663353458]
+snrb_c_s3 = f_sky * np.array([18.14225774292767, 104.80963360462812, 246.33294403758597, 406.55575384332985, 565.893477899891, 723.8984338793172, 883.4983606420614, 1050.0227116349033])
 
-snrb_c_s4 = [23.083226199124194, 160.95563855174456, 453.86854903039114, 846.2154979047796, 1236.5912661335751, 1556.0849450590474, 1785.694771513511, 1939.702697620804]
+snrb_s_s3 = f_sky * np.array([463.85614017322854, 1522.8225121434366, 2593.8277015312924, 3530.456621967403, 4305.844778597493, 4941.576011585971, 5457.833730995741, 5876.8344414029])
 
-snrb_s_s4 = [1841.1019217665996, 13139.433138837652, 35830.00873208229, 66306.94638041034, 100272.43117960829, 134537.47452003675, 167174.8737905056, 197196.08660044952]
+snrb_c_s4 = f_sky * np.array([21.634443589995616, 166.04498908145288, 510.24879700176723, 1052.5932963778246, 1725.518886984883, 2466.6245011624437, 3231.621124084956, 4010.7797969493395])
+
+snrb_s_s4 = f_sky * np.array([1613.0939430053656, 11693.189598379018, 31673.989578902532, 58205.021807229314, 87412.42982093079, 116807.14803651144, 144721.82601691395, 170335.10666953446])
 
 ls = np.linspace(2, 2000, 201)
 plot_datas = np.zeros((4, len(ls)))
 datas = (snrb_c_s3, snrb_s_s3, snrb_c_s4, snrb_s_s4)
 for i in range(4):
     data = datas[i]
-    interp_func = interp1d([0] + ell_max, np.sqrt([0] + data), kind='cubic')
[…]

### 2025-06-06 paper/playing_with_fish.py

@@ -46,7 +46,11 @@ planck_prior = np.diag([12 / 60**2, 1/(0.0005)**2, 12 / 1**2, 1/(0.02)**2, 0.01,
 # CMB Stage 4
 ######################
 
-viscmb_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_both_cmb.txt')
+viscmb_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_both_cmb_lmin30.txt')
+
+viscmb_f_lmin2 = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_both_cmb.txt')
+
+# these still have lmin = 2
 
 viscmb_t = np.loadtxt(fisher_matrices_dir + '/fish_mat_powersp_t_cmb.txt')
 
@@ -175,7 +179,7 @@ sigmaomm_ders = [
 
 keep_indices = [4, 7, 8, 10]
 #keep_indices=[10]
-#keep_indices = [0, 1, 2, 3, 4, 5, 6, 7]
+#keep_indices = [0, 1, 2, 3, 5, 6]
 #keep_indices = [4, 6]
 
 def process_fishes(fishes, keep_indices, derived_param_derivss = [s8_ders, S8_ders, omm_ders, sigmaomm_ders]):
@@ -470,7 +474,7 @@ if __name__ == "__main__":
     # ]
 
     # labels = ['CMB + Gal Powersp', 'CMB + Gal Bisp', 'CMB Power- + Bisp', 'Gal Power- + Bisp', 'CMB + Gal Power- + Bisp']
-    fish_pond_number = 3
+    fish_pond_number = 4
     f_sky = 0.5
     if fish_pond_number == 0:
         fish_matrices = [
@@ -487,6 +491,14 @@ if __name__ == "__main__":
         ]
 
         labels = ['cmb prior', 'lps cmb', 'lbs cmb', 'lps + lbs cmb', 'lps s', 'lbs s', 'lps + lbs s', 'lps cmb + s', 'lbs cmb + s', 'lps + lbs, s + cmb']
+    
+    if fish_pond_number == 7:
+        fish_matrices = [
+            f_sky * viscmb_f,
+            f_sky * viscmb_f_lmin30
+        ]
+
+        labels = ['cmb prior lmin 2', 'cmb prior lmin 30']
     
     if fish_pond_number == 6:
         fish_matrices = [
@@ -531,7 +543,7 @@ if __name__ == "__main__":
 
[…]

### 2025-06-06 paper/paper_draft.tex
Updates from Overleaf

--- 2025-06-04 paper/paper_draft.tex
@@
 $\Omega_m$ is the present day matter density parameter.
-$H_0$ is the present-day Hubble constant.
-$a(\eta)$ is the scale factor.
-$\eta_0$ is the conformal time today.
-$\chi$ is the comoving radial distance.
-$\chi_*$ is the distance to surface of last scattering.
+$H_0$ is the present-day Hubble constant, $a(\eta)$ is the scale factor, $\eta_0$ is the conformal time today and $\chi$ is the comoving radial distance.
+$\chi_*$ presents the distance to surface of last scattering.
 $P^\delta(k, \eta)$ is the matter power spectrum.

### 2025-06-06 code/cosm_setup.py
paper changes

@@ -484,4 +484,4 @@ class lensing_spectra:
 if __name__ == '__main__':
   # for testing purposes
   results = lensing_spectra()
-  print(results.window_func(200, 's'))
+  print(results.window_func(200, 's

### 2025-06-06 paper/figures/snr_plots.py

@@ -68,5 +68,4 @@ axs[1].legend(fontsize=10, loc='upper left')
 plt.tight_layout()
 
 # Save the figure
-plt.savefig('/Users/jonasfrugte/Desktop/fisher_calc_weak_lensing/paper/figures/snrplots.pdf', dpi=300)
-# plt.show()
+plt.savefig('/Users/jonasfrugte/Desktop/fisher_calc_weak_lensing/paper/figures/snrplots.

### 2025-06-18 paper/paper_draft.tex
added acknowledgment
added abstract

@@
 \emailAdd{p.d.meerburg@rug.nl}
-\abstract{}
+\abstract{
+Upcoming Stage 4 surveys from the Simons Observatory, LSST, and Euclid are poised to measure weak gravitational lensing of the Cosmic Microwave Background (CMB) and galaxies with unprecedented precision.
+While the power spectrum is the standard statistic used to analyze weak lensing data, non-Gaussianity induced by the non-linear growth of structure imparts significant cosmological information into higher-order statistics.
+
+In this paper, we perform a Fisher matrix analysis to forecast the ability of future surveys to constrain cosmological parameters using the weak lensing power spectrum and bispectrum from both CMB and galaxy probes, as well as their cross-correlations.
+We analyze an eight-parameter cosmological model ($\Lambda$CDM + $\sum m_{\nu}$ + $w_{0}$) and assess parameter constraints for Stage 4 survey specifications.
+
+Our analysis shows that, in the absence of systematics, both the CMB and galaxy lensing bispectra will be detectable at high signal-to-noise.
+We consider parameter constraints with a ``weak'' prior and with a stronger prior from the primary CMB anisotropies.
+In the weak prior case the lensing bispectrum substantially improves parameter constraints over using the lensing power spectrum alone, primarily by breaking key parameter degeneracies.
+For the stronger prior we see less significant gains, particularly for the CMB lensing bispectrum.
+We further find that for stage 4 surveys that are able to measure lensing on small scales where non-linear structure formation becomes important, the constraining power of the lensing bispectrum can become comparable to that of the lensing power spectrum.
+Furthermore, we demonstrate that, both for our weak and strong priors, the synergy between CMB and galaxy lensing is substantial.
+Combining the two probes, particularly with both their lensing power- and bi-spectra, leads to significantly tighter constraints than achievable with either probe individually, with the greatest improvement seen for the sum of neutrino masses.
+}
 
@@
 Addressing these limitations will be crucial for obtaining robust cosmological constraints from upcoming observational data.
+
+\section*{Acknowledgements}
+We would like to thank prof.
+Toshiya Namikawa for providing extra details on his paper on the CMB lensing bispectrum \cite{Namikawa_2016}, which, in particular, allowed us to cross check (intermediate) results.
 

### 2025-07-06 code/Fisher_calc.pyx
loads of new stuff

@@ -13,7 +13,11 @@ from itertools import *
 cdef int dd = 0
 cdef bint pb_correction = False
 
-print(f'Post-Born corrections: {pb_correction}')
+def set_pb_correction(val):
+    global pb_correction
+    pb_correction = val
+    print(f'Post-Born corrections are now: {pb_correction}')
+    pass
 
 
 k_max = get_k_max()
@@ -67,27 +71,20 @@ cdef double Fisher_mat_single_term(int l1, int l2, int l3, char* type, char* par
         return lbs_der(l1, l2, l3, type, type, type, num_samples, pb_correction, par1, dd) * lps_f_obs(l1, type, type)**(-1) \
         * lps_f_obs(l2, type, type)**(-1) * lps_f_obs(l3, type, type)**(-1) * lbs_der(l1, l2, l3, type, type, type, num_samples, pb_correction, par2, dd) / Delta(l1, l2, l3)
 
-def singlefishtest(l1,  l2,  l3,  type,  par1,  par2,  num_samples):
-    print('wanoo')
-    return Fisher_mat_single_term( l1,  l2,  l3, type, par1, par2,  num_samples)
-
 cpdef double Fisher_mat_single(int lmin, int lminbin, int lmax, int triangle_step_size, int num_bispec_samples, char* par1, char* par2, int num_cores, char* type):
+    # print(f'Post-Born corrections: {pb_correction}')
+
     if lmax > k_max:
         raise Exception('lmax should be less than upper bound for interpolation (k_max)')
-
     cdef int[:, :] triangles
     cdef double prop_calc
     triangles, prop_calc = generate_triangles(lmin, lminbin, lmax, triangle_step_size, ordermatters = False)
-
     cdef int num_samples = int(len(triangles))
-
     # Create a local result array for each thread
     #cdef double[:] local_results = np.zeros(num_cores, dtype=np.float64)
     cdef double result = 0
-
     cdef int k1, k2, k3
     cdef int index
-    print('wapoo')
 
     for index in prange(0, num_samples, schedule='static', num_threads=num_cores, nogil=True):
         k1 = triangles[index, 0]
@@ -99,19 +96,21 @@ cpdef double Fisher_mat_single(int lmin, int lminbin, int lmax, int triangle_ste
 
 # CAUTION: function doesn't check for invalid type1, 2 input to optimize performance
[…]

### 2025-07-06 code/Fisher_matrix.py

@@ -24,7 +24,6 @@ def fisher_calc_wrapper(args, tracers):
         return vis.Fisher_mat_full(lmin, 0, 200, stepsizes[0], num_bispec_samples, pars[i], pars[j], num_cores) + vis.Fisher_mat_full(lmin, 200, 1000, stepsizes[1], num_bispec_samples, pars[i], pars[j], num_cores) + vis.Fisher_mat_full(lmin, 1000, lmax, stepsizes[2], num_bispec_samples, pars[i], pars[j], num_cores)
 
 num_pars = len(pars)
-print('koopie')
 def main():
     for tracer in ['c', 's', 'both']:
         mat = np.zeros((num_pars, num_pars))
@@ -41,6 +40,7 @@ def main():
         print(mat)
 
 if __name__ == '__main__':
+    vis.set_pb_correction(False)
     main()
 
 

### 2025-07-06 code/Fisher_matrix_approx.py

@@ -40,7 +40,8 @@ def fisher_calc_wrapper_takada_jain(args, tracers):
         # lmax should actually be 3000, but interpolation currently doesn't go that high
         return vis.Fisher_mat_single(50, 0, 200, stepsizes[0], num_bispec_samples, pars[i], pars[j], num_cores, b's') + vis.Fisher_mat_single(50, 200, 1000, stepsizes[1], num_bispec_samples, pars[i], pars[j], num_cores, b's') + vis.Fisher_mat_single(50, 1000, 2000, stepsizes[2], num_bispec_samples, pars[i], pars[j], num_cores, b's')
 num_pars = len(pars)
-print('snoopie2')
+
+pb_corrections = True
 
 def main():
     for tracer in ['both', 's', 'c']:
@@ -53,9 +54,17 @@ def main():
                 mat[j, i] = result  # Symmetric assignment
                 print(f'{tracer}:', counter, '/', num_pars * (num_pars + 1) / 2)
                 counter += 1
+
+        if pb_corrections:
+            filepath = f'fisher_matrices/fish_mat_bisp_{tracer}_pb.txt'
+        else:
+            filepath = f'fisher_matrices/fish_mat_bisp_{tracer}.txt'
         
-        np.savetxt(f'fisher_matrices/fish_mat_bisp_{tracer}_pb_exp_fix.txt', mat)
+        np.savetxt(filepath, mat)
+        print(f'Created Fisher matrix at {filepath}')
         print(mat)
+    pass
 
 if __name__ == '__main__':
+    vis.set_pb_correction(pb_corrections)
     main()

### 2025-07-06 code/Fisher_matrix_powerspectra.py

@@ -3,7 +3,7 @@ import Fisher_calc_python_imp as vispy
 import numpy as np
 from itertools import *
 
-lmin = 30
+lmin = 100
 lmax = 2000
 
 pars = [b'H', b'ombh2', b'omch2', b'ns', b'mnu', b'tau', b'As', b'w0'] # add tau back in later!
@@ -34,6 +34,7 @@ def fisher_calc_wrapper_takada_jain(args, tracers):
         return vispy.Fisher_powersp_single(50, 2000, b's', par1 = pars[i], par2 = pars[j])
     
 def fisher_calc_wrapper_cmb(args, tracers):
+    print(lmin)
     i, j = args
     # Fisher_mat_full(int lmin, int lminbin, int lmax, int triangle_step_size, int num_bispec_samples, char* par1, char* par2, int num_cores)
     # Fisher_mat_single(int lmin, int lminbin, int lmax, int triangle_step_size, int num_bispec_samples, char* par1, char* par2, int num_cores, char* type)
@@ -45,16 +46,19 @@ def fisher_calc_wrapper_cmb(args, tracers):
         # lmax should actually be 3000, but interpolation currently doesn't go that high
         return vispy.Fisher_powersp_cmb(lmin, lmax, par1 = pars[i], par2 = pars[j])
 
-def main():
-    for tracer in ['both']: #['c', 's', 'both']:
+def main(lmin_loc):
+    global lmin
+    lmin = lmin_loc
+    for tracer in ['t']: #['c', 's', 'both']:
         mat = np.zeros((len(pars), len(pars)))
         for i, j in product(range(len(pars)), repeat = 2):
             result = fisher_calc_wrapper_cmb((i, j), tracer)
             mat[i, j] = result
             mat[j, i] = result  # Symmetric assignment
 
-        np.savetxt(f'fisher_matrices/fish_mat_powersp_{tracer}_cmb_lmin30.txt', mat)
+        #np.savetxt(f'fisher_matrices/fish_mat_powersp_{tracer}_cmb_lmin30.txt', mat)
         print(np.linalg.inv(mat))
+        return mat
 
     # for tracer in ['c', 's', 'both']: #['c', 's', 'both']:
     #     mat = np.zeros((len(pars), len(pars)))

### 2025-07-06 code/SNR.py

@@ -1,4 +1,4 @@
-from Fisher_calc import Fisher_mat_single, Fisher_mat_full
+from Fisher_calc import Fisher_mat_single, Fisher_mat_full, set_pb_correction
 from Fisher_calc_python_imp import Fisher_powersp_single, Fisher_powersp
 
 import numpy as np
@@ -6,6 +6,7 @@ import numpy as np
 num_cores = 64
 stepsize = 49
 num_samples = 50
+print(f'Stepsize: {stepsize}')
 
 # ls = [0] + ls
 ls = [0, 250, 500, 750, 1000, 1250, 1500, 1750, 2000]
@@ -67,6 +68,8 @@ def all_bs_SNR2():
 
     pass
 
-#ps_SNR2()
+set_pb_correction(False)
 single_bs_SNR2()
-# all_bs_SNR2()
+
+set_pb_correction(True)
+single_bs_SNR2()

### 2025-07-06 code/cmbps_vs_noise_plot.py

@@ -1,0 +1,20 @@
+import data_importer_new as di
+import numpy as np
+import matplotlib.pyplot as plt
+
+ls = np.arange(5, 3000)
+
+sigma = 3
+Delta_T = 1
+
+cmbps_TT = np.array([di.lps_der_test(l, b't', b't', b'snr') for l in ls])
+cmbps_TT_noise_namikawa = np.array([di.cmbps_noise_test(l, sigma, Delta_T) for l in ls])
+cmbps_TT_noise_hu = np.array([di.cmbps_noise_test(l, 4, 1) for l in ls])
+
+
+plt.loglog(ls, ls**2 * cmbps_TT, label = 'C^TT')
+plt.loglog(ls, ls**2 * cmbps_TT_noise_namikawa, label = f'noise, sigma = {sigma}, Delta_T = {Delta_T}')
+plt.loglog(ls, ls**2 * cmbps_TT_noise_hu, label = f'noise, sigma = {4}, Delta_T = {1}')
+
+plt.legend()
+plt.savefig('cmbps_vs_noise.png')

### 2025-07-06 code/cosm_setup.py

@@ -66,7 +66,8 @@ class lensing_spectra:
     beta = 1.5
     normalization = scipy.integrate.quad(lambda z : f(z) * np.exp(float(-(z / z0)**beta)), 1e-8, self.results.get_derived_params()['zstar'])[0]
 
-    self.galaxy_density = lambda z : f(z) * np.exp(-(max(1e-12, z / z0))**beta) / normalization # normalizes integral of func to 1
+    self.galaxy_density_z = lambda z : f(z) * np.exp(-(max(1e-12, z / z0))**beta) / normalization # normalizes integral of func to 1
+    self.galaxy_density_chi = lambda chi : self.galaxy_density_z(self.results.redshift_at_comoving_radial_distance(chi)) * np.abs((self.results.redshift_at_comoving_radial_distance(chi + 0.1) - self.results.redshift_at_comoving_radial_distance(chi)) / 0.1)
 
     # source redshift and comoving dist of galaxies using Luna's convention
     self.z_galaxy_source = 5
@@ -129,30 +130,6 @@ class lensing_spectra:
     
     pass
 
-  # def lps_cc_CAMB(self, lmax=10000):
-  #   """
-  #   Calculate the lensing potential power spectrum (C_phi_phi) using CAMB for the CMB (convergence) case.
-
-  #   Parameters:
-  #   -----------
-  #   lmax : int
-  #       Maximum multipole for the calculation.
-
-  #   Returns:
-  #   --------
-  #   ell : ndarray
-  #       Array of multipole moments.
-  #   C_phi_phi : ndarray
-  #       Array of lensing potential power spectrum values.
-  #   """
-  #   # Get the lensing potential power spectrum
-  #   # if raw_cl = false then returns L^2(L+1)^2 / (2pi) times the result
-  #   lens_potential_cls = self.results.get_lens_potential_cls(lmax=lmax, raw_cl=True)
-  #   ell = np.arange(lmax + 1)
-  #   C_phi_phi = lens_potential_cls[:, 0]
-
-  #   return ell, C_phi_phi
-  
   def ns_eff(self, k, z):
     '''
     Parameters:
@@ -476,7 +453,7 @@ class lensing_spectra:
       return (self.chi_last_scatter - chi)/(self.chi_last_scatter * chi)
     elif type[0] == 's':
       return scipy.integrate.quad(
[…]

### 2025-07-06 code/data_exporter_new.py

@@ -13,7 +13,7 @@ import multiprocessing
 import interp_settings
 exp_par = interp_settings.exp_par
 
-def data_export(folder_name, cosm_par, lps = True, a_create = True, b_create = True, c_create = True, mps = True, rest = True, window_func_pb = True):
+def data_export(folder_name, cosm_par, lps = True, a_create = True, b_create = True, c_create = True, mps = True, rest = True):
     cosm_data = cosm_setup.lensing_spectra(*cosm_par)
 
     filepath = '/scratch/p319950/data_toshiya_like/' + folder_name
@@ -41,11 +41,7 @@ def data_export(folder_name, cosm_par, lps = True, a_create = True, b_create = T
             'mnu': cosm_par[6],
             'w0' : cosm_par[7]
             }
-
-    if window_func_pb:
-        np.save(filepath + '/window_func_pb', [[cosm_data.window_func_pbs(chi, chi_s) for chi_s in chis_log_pb] for chi in chis_log_pb])
-        print(f'window_func_pb created to {folder_name}')
-
+    
     if rest:
         np.save(filepath + '/cosm_par', cosm_par)
         print(cosm_par)
@@ -170,7 +166,7 @@ delta_delta = 0.05
 # data_export(folder_name, cosm_par, lps = True, a_create = True, b_create = True, c_create = True, mps = True, rest = True, exp_par = exp_par)
 
 # allows control over which data to create
-which_to_create = [[False, False, False, False, False, False, True]]
+which_to_create = [[True, False, False, False, False, True]]
 
 exports = [['data_fiducial', fiducial_cosm_par, *create_settings] for create_settings in which_to_create]
 

### 2025-07-06 code/data_importer_new.pyx

@@ -1,4 +1,3 @@
-print('wahoo!')
 import json
 cimport interpolation as interp
 import math
@@ -23,8 +22,10 @@ cdef double z_max = exp_par['z_max']
 cdef int z_num = exp_par['z_num'] # * 2 # for finer data option
 cdef int z_num_fine = z_num * 25
 
+
 def get_k_max():
     return k_max
+
 
 # specifies the noise that will be used for cmb lensing
 cdef int cmb_noise_type = 2
@@ -46,7 +47,6 @@ cdef str folder_file_path = '/scratch/p319950/data/'
 # cdef str filepath_convergence_noise_file_path = '/scratch/p319950/data_rough/' + 'conv_noise.dat'
 # conv_noise_data_array = np.loadtxt(filepath_convergence_noise_file_path)
 # cdef double[:, :] conv_noise_data = conv_noise_data_array
-print('wahoo2')
 # conv noise from quadratic estimator
 cdef double[:] ls_cmbn = np.loadtxt('cmb_noise_files/ls_1_3000_64.txt')
 ls_cmbn_np_array = np.loadtxt('cmb_noise_files/ls_1_3000_64.txt') # need to also have a np array version to easily convert noise values from convergence to lens potential below
@@ -65,7 +65,6 @@ toshiya_derivatives = True
 def data_import_func(folder_name):
     cdef str filepath
     if toshiya_derivatives:# and (folder_name[-1] == '0' or folder_name == 'data_fiducial'):
-        print(folder_name, 'toshiya')
         filepath = '/scratch/p319950/data_toshiya_like/' + folder_name
 
     else:
@@ -174,18 +173,23 @@ cpdef double mbs(double k1, double k2, double k3, double z, double[:, :] mps_dat
 
 # post born corrections, based on https://arxiv.org/pdf/1605.05662
 
-# from cosm_setup:
-# # galaxy density dist as a function of redshift
+# from cosm_setup.py:
+
 # f = lambda z : float(z)**2
 # z0 = 0.64
 # beta = 1.5
 # normalization = scipy.integrate.quad(lambda z : f(z) * np.exp(float(-(z / z0)**beta)), 1e-8, self.results.get_derived_params()['zstar'])[0]
 
+# self.galaxy_density_z = lambda z : f(z) * np.exp(-(max(1e-12, z / z0))**beta) / normalization # normalizes integral of func to 1
[…]

### 2025-07-06 code/lps_vs_noise_plot.py

@@ -30,16 +30,16 @@ names = ['TT', 'TE', 'TB', 'EE', 'EB']
  
 # SIMONS OBSERVATORY NOISE LEVELS
 # SO_noise_baseline_file_path = '/scratch/p319950/data/conv_noise.dat'
-SO_noise_goal_file_path = '/Users/jonasfrugte/Desktop/fisher_calc_weak_lensing/code/cmb_noise_files/nlkk_v3_1_0_deproj0_SENS2_fsky0p4_it_lT30-3000_lP30-5000.dat'
+#SO_noise_goal_file_path = '/Users/jonasfrugte/Desktop/fisher_calc_weak_lensing/code/cmb_noise_files/nlkk_v3_1_0_deproj0_SENS2_fsky0p4_it_lT30-3000_lP30-5000.dat'
 
 # conv_noise_data_array = np.loadtxt(SO_noise_baseline_file_path)
-conv_noise_goal_data_array = np.loadtxt(SO_noise_goal_file_path)
+#conv_noise_goal_data_array = np.loadtxt(SO_noise_goal_file_path)
 
 # SO_noise = lambda l : conv_noise_data_array[l-2, 7] * 4. # * (l * 1.0)**(-2) * (l + 1.0)**(-2)
 # SO_noise_values = np.array([SO_noise(int(l)) for l in l_values])
 
-SO_noise_goal = lambda l : 0.4 * conv_noise_goal_data_array[l-2, 7] * 4. # * (l * 1.0)**(-2) * (l + 1.0)**(-2)
-SO_noise_goal_values = np.array([SO_noise_goal(int(l)) for l in l_values])
+#SO_noise_goal = lambda l : 0.4 * conv_noise_goal_data_array[l-2, 7] * 4. # * (l * 1.0)**(-2) * (l + 1.0)**(-2)
+#SO_noise_goal_values = np.array([SO_noise_goal(int(l)) for l in l_values])
 
 # GALAXY LENSING NOISE
 sigma_rms_S3 = 0.3
@@ -48,8 +48,8 @@ sigma_rms_S4 = 0.3  # e.g. the Euclid ellipticity dispersion
 # Number density per steradian (converted from per arcmin^2)
 # actually shouldn't be computed to get proper ratio with galaxy lensing spectrum, no idea why
 arcmin2_to_steradian = (np.pi / (180 * 60)) ** 2
-n_g_S3 = 5
[…]

### 2025-07-06 code/tau_plots.py

@@ -1,0 +1,33 @@
+import cosm_setup as cs
+import matplotlib.pyplot as plt
+import numpy as np
+
+tau_vals = [0.063*0.85, 0.063 * 0.9, 0.063 * 0.95, 0.063, 0.063 * 1.05, 0.063 * 1.1, 0.063 * 1.15]
+As_fid = 2.13e-9
+As_vals = [As_fid*0.85, As_fid*0.9, As_fid*0.95, As_fid, As_fid * 1.05, As_fid * 1.1, As_fid * 1.15]
+ls = np.arange(2000+1)
+
+fig, axs = plt.subplots()
+
+spectra_fid = cs.lensing_spectra()
+cls_fid = spectra_fid.results.get_lensed_scalar_cls(lmax=2000, raw_cl=True)
+
+def tau_plot():
+    for tau in tau_vals:
+        spectra = cs.lensing_spectra(tau = tau)
+        cls = spectra.results.get_lensed_scalar_cls(lmax=2000, raw_cl=True)
+        axs.semilogx(ls[2:], cls[2:,0] / cls_fid[2:,0], label = f'TT, tau = {tau}')
+
+    fig.legend()
+    fig.savefig('taus_perturbed.png')
+    pass
+
+def As_plot():
+    for As in As_vals:
+        spectra = cs.lensing_spectra(As = As)
+        cls = spectra.results.get_lensed_scalar_cls(lmax=2000, raw_cl=True)
+        axs.semilogx(ls[2:], cls[2:,0] / cls_fid[2:,0], label = f'TT, As = {As}')
+
+    fig.legend()
+    fig.savefig('As_perturbed.png')
+    pass

### 2025-07-06 paper/figures/snr_plots.py

@@ -7,18 +7,26 @@ ell_max = [250, 500, 750, 1000, 1250, 1500, 1750, 2000]
 
 f_sky = 0.5
 
-snrb_c_s3 = f_sky * np.array([18.14225774292767, 104.80963360462812, 246.33294403758597, 406.55575384332985, 565.893477899891, 723.8984338793172, 883.4983606420614, 1050.0227116349033])
+snrb_c_s3 = f_sky * np.array([12.672454984617664, 83.7377129725408, 206.95352488728747, 350.9902759829114, 498.35684081032775, 645.7108221860103, 796.0813174568198, 953.7584589205301])
 
-snrb_s_s3 = f_sky * np.array([463.85614017322854, 1522.8225121434366, 2593.8277015312924, 3530.456621967403, 4305.844778597493, 4941.576011585971, 5457.833730995741, 5876.8344414029])
+snrb_s_s3 = f_sky * np.array([477.5259136914146, 2533.295459864549, 5493.551176152028, 8644.382159891089, 11636.918745628755, 14316.53223335389, 16656.336652636706, 18662.00612425752])
 
-snrb_c_s4 = f_sky * np.array([21.634443589995616, 166.04498908145288, 510.24879700176723, 1052.5932963778246, 1725.518886984883, 2466.6245011624437, 3231.621124084956, 4010.7797969493395])
+snrb_c_s3_pb = f_sky * np.array([7.825690349725596, 46.151237529667654, 107.18588258187486, 177.07130085865384, 247.7876624988021, 319.56645397880607, 394.29703253136023, 474.84034728265567])
 
-snrb_s_s4 = f_sky * np.array([1613.0939430053656, 11693.189598379018, 31673.989578902532, 58205.021807229314, 87412.42982093079, 116807.14803651144, 144721.82601691395, 170335.10666953446])
+snrb_s_s3_pb = f_sky * np.array([464.03174572011704, 2474.5313315586527, 5380.676014059787, 8479.734824268631, 11425.708268033992, 14064.629823584384, 16368.973212497156, 18344.038048655635])
+
[…]

### 2025-07-06 paper/paper_draft.tex

@@
 \abstract{
-Upcoming Stage 4 surveys from the Simons Observatory, LSST, and Euclid are poised to measure weak gravitational lensing of the Cosmic Microwave Background (CMB) and galaxies with unprecedented precision.
+Upcoming Stage 4 surveys, such as the Simons Observatory, LSST, and Euclid are poised to measure weak gravitational lensing of the Cosmic Microwave Background (CMB) and galaxies with unprecedented precision.
 While the power spectrum is the standard statistic used to analyze weak lensing data, non-Gaussianity induced by the non-linear growth of structure imparts significant cosmological information into higher-order statistics.
-
-In this paper, we perform a Fisher matrix analysis to forecast the ability of future surveys to constrain cosmological parameters using the weak lensing power spectrum and bispectrum from both CMB and galaxy probes, as well as their cross-correlations.
-We analyze an eight-parameter cosmological model ($\Lambda$CDM + $\sum m_{\nu}$ + $w_{0}$) and assess parameter constraints for Stage 4 survey specifications.
-
+In this paper, we forecast the ability of future surveys to constrain cosmological parameters using the weak lensing power spectrum and bispectrum from both CMB and galaxy surveys, as well as their cross-correlations.
+We consider an eight-parameter cosmological model ($\Lambda$CDM + $\sum m_{\nu}$ + $w_{0}$) and assess parameter constraints for Stage 4 survey specifications.
 Our analysis shows that, in the absence of systematics, both the CMB and galaxy lensing bispectra will be detectable at high signal-to-noise.
 We consider parameter constraints with a ``weak'' prior and with a stronger prior from the primary CMB anisotropies.
-In the weak prior case the lensing bispectrum substantially improves parameter constraints over using the lensing power spectrum alone, primarily by breaking key parameter degeneracies.
+In the weak prior case the lensing bispectrum substantially improves parameter constraints over using the lensing power spectrum alone, primarily by breaking parameter degeneracies.
 For the stronger prior we see less significant gains, particularly for the CMB lensing bispectrum.
@@
 This can be measured in terms of the lensing convergence $\kappa$, lensing shear $\gamma$, or lensing potential $\psi$.
-In the weak lensing regime they are all equivalent, in the sense that they can be directly converted into one another (appendix \ref{sec:weaklensstats}).
+In the weak lensing regime and under the Born approximation they are all equivalent, in the sense that they can be directly converted into one another (appendix \ref{sec:weaklensstats}).
 In CMB surveys, the temperature and polarization anisotropies are used to reconstruct the lensing potential \cite{Hu_2002}, while with galaxies one instead calculates lensing effects by measuring the lensing shear, which can be deduced from the ellipticity distribution of the observed galaxies \cite{HoekstraJain2008}.
@@
 
-In this paper, we aim to answer this question by adopting experimental parameters similar to those of next-generation surveys.
+Additionally, we will also look at the effects of post-Born corrections.
+The CMB and galaxy lensing power spectra are typically affected by less than $1\%$ \cite{postborn_pratten_lewis, galpostbornlpscorr}.
+The galaxy lensing bispectrum is only affected by a few percent at most \cite{galpostbornlbscorr}.
+The only significant effect is on the CMB lensing bispectrum, as was shown in a 2024 paper \cite{postborn_pratten_lewis}.
+In particular, for the folded configuration the effect can be of the same magnitude as the lensing bispectrum due to nonlinear large scale structure \eqref{eq:lensingbispectrum}.
+It is thus worth considering how including first order corrections changes the detectability of the CMB lensing bispectrum and the associated parameter constraints.
+This analysis will shed light on whether or not the the Born approximation is still appropriate to use if one were to calculate and use the CMB lensing bispectrum using future data.
+
+In this paper, we aim to answer these questions by adopting experimental parameters similar to those of next-generation surveys.
 We are especially interested in seeing if approximate parameter degeneracies can be broken by combining CMB and galaxy weak lensing power- and bi-spectra.
@@
 
-All code used to derive the results in this paper is publically available on GitHub at \url{https://github.com/Jonas-Frugte/fisher_calc_weak_lensing}.
+All code used to derive the results in this paper is publicly available on GitHub at \url{https://github.com/Jonas-Frugte/fisher_calc_weak_lensing}.
 
@@
 Radiation from cosmological objects is distorted due to gravitational lensing.
-Due to the low density of the cosmic web, the average deflection of a photon as it propagates through the universe is relatively weak and is thus referred to as weak lensing.
+Due to the low density of the cosmic web, the average deflection of a photon as it propagates through the universe is relatively weak.
+We thus work under the assumption that all deflection angles are small, this is referred to as the weak lensing regime.
 Weak lensing is quantified through the deflection field $\mathbf d(\hat {\mathbf n})$ which equals the difference between the observed angle of a point in the sky and the true (unlensed) angle.
@@
 
-In the case of CMB surveys, lensing alters the statistical properties of the temperature and polarization fields and can thus be calculated by comparing the observed signal to the expected unlensed signal(see e.g.
-\cite{cmblensingestimator}).
+In the case of CMB surveys, lensing alters the statistical properties of the temperature and polarization fields and can thus be calculated by comparing the observed signal to the expected unlensed signal (see e.g.
+\cite{cmblensingestimator, Maniyar2021_QE_CMBLensing, Hirata_2003}).
 In galaxy surveys, lensing alters the ellipticities of observed galaxies.
@@
     C^{\psi_X\psi_Y}_\ell
-    & = \frac{9}{\ell^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)P^\delta(\ell/\chi,\eta_0-\chi),\\
-    B^{\psi_X\psi_Y\psi_Z}_{\ell\ell_1\ell\ell_2\ell\ell_3} &= \sqrt{\frac{(2\ell_1 + 1)(2\ell_2 + 1)(2\ell_3 + 1)}{4\pi}} \begin{pmatrix} \ell_1 & \ell_2 & \ell_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{\ell_1^2\ell_2^2\ell_3^2}\Omega_m^3H_0^6\\
-    & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{\ell_i/\chi\}, \eta_0-\chi),
+    & = \frac{9}{\ell^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)P^\delta(\ell/\chi,\eta_0-\chi), \label{eq:lensingpowerspectrum}\\
+    B^{\psi_X\psi_Y\psi_Z}_{\ell\ell_1\ell\ell_2\ell\ell_3} &= \sqrt{\frac{(2\ell_1 + 1)(2\ell_2 + 1)(2\ell_3 + 1)}{4\pi}} \begin{pmatrix} \ell_1 & \ell_2 & \ell_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{\ell_1^2\ell_2^2\ell_3^2}\Omega_m^3H_0^6 \notag\\
+    & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{\ell_i/\chi\}, \eta_0-\chi), \label{eq:lensingbispectrum}
 \end{align}
@@
 is the Wigner 3-j symbol. The difference between CMB and galaxy lensing is the window function, defined as
-\begin{equation*}
+\begin{equation}
     W_X(\chi) = \int_\chi^{\infty}\d\chi' p_X(\chi')\frac{\chi'-\chi}{\chi'\chi},
-\end{equation*}
-with $p_X$ the radius distribution of the radiation source.
+    \label{eq:windowfuncdef}
+\end{equation}
+with $p_X$ the radial distribution of the radiation source.
 For the CMB we take $p(\chi) = \delta(\chi - \chi_*)$.
 The window function presents the distribution of redshifts at which a CMB photon or galaxy photon is deflected.
-As a result, the CMB lensing window function is broader and peaks at lower redshift compared to the galaxy window function.
+As a result, the CMB lensing window function is broader and peaks at higher redshift compared to the galaxy window function.
 
@@
 \begin{equation*}
-    \langle \delta(\mathbf k, \eta) \delta(\mathbf{k}',\eta) \rangle = (2\pi)^3\delta(\mathbf k + \mathbf k ')P^\delta(\mathbf k, \eta),
+    \langle \delta(\mathbf k, \eta) \delta(\mathbf{k}',\eta) \rangle = (2\pi)^3\delta_D(\mathbf k + \mathbf k ')P^\delta(\mathbf k, \eta),
 \end{equation*}
@@
 \begin{equation*}
-    \langle \delta(\mathbf k_1, \eta) \delta(\mathbf k_2, \eta)\delta(\mathbf k_3, \eta) \rangle = (2\pi)^3\delta(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\delta(k_1,k_2,k_3, \eta),
-\end{equation*}
-with $\delta$ the fractional matter density perturbation field.
+    \langle \delta(\mathbf k_1, \eta) \delta(\mathbf k_2, \eta)\delta(\mathbf k_3, \eta) \rangle = (2\pi)^3\delta_D(\mathbf k_1 + \mathbf k_2 + \mathbf k_3)B^\delta(k_1,k_2,k_3, \eta),
+\end{equation*}
+with $\delta$ the fractional matter density perturbation field and $\delta_D$ the Dirac delta function.
 
@@
 \end{equation}
-where \( C^\delta \) is the nonlinear matter power spectrum\footnote{Compare to the tree level bispectrum where we instead use the linear power spectrum.}, and the kernel \( F_2 \) is modified from the tree level result with factors \( a(k, z) \), \( b(k, z) \), and \( c(k, z) \):
+where \( P^\delta \) is the nonlinear matter power spectrum\footnote{Compare to the tree level bispectrum where we instead use the linear power spectrum.}, and the kernel \( F_2 \) is modified from the tree level result with factors \( a(k, z) \), \( b(k, z) \), and \( c(k, z) \):
 \begin{equation}
@@
 
+\subsection{Post-Born corrections}
+One of the approximations made in the derivation of the weak lensing power- and bispectra (equations \eqref{eq:lensingpowerspectrum} and \eqref{eq:lensingbispectrum}) is that the integral is performed along the line of sight to the light source.
+To get an analytically correct result, one would instead need to solve a differential equation with a feedback loop where the effects of gravitational lensing affect the trajectory and shape of given bundle of light rays which in turn changes the way future gravitational lenses affect the bundle.
+One can also choose to make order by order corrections to the Born approximation.
+There is motivation to understand the effects of these corrections (see the introduction) so we present them here.
+
+Under the Born approximation, the elements of the distortion tensor, $\psi_{ab}$, (from which shear, convergence, and rotation are constructed) equal second derivatives of the lensing potential.
+We thus calculate the potential instead of the distortion tensor itself.
+When post-born effects are taken into account this is no longer true and we are forced to work with $\psi_{ab}$ directly.
+Following \cite{postborn_pratten_lewis}, the leading order post born correction to the distortion tensor for a point source at distance $\chi_s$ is given by
+\begin{equation}
+    \left( \psi_{ab}^{\text{p.s.}}  \right)^{(2)}(\chi_s) := 2 \int^{\infty}_0 d \chi^{} \, \chi^2 \, \frac{\chi_s - \chi}{\chi_s\chi}\Theta(\chi_s - \chi) \, \left[ - \Psi_{,ac} (\chi) \left(\psi^{\text{p.s.}}_{cb}\right)^{(1)} (\chi) + \Psi_{,acd} (\chi) \, \delta x_d^{(1)} (\chi^{}) \right].
+    \label{eq:pbcorrectionpointsorce}
+\end{equation}
+Here $\left(\psi_{cb}^{(p.s.)}\right)^{(1)}(\chi)$ is the deformation for a point source at $\chi$ under the Born approximation as defined earlier.
+$\delta x_d^{(1)}(\chi)$ is the first order correction to the location of the lensed light ray at a distance $\chi$, given by
+\begin{equation}
+\delta x_a^{(1)} = -2 \int^{\chi}_0 d \chi^{\prime} \, \frac{\chi-\chi'}{\chi\chi'} \, \chi \chi' \, \Psi_{,a} (\chi').
+\end{equation}
+To adapt the above to a source with radial distribution $p(\chi)$ (which in our case will be a population of galaxies) we integrate over $\chi_s$ as follows:
+\begin{equation}
+    (\psi_{ab}^{\text{gal}})^{(i)} = \int_0^\infty d \chi_s p(\chi_s)\left(\psi_{cb}^{(p.s.)}\right)^{(i)}(\chi_s).
+\end{equation}
+This step essentially just changes $(\chi_s - \chi)/(\chi_s\chi)\Theta(\chi_s-\chi)$ in equation \eqref{eq:pbcorrectionpointsorce} to the window function $W(\chi)$ defined in equation \eqref{eq:windowfuncdef}.
+
+The leading order correction to the convergence bispectrum of sources 1, 2, 3 is given as
+\begin{equation}
+    B^{\kappa_{(1)}\kappa_{(2)}\kappa_{(3)}}_{L_1 L_2 L_3} = 2\frac{\mathbf L_1\cdot\mathbf L_2}{L_1^2L_2^2}\left[ \mathbf L_1\cdot \mathbf L_3 \mathcal M_s^{(1)(2)(3)}(L_1,L_2)
+ +\mathbf L_2\cdot \mathbf L_3 \mathcal M_s^{(2)(1)(3)}(L_2,L_1)\right] + \text{cyc. perm.},
+\end{equation}
+where
+\begin{equation}
+    \mathcal M_s^{(1)(2)(3)}(L_1,L_2) := L_1^4\int_0^{\chi_s}d\chi \frac{W^{(1)}(\chi)W^{(3)}(\chi)}{\chi^2} P_{\Psi\Psi}\left(\frac{L_1}{\chi}, z(\chi)\right) C_{L_2}^{\kappa_{(2)}\kappa_{(3)}}(\chi,\chi_s),
+\end{equation}
+and
+\begin{equation}
+    C_{L_2}^{\kappa_{(2)}\kappa_{(3)}}(\chi,\chi_s) := \int_0^{\infty}d \chi' W^{(2)}(\chi')\frac{\chi_s-\chi}{\chi_s\chi}\Theta(\chi_s-\chi)P_{\Psi\Psi}(\frac{L_2}{\chi'},\chi').
+\end{equation}
+
 \subsection{Fisher matrix analysis}
@@
     C_{\ell} := \begin{pmatrix}
-        C^{\psi_\kappa\psi_\kappa}_\ell & C^{\psi_\kappa\psi_\gamma}_\ell \\
-        C^{\psi_\kappa\psi_\gamma}_\ell & C^{\psi_\gamma\psi_\gamma}_\ell
+        C^{\psi_{\text{CMB}}\psi_{\text{CMB}}}_\ell & C^{\psi_{\text{CMB}}\psi_{\text{gal}}}_\ell \\
+        C^{\psi_{\text{CMB}}\psi_{\text{gal}}}_\ell & C^{\psi_{\text{gal}}\psi_{\text{gal}}}_\ell
     \end{pmatrix}.
@@
 
-\section{Calculation Details}\label{sec:calcdetails}
+\section{Experimental parameters and priors}\label{sec:calcdetails}
 % TODO: find proper references for stage 3 and stage 4 names as well as experimental parameters and real life surveys
+\subsection{Fiducial cosmology}
 Our fiducial cosmology is based on the Planck results \cite{planckresults} (see table \ref{tab:fiducialpars}).
+We choose to constrain the standard parameters of the $\Lambda$CDM model as well as neutrino mass, $m_\nu$, and dark energy equation of state parameter, $w_0$.
+Our reason for including these extra degrees of freedom is that we would like to see to what degree future surveys will be able to tell us more about the mass of neutrinos and the nature of dark energy.
+For example, a paper from the DESI collaboration \cite{Roy2024DESI} recently showed that measurements of baryon acoustic oscillations give a 2.5 - 4 $\sigma$ tension with a non-evolving dark energy equation of state model, so it is worth considering if future weak lensing experiments could tell us more about this tension.
+Regarding neutrinos, depending on how stage 4 surveys will be able to constrain neutrino mass, the results could tell us more about their mass hierarchy and serve as a test for constraints obtained from other probes, such as baryon acoustic oscillations.
+
+\subsection{Choice of priors}
 Throughout this paper we consider two priors.
-The first we refer to as our ``weak prior'' and is similar to the one used in \cite{Planck2018Lensing} (see table \ref{tab:fiducialpars}).
-This prior only restricts $\Omega_b h^2$ and $n_s$ significantly.
+The first we refer to as our ``weak prior'' and is based on the one used in \cite{Planck2018Lensing} (see table \ref{tab:fiducialpars}).
+This prior only restricts $\Omega_b h^2$ and $n_s$ significantly, as those are not well constrained by lensing spectra alone.
+Our differences with \cite{Planck2018Lensing} are that (i) we do not fix $\tau$, but instead take it to have a SNR of 1 and added a loose $A_s$ prior as well, (ii) instead of a flat prior for the Hubble constant ($40 < H_0 < 100$) we use a gaussian distribution with the same standard deviation, (iii) we added a weak constraint on $\Omega_c h^2$, and (iv) we also vary neutrino mass and $w_0$ but do not assume any constraints on them in our priors.
 The purpose of the weak prior is as follows.
-The power spectra, in particular, exhibit complete insensitivity to certain parameters.
+The weak lensing power spectra, in particular, exhibit complete insensitivity to certain parameters.
 Consequently, relying solely on them to constrain the entire cosmological model results in very poor constraints.
-Future lensing surveys will of course use information from surveys measuring other statistics such as galaxy distributions or primary CMB anisotropies, so these poor constraints are not representative of what stage 3 and 4 experiments could achieve.
-By adding a prior our calculations become more realistic.
-Additionally, by having the prior be conservative, it stays clear how the different lensing spectra complement each other.
-The other prior we will use is based on the CMB temperature and $E$-mode polarization cross- and auto-powerspectra from $\ell=30$ to $\ell = 2000$.
-With the same noise values as used for the CMB weak lensing spectra.
-These results will be more realistic and serve to test whether weak lensing spectra are able to further tighten constraints in the presence of surveys measuring more standard statistics (such as the primary CMB anisotropies) with similar noise values.
+Future lensing surveys will of course use information from surveys measuring other statistics such as those derived from galaxy clustering and the primary CMB anisotropies.
+By adding a prior based on these (primary) statistics our results become more representative.
+Additionally, by considering a conservative prior, it remains clear how the different lensing spectra complement each other.
+The other prior we consider is based on the CMB temperature and $E$-mode polarization cross- and auto-powerspectra from $\ell=30$ to $\ell = 2000$ assuming the same noise properties as used for the CMB weak lensing reconstruction noise.
+Adopting these priors  serve to test how weak lensing statistics are able to further tighten constraints beyond measuring primary statistics.
+It is worth noting that the constraint for $\tau$ and $w_0$ are tighter than in other literature (see \cite{cmbs4sciencebook} and \cite{Namikawa_2016} for comparable $\tau$ and $w_0$ constraints, respectively).
+Based on a number of internal tests\footnote{
+This included looking at the accuracy of numerical derivatives w.r.t.
+cosmological parameters, testing the numerical stability of Fisher matrices wr.r.t.
+matrix inversion, and varying our values for detector noise and beam width.
+} we suspect that this is mainly due us not taking into account foregrounds in our noise models.
+
 \begin{table}[h!]
@@
 
-This paper considers noise levels for ``stage 3'' and ``stage 4'' weak lensing surveys.
-Values can be found in table \ref{tab:noiselevels}, including the experiments they are representative of.
+\subsection{Noise Modeling}
+This paper considers noise levels for ``stage 3'' and ``stage 4'' (weak lensing) surveys.
+Noise properties considered for this analysis can be found in table \ref{tab:noiselevels}, including experiments they are representative of.
 A comparison of the noise power spectra to the lensing power spectra are shown     in figure \ref{fig:lpsplusnoise}.
@@
     \multirow{2}{*}{CMB} 
-        & stage 3 & $\sigma = 1'$, $\Delta_P = 6' \; \mu \text{K}$ & Advanced ACTPol, SPT-3G \\
+        & stage 3 & $\sigma = 1'$, $\Delta_P = 6' \; \mu \text{K}$ & Advanced ACTPol, Simons Observatory \\
     \cline{2-4}
-        & stage 4 & $\sigma = 3'$, $\Delta_P = 1' \; \mu \text{K}$ & CMB-S4, LiteBIRD \\
+        & stage 4 & $\sigma = 3'$, $\Delta_P = 1' \; \mu \text{K}$ & CMB-S4 \\
     \hline
@@
 
-CMB lensing noise is estimated using the quadratic estimator from \cite{cmblensingestimator}.
+The CMB lensing noise is estimated using a quadratic estimator \cite{cmblensingestimator}.
 % We consider experiments that only measure polarization noise. This is similarly done in \cite{Namikawa_2016}. 
@@
 The noise in this measurement is dominated by scale-independent shot noise and has associated noise power spectrum $N_l^{\text{shear}} = \sigma_{\text{rms}}^2 / n_g$, \cite{bartelmann2001weak}.
-$n_g$ is the amount of galaxies observed per unit angle.
+$n_g$ is the amount of galaxies observed per unit solid angle.
 The noise power spectrum for the lensing potential then equals
 \begin{equation*}
-    N_\ell^{\text{lens. potential}} = \frac{(\ell-1)\ell(\ell+1)(\ell+2)}{4}N_\ell^{\text{shear}}.
+    N_\ell^{\text{lens. potential}} = \frac{4}{(\ell-1)\ell(\ell+1)(\ell+2)}N_\ell^{\text{shear}}.
 \end{equation*}
@@
     \caption{CMB (right) and galaxy (left) lensing potential power spectra compared to associated experimental noise.
-Current (stage 3) noise values are displayed as well as near future (stage 4) noise values.
+Current (stage 3) noise values are displayed as well as near future\footnote{Some of the stage 4 experiments are already in operation, but expected noise levels are only achieved after several years of integration.} (stage 4) noise values.
 The CMB lensing experiment uses only polarization.
 CMB noise values are chosen in accordance with \cite{Namikawa_2016}.
-For comparison we also how the noise reconstruction for the Simons Observatory.
+For comparison we also how the reconstruction noise for the Simons Observatory.
 Shear noise values are chosen to be similar to e.g.
@@
 A comparison of stage 3 noise, stage 4 noise, and lensing signal can be seen in figure \ref{fig:lpsplusnoise}.
-
+\subsection{Other details}
 The redshift distribution of the observed galaxies is commonly parameterized as \cite{bartelmann2001weak}
@@
 \section{Results}\label{sec:results}
-Lensing power spectra have high signal-to-noise in all cases. 
+As expected, lensing power spectra should be detectable at high signal-to-noise. 
 The signal-to-noise of the CMB and galaxy lensing bispectra versus the maximum multipole moment measured is shown in figure \ref{fig:snrplots}.
@@
 \cite{Kalaja_2023} for the effect of noise biases) and with high signal-to-noise with a stage 4 experiment.
-CMB $S/N$ results match those of reference \cite{Namikawa_2016}.
+Our results for CMB weak lensing $S/N$ match those of reference \cite{Namikawa_2016}.
 %It can be seen that the CMB bispectrum signal to noise never exceeds 2 even for $l_{\max} = 2000$, meaning that this signal is not likely to be able to contribute to parameter constraints in future surveys.
@@
     \includegraphics[width=\textwidth]{figures/snrplots.pdf}
-    \caption{Signal-to-noise ratios for galaxy lensing (left) and CMB lensing (right) for stage 3 (dashed) and stage 4 (not dashed) surveys as a function of maximum multipole measured.
+    \caption{Signal-to-noise ratios for galaxy lensing (left) and CMB lensing (right) bispectra for stage 3 (dashed) and stage 4 (not dashed) surveys as a function of maximum multipole measured.
 The minimum multipole is always $l_{\text{min}}=2$.
@@
 We look at two extension of $\Lambda CDM$, $m_\nu$ and $w_0$ as well as the derived parameters $\sigma_8$, and $\Omega_m$.
-Constraints on the full set of $\Lambda$CDM parameters can be found found in appendix \ref{sec:lambdacdmconstraints}.
+Constraints on the full set of $\Lambda$CDM parameters can be found in appendix \ref{sec:lambdacdmconstraints}.
 Results shown include marginalization over all parameters.
-Figures \ref{fig:paramconstraintstightcmb} and \ref{fig:paramconstraintstightgal} show confidence ellipses from stage 4 CMB and galaxy lensing surveys with the weak prior, respectively.
-Figures \ref{fig:paramconstraintstightcmbcmbprior} and \ref{fig:paramconstraintstightgalcmbprior} show confidence ellipses in the case of our CMB $T+E$ prior.
-Table \ref{tab:paramconstraintstight} lists all forecasted constraints .
+Figures \ref{fig:paramconstraintstightcmb} and \ref{fig:paramconstraintstightgal} show confidence ellipses from stage 4 CMB and galaxy lensing surveys assuming a weak prior, respectively.
+Figures \ref{fig:paramconstraintstightcmbcmbprior} and \ref{fig:paramconstraintstightgalcmbprior} show confidence ellipses in the case of our primary CMB $T+E$ prior.
+Table \ref{tab:paramconstraintstight} summerizes all forecasted constraints.
 
 Constraints are significantly improved by including bispectrum information in nearly all cases when using weak priors.
-When including the CMB $T+E$ priors, both the CMB lensing power- and bispectra significantly improve prior constraints, however combining them doesn't lead to significant improvements compared to only using lensing power spectra.
+When including the CMB $T+E$ priors, both the CMB lensing power- and bispectra significantly improve prior constraints, however combining them does not lead to significant improvements compared to only using lensing power spectra.
 For galaxy lensing, the power- and bispectra yield nearly identical improvements compared to the CMB $T+E$ prior.
 In this case, combining both statistics leads to significantly tighter constraints (in contrast to CMB lensing).
-When combining CMB and galaxy lensing we see that the bispectrum is slightly less competitive than the lensing power spectrum, but can still significantly improve constraints compared to only using the lensing power spectrum.
+When combining CMB and galaxy lensing we see that the bispectrum is slightly less competitive compared to the lensing power spectrum, but can still significantly improve constraints compared to only using the lensing power spectrum.
 % Again in agreement with reference \cite{Namikawa_2016}, in which case a 33 percent and 35 percent improvement was found for $m_\nu$ and $w_0$, respectively, by measuring the CMB lensing bispectrum in addition to the power spectrum.
 
-Furthermore, it is clear that, in general, combining CMB and galaxy lensing surveys also leads to significant improvements in constraining power, even in the case of the CMB $T+E$ prior.
-The improvement is strongest for the neutrino constraint.
+Furthermore, it is clear that, in general, combining CMB and galaxy lensing surveys also leads to significant improvements in constraining power, even in the case of a strong CMB $T+E$ prior.
+The improvement is most significant for the neutrino mass constraint.
 
@@
 \begin{itemize}
-    \item Lensing powers pectra alone are not sufficient to competitively constrain $\Lambda\mathrm{CDM} + w_0 + m_\nu$ without priors, even in the case of stage 4 surveys.
-
-    \item For lower noise values (in particular for stage 4 noise) the bispectra becomes increasingly important compared to the powerspectra.
+    \item Lensing power spectra alone are not sufficient to competitively constrain $\Lambda\mathrm{CDM} + w_0 + m_\nu$ without priors, even in the case of stage 4 surveys.
+
+    \item For lower noise values (in particular for stage 4 noise) the bispectra become increasingly important compared to the powerspectra.
 The lensing bispectra are directly related to the amount of non-gaussianity in the matter distribution.
-As noise levels becomes lower, we are are able to measure the matter distribution on small enough scales where nonlinear matter evolution becomes apparent.
-Additionally, for galaxy lensing, the window function peaks at earlier redshifts than compared to CMB lensing.
+As noise levels become lower, we are able to measure the matter distribution on small enough scales where nonlinear matter evolution becomes important.
+Additionally, for galaxy lensing, the window function peaks at later redshifts than compared to CMB lensing.
 The bispectrum is largest at late times, and thus the galaxy lensing bispectrum is significantly easier to detect and is better at constraining parameters then the CMB lensing bispectrum.
@@
 The main parameter combinations where approximate degeneracies are broken are $(n_s,\Omega_b h^2),\;(\Omega_b h^2,H_0),\;(\Omega_c h^2,H_0),\;(\Omega_c h^2,\Omega_b h^2)\;(n_s,\Omega_c h^2),\;(m_\nu,H_0),\;(m_\nu,n_s),$ $(A_s,\Omega_b h^2)$ (see also appendix \ref{sec:lambdacdmconstraints}).
-This means that even if stage 4 surveys aren't able to reach the noise levels assumed in this paper, the bispectra will likely still lead to significantly tighter constraints.
+This means that even if stage 4 surveys are not able to reach the noise levels assumed in this paper, the bispectra will likely still lead to significantly tighter constraints.
 
@@
 In this work, we have presented parameter forecasts from CMB lensing and galaxy lensing power- and bispectra, considering experimental parameters representative of Stage 3 and Stage 4 CMB and galaxy surveys.
-Our analysis shows that while lensing power spectra are well detectable, the inclusion of bispectra generally offers significant improvements when considering parameter constraints.
-Notably, both CMB and galaxy bispectra are found to be measurable, even with Stage 3 experimental noise levels.
-
-The primary impact of incorporating bispectrum information is a tightening of cosmological parameter constraints, particularly for parameters like $H_0$, $\sigma_8$, and $\Omega_m$, and in breaking key parameter degeneracies.
-We found that for sufficiently low noise levels, characteristic of Stage 4 surveys, the bispectra themselves can offer constraining power comparable to or even exceeding that of the power spectra, due to a more favorable scaling of information content with the maximum usable multipole\footnote{The multipole at which the noise power spectrum starts exceeding the cosmological power spectrum.}.
+Our analysis shows that while lensing power spectra are detectable at high significance, the inclusion of bispectra generally offers significant improvements when considering parameter constraints when considering an LCDM+ model of the universe.
+Notably, both CMB and galaxy bispectra are found to be detectable, even with Stage 3 experimental noise levels.
+
+The primary impact of incorporating bispectrum information is a tightening of cosmological parameter constraints, particularly for parameters such as $H_0$, $\sigma_8$, and $\Omega_m$, and in breaking key parameter degeneracies.
+We found that for sufficiently low noise levels, characteristic of Stage 4 surveys, the bispectra themselves can offer constraining power comparable to that of the power spectra, due to nonlinear structure formation on small scales being better detected by these future surveys.
 Furthermore, the combination of CMB and galaxy lensing probes, especially when both power spectra and bispectra are utilized, further enhances constraints and aids in mitigating degeneracies.
@@
 This intrinsic inaccuracy in the model for $B^{\delta}$ will propagate to the lensing bispectra.
+
      \item Certain cosmological parameters that affect nonlinear large-scale structure, such as neutrino masses, likely alter the fitting parameters of the matter bispectrum formula when varied.
 In the literature, this is currently not considered, which can make the derivatives of the bispectrum with respect to these cosmological parameters less accurate.
+
      \item We assumed throughout this paper that surveys would measure multipoles as low as $\ell = 2$.
@@
 we ignored any contributions not arising from standard Wick contractions.
-A next step could be to account for these additional contributions, perhaps through N-body and ray-tracing simulations.
+We know that the estimated lensing potential is not Gaussian in general, for two reasons: (i) The matter distribution is not Gaussian.
+(ii) Even with a Gaussian matter distribution, the quadratic estimator we used will not be Gaussian.
+Consider, for example, $\langle \hat \phi \hat \phi \hat \phi \rangle$.
+Each $\hat \phi$ is given as a sum over products of the lensing CMB anisotropies, $T$, $E$, $B$.
+Writing out each $\hat \phi$ in this way would yield terms such as $\langle TTTTTT\rangle$ (i.e.
+6 point functions).
+These terms are non-zero.
+In fact, not only would they give products of 2 point functions due to Wick contractions, but also connected $n$-point functions with $n>2$ because the lensed fields are no longer Gaussian even if the original fields are.
+A next step could thus be to account for these additional contributions, perhaps through N-body and ray-tracing simulations.
 Notably, such simulations could potentially also allow for a more accurate calculation of the lensing spectra themselves, without relying on fitting formulas for the matter bispectrum or the Limber approximation, depending on simulation accuracy.
-This might also explain the fact that in some instances the bispectra appear more constraining compared to the power spectra.
+
+     \item Related to the previous point, for a completely correct analysis we would also need to take into account correlations between a CMB prior and the lensing power- and bispectra.
+Currently we simply add Fisher matrices, however taking these correlations into account means computing a covariance matrix for combinations of $T$, $E$, $\phi$ with nonzero entries for correlations between $T$/$E$ and $\phi$.
+For instance, consider the correlation between the estimator of the temperature powerspectrum and the lensing powerspectrum at arbitrary multipoles.
+We will get terms like $\langle TT\hat\phi\hat\phi \rangle$, which will be a sum over 6 point functions like $\langle TT TETE\rangle$ due to our quadratic estimator.
+These 6 point functions are clearly non-zero, as explained earlier.
 
@@
 While standard, such estimators may not be strictly optimal, especially with complex foregrounds or non-Gaussianities.
-Better estimators exist; for example, the iterative estimator developed in \cite{Smith_2012} can lead to a significantly better signal-to-noise ratio for the CMB lensing bispectrum, as shown in \cite{Namikawa_2016}.
+Estimators that, given our Stage 4 noise parameters, perform better do exist; for example, the iterative estimator developed in \cite{Smith_2012} can lead to a significantly better signal-to-noise ratio for the CMB lensing bispectrum, as shown in \cite{Namikawa_2016}.
 Using such an iterative estimator may also result in the CMB bispectrum being able to significantly improve parameter constraints when combined with lensing power spectra and a CMB $T+E$ prior.
@@
 
-    \item A recent paper \cite{postborn} has shown that the Born approximation used to derive the formulas for the lensing power and bispectra, is no longer negligible for the lensing bispectrum.
+    \item A 2024 paper \cite{postborn_pratten_lewis} has shown that the Born approximation used to derive the formulas for the lensing power and bispectra, is no longer negligible for the lensing bispectrum.
 The post-Born corrections calculated in that paper only changed the lensing power spectra by $\leq 0.2\%$, while for the bispectra this effect was much larger, in particular for folded configurations.
 For future work it would thus be worthwile to analyze how these post-Born corrections affect the results presented in this paper.
+
+    \item We did not model foreground contamination such as the cosmic infrared background (CIB), thermal Sunyaev-Zel'dovich (tSZ) effect, or point sources.
+These foregrounds can bias lensing reconstructions and bispectra, particularly in CMB–galaxy cross-correlations, and should be accounted for in future analyses.
+
 \end{itemize}
@@
 \section*{Acknowledgements}
-We would like to thank prof.
+We would like to thank Prof.
 Toshiya Namikawa for providing extra details on his paper on the CMB lensing bispectrum \cite{Namikawa_2016}, which, in particular, allowed us to cross check (intermediate) results.
+The authors also like to thank Luna Jansma, who wrote a MSc thesis on the same subject, which served as the foundation of this study.
 
@@
 \end{equation}
-At 1st order, $A$ is a symmetric matrix by definition and $\omega$ vanishes, we will ignore it from here on out.
+In the weak lensing regime and under the Born approximation, $A$ is a symmetric matrix by definition and $\omega$ vanishes, we will ignore it from here on out.
 Intuitively, $A$ tells you how a small patch in the sky transforms due to lensing.

### 2025-07-06 code/lps_vs_noise_plot.py
added SO goal curves

@@ -30,16 +30,16 @@ names = ['TT', 'TE', 'TB', 'EE', 'EB']
  
 # SIMONS OBSERVATORY NOISE LEVELS
 # SO_noise_baseline_file_path = '/scratch/p319950/data/conv_noise.dat'
-#SO_noise_goal_file_path = '/Users/jonasfrugte/Desktop/fisher_calc_weak_lensing/code/cmb_noise_files/nlkk_v3_1_0_deproj0_SENS2_fsky0p4_it_lT30-3000_lP30-5000.dat'
+SO_noise_goal_file_path = './cmb_noise_files/nlkk_v3_1_0_deproj0_SENS2_fsky0p4_it_lT30-3000_lP30-5000.dat'
 
 # conv_noise_data_array = np.loadtxt(SO_noise_baseline_file_path)
-#conv_noise_goal_data_array = np.loadtxt(SO_noise_goal_file_path)
+conv_noise_goal_data_array = np.loadtxt(SO_noise_goal_file_path)
 
 # SO_noise = lambda l : conv_noise_data_array[l-2, 7] * 4. # * (l * 1.0)**(-2) * (l + 1.0)**(-2)
 # SO_noise_values = np.array([SO_noise(int(l)) for l in l_values])
 
-#SO_noise_goal = lambda l : 0.4 * conv_noise_goal_data_array[l-2, 7] * 4. # * (l * 1.0)**(-2) * (l + 1.0)**(-2)
-#SO_noise_goal_values = np.array([SO_noise_goal(int(l)) for l in l_values])
+SO_noise_goal = lambda l : 0.4 * conv_noise_goal_data_array[l-2, 7] * 4. # * (l * 1.0)**(-2) * (l + 1.0)**(-2)
+SO_noise_goal_values = np.array([SO_noise_goal(int(l)) for l in l_values])
 
 # GALAXY LENSING NOISE
 sigma_rms_S3 = 0.3
@@ -71,7 +71,7 @@ axs[0].grid(True, which='both', linestyle='--', linewidth=0.5)
 axs[1].loglog(l_values, l_values**4 * cmb_lps, label='CMB lensing potential powerspectrum', color='black')
 # CHECK CONVERSION FACTOR HERE
 # axs[1].loglog(l_values, SO_noise_values, label='SO Noise, baseline', linestyle='dashed', color='blue')
[…]

### 2025-07-06 paper/paper_draft.tex

@@
 \section{Results}\label{sec:results}
+\subsection{Detectability}
 As expected, lensing power spectra should be detectable at high signal-to-noise. 
@@
 Our results for CMB weak lensing $S/N$ match those of reference \cite{Namikawa_2016}.
-%It can be seen that the CMB bispectrum signal to noise never exceeds 2 even for $l_{\max} = 2000$, meaning that this signal is not likely to be able to contribute to parameter constraints in future surveys.
-All other quantities should be well detectable for $l_{\max}\geq 500$.
+
+Regarding post-Born corrections, we found that the effect on the galaxy lensing bispectrum $S/N$ is a few percent at most, in agreement with \cite{galpostbornlbscorr}.
+We thus did not include additional plots for these corrections.
+For the CMB the $S/N$ is reduced significantly (in agreement with \cite{postborn_pratten_lewis}) but not so severely that they are no longer detectable.
+Our conclusions thus remain the same: the CMB bispectrum is probably detectable with stage 3 surveys and definitely detectable with stage 4 surveys.
+
 \begin{figure}[h!]
@@
 
+\subsection{Parameter Constraints}
 Next, we consider cosmological parameter constraints.

### 2025-07-07 code/Fisher_matrix_approx.py
new fish pics

@@ -4,7 +4,7 @@ import Fisher_calc as vis
 from itertools import *
 lmin = 2
 lmax = 2000
-stepsizes = [33, 133, 166]
+stepsizes = [11, 44, 55]
 
 print(f'Stepsizes are: {stepsizes}')
 
@@ -41,7 +41,7 @@ def fisher_calc_wrapper_takada_jain(args, tracers):
         return vis.Fisher_mat_single(50, 0, 200, stepsizes[0], num_bispec_samples, pars[i], pars[j], num_cores, b's') + vis.Fisher_mat_single(50, 200, 1000, stepsizes[1], num_bispec_samples, pars[i], pars[j], num_cores, b's') + vis.Fisher_mat_single(50, 1000, 2000, stepsizes[2], num_bispec_samples, pars[i], pars[j], num_cores, b's')
 num_pars = len(pars)
 
-pb_corrections = True
+pb_corrections = False
 
 def main():
     for tracer in ['both', 's', 'c']:

### 2025-07-07 code/Fisher_matrix_powerspectra.py

@@ -3,10 +3,10 @@ import Fisher_calc_python_imp as vispy
 import numpy as np
 from itertools import *
 
-lmin = 100
+lmin = 2
 lmax = 2000
 
-pars = [b'H', b'ombh2', b'omch2', b'ns', b'mnu', b'tau', b'As', b'w0'] # add tau back in later!
+pars = [b'H', b'ombh2', b'omch2', b'ns', b'mnu', b'tau', b'As', b'w0']
 # Wrapper function for multiprocessing
 def fisher_calc_wrapper(args, tracers):
     i, j = args
@@ -46,29 +46,19 @@ def fisher_calc_wrapper_cmb(args, tracers):
         # lmax should actually be 3000, but interpolation currently doesn't go that high
         return vispy.Fisher_powersp_cmb(lmin, lmax, par1 = pars[i], par2 = pars[j])
 
-def main(lmin_loc):
-    global lmin
-    lmin = lmin_loc
-    for tracer in ['t']: #['c', 's', 'both']:
+def main():
+    for tracer in ['c', 's', 'both']:
         mat = np.zeros((len(pars), len(pars)))
         for i, j in product(range(len(pars)), repeat = 2):
-            result = fisher_calc_wrapper_cmb((i, j), tracer)
+            result = fisher_calc_wrapper((i, j), tracer)
             mat[i, j] = result
             mat[j, i] = result  # Symmetric assignment
 
-        #np.savetxt(f'fisher_matrices/fish_mat_powersp_{tracer}_cmb_lmin30.txt', mat)
-        print(np.linalg.inv(mat))
-        return mat
-
-    # for tracer in ['c', 's', 'both']: #['c', 's', 'both']:
-    #     mat = np.zeros((len(pars), len(pars)))
-    #     for i, j in product(range(len(pars)), repeat = 2):
-    #         result = fisher_calc_wrapper((i, j), tracer)
-    #         mat[i, j] = result
-    #         mat[j, i] = result  # Symmetric assignment
-
-    #     np.savetxt(f'fisher_matrices/fish_mat_powersp_{tracer}.txt', mat)
-    #     print(mat)
+        filepath = f'fisher_matrices/fish_mat_powersp_{tracer}.txt'
+        np.savetxt(filepath, mat)
+        print(f'Created Fisher matrix at {filepath}')
+        print(mat)
+    pass
 
 if __name__ == '__main__':
     main()

### 2025-07-07 paper/playing_with_fish.py

@@ -177,9 +177,9 @@ sigmaomm_ders = [
     (0.25 * sigma8 * omm**(-0.75))]
 
 
-keep_indices = [4, 7, 8, 10]
+#keep_indices = [4, 7, 8, 10]
 #keep_indices=[10]
-#keep_indices = [0, 1, 2, 3, 5, 6]
+keep_indices = [0, 1, 2, 3, 5, 6] # lcdm pars
 #keep_indices = [4, 6]
 
 def process_fishes(fishes, keep_indices, derived_param_derivss = [s8_ders, S8_ders, omm_ders, sigmaomm_ders]):
@@ -187,9 +187,6 @@ def process_fishes(fishes, keep_indices, derived_param_derivss = [s8_ders, S8_de
 
     # extend to include derived params
     for derived_param_derivs in derived_param_derivss:
-        print(np.shape(fishes_inv))
-        print(np.shape(derived_param_derivs))
-
         fishes_inv = append_row_column(fishes_inv, derived_param_derivs)
     
     # marginalize over unwanted params
@@ -346,7 +343,6 @@ def plot_corner(
     # Off-diagonal subplots: 2D ellipses
     for i in range(N-1):
         for j in range(i, N-1):
-            print(i, j)
             ax_low = axes[j, i]  # lower triangle
             ax_high = axes[i, j] # upper triangle
             if i != j:
@@ -396,8 +392,6 @@ def plot_corner(
 from tabulate import tabulate
 
 def save_table(param_names, param_vals, which_pars, constraints, labels):
-    print(param_names)
-    print(which_pars)
     table = [[0 for i in range(len(constraints) + 1)] for j in range(len(which_pars))]
     for i in range(len(which_pars)):
         # param names and fiducial values
@@ -407,7 +401,7 @@ def save_table(param_names, param_vals, which_pars, constraints, labels):
         for j in range(len(constraints)):
             table[i][j+1] = np.abs(np.sqrt(constraints[j][i][i]))
     labels = ['Par'] + labels
[…]

### 2025-07-11 code/lps_vs_noise_plot.py
changed figures slightly

@@ -2,7 +2,7 @@ import cosm_setup as cs
 import matplotlib.pyplot as plt
 import numpy as np
 
-fig, axs = plt.subplots(1, 2, figsize = (10, 5))
+fig, axs = plt.subplots(1, 2, figsize = (8.3, 4))
 
 # LENSING SPECTRA
 l_values = np.logspace(np.log10(2), np.log10(3000), 100)

### 2025-07-11 paper/figures/snr_plots.py

@@ -63,9 +63,9 @@ axs[1].plot([ls[-1]], [plot_datas[6][-1]], marker='o', color='blue', label='_nol
 axs[0].minorticks_on()
 axs[1].minorticks_on()
 
-axs[0].set_xlabel(r'$l_{\max}$', fontsize=12)
-axs[0].set_ylabel(r'$S/N$', fontsize=12)
-axs[1].set_xlabel(r'$l_{\max}$', fontsize=12)
+axs[0].set_xlabel(r'$l_{\max}$', fontsize=10)
+axs[0].set_ylabel(r'$S/N$', fontsize=10)
+axs[1].set_xlabel(r'$l_{\max}$', fontsize=10)
 # axs[1].set_ylabel(r'$S/N$', fontsize=12)
 
 # Increase the x-axis limit to add extra space on the right

### 2025-07-11 paper/playing_with_fish.py

@@ -23,6 +23,12 @@ visb_s = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_s.txt')
 
 visb_f = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_both.txt')
 
+visb_c_pb = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_c_pb.txt')
+
+visb_s_pb = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_s_pb.txt')
+
+visb_f_pb = np.loadtxt(fisher_matrices_dir + '/fish_mat_bisp_both_pb.txt')
+
 ########################
 # Powerspectra Matrices
 ########################
@@ -177,9 +183,9 @@ sigmaomm_ders = [
     (0.25 * sigma8 * omm**(-0.75))]
 
 
-keep_indices = [4, 7, 8, 10]
+# keep_indices = [4, 7, 8, 10]
 #keep_indices=[10]
-#keep_indices = [0, 1, 2, 3, 5, 6] # lcdm pars
+keep_indices = [0, 1, 2, 3, 5, 6] # lcdm pars
 #keep_indices = [4, 6]
 
 def process_fishes(fishes, keep_indices, derived_param_derivss = [s8_ders, S8_ders, omm_ders, sigmaomm_ders]):
@@ -441,34 +447,15 @@ if __name__ == "__main__":
     param_names_latex_kept = [param_names_latex[i] for i in keep_indices]
     param_values_kept      = [param_values[i] for i in keep_indices]
 
-    # cov_matrices = [
-    #     inv(visb_c_reduced + visp_c_reduced),
-    #     inv(visb_s_reduced + visp_s_reduced),
-    #     inv(visb_f_reduced + visp_f_reduced)
-    # ]
+    # fish pond numbers correspond to:
+    # 0: for the table, strong prior
+    # 1: for the table, weak prior
+    # 2, 3, 4, 5: for plots, tight parameters
+    # 6, 7, 8, 9: for plots, lcdm parameters
+    # 0.5: for the table, strong prior, with pb corrections
+    # 1.5: for the table, weak prior, with pb corrections
 
-    # labels = ["CMB", "Galaxy", "CMB + Galaxy"]
-
-    # fish_matrices = [
-    #     0.5 * viscmb_f,
-    #     0.5 * (viscmb_f + 1.5 * visp_c),
[…]

### 2025-07-11 paper/paper_draft.tex
Updates from Overleaf

@@
 \usepackage{booktabs}
+\usepackage[table]{xcolor}
 
@@
 \abstract{
-Upcoming Stage 4 surveys, such as the Simons Observatory, LSST, and Euclid are poised to measure weak gravitational lensing of the Cosmic Microwave Background (CMB) and galaxies with unprecedented precision.
+Upcoming stage 4 surveys, such as the Simons Observatory, LSST, and Euclid are poised to measure weak gravitational lensing of the Cosmic Microwave Background (CMB) and galaxies with unprecedented precision.
 While the power spectrum is the standard statistic used to analyze weak lensing data, non-Gaussianity induced by the non-linear growth of structure imparts significant cosmological information into higher-order statistics.
 In this paper, we forecast the ability of future surveys to constrain cosmological parameters using the weak lensing power spectrum and bispectrum from both CMB and galaxy surveys, as well as their cross-correlations.
-We consider an eight-parameter cosmological model ($\Lambda$CDM + $\sum m_{\nu}$ + $w_{0}$) and assess parameter constraints for Stage 4 survey specifications.
+We consider an eight-parameter cosmological model ($\Lambda$CDM + $\sum m_{\nu}$ + $w_{0}$) and assess parameter constraints for stage 4 survey specifications.
 Our analysis shows that, in the absence of systematics, both the CMB and galaxy lensing bispectra will be detectable at high signal-to-noise.
@@
 Combining the two probes, particularly with both their lensing power- and bi-spectra, leads to significantly tighter constraints than achievable with either probe individually, with the greatest improvement seen for the sum of neutrino masses.
+Additionally, we also analyze the changes in detectability and parameter constraints when taking into account post-Born corrections to the lensing bispectra, as these were shown to be significant, especially for the CMB lensing bispectrum \cite{postborn_pratten_lewis}.
+We show that the conclusions of this paper remain largely the same with this correction.
 }
@@
 This can be measured in terms of the lensing convergence $\kappa$, lensing shear $\gamma$, or lensing potential $\psi$.
-In the weak lensing regime and under the Born approximation they are all equivalent, in the sense that they can be directly converted into one another (appendix \ref{sec:weaklensstats}).
-In CMB surveys, the temperature and polarization anisotropies are used to reconstruct the lensing potential \cite{Hu_2002}, while with galaxies one instead calculates lensing effects by measuring the lensing shear, which can be deduced from the ellipticity distribution of the observed galaxies \cite{HoekstraJain2008}.
+In the weak lensing regime and under the Born approximation, they are all equivalent, in the sense that they can be directly converted into one another (appendix \ref{sec:weaklensstats}).
+In CMB surveys, the temperature and polarization anisotropies are used to reconstruct the lensing potential \cite{Hu_2002}, while with galaxies, one instead calculates lensing effects by measuring the lensing shear, which can be deduced from the ellipticity distribution of the observed galaxies \cite{HoekstraJain2008}.
 
 Even with Gaussian initial conditions, there are significant amounts of non-Gaussianity in the weak lensing signal, especially in galaxy lensing due to it being a probe of the more recent distribution of matter \cite{Bernardeau1997,Takada2003}.
-In addition to looking at the power spectrum, a natural next step is thus to measure weak lensing bispectra.
+In addition to looking at the power spectrum, a natural next step is to measure weak lensing bispectra.
 While the bispectrum has been measured for galaxy lensing signals \cite{vanWaerbeke2002}, a definitive detection of the CMB lensing bispectrum is still pending \cite{Kalaja_2023,Namikawa_2016}.
-Upcoming surveys, such as the ones mentioned earlier may be able to detect this signal.
-Therefore it is timely to consider the potential of higher order spectra to constrain cosmological parameters.
+Upcoming surveys, such as the ones mentioned earlier, may be able to detect this signal.
+Therefore it is timely to consider the potential of higher-order spectra to constrain cosmological parameters.
 
@@
 Results shown include marginalization over all parameters.
-Figures \ref{fig:paramconstraintstightcmb} and \ref{fig:paramconstraintstightgal} show confidence ellipses from stage 4 CMB and galaxy lensing surveys assuming a weak prior, respectively.
-Figures \ref{fig:paramconstraintstightcmbcmbprior} and \ref{fig:paramconstraintstightgalcmbprior} show confidence ellipses in the case of our primary CMB $T+E$ prior.
-Table \ref{tab:paramconstraintstight} summerizes all forecasted constraints.
+Figures \ref{fig:paramconstraintstightcmb} and \ref{fig:paramconstraintstightgal} show confidence ellipses assuming a weak prior from stage 4 CMB and galaxy lensing surveys, respectively.
+Figures \ref{fig:paramconstraintstightcmbcmbprior} and \ref{fig:paramconstraintstightgalcmbprior} similarly show confidence ellipses in the case of our primary CMB $T+E$ prior.
+Table \ref{tab:paramconstraintstight} summarizes all forecasted constraints.
 
@@
 
-    \item For lower noise values (in particular for stage 4 noise) the bispectra become increasingly important compared to the powerspectra.
-The lensing bispectra are directly related to the amount of non-gaussianity in the matter distribution.
+    \item For lower noise values (in particular for stage 4 noise), the bispectra become increasingly important compared to the power spectra.
+The lensing bispectra are directly related to the amount of non-Gaussianity in the matter distribution.
 As noise levels become lower, we are able to measure the matter distribution on small enough scales where nonlinear matter evolution becomes important.
-Additionally, for galaxy lensing, the window function peaks at later redshifts than compared to CMB lensing.
-The bispectrum is largest at late times, and thus the galaxy lensing bispectrum is significantly easier to detect and is better at constraining parameters then the CMB lensing bispectrum.
+Additionally, for galaxy lensing, the window function peaks at later redshifts compared to CMB lensing.
+The bispectrum is largest at late times, and thus the galaxy lensing bispectrum is significantly easier to detect and is better at constraining parameters than the CMB lensing bispectrum.
 
@@
     \item Stage 4 galaxy weak lensing surveys generally contain more information than CMB surveys. 
-    % Comparing the confidence ellipses, it is clear that the two types of surveys depend on the underlying parameters in different ways.
+    Comparing the confidence ellipses, it is clear that the two types of surveys depend on the underlying parameters in different ways.
 This leads to substantially better constraints for $\sigma_8$, $\Omega_m$, $m_\nu$, and $w_0$ when combining CMB and galaxy lensing.
+
+    \item It is clear that post-Born corrections do not change the constraints drastically.
+As expected, they matter less for galaxy lensing and when using the CMB $T+E$ prior.
+Interestingly, although the constraints due to bispectra alone are always worse, for some parameters, they are instead improved when combining lensing bispectrum information with that of the lensing power spectrum.
+This implies that, at least in some cases, the post-Born contribution to the lensing bispectrum depends on the parameters in a different way than the lensing power spectrum and thus improves the amount of information we can extract from survey data despite making the lensing bispectrum less detectable.
+Besides these observations, we conclude that taking into account post-Born corrections does not lead to different conclusions about the overall sensitivity of the lensing spectra.
 \end{itemize}
@@
     \scriptsize
+%     \begin{tabular}{|l|r|rrr|rrr|rrr|}
+%         \hline
+%         \multicolumn{11}{|c|}{weak priors} \\
+%         \hline
+%         && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
+%         \hline
+%         Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
+% \hline
+%  $m_\nu$    & 1.000e+01 &  2.815e-01 &                  1.104e+00 &                           2.247e-01 &  9.954e-01 &                  2.646e-01 &                           1.551e-01 &  1.042e-01 &                  1.317e-01 &                           6.537e-02 \\
+%  $w_0$      & 1.000e+01 &  5.415e-01 &                  1.405e+00 &                           3.683e-01 &  2.638e-01 &                  7.817e-02 &                           2.475e-02 &  5.936e-02 &                  6.629e-02 &                           1.899e-02 \\
+%  $\sigma_8$ & 2.457e+00 &  1.181e-01 &                  1.872e-01 &                           5.486e-02 &  8.069e-02 &                  3.668e-02 &                           1.299e-02 &  1.375e-02 &                  1.587e-02 &                           8.048e-03 \\
+%  $\Omega_m$ & 6.702e-01 &  1.948e-01 &                  1.790e-01 &                           7.270e-02 &  1.434e-01 &                  4.831e-02 &                           1.667e-02 &  1.619e-02 &                  1.607e-02 &                           7.233e-03 \\
+%        \hline
+%         \multicolumn{11}{|c|}{CMB $T + E$ priors} \\
+%         \hline
+%         && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
+%         \hline
+%          Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
+%         \hline
+%  $m_\nu$    & 2.157e-01 &  7.445e-02 &                  1.178e-01 &                           7.283e-02 &  1.406e-01 &                  1.042e-01 &                           8.592e-02 &  4.340e-02 &                  8.965e-02 &                           3.735e-02 \\
+%  $w_0$      & 5.781e-02 &  4.606e-02 &                  5.285e-02 &                           4.578e-02 &  4.756e-02 &                  3.934e-02 &                           2.080e-02 &  3.231e-02 &                  3.667e-02 &                           1.741e-02 \\
+%  $\sigma_8$ & 1.547e-02 &  1.360e-02 &                  1.492e-02 &                           1.349e-02 &  1.252e-02 &                  1.136e-02 &                           7.881e-03 &  9.682e-03 &                  1.023e-02 &                           6.126e-03 \\
+%  $\Omega_m$ & 1.495e-02 &  1.381e-02 &                  1.457e-02 &                           1.376e-02 &  1.336e-02 &                  1.124e-02 &                           8.715e-03 &  1.028e-02 &                  9.616e-03 &                           5.484e-03 \\
+%        \hline
+%             \multicolumn{11}{|c|}{weak priors, with post-Born corrections} \\
+%         \hline
+%         && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
+%         \hline
+%         Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
+% \hline
+% $m_\nu$    & 1.000e+01 &  2.815e-01 &                  1.158e+00 &                           1.805e-01 &  9.954e-01 &                  3.349e-01 &                           2.073e-01 &  1.042e-01 &                  1.775e-01 &                           6.256e-02 \\
+%  $w_0$      & 1.000e+01 &  5.415e-01 &                  8.193e-01 &                           3.051e-01 &  2.638e-01 &                  8.328e-02 &                           2.825e-02 &  5.936e-02 &                  6.271e-02 &                           1.988e-02 \\
+%  $\sigma_8$ & 2.457e+00 &  1.181e-01 &                  1.846e-01 &                           5.660e-02 &  8.069e-02 &                  4.047e-02 &                           1.317e-02 &  1.375e-02 &                  1.782e-02 &                           8.369e-03 \\
+%  $\Omega_m$ & 6.702e-01 &  1.948e-01 &                  1.810e-01 &                           7.086e-02 &  1.434e-01 &                  5.474e-02 &                           1.811e-02 &  1.619e-02 &                  1.799e-02 &                           7.325e-03 \\
+%        \hline
+%         \multicolumn{11}{|c|}{CMB $T + E$ priors, with post-Born corrections} \\
+%         \hline
+%         && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
+%         \hline
+%          Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
+%         \hline
+%  $m_\nu$    & 2.157e-01 &  7.445e-02 &                  1.298e-01 &                           7.313e-02 &  1.406e-01 &                  1.166e-01 &                           9.869e-02 &  4.340e-02 &                  1.034e-01 &                           3.766e-02 \\
+%  $w_0$      & 5.781e-02 &  4.606e-02 &                  5.308e-02 &                           4.596e-02 &  4.756e-02 &                  4.163e-02 &                           2.220e-02 &  3.231e-02 &                  3.911e-02 &                           1.798e-02 \\
+%  $\sigma_8$ & 1.547e-02 &  1.360e-02 &                  1.498e-02 &                           1.354e-02 &  1.252e-02 &                  1.168e-02 &                           7.717e-03 &  9.682e-03 &                  1.073e-02 &                           6.285e-03 \\
+%  $\Omega_m$ & 1.495e-02 &  1.381e-02 &                  1.467e-02 &                           1.379e-02 &  1.336e-02 &                  1.141e-02 &                           8.530e-03 &  1.028e-02 &                  1.017e-02 &                           5.687e-03 \\
+%        \hline
+%         \end{tabular}
  \begin{tabular}{|l|r|rrr|rrr|rrr|}
         \hline
-        \multicolumn{11}{|c|}{weak priors} \\
-        \hline
-        && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
+        \multicolumn{11}{|c|}{weak priors (with and without post-Born corrections)} \\
+        \hline
+        && \multicolumn{3}{c|}{CMB lensing} & \multicolumn{3}{c|}{Gal. lensing} & \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
         \hline
@@
         \hline
-        $m_\nu$    &     1.00e+01 &  2.82e-01 &  1.13e+00 &        2.26e-01 & 2.60e-01 & 4.07e-01 &      1.85e-01 &      9.54e-02 &      1.81e-01 &             6.46e-02 \\
-        $w_0$      &     1.00e+01 &  5.42e-01 &  1.44e+00 &        3.87e-01 & 2.14e-01 & 1.28e-01 &      3.22e-02 &      1.51e-01 &      1.13e-01 &             2.45e-02 \\
-        $\sigma_8$ &     2.46e+00 &  1.18e-01 &  1.95e-01 &        5.69e-02 & 5.36e-02 & 5.43e-02 &      1.73e-02 &      3.48e-02 &      3.32e-02 &             9.20e-03 \\
-        $\Omega_m$ &     6.70e-01 &  1.95e-01 &  1.84e-01 &        7.73e-02 & 7.57e-02 & 5.88e-02 &      1.81e-02 &      5.15e-02 &      3.76e-02 &             8.57e-03 \\
-       \hline
-        \multicolumn{11}{|c|}{CMB $T + E$ priors} \\
-        \hline
-        && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
+$m_\nu$    & 10 & 0.28 & 1.1 & 0.23 & 1.0 & 0.26 & 0.16 & 0.10 & 0.13 & 0.065 \\
+\rowcolor{blue!10}
+           &    &  & 1.2 & 0.18 & & 0.33 & 0.21 & & 0.18 & 0.063 \\
+$w_0$      & 10 & 0.54 & 1.4 & 0.37 & 0.26 & 0.078 & 0.025 & 0.059 & 0.066 & 0.019 \\
+\rowcolor{blue!10}
+           &    &  & 0.82 & 0.31 &  & 0.083 & 0.028 & & 0.063 & 0.020 \\
+$\sigma_8$ & 2.5 & 0.12 & 0.19 & 0.055 & 0.081 & 0.037 & 0.013 & 0.014 & 0.016 & 0.0080 \\
+\rowcolor{blue!10}
+           &     &  & 0.18 & 0.057 & & 0.040 & 0.013 & & 0.018 & 0.0084 \\
+$\Omega_m$ & 0.67 & 0.19 & 0.18 & 0.073 & 0.14 & 0.048 & 0.017 & 0.016 & 0.016 & 0.0072 \\
+\rowcolor{blue!10}
+           &     &  & 0.18 & 0.071 & & 0.055 & 0.018 & & 0.018 & 0.0073 \\
+        \hline
+        \multicolumn{11}{|c|}{CMB $T + E$ priors (with and without post-Born corrections)} \\
+        \hline
+        && \multicolumn{3}{c|}{CMB lensing} & \multicolumn{3}{c|}{Gal. lensing} & \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
         \hline
@@
         \hline
-        $m_\nu$    &    2.16e-01 &  5.28e-02 &  1.21e-01 &        5.22e-02 & 1.48e-01 & 1.49e-01 &      1.12e-01 &      1.88e-02 &      1.02e-01 &             1.62e-02 \\
-        $w_0$      &    5.78e-02 &  4.19e-02 &  5.30e-02 &        4.19e-02 & 4.66e-02 & 4.77e-02 &      2.55e-02 &      3.35e-02 &      4.33e-02 &             1.99e-02 \\
-        $\sigma_8$ &    1.55e-02 &  1.06e-02 &  1.50e-02 &        1.06e-02 & 1.18e-02 & 1.21e-02 &      8.18e-03 &      8.99e-03 &      1.19e-02 &             5.28e-03 \\
-        $\Omega_m$ &    1.50e-02 &  1.20e-02 &  1.46e-02 &        1.20e-02 & 1.34e-02 & 1.23e-02 &      9.32e-03 &      1.13e-02 &      1.19e-02 &             6.16e-03 \\
+$m_\nu$    & 0.22 & 0.074 & 0.12 & 0.073 & 0.14 & 0.10 & 0.086 & 0.043 & 0.090 & 0.037 \\
+\rowcolor{blue!10}
+           &      & & 0.13 & 0.073 & & 0.12 & 0.099 & & 0.10 & 0.038 \\
+$w_0$      & 0.058 & 0.046 & 0.053 & 0.046 & 0.048 & 0.039 & 0.021 & 0.032 & 0.037 & 0.017 \\
+\rowcolor{blue!10}
+           &       & & 0.053 & 0.046 & & 0.042 & 0.022 &  & 0.039 & 0.018 \\
+$\sigma_8$ & 0.015 & 0.014 & 0.015 & 0.013 & 0.013 & 0.011 & 0.0079 & 0.0097 & 0.010 & 0.0061 \\
+\rowcolor{blue!10}
+           &       & & 0.015 & 0.014 &  & 0.012 & 0.0077 & & 0.011 & 0.0063 \\
+$\Omega_m$ & 0.015 & 0.014 & 0.015 & 0.014 & 0.013 & 0.011 & 0.0087 & 0.010 & 0.0096 & 0.0055 \\
+\rowcolor{blue!10}
+           &       & & 0.015 & 0.014 & & 0.011 & 0.0085 & & 0.010 & 0.0057 \\
         \hline
     \end{tabular}
+
 \caption{Parameter constraints for different combinations of weak lensing information.
-$C$ stands for powerspectra, and $B$ for bispectra.
+$C$ stands for power spectra, and $B$ for bispectra.
 All surveys are assumed to be stage 4 surveys with multipole range $2 \leq \ell \leq 2000$.
-We use the weak and CMB priors described in table \ref{tab:fiducialpars}.}
+We use the weak and CMB priors described in table \ref{tab:fiducialpars}.
+The blue colored rows show constraints from bispectra with post-Born corrections.
+Post-Born corrections are not taken into account for the lensing power spectra as mentioned earlier, so these values are omitted in the blue rows.}
 \label{tab:paramconstraintstight}
@@
 Using ``stage 4'' noise with multipole range from $2$ to $2000$.
-The colored plots show CMB lensing power- and/or bispectrum constraints.
+The colored plots show CMB lensing power and/or bispectrum constraints.
 We use the weak priors listed in table \ref{tab:fiducialpars}.
@@
 They show approximate degeneracies in the information gained from a survey if they are ``stretched''.
-When the information of ellipses with degeneracies in different directions is combined, the degeneracies are removed and the constraints typically becomes much better on the relevant parameters.
+When the information of ellipses with degeneracies in different directions is combined, the degeneracies are removed and the constraints typically become much better on the relevant parameters.
 The black plots show the constraints using all lensing information, i.e.
-CMB lensing spectra, galaxy lensing spectram, and all cross spectra.}
+CMB lensing spectra, galaxy lensing spectra, and all cross spectra.}
     \label{fig:paramconstraintstightcmb}
@@
 \section{Discussion and conclusion}\label{sec:discussion}
-In this work, we have presented parameter forecasts from CMB lensing and galaxy lensing power- and bispectra, considering experimental parameters representative of Stage 3 and Stage 4 CMB and galaxy surveys.
+In this work, we have presented parameter forecasts from CMB lensing and galaxy lensing power and bispectra, considering experimental parameters representative of stage 3 and stage 4 CMB and galaxy surveys.
 Our analysis shows that while lensing power spectra are detectable at high significance, the inclusion of bispectra generally offers significant improvements when considering parameter constraints when considering an LCDM+ model of the universe.
-Notably, both CMB and galaxy bispectra are found to be detectable, even with Stage 3 experimental noise levels.
+Notably, both CMB and galaxy bispectra are found to be detectable, even with stage 3 experimental noise levels.
 
 The primary impact of incorporating bispectrum information is a tightening of cosmological parameter constraints, particularly for parameters such as $H_0$, $\sigma_8$, and $\Omega_m$, and in breaking key parameter degeneracies.
-We found that for sufficiently low noise levels, characteristic of Stage 4 surveys, the bispectra themselves can offer constraining power comparable to that of the power spectra, due to nonlinear structure formation on small scales being better detected by these future surveys.
+We found that for sufficiently low noise levels, characteristic of stage 4 surveys, the bispectra themselves can offer constraining power comparable to that of the power spectra, due to nonlinear structure formation on small scales being better detected by these future surveys.
 Furthermore, the combination of CMB and galaxy lensing probes, especially when both power spectra and bispectra are utilized, further enhances constraints and aids in mitigating degeneracies.
-However, lensing power spectra alone, even for Stage 4 surveys, appear insufficient to competitively constrain an 8-parameter $\Lambda\text{CDM} + w_0 + \sum m_\nu$ model without priors.
+If all information were to be combined\footnote{And using the CMB $T+E$ prior.} we find particularly tight constraints on neutrino mass and dark energy equation of state in particular, i.e.
+$\sigma(m_\nu) = 55$ meV and $\sigma(w_0) = 0.017$.
+However, lensing power spectra alone, even for stage 4 surveys, appear insufficient to competitively constrain an 8-parameter $\Lambda\text{CDM} + w_0 + \sum m_\nu$ model without any priors.
 
@@
 
-     \item We assumed throughout this paper that surveys would measure multipoles as low as $\ell = 2$.
+     \item We assumed throughout this paper that lensing surveys would use data from multipoles as low as $\ell = 2$.
 However, our analytical formulas to calculate lensing spectra utilize the Limber approximation, which is only robust for $\ell \gtrsim 50$.
-Avoiding the Limber approximation would complicate calculations, requiring knowledge of unequal-time matter power and bispectra, and would be computationally much more demanding due to the need to evaluate double integrals instead of single integrals.
-With the methods used in this paper, recalculating Fisher matrices without the Limber approximation would not be feasible, even if unequal-time matter spectra were readily available.
-Methods such as those explored in reference \cite{Chen:2021vba} might offer a path to address this.
-
-     \item As noted in appendix \ref{sec:fisher}, the Gaussian approximation was used for the covariance of the spectra estimators, i.e.
+Avoiding the Limber approximation would complicate calculations, requiring knowledge of unequal-time matter power and bispectra\footnote{Technically, this would be computationally much more demanding due to the need to evaluate double integrals instead of single integrals}.
+Moving beyond the Limber approximation can be achieved using methods such as those explored in reference \cite{Chen:2021vba} and might offer a path to address this.
+
+     \item As noted in appendix \ref{sec:fisher}, a Gaussian approximation was used for the covariance of the spectra estimators, i.e.
 we ignored any contributions not arising from standard Wick contractions.
-We know that the estimated lensing potential is not Gaussian in general, for two reasons: (i) The matter distribution is not Gaussian.
+We know that the estimated lensing potential is not Gaussian in general, for two reasons: (i) the matter distribution is not Gaussian.
 (ii) Even with a Gaussian matter distribution, the quadratic estimator we used will not be Gaussian.
@@
 Writing out each $\hat \phi$ in this way would yield terms such as $\langle TTTTTT\rangle$ (i.e.
-6 point functions).
+6-point functions).
 These terms are non-zero.
-In fact, not only would they give products of 2 point functions due to Wick contractions, but also connected $n$-point functions with $n>2$ because the lensed fields are no longer Gaussian even if the original fields are.
+In fact, not only would they give products of 2-point functions due to Wick contractions, but also connected $n$-point functions with $n>2$ because the lensed fields are no longer Gaussian even if the original fields are.
 A next step could thus be to account for these additional contributions, perhaps through N-body and ray-tracing simulations.
 Notably, such simulations could potentially also allow for a more accurate calculation of the lensing spectra themselves, without relying on fitting formulas for the matter bispectrum or the Limber approximation, depending on simulation accuracy.
-
-     \item Related to the previous point, for a completely correct analysis we would also need to take into account correlations between a CMB prior and the lensing power- and bispectra.
-Currently we simply add Fisher matrices, however taking these correlations into account means computing a covariance matrix for combinations of $T$, $E$, $\phi$ with nonzero entries for correlations between $T$/$E$ and $\phi$.
-For instance, consider the correlation between the estimator of the temperature powerspectrum and the lensing powerspectrum at arbitrary multipoles.
-We will get terms like $\langle TT\hat\phi\hat\phi \rangle$, which will be a sum over 6 point functions like $\langle TT TETE\rangle$ due to our quadratic estimator.
-These 6 point functions are clearly non-zero, as explained earlier.
+Obviously, such simulations are expensive.
+
+     \item Related to the previous point, for a completely correct analysis, we would also need to take into account correlations between a CMB prior and the lensing power and bispectra.
+Currently, we simply add Fisher matrices; however, taking these correlations into account means computing a covariance matrix for combinations of $T$, $E$, $\phi$ with nonzero entries for correlations between $T$/$E$ and $\phi$.
+For instance, consider the correlation between the estimator of the temperature power spectrum and the lensing power spectrum at arbitrary multipoles.
+We will get terms like $\langle TT\hat\phi\hat\phi \rangle$, which will be a sum over 6-point functions like $\langle TT TETE\rangle$ due to our quadratic estimator.
+These 6-point functions are clearly non-zero, as explained earlier.
+
+     \item Another consideration is that non-Gaussianity of the true lensing field doesn't only change the covariance matrices used to calculate Fisher matrix elements, but also introduces an additional bias to the estimated lensing powerspectrum, denoted as $N^{(3/2)}_\ell$ \cite{B_hm_2016, B_hm_2018}.
+This is because the weights in quadratic estimator from ref.
+\cite{Hu_2002} are calculated while assuming a gaussian lensing field.
+The bias is significant at the percent level for CMB lensing, however it may be larger for cross correlations of the CMB lensing field with low redshift mass tracers (such as galaxy lensing).
+Fitting $N^{(3/2)}_\ell$ to survey data will loosen constraints for the cosmological parameters.
+
+     \item Other than the lensing powerspectrum and bispectrum, there are also other statistics from the lensing field that can be used for parameter constraints.
+For example, ref.
+\cite{Liu_2016} showed that the lensing convergence one-point probability distribution function (PDF) and peak counts can tighten constraints in the $\Omega_m$ - $\sigma_8$ plane by $\sim30\%$ over just using the lensing powerspectrum for a stage 3 experiment\footnote{Specifically, for the AdvACT experiment.}.
+It is thus clear that future work could combine the power- and bispectrum with these additional statistics to further tighten forecasted constraints.
 
     \item The CMB lensing noise was estimated via a quadratic estimator.
-While standard, such estimators may not be strictly optimal, especially with complex foregrounds or non-Gaussianities.
-Estimators that, given our Stage 4 noise parameters, perform better do exist; for example, the iterative estimator developed in \cite{Smith_2012} can lead to a significantly better signal-to-noise ratio for the CMB lensing bispectrum, as shown in \cite{Namikawa_2016}.
+While standard, such estimators may not be strictly optimal, especially with complex foregrounds, non-Gaussianities and on small scales.
+Estimators that, given our stage 4 noise parameters, perform better do exist; for example, the iterative estimator developed in \cite{Smith_2012} can lead to a significantly better signal-to-noise ratio for the CMB lensing bispectrum, as shown in \cite{Namikawa_2016}.
 Using such an iterative estimator may also result in the CMB bispectrum being able to significantly improve parameter constraints when combined with lensing power spectra and a CMB $T+E$ prior.
@@
 
-    \item A 2024 paper \cite{postborn_pratten_lewis} has shown that the Born approximation used to derive the formulas for the lensing power and bispectra, is no longer negligible for the lensing bispectrum.
-The post-Born corrections calculated in that paper only changed the lensing power spectra by $\leq 0.2\%$, while for the bispectra this effect was much larger, in particular for folded configurations.
-For future work it would thus be worthwile to analyze how these post-Born corrections affect the results presented in this paper.
-
     \item We did not model foreground contamination such as the cosmic infrared background (CIB), thermal Sunyaev-Zel'dovich (tSZ) effect, or point sources.
 These foregrounds can bias lensing reconstructions and bispectra, particularly in CMB–galaxy cross-correlations, and should be accounted for in future analyses.
+Baryonic feedback was also not accounted for.
+Given that future surveys will probe the matter distribution on smaller scales, this will become increasingly relevant \cite{McCarthy_2022}, particularly for estimating neutrino mass \cite{PhysRevD.101.063534}.
+Bias due to baryonic feedback can be mitigated, e.g.
+by using the Principal Component Analysis method \cite{s2024impactcosmologydependencebaryonic}, discarding small scale $T$-modes, or fitting additional parameters that capture the effects of baryonic feedback \cite{McCarthy_2022}.
+Some of these methods may be able to mitigate baryonic feedback effects without increasing the uncertainty of parameter constraints \cite{McCarthy_2021}.
+Finally, there are still other well-known sources of uncertainty that we did not take into account, such as photometric redshift errors for galaxies or instrumental systematics.
 
@@
 \section*{Acknowledgements}
-We would like to thank Prof.
-Toshiya Namikawa for providing extra details on his paper on the CMB lensing bispectrum \cite{Namikawa_2016}, which, in particular, allowed us to cross check (intermediate) results.
-The authors also like to thank Luna Jansma, who wrote a MSc thesis on the same subject, which served as the foundation of this study.
+We would like to thank Toshiya Namikawa for providing extra details on his paper on the CMB lensing bispectrum \cite{Namikawa_2016}, which, in particular, allowed us to cross-check (intermediate) results.
+We would also like to thank Antony Lewis for confirming the correctness of the formulas for post-Born corrections.
+The authors would like to thank Alex van Engelen for detailed comments and suggestions to improve the manuscript.
+Finally, the authors thank Luna Jansma, who wrote an MSc thesis on the same subject, which served as the foundation of this study.
 
@@
 \end{equation}
-In the weak lensing regime and under the Born approximation, $A$ is a symmetric matrix by definition and $\omega$ vanishes, we will ignore it from here on out.
+In the weak lensing regime and under the Born approximation, $A$ is a symmetric matrix by definition, and $\omega$ vanishes; we will ignore it from here on out.
 Intuitively, $A$ tells you how a small patch in the sky transforms due to lensing.
@@
 \end{gather}
-Now define the power spectrum as
+Now, define the power spectrum as
 \begin{equation}
@@
 \end{equation}
-In combination with the Limber approximation we thus find
+In combination with the Limber approximation, we thus find
 \begin{equation}
@@
 \end{gather*}
-Now using equation \ref{eq:complexexp} to rewrite the complex exponential:
+Now using equation \eqref{eq:complexexp} to rewrite the complex exponential:
 %TODO: check \ell m here
@@
 When all $m_i$ equal zero, the Wigner 3j-symbol gains a number of useful properties % TODO: discuss all of them
-In particular, it vanishes if $\ell_1+\ell_2+\ell_3$ is odd, meaning we can drop the $(-1)^{\ell_1+\ell_2+\ell_3}$ factor. Additionally cancelling common factors then gives
+In particular, it vanishes if $\ell_1+\ell_2+\ell_3$ is odd, meaning we can drop the $(-1)^{\ell_1+\ell_2+\ell_3}$ factor. Additionally, cancelling common factors then gives
 \begin{align}
@@
 % TODO: fix references and citations 
-We can rewrite equations \ref{eq:lpsintermsofpsi} and \ref{eq:lbsintermsofpsi} in terms of the matter spectra instead of the $\psi$ spectra using the poisson equation.
+We can rewrite equations \ref{eq:lpsintermsofpsi} and \ref{eq:lbsintermsofpsi} in terms of the matter spectra instead of the $\psi$ spectra using the Poisson equation.
 This allows us to numerically evaluate these lensing spectra using CAMB.
@@
 \end{equation}
-and the matter spectra are defined in terms of the fourier transformed density contrast $\delta(\mathbf k)$ as 
+and the matter spectra are defined in terms of the Fourier transformed density contrast $\delta(\mathbf k)$ as 
 \begin{gather*}
@@
 $$
-where $a(\eta)$ is the only time dependent factor on the right hand side.
-The poisson equation relates $\Psi$ to the density contrast as \cite{dodelson2020modern}
+where $a(\eta)$ is the only time-dependent factor on the right-hand side.
+The Poisson equation relates $\Psi$ to the density contrast as \cite{dodelson2020modern}
 % in modern cosm eq 6.41 they say - nabla^2 Phi = ..., however here we use Psi instead and Psi = - Phi in our regime so hence why we don't take the - sign
@@
 \end{equation}
-where $\Psi(k,\eta)$ and $\delta(k,\eta)$ are functions in Fourier space. For the power- and bispectra we find
+where $\Psi(k,\eta)$ and $\delta(k,\eta)$ are functions in Fourier space. For the power- and bispectra, we find
 \begin{gather*}
@@
 \section{Fisher Matrix Analysis} \label{sec:fisher}
-\subsection{Determing uncertainty in experimental parameters}
+\subsection{Determining uncertainty in experimental parameters}
 
 The Fisher matrix formalism is used to find a lower bound on the constraints we can place on experimental parameters.
-It combines the Cramer-Rao Inequality \cite{casella2002statistical} with the assumption that we have unbiased estimators  following a gaussian distribution \cite{dodelson2020modern}.
+It combines the Cramer-Rao Inequality \cite{casella2002statistical} with the assumption that we have unbiased estimators following a Gaussian distribution \cite{dodelson2020modern}.
 In particular, denoting the Fisher matrix by $F_{\theta_i\theta_j}$, the parameters as $\theta_i$, and their estimators as $\hat\theta_i$, it can be shown that
@@
 where Cov is the covariance matrix associated with the random vector $(x_1, ..., x_n)$, $\text{Cov}_{x_px_q}:=\text{Cov}(x_p,x_q)$.
-The derivative of the mean measurement outcomes and measurement covariances will in general depend on the true, unknown, values of the experimental parameters.
-We therefore evaluate these quantities with for our best guess of the experimental parameters given some outside information.
+The derivative of the mean measurement outcomes and measurement covariances will in general, depend on the true, unknown values of the experimental parameters.
+We therefore evaluate these quantities for our best guess of the experimental parameters given some outside information.
 These are known as the ``fiducial'' values.
@@
 \end{equation*}
-Here the covariance matrix is given as
+Here, the covariance matrix is given as
 \begin{equation*}
@@
 $\hat X(\ell) \hat Y(\ell) = \hat Y(\ell) \hat X(\ell)$.
-In fact, if we were to count these permutations as distinct signals we would get identical columns and/or rows in our covariance matrix making inversion impossible:
+In fact, if we were to count these permutations as distinct signals, we would get identical columns and/or rows in our covariance matrix, making inversion impossible:
 \begin{equation*}
@@
 
-To evaluate the covariance matrix we again assume that the estimators are Gaussian distributed so that we can apply a wick contraction, as is commonly done \cite{tegmark1997cosmic}.
+To evaluate the covariance matrix, we again assume that the estimators are Gaussian distributed so that we can apply a Wick contraction, as is commonly done \cite{tegmark1997cosmic}.
 In this case we get
@@
 This results in the $(2\ell+1)^{-1}$ factor in the power spectrum covariance.
-    \item The tilde is used to denote the power spectrum as calculated earlier plus the noise power spectrum, $N^{XX'}_\ell$, which is the power spectrum of the noise associated with the estimator of the field.
+    \item The tilde is used to denote the power spectrum as calculated earlier, plus the noise power spectrum, $N^{XX'}_\ell$, which is the power spectrum of the noise associated with the estimator of the field.
 This is where experimental noise is incorporated into the calculation.
 \end{enumerate}
-Under the Gaussian approximation, the covariance matrix vanishes except for $3\times 3$ block matrices (in the case of 2 tracers) on the diagonal. The fisher matrix is then
+Under the Gaussian approximation, the covariance matrix vanishes except for $3\times 3$ block matrices (in the case of 2 tracers) on the diagonal. The Fisher matrix is then
 \begin{equation*}
@@
 Counting only distinct signals requires more care compared to the power spectra.
-The rule is that $B^{XYZ}_{\ell_1\ell_2\ell_3}$ and $B^{X'Y'Z'}_{\ell_1'\ell_2'\ell_3'}$ are not distinct signals if there exists a permutation $\sigma$ that simultaniously maps $X'Y'Z'$ to $XYZ$ and $\ell_1'\ell_2'\ell_3'$ to $\ell_1\ell_2\ell_3$.
+The rule is that $B^{XYZ}_{\ell_1\ell_2\ell_3}$ and $B^{X'Y'Z'}_{\ell_1'\ell_2'\ell_3'}$ are not distinct signals if there exists a permutation $\sigma$ that simultaneously maps $X'Y'Z'$ to $XYZ$ and $\ell_1'\ell_2'\ell_3'$ to $\ell_1\ell_2\ell_3$.
 It turns out that we can write a sum over distinct signals explicitly as
@@
 It follows to show that the sets that these sums sum over form a partition of the set of all distinct signals.
-Clearly all 3 sets are pairwise disjoint (no common elements) because of the criteria for the $l_i$'s.
+Clearly, all 3 sets are pairwise disjoint (no common elements) because of the criteria for the $l_i$'s.
 To show that their union covers the set of distinct signals, start by considering an arbitrary signal.
 Its $l$ configuration will trivially correspond to exactly one of the three sums.
-If it is sum 1, then we are free to permute the $XYZ$'s by virtue of the $l$'s being identical so we will be able to match the $XYZ$ configuration to one of the elements of $\{\{XYZ\}\}$.
+If it is sum 1, then we are free to permute the $XYZ$'s by virtue of the $l$'s being identical, so we will be able to match the $XYZ$ configuration to one of the elements of $\{\{XYZ\}\}$.
 Similarly, if the $l$ configuration corresponds to sum 2, then we are free to permute the $XY$ configuration to match with one of the elements of $\{\{XY\}Z\}$.
 The $Z$ value does not matter because any $Z$ value is accounted for.
-For sum 3 we can argue that we can switch around the order of the $l$'s to satisfy $\ell_1 < l2 < l3$ and the corresponding $XYZ$ configuration will be accounted for in sum 3 because all $XYZ$ combinations are counted.
+For sum 3, we can argue that we can switch around the order of the $l$'s to satisfy $\ell_1 < l2 < l3$ and the corresponding $XYZ$ configuration will be accounted for in sum 3 because all $XYZ$ combinations are counted.
 Finally, it is simple to verify that no distinct signal is counted more than once within each sum.
 
-To calculate the elements of the covariance matrix consider the following.
+To calculate the elements of the covariance matrix, consider the following.
 Every element of the Fisher matrix can be seen as an inner product weighted by the inverse covariance matrix.
 We can choose how we order the vectors\footnote{The entries are the derivatives of the bispectra}.
-We organize the vectors according to the sum 1, 2, and 3 parts first.
+We organize the vectors according to the sum of 1, 2, and 3 parts first.
 Then by $l$ configuration.
-Within each $l$ configuration we can choose any ordering for the $XYZ$ configurations.
+Within each $l$ configuration, we can choose any ordering for the $XYZ$ configurations.
 The covariance matrix now becomes a block matrix with each block corresponding to an $l_i$ and $l_i'$ configuration.
-When wick contracting using the gaussian approximation, every block matrix where $(\ell_1,\ell_2,\ell_3)\neq (\ell_1',\ell_2',\ell_3')$ vanishes.
+When wick contracting using the Gaussian approximation, every block matrix where $(\ell_1,\ell_2,\ell_3)\neq (\ell_1',\ell_2',\ell_3')$ vanishes.
 It can then be shown that the entries of each block matrix are given as
@@
 \end{gather*}
-With our ordering this means that the covariance matrix is again a diagonal block matrix, now with blocks of size $4 \times 4$ (sum 1), $6 \times 6$ (sum 2), and $8 \times 8$ (sum 3).
+With our ordering, this means that the covariance matrix is again a diagonal block matrix, now with blocks of size $4 \times 4$ (sum 1), $6 \times 6$ (sum 2), and $8 \times 8$ (sum 3).
 
@@
 
-To show that the above is the same as the formula for the Fisher matrix given earlier, first parition the sum in the same way and collect all terms that fit in the different categories.
+To show that the above is the same as the formula for the Fisher matrix given earlier, first partition the sum in the same way and collect all terms that fit in the different categories.
 \begin{align*}
@@
 The sum over the different wick contractions will similarly cancel with the $\mathcal P_{\ell_1\ell_2\ell_3} / 6$ factor for the $\ell_1 = \ell_2 \neq \ell_3$ sum.
-For the $\ell_1 < \ell_2 < \ell_3$ sum the proof is similar as well except no cancellation is required.
+For the $\ell_1 < \ell_2 < \ell_3$ sum, the proof is similar as well, except no cancellation is required.
 
@@
 \subsection{Signal to Noise Ratio (SNR)}
-To quantify the detectability of the lensing spectra, we introduce an overal amplitude of our signal, $A$, with fiducial value 1 as experimental parameter and compute $F_{AA}$.
+To quantify the detectability of the lensing spectra, we introduce an overall amplitude of our signal, $A$, with fiducial value 1 as an experimental parameter and compute $F_{AA}$.
 Obviously, $\partial_A\br{A B^{XYZ}_{\ell_1\ell_2\ell_3}}|_{A=1} = B^{XYZ}_{\ell_1\ell_2\ell_3}$, so we find
@@
 \subsection{Fisher matrix of power- + bispectra}
-To compute the Fisher matrix of an experiment measuring both lensing power- and bispectra we are required to compute and invert the full covariance matrix.
-If we keep assuming that the measurements are close enough to Gaussian to be able to use wick contractions to a good approximation, the full covariance matrix simplifies trivially.
-The correlation between a power- and bispectrum estimator contains an odd (5) amount of fields and thus always vanishes.
-We are thus allowed to simply add the Fisher matrices of the power- and bispectra to compute the combined Fisher matrix.
+To compute the Fisher matrix of an experiment measuring both lensing power and bispectra, we are required to compute and invert the full covariance matrix.
+If we keep assuming that the measurements are close enough to Gaussian to be able to use Wick contractions to a good approximation, the full covariance matrix simplifies trivially.
+The correlation between a power and bispectrum estimator contains an odd (5) amount of fields and thus always vanishes.
+We are thus allowed to simply add the Fisher matrices of the power and bispectra to compute the combined Fisher matrix.
 
@@
 Each change of the cosmological parameters requires a recalculation of the entire cosmology, making it computationally expensive.
-For the approximation to be accurate a balance needs to be found between numerical errors for small $h$ and a larger $\mathcal O(h^2)$ error for larger $h$.
+For the approximation to be accurate, a balance needs to be found between numerical errors for small $h$ and a larger $\mathcal O(h^2)$ error for larger $h$.
 The $h$ values chosen for each parameter are shown in table \ref{tab:cosmo-params-diff} and are similar to the values used in \cite{Namikawa_2016}\footnote{As confirmed during a conversation with the author.}.
@@
 
-To test the accuracy, we varied $h$ by $\pm 5\%$ and $\pm 10\%$ and compared the relative change in the derivative of the equilateral lensing bispectra and the lensing powerspectra.
+To test the accuracy, we varied $h$ by $\pm 5\%$ and $\pm 10\%$ and compared the relative change in the derivative of the equilateral lensing bispectra and the lensing power spectra.
 As long as numerical noise doesn't dominate, it can be shown that the relative error in our approximation is approximately 5 times the relative difference that taking $h \rightarrow h(1 \pm 0.1)$ leads to.
@@
 % \end{figure}
-
+\pagebreak
 \section{$\Lambda$CDM constraints}
@@
 We do not expect to use weak lensing surveys to competitively constrain the main $\Lambda$CDM parameters.
-Despite this, we include these constraints for completeness in figures \ref{fig:paramconstraintsallcmb} and \ref{fig:paramconstraintsallgal} for CMB and galaxy lensing, respectively, and in table \ref{tab:paramconstraintsall}.
-CMB and galaxy lensing respectively.
+Despite this, we include these constraints for completeness in figures \ref{fig:paramconstraintsallcmb}, \ref{fig:paramconstraintsallcmbcmbprior}, \ref{fig:paramconstraintsallgal}, \ref{fig:paramconstraintsallgalcmbprior}, and in table \ref{tab:paramconstraintsall}.
 All survey parameters are the same as in section \ref{sec:results}.
@@
     \hline
-        && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
-        \hline
-        Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
-        \hline
-        $H_0$          &     1.73e+01 &  1.65e+01 &  1.65e+01 &        6.91e+00 & 6.39e+00 & 5.04e+00 &      1.65e+00 &      4.13e+00 &      2.88e+00 &             6.65e-01 \\
-        $\Omega_b h^2$ &     5.00e-04 &  5.00e-04 &  5.00e-04 &        5.00e-04 & 5.00e-04 & 4.98e-04 &      4.95e-04 &      4.86e-04 &      4.94e-04 &             4.65e-04 \\
-        $\Omega_c h^2$ &     2.89e-01 &  6.99e-03 &  1.35e-02 &        6.42e-03 & 5.22e-03 & 4.74e-03 &      2.04e-03 &      1.07e-03 &      2.63e-03 &             8.62e-04 \\
-        $n_s$          &     2.00e-02 &  1.90e-02 &  1.99e-02 &        1.82e-02 & 8.04e-03 & 1.89e-02 &      5.54e-03 &      2.98e-03 &      1.52e-02 &             2.34e-03 \\
-        $\tau$         &     6.30e-02 &  4.69e-04 &  6.30e-02 &        4.46e-04 & 6.30e-02 & 6.30e-02 &      6.30e-02 &      5.20e-05 &      6.30e-02 &             5.07e-05 \\
-        $A_s$          &     1.00e-09 &  3.64e-10 &  9.29e-10 &        3.07e-10 & 2.56e-10 & 1.16e-10 &      7.27e-11 &      8.39e-11 &      1.10e-10 &             4.51e-11 \\
+    && \multicolumn{3}{c|}{CMB lensing} & \multicolumn{3}{c|}{Gal. lensing} & \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
+    \hline
+    Par & prior & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_2}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
+    \hline
+$H_0$              & 17    & 17    & 16    & 6.5   & 13    & 4.2   & 1.8   & 1.6   & 1.5   & 1.0   \\
+$10^3 \Omega_b h^2$ & 0.50  & 0.50  & 0.50  & 0.50  & 0.50  & 0.48  & 0.48  & 0.50  & 0.48  & 0.46  \\
+$\Omega_c h^2$     & 0.29  & 0.0070& 0.012 & 0.0063& 0.0095& 0.0049& 0.0041& 0.0050& 0.0044& 0.0033\\
+$n_s$              & 0.020 & 0.019 & 0.020 & 0.018 & 0.020 & 0.017 & 0.011 & 0.015 & 0.016 & 0.010 \\
+$\tau$             & 0.063 & 0.063 & 0.063 & 0.063 & 0.063 & 0.063 & 0.063 & 0.063 & 0.063 & 0.063 \\
+$10^9 A_s$         & 1.0   & 0.36  & 0.91  & 0.30  & 0.53  & 0.10  & 0.10  & 0.081 & 0.098 & 0.063 \\
     \hline
@@
     \hline
-        && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
-        \hline
-         Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
-        \hline
-        $H_0$          &    1.21e+00 &  1.00e+00 &  1.18e+00 &        1.00e+00 & 1.07e+00 & 9.93e-01 &      7.20e-01 &      9.32e-01 &      9.68e-01 &             5.17e-01 \\
-        $\Omega_b h^2$ &    5.69e-05 &  5.16e-05 &  5.67e-05 &        5.15e-05 & 5.49e-05 & 5.29e-05 &      5.18e-05 &      4.74e-05 &      5.15e-05 &             4.69e-05 \\
-        $\Omega_c h^2$ &    8.27e-04 &  6.08e-04 &  8.15e-04 &        5.99e-04 & 5.89e-04 & 6.57e-04 &      5.27e-04 &      1.63e-04 &      6.03e-04 &             1.54e-04 \\
-        $n_s$          &    2.50e-03 &  2.17e-03 &  2.50e-03 &        2.15e-03 & 2.12e-03 & 2.42e-03 &      2.11e-03 &      1.44e-03 &      2.33e-03 &             1.38e-03 \\
-        $\tau$         &    1.34e-02 &  4.37e-05 &  1.22e-02 &        4.33e-05 & 1.08e-02 & 1.13e-02 &      9.73e-03 &      2.94e-05 &      1.08e-02 &             2.82e-05 \\
-        $A_s$          &    5.45e-11 &  4.82e-12 &  4.94e-11 &        4.77e-12 & 4.39e-11 & 4.58e-11 &      3.92e-11 &      2.83e-12 &      4.32e-11 &             2.66e-12 \\
+    && \multicolumn{3}{c|}{CMB lensing} & \multicolumn{3}{c|}{Gal. lensing} & \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
+    \hline
+Par & prior & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_2}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
+\hline
+$H_0$              & 1.2   & 1.1   & 1.2   & 1.1   & 1.1   & 0.90  & 0.66  & 0.84  & 0.78  & 0.43  \\
+$10^3 \Omega_b h^2$ & 0.057 & 0.053 & 0.057 & 0.053 & 0.056 & 0.053 & 0.052 & 0.051 & 0.052 & 0.050 \\
+$\Omega_c h^2$     & 0.00083& 0.00076& 0.00081& 0.00076& 0.00064& 0.00068& 0.00053& 0.00048& 0.00061& 0.00043 \\
+$n_s$              & 0.0025& 0.0025& 0.0025& 0.0025& 0.0022& 0.0024& 0.0022& 0.0020& 0.0022& 0.0018 \\
+$\tau$             & 0.013 & 0.011 & 0.012 & 0.011 & 0.011 & 0.010 & 0.0093& 0.0091& 0.0098& 0.0082 \\
+$10^9 A_s$         & 0.055 & 0.044 & 0.049 & 0.043 & 0.045 & 0.041 & 0.037 & 0.036 & 0.039 & 0.032 \\
 \hline
@@
     \centering
-    \includegraphics[width=\textwidth]{figures/param_constraints_all_cmb.pdf}
+    \includegraphics[width=\textwidth]{paper/figures/param_constraints_lcdm_cmb_weak_prior.pdf}
     \caption{Same as figure \ref{fig:paramconstraintstightcmb}, except for the standard $\Lambda$CDM parameters.}
@@
 \end{figure}
+
 \begin{figure}[h!]
     \centering
-    \includegraphics[width=\textwidth]{figures/param_constraints_all_gal.pdf}
+    \includegraphics[width=\textwidth]{paper/figures/param_constraints_lcdm_cmb_cmb_prior.pdf}
+    \caption{Same as figure \ref{fig:paramconstraintstightcmbcmbprior}, except for the standard $\Lambda$CDM parameters.}
+    \label{fig:paramconstraintsallcmbcmbprior}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{paper/figures/param_constraints_lcdm_gal_weak_prior.pdf}
     \caption{Same as figure \ref{fig:paramconstraintstightgal}, except for the standard $\Lambda$CDM parameters.}
     \label{fig:paramconstraintsallgal}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{paper/figures/param_constraints_lcdm_gal_cmb_prior.pdf}
+    \caption{Same as figure \ref{fig:paramconstraintstightgalcmbprior}, except for the standard $\Lambda$CDM parameters.}
+    \label{fig:paramconstraintsallgalcmbprior}
 \end{figure}

### 2025-07-11 code/lps_vs_noise_plot.py
figure cahnges

@@ -2,94 +2,75 @@ import cosm_setup as cs
 import matplotlib.pyplot as plt
 import numpy as np
 
-fig, axs = plt.subplots(1, 2, figsize = (8.3, 4))
+fig, axs = plt.subplots(1, 2, figsize=(8.3, 4))
 
 # LENSING SPECTRA
 l_values = np.logspace(np.log10(2), np.log10(3000), 100)
 spectra = cs.lensing_spectra()
-gal_lps = np.array([spectra.lps(l, ('s', 's')) for l in l_values])  # Shear-shear power spectrum
+gal_lps = np.array([spectra.lps(l, ('s', 's')) for l in l_values])
 cmb_lps = np.array([spectra.lps(l, ('c', 'c')) for l in l_values])
 
 # CMB NOISE QUAD ESTIMATOR
-# original noise is of deflection signal, so divide by l(l+1) to convert to lens potential
 ls_cmbn = np.loadtxt('cmb_noise_files/ls_1_3000_64.txt')
-cmbn_301 = np.abs(np.loadtxt('cmb_noise_files/Ns_sigma3_DeltaT0.71_DeltaP1.txt')) / (ls_cmbn * (ls_cmbn + 1)) # S4
-cmbn_106 = np.abs(np.loadtxt('cmb_noise_files/Ns_sigma1_DeltaT4.2_DeltaP6.txt')) / (ls_cmbn * (ls_cmbn + 1)) # S3
-
-# to cross check with paper
-cmbn_41sqrt2 = np.abs(np.loadtxt('cmb_noise_files/Ns_sigma4_DeltaT1_DeltaP1.41.txt')) / (ls_cmbn * (ls_cmbn + 1)) 
-
+cmbn_301 = np.abs(np.loadtxt('cmb_noise_files/Ns_sigma3_DeltaT0.71_DeltaP1.txt')) / (ls_cmbn * (ls_cmbn + 1))
+cmbn_106 = np.abs(np.loadtxt('cmb_noise_files/Ns_sigma1_DeltaT4.2_DeltaP6.txt')) / (ls_cmbn * (ls_cmbn + 1))
 cmbn_TT = np.abs(np.loadtxt('cmb_noise_files/Ns_sigma4_DeltaT1_DeltaP1.41_TT.txt')) / (ls_cmbn * (ls_cmbn + 1))
[…]

### 2025-07-11 paper/playing_with_fish.py

@@ -455,7 +455,7 @@ if __name__ == "__main__":
     # 0.5: for the table, strong prior, with pb corrections
     # 1.5: for the table, weak prior, with pb corrections
 
-    fish_pond_number = 1
+    fish_pond_number = 9
     f_sky = 0.5
     if fish_pond_number == 0:
         fish_matrices = [
@@ -640,7 +640,7 @@ if __name__ == "__main__":
         labels=labels,
         colors=["tab:blue", "tab:orange", 'tab:green', 'black'],
         nstd=1.0,
-        figsize=(8.3,8.3),
+        figsize=(8,8),
         dpi=100
     )
 
@@ -650,7 +650,7 @@ if __name__ == "__main__":
     print(f'Fish pond number: {fish_pond_number}')
     print(f'parameters used: {keep_indices}')
     
-    #plt.savefig('paper/figures/' + plt_name, dpi = 300)
-    #print('created:', plt_name)
+    plt.savefig('paper/figures/' + plt_name, dpi = 300)
+    print('created:', plt_name)
 
-    save_table(param_names_latex_kept, param_values_kept, which_pars=keep_indices,constraints=cov_matrices, labels=labels)
+    #save_table(param_names_latex_kept, param_values_kept, which_pars=keep_indices,constraints=cov_matrices, labels=labels)

### 2025-07-25 paper/playing_with_fish.py
plots smaller

@@ -183,9 +183,9 @@ sigmaomm_ders = [
     (0.25 * sigma8 * omm**(-0.75))]
 
 
-# keep_indices = [4, 7, 8, 10]
+keep_indices = [4, 7, 8, 10]
 #keep_indices=[10]
-keep_indices = [0, 1, 2, 3, 5, 6] # lcdm pars
+# keep_indices = [0, 1, 2, 3, 5, 6] # lcdm pars
 #keep_indices = [4, 6]
 
 def process_fishes(fishes, keep_indices, derived_param_derivss = [s8_ders, S8_ders, omm_ders, sigmaomm_ders]):
@@ -455,7 +455,7 @@ if __name__ == "__main__":
     # 0.5: for the table, strong prior, with pb corrections
     # 1.5: for the table, weak prior, with pb corrections
 
-    fish_pond_number = 9
+    fish_pond_number = 5
     f_sky = 0.5
     if fish_pond_number == 0:
         fish_matrices = [
@@ -640,7 +640,7 @@ if __name__ == "__main__":
         labels=labels,
         colors=["tab:blue", "tab:orange", 'tab:green', 'black'],
         nstd=1.0,
-        figsize=(8,8),
+        figsize=(4.15, 4.15),
         dpi=100
     )
 

### 2025-07-25 paper/paper_draft.tex

@@
 
-\usepackage[backend=bibtex]{biblatex}
+\usepackage[backend=bibtex, sorting=none]{biblatex}
 \addbibresource{sources.bib}
@@
 \abstract{
-Upcoming stage 4 surveys, such as the Simons Observatory, LSST, and Euclid are poised to measure weak gravitational lensing of the Cosmic Microwave Background (CMB) and galaxies with unprecedented precision.
+Upcoming stage 4 surveys, such as the Simons Observatory, LSST, and \textit{Euclid}, are poised to measure weak gravitational lensing of the Cosmic Microwave Background (CMB) and galaxies with unprecedented accuracy and precision.
 While the power spectrum is the standard statistic used to analyze weak lensing data, non-Gaussianity induced by the non-linear growth of structure imparts significant cosmological information into higher-order statistics.
@@
 Our analysis shows that, in the absence of systematics, both the CMB and galaxy lensing bispectra will be detectable at high signal-to-noise.
-We consider parameter constraints with a ``weak'' prior and with a stronger prior from the primary CMB anisotropies.
-In the weak prior case the lensing bispectrum substantially improves parameter constraints over using the lensing power spectrum alone, primarily by breaking parameter degeneracies.
-For the stronger prior we see less significant gains, particularly for the CMB lensing bispectrum.
-We further find that for stage 4 surveys that are able to measure lensing on small scales where non-linear structure formation becomes important, the constraining power of the lensing bispectrum can become comparable to that of the lensing power spectrum.
+We consider two different choices of prior: one ``strong'' as could be expected from external information from a primary CMB experiment; and one ``weak'' containing minimal information.
+In the weak prior case, the lensing bispectrum substantially improves parameter constraints over the lensing power spectrum alone, primarily by breaking parameter degeneracies.
+For the stronger prior, we find less significant gains, particularly for the CMB lensing bispectrum.
+For stage 4 surveys that can measure lensing on small scales where non-linear structure formation becomes important, the constraining power of the lensing bispectrum can become comparable to that of the lensing power spectrum.
 Furthermore, we demonstrate that, both for our weak and strong priors, the synergy between CMB and galaxy lensing is substantial.
-Combining the two probes, particularly with both their lensing power- and bi-spectra, leads to significantly tighter constraints than achievable with either probe individually, with the greatest improvement seen for the sum of neutrino masses.
+Combining the two probes, particularly with both their lensing power- and bispectra, leads to significantly tighter constraints achievable with either probe individually, with the greatest improvement seen for the sum of neutrino masses.
 Additionally, we also analyze the changes in detectability and parameter constraints when taking into account post-Born corrections to the lensing bispectra, as these were shown to be significant, especially for the CMB lensing bispectrum \cite{postborn_pratten_lewis}.
-We show that the conclusions of this paper remain largely the same with this correction.
+Despite this correction, we show that the conclusions of this analysis remain largely the same.
+With these findings, we have established the potential of higher-order lensing statistics for cosmology.
+They should serve as a motivation for further studies that are required to address several things we have neglected here, such as non-Gaussian covariance, instrumental systematics, and baryonic feedback.
 }
@@
 
-Weak lensing has been used as a probe to constrain cosmological parameters for over a decade \cite{kilbinger2015cosmology,Bacon2000,Kaiser2000,Waerbeke2000,Bartelmann2001,Planck2018Lensing}.
-A range of upcoming surveys, such as the survey by the Simons Observatory (SO) \cite{Ade2019}, the Legacy Survey of Space and Time (LSST) \cite{Ivezic2019}, and Euclid \cite{Laureijs2011} aims to measure weak lensing of the cosmic microwave background (CMB) and galaxies respectively.
+Weak lensing, first detected in 2007 for radiation from the CMB \cite{Smith2007} and in 2000 for that of galaxies \cite{Bacon2000, Kaiser2000, Wittman_2000, vanwaerbeke2000detectioncorrelatedgalaxyellipticities}, has been used as a probe to constrain cosmological parameters for over a decade \cite{kilbinger2015cosmology, Waerbeke2000,Bartelmann2001,Planck2018Lensing}.
+A range of upcoming surveys, such as the Simons Observatory (SO) \cite{Ade2019}, the Legacy Survey of Space and Time (LSST) \cite{Ivezic2019}, and \textit{Euclid} \cite{Laureijs2011} aims to measure weak lensing of the cosmic microwave background (CMB) and galaxies respectively.
 These upcoming surveys are expected to achieve significantly higher accuracy than their predecessors, leading to tighter constraints on cosmological parameters and advancing our overall understanding of cosmology.
-Typically, the standard summary statistic of interest is the power spectrum (or equivalently, two-point function) of the lensing potential (as well as cross-correlations with galaxy clustering).
+Typically, the standard summary statistic of interest is the power spectrum (or equivalently, the two-point function) of the lensing potential (as well as cross-correlations with galaxy clustering).
 This can be measured in terms of the lensing convergence $\kappa$, lensing shear $\gamma$, or lensing potential $\psi$.
@@
 
-Even with Gaussian initial conditions, there are significant amounts of non-Gaussianity in the weak lensing signal, especially in galaxy lensing due to it being a probe of the more recent distribution of matter \cite{Bernardeau1997,Takada2003}.
-In addition to looking at the power spectrum, a natural next step is to measure weak lensing bispectra.
-While the bispectrum has been measured for galaxy lensing signals \cite{vanWaerbeke2002}, a definitive detection of the CMB lensing bispectrum is still pending \cite{Kalaja_2023,Namikawa_2016}.
-Upcoming surveys, such as the ones mentioned earlier, may be able to detect this signal.
+Even with Gaussian initial conditions, there are significant amounts of non-Gaussianity in the weak lensing signal, especially in galaxy lensing due to the fact that lensing indirectly traces the more recent distribution of density \cite{Bernardeau1997,Takada2003}.
+Higher order spectra, such as the bispectrum, have already been shown to be useful for constraining cosmological parameters and breaking parameter degeneracies.
+A first detection of non-Gaussian clustering using lensing has recently been achieved by cross correlating the matter overdensity squared with CMB lensing convergence \cite{Farren:2023yna, Harscouet:2025ksm}.
+In addition to looking at power spectra, a natural next step is to measure weak lensing bispectra.
+Both the CMB and galaxy lensing power spectra are detectable with current surveys \cite{Das2011, Planck2013Lensing, Planck2015Lensing, Planck2018Lensing}.
+On the other hand, while the bispectrum has been measured for galaxy lensing signals \cite{Simon_2013}, a detection of the CMB lensing bispectrum is still pending \cite{Kalaja_2023, Namikawa_2016}.
+Upcoming surveys, such as the ones mentioned earlier, may be able to detect this higher order spectrum.
 Therefore it is timely to consider the potential of higher-order spectra to constrain cosmological parameters.
 
-Additionally, we will also look at the effects of post-Born corrections.
+In our analysis we will include post-Born corrections.
 The CMB and galaxy lensing power spectra are typically affected by less than $1\%$ \cite{postborn_pratten_lewis, galpostbornlpscorr}.
 The galaxy lensing bispectrum is only affected by a few percent at most \cite{galpostbornlbscorr}.
-The only significant effect is on the CMB lensing bispectrum, as was shown in a 2024 paper \cite{postborn_pratten_lewis}.
+The only significant effect is on the CMB lensing bispectrum, as was shown in Ref.~ \cite{postborn_pratten_lewis}.
 In particular, for the folded configuration the effect can be of the same magnitude as the lensing bispectrum due to nonlinear large scale structure \eqref{eq:lensingbispectrum}.
-It is thus worth considering how including first order corrections changes the detectability of the CMB lensing bispectrum and the associated parameter constraints.
-This analysis will shed light on whether or not the the Born approximation is still appropriate to use if one were to calculate and use the CMB lensing bispectrum using future data.
-
-In this paper, we aim to answer these questions by adopting experimental parameters similar to those of next-generation surveys.
+Including post-Born effects  will allow us to determine how future parameter constraints are by more accurate modelling.
+
+In this paper, we aim to answer these questions by adopting experimental configurations similar to those of next-generation surveys.
 We are especially interested in seeing if approximate parameter degeneracies can be broken by combining CMB and galaxy weak lensing power- and bi-spectra.
@@
 Useful derivations relevant to this paper can be found in the appendices.
-In particular, they include a derivation for the Fisher matrix formalism for the non-trivial context of bispectra with multiple tracers.
 
@@
 A derivation can be found in appendices \ref{sec:weaklensing} and \ref{sec:weaklensstats}, or in the literature; see, e.g.
-\cite{bartelmann2001weak}.
-$\Omega_m$ is the present day matter density parameter.
-$H_0$ is the present-day Hubble constant, $a(\eta)$ is the scale factor, $\eta_0$ is the conformal time today and $\chi$ is the comoving radial distance.
+\cite{Bartelmann2001}.
+Here, $\Omega_m$ is the present day matter density parameter, $H_0$ is the present-day Hubble constant, $a(\eta)$ is the scale factor, $\eta_0$ is the conformal time today and $\chi$ is the comoving radial distance.
 $\chi_*$ presents the distance to surface of last scattering.
@@
 
-Determining the nonlinear matter power spectrum can be done using numerical codes such as CAMB \cite{Lewis2000}.
+An approximate form of nonlinear matter power spectrum can be obtained using numerical codes such as CAMB \cite{Lewis2000}.
 The nonlinear matter bispectrum was calculated from the power spectrum using a fitting formula based on perturbation theory in \cite{bispfit}.
@@
 One can also choose to make order by order corrections to the Born approximation.
-There is motivation to understand the effects of these corrections (see the introduction) so we present them here.
+As outlined in the introduction there is motivation to understand the effects of these corrections, so we present them here.
 
@@
 We thus calculate the potential instead of the distortion tensor itself.
-When post-born effects are taken into account this is no longer true and we are forced to work with $\psi_{ab}$ directly.
-Following \cite{postborn_pratten_lewis}, the leading order post born correction to the distortion tensor for a point source at distance $\chi_s$ is given by
+When post-Born effects are taken into account this is no longer true and we are forced to work with $\psi_{ab}$ directly.
+Following \cite{postborn_pratten_lewis}, the leading order post-Born correction to the distortion tensor for a point source at distance $\chi_s$ is given by
 \begin{equation}
@@
 \subsection{Fisher matrix analysis}
-The signal-to-noise (SNR) and parameter constraints are obtained through a Fisher matrix analysis as is standard in cosmology\cite{dodelson2020modern}.
+The signal-to-noise (SNR) and parameter constraints are obtained through a Fisher matrix analysis as is standard in cosmology \cite{dodelson2020modern, Knox1995, Tegmark1997}.
 A full derivation of the equations used below can be found in appendix \ref{sec:fisher}.
@@
 \subsection{Fiducial cosmology}
-Our fiducial cosmology is based on the Planck results \cite{planckresults} (see table \ref{tab:fiducialpars}).
+Our fiducial cosmology is based on the \textit{Planck} results \cite{planckresults} (see table \ref{tab:fiducialpars}).
 We choose to constrain the standard parameters of the $\Lambda$CDM model as well as neutrino mass, $m_\nu$, and dark energy equation of state parameter, $w_0$.
-Our reason for including these extra degrees of freedom is that we would like to see to what degree future surveys will be able to tell us more about the mass of neutrinos and the nature of dark energy.
-For example, a paper from the DESI collaboration \cite{Roy2024DESI} recently showed that measurements of baryon acoustic oscillations give a 2.5 - 4 $\sigma$ tension with a non-evolving dark energy equation of state model, so it is worth considering if future weak lensing experiments could tell us more about this tension.
-Regarding neutrinos, depending on how stage 4 surveys will be able to constrain neutrino mass, the results could tell us more about their mass hierarchy and serve as a test for constraints obtained from other probes, such as baryon acoustic oscillations.
+The reason to include these extra degrees of freedom is to explore the potential of future surveys on these extensions of $\Lambda$CDM.
+For example, the DESI collaboration \cite{Roy2024DESI} recently showed that measurements of baryon acoustic oscillations give a 2.5 - 4 $\sigma$ tension with a non-evolving dark energy equation of state model, so it is worth considering if future weak lensing experiments could tell us more about this tension and if higher order statistics would benefit the constraints on these tensions.
+Regarding neutrinos, depending on how tightly stage 4 surveys will be able to constrain neutrino mass, the results could tell us more about their mass hierarchy and serve as a test for neutrino mass constraints obtained from other probes, such as baryon acoustic oscillations.
 
@@
 This prior only restricts $\Omega_b h^2$ and $n_s$ significantly, as those are not well constrained by lensing spectra alone.
-Our differences with \cite{Planck2018Lensing} are that (i) we do not fix $\tau$, but instead take it to have a SNR of 1 and added a loose $A_s$ prior as well, (ii) instead of a flat prior for the Hubble constant ($40 < H_0 < 100$) we use a gaussian distribution with the same standard deviation, (iii) we added a weak constraint on $\Omega_c h^2$, and (iv) we also vary neutrino mass and $w_0$ but do not assume any constraints on them in our priors.
+Our differences with \cite{Planck2018Lensing} are that (i) we do not fix $\tau$, but instead take it to have a SNR of 1 and added a weak $A_s$ prior as well, (ii) instead of a flat prior for the Hubble constant ($40 < H_0 < 100$) we use a gaussian distribution with the same standard deviation, (iii) we added a weak constraint on $\Omega_c h^2$, and (iv) we also vary neutrino mass and $w_0$ but do not assume any constraints on them in our priors.
 The purpose of the weak prior is as follows.
@@
 Future lensing surveys will of course use information from surveys measuring other statistics such as those derived from galaxy clustering and the primary CMB anisotropies.
-By adding a prior based on these (primary) statistics our results become more representative.
+By adding a prior based on these (primary) statistics our results become more representative of we should expect is possible.
 Additionally, by considering a conservative prior, it remains clear how the different lensing spectra complement each other.
-The other prior we consider is based on the CMB temperature and $E$-mode polarization cross- and auto-powerspectra from $\ell=30$ to $\ell = 2000$ assuming the same noise properties as used for the CMB weak lensing reconstruction noise.
-Adopting these priors  serve to test how weak lensing statistics are able to further tighten constraints beyond measuring primary statistics.
+The other prior we consider is based on the CMB temperature and $E$-mode polarization cross- and auto-power spectra from $\ell=30$ to $\ell = 2000$ assuming the same noise properties as used for the CMB weak lensing reconstruction noise.
+Adopting these priors  serves to test how weak lensing statistics are able to further tighten constraints beyond measuring primary statistics.
 It is worth noting that the constraint for $\tau$ and $w_0$ are tighter than in other literature (see \cite{cmbs4sciencebook} and \cite{Namikawa_2016} for comparable $\tau$ and $w_0$ constraints, respectively).
@@
 This included looking at the accuracy of numerical derivatives w.r.t.
-cosmological parameters, testing the numerical stability of Fisher matrices wr.r.t.
+cosmological parameters, testing the numerical stability of Fisher matrices w.r.t.
 matrix inversion, and varying our values for detector noise and beam width.
-} we suspect that this is mainly due us not taking into account foregrounds in our noise models.
+} we suspect that this is partly due to us not taking into account foregrounds in our noise models.
 
@@
     \caption{Cosmological parameters varied ($\Lambda \text{CDM} + w_0 + m_\nu$).
-We use a prior similar to the one used by the Planck weak lensing results \cite{Planck2018Lensing} which essentially only restricts $\Omega_bh^2$ and $n_s$ significantly.}
+We use a prior similar to the one used by the \textit{Planck} weak lensing results \cite{Planck2018Lensing} which essentially only restricts $\Omega_bh^2$ and $n_s$ significantly.}
     \label{tab:fiducialpars}
@@
 This paper considers noise levels for ``stage 3'' and ``stage 4'' (weak lensing) surveys.
-Noise properties considered for this analysis can be found in table \ref{tab:noiselevels}, including experiments they are representative of.
+Noise properties considered for this analysis can be found in table \ref{tab:noiselevels}, including a list of experiments with comparable noise properties.
 A comparison of the noise power spectra to the lensing power spectra are shown     in figure \ref{fig:lpsplusnoise}.
@@
     \cline{2-4}
-        & stage 4 & $\sigma_{\text{rms}} = 0.3$, $n_g = 30 \text{arcmin}^{-2}$ & LSST, Euclid \\
+        & stage 4 & $\sigma_{\text{rms}} = 0.3$, $n_g = 30 \text{arcmin}^{-2}$ & LSST, \textit{Euclid} \\
     \hline
@@
 Galaxy lensing is determined by measuring lensing shear (see appendices \ref{sec:weaklensing} and \ref{sec:shear}).
-The noise in this measurement is dominated by scale-independent shot noise and has associated noise power spectrum $N_l^{\text{shear}} = \sigma_{\text{rms}}^2 / n_g$, \cite{bartelmann2001weak}.
+The noise in this measurement is dominated by scale-independent shot noise and has associated noise power spectrum $N_l^{\text{shear}} = \sigma_{\text{rms}}^2 / n_g$, \cite{Bartelmann2001}.
 $n_g$ is the amount of galaxies observed per unit solid angle.
@@
     \includegraphics[width=\textwidth]{../code/plots/spectraplusnoise.pdf}
-    \caption{CMB (right) and galaxy (left) lensing potential power spectra compared to associated experimental noise.
-Current (stage 3) noise values are displayed as well as near future\footnote{Some of the stage 4 experiments are already in operation, but expected noise levels are only achieved after several years of integration.} (stage 4) noise values.
+    \caption{ CMB (right) and galaxy (left) lensing potential power spectra compared to associated experimental noise.
+Current (stage 3) noise values are displayed as well as near future (some of the stage 4 experiments are already in operation, but expected noise levels are only achieved after several years of integration) (stage 4) noise values.
 The CMB lensing experiment uses only polarization.
 CMB noise values are chosen in accordance with \cite{Namikawa_2016}.
-For comparison we also how the reconstruction noise for the Simons Observatory.
+For comparison we also show the reconstruction noise for the Simons Observatory \cite{Ade2019}.
 Shear noise values are chosen to be similar to e.g.
-Euclid measurements for stage 4 and e.g.
-KiDS for stage 3.}
+\textit{Euclid} measurements \cite{laureijs2009} for stage 4 and e.g.
+KiDS \cite{kuijken2021} for stage 3.}
     \label{fig:lpsplusnoise}
@@
 % and $\sigma^2_\gamma$ is the intrinsic ellipticity dispersion. 
-% Experiments such as Euclid and the LSST both have $\sigma_{rms} \approx 0.3$ \cite{euclid1overview} \cite{lsstsciencebook} and 
+% Experiments such as \textit{Euclid} and the LSST both have $\sigma_{rms} \approx 0.3$ \cite{\textit{Euclid}1overview} \cite{lsstsciencebook} and 
 % $ n_g\approx 30 \text{ arcmin}^{-2}$.
@@
 \subsection{Other details}
-The redshift distribution of the observed galaxies is commonly parameterized as \cite{bartelmann2001weak}
+The redshift distribution of the observed galaxies is commonly parameterized as \cite{Bartelmann2001}
 \begin{equation*}
@@
 \end{equation*}
-We choose the parameter combination $a = 2$, $b = 3/2$ and $z_0 = 0.64$ which is similar to the expected distributions of Euclid, (which will probe primarily in the 0.2 - 2.6 redshift range \cite{euclidprep10})
+We choose the parameter combination $a = 2$, $b = 3/2$ and $z_0 = 0.64$ which is similar to the expected distributions of \textit{Euclid}, (which will probe primarily in the 0.2 - 2.6 redshift range \cite{euclidprep10})
 and the LSST mission (which has $a \approx 2$, $b \approx 1$, and $z_0 \approx 0.3$ from predictions for the obtained data \cite{lsstsciencebookchapter3}).
@@
 For the CMB the $S/N$ is reduced significantly (in agreement with \cite{postborn_pratten_lewis}) but not so severely that they are no longer detectable.
-Our conclusions thus remain the same: the CMB bispectrum is probably detectable with stage 3 surveys and definitely detectable with stage 4 surveys.
+Our conclusions thus remain the same: the CMB bispectrum is possibly detectable with stage 3 surveys and definitely detectable with stage 4 surveys in the absence of any further systematics.
 
@@
 \begin{itemize}
-    \item Lensing power spectra alone are not sufficient to competitively constrain $\Lambda\mathrm{CDM} + w_0 + m_\nu$ without priors, even in the case of stage 4 surveys.
+    % \item Lensing power spectra combined with the weak prior is close in constrainting alone are not sufficient to competitively constrain $\Lambda\mathrm{CDM} + w_0 + m_\nu$ without priors, even in the case of stage 4 surveys.
 
@@
 
-    \item It is clear that post-Born corrections do not change the constraints drastically.
-As expected, they matter less for galaxy lensing and when using the CMB $T+E$ prior.
-Interestingly, although the constraints due to bispectra alone are always worse, for some parameters, they are instead improved when combining lensing bispectrum information with that of the lensing power spectrum.
+    \item Post-Born corrections do not change the constraints drastically.
+As expected, they matter less in particular for galaxy lensing and when using the CMB $T+E$ prior.
+We thus conclude that taking into account post-Born corrections does not lead to significantly different conclusions about the general sensitivity of the lensing spectra.
+
+    \item Interestingly, when comparing the post-Born results with the Born-approximate results, although the constraints due to bispectra alone are always worse, for some parameters, they are instead improved when combining lensing bispectrum information with that of the lensing power spectrum.
 This implies that, at least in some cases, the post-Born contribution to the lensing bispectrum depends on the parameters in a different way than the lensing power spectrum and thus improves the amount of information we can extract from survey data despite making the lensing bispectrum less detectable.
-Besides these observations, we conclude that taking into account post-Born corrections does not lead to different conclusions about the overall sensitivity of the lensing spectra.
 \end{itemize}
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{figures/param_constraints_tight_cmb_cmb_prior.pdf}
+    \includegraphics[width=0.5\textwidth]{figures/param_constraints_tight_cmb_cmb_prior.pdf}
     \caption{Same as figure \ref{fig:paramconstraintstightcmb}, except we use the CMB temperature and polarization based priors listed in table \ref{tab:fiducialpars}.}
@@
 In this work, we have presented parameter forecasts from CMB lensing and galaxy lensing power and bispectra, considering experimental parameters representative of stage 3 and stage 4 CMB and galaxy surveys.
-Our analysis shows that while lensing power spectra are detectable at high significance, the inclusion of bispectra generally offers significant improvements when considering parameter constraints when considering an LCDM+ model of the universe.
+Our analysis shows that while lensing power spectra are detectable at high significance, the inclusion of bispectra generally offers significant improvements when considering parameter constraints when considering an $\Lambda$CDM+ model of the universe.
 Notably, both CMB and galaxy bispectra are found to be detectable, even with stage 3 experimental noise levels.
@@
 Furthermore, the combination of CMB and galaxy lensing probes, especially when both power spectra and bispectra are utilized, further enhances constraints and aids in mitigating degeneracies.
-If all information were to be combined\footnote{And using the CMB $T+E$ prior.} we find particularly tight constraints on neutrino mass and dark energy equation of state in particular, i.e.
-$\sigma(m_\nu) = 55$ meV and $\sigma(w_0) = 0.017$.
+If all information were to be combined\footnote{And using the CMB $T+E$ prior.} we find particularly tight constraints on neutrino mass and dark energy equation of state in particular, i.e., $\sigma(m_\nu) = 55$ meV and $\sigma(w_0) = 0.017$.
 However, lensing power spectra alone, even for stage 4 surveys, appear insufficient to competitively constrain an 8-parameter $\Lambda\text{CDM} + w_0 + \sum m_\nu$ model without any priors.
@@
 This intrinsic inaccuracy in the model for $B^{\delta}$ will propagate to the lensing bispectra.
-
-     \item Certain cosmological parameters that affect nonlinear large-scale structure, such as neutrino masses, likely alter the fitting parameters of the matter bispectrum formula when varied.
+A correct modeling using e.g.
+effective field theory would be preferred, and would also allow us to marginalize over bias parameters.
+
+     \item Certain cosmological parameters that affect nonlinear large-scale structure, such as neutrino masses or the dark energy equation of state, likely alter the fitting parameters of the matter bispectrum formula when varied.
 In the literature, this is currently not considered, which can make the derivatives of the bispectrum with respect to these cosmological parameters less accurate.
@@
 Moving beyond the Limber approximation can be achieved using methods such as those explored in reference \cite{Chen:2021vba} and might offer a path to address this.
+It is also worth mentioning that including foregrounds and other systematics will especially raise the noise for lower multipoles (see e.g.
+\cite{de_Oliveira_Costa_2006}).
 
@@
 
-     \item Related to the previous point, for a completely correct analysis, we would also need to take into account correlations between a CMB prior and the lensing power and bispectra.
+     \item Related to the previous point, for a completely correct analysis, we would also need to take into account correlations between a CMB prior and the lensing power and bispectra (see e.g.
+\cite{schmittfull2018}).
 Currently, we simply add Fisher matrices; however, taking these correlations into account means computing a covariance matrix for combinations of $T$, $E$, $\phi$ with nonzero entries for correlations between $T$/$E$ and $\phi$.
@@
 
-     \item Another consideration is that non-Gaussianity of the true lensing field doesn't only change the covariance matrices used to calculate Fisher matrix elements, but also introduces an additional bias to the estimated lensing powerspectrum, denoted as $N^{(3/2)}_\ell$ \cite{B_hm_2016, B_hm_2018}.
+     \item Another consideration is that non-Gaussianity of the true lensing field doesn't only change the covariance matrices used to calculate Fisher matrix elements, but also introduces an additional bias to the estimated lensing power spectrum, denoted as $N^{(3/2)}_\ell$ \cite{B_hm_2016, B_hm_2018}.
 This is because the weights in quadratic estimator from ref.
@@
 
-     \item Other than the lensing powerspectrum and bispectrum, there are also other statistics from the lensing field that can be used for parameter constraints.
+     \item Other than the lensing power spectrum and bispectrum, there are also other statistics from the lensing field that can be used for parameter constraints.
 For example, ref.
-\cite{Liu_2016} showed that the lensing convergence one-point probability distribution function (PDF) and peak counts can tighten constraints in the $\Omega_m$ - $\sigma_8$ plane by $\sim30\%$ over just using the lensing powerspectrum for a stage 3 experiment\footnote{Specifically, for the AdvACT experiment.}.
+\cite{Liu_2016} showed that the lensing convergence one-point probability distribution function (PDF) and peak counts can tighten constraints in the $\Omega_m$ - $\sigma_8$ plane by $\sim30\%$ over just using the lensing power spectrum for a stage 3 experiment\footnote{Specifically, for the AdvACT experiment \cite{Liu_2016}.}.
 It is thus clear that future work could combine the power- and bispectrum with these additional statistics to further tighten forecasted constraints.
@@
 
-    \item To calculate the Fisher matrices, derivatives of the lensing spectra with respect to cosmological parameters were computed using a central difference formula.
-This introduces errors (estimated at most a few percent, though larger for neutrino mass derivatives, see appendix \ref{sec:derivatives}).
-Functions such as the matter power spectrum and fitting parameters for the matter bispectrum were approximated using pre-generated data and interpolation, also introducing errors of at most a few percent.
-Currently, these numerical errors are not considered significant because this paper primarily aims to provide a qualitative understanding of the utility of CMB and galaxy weak lensing spectra.
-However, for precise quantitative results, these factors would need to be revisited and improved.
+    \item In addition to the general limitations of the Fisher matrix formalism, our implementation also introduces numerical errors due to the use of finite-difference derivatives and interpolation of spectra, leading to inaccuracies at the few percent level (see appendix \ref{sec:derivatives}).
+These are not expected to qualitatively affect our conclusions.
 

### 2025-07-25 paper/playing_with_fish.py
plots bigger

@@ -640,7 +640,7 @@ if __name__ == "__main__":
         labels=labels,
         colors=["tab:blue", "tab:orange", 'tab:green', 'black'],
         nstd=1.0,
-        figsize=(4.15, 4.15),
+        figsize=(6, 6),
         dpi=100
     )
 

### 2025-07-26 paper/paper_draft.tex
paper final version 1

@@
 \usepackage{listings}
-\usepackage{xcolor} % For color customization (optional)
+\usepackage{xcolor}
+\usepackage{pgfmath}
+\newcommand{\greenfactor}[1]{%
+    \pgfmathsetmacro{\value}{min(100,round(100*(#1-1)))}%
+    \textcolor{green!80!black!\value!black}{#1}%
+}
 
@@
 
-In our analysis we will include post-Born corrections.
+In our analysis, we will include post-Born corrections.
 The CMB and galaxy lensing power spectra are typically affected by less than $1\%$ \cite{postborn_pratten_lewis, galpostbornlpscorr}.
@@
 In particular, for the folded configuration the effect can be of the same magnitude as the lensing bispectrum due to nonlinear large scale structure \eqref{eq:lensingbispectrum}.
-Including post-Born effects  will allow us to determine how future parameter constraints are by more accurate modelling.
+Including post-Born effects  will allow us to determine how future parameter constraints are affected by more accurate modeling.
 
@@
 In the case of CMB surveys, lensing alters the statistical properties of the temperature and polarization fields and can thus be calculated by comparing the observed signal to the expected unlensed signal (see e.g.
-\cite{cmblensingestimator, Maniyar2021_QE_CMBLensing, Hirata_2003}).
+\cite{Maniyar2021_QE_CMBLensing, Hirata_2003}).
 In galaxy surveys, lensing alters the ellipticities of observed galaxies.
@@
     C^{\psi_X\psi_Y}_\ell
-    & = \frac{9}{\ell^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)P^\delta(\ell/\chi,\eta_0-\chi), \label{eq:lensingpowerspectrum}\\
+    & = \frac{9}{\ell^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)P^\delta(\ell/\chi,\eta_0-\chi), \label{eq:lensingpower spectrum}\\
     B^{\psi_X\psi_Y\psi_Z}_{\ell\ell_1\ell\ell_2\ell\ell_3} &= \sqrt{\frac{(2\ell_1 + 1)(2\ell_2 + 1)(2\ell_3 + 1)}{4\pi}} \begin{pmatrix} \ell_1 & \ell_2 & \ell_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{\ell_1^2\ell_2^2\ell_3^2}\Omega_m^3H_0^6 \notag\\
@@
 \subsection{Post-Born corrections}
-One of the approximations made in the derivation of the weak lensing power- and bispectra (equations \eqref{eq:lensingpowerspectrum} and \eqref{eq:lensingbispectrum}) is that the integral is performed along the line of sight to the light source.
-To get an analytically correct result, one would instead need to solve a differential equation with a feedback loop where the effects of gravitational lensing affect the trajectory and shape of given bundle of light rays which in turn changes the way future gravitational lenses affect the bundle.
+One of the approximations made in the derivation of the weak lensing power- and bispectra (equations \eqref{eq:lensingpower spectrum} and \eqref{eq:lensingbispectrum}) is that the integral is performed along the line of sight to the light source.
+To get an analytically correct result, one would instead need to solve a differential equation with a feedback loop where the effects of gravitational lensing affect the trajectory and shape of given bundle of light rays, which in turn changes the way future gravitational lenses affect the bundle.
 One can also choose to make order by order corrections to the Born approximation.
@@
 We thus calculate the potential instead of the distortion tensor itself.
-When post-Born effects are taken into account this is no longer true and we are forced to work with $\psi_{ab}$ directly.
+When post-Born effects are taken into account this is no longer true and we are required to work with $\psi_{ab}$ directly.
 Following \cite{postborn_pratten_lewis}, the leading order post-Born correction to the distortion tensor for a point source at distance $\chi_s$ is given by
@@
 This prior only restricts $\Omega_b h^2$ and $n_s$ significantly, as those are not well constrained by lensing spectra alone.
-Our differences with \cite{Planck2018Lensing} are that (i) we do not fix $\tau$, but instead take it to have a SNR of 1 and added a weak $A_s$ prior as well, (ii) instead of a flat prior for the Hubble constant ($40 < H_0 < 100$) we use a gaussian distribution with the same standard deviation, (iii) we added a weak constraint on $\Omega_c h^2$, and (iv) we also vary neutrino mass and $w_0$ but do not assume any constraints on them in our priors.
+Our differences with \cite{Planck2018Lensing} are that (i) we do not fix $\tau$, but instead take it to have a SNR of 1 and added a weak $A_s$ prior as well, (ii) instead of a flat prior for the Hubble constant ($40 < H_0 < 100$) we use a Gaussian distribution with the same standard deviation, (iii) we added a weak constraint on $\Omega_c h^2$, and (iv) we also vary neutrino mass and $w_0$ but do not assume any constraints on them in our priors.
 The purpose of the weak prior is as follows.
@@
 Future lensing surveys will of course use information from surveys measuring other statistics such as those derived from galaxy clustering and the primary CMB anisotropies.
-By adding a prior based on these (primary) statistics our results become more representative of we should expect is possible.
+By adding a prior based on these (primary) statistics our results become more representative of what we should expect is possible.
 Additionally, by considering a conservative prior, it remains clear how the different lensing spectra complement each other.
@@
 
-The CMB lensing noise is estimated using a quadratic estimator \cite{cmblensingestimator}.
+The CMB lensing noise is estimated using a quadratic estimator \cite{Maniyar2021_QE_CMBLensing}.
 % We consider experiments that only measure polarization noise. This is similarly done in \cite{Namikawa_2016}. 
@@
 
-% For noise in CMB lensing we consider lensing estimated by an experiment measuring polarization with white noise $\Delta_P$ and beam spread $\sigma$.
-We then use the quadratic estimator from \cite{Hu_2002} to estimate the resulting noise in the lensing power spectrum.
-
-% For shear measurements, the noise is simpler to model as the primary source is shot noise due to the discrete number of galaxies in each patch of the survey.
-The mean ellipiticity of the galaxies in a patch is calculated and a deviation from the mean is interpreted as lensing shear.
-The greater the number of galaxies the more accurately can the mean ellipticity be determined.
-Choosing the size of the patch is a balancing between signal to noise and resolution.
-Smaller patches give more uncertain shear measurements while providing better overal resolution for the shear map and thus allowing measurements of shear bi- and powerspectra to higher multipole numbers.
-The opposite is true for larger patch sizes.
-The power spectrum of the shear noise is given by
-% \begin{equation*}
-%     N^{\gamma\gamma}_{l} = \frac{\sigma^2_{rms}}{\bar n_g}
-% \end{equation*}
-% where $\bar n_g$ is the mean number of galaxies observed per sterradian accross the whole survey \cite{weaklensingnotes2020}.
-% and $\sigma^2_\gamma$ is the intrinsic ellipticity dispersion. 
-% Experiments such as \textit{Euclid} and the LSST both have $\sigma_{rms} \approx 0.3$ \cite{\textit{Euclid}1overview} \cite{lsstsciencebook} and 
-% $ n_g\approx 30 \text{ arcmin}^{-2}$.
-
-% The noise values are calculated for stage 3 (current) surveys and stage 4 (near future) surveys.
-We will use the stage 4 noise values for parameter constraints.
-A comparison of stage 3 noise, stage 4 noise, and lensing signal can be seen in figure \ref{fig:lpsplusnoise}.
 \subsection{Other details}
@@
 We include both the power spectrum and the bispectrum of the weak lensing.
-We look at two extension of $\Lambda CDM$, $m_\nu$ and $w_0$ as well as the derived parameters $\sigma_8$, and $\Omega_m$.
+We look at two extensions of $\Lambda CDM$, $m_\nu$ and $w_0$ as well as the derived parameters $\sigma_8$, and $\Omega_m$.
 Constraints on the full set of $\Lambda$CDM parameters can be found in appendix \ref{sec:lambdacdmconstraints}.
@@
 When combining CMB and galaxy lensing we see that the bispectrum is slightly less competitive compared to the lensing power spectrum, but can still significantly improve constraints compared to only using the lensing power spectrum.
-% Again in agreement with reference \cite{Namikawa_2016}, in which case a 33 percent and 35 percent improvement was found for $m_\nu$ and $w_0$, respectively, by measuring the CMB lensing bispectrum in addition to the power spectrum.
 
@@
 \begin{itemize}
-    % \item Lensing power spectra combined with the weak prior is close in constrainting alone are not sufficient to competitively constrain $\Lambda\mathrm{CDM} + w_0 + m_\nu$ without priors, even in the case of stage 4 surveys.
-
     \item For lower noise values (in particular for stage 4 noise), the bispectra become increasingly important compared to the power spectra.
@@
 
-    \item Interestingly, when comparing the post-Born results with the Born-approximate results, although the constraints due to bispectra alone are always worse, for some parameters, they are instead improved when combining lensing bispectrum information with that of the lensing power spectrum.
+    \item Interestingly, when comparing the post-Born results with the Born-approximate results, although bispectrum-only constraints are generally weaker, their combination with power spectra information instead improves constraints for some parameters.
 This implies that, at least in some cases, the post-Born contribution to the lensing bispectrum depends on the parameters in a different way than the lensing power spectrum and thus improves the amount of information we can extract from survey data despite making the lensing bispectrum less detectable.
@@
     \scriptsize
-%     \begin{tabular}{|l|r|rrr|rrr|rrr|}
-%         \hline
-%         \multicolumn{11}{|c|}{weak priors} \\
-%         \hline
-%         && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
-%         \hline
-%         Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
-% \hline
-%  $m_\nu$    & 1.000e+01 &  2.815e-01 &                  1.104e+00 &                           2.247e-01 &  9.954e-01 &                  2.646e-01 &                           1.551e-01 &  1.042e-01 &                  1.317e-01 &                           6.537e-02 \\
-%  $w_0$      & 1.000e+01 &  5.415e-01 &                  1.405e+00 &                           3.683e-01 &  2.638e-01 &                  7.817e-02 &                           2.475e-02 &  5.936e-02 &                  6.629e-02 &                           1.899e-02 \\
-%  $\sigma_8$ & 2.457e+00 &  1.181e-01 &                  1.872e-01 &                           5.486e-02 &  8.069e-02 &                  3.668e-02 &                           1.299e-02 &  1.375e-02 &                  1.587e-02 &                           8.048e-03 \\
-%  $\Omega_m$ & 6.702e-01 &  1.948e-01 &                  1.790e-01 &                           7.270e-02 &  1.434e-01 &                  4.831e-02 &                           1.667e-02 &  1.619e-02 &                  1.607e-02 &                           7.233e-03 \\
-%        \hline
-%         \multicolumn{11}{|c|}{CMB $T + E$ priors} \\
-%         \hline
-%         && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
-%         \hline
-%          Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
-%         \hline
-%  $m_\nu$    & 2.157e-01 &  7.445e-02 &                  1.178e-01 &                           7.283e-02 &  1.406e-01 &                  1.042e-01 &                           8.592e-02 &  4.340e-02 &                  8.965e-02 &                           3.735e-02 \\
-%  $w_0$      & 5.781e-02 &  4.606e-02 &                  5.285e-02 &                           4.578e-02 &  4.756e-02 &                  3.934e-02 &                           2.080e-02 &  3.231e-02 &                  3.667e-02 &                           1.741e-02 \\
-%  $\sigma_8$ & 1.547e-02 &  1.360e-02 &                  1.492e-02 &                           1.349e-02 &  1.252e-02 &                  1.136e-02 &                           7.881e-03 &  9.682e-03 &                  1.023e-02 &                           6.126e-03 \\
-%  $\Omega_m$ & 1.495e-02 &  1.381e-02 &                  1.457e-02 &                           1.376e-02 &  1.336e-02 &                  1.124e-02 &                           8.715e-03 &  1.028e-02 &                  9.616e-03 &                           5.484e-03 \\
-%        \hline
-%             \multicolumn{11}{|c|}{weak priors, with post-Born corrections} \\
-%         \hline
-%         && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
-%         \hline
-%         Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
-% \hline
-% $m_\nu$    & 1.000e+01 &  2.815e-01 &                  1.158e+00 &                           1.805e-01 &  9.954e-01 &                  3.349e-01 &                           2.073e-01 &  1.042e-01 &                  1.775e-01 &                           6.256e-02 \\
-%  $w_0$      & 1.000e+01 &  5.415e-01 &                  8.193e-01 &                           3.051e-01 &  2.638e-01 &                  8.328e-02 &                           2.825e-02 &  5.936e-02 &                  6.271e-02 &                           1.988e-02 \\
-%  $\sigma_8$ & 2.457e+00 &  1.181e-01 &                  1.846e-01 &                           5.660e-02 &  8.069e-02 &                  4.047e-02 &                           1.317e-02 &  1.375e-02 &                  1.782e-02 &                           8.369e-03 \\
-%  $\Omega_m$ & 6.702e-01 &  1.948e-01 &                  1.810e-01 &                           7.086e-02 &  1.434e-01 &                  5.474e-02 &                           1.811e-02 &  1.619e-02 &                  1.799e-02 &                           7.325e-03 \\
-%        \hline
-%         \multicolumn{11}{|c|}{CMB $T + E$ priors, with post-Born corrections} \\
-%         \hline
-%         && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
-%         \hline
-%          Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
-%         \hline
-%  $m_\nu$    & 2.157e-01 &  7.445e-02 &                  1.298e-01 &                           7.313e-02 &  1.406e-01 &                  1.166e-01 &                           9.869e-02 &  4.340e-02 &                  1.034e-01 &                           3.766e-02 \\
-%  $w_0$      & 5.781e-02 &  4.606e-02 &                  5.308e-02 &                           4.596e-02 &  4.756e-02 &                  4.163e-02 &                           2.220e-02 &  3.231e-02 &                  3.911e-02 &                           1.798e-02 \\
-%  $\sigma_8$ & 1.547e-02 &  1.360e-02 &                  1.498e-02 &                           1.354e-02 &  1.252e-02 &                  1.168e-02 &                           7.717e-03 &  9.682e-03 &                  1.073e-02 &                           6.285e-03 \\
-%  $\Omega_m$ & 1.495e-02 &  1.381e-02 &                  1.467e-02 &                           1.379e-02 &  1.336e-02 &                  1.141e-02 &                           8.530e-03 &  1.028e-02 &                  1.017e-02 &                           5.687e-03 \\
-%        \hline
-%         \end{tabular}
  \begin{tabular}{|l|r|rrr|rrr|rrr|}
@@
         \hline
-        Par & prior & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
-        \hline
-$m_\nu$    & 10 & 0.28 & 1.1 & 0.23 & 1.0 & 0.26 & 0.16 & 0.10 & 0.13 & 0.065 \\
+        Par & prior & $C_\ell$ & $B$ & $C_\ell + B$ & $C_\ell$ & $B$ & $C_\ell + B$ & $C_\ell$ & $B$ & $C_\ell + B$ \\
+        \hline
+$m_\nu$    & 10 & 0.28 & 1.1 & 0.23 (\greenfactor{1.22}) & 1.0 & 0.26 & 0.16 (\greenfactor{6.25}) & 0.10 & 0.13 & 0.065 (\greenfactor{1.54}) \\
 \rowcolor{blue!10}
-           &    &  & 1.2 & 0.18 & & 0.33 & 0.21 & & 0.18 & 0.063 \\
-$w_0$      & 10 & 0.54 & 1.4 & 0.37 & 0.26 & 0.078 & 0.025 & 0.059 & 0.066 & 0.019 \\
+           &    &  & 1.2 & 0.18 (\greenfactor{1.56}) & & 0.33 & 0.21 (\greenfactor{4.76}) & & 0.18 & 0.063 (\greenfactor{1.59}) \\
+$w_0$      & 10 & 0.54 & 1.4 & 0.37 (\greenfactor{1.46}) & 0.26 & 0.078 & 0.025 (\greenfactor{10.4}) & 0.059 & 0.066 & 0.019 (\greenfactor{3.11}) \\
 \rowcolor{blue!10}
-           &    &  & 0.82 & 0.31 &  & 0.083 & 0.028 & & 0.063 & 0.020 \\
-$\sigma_8$ & 2.5 & 0.12 & 0.19 & 0.055 & 0.081 & 0.037 & 0.013 & 0.014 & 0.016 & 0.0080 \\
+           &    &  & 0.82 & 0.31 (\greenfactor{1.74}) &  & 0.083 & 0.028 (\greenfactor{9.29}) & & 0.063 & 0.020 (\greenfactor{2.95}) \\
+$\sigma_8$ & 2.5 & 0.12 & 0.19 & 0.055 (\greenfactor{2.18}) & 0.081 & 0.037 & 0.013 (\greenfactor{6.23}) & 0.014 & 0.016 & 0.0080 (\greenfactor{1.75}) \\
 \rowcolor{blue!10}
-           &     &  & 0.18 & 0.057 & & 0.040 & 0.013 & & 0.018 & 0.0084 \\
-$\Omega_m$ & 0.67 & 0.19 & 0.18 & 0.073 & 0.14 & 0.048 & 0.017 & 0.016 & 0.016 & 0.0072 \\
+           &     &  & 0.18 & 0.057 (\greenfactor{2.11}) & & 0.040 & 0.013 (\greenfactor{6.23}) & & 0.018 & 0.0084 (\greenfactor{1.67}) \\
+$\Omega_m$ & 0.67 & 0.19 & 0.18 & 0.073 (\greenfactor{2.60}) & 0.14 & 0.048 & 0.017 (\greenfactor{8.24}) & 0.016 & 0.016 & 0.0072 (\greenfactor{2.22}) \\
 \rowcolor{blue!10}
-           &     &  & 0.18 & 0.071 & & 0.055 & 0.018 & & 0.018 & 0.0073 \\
+           &     &  & 0.18 & 0.071 (\greenfactor{2.68}) & & 0.055 & 0.018 (\greenfactor{7.78}) & & 0.018 & 0.0073 (\greenfactor{2.19}) \\
         \hline
@@
         \hline
-        Par & prior & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
-        \hline
-$m_\nu$    & 0.22 & 0.074 & 0.12 & 0.073 & 0.14 & 0.10 & 0.086 & 0.043 & 0.090 & 0.037 \\
+        Par & prior & $C_\ell$ & $B$ & $C_\ell + B$ & $C_\ell$ & $B$ & $C_\ell + B$ & $C_\ell$ & $B$ & $C_\ell + B$ \\
+        \hline
+$m_\nu$    & 0.22 & 0.074 & 0.12 & 0.073 (\greenfactor{1.01}) & 0.14 & 0.10 & 0.086 (\greenfactor{1.63}) & 0.043 & 0.090 & 0.037 (\greenfactor{1.16}) \\
 \rowcolor{blue!10}
-           &      & & 0.13 & 0.073 & & 0.12 & 0.099 & & 0.10 & 0.038 \\
-$w_0$      & 0.058 & 0.046 & 0.053 & 0.046 & 0.048 & 0.039 & 0.021 & 0.032 & 0.037 & 0.017 \\
+           &      & & 0.13 & 0.073 (\greenfactor{1.01}) & & 0.12 & 0.099 (\greenfactor{1.41}) & & 0.10 & 0.038 (\greenfactor{1.13}) \\
+$w_0$      & 0.058 & 0.046 & 0.053 & 0.046 (\greenfactor{1.00}) & 0.048 & 0.039 & 0.021 (\greenfactor{2.29}) & 0.032 & 0.037 & 0.017 (\greenfactor{1.88}) \\
 \rowcolor{blue!10}
-           &       & & 0.053 & 0.046 & & 0.042 & 0.022 &  & 0.039 & 0.018 \\
-$\sigma_8$ & 0.015 & 0.014 & 0.015 & 0.013 & 0.013 & 0.011 & 0.0079 & 0.0097 & 0.010 & 0.0061 \\
+           &       & & 0.053 & 0.046 (\greenfactor{1.00}) & & 0.042 & 0.022 (\greenfactor{2.18}) &  & 0.039 & 0.018 (\greenfactor{1.78}) \\
+$\sigma_8$ & 0.015 & 0.014 & 0.015 & 0.013 (\greenfactor{1.08}) & 0.013 & 0.011 & 0.0079 (\greenfactor{1.65}) & 0.0097 & 0.010 & 0.0061 (\greenfactor{1.59}) \\
 \rowcolor{blue!10}
-           &       & & 0.015 & 0.014 &  & 0.012 & 0.0077 & & 0.011 & 0.0063 \\
-$\Omega_m$ & 0.015 & 0.014 & 0.015 & 0.014 & 0.013 & 0.011 & 0.0087 & 0.010 & 0.0096 & 0.0055 \\
+           &       & & 0.015 & 0.014 (\greenfactor{1.00}) &  & 0.012 & 0.0077 (\greenfactor{1.69}) & & 0.011 & 0.0063 (\greenfactor{1.54}) \\
+$\Omega_m$ & 0.015 & 0.014 & 0.015 & 0.014 (\greenfactor{1.00}) & 0.013 & 0.011 & 0.0087 (\greenfactor{1.49}) & 0.010 & 0.0096 & 0.0055 (\greenfactor{1.82}) \\
 \rowcolor{blue!10}
-           &       & & 0.015 & 0.014 & & 0.011 & 0.0085 & & 0.010 & 0.0057 \\
+           &       & & 0.015 & 0.014 (\greenfactor{1.00}) & & 0.011 & 0.0085 (\greenfactor{1.53}) & & 0.010 & 0.0057 (\greenfactor{1.75}) \\
         \hline
@@
 The blue colored rows show constraints from bispectra with post-Born corrections.
-Post-Born corrections are not taken into account for the lensing power spectra as mentioned earlier, so these values are omitted in the blue rows.}
+Post-Born corrections are not taken into account for the lensing power spectra as mentioned earlier, so these values are omitted in the blue rows.
+The values between brackets indicate the fractional improvement in constraining power when combining the power- and bispectra compared to only using the power spectrum.
+They are color-coded such that greener values indicate stronger improvements.}
 \label{tab:paramconstraintstight}
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figures/param_constraints_tight_cmb_cmb_prior.pdf}
+    \includegraphics[width=0.7\textwidth]{figures/param_constraints_tight_cmb_cmb_prior.pdf}
     \caption{Same as figure \ref{fig:paramconstraintstightcmb}, except we use the CMB temperature and polarization based priors listed in table \ref{tab:fiducialpars}.}
@@
 \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \tiny
-%     \begin{tabular}{lrrrrr}
-%         \hline
-%          Par        &   CMB + Gal Powersp &   CMB + Gal Bisp &   CMB Power- + Bisp &   Gal Power- + Bisp &   CMB + Gal Power- + Bisp \\
-%         \hline
-%          $H_0$      &      102.8657993427 &     1.9383910910 &        9.6317154638 &        5.2219296015 &              0.9395018958 \\
-%          $\sigma_8$ &       31.8199604416 &     3.8499102909 &        9.1631304725 &        4.5639390699 &              1.8357116464 \\
-%          $\Omega_m$ &      172.9940493951 &     5.6526767874 &       27.2943604041 &       14.4462387878 &              3.5663315554 \\
-%         \hline
-%     \end{tabular}
-% \caption{Same as table \ref{tab:paramconstraintstight}, except for a smaller selection of parameters well constrained by lensing.}
-% \label{tab:paramconstraintstight}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[width=\textwidth]{figures/param_constraints_tight.pdf}
-%     \caption{Same as \ref{fig:paramconstraintsall}, except showing constraints for a set of parameters that can be determined particularly accurately with weak lensing surveys.
-We chose Hubble's constant, $H_0$, baryonic + dark matter density, $\Omega_m$, and the root mean square density variation when the matter distribution is smoothed on a scale of 8 megaparsecs, commonly known as $\sigma_8$.}
-%     \label{fig:paramconstraintstight}
-% \end{figure}
 
@@
 This is because the weights in quadratic estimator from ref.
-\cite{Hu_2002} are calculated while assuming a gaussian lensing field.
+\cite{Hu_2002} are calculated while assuming a Gaussian lensing field.
 The bias is significant at the percent level for CMB lensing, however it may be larger for cross correlations of the CMB lensing field with low redshift mass tracers (such as galaxy lensing).
@@
 
-% The spin weighted spherical harmonics, $_sY_l^m$ are eigenfunctions of the raising/lowering operator, in particular
-% \begin{equation}
-%     \eth_sY_l^m = \sqrt{(l-s)(l+s+1)} _{s+1}Y_l^m.
-% \end{equation}
 The spherical harmonics are eigenfunctions of $\nabla^2$ and the spin raising and lowering operators. Using this property, the corresponding relations in spherical harmonic space can be shown to be 
@@
 \end{equation}
-We can then relate the multipole modes of $\psi$ to the Fourier modes of $\Psi$ through % TODO: check sign convention here
+We can then relate the multipole modes of $\psi$ to the Fourier modes of $\Psi$ through
 \begin{gather}
@@
 with $\eta$ denoting the conformal time. This gives 
-%TODO: define the window function
 \begin{equation}
@@
 \end{equation}
-Note that the correlation is independent of $m$ and $m'$. %TODO: explain how this relates to statistical isotropy
+Note that the correlation is independent of $m$ and $m'$.
 We can thus read from equation \ref{eq:lenspotcorrexpanded} that
@@
 \end{equation}
-%TODO: how high? quantify limber approx, perhaps using convergence x LSS paper
 It allows us to write the Limber-approximate angular spectrum as
@@
 \end{gather}
-% The power spectrum of the potential can be related to the power spectrum of the matter density using the Poisson equation %TODO: create ref
-% For a flat universe in atter or dark energy domination we thus find the relation
-% \begin{equation}
-%     P_\Psi(k;\eta, \eta)=\frac{9\omega_m^2(\eta)H^4(\eta)}{8\pi^2} \frac{P_m(k, \eta)}{k},
-% \end{equation}
-% where we use the shorthand $P_m(k, \eta):=P_m(k,\eta,\eta)$. The lensing power spectrum is thus finally given in terms of the matter power spectrum as
-% \begin{equation}
-%     C_l^\psi=\underbrace{9\omega_m^2(\eta)H^4(\eta)}_{:=C(\eta)}\frac{1}{l^4}\int_0^{\chi_*}\d\chi\chi^2 W(\chi)^2P_m(l/\chi, \eta_0-\chi).
-% \end{equation}
-% We will henceforth denote the $\eta$ dependent factor outside of the integral as $C(\eta)$.
+
 \subsection{Lensing potential bispectrum}
@@
 
-% TODO: mention that the definition of the bispectrum is different form power spectrum of Psi in terms of normalization constant
-% TODO: explain 3d dirac delta in terms of spherical harmonics (eq 78 thesis Luna)
-% TODO: explain that we used that int d^2 Omega Y Y Y = A_vec l^vec m (eq 82 thesis Luna)
-% using that d^3 k = k^2 d k d^2 Omega_k
-% TODO: fix this
 \subsection{Gravitational potential spectra in terms of matter spectra}
-% TODO: fix references and citations 
 We can rewrite equations \ref{eq:lpsintermsofpsi} and \ref{eq:lbsintermsofpsi} in terms of the matter spectra instead of the $\psi$ spectra using the Poisson equation.
@@
 
-% \subsection{Fisher Matrices and Eigenvalues}
-% As explained earlier, inverting the Fisher matrix gives us a covariance matrix of the estimators, $\hat \theta_i$ of our experimental parameters. For any vector $\mathbf v$,
-% \begin{equation*}
-%     \mathbf{v}^T_j \Cov_{ij} \mathbf v_i = \langle (v_i \hat \theta_i)^2 \rangle.
-% \end{equation*}
-% In particular, if the covariance matrix thus has some large eigenvalue $\lambda_L$ (large compared to the overal accuracy of our experiment), with eigenvector $v_i$, then the corresponding estimator, $v_i\hat \theta_i$, has a large variance and thus represents an approximate degeneracy in our parameters.
-Clearly, when looking at the Fisher matrix we thus instead look for the smallest eigenvalue.
-
-% If we have such an abnormally small eigenvalue we are likely to run into problems with calculating and inverting the Fisher matrix due to numerical errors.
-These problems are that inversion becomes unstable and that the matrix as a whole might not be PSD because the small eigenvalue was (wrongly) calculated to be negative.
-These problems are likely because (1) the smallest eigenvalue is sensitive to numerical errors due to being small and (2) the errors on the eigenvalues is proportial to the condition number of the matrix \cite{horn2012matrix}, where the condition number is the ratio between the largest and smallest eigenvalue of the matrix \cite{trefethen1997numerical}.
-Specifically, for a matrix $A$ with some perturbation $\Delta A$, the corresponding  perturbation of eigenvalue $\lambda_i$ is bounded above as
-% \begin{equation*}
-%     |\Delta \lambda_i | \leq ||\Delta A || \kappa(A)|,
-% \end{equation*}
-% with $\kappa(A)$ the condition number.
-
-% Before inverting a Fisher matrix, it is thus recommended to remove rows and columns corresponding to parameters that are poorly constrained by the experiment, which can be determined by checking for small eigenvalues and their eigenvectors.
-
-% Computing the inverse is then reduced to computing the inverse of the individual block matrices.
-The block matrices for sum 1 and 2 don't have a particularly nice form (cite paper about multiple tracer fisher matrices for bispectra) but for sum 3 the inverse of the block matrix corresponding to $\ell_1\ell_2\ell_3$ is
-% \begin{equation*}
-%     (\Cov_{\ell_1\ell_2\ell_3}^{-1})^{XYZ, X'Y'Z'} = \hat C_{\ell_1}^{X,X'}\hat C_{\ell_2}^{Y,Y'}\hat C_{\ell_3}^{Z,Z'}.
-% \end{equation*}
-% \begin{equation*}
-%     (\Cov)^{XYZ, X'Y'Z'} = C_{\ell_1}^{X,X'} C_{\ell_2}^{Y,Y'} C_{\ell_3}^{Z,Z'}.
-% \end{equation*}
-% As can be seen from
-% \begin{equation*}
-%     \sum_{X'Y'Z'} = \hat C_{\ell_1}^{X,X'}\hat C_{\ell_2}^{Y,Y'}\hat C_{\ell_3}^{Z,Z'} C_{\ell_1}^{X'X''}C_{\ell_2}^{Y'Y''}C_{\ell_3}^{Z'Z''} = \delta_{XX''}\delta_{YY''}\delta_{ZZ''} = \delta_{XYZ, X''Y''Z''}
-% \end{equation*}
-
-% It's worth mentioning that due to all $l$'s being distinct in each configuration of sum 3 the Wick contraction corresponding to each element of the covariance matrix can only be done in one way and we thus only have one term.
-In the sum 2 we get two terms due to $\ell_1$ and $\ell_2$ being identical.
-In sum 3 we similarly get 6 terms corresponding to all possible Wick contractions being nonzero.
-
-% \subsection{Shear noise}
-% In the absence of lensing, galaxies will generally still look elliptical rather than perfectly spherical.
-This \textbf{intrinsic ellipticity} is a major cause of noise when measuring lensing shear.
-We can define the ellipticity of a galaxy image (or any type of object, for that matter) in terms of \textbf{quadrupole moments} of the surface brightness as
-% \begin{equation*}
-% 	e_1 = \frac{I_{xx} - I_{yy}}{I_{xx} + I_{yy}}, \quad e_2 = \frac{2I_{xy}}{I_{xx} + I_{yy}},
-% \end{equation*}
-% where
-% \begin{equation*}
-% 	I_{ij} = \int \d x\d y I(x,y)ij, \quad i,j\in\{x,y\},
-% \end{equation*}
-% and the galaxy is centered on $(x,y)=0$.
-
-% The connection with lensing shear is then made by taylor expanding and using the lensing matrix.
-If we have unlensed ellipiticities $e_i^\text{true}$ and lensed ellipticities $e_i^{\text{obs}}$, then \cite{dodelson_modern_cosmology}
-% \begin{gather*}
-% 	e_1^{\text{obs}} = \br{1-2\sbr{e_1^{\text{true}}\gamma_1+e_2^{\text{true}}\gamma_2}}e_1^{\text{true}} + 2\gamma_1,\\
-% 	e_2^{\text{obs}} = \br{1-2\sbr{e_1^{\text{true}}\gamma_1+e_2^{\text{true}}\gamma_2}}e_2^{\text{true}} + 2\gamma_2.
-% \end{gather*}
-% In the case of weak lensing, $\gamma_i, \kappa \ll 1$, thus the above reduces to $\gamma_i \approx (e_i^{\text{obs}} - e^{\text{true}}_i)/2$.
-The analysis thusfar was quietly done in real space, but taking the harmonic transform of $\gamma_i$ is trivial under the approximation we just made.
-We thus find the error due to intrinsic ellipticity in harmonic space to equal $ \langle \gamma_{i,l}^2\rangle \approx \langle \br{e_{i,l}^\text{true}}^2\rangle/4$.
-In a real survey we don't measure the shear for individual galaxies, but instead average it over small patches of the sky.
-The size of each patch is chosen to balance between signal to noise ratio of the shear measurement and resolution of the shear map.
-If there are $N$ galaxies observed per patch, then the noise is given by
-% $$
-% \langle \br{\hat \gamma_{i,l}  - \gamma_{i,l}^{\text{true}}}^2\rangle \approx \frac{\langle \br{e_{i,l}^\text{true}}^2\rangle}{4N}.
-% $$
-% We don't know the exact value for the right hand side either, so we use the standard estimator for that.
-There is no correlation between two patches, i.e.
-the noise is white noise and thus scale independent.
-Finally, we want to look at the noise in $\gamma_l\gamma_{l'}^*$.
-We thus get
-% \begin{equation*}
-% 	C^{\gamma\gamma,\;\text{obs}}_l = C^{\gamma\gamma, \;\text{true}}_l + N_l^{\gamma\gamma},
-% \end{equation*} 
-% with
-% \begin{equation*}
-% 	 N_l^{\gamma\gamma} = \frac{1}{4}\frac{e_{\text{RMS}}}{N}, \quad e_{\text{RMS}}:=\sqrt{\frac{1}{N}\sum^N_{i=1}\sbr{(e_{1, i} - \bar e_1)^2 + (e_{2,i}-\bar e_2)^2}}.
-% \end{equation*}
-
-% Typically $e_{RMS} / 4 \sim 0.3$ \cite{Kilbinger_2015}.
-
-% The EUCLID team aims for a noise of about 20 percent and thus require at least $\sim 4 \times 10^8$ galaxies measured per patch, giving an absolute error of \cite{2024}
-% $$
-% N_l^{\gamma\gamma} \approx 8 \times 10^{-10} \quad \text{(EUCLID)}.
-% $$
-
-% \subsection{Convergence Noise}
-% \subsubsection{Optimal Estimator}
-% In practice, the lensing potential cannot be directly measured but is instead estimated based on the temperature, $\Theta$, $E$-mode polarization, $E$, and $B$-mode polarization, $B$, fields.
-
-% Define some axes $x$ and $y$ on a small patch of the sky and let $\mathbf E$ be the electric field of the CMB radiation, then the \textbf{linear polarization Stokes parameters} are defined as
-% \begin{equation*}
-% 	Q := E_xE^*_x - E_yE_y^*, \quad U := E_xE_y^* + E_yE_x^*.
-% \end{equation*} 
-% By looking at a sufficiently small patch of the sky we can apply the \textbf{flat sky limit}, in which case we take a 2D Fourier transform over the approximately flat patch instead of working with spherical harmonics.
-In this case we find $Q(\mathbf k)$ and $U(\mathbf k)$.
-In Fourier space we can instead work with the more convenient $E$ and $B$ modes, defined as
-% \begin{gather*}
-% 	E(\mathbf k)=\cos 2\phi_{\mathbf k} Q(\mathbf k) + \sin 2\phi_{\mathbf k} U(\mathbf k),\\
-% 	B(\mathbf k)=-\sin 2\phi_{\mathbf k} Q(\mathbf k) + \cos 2\phi_{\mathbf k} U(\mathbf k),
-% \end{gather*}
-% where $\phi_\textbf{k}$ is the angle of $\mathbf k$ with the $(1,0)$  vector (i.e. $k_{\hat{\mathbf{x}}}$).
-% $E$ corresponds to the curl free component of the polarization field and behaves like a scalar under transformations.
-$B$ corresponds to the rest of the polarization field and behaves like a semi-scalar under transformations.
-The Fourier components are thus related to the real space fields as
-% \begin{gather*}
-% \Theta(\mathbf{\hat{n}}) = \int \frac{d^2 k}{(2\pi)^2} \Theta(\mathbf{k}) e^{i \mathbf{k} \cdot \mathbf{\hat{n}}}, \\
-% [Q \pm iU](\mathbf{\hat{n}}) = -\int \frac{d^2 k}{(2\pi)^2} \left[ E(\mathbf{k}) \pm iB(\mathbf{k}) \right] e^{\pm 2i \varphi_{\mathbf{k}}} e^{i \mathbf{k} \cdot \mathbf{\hat{n}}}, \\
-% \phi(\mathbf{\hat{n}}) = \int \frac{d^2 K}{(2\pi)^2} \phi(\mathbf{K}) e^{i \mathbf{K} \cdot \mathbf{\hat{n}}}.
-% \end{gather*}
-
-% As discussed earlier, lensing acts in real space as 
-% \begin{equation*}
-%     X(\hat n) = \tilde X(\hat {\mathbf n} + \nabla \phi(\hat{\mathbf n})) \approx \tilde X(\hat{\mathbf n}) + \nabla \phi \cdot \nabla \tilde X |_{\hat{\mathbf{n}}},
-% \end{equation*}
-% with $X$ any of $\Theta$, $Q \pm iU$ and $\tilde X$ the corresponding \textit{unlensed} field.
-In Fourier space this naturally takes the form of a convolution.
-The changes in Fourier moments due to lensing are thus given by \cite{Hu_2002}
-% \begin{gather*}
-%     \delta \Theta(\mathbf{k}) = \int \frac{d^2 k'}{(2\pi)^2} \tilde{\Theta}(\mathbf{k'}) W(\mathbf{k'}, \mathbf{K}), \\[8pt]
-%     \delta E(\mathbf{k}) = \int \frac{d^2 k'}{(2\pi)^2} 
-%     \left[ \tilde{E}(\mathbf{k'}) \cos 2\varphi_{k'k} - \tilde{B}(\mathbf{k'}) \sin 2\varphi_{k'k} \right] W(\mathbf{k'}, \mathbf{K}), \\[8pt]
-%     \delta B(\mathbf{k}) = \int \frac{d^2 k'}{(2\pi)^2} 
-%     \left[ \tilde{B}(\mathbf{k'}) \cos 2\varphi_{k'k} + \tilde{E}(\mathbf{k'}) \sin 2\varphi_{k'k} \right] W(\mathbf{k'}, \mathbf{K}),
-% \end{gather*}
-
-% \noindent where \( \varphi_{k'k} \equiv \varphi_{k'} - \varphi_k \), \( \mathbf{K} = \mathbf{k} - \mathbf{k'} \), and
-% \begin{gather*}
-%     W(\mathbf{k}, \mathbf{K}) = -[\mathbf{k} \cdot \mathbf{K}] \phi(\mathbf{K}).
-% \end{gather*}
-
-% For power spectra in 2D Fourier space we use the definition
-% \begin{equation*}
-%     \langle x(\mathbf k)x'(\mathbf k') \rangle = (2\pi)^2\delta(\mathbf k + \mathbf k')C_k^{xx'}
-% \end{equation*}
-% unless stated otherwise. We will denote unlensed powerspectra with a tilde on top. $C^{xx'}(k)$ denotes the observed power spectrum is related to the true (but still lensed) spectrum as
-% \begin{equation*}
-%     C^{xx', \text{ obs}}(k) = B(k)^2C^{xx', \text{ true}}(k) + \delta_{xx'}N^x(k).
-% \end{equation*}
-% We are then able to express the power spectra of the lensed fields in terms of the unlensed fields as
-% \begin{equation*}
-%     \langle x(\mathbf k)x'(\mathbf k') \rangle = f_{xx'}(\mathbf k, \mathbf k') \phi(\mathbf k + \mathbf k'),
-% \end{equation*}
-% where $f_{xx'}$ is a function of 2 wavenumbers and unlensed power spectra. Specific values of $f$ for different combinations of $x$ and $x'$ can be found in ... %TODO: fix this
-% We can then set up an estimator for $\phi$ using a weighted integral of the Fourier modes measured for $x$ and $x'$ by allowing for a filter function $F_{xx'}(\mathbf k, \mathbf k')$.
-This gives the estimator
-% \begin{equation*}
-%     \hat\phi_{xx'}(\mathbf k) = \frac{\int \frac{\d^2k'}{(2\pi)^2}x(\mathbf k')x'(\mathbf k - \mathbf k') F_{xx'}(\mathbf k', \mathbf k - \mathbf k')}{\int \frac{\d^2k'}{(2\pi)^2}f_{xx'}(\mathbf k', \mathbf k - \mathbf k') F_{xx'}(\mathbf k', \mathbf k - \mathbf k')}.
-% \end{equation*}
-% It's easily checked that $\langle \hat\phi_{xx'} \rangle = \phi$ regardless of the filter function used. Minimizing for the variance of the estimator gives
-% \begin{equation}
-%     F_{xx'}(\mathbf{k}_1, \mathbf{k}_2) = 
-%     \frac{C_{k_1}^{x'x'} C_{k_2}^{xx} f_{xx'}(\mathbf{k}_1, \mathbf{k}_2) 
-%     - C_{k_1}^{xx'} C_{k_2}^{xx'} f_{xx'}(\mathbf{k}_2, \mathbf{k}_1)}
-%     {C_{k_1}^{xx} C_{k_2}^{x'x'} C_{k_1}^{x'x'} C_{k_2}^{xx} 
-%     - \left( C_{k_1}^{xx'} C_{k_2}^{xx'} \right)^2}.
-% \end{equation}
-% With this estimator we assume that the unlensed powerspectra have been determined through other means (e.g. a simulation).
-% With optimal weighting the noise becomes
-% \begin{equation*}
-%     \langle \hat\phi_{xy}(\mathbf K) \hat\phi_{x'y'}(\mathbf K') \rangle = (2\pi)^2\delta(\mathbf K - \mathbf K') [C^{\phi \phi}_K + K^{-2}N_{xyx'y'}(K)],
-% \end{equation*}
-% where
-% \begin{equation*}
-%     N_{xyx'y'}(K):=K^{-2}A_{xy}(K)A_{x'y'}(K)\int\frac{\d^2 k'}{(2\pi)^2}F_{xy}(\mathbf k', \mathbf k - \mathbf k')\br{
-%         F_{x'y'}(\mathbf k', \mathbf k - \mathbf k') C_{k'}^{xx'}C^{yy'}_{k - k'} + F_{x'y'}(\mathbf k - \mathbf k', \mathbf k') C_{k'}^{xy'}C^{yx'}_{k - k'}
-%     }.
-% \end{equation*}
-
-% \subsubsection{Noise}
-% We partition the sky into a finite number of pixels of size (in solid angle) $\Omega_{\text{pix}}$.
-We take into account detector beam width and homogenous noise, but not boundary effects or inhomogeneous noise.
-%TODO: justify this or take it into account
-% In real space the observed field relates to the true (but still lensed) field as
-% \begin{equation*}
-%     X^{\text{obs}}(\hat{\mathbf n}) = \int \d\Omega' X^{\text{true}}(\hat{\mathbf n}') B(\hat{\mathbf n}, \hat{\mathbf n}') + \eta(\hat{\mathbf n}),
-% \end{equation*}
-% where $B$ is the beam pattern and $\eta$ is the noise map stemming from any number of experimental sources of error.
-
-% In the case of an isotropic uniform beam, the beam pattern can be written as a function of $\hat{\mathbf n} - \hat{\mathbf n}'$ and the Fourier transform gives 
-% \begin{equation*}
-%     X^{\text{obs}}(\mathbf k) = X^{\text{true}}(\mathbf k)B(\mathbf k) + \eta(\mathbf k),
-% \end{equation*}
-% with
-% \begin{equation*}
-%     B(\mathbf k) = e^{-\frac{k^2\theta_{\text{FWHM}}^2}{4\ln 2}},
-% \end{equation*}
-% and $\theta_{\text{FWHM}}$ the angle for full width at half minimum.
-% By using a likelihood approach we obtain the estimator of the lensed power spectrum given by
-% \begin{equation*}
-%     \hat C^{XX}(k) = B(\mathbf k)^{-2}\br{X^{\text{obs}}(\mathbf k)X^{\text{obs}}(\mathbf k)^* - N(k)}
-% \end{equation*}
-% with variance
-% \begin{equation*}
-%     \Cov(\hat C^{XX}(k), \hat C^{XX}(k')) = 2\sbr{C(k) + N(k)B(k)^{-2}}\delta(\mathbf k - \mathbf k').
-% \end{equation*}
-% The noise as calculated in eq ... % TODO: ref 
-% is thus given by using
-% \begin{equation*}
-%     C^{XY, \text{ obs}}(k) = \langle \hat C^{XY}(k) \rangle = B(\mathbf k)^{-2}(C)
-% \end{equation*}
-% \subsection{Power and bispectrum estimators}
-% \subsection{Combining experimental and theoretical (shot noise?) uncertainty}
-% %Let's say you are looking at a field $X_l$ and have some experimental error $E_l$, uncorrelated with $X_l$ and with mean $0$ (so no systematic error). Then
-% %\begin{equation*}
-% %	\langle (X+E)(X+E)^*\rangle = \langle X X^*\rangle + \langle EE^* \rangle.
-% %\end{equation*}
-
-% In the case of an error expressed as
-% \begin{gather*}
-% 	\text{``We measure the amplitude of the CMB lensing} \\\text{
-% 	power spectrum at state-of-the-art 2.3\% precision, [...]"}
-% \end{gather*}
-% we can quantify this as
-% \begin{equation*}
-% 	\hat P^X_l = P_l^X(1+E),
-% \end{equation*}
-% where $E\sim \mathcal N(0, \sigma^2)$ and in the example above $\sigma = 0.023$. It thus follows that
-
-%TODO: mention SNR
-
 \section{Shear equals twice spin raised lensing potential}\label{sec:shear}
@@
 
-% \begin{figure}[t]
-%     \centering
-%     \includegraphics[height=\textheight]{../code/plots/derivative_accuracy.png}
-%     \caption{Central difference approximation of derivative of equilateral CMB weak lensing bispectrum (left) and power spectrum (right) for differences as specified in table \ref{tab:cosmo-params-diff} as well as for differences $5\%$ and $10\%$ larger and smaller.}
-%     \label{fig:deraccuracy}
-% \end{figure}
 \pagebreak
@@
 
-% NOT SUPPOSED TO STAY IN THE PAPER, JUST FOR REFERENCE
-% \section{Papers compared to}
-% Takada Jain 2005:
-% \begin{itemize}
-%     \item $ 50 < l < 3000$
-%     \item powerspectrum + bispectrum
-%     \item galaxy lensing
-%     \item $\epsilon = 0.4$, $n_g = 100$, $f_{sky}=0.1$
-%     \item derivative step: 5 percent.
-%     \item priors:
-%     \begin{gather*}
-%         \sigma(\ln \Omega_B h^2) = 0.01\\
-%         \sigma(n) = 0.008\\
-%         \sigma(H_0) = 13
-%         \sigma(m_\nu) = \infty
-%     \end{gather*}
-%     \item constraints obtained by paper:
-%     \begin{table}[h!]
-%         \centering
-%         \begin{tabular}{lccc}
-%           \hline
-%           Quantity                & PS   & Bisp & PS+Bisp \\
-%           \hline
-%           $\sigma(\Omega_{\rm de})$ & 0.14 & 0.13 & 0.023   \\
-%           $\sigma(\sigma_8)$        & 0.12 & 0.12 & 0.023   \\
-%           $\sigma(w_0)$             & 4.0  & 3.6  & 1.1     \\
-%           $\sigma(w_a)$             & 20   & 15   & 5.1     \\
-%           \hline
-%         \end{tabular}
-%       \end{table}
-%       \item constraints we got:
-%       \begin{table}[h!]
-%         \centering
-%         \begin{tabular}{lccc}
-%           \hline
-%           Quantity                & PS   & Bisp & PS+Bisp \\
-%           \hline
-%           $\sigma(\sigma_8)$        & 0.09 & 0.09 & 0.05   \\
-%           $\sigma(w_0)$             & 0.37  & 0.37  & 0.21     \\
-%           \hline
-%         \end{tabular}
-%       \end{table}
-% \end{itemize}
-
-% Planck VII:
-% \begin{itemize}
-%     \item $ 8 < l < 400$
-%     \item powerspectrum only
-%     \item cmb lensing
-%     \item noise: 5 beamwidth, $\Delta_T = 30$, $\Delta_P = 52$.
-%     \item derivative step: ??
-%     \item priors:
-%     \begin{gather*}
-%         \sigma(\Omega_b h^2) = 0.0005\\
-%         \sigma(\Omega_c h^2) = [0.01, 0.99]\\
-%         \sigma(n_s) = 0.02\\
-%         \sigma(10^10A_s) = [1.91, 3.91]\\
-%         \sigma(H_0) = [0.4, 1.0]\\
-%         \sigma(m_\nu) = \infty\\
-%         \sigma(w_0) = \infty
-%     \end{gather*}
-%     \item constraints obtained by paper:
-%     \begin{table}[h!]
-%         \centering
-%         \begin{tabular}{lc}
-%           \hline
-%           Quantity & PS  \\
-%           \hline
-%           $\sigma(\sigma_8\Omega_m^{0.25})$ & 0.020 \\
-%           \hline
-%         \end{tabular}
-%       \end{table}
-%       \item constrain that we obtained: 0.0275
-%       \item notes: actually different beamwidth for polarization and temperature
-% \end{itemize}
-
 \end{document}
